### 2013-06-27 mt-marathon.2012/PBML_article.tex
from google doc

%% Sample file $Id: PBML_article.tex 134 2010-07-01 19:28:04Z popel $
\documentclass{pbml}
%\documentclass[nofonts]{pbml} % for XeLaTeX without Pagella and DejaVu fonts
%\documentclass[color]{pbml} % for color images and hypertext links

% This is a sample file for the PBML article.
% You can compile it with (ordered by preference)
%  1. XeLaTeX with installed fonts TeX Gyre Pagella and DejaVu
%  2. XeLaTeX without the fonts -> you must use \documentclass[nofonts]{pbml}
%  3. pdfLaTeX
%
% In all three methods, you can use Unicode (utf8) encoding for special letters
% (so instead of na\"{i}ve you can write directly naïve).
% Note that with methods 2 and 3, your output will be slightly DIFFERENT than
% our final print version (different line breaks and page breaks).
% See the end of this file for more information.

% Packages
% ========
% In this place you can load required packages by the \usepackage commands.
% The following packages are loaded automatically by the class:
% euler    (for math fonts)
% graphicx (for inclusion of images)
% multicol (for multicolumn typesetting)
% natbib   (for bibliography citations)
% amssymb  (for various symbols)
% If XeLaTeX is used, fontspec and xltxtra are loaded as well.

% Definitions
% ===========
% You can use your own macros defined by \newcommand, \providecommand,
% \DeclareRobustCommand (or even by \def) as well as environments declared by
% \newenvironment. You can also use \newcounter in order to declare your own counters.

\begin{document}

% Document title and authors
% ==========================
% Due to journal organization, the article title and authors must be specified
% AFTER \begin{document}. Note that \subtitle is not allowed anymore due to
% problems with indexing in science databases. If needed use colon in the title.

\title{Title of My Article\titlelinebreak{} with a Forced Linebreak}

% Now put the affiliated institutes first, then author names and the labels
% of their institutes in the "institute" field. The order of institutes and
% authors printed below the title will be the same as the order of your commands.
% Each author can be associated with more than one institute.
% One author must be chosen as the "corresponding author" (using attribute
% corresponding) and his or her email and full address must be provided.
% PLEASE, use Unicode (utf8) encoding (e.g ï instead of \"{i}).

\institute{label1}{IBM}
\institute{label2}{CNGL, Dublin City University}
\institute{label3}{Center for Language and Speech Processing, Johns Hopkins University}
\institute{label4}{Carnegie Mellon University}

\author{
  firstname=Humpty,
  surname=Dumpty,
  institute=label1,
}
\author{
  firstname=Mock,
  surname=Turtle,
  institute={label2,label4},
  corresponding=yes,
  email={turtle@seacoast.wl},
  address={Carnegie Mellon University\\407 S. Craig Street, Pittsburgh, PA 15213, United States}
}
\author{
  firstname=Cheshire,
  initials=C.,
  surname=Cat,
  institute=label3,
}

% If all authors belong to the same institute, you can use simpler syntax:
% \institute{}{Charles University in Prague, Faculty of Mathematics and Physics, Institute of Formal and Applied Linguistics}
% \author{firstname=Humpty, surname=Dumpty}
% \author{firstname=Mock, surname=Turtle,
%   corresponding=yes,
%   email={turtle@seacoast.wl},
%   address={Institute of Formal and Applied Linguistics\\
%            Faculty of Mathematics and Physics,\\
%            Charles University in Prague\\
%            Malostranské náměstí 25\\
%            118 00 Praha 1, Czech Republic}}
% \author{firstname=Cheshire, surname=Cat}

% The title and authors' names are used in the running head. If they are
% long, you should define short versions. These definitions are optional. You
% define them only if they are needed. The example follows:
\shorttitle{Short title}
\shortauthor{H. Dumpty, M. Turtle, C. Cat}

% Now print the title by:
\PBMLmaketitle

% Abstract
% ========
% The abstract is placed within the "abstract" environment. It is a mandatory
% part of the article. PLEASE, do not use your own macros in abstract, if possible.

\begin{abstract}
Best practises for developing a SMT decoder has progressed enormously since the initial release of the Moses decoder in 2006. The emergence of syntax-inspired models, sparse features and more features, have been the driver for the major changes in the framework of the decoder.

In this paper, we describe the changes in the Moses decoder to support such changes. How we have updated the decoder, and dependent programs, while preserving the rich features in the toolkit. Particular emphasis will be place in describing the feature function framework.

The open-source community has also changed from the dominance of phrase-based model and virtual Moses monopoly, to a situation where syntactic models are leading the research agenda, and other projects such as cdec and Joshua have equal prominance.
We take this opportunity to compare the implementations of these other decoders with Moses. We find that there are many similarities between the decoders, but also some surprising differences in implementation of even very fundamental features in each decoder.

\end{abstract}

\section{Introduction}
% The body of the article
% =======================
% The PBML class is modelled after the standard article class. This means
% that you can use almost everything that is allowed in articles as described
% in the textbooks of LaTeX. We support sectioning commands \section,
% \subsection and \subsubsection.

% In addition to the \cite command, you can use natbib style of citations.
% PLEASE, use \citet instead of \cite if the author names are part of the sentence.
%\citet{PDT2} show \ldots \\ % This renders as "Hajič et al. (2006) show" instead of "(Hajič et al., 2006) show".
%\ldots tree-based annotation (e.g.~\citealp{PDT2}).

The Moses decoder was initially created for the phrase-based model. It was a vast improvement over its immediate predecessor, Pharoah, in being extensible. It allowed phrase-tables and language models. It allowed multiple phrase-table and language model implementations. The first has enabled researchers to combine multiple datasets in novel ways, as well as the use of secondary models on linguistic information (???). The second has spurred on improvements in the implementation of language models and phrase tables, which are ever faster, while using less memory (RandLM, KenLM, Compact PT).

Over time, Moses also required other search algorithms and formalisms. Cube pruning was implemented for phrase-based models in 2???. Hierarchical phrase-based and syntactic models were introduced in 2???.

However, MT research had also pushing in the direction of being able to create your own ad-hoc feature functions, and of using sparse features. Interest in incorporating syntactic features has been the most biggest driver for this, for example, (Chiang, 2009 - 11,001 features) presented a decoder able to cope with many sparse features.

This required a more general framework for incorporating feature functions into the decoder. That would make adding new features easier and more transparent. However, the new framework must be able to incorporate the existing features efficiently.

It also required tuning algorithms which are able to deal with a large number of features. It is generally known that MERT is unstable with a large number of features.

Moses now implements a new feature function framework which will allow researchers to add new feature functions much more easily. Other tuning algorithms such as MIRA, kbMIRA, and PRO have been added in order to better support a large number of sparse features  Of course, these algorithms will also work with the standard features, and MERT is still available.

This paper describes the new feature function framework.

\section{Feature Function Framework}
\subsection{Old Framework}
The built-in feature functions in the phrase-based models are:

i. Phrase-table

ii. Language model

iii. Distortion model

iii. Lexicalized Reordering model

    iv. Word penalty

    v. Unknown word penalty

The hierarchical and syntax model also uses these models, except for the distortion and lexicalized reordering model.

Each feature function was implemented in a separate class, with no common class hierarchy. The initialisation and evaluation of each feature function was entirely bespoked.

There is also no centralized store of feature functions. A new feature function requires a new variable to store the object.

Feature function state information is an essential component of the search procedure in decoding. State information control hypothesis recombination, enabling the decoder to effectively search through many more hypotheses than it actually has. State information is emitted from stateful feature functions.

The state information in Moses has evolved since inception. Initially, state information was only available from language models, therefore, only language models states were compare, Figure (1).

    CompareHypotheses(a, b)

       if Coverage(a) != Coverage(b)

         return not equal

        for each language model state s

          if (state type s for hypothesis a) != (state type s for hypothesis b)

             return not equal

        end-for
        return equal

\subsection{New Framework}        
 All feature functions inherit indirectly from the class

    FeatureFunction

which provides basic services to all feature functions such as storing the name and number of dense features, and providing a central repository of all feature functions.

A feature function can be stateless or stateful, depending on whether it inherits from

StatelessFeatureFunction

StatefulFeatureFunction

The state comparison routine of Figure (1) evolved to a more general comparison of states from all stateful feature functions, Figure (2).

    CompareHypotheses(a, b)

       if Coverage(a) != Coverage(b)

         return not equal

        for each feature function state s

          if (state type s for hypothesis a) != (state type s for hypothesis b)

             return not equal

        end-for
        return equal

\subsection{CDEC and Joshua}
CDEC and Joshua both have modern feature function frameworks similar to the Moses new framework. CDEC has no distinction between stateful and stateless features; stateless features have zero size states.

\section{Scoring}
\subsection{Old Framework}

% For figures and tables always use "figure" resp. "table" floating environments
% and always supply a caption. See the paragraph about color images below.
%\begin{figure}
% \begin{center}
%  \includegraphics[width=\textwidth]{example_figure}
%  \caption{Example figure}\label{fig:example}
% \end{center}
%\end{figure}

\section*{Acknowledgements}
This research is supported by \ldots

% Bibliography
% ============
% You may either enter the bibliography manually, possibly making use of
% \label and \ref, or use BibTeX. The bibliography style is set
% automatically. You process the bibliography by BibTeX in the
% standard way and include it by:
\bibliography{mybib}

% If needed, add appendices here
%\section*{Appendix A: \ldots}

\correspondingaddress
\end{document}

% ======= Additional information ==========

% XeLaTeX
% =======
% The Prague Bulletin of Mathematical Linguistics is typeset by XeLaTeX.
% It is not required that you use XeLaTeX and the same fonts as will be used in the
% journal but it is better to do so if you could (so you have the same line and page breaks).
% If XeLaTeX is not available on your computer, you can even used standard LaTeX.

% Fonts
% =====
% In the printed version, fonts TeX Gyre Pagella (shipped with TeX Live 2008 or later)
% and DejaVu (http://dejavu.sourceforge.net/wiki/index.php/Main_Page)
% will be used. If you have XeLaTeX but not these fonts, you may need to install
% the fonts somewhere where your system (fontconfig) can find them. On Linux,
% try something like this:
% ln -s YOUR-TEXLIVE/2008/texmf-dist/fonts/opentype/public/ \
%  ${HOME}/.fonts/texlive-2008-otf-public
% If you are not able to install these fonts, you can instruct XeLaTeX
% to use its default fonts by \document[nofonts]{pbml}.

% If you want to typset examples in east Asian scripts, you have to use
% OpenType Unicode fonts that are freely redistributable and you have to
% include them with your article. If you must use nonfree or non-Unicode
% fonts, you must supply the examples as EPS or PDF with fonts embedded in
% the files (as the required subset).
% If encountering problems with the "bidi" package, contact pbml@ufal.mff.cuni.cz.

% Vector Graphics
% ===============
% The most portable graphics package is tikz which is a front end to pgf.
% This package is fully supported both in standard LaTeX and XeLaTeX.
% PSTricks is not currently available for XeLaTeX.

% Color Images
% ============
% The printed version of PBML does not allow color figures, so please convert
% all your figures to grayscale (or black&white), make sure they are legible
% and put them in a subdirectory called "grayscale".
% When specifying the filename in \includegraphics DO NOT include the
% subdirectory.
% Optionally, in addition to the default grayscale images, you can add also
% color versions of the figures to be published in the online version.
% Put such images in a subdirectory called "color" with the same filenames
% as the grayscale ones.
% To temporarily switch to color output and hypertext links, use
% \documentclass[color]{pbml}

% Including images and other files
% ===================================
% When loading the file you must use a relative path, never the full path.
% Load the images by \includegraphics
% and other files by \input, never use \include.

### 2013-06-27 mt-marathon.2012/PBML_article.tex
from google doc

@@
 \subsection{Old Framework}
+The types of feature functions were hardcoded when evaluating the scores of translation rules and hypotheses. For example, a translation rule is scored according to Equation (1)
+
+\begin{equation}
+%  score = Sum_i _member_T(weight_i * score_i) + Sum_i _member_L(weight_i * score_i) + weight_wp * word-penalty
+\end{equation}
+
+where T is the set of translation models, and L is the set of sequence models.
+
+Similarly, when evaluating a hypothesis, the score is calculated as follows:
+
+\begin{equation}
+% score =  Sum_i _member_T(weight_i * score_i)
+%    + Sum_i _member_L(weight_i * score_i)
+%    + weight_wp * word-penalty
+%    + weight_d * score_d
+%    + Sum_i _member_R(weight_i * score_i)
+%    + future_score
+\end{equation}
+
+where $weight_d$ and $score_d$ is the distortion weight and score respectively, R is the set of lexicalized reordering models.
+
+When adding a new feature function, it would not be included in the scoring of the translation rule and hypotheses until Equation (1) and (2) are update.
+
+Translation rule always evaluated occurs before decoding, either during loading of the feature function, or just before the translation rule is used in decoding.
+
+Hypotheses are evaluated when they are created during the search process.
+
+\subsection{New Framework}
+The new framework better supports the creation of arbitary feature functions by generalizing all code that deals with feature functions.
+
+Equation (1) becomes Equation (...)
+\begin{equation}
+%  score = Sum_i _member_FF(weight_i * score_i)
+\end{equation}
+
+where FF is the set of feature functions.
+
+Equation (2) becomes Equation (...)
+\begin{equation}
+%  score =  Sum_i _member_FF(weight_i * score_i)
+%  + future_score
+\end{equation}
+
+Furthermore, there are three places where each feature function can be evaluated. At each place, a different range of information is available to the feature function to access for information.
+
+    1. During creation of translation rules. Access to the words in the source and target side of translation rule.
+
+    2. Just before decoding. Access to the entire source sentence to be decoded.
+
+    3.
+During decoding.
+Access to the context information is available, containing the hypotheses create to translate the sentence so far.
+The exact information contained in the hypotheses is dependent on whether the formalism used.
+
+Scores that are evaluated at stage one can be cached and reused inter-sentence.
+Scores evaluted during stage (2) and (3) cannot be cached as they depend on the entire source sentence.
+Therefore, it is more efficient to evaluate as early as possible.
+
+Each feature function can override any or all of the method to evaluate at each of these stage.
+For example.
+a simple word penalty feature would typically implement its evaluation in stage (1) as access to the translation rule is sufficient information to allow it to evaluate the word penalty score.
+
+On the other hand, a feature that depends on the bag of words of the input sentence would override the method for stage (2).
+
+Stateful features, and features that depend on the segmentation of the input sentence, need to override stage (3).
+
+The most efficient way to implement language models is to override stage (1) and (3).
+As a stateful feature, it must override stage (3) to access adjacent target words to score overlapping n-grams.
+However, it is able to score the n-grams within each translation rule, and also calculate the estimated future cost, without the context.
+By evaluating the stateless portion of the language model score during stage (1) mitigate these score from having to be re-evaluated each time the translation rule is applied.
+This behaviour codify the existing practises by building the framework to support it, alllowing other feature function to take advantage of this efficient mechanism.
+
+Additionally, some feature functions do not override any of the stages.
+These feature is so intertwined with the search operation that their scores are set by search operation themselves.
+The feature function exist as an identifier to locate the score.
+Features functions such as the translation models, unknown word penalty, and generation models score fall into this category.
+
+\subsection{CDEC and Joshua}
+Moses implements a one-pass search strategy. All feature functions are evaluated when a hypothesis is created.
+
+Joshua and CDEC employs a multi-pass search strategy to translate each sentence.
+Each pass creates a hypergraph that is used as the input for the next pass.
+Both decoders evaluates stateless feature functions in the first pass, then stateful feature functions in the second.
+
+CDEC also supports an arbitary number of passes and split the feature functions between those passes.
+
+\section{File Format}
+\subsection{Old Framework}
+Each feature function also required a separate section in the configuration (moses.ini) file to be created, along with the code to read the new section.
+For example, translation models are specified in the section
+
+  [ttable-file]
+
+Language models in
+
+  [lmodel-file]
+
+Furthermore, feature function also need a separate weight section, for example, for translation models:
+
+  [weight-t]
+
+and language models:
+
+ [weight-l]
+
+New feature functions need to create and manage their own specification and sections.
+
+The features that can be tuned by MERT is also hardcoded in the tuning script. The MERT script has to be altered every time a new feature function is integrated into Moses.
+
+\subsection{New Framework}
+The configuration file format has been changed to minimize the amount of boilerplate code necessary when adding new feature functions.
+All feature functions are declared in the configuration file in the section
+
+  [feature]
+
+In the fixed format
+
+  Feature-Name [key=value]*
+
+For example,
+
+  KENLM factor=0 path=lm.arpa order=5
+
+All dense weights are put into the section
+
+  [weight]
+
+In the fixed format
+
+  Feature-Instance= [value]*
+
+Similarly, weights for sparse features are in separate file referenced by the section
+
+  [weight-file]
+
+in the format
+  Feature-Instance\_feature value
+
+\section{Sparse Feature Function}
+
+Sparse features contain scores that may only be present occasionally.
+Also, there may be many more sparse features than dense features.
+There are typically 12 scores in a standard phrase-based model and 8 scores in a hierarchical phrase-based model.
+There may be millions of sparse score in a system with sparse features.
+However, only a few will be present in a particular hypothesis.
+
+This presents efficiency challenges which each decoder has attempted to solve in a different way.
+
+\subsection{Moses}
+Moses explicitly implements both dense and sparse features, and each score must be identifiable as one or the other.
+The intuition behind this is that most systems will still contain dense features, with a smattering of sparse features. Therefore, it would be more efficient to use dedicated
 

### 2013-06-27 mt-marathon.2012/PBML_article.tex
from google doc

@@
 
-\institute{label1}{IBM}
-\institute{label2}{CNGL, Dublin City University}
-\institute{label3}{Center for Language and Speech Processing, Johns Hopkins University}
-\institute{label4}{Carnegie Mellon University}
+\institute{label1}{University of Edinburgh}
+\institute{label2}{Center for Language and Speech Processing, Johns Hopkins University}
+\institute{label3}{Carnegie Mellon University}
 
 \author{
-  firstname=Humpty,
-  surname=Dumpty,
+  firstname=Hieu,
+  surname=Hoang,
   institute=label1,
@@
 \author{
-  firstname=Mock,
-  surname=Turtle,
-  institute={label2,label4},
-  corresponding=yes,
-  email={turtle@seacoast.wl},
-  address={Carnegie Mellon University\\407 S. Craig Street, Pittsburgh, PA 15213, United States}
+  firstname=Matt,
+  surname=Post,
+  institute=label2,
 }
 \author{
-  firstname=Cheshire,
-  initials=C.,
-  surname=Cat,
+  firstname=Chris,
+  surname=Dyer,
   institute=label3,
+}
+\author{
+  firstname=Eva,
+  surname=Hasler,
+  institute=label1,
+}
+\author{
+  firstname=Barry,
+  surname=Haddow,
+  institute=label1,
+}
+\author{
+  firstname=Philipp,
+  surname=Koehn,
+  institute=label1,
 }

### 2013-06-27 mt-marathon.2012/PBML_article.tex
from google doc

@@
 
-\title{Title of My Article\titlelinebreak{} with a Forced Linebreak}
+\title{A Comparison of the Synchronous-CFG \titlelinebreak{} Decoder Implementations}
 
@@
 \begin{equation}
-%  score = Sum_i _member_T(weight_i * score_i) + Sum_i _member_L(weight_i * score_i) + weight_wp * word-penalty
+  score = \sum_{i \in T}(weight_i * score_i) + \sum_{i  \in L} (weight_i * score_i) + weight_{wp} * word-penalty
 \end{equation}
 
-where T is the set of translation models, and L is the set of sequence models.
+where $T$ is the set of translation models, and $L$ is the set of sequence models.
 

### 2013-06-27 mt-marathon.2012/PBML_article.tex
from google doc

@@
 \begin{equation}
-  score = \sum_{i \in T}(weight_i * score_i) + \sum_{i  \in L} (weight_i * score_i) + weight_{wp} * word-penalty
+  score = \sum_{i \in T}(\lambda_i * score_i) + \sum_{i  \in L} (\lambda_i * score_i) + \lambda_{wp} * words
 \end{equation}
@@
 \begin{equation}
-% score =  Sum_i _member_T(weight_i * score_i)
-%    + Sum_i _member_L(weight_i * score_i)
-%    + weight_wp * word-penalty
-%    + weight_d * score_d
-%    + Sum_i _member_R(weight_i * score_i)
-%    + future_score
+ score =  \sum_{i \in T} (\lambda_i * score_i)
+    + \sum_{i \in L} (\lambda_i * score_i)
+    + \lambda_{wp} * words
+    + \lambda_d * score_d
+    + \sum_{i \in R}(\lambda_i * score_i)
+    + score_{future}
 \end{equation}
 
-where $weight_d$ and $score_d$ is the distortion weight and score respectively, R is the set of lexicalized reordering models.
+where $\lambda_d$ and $score_d$ is the distortion weight and score respectively, $R$ is the set of lexicalized reordering models.
 
@@
 \subsection{New Framework}
-The new framework better supports the creation of arbitary feature functions by generalizing all code that deals with feature functions.
+The new framework better supports the creation of arbitrary feature functions by generalizing all code that deals with feature functions.
 
@@
 \begin{equation}
-%  score = Sum_i _member_FF(weight_i * score_i)
+  score = \sum_{i \in FF}(\lambda_i * score_i)
 \end{equation}
-
-where FF is the set of feature functions.
+where $FF$ is the set of feature functions.
 
@@
 \begin{equation}
-%  score =  Sum_i _member_FF(weight_i * score_i)
-%  + future_score
+  score =  \sum_{i \in FF}(\lambda_i * score_i)
+	  + score_{future}
 \end{equation}

### 2013-06-28 mt-marathon.2012/PBML_article.tex
from google doc
blah blah blah

@@
 \institute{label2}{Center for Language and Speech Processing, Johns Hopkins University}
-\institute{label3}{Carnegie Mellon University}
+%\institute{label3}{Carnegie Mellon University}
 
@@
 \author{
-  firstname=Chris,
-  surname=Dyer,
-  institute=label3,
+  firstname=Kenneth,
+  surname=Heafield,
+  institute=label1,
 }
+%\author{
+%  firstname=Chris,
+%  surname=Dyer,
+%  institute=label3,
+%}
 \author{
@@
   surname=Hasler,
+  institute=label1,
+}
+\author{
+  firstname=Phil,
+  surname=Williams,
   institute=label1,
@@
 % define them only if they are needed. The example follows:
-\shorttitle{Short title}
+\shorttitle{SCFG Decoder Implementations}
 \shortauthor{H. Dumpty, M. Turtle, C. Cat}
@@
 Moses explicitly implements both dense and sparse features, and each score must be identifiable as one or the other.
-The intuition behind this is that most systems will still contain dense features, with a smattering of sparse features. Therefore, it would be more efficient to use dedicated
+The intuition behind this is that most systems will still contain dense features. Therefore, it would be more efficient to use dedicated dense vectors for these scores.
+
+A sparse feature map was added to handle sparse scores.
+
+\subsection{CDEC}
+CDEC eschew the dense and sparse hybrid approach and implement only sparse maps. Dense features can still be modelled in the framework, with some loss of efficiency.
+
+To improve speed and memory usage, hypotheses only store deltas of sparse features differences with antecedent hypotheses.
+
+\subsection{Joshua}
+As with CDEC, Joshua only supports sparse features, however, it uses a different tack to improve efficiency.
+
+Hypotheses only retain the weighted sum of the feature scores.
+They do not retain the individual score components unless required, for example, in the n-best list.
+The intuition behind this approach is that the score breakdown is only required in some situations such as producing n-best lists for tuning.
+In most other circumstances, they can be traded for better performance.
 

### 2013-07-01 mt-marathon.2013/PBML_article.tex
mv
start adding bibs

--- 2013-06-28 mt-marathon.2012/PBML_article.tex
@@
 
-\title{A Comparison of the Synchronous-CFG \titlelinebreak{} Decoder Implementations}
+\title{A Comparison of Synchronous-CFG \titlelinebreak{} Decoder Implementations}
 
@@
 \begin{abstract}
-Best practises for developing a SMT decoder has progressed enormously since the initial release of the Moses decoder in 2006.
+Best practise for developing a SMT decoder has progressed enormously since the initial release of the Moses decoder in 2006.
 The emergence of syntax-inspired models, sparse features and more features, have been the driver for the major changes in the framework of the decoder.
@@
 
-The open-source community has also changed from the dominance of phrase-based model and virtual Moses monopoly, to a situation where syntactic models are leading the research agenda, and other projects such as cdec and Joshua have equal prominance.
+The open-source community has also changed from the dominance of phrase-based model and virtual Moses monopoly, to a situation where syntactic models are leading the research agenda, and other projects such as cdec and Joshua have equal prominence.
 We take this opportunity to compare the implementations of these other decoders with Moses.
@@
 
-The Moses decoder was initially created for the phrase-based model.
-It was a vast improvement over its immediate predecessor, Pharoah, in being extensible.
-It allowed phrase-tables and language models.
-It allowed multiple phrase-table and language model implementations.
-The first has enabled researchers to combine multiple datasets in novel ways, as well as the use of secondary models on linguistic information (???).
+In terms of extensibility, the Moses\cite{koehn-EtAl:2007:PosterDemo} decoder was a vast improvement over its immediate predecessor, Pharoah\cite{Koehn:Pharaoh:2004}.
+
+Moses allows multiple phrase-tables and language models instances and implementations.
+The first has enabled researchers to combine multiple datasets in novel ways, as well as the use of secondary models on linguistic information\cite{Koehn:Philipp:2007, Sennrich:Rico:2011,Koehn:Philipp:2012}.
 The second has spurred on improvements in the implementation of language models and phrase tables, which are ever faster, while using less memory (RandLM, KenLM, Compact PT).

### 2013-07-01 mt-marathon.2013/PBML_article.tex
start adding bibs

@@
 
-Moses allows multiple phrase-tables and language models instances and implementations.
-The first has enabled researchers to combine multiple datasets in novel ways, as well as the use of secondary models on linguistic information\cite{Koehn:Philipp:2007, Sennrich:Rico:2011,Koehn:Philipp:2012}.
-The second has spurred on improvements in the implementation of language models and phrase tables, which are ever faster, while using less memory (RandLM, KenLM, Compact PT).
+Moses allows multiple translation models and language models instances and implementations.
+Allowing multiple instances has enabled researchers to combine multiple datasets in novel ways, as well as the use of secondary models on linguistic information\cite{Koehn:Philipp:2007, Sennrich:Rico:2011,Koehn:Philipp:2012}.
+Allowing multiple implementations has spurred on the developments of smaller, faster, better implementation of language models and phrase tables\cite{talbot-osborne:2007:ACLMain, Heafield-kenlm, Junczys-Dowmunt:Marcin:2012}.
 
-Over time, Moses also required other search algorithms and formalisms.
-Cube pruning was implemented for phrase-based models in 2???.
-Hierarchical phrase-based and syntactic models were introduced in 2???.
+However, a direction of machine translation research in the last few years has been improving translation quality with the addition of ad-hoc feature functions outside of the translation model and language model types.
+Also, the use of sparse features during decoding has also been popular.
+Research such as \cite{chiang-knight-wang:2009:NAACLHLT09, cherry:2013:NAACL-HLT} typifies this approach.
 
-However, MT research had also pushing in the direction of being able to create your own ad-hoc feature functions, and of using sparse features.
-Interest in incorporating syntactic features has been the most biggest driver for this, for example, (Chiang, 2009 - 11,001 features) presented a decoder able to cope with many sparse features.
-
-This required a more general framework for incorporating feature functions into the decoder.
-That would make adding new features easier and more transparent.
-However, the new framework must be able to incorporate the existing features efficiently.
+Enabling ad-hoc features and sparse features requires a more extensible framework that incorporate feature functions into the decoder.
 
@@
 
-Moses now implements a new feature function framework which will allow researchers to add new feature functions much more easily.
-Other tuning algorithms such as MIRA, kbMIRA, and PRO have been added in order to better support a large number of sparse features  Of course, these algorithms will also work with the standard features, and MERT is still available.
-
-This paper describes the new feature function framework.
+This paper describes this new feature function framework as implemented in the Moses decoder, a compares it with that implemented in cdec\cite{Dyer_etal_2010} and Joshua\cite{Joshua-Decoder}.
 

### 2013-07-01 mt-marathon.2013/PBML_article.tex
start adding bibs

@@
 
-This paper describes this new feature function framework as implemented in the Moses decoder, a compares it with that implemented in cdec\cite{Dyer_etal_2010} and Joshua\cite{Joshua-Decoder}.
+This paper describes this new feature function framework as implemented in the Moses decoder, and compares it with that implemented in cdec\cite{Dyer_etal_2010} and Joshua\cite{Joshua-Decoder}.
 

### 2013-07-01 mt-marathon.2013/PBML_article.tex
start adding bibs

--- 2013-07-01 mt-marathon.2013/PBML_article.tex
@@
 % If XeLaTeX is used, fontspec and xltxtra are loaded as well.
+\usepackage{algorithmic}
 
@@
 
-This paper describes this new feature function framework as implemented in the Moses decoder, a compares it with that implemented in cdec\cite{Dyer_etal_2010} and Joshua\cite{Joshua-Decoder}.
+This paper describes the changes to the Moses decoder in order to incorporate more feature functions, and compares it with that implemented in cdec\cite{Dyer_etal_2010} and Joshua\cite{Joshua-Decoder}.
 
@@
 \subsection{Old Framework}
+
 The built-in feature functions in the phrase-based models are:
 
-i. Phrase-table
+\begin{itemize}
+  \item Translation model
+  \item Language model
+  \item Distortion model
+  \item Lexicalized reordering model
+  \item Word penalty
+  \item Unknown word penalty
+\end{itemize}
 
-ii. Language model
+The hierarchical and syntax model also uses these feature functions, except for the distortion and lexicalized reordering model.
 
-iii. Distortion model
-
-iii. Lexicalized Reordering model
-
-    iv. Word penalty
-
-    v. Unknown word penalty
-
-The hierarchical and syntax model also uses these models, except for the distortion and lexicalized reordering model.
-
-Each feature function was implemented in a separate class, with no common class hierarchy. The initialisation and evaluation of each feature function was entirely bespoked.
-
-There is also no centralized store of feature functions. A new feature function requires a new variable to store the object.
+Each feature function is implemented in a separate class, with no common class hierarchy. The initialisation and evaluation of each feature function was entirely bespoked.
 
@@
 
-The state information in Moses has evolved since inception. Initially, state information was only available from language models, therefore, only language models states were compare, Figure (1).
+The state information in Moses has evolved since inception.
+Initially, state information was only available from language models, therefore, only language models states are compared, Figure~\ref{algo:CompareHypoStates}.
 
-    CompareHypotheses(a, b)
+\begin{figure} [h!]
+\small 
+\begin{algorithmic}
+\REQUIRE hypothesis $a$, hypothesis b
 
-       if Coverage(a) != Coverage(b)
+       \IF {coverage of a != coverage of b }
+	\RETURN not equal
+       \ENDIF
 
-         return not equal
+        \FORALL {LM state s}
+          \IF {state s of hypothesis a != state s for hypothesis b}
+	    \RETURN not equal
+	 \ENDIF
+        \ENDFOR
+        \RETURN equal
 
-        for each language model state s
-
-          if (state type s for hypothesis a) != (state type s for hypothesis b)
-
-             return not equal
-
-        end-for
-        return equal
+\end{algorithmic}
+\caption{Comparing Hypothesis States (old framework}
+\label{algo:CompareHypoStates}
+\end{figure}
 

### 2013-07-01 mt-marathon.2013/PBML_article.tex
start adding bibs

@@
 \end{algorithmic}
-\caption{Comparing Hypothesis States (old framework}
+\caption{Comparing Hypothesis States (old framework)}
 \label{algo:CompareHypoStates}
@@
 \subsection{New Framework}        
- All feature functions inherit indirectly from the class
+ All feature functions inherit indirectly from the class \emph{FeatureFunction}, which provides basic services to all feature functions such as storing the name and number of dense features, and providing a central repository of all feature functions.
 
-    FeatureFunction
+A feature function can be stateless or stateful, depending on whether it inherits from \emph{StatelessFeatureFunction} or \emph{StatefulFeatureFunction}, respectively.
 
-which provides basic services to all feature functions such as storing the name and number of dense features, and providing a central repository of all feature functions.
+The state comparison routine of Figure~\ref{algo:CompareHypoStates} evolved to compare the state of all stateful feature functions, Figure~\ref{algo:CompareHypoStates.new}.
 
-A feature function can be stateless or stateful, depending on whether it inherits from
+\begin{figure} [h!]
+\small 
+\begin{algorithmic}
+\REQUIRE hypothesis $a$, hypothesis b
 
-StatelessFeatureFunction
+       \IF {coverage of a != coverage of b }
+	\RETURN not equal
+       \ENDIF
 
-StatefulFeatureFunction
+        \FORALL {feature function state s}
+          \IF {state s of hypothesis a != state s for hypothesis b}
+	    \RETURN not equal
+	 \ENDIF
+        \ENDFOR
+        \RETURN equal
 
-The state comparison routine of Figure (1) evolved to a more general comparison of states from all stateful feature functions, Figure (2).
+\end{algorithmic}
+\caption{Comparing Hypothesis States (new framework)}
+\label{algo:CompareHypoStates.new}
+\end{figure}
 
-    CompareHypotheses(a, b)
+\subsection{cdec and Joshua}
+cdec and Joshua both have modern feature function frameworks similar to the Moses new framework.
+cdec has no distinction between stateful and stateless features; stateless features have zero size states.
 
-       if Coverage(a) != Coverage(b)
-
-         return not equal
-
-        for each feature function state s
-
-          if (state type s for hypothesis a) != (state type s for hypothesis b)
-
-             return not equal
-
-        end-for
-        return equal
-
-\subsection{CDEC and Joshua}
-CDEC and Joshua both have modern feature function frameworks similar to the Moses new framework.
-CDEC has no distinction between stateful and stateless features; stateless features have zero size states.
-
-\section{Scoring}
+\section{Hypothesis Evaluation}
 \subsection{Old Framework}

### 2013-07-01 mt-marathon.2013/PBML_article.tex
start adding bibs

@@
 
-\section{Hypothesis Evaluation}
+\section{Hypothesis and Translation Rule Evaluation}
 \subsection{Old Framework}
-The types of feature functions were hardcoded when evaluating the scores of translation rules and hypotheses. For example, a translation rule is scored according to Equation (1)
+The range of feature functions were hardcoded when evaluating the scores of hypotheses and translation rules. Translation rules are evaluated according to Equation~\ref{eq:translation rule score}
+\begin{equation}
+  score = \sum_{m \in T} \sum_{i \in m} (\lambda_{m,i} * score_i) + \sum_{l  \in L} (\lambda_l * score_i) + \lambda_{wp} * words
+\label{eq:translation rule score}
+\end{equation}
+where $T$ is the set of translation models, and $L$ is the set of sequence models, $\lambda_{m,i}$ is the i'th weight for translation model $m$, $\lambda_l$ is the weight for language model $l$, $\lambda_{wp}$ is the word penalty weight.
 
-\begin{equation}
-  score = \sum_{i \in T}(\lambda_i * score_i) + \sum_{i  \in L} (\lambda_i * score_i) + \lambda_{wp} * words
-\end{equation}
-
-where $T$ is the set of translation models, and $L$ is the set of sequence models.
-
-Similarly, when evaluating a hypothesis, the score is calculated as follows:
-
+Similarly, hypotheses are evaluated as follows, Equation~\ref{hypothesis rule score}:
 \begin{equation}
@@
     + score_{future}
+\label{hypothesis rule score}
 \end{equation}

### 2013-07-02 mt-marathon.2013/PBML_article.tex
single threaded build compiles & passes regression

@@
 \usepackage{algorithmic}
+\usepackage{amsmath}
 
@@
 \begin{equation}
-  score = \sum_{m \in T} \sum_{i \in m} (\lambda_{m,i} * score_i) + \sum_{l  \in L} (\lambda_l * score_i) + \lambda_{wp} * words
+  score = \sum_{m \in T} \sum_{i \in m} (\lambda_{m,i} * score_{m,i}) + \sum_{l  \in L} (\lambda_l * score_i) + \lambda_{wp} * words
 \label{eq:translation rule score}
@@
 
-Similarly, hypotheses are evaluated as follows, Equation~\ref{hypothesis rule score}:
+Similarly, hypotheses are evaluated as follows, Equation~\ref{eq:hypothesis rule score}:
 \begin{equation}
- score =  \sum_{i \in T} (\lambda_i * score_i)
-    + \sum_{i \in L} (\lambda_i * score_i)
-    + \lambda_{wp} * words
-    + \lambda_d * score_d
-    + \sum_{i \in R}(\lambda_i * score_i)
-    + score_{future}
-\label{hypothesis rule score}
+\begin{split}
+ score & =  \sum_{m \in T} \sum_{i \in m} (\lambda_{m,i} * score_{m,i}) \\
+    & + \sum_{l \in L} (\lambda_l * score_l) \\
+    & + \lambda_{wp} * words \\
+    & + \lambda_d * score_d \\
+    & + \sum_{m \in R} \sum_{i \in m} (\lambda_{m,i} * score_{m,i}) %\\
+%    & + score_{future} \\
+\label{eq:hypothesis rule score}
+\end{split}
 \end{equation}
+where $\lambda_d$ is the distortion weight, and $R$ is the set of lexicalized reordering models.
+%, and $score_{future}$ is the estimate of the future score for the hypothesis.
 
-where $\lambda_d$ and $score_d$ is the distortion weight and score respectively, $R$ is the set of lexicalized reordering models.
-
-When adding a new feature function, it would not be included in the scoring of the translation rule and hypotheses until Equation (1) and (2) are update.
+When adding a new feature function, its scores would not be included in the translation rule or hypothesis scores until the implementation of Equation~\ref{eq:translation rule score} and Equation~\ref{eq:hypothesis rule score} are update.
 

### 2013-07-02 mt-marathon.2013/PBML_article.tex
script to replace numbers with placeholder. /Achim Ruopp

@@
 \subsection{Old Framework}
-The range of feature functions were hardcoded when evaluating the scores of hypotheses and translation rules. Translation rules are evaluated according to Equation~\ref{eq:translation rule score}
+The range of feature functions are hardcoded when evaluating the scores of hypotheses and translation rules. Translation rules are evaluated according to Equation~\ref{eq:translation rule score}:
 \begin{equation}
@@
     & + \lambda_d * score_d \\
-    & + \sum_{m \in R} \sum_{i \in m} (\lambda_{m,i} * score_{m,i}) %\\
-%    & + score_{future} \\
+    & + \sum_{m \in R} \sum_{i \in m} (\lambda_{m,i} * score_{m,i})  \\
+    & + score_{future} \\
 \label{eq:hypothesis rule score}
@@
 \end{equation}
-where $\lambda_d$ is the distortion weight, and $R$ is the set of lexicalized reordering models.
-%, and $score_{future}$ is the estimate of the future score for the hypothesis.
+where $\lambda_d$ is the distortion weight, and $R$ is the set of lexicalized reordering models, and $score_{future}$ is the estimate of the score to completion for the hypothesis.
 
@@
 
-Translation rule always evaluated occurs before decoding, either during loading of the feature function, or just before the translation rule is used in decoding.
+%Translation rule always evaluated occurs before decoding, either during loading of the feature function, or just before the translation rule is used in decoding.
 
-Hypotheses are evaluated when they are created during the search process.
+%Hypotheses are evaluated when they are created during the search process.
 
@@
 
-Equation (1) becomes Equation (...)
+Equation~\ref{eq:translation rule score} becomes Equation~\ref{eq:translation rule score.new}
 \begin{equation}
-  score = \sum_{i \in FF}(\lambda_i * score_i)
+  score = \sum_{m \in FF} \sum_{i \in m} (\lambda_{m,i} * score_{m,i})
+\label{eq:translation rule score.new}
 \end{equation}
@@
 
-Equation (2) becomes Equation (...)
+Similarly, Equation~\ref{eq:hypothesis rule score} becomes Equation~\ref{eq:hypothesis rule score.new}:
 \begin{equation}
-  score =  \sum_{i \in FF}(\lambda_i * score_i)
+  score =  \sum_{m \in FF} \sum_{i \in m} (\lambda_{m,i} * score_{m,i})
 	  + score_{future}
+\label{eq:hypothesis rule score.new}
 \end{equation}
 
-Furthermore, there are three places where each feature function can be evaluated. At each place, a different range of information is available to the feature function to access for information.
-
-    1. During creation of translation rules. Access to the words in the source and target side of translation rule.
-
-    2. Just before decoding. Access to the entire source sentence to be decoded.
-
-    3.
-During decoding.
+Furthermore, there are three stages during decoding where each feature function can be evaluated:
+\begin{enumerate}
+    \item During creation of translation rules. Access to the words in the source and target side of translation rule.
+    \item Just before decoding. Access to the entire source sentence to be decoded.
+    \item During decoding.
 Access to the context information is available, containing the hypotheses create to translate the sentence so far.
 The exact information contained in the hypotheses is dependent on whether the formalism used.
-
-Scores that are evaluated at stage one can be cached and reused inter-sentence.
+\end{enumerate}
+At each stage, a different range of information is available to the feature function to access for information.
+Scores that are evaluated at stage (1) can be cached and reused inter-sentence.
 Scores evaluted during stage (2) and (3) cannot be cached as they depend on the entire source sentence.
@@
 
-Each feature function can override any or all of the method to evaluate at each of these stage.
+Each feature function can override any or all of the methods to evaluate at each of these stage.
 For example.
@@
 
-Stateful features, and features that depend on the segmentation of the input sentence, need to override stage (3).
+Stateful features need to override stage (3). Features that depend on the segmentation of the input sentence also need to override this stage.
 
@@
 
-\subsection{CDEC and Joshua}
+\subsection{cdec and Joshua}
 Moses implements a one-pass search strategy. All feature functions are evaluated when a hypothesis is created.
 
-Joshua and CDEC employs a multi-pass search strategy to translate each sentence.
+Joshua and cdec employs a multi-pass search strategy to translate each sentence.
 Each pass creates a hypergraph that is used as the input for the next pass.
@@
 
-CDEC also supports an arbitary number of passes and split the feature functions between those passes.
+cdec also supports an arbitrary number of passes, feature functions exactly to a particular pass.
 

### 2013-07-02 mt-marathon.2013/PBML_article.tex
script to replace numbers with placeholder. /Achim Ruopp

@@
 For example, translation models are specified in the section
-
+\begin{verbatim}
      [ttable-file]
-
+\end{verbatim}
 Language models in
-
+\begin{verbatim}
      [lmodel-file]
-
+\end{verbatim}
 Furthermore, feature function also need a separate weight section, for example, for translation models:
-
+\begin{verbatim}
      [weight-t]
-
+\end{verbatim}
 and language models:
-
+\begin{verbatim}
      [weight-l]
-
+\end{verbatim}
 New feature functions need to create and manage their own specification and sections.
@@
 All feature functions are declared in the configuration file in the section
-
+\begin{verbatim}
   [feature]
-
+\end{verbatim}
 In the fixed format
-
+\begin{verbatim}
   Feature-Name [key=value]*
-
+\end{verbatim}
 For example,
-
+\begin{verbatim}
   KENLM factor=0 path=lm.arpa order=5
-
-All dense weights are put into the section
-
+\end{verbatim}
+Similarly, all dense \emph{weights} are put into the section
+\begin{verbatim}
   [weight]
-
+\end{verbatim}
 In the fixed format
-
+\begin{verbatim}
   Feature-Instance= [value]*
-
-Similarly, weights for sparse features are in separate file referenced by the section
-
+\end{verbatim}
+Weights for sparse features are in separate file referenced by the section
+\begin{verbatim}
   [weight-file]
-
+\end{verbatim}
 in the format
-  Feature-Instance\_feature value
+\begin{verbatim}
+  Feature-Instance feature value
+\end{verbatim}
 

### 2013-07-02 mt-marathon.2013/PBML_article.tex
script to replace numbers with placeholder. /Achim Ruopp

@@
 The built-in feature functions in the phrase-based models are:
-
 \begin{itemize}
@@
 Also, there may be many more sparse features than dense features.
-There are typically 12 scores in a standard phrase-based model and 8 scores in a hierarchical phrase-based model.
+There are typically 14 scores from five feature functions in a standard phrase-based model and 7 scores in a hierarchical phrase-based model.
 There may be millions of sparse score in a system with sparse features.
@@
 \subsection{Moses}
-Moses explicitly implements both dense and sparse features, and each score must be identifiable as one or the other.
+Moses explicitly implements both dense and sparse features. Each score must be identifiable as one or the other.
 The intuition behind this is that most systems will still contain dense features. Therefore, it would be more efficient to use dedicated dense vectors for these scores.
@@
 \subsection{CDEC}
-CDEC eschew the dense and sparse hybrid approach and implement only sparse maps. Dense features can still be modelled in the framework, with some loss of efficiency.
+CDEC eschew the dense and sparse hybrid approach and implement only sparse maps. Dense features can still be modelled in the framework with sparse features, with some loss of efficiency.
 
-To improve speed and memory usage, hypotheses only store deltas of sparse features differences with antecedent hypotheses.
+To improve speed and memory usage, hypotheses only store deltas of feature differences with antecedent hypotheses.
 
 \subsection{Joshua}
-As with CDEC, Joshua only supports sparse features, however, it uses a different tack to improve efficiency.
+As with CDEC, Joshua only supports sparse features, however, it uses a different strategy to improve efficiency.
 
@@
 \section*{Acknowledgements}
-This research is supported by \ldots
+
+The research leading to these results has received funding from the European Union Seventh Framework Programme (FP7/2007-2013) under grant agreement 288487 (MosesCore).
 

### 2013-07-02 mt-marathon.2013/PBML_article.tex
Add plots
more

@@
 
-\subsection{CDEC}
-CDEC eschew the dense and sparse hybrid approach and implement only sparse maps. Dense features can still be modelled in the framework with sparse features, with some loss of efficiency.
+\subsection{cdec}
+cdec eschew the dense and sparse hybrid approach and implement only sparse maps. Dense features can still be modelled in the framework with sparse features, with some loss of efficiency.
 
@@
 \subsection{Joshua}
-As with CDEC, Joshua only supports sparse features, however, it uses a different strategy to improve efficiency.
+As with cdec, Joshua only supports sparse features, however, it uses a different strategy to improve efficiency.
 
@@
 %\end{figure}
+
+\section{Results}
+
+Given the same formalism and
 

### 2013-07-02 mt-marathon.2013/PBML_article.tex
more

@@
 \usepackage{amsmath}
+\usepackage{tikz,gnuplot-lua-tikz}
 
@@
 
-% For figures and tables always use "figure" resp. "table" floating environments
-% and always supply a caption. See the paragraph about color images below.
-%\begin{figure}
-% \begin{center}
-%  \includegraphics[width=\textwidth]{example_figure}
-%  \caption{Example figure}\label{fig:example}
-% \end{center}
-%\end{figure}
-
 \section{Results}
 
-Given the same formalism and
+Given the same formalism and model data, the results from each decoder is comparable, after adjusting for scaling differences and constants.
+Figure~\ref{fig:model} shows the comparison of model scores and BLEU scores for a hierarchical model.
+Moses and cdec are roughly comparable in terms of speed.
+Also, Moses and cdec converges to roughly the same model scores, while Joshua makes more search errors.
+This is due to two issues that have been resolved in Moses and cdec, but not yet in Joshua.
+
+\begin{figure}
+%run report.gnuplot to update these.
+\input{1111-8/model.tex}\input{1111-8/bleu.tex}
+\caption{\label{fig:model}Model scores and BLEU using cube pruning with pop limits ranging from 5 to 1000.}
+\end{figure}
+
+Firstly, the language model state is delegated to the language model implementation by Moses and cdec, which uses the internal data-structure of the language model to determine state, in a similar manner to (???
+Ney paper from yrs ago).
+Joshua, on the other hand, uses the words in the left and right context to determine language model state.
+This results in more aggressive hypothesis recombination in Moses and cdec, leading to better model score.
+
+Secondly, the glue rule formulation in Moses and cdec is different from Joshua's. In Moses and cdec, the glue rule grammar is:
+\begin{verbatim}
+   S -> <s>
+   S -> S X
+   GOAL -> S </s> 
+\end{verbatim}
+where <s> and </s> are first-class words like any other.
+
+In this formulation, glued hypotheses know that they are bound to the beginning of sentence <s>.
+Therefore, their left language model state is empty and more hypotheses can recombine.
+This is better than Joshua, where hypotheses are informed about <s> and </s> only after forming a complete sentence, so the left states are spuriously ambiguous.
+Moreover, hypothesis score estimates are more accurate when they know about <s> earlier.
+Figure~\ref{fig:glue} shows the improvement in cdec by changing the glue rule formulation.
+This improvement was the result of the comparison between the three decoders.
+
+\begin{figure}[h]
+\begin{center}
+\includegraphics[scale=0.50]{glue.pdf}
+\end{center}
+\caption{Affect on model score by changing glue-rule formulation}
+\label{fig:glue}
+\end{figure}
 

### 2013-07-03 mt-marathon.2013/PBML_article.tex
more

@@
 
-Given the same formalism and model data, the results from each decoder is comparable, after adjusting for scaling differences and constants.
+Given the same formalism and model data, the results from the decoders are comparable, after adjusting for scaling differences and constants.
 Figure~\ref{fig:model} shows the comparison of model scores and BLEU scores for a hierarchical model.
@@
 
-Firstly, the language model state is delegated to the language model implementation by Moses and cdec, which uses the internal data-structure of the language model to determine state, in a similar manner to (???
+Firstly, in Moses and cdec, the calculation of language model state is delegated to the language model implementation (???
+Kenlm) which uses the internal data-structure of the language model to determine state (???
 Ney paper from yrs ago).
@@
 \end{verbatim}
-where <s> and </s> are first-class words like any other.
-
-In this formulation, glued hypotheses know that they are bound to the beginning of sentence <s>.
+where <s> and </s> are first-class words like any other. In Joshua, the glue rule grammar is:
+\begin{verbatim}
+   S -> X
+   S -> S X
+   GOAL -> <s> S </s> 
+\end{verbatim}
+In Moses and cdec formulation, glued hypotheses know that they are bound to the beginning of sentence <s>.
 Therefore, their left language model state is empty and more hypotheses can recombine.
@@
 Figure~\ref{fig:glue} shows the improvement in cdec by changing the glue rule formulation.
-This improvement was the result of the comparison between the three decoders.
+This improvement was the direct result of research comparing the three decoders.
 
@@
 \end{figure}
+
+\section{Conclusion}
+
+???
 

### 2013-07-03 mt-marathon.2013/PBML_article.tex
more

@@
 \institute{label2}{Center for Language and Speech Processing, Johns Hopkins University}
-%\institute{label3}{Carnegie Mellon University}
+\institute{label3}{Carnegie Mellon University}
 
@@
 }
-%\author{
-%  firstname=Chris,
-%  surname=Dyer,
-%  institute=label3,
-%}
 \author{
@@
   institute=label1,
+}
+\author{
+  firstname=Chris,
+  surname=Dyer,
+  institute=label3,
 }
@@
 
-Firstly, in Moses and cdec, the calculation of language model state is delegated to the language model implementation (???
-Kenlm) which uses the internal data-structure of the language model to determine state (???
+Firstly, in Moses and cdec, the calculation of language model state is delegated to the language model implementation~\cite{Heafield-left} which uses the internal data-structure of the language model to determine state (???
 Ney paper from yrs ago).

### 2013-07-03 mt-marathon.2013/PBML_article.tex
more
Glue comparison

@@
 %\documentclass[color]{pbml} % for color images and hypertext links
+% test
 
@@
 \end{verbatim}
-In Moses and cdec formulation, glued hypotheses know that they are bound to the beginning of sentence <s>.
+In the Moses and cdec formulation, glued hypotheses know that they are bound to the beginning of sentence <s>.
 Therefore, their left language model state is empty and more hypotheses can recombine.
@@
 \begin{center}
-\includegraphics[scale=0.50]{glue.pdf}
+\input{1111-8/glue_model.tex}
 \end{center}
-\caption{Affect on model score by changing glue-rule formulation}
+\caption{Effect on model score by changing glue-rule formulation}
 \label{fig:glue}

### 2013-07-08 mt-marathon.2013/PBML_article.tex
corrected statement about sentence markers in joshua

@@
 
-Given the same formalism and model data, the results from the decoders are comparable, after adjusting for scaling differences and constants.
-Figure~\ref{fig:model} shows the comparison of model scores and BLEU scores for a hierarchical model.
-Moses and cdec are roughly comparable in terms of speed.
-Also, Moses and cdec converges to roughly the same model scores, while Joshua makes more search errors.
-This is due to two issues that have been resolved in Moses and cdec, but not yet in Joshua.
+Given the same formalism and model data, the results from the decoders are comparable, after
+adjusting for scaling differences and constants. Figure~\ref{fig:model} shows the comparison of
+model scores and BLEU scores for a hierarchical model. Moses and cdec are roughly comparable in
+terms of speed. Also, Moses and cdec converges to roughly the same model scores, while Joshua makes
+more search errors.
 
@@
 
-Firstly, in Moses and cdec, the calculation of language model state is delegated to the language model implementation~\cite{Heafield-left} which uses the internal data-structure of the language model to determine state (???
+This is likely due to the fact that in Moses and cdec, the calculation of language model state is delegated to the language model implementation~\cite{Heafield-left} which uses the internal data-structure of the language model to determine state (???
 Ney paper from yrs ago).
@@
 
-Secondly, the glue rule formulation in Moses and cdec is different from Joshua's. In Moses and cdec, the glue rule grammar is:
-\begin{verbatim}
-   S -> <s>
-   S -> S X
-   GOAL -> S </s> 
-\end{verbatim}
-where <s> and </s> are first-class words like any other. In Joshua, the glue rule grammar is:
-\begin{verbatim}
-   S -> X
-   S -> S X
-   GOAL -> <s> S </s> 
-\end{verbatim}
-In the Moses and cdec formulation, glued hypotheses know that they are bound to the beginning of sentence <s>.
+%% (2013-07-08 Matt Post): This isn't true; Joshua does include the <s> and </s> symbols and allows
+%% early combination with <s> (this was implemented last summer). Model scores between Moses and
+%% Joshua are exact, now, with a properly converted model. The main decoder difference seems to be
+%% that Joshua does not incorporate lefthand-side LM state collapsing for OOVs. I've commented out
+%% the below because I think you'll want to include it but it will have to be moved somewhere
+%% outside the discussion of differences from Joshua.
+
+%% Secondly, the glue rule formulation in Moses and cdec is different from Joshua's. In Moses and cdec, the glue rule grammar is:
+%% \begin{verbatim}
+%%    S -> <s>
+%%    S -> S X
+%%    GOAL -> S </s> 
+%% \end{verbatim}
+%% where <s> and </s> are first-class words like any other. In Joshua, the glue rule grammar is:
+%% \begin{verbatim}
+%%    S -> X
+%%    S -> S X
+%%    GOAL -> <s> S </s> 
+%% \end{verbatim}
+%% In Moses and cdec formulation, glued hypotheses know that they are bound to the beginning of sentence <s>.
 Therefore, their left language model state is empty and more hypotheses can recombine.
@@
 
-\begin{figure}[h]
-\begin{center}
-\input{1111-8/glue_model.tex}
-\end{center}
-\caption{Effect on model score by changing glue-rule formulation}
-\label{fig:glue}
-\end{figure}
+%% \begin{figure}[h]
+%% \begin{center}
+%% \input{1111-8/glue_model.tex}
+%% \end{center}
+%% \caption{Affect on model score by changing glue-rule formulation}
+%% \label{fig:glue}
+%% \end{figure}
 

### 2013-07-15 mt-marathon.2013/PBML_article.tex

--- 2013-07-03 mt-marathon.2013/PBML_article.tex
@@
 
-Firstly, in Moses and cdec, the calculation of language model state is delegated to the language model implementation~\cite{Heafield-left} which uses the internal data-structure of the language model to determine state (???
-Ney paper from yrs ago).
+This is likely due to the fact that in Moses and cdec, the calculation of language model state is delegated to the language model implementation~\cite{Heafield-left} which uses the internal data-structure of the language model to determine state.
 Joshua, on the other hand, uses the words in the left and right context to determine language model state.
 This results in more aggressive hypothesis recombination in Moses and cdec, leading to better model score.
+Joshua will soon be updated~\footnote{\url{https://github.com/hieuhoang/papers/commit/e417a6adcf4f1e49954836b14799509fecdea11c#mt-marathon.2013} } to incorporate better left-state minimisation to match the perform of Moses and cdec.
 
-Secondly, the glue rule formulation in Moses and cdec is different from Joshua's. In Moses and cdec, the glue rule grammar is:
+\subsection{Glue Rules}
+
+The glue rule formulation in Moses and cdec is different from Joshua's. In Moses and cdec, the glue rule grammar is:
  \begin{verbatim}
@@
  \begin{center}
-\includegraphics[scale=0.50]{glue.pdf}
+ \input{1111-8/glue_model.tex}
  \end{center}
@@
  \end{figure}
+>>>>>>> e417a6adcf4f1e49954836b14799509fecdea11c
 

### 2013-07-15 mt-marathon.2013/PBML_article.tex
changes

@@
 \author{
+  firstname=Kenneth,
+  surname=Heafield,
+  institute=label1,
+}
+\author{
   firstname=Matt,
@@
   institute=label2,
-}
-\author{
-  firstname=Kenneth,
-  surname=Heafield,
-  institute=label1,
 }
@@
 
-The glue rule formulation in Moses and cdec is different from Joshua's. In Moses and cdec, the glue rule grammar is:
+The glue grammar can be constructed in a number of ways. In formulation (1), the grammar is:
  \begin{verbatim}
@@
  \end{verbatim}
- where <s> and </s> are first-class words like any other. In Joshua, the glue rule grammar is:
+ where <s> and </s> are first-class words like any other.
+In this formulation, glued hypotheses know that they are bound to the beginning of sentence <s>.
+Therefore, their left language model state is empty and more hypotheses can recombine.
+
+The grammar for formulation (2) is:
  \begin{verbatim}
@@
  \end{verbatim}
- In Moses and cdec formulation, glued hypotheses know that they are bound to the beginning of sentence <s>.
-Therefore, their left language model state is empty and more hypotheses can recombine.
-This is better than Joshua, where hypotheses are informed about <s> and </s> only after forming a complete sentence, so the left states are spuriously ambiguous.
-Moreover, hypothesis score estimates are more accurate when they know about <s> earlier.
+Glued hypotheses are informed about <s> and </s> only after forming a complete sentence, so the left states are spuriously ambiguous.
+Moreover, hypothesis score estimates are more accurate in (1) as they know about <s> earlier.
+
 Figure~\ref{fig:glue} shows the improvement in cdec by changing the glue rule formulation.
-This improvement was the direct result of research comparing the three decoders.
+The glue grammar for each decoder had varied since their inception but the convergence on formulation (1) is the direct result of research comparing the three decoders.
 
@@
  \end{figure}
->>>>>>> e417a6adcf4f1e49954836b14799509fecdea11c
 

### 2013-07-15 mt-marathon.2013/PBML_article.tex
changes

@@
 
-Given the same formalism and model data, the results from the decoders are comparable, after
-adjusting for scaling differences and constants. Figure~\ref{fig:model} shows the comparison of
-model scores and BLEU scores for a hierarchical model. Moses and cdec are roughly comparable in
-terms of speed. Also, Moses and cdec converges to roughly the same model scores, while Joshua makes
-more search errors.
+The results from the three decoders are comparable, given the same formalism and model data, and adjusting for scaling differences and constants.
+.
+Figure~\ref{fig:model} shows the comparison of model scores and BLEU scores for a hierarchical model.
+Moses and cdec are roughly comparable in terms of speed.
+Also, Moses and cdec converges to roughly the same model scores, while Joshua makes more search errors.
 
@@
 
-This is likely due to the fact that in Moses and cdec, the calculation of language model state is delegated to the language model implementation~\cite{Heafield-left} which uses the internal data-structure of the language model to determine state.
+Joshua's lower model score is likely due to the fact that in Moses and cdec, the calculation of language model state is delegated to the language model implementation~\cite{Heafield-left} which uses the internal data-structure of the language model to determine state.
 Joshua, on the other hand, uses the words in the left and right context to determine language model state.
 This results in more aggressive hypothesis recombination in Moses and cdec, leading to better model score.
-Joshua will soon be updated~\footnote{\url{https://github.com/hieuhoang/papers/commit/e417a6adcf4f1e49954836b14799509fecdea11c#mt-marathon.2013} } to incorporate better left-state minimisation to match the perform of Moses and cdec.
+Joshua will soon be updated~\footnote{\url{https://github.com/hieuhoang/papers/commit/e417a6adcf4f1e49954836b14799509fecdea11c#mt-marathon.2013} } to incorporate better left-state minimisation to match the performance of Moses and cdec.
 

### 2013-07-15 mt-marathon.2013/PBML_article.tex
changes

@@
 \shorttitle{SCFG Decoder Implementations}
-\shortauthor{H. Dumpty, M. Turtle, C. Cat}
+\shortauthor{Hoang, Heafield et al}
 
@@
 
-In this paper, we describe the changes in the Moses decoder to support such changes.
-How we have updated the decoder, and dependent programs, while preserving the rich features in the toolkit.
+In this paper, we describe the changes in the Moses decoder and supporting programs to support such changes, while preserving the rich features in the Moses toolkit.
 Particular emphasis will be place in describing the feature function framework.
 
-The open-source community has also changed from the dominance of phrase-based model and virtual Moses monopoly, to a situation where syntactic models are leading the research agenda, and other projects such as cdec and Joshua have equal prominence.
+The open-source community has also changed from the dominance of the phrase-based model and virtual Moses monopoly, to a situation where syntactic models are leading the research agenda, and other projects such as cdec and Joshua have equal prominence.
 We take this opportunity to compare the implementations of these other decoders with Moses.
@@
 In terms of extensibility, the Moses\cite{koehn-EtAl:2007:PosterDemo} decoder was a vast improvement over its immediate predecessor, Pharoah\cite{Koehn:Pharaoh:2004}.
+For instance, Moses allows multiple translation models and language models \emph{instances} and \emph{implementations}.
 
-Moses allows multiple translation models and language models instances and implementations.
-Allowing multiple instances has enabled researchers to combine multiple datasets in novel ways, as well as the use of secondary models on linguistic information\cite{Koehn:Philipp:2007, Sennrich:Rico:2011,Koehn:Philipp:2012}.
+Allowing multiple instances has enabled researchers to combine multiple datasets in novel ways such as the use of secondary models with differing linguistic information\cite{Koehn:Philipp:2007, Sennrich:Rico:2011,Koehn:Philipp:2012}.
 Allowing multiple implementations has spurred on the developments of smaller, faster, better implementation of language models and phrase tables\cite{talbot-osborne:2007:ACLMain, Heafield-kenlm, Junczys-Dowmunt:Marcin:2012}.
 
-However, a direction of machine translation research in the last few years has been improving translation quality with the addition of ad-hoc feature functions outside of the translation model and language model types.
-Also, the use of sparse features during decoding has also been popular.
+However, a direction of machine translation research in the last few years has been on improving translation quality with the addition of ad-hoc feature functions to soft-constrain the search process.
+The use of sparse features during decoding has also been popular.
 Research such as \cite{chiang-knight-wang:2009:NAACLHLT09, cherry:2013:NAACL-HLT} typifies this approach.
 
-Enabling ad-hoc features and sparse features requires a more extensible framework that incorporate feature functions into the decoder.
+Enabling ad-hoc features in Moses requires a framework where new feature functions can be easily added.
+Adding support for sparse features entails fundamental changes in the way features function scores are stored and calculated.
+It is generally known that MERT ???
+is unstable with a large number of scores, using sparse features will also require tuning algorithms which are able to deal with a large, non-predetermined number of feature scores.
 
-It also required tuning algorithms which are able to deal with a large number of features. It is generally known that MERT is unstable with a large number of features.
-
-This paper describes the changes to the Moses decoder in order to incorporate more feature functions, and compares it with that implemented in cdec\cite{Dyer_etal_2010} and Joshua\cite{Joshua-Decoder}.
+This paper describes the changes to the Moses decoder in order to incorporate more feature functions, sparse feature scores, and compares it with that implemented in cdec\cite{Dyer_etal_2010} and Joshua\cite{Joshua-Decoder}.
 
@@
 
-???
+Statistical machine translation is a dynamic field which is constantly evolving.
+A major challenge for the Moses project, in common with many long-term software projects, is to maintain and update the the underlying framework and infrastructure to enable this evolution to continue.
 

### 2013-07-15 mt-marathon.2013/PBML_article.tex
changes

@@
 Adding support for sparse features entails fundamental changes in the way features function scores are stored and calculated.
-It is generally known that MERT ???
-is unstable with a large number of scores, using sparse features will also require tuning algorithms which are able to deal with a large, non-predetermined number of feature scores.
+It is generally known that MERT\cite{Och:MERT} is unstable with a large number of scores, using sparse features will also require tuning algorithms which are able to deal with a large, non-predetermined number of feature scores.
 
@@
 
-The built-in feature functions in the phrase-based models are:
+The standard feature functions in the phrase-based models are:
 \begin{itemize}
@@
 \end{itemize}
-
 The hierarchical and syntax model also uses these feature functions, except for the distortion and lexicalized reordering model.

### 2013-07-15 mt-marathon.2013/PBML_article.tex
changes

@@
 
-Each feature function is implemented in a separate class, with no common class hierarchy. The initialisation and evaluation of each feature function was entirely bespoked.
+Each feature function is implemented as a separate class, with no common class hierarchy. The initialisation and evaluation of each feature function was entirely bespoked.
 
@@
 
-The state information in Moses has evolved since inception.
-Initially, state information was only available from language models, therefore, only language models states are compared, Figure~\ref{algo:CompareHypoStates}.
+The state information in Moses has evolved since its inception.
+Initially, state information was only available for language models, therefore, only language models states are compared, Figure~\ref{algo:CompareHypoStates}.
 
@@
 
-The state comparison routine of Figure~\ref{algo:CompareHypoStates} evolved to compare the state of all stateful feature functions, Figure~\ref{algo:CompareHypoStates.new}.
+The state comparison routine of Figure~\ref{algo:CompareHypoStates} was generalised to compare the state of all stateful feature functions, Figure~\ref{algo:CompareHypoStates.new}.
 
@@
 \subsection{Old Framework}
-The range of feature functions are hardcoded when evaluating the scores of hypotheses and translation rules. Translation rules are evaluated according to Equation~\ref{eq:translation rule score}:
+The set of feature functions are hardcoded when evaluating the scores of hypotheses and translation rules. Translation rules are evaluated according to Equation~\ref{eq:translation rule score}:
 \begin{equation}
@@
 \end{equation}
-where $FF$ is the set of feature functions.
+where $FF$ is the set of all feature functions.
 
@@
 
-Furthermore, there are three stages during decoding where each feature function can be evaluated:
+Furthermore, there are three juncture during decoding where each feature function can be evaluated. At each point, a different range of information is available to the feature function. 
 \begin{enumerate}
-    \item During creation of translation rules. Access to the words in the source and target side of translation rule.
-    \item Just before decoding. Access to the entire source sentence to be decoded.
+    \item During creation of translation rules. Has access to the words in the source and target side of translation rule, but not the entire input sentence.
+    \item Just before decoding. Has access to the entire source sentence to be decoded.
     \item During decoding.
-Access to the context information is available, containing the hypotheses create to translate the sentence so far.
-The exact information contained in the hypotheses is dependent on whether the formalism used.
+Access to the context information is possible, containing the current and antecedent hypotheses.
+The exact information contained in the hypotheses is dependent on the formalism used.
 \end{enumerate}
-At each stage, a different range of information is available to the feature function to access for information.
+However, the stage at which the feature function is evaluate determines whether the feature function is cacheable between sentences.
 Scores that are evaluated at stage (1) can be cached and reused inter-sentence.
-Scores evaluted during stage (2) and (3) cannot be cached as they depend on the entire source sentence.
+Scores evaluated during stage (2) and (3) cannot be cached as they depend on the entire source sentence.
 Therefore, it is more efficient to evaluate as early as possible.
+
+Additionally, some feature functions do not override any of the stages.
+These feature is so intertwined with the search operation that their scores are set by search operation themselves.
+The feature function exist as an identifier to locate the score.
+Features functions such as the translation models, unknown word penalty,  generation models, and input feature score fall into this category.
+
+\subsection{Examples}
 
@@
 
-On the other hand, a feature that depends on the bag of words of the input sentence would override the method for stage (2).
+On the other hand, a feature that depends on the bag of words of the input sentence must override the evaluation method for stage (2).
 
@@
 The most efficient way to implement language models is to override stage (1) and (3).
-As a stateful feature, it must override stage (3) to access adjacent target words to score overlapping n-grams.
-However, it is able to score the n-grams within each translation rule, and also calculate the estimated future cost, without the context.
-By evaluating the stateless portion of the language model score during stage (1) mitigate these score from having to be re-evaluated each time the translation rule is applied.
-This behaviour codify the existing practises by building the framework to support it, alllowing other feature function to take advantage of this efficient mechanism.
-
-Additionally, some feature functions do not override any of the stages.
-These feature is so intertwined with the search operation that their scores are set by search operation themselves.
-The feature function exist as an identifier to locate the score.
-Features functions such as the translation models, unknown word penalty, and generation models score fall into this category.
+As a stateful feature, it must override stage (3) to access adjacent target words to score overlapping n-grams, and to set the current state.
+However, it is able to evaluate the n-grams within each translation rule without the context, therefore, language models should also override stage (1).
+The calculation of estimate future cost should also occur in stage (1).
+This behaviour codify existing practises by building the framework to support it, allowing other feature functions to take advantage of this efficient mechanism.
 

### 2013-07-16 mt-marathon.2013/PBML_article.tex
changes

@@
 \subsection{cdec and Joshua}
-Moses implements a one-pass search strategy. All feature functions are evaluated when a hypothesis is created.
+In contrast to Moses which uses a one-pass search strategy, cdec and Joshua uses a multi-pass search strategy.
+That is, for each input sentence, all feature functions are evaluated when a hypothesis is created in Moses.
 
-Joshua and cdec employs a multi-pass search strategy to translate each sentence.
-Each pass creates a hypergraph that is used as the input for the next pass.
-Both decoders evaluates stateless feature functions in the first pass, then stateful feature functions in the second.
+In cdec and Joshua, each pass creates a hypergraph that is used as the input for the next pass.
+Both decoders evaluates stateless feature functions in the first pass, followed by stateful feature functions in the second.
 
-cdec also supports an arbitrary number of passes, feature functions exactly to a particular pass.
+cdec also supports an arbitrary number of passes, feature functions can be assigned to a particular pass.
+
+cdec and Joshua do not partially evaluate stageful feature functions to estimate future cost, as done in Moses.
 
@@
 \end{verbatim}
-Furthermore, feature function also need a separate weight section, for example, for translation models:
+Similarly, each feature function also need a separate weight section, for example, for translation models:
 \begin{verbatim}
@@
 \end{verbatim}
-New feature functions need to create and manage their own specification and sections.
+New feature functions need to create and manage their own specification and weight sections.
 
-The features that can be tuned by MERT is also hardcoded in the tuning script. The MERT script has to be altered every time a new feature function is integrated into Moses.
+The tuning script must also be informed of new weight sections by manually editing the MERT script.
 
 \subsection{New Framework}
-The configuration file format has been changed to minimize the amount of boilerplate code necessary when adding new feature functions.
+The configuration file format has been changed to minimize the amount of code changes necessary when adding new feature functions.
 All feature functions are declared in the configuration file in the section

### 2013-07-16 mt-marathon.2013/PBML_article.tex
changes

@@
 Both decoders evaluates stateless feature functions in the first pass, followed by stateful feature functions in the second.
+Neither decoder partially evaluate stateful feature functions to estimate future cost, as done in Moses.
 
 cdec also supports an arbitrary number of passes, feature functions can be assigned to a particular pass.
-
-cdec and Joshua do not partially evaluate stageful feature functions to estimate future cost, as done in Moses.
 
@@
 \end{verbatim}
-Language models in
+Language models are in
 \begin{verbatim}
@@
 \end{verbatim}
-For example,
-\begin{verbatim}
-  KENLM factor=0 path=lm.arpa order=5
-\end{verbatim}
 Similarly, all dense \emph{weights} are put into the section
@@
 \end{verbatim}
+The following is an example section from a vanilla phrase-based configuration file:
+\begin{verbatim}
+  [feature]
+  UnknownWordPenalty
+  WordPenalty
+  PhraseDictionaryMemory name=TranslationModel0 table-limit=20 num-features=5...
+  LexicalReordering name=LexicalReordering0 num-features=6 type=wbe-msd...
+  Distortion
+  KENLM lazyken=1 name=LM0 factor=0 path=europarl.binlm.1 order=5
+\end{verbatim}
+Notice that feature functions \emph{UnknownWordPenalty}, \emph{WordPenalty}, and \emph{Distortion} must be explicitly specified.
+In the old format, these features were implicitly included for all phrase-based models.
+
+The weights section from the same moses.ini file is shown below:
+\begin{verbatim}
+  [weight]
+  UnknownWordPenalty0= 1
+  WordPenalty0= -1
+  TranslationModel0= 0.2 0.2 0.2 0.2 0.2
+  LexicalReordering0= 0.3 0.3 0.3 0.3 0.3 0.3
+  Distortion0= 0.3
+  LM0= 0.5
+\end{verbatim}
 
@@
 Sparse features contain scores that may only be present occasionally.
-Also, there may be many more sparse features than dense features.
-There are typically 14 scores from five feature functions in a standard phrase-based model and 7 scores in a hierarchical phrase-based model.
-There may be millions of sparse score in a system with sparse features.
-However, only a few will be present in a particular hypothesis.
+However, whereas there are typically 14 scores from five feature functions in a standard phrase-based model and 7 scores in a hierarchical phrase-based model, there may be millions of sparse scores in a system with sparse features.
+Only a few will be present in a particular hypothesis or translation rule.
 
-This presents efficiency challenges which each decoder has attempted to solve in a different way.
+Storing and using sparse feature efficiently present a challenge which each decoder has attempted to solve differently.
 
 \subsection{Moses}
-Moses explicitly implements both dense and sparse features. Each score must be identifiable as one or the other.
-The intuition behind this is that most systems will still contain dense features. Therefore, it would be more efficient to use dedicated dense vectors for these scores.
+Moses implements supports for both dense and sparse features. Each score must be identifiable as one or the other.
+The intuition behind this is that most systems will still contain dense features. Therefore, it would be more efficient to use existing dedicated data-structures to efficiently handle dense vectors.
 
@@
 \subsection{cdec}
-cdec eschew the dense and sparse hybrid approach and implement only sparse maps. Dense features can still be modelled in the framework with sparse features, with some loss of efficiency.
+cdec eschew Moses' dense and sparse hybrid approach and implement only sparse maps. Dense features can still be modelled in the framework as sparse features, whose name are suffixed with indices.
 
@@
 
-Hypotheses only retain the weighted sum of the feature scores.
-They do not retain the individual score components unless required, for example, in the n-best list.
-The intuition behind this approach is that the score breakdown is only required in some situations such as producing n-best lists for tuning.
-In most other circumstances, they can be traded for better performance.
+Hypotheses only retain the weighted sum of the feature scores, .
+The individual score components are not stored unless required, for example, in the n-best list.
+The intuition behind this approach is that the score breakdown is only required in some situations.
+In most other circumstances, they can be traded for better speed and lower memory consumption.
 
@@
 Moses and cdec are roughly comparable in terms of speed.
-Also, Moses and cdec converges to roughly the same model scores, while Joshua makes more search errors.
+Also, Moses and cdec converges to roughly the same model scores, while Joshua makes more search errors, leading to lower BLEU scores.
 
@@
 
-Figure~\ref{fig:glue} shows the improvement in cdec by changing the glue rule formulation.
-The glue grammar for each decoder had varied since their inception but the convergence on formulation (1) is the direct result of research comparing the three decoders.
+Figure~\ref{fig:glue} shows the improvement in cdec by changing the glue rule formulation from (2) to (1).
+The glue grammar for each decoder had varied since their inceptions.
+The convergence on using formulation (1) is the direct result of the comparison between the three decoders.
 

### 2013-07-16 mt-marathon.2013/PBML_article.tex
changes

@@
 A major challenge for the Moses project, in common with many long-term software projects, is to maintain and update the the underlying framework and infrastructure to enable this evolution to continue.
+The paper documents the major changes that were made to Moses to enable this to happen and compare it two other decoders which implements the same decoding formalism.
+
+Even though each decoder was created independently, their output are comparable.
+We see that be doing an indepth analysis of the working of each decoder and comparing their model score, we are able to transfer the best optmization technique from each decoder to the others.
 

### 2013-07-16 mt-marathon.2013/PBML_article.tex
changes

@@
 
-\section{File Format}
-\subsection{Old Framework}
-Each feature function also required a separate section in the configuration (moses.ini) file to be created, along with the code to read the new section.
+% \section{File Format}
+% \subsection{Old Framework}
+% Each feature function also required a separate section in the configuration (moses.ini) file to be created, along with the code to read the new section.
 For example, translation models are specified in the section
-\begin{verbatim}
-     [ttable-file]
-\end{verbatim}
-Language models are in
-\begin{verbatim}
-     [lmodel-file]
-\end{verbatim}
-Similarly, each feature function also need a separate weight section, for example, for translation models:
-\begin{verbatim}
-     [weight-t]
-\end{verbatim}
-and language models:
-\begin{verbatim}
-     [weight-l]
-\end{verbatim}
-New feature functions need to create and manage their own specification and weight sections.
-
-The tuning script must also be informed of new weight sections by manually editing the MERT script.
-
-\subsection{New Framework}
-The configuration file format has been changed to minimize the amount of code changes necessary when adding new feature functions.
+% \begin{verbatim}
+%      [ttable-file]
+% \end{verbatim}
+% Language models are in
+% \begin{verbatim}
+%      [lmodel-file]
+% \end{verbatim}
+% Similarly, each feature function also need a separate weight section, for example, for translation models:
+% \begin{verbatim}
+%      [weight-t]
+% \end{verbatim}
+% and language models:
+% \begin{verbatim}
+%      [weight-l]
+% \end{verbatim}
+% New feature functions need to create and manage their own specification and weight sections.
+% 
+% The tuning script must also be informed of new weight sections by manually editing the MERT script.
+% 
+% \subsection{New Framework}
+% The configuration file format has been changed to minimize the amount of code changes necessary when adding new feature functions.
 All feature functions are declared in the configuration file in the section
-\begin{verbatim}
-  [feature]
-\end{verbatim}
-In the fixed format
-\begin{verbatim}
-  Feature-Name [key=value]*
-\end{verbatim}
-Similarly, all dense \emph{weights} are put into the section
-\begin{verbatim}
-  [weight]
-\end{verbatim}
-In the fixed format
-\begin{verbatim}
-  Feature-Instance= [value]*
-\end{verbatim}
-Weights for sparse features are in separate file referenced by the section
-\begin{verbatim}
-  [weight-file]
-\end{verbatim}
-in the format
-\begin{verbatim}
-  Feature-Instance feature value
-\end{verbatim}
-The following is an example section from a vanilla phrase-based configuration file:
-\begin{verbatim}
-  [feature]
-  UnknownWordPenalty
-  WordPenalty
-  PhraseDictionaryMemory name=TranslationModel0 table-limit=20 num-features=5...
-  LexicalReordering name=LexicalReordering0 num-features=6 type=wbe-msd...
-  Distortion
-  KENLM lazyken=1 name=LM0 factor=0 path=europarl.binlm.1 order=5
-\end{verbatim}
-Notice that feature functions \emph{UnknownWordPenalty}, \emph{WordPenalty}, and \emph{Distortion} must be explicitly specified.
+% \begin{verbatim}
+%   [feature]
+% \end{verbatim}
+% In the fixed format
+% \begin{verbatim}
+%   Feature-Name [key=value]*
+% \end{verbatim}
+% Similarly, all dense \emph{weights} are put into the section
+% \begin{verbatim}
+%   [weight]
+% \end{verbatim}
+% In the fixed format
+% \begin{verbatim}
+%   Feature-Instance= [value]*
+% \end{verbatim}
+% Weights for sparse features are in separate file referenced by the section
+% \begin{verbatim}
+%   [weight-file]
+% \end{verbatim}
+% in the format
+% \begin{verbatim}
+%   Feature-Instance feature value
+% \end{verbatim}
+% The following is an example section from a vanilla phrase-based configuration file:
+% \begin{verbatim}
+%   [feature]
+%   UnknownWordPenalty
+%   WordPenalty
+%   PhraseDictionaryMemory name=TranslationModel0 table-limit=20 num-features=5...
+%   LexicalReordering name=LexicalReordering0 num-features=6 type=wbe-msd...
+%   Distortion
+%   KENLM lazyken=1 name=LM0 factor=0 path=europarl.binlm.1 order=5
+% \end{verbatim}
+% Notice that feature functions \emph{UnknownWordPenalty}, \emph{WordPenalty}, and \emph{Distortion} must be explicitly specified.
 In the old format, these features were implicitly included for all phrase-based models.
-
-The weights section from the same moses.ini file is shown below:
-\begin{verbatim}
-  [weight]
-  UnknownWordPenalty0= 1
-  WordPenalty0= -1
-  TranslationModel0= 0.2 0.2 0.2 0.2 0.2
-  LexicalReordering0= 0.3 0.3 0.3 0.3 0.3 0.3
-  Distortion0= 0.3
-  LM0= 0.5
-\end{verbatim}
+% 
+% The weights section from the same moses.ini file is shown below:
+% \begin{verbatim}
+%   [weight]
+%   UnknownWordPenalty0= 1
+%   WordPenalty0= -1
+%   TranslationModel0= 0.2 0.2 0.2 0.2 0.2
+%   LexicalReordering0= 0.3 0.3 0.3 0.3 0.3 0.3
+%   Distortion0= 0.3
+%   LM0= 0.5
+% \end{verbatim}
 

### 2013-07-16 mt-marathon.2013/PBML_article.tex
changes

@@
 Even though each decoder was created independently, their output are comparable.
-We see that be doing an indepth analysis of the working of each decoder and comparing their model score, we are able to transfer the best optmization technique from each decoder to the others.
+We see that by doing an indepth analysis of the working of each decoder and comparing their model scores, we are able to transfer the best optmization techniques from each decoder to the others.
 

### 2013-07-17 mt-marathon.2013/PBML_article.tex
changes
merge

@@
 \author{
+  firstname=Barry,
+  surname=Haddow,
+  institute=label1,
+}
+\author{
   firstname=Matt,
@@
   surname=Williams,
-  institute=label1,
-}
-\author{
-  firstname=Barry,
-  surname=Haddow,
   institute=label1,
@@
 \begin{abstract}
-Best practise for developing a SMT decoder has progressed enormously since the initial release of the Moses decoder in 2006.
+Best practices for developing a SMT decoder have progressed enormously since the initial release of the Moses decoder in 2006.
 The emergence of syntax-inspired models, sparse features and more features, have been the driver for the major changes in the framework of the decoder.
@@
 In terms of extensibility, the Moses\cite{koehn-EtAl:2007:PosterDemo} decoder was a vast improvement over its immediate predecessor, Pharoah\cite{Koehn:Pharaoh:2004}.
-For instance, Moses allows multiple translation models and language models \emph{instances} and \emph{implementations}.
+For instance, Moses allows multiple translation model and language model \emph{instances} and \emph{implementations}.
 
@@
 
-However, a direction of machine translation research in the last few years has been on improving translation quality with the addition of ad-hoc feature functions to soft-constrain the search process.
+However, a direction of machine translation research in the last few years has been on improving translation quality with the addition of novel feature functions, for example, \cite{Gimpel:2008:RSC:1626394.1626396}.
 The use of sparse features during decoding has also been popular.
@@
 
-Enabling ad-hoc features in Moses requires a framework where new feature functions can be easily added.
+Enabling novel features functions in Moses requires a framework where new feature functions can be easily added.
 Adding support for sparse features entails fundamental changes in the way features function scores are stored and calculated.
@@
 \section{Feature Function Framework}
-\subsection{Old Framework}
+\subsection{Original Framework}
 
@@
 
-\subsection{New Framework}        
+\subsection{Current Framework}        
  All feature functions inherit indirectly from the class \emph{FeatureFunction}, which provides basic services to all feature functions such as storing the name and number of dense features, and providing a central repository of all feature functions.
 
-A feature function can be stateless or stateful, depending on whether it inherits from \emph{StatelessFeatureFunction} or \emph{StatefulFeatureFunction}, respectively.
+Feature function \emph{state} plays an important role in hypothesis recombination.
+Two hypotheses can be recombined if their states are equivalent.
+Higher recombination results in increased effective number of hypotheses created, leading to better model score.
+
+Each feature function can be stateless or stateful, depending on whether it emits a state for each hypothesis.
+Stateful features imherits from \emph{StatefulFeatureFunction}, otherwise they it inherits from \emph{StatelessFeatureFunction}.
 
@@
 \section{Hypothesis and Translation Rule Evaluation}
-\subsection{Old Framework}
+\subsection{Original Framework}
 The set of feature functions are hardcoded when evaluating the scores of hypotheses and translation rules. Translation rules are evaluated according to Equation~\ref{eq:translation rule score}:
@@
 \end{equation}
-where $T$ is the set of translation models, and $L$ is the set of sequence models, $\lambda_{m,i}$ is the i'th weight for translation model $m$, $\lambda_l$ is the weight for language model $l$, $\lambda_{wp}$ is the word penalty weight.
+where $T$ is the set of translation models, and $L$ is the set of language models, $\lambda_{m,i}$ is the i'th weight for translation model $m$, $\lambda_l$ is the weight for language model $l$, $\lambda_{wp}$ is the word penalty weight.
+Translation rules are evaluated before the search process.
+The score are used to inform table pruning and cube pruning.
 
-Similarly, hypotheses are evaluated as follows, Equation~\ref{eq:hypothesis rule score}:
+Similarly, hypotheses are evaluated as follows during the search process, Equation~\ref{eq:hypothesis rule score}:
 \begin{equation}
@@
 
-When adding a new feature function, its scores would not be included in the translation rule or hypothesis scores until the implementation of Equation~\ref{eq:translation rule score} and Equation~\ref{eq:hypothesis rule score} are update.
+When adding a new feature function, its scores would not be included in the translation rule or hypothesis scores until the implementation of Equation~\ref{eq:translation rule score} and Equation~\ref{eq:hypothesis rule score} are updated.
 
@@
 
-\subsection{New Framework}
+\subsection{Current Framework}
 The new framework better supports the creation of arbitrary feature functions by generalizing all code that deals with feature functions.
@@
 
-Furthermore, there are three juncture during decoding where each feature function can be evaluated. At each point, a different range of information is available to the feature function. 
+Furthermore, there are three junctures during decoding where each feature function can be evaluated. At each point, a different range of information is available to the feature function. 
 \begin{enumerate}
@@
 Additionally, some feature functions do not override any of the stages.
-These feature is so intertwined with the search operation that their scores are set by search operation themselves.
+These feature is so intertwined with the search mechanism that their scores are set by search mechanism themselves.
 The feature function exist as an identifier to locate the score.
-Features functions such as the translation models, unknown word penalty,  generation models, and input feature score fall into this category.
+Features functions such as the translation models, unknown word penalty,  generation models fall into this category.
+The input feature, which scores lattices and confusion network input is also hardcoded into the decoder.
 
@@
 Both decoders evaluates stateless feature functions in the first pass, followed by stateful feature functions in the second.
-Neither decoder partially evaluate stateful feature functions to estimate future cost, as done in Moses.
+Neither decoder partially evaluate stateful feature functions, as done in Moses.
 
@@
 \subsection{Moses}
-Moses implements supports for both dense and sparse features. Each score must be identifiable as one or the other.
+Moses implements supports for both dense and sparse features. Each feature must be identifiable as one or the other.
 The intuition behind this is that most systems will still contain dense features. Therefore, it would be more efficient to use existing dedicated data-structures to efficiently handle dense vectors.
@@
 \subsection{cdec}
-cdec eschew Moses' dense and sparse hybrid approach and implement only sparse maps. Dense features can still be modelled in the framework as sparse features, whose name are suffixed with indices.
+cdec eschews Moses' dense and sparse hybrid approach and implement only sparse maps. Dense features can still be modelled in the framework as sparse features, whose name are suffixed with indices.
 
@@
 The results from the three decoders are comparable, given the same formalism and model data, and adjusting for scaling differences and constants.
-.
 Figure~\ref{fig:model} shows the comparison of model scores and BLEU scores for a hierarchical model.

### 2013-07-17 mt-marathon.2013/PBML_article.tex
merge
made some changes to the description of joshua's decoding algorithm

@@
 \institute{label1}{University of Edinburgh}
-\institute{label2}{Center for Language and Speech Processing, Johns Hopkins University}
+\institute{label2}{Human Language Technology Center of Excellence, Johns Hopkins University}
 \institute{label3}{Carnegie Mellon University}
@@
 \REQUIRE hypothesis $a$, hypothesis b
-
        \IF {coverage of a != coverage of b }
@@
         \RETURN equal
-
 \end{algorithmic}
@@
 \subsection{cdec and Joshua}
-cdec and Joshua both have modern feature function frameworks similar to the Moses new framework.
-cdec has no distinction between stateful and stateless features; stateless features have zero size states.
+cdec and Joshua both have modern feature function frameworks similar to the Moses new
+framework. Both of them have no distinction between stateful and stateless features; stateless
+features have zero size states.
 
@@
 
-\subsection{cdec and Joshua}
-In contrast to Moses which uses a one-pass search strategy, cdec and Joshua uses a multi-pass search strategy.
+\subsection{Joshua}
+
+Joshua's decoding algorithm is an implementation of CKY+ \cite{chappelier1998generalized}. It
+supports decoding with arbitrary synchronous CFG productions, imposing no limits on the rank of
+rules or on the intermingling of terminals and nonterminals.
+
+Joshua makes only a single CKY pass over the source language sentence; it does not build a
+``stateless'' hypergraph (or --LM forest, in the language of \citet{chiang2007hierarchical}). To
+enable this, Joshua interleaves the extension of partially-constructed rules with the application of
+complete rules while building the hypergraph, depicted here with pseudocode:
+
+\begin{figure} [h!]
+\small 
+\begin{algorithmic}
+  \FORALL {spans $(i,j)$ in bottom-up order}
+    \STATE {extend incomplete rules seeking the terminal at $(j-1,j)$}
+    \FORALL {split points $k$}
+      \STATE {extend incomplete rules over $(i,k)$ seeking a nonterminal $N$ such that chart item $(N,k,j)$ exists in the hypergraph}
+    \ENDFOR
+    \STATE {construct new hypergraph node by applying cube pruning to all completed rules over $(i,j)$}
+  \ENDFOR
+\end{algorithmic}
+\caption{Interweaving of partially-completed and completed rules in Joshua.}
+\label{algorithm:joshua}
+\end{figure}
+
+The chart containing partially-completed rules is called the \emph{dot chart}, and no pruning is
+applied to it; rule features are computed only when complete rules are applied.
+
+The efficiency gained from this interleaving allows Joshua to decode with SCFGs with rich
+nonterminal sets \cite{galley2004whats,zollmann2006syntax}, although the time and memory
+requirements increase dramatically. Incorporating feature scores into the dot chart for more
+efficient decoding with SCFGs with rich nonterminal sets is a direction of interest.
+
+\subsection{cdec}
+In contrast to Moses and Joshua which use a one-pass search strategy, cdec uses a multi-pass search strategy.
 That is, for each input sentence, all feature functions are evaluated when a hypothesis is created in Moses.
 
-In cdec and Joshua, each pass creates a hypergraph that is used as the input for the next pass.
-Both decoders evaluates stateless feature functions in the first pass, followed by stateful feature functions in the second.
+In cdec, each pass creates a hypergraph that is used as the input for the next pass.
+cdec evaluates stateless feature functions in the first pass, followed by stateful feature functions in the second.
 Neither decoder partially evaluate stateful feature functions, as done in Moses.
 
-cdec also supports an arbitrary number of passes, feature functions can be assigned to a particular pass.
+cdec also supports an arbitrary number of passes, feature functions can be assigned to a particular
+pass.
 
@@
 The results from the three decoders are comparable, given the same formalism and model data, and adjusting for scaling differences and constants.
-Figure~\ref{fig:model} shows the comparison of model scores and BLEU scores for a hierarchical model.
-Moses and cdec are roughly comparable in terms of speed.
-Also, Moses and cdec converges to roughly the same model scores, while Joshua makes more search errors, leading to lower BLEU scores.
-
-\begin{figure}
-%run report.gnuplot to update these.
-\input{1111-8/model.tex}\input{1111-8/bleu.tex}
-\caption{\label{fig:model}Model scores and BLEU using cube pruning with pop limits ranging from 5 to 1000.}
-\end{figure}
-
-Joshua's lower model score is likely due to the fact that in Moses and cdec, the calculation of language model state is delegated to the language model implementation~\cite{Heafield-left} which uses the internal data-structure of the language model to determine state.
-Joshua, on the other hand, uses the words in the left and right context to determine language model state.
-This results in more aggressive hypothesis recombination in Moses and cdec, leading to better model score.
-Joshua will soon be updated~\footnote{\url{https://github.com/hieuhoang/papers/commit/e417a6adcf4f1e49954836b14799509fecdea11c#mt-marathon.2013} } to incorporate better left-state minimisation to match the performance of Moses and cdec.
 
@@
  \end{figure}
+
+\subsection{Head-to-head}
+
+Figure~\ref{fig:model} shows the comparison of model scores and BLEU scores for a hierarchical model.
+Moses and cdec are roughly comparable in terms of speed.
+Also, Moses and cdec converges to roughly the same model scores, while Joshua makes more search errors, leading to lower BLEU scores.
+
+\begin{figure}
+%run report.gnuplot to update these.
+\input{1111-8/model.tex}\input{1111-8/bleu.tex}
+\caption{\label{fig:model}Model scores and BLEU using cube pruning with pop limits ranging from 5 to 1000.}
+\end{figure}
+
+Joshua's lower model score is likely due to the fact that in Moses and cdec, the calculation of language model state is delegated to the language model implementation~\cite{Heafield-left} which uses the internal data-structure of the language model to determine state.
+Joshua, on the other hand, uses the words in the left and right context to determine language model state.
+This results in more aggressive hypothesis recombination in Moses and cdec, leading to better model score.
+Joshua will soon be updated~\footnote{\url{https://github.com/hieuhoang/papers/commit/e417a6adcf4f1e49954836b14799509fecdea11c#mt-marathon.2013} } to incorporate better left-state minimisation to match the performance of Moses and cdec.
 

### 2013-07-18 mt-marathon.2013/PBML_article.tex
merge

@@
 
-\title{A Comparison of Synchronous-CFG \titlelinebreak{} Decoder Implementations}
+\title{Updating the Feature Function Framework \titlelinebreak{} in the Moses Decoder}
 
@@
 
-The open-source community has also changed from the dominance of the phrase-based model and virtual Moses monopoly, to a situation where syntactic models are leading the research agenda, and other projects such as cdec and Joshua have equal prominence.
-We take this opportunity to compare the implementations of these other decoders with Moses.
-We find that there are many similarities between the decoders, but also some surprising differences in implementation of even very fundamental features in each decoder.
+We compare the implementation of the Moses SCFG decoder to that of cdec and Joshua and find that there are some surprising differences in implementation of even very fundamental features in each decoder.
 
@@
 
-Additionally, some feature functions do not override any of the stages.
+Additionally, the evaluation at stage (1) can return an estimate of the future cost to guide search. This is a technique used by the phrase-based decoder but not by the SCFG decoder.
+
+Some feature functions do not override any of the stages.
 These feature is so intertwined with the search mechanism that their scores are set by search mechanism themselves.
@@
 
-Joshua's decoding algorithm is an implementation of CKY+ \cite{chappelier1998generalized}. It
-supports decoding with arbitrary synchronous CFG productions, imposing no limits on the rank of
-rules or on the intermingling of terminals and nonterminals.
+As with cdec and Moses' SCFG decoder, Joshua's decoding algorithm is an implementation of CKY+ \cite{chappelier1998generalized}.
 
@@
 enable this, Joshua interleaves the extension of partially-constructed rules with the application of
-complete rules while building the hypergraph, depicted here with pseudocode:
+complete rules while building the hypergraph, Figure~\ref{algorithm:joshua}:
 
@@
 
-The efficiency gained from this interleaving allows Joshua to decode with SCFGs with rich
-nonterminal sets \cite{galley2004whats,zollmann2006syntax}, although the time and memory
-requirements increase dramatically. Incorporating feature scores into the dot chart for more
-efficient decoding with SCFGs with rich nonterminal sets is a direction of interest.
+% The efficiency gained from this interleaving allows Joshua to decode with SCFGs with rich
+% nonterminal sets \cite{galley2004whats,zollmann2006syntax}, although the time and memory
+% requirements increase dramatically. Incorporating feature scores into the dot chart for more
+% efficient decoding with SCFGs with rich nonterminal sets is a direction of interest.
 
@@
 In contrast to Moses and Joshua which use a one-pass search strategy, cdec uses a multi-pass search strategy.
-That is, for each input sentence, all feature functions are evaluated when a hypothesis is created in Moses.
-
-In cdec, each pass creates a hypergraph that is used as the input for the next pass.
-cdec evaluates stateless feature functions in the first pass, followed by stateful feature functions in the second.
-Neither decoder partially evaluate stateful feature functions, as done in Moses.
+%That is, for each input sentence, all feature functions are evaluated when a hypothesis is created in Moses.
+Each pass creates a hypergraph that is used as the input for the next pass. cdec evaluates stateless feature functions in the first pass, followed by stateful feature functions in the second.
 
@@
 \subsection{Moses}
-Moses implements supports for both dense and sparse features. Each feature must be identifiable as one or the other.
-The intuition behind this is that most systems will still contain dense features. Therefore, it would be more efficient to use existing dedicated data-structures to efficiently handle dense vectors.
-
+Moses implements supports for both dense and sparse features.
+Each feature must be identifiable as one or the other.
+The intuition behind this is that most systems will still contain dense features.
+Therefore, it would be more efficient to use existing dedicated data-structures to efficiently handle dense vectors.
 A sparse feature map was added to handle sparse scores.
@@
 The results from the three decoders are comparable, given the same formalism and model data, and adjusting for scaling differences and constants.
+Figure~\ref{fig:model} shows the comparison of model scores and BLEU scores for a hierarchical model.
+Moses and cdec are roughly comparable in terms of speed.
+Also, Moses and cdec converges to roughly the same model scores, while Joshua makes more search errors, leading to overall lower BLEU scores.
+
+\begin{figure}
+%run report.gnuplot to update these.
+\input{1111-8/model.tex}\input{1111-8/bleu.tex}
+\caption{\label{fig:model}Model scores and BLEU using cube pruning with pop limits ranging from 5 to 1000.}
+\end{figure}
+
+Joshua's lower model score is likely due to the fact that in Moses and cdec, the calculation of language model state is delegated to the language model implementation~\cite{Heafield-left} which uses the internal data-structure of the language model to determine state.
+Joshua, on the other hand, uses the words in the left and right context to determine language model state.
+This results in more aggressive hypothesis recombination in Moses and cdec, leading to better model score.
+Joshua will soon be updated~\footnote{\url{https://github.com/hieuhoang/papers/commit/e417a6adcf4f1e49954836b14799509fecdea11c#mt-marathon.2013} } to incorporate better left-state minimisation to match the performance of Moses and cdec.
 
@@
 The convergence on using formulation (1) is the direct result of the comparison between the three decoders.
+(Result for cdec and Joshua in Figure~\ref{fig:model} were done with the improved glue grammar formulation.)
 
@@
 
-\subsection{Head-to-head}
-
-Figure~\ref{fig:model} shows the comparison of model scores and BLEU scores for a hierarchical model.
-Moses and cdec are roughly comparable in terms of speed.
-Also, Moses and cdec converges to roughly the same model scores, while Joshua makes more search errors, leading to lower BLEU scores.
-
-\begin{figure}
-%run report.gnuplot to update these.
-\input{1111-8/model.tex}\input{1111-8/bleu.tex}
-\caption{\label{fig:model}Model scores and BLEU using cube pruning with pop limits ranging from 5 to 1000.}
-\end{figure}
-
-Joshua's lower model score is likely due to the fact that in Moses and cdec, the calculation of language model state is delegated to the language model implementation~\cite{Heafield-left} which uses the internal data-structure of the language model to determine state.
-Joshua, on the other hand, uses the words in the left and right context to determine language model state.
-This results in more aggressive hypothesis recombination in Moses and cdec, leading to better model score.
-Joshua will soon be updated~\footnote{\url{https://github.com/hieuhoang/papers/commit/e417a6adcf4f1e49954836b14799509fecdea11c#mt-marathon.2013} } to incorporate better left-state minimisation to match the performance of Moses and cdec.
-
 \section{Conclusion}
@@
 A major challenge for the Moses project, in common with many long-term software projects, is to maintain and update the the underlying framework and infrastructure to enable this evolution to continue.
-The paper documents the major changes that were made to Moses to enable this to happen and compare it two other decoders which implements the same decoding formalism.
+This paper documents the major changes that were made to Moses to enable this to happen and compare it two other decoders which implements the same decoding formalism.
 
-Even though each decoder was created independently, their output are comparable.
+%Even though each decoder was created independently, their output are comparable.
 We see that by doing an indepth analysis of the working of each decoder and comparing their model scores, we are able to transfer the best optmization techniques from each decoder to the others.

### 2013-07-18 mt-marathon.2013/PBML_article.tex
merge

@@
 
-In this paper, we describe the changes in the Moses decoder and supporting programs to support such changes, while preserving the rich features in the Moses toolkit.
+In this paper, we describe the changes in the Moses decoder and supporting programs to support such changes, while preserving the toolkit's existing functionality.
 Particular emphasis will be place in describing the feature function framework.
@@
 
-Feature function state information is an essential component of the search procedure in decoding.
+Feature function \emph{state} information is an essential component of the search procedure in decoding.
 State information control hypothesis recombination, enabling the decoder to effectively search through many more hypotheses than it actually has.
-State information is emitted from stateful feature functions.
+Higher recombination results in higher effective number of hypotheses created, leading to better model score.
 
-The state information in Moses has evolved since its inception.
-Initially, state information was only available for language models, therefore, only language models states are compared, Figure~\ref{algo:CompareHypoStates}.
+Initially, state information was only available for language models, therefore, only language models states are compared.
+%, Figure~\ref{algo:CompareHypoStates}.
 
-\begin{figure} [h!]
-\small 
-\begin{algorithmic}
-\REQUIRE hypothesis $a$, hypothesis b
-       \IF {coverage of a != coverage of b }
-         \RETURN not equal
-       \ENDIF
-
-        \FORALL {LM state s}
-          \IF {state s of hypothesis a != state s for hypothesis b}
-            \RETURN not equal
-          \ENDIF
-        \ENDFOR
-        \RETURN equal
-\end{algorithmic}
-\caption{Comparing Hypothesis States (old framework)}
-\label{algo:CompareHypoStates}
-\end{figure}
+% \begin{figure} [h!]
+% \small 
+% \begin{algorithmic}
+% \REQUIRE hypothesis $a$, hypothesis b
+%        \IF {coverage of a != coverage of b }
+%          \RETURN not equal
+%        \ENDIF
+%        
+%         \FORALL {LM state s}
+%           \IF {state s of hypothesis a != state s for hypothesis b}
+%             \RETURN not equal
+%           \ENDIF
+%         \ENDFOR
+%         \RETURN equal
+% \end{algorithmic}
+% \caption{Comparing Hypothesis States (old framework)}
+% \label{algo:CompareHypoStates}
+% \end{figure}
 
@@
 
-Feature function \emph{state} plays an important role in hypothesis recombination.
-Two hypotheses can be recombined if their states are equivalent.
-Higher recombination results in increased effective number of hypotheses created, leading to better model score.
+Each feature function can be stateless or stateful, depending on whether it emits a state for each hypothesis.
+Stateful features inherits from \emph{StatefulFeatureFunction}, otherwise they inherits from \emph{StatelessFeatureFunction}.
 
-Each feature function can be stateless or stateful, depending on whether it emits a state for each hypothesis.
-Stateful features imherits from \emph{StatefulFeatureFunction}, otherwise they it inherits from \emph{StatelessFeatureFunction}.
-
-The state comparison routine of Figure~\ref{algo:CompareHypoStates} was generalised to compare the state of all stateful feature functions, Figure~\ref{algo:CompareHypoStates.new}.
+The state comparison of the original framework was generalised to compare the state of all stateful feature functions, Figure~\ref{algo:CompareHypoStates.new}.
 
@@
 \end{algorithmic}
-\caption{Comparing Hypothesis States (new framework)}
+\caption{Comparing Hypothesis States}
 \label{algo:CompareHypoStates.new}
@@
 \subsection{cdec and Joshua}
-cdec and Joshua both have modern feature function frameworks similar to the Moses new
-framework. Both of them have no distinction between stateful and stateless features; stateless
-features have zero size states.
+cdec and Joshua both have modern feature function frameworks. Both of them have no distinction between stateful and stateless features; stateless features have zero size states.
 
@@
 \end{enumerate}
-However, the stage at which the feature function is evaluate determines whether the feature function is cacheable between sentences.
-Scores that are evaluated at stage (1) can be cached and reused inter-sentence.
-Scores evaluated during stage (2) and (3) cannot be cached as they depend on the entire source sentence.
-Therefore, it is more efficient to evaluate as early as possible.
+However, the stage at which the feature function is evaluated determines whether the feature function needs to be evaluated, given a different sentence.
+Scores that are evaluated at stage (1) are only dependent the source phrase, not the entire source sentence, and therefore can be reused inter-sentence.
+Scores evaluated during stage (2) and (3) must be re-evaluated given a different sentence.
+Given a simple, source phrase-level cache, only feature function evaluated at stage (1) can be cached.
 
@@
 These feature is so intertwined with the search mechanism that their scores are set by search mechanism themselves.
-The feature function exist as an identifier to locate the score.
 Features functions such as the translation models, unknown word penalty,  generation models fall into this category.
@@
 
+Joshua's lower model score is likely due to the fact that in Moses and cdec, the calculation of language model state is delegated to the language model implementation~\cite{Heafield-left} which uses the internal data-structure of the language model to determine state.
+Joshua, on the other hand, uses the words in the left and right context to determine language model state.
+This results in more aggressive hypothesis recombination in Moses and cdec, leading to better model score.
+Joshua will soon be updated~\footnote{\url{https://github.com/hieuhoang/papers/commit/e417a6adcf4f1e49954836b14799509fecdea11c#mt-marathon.2013} } to incorporate better left-state minimisation to match the performance of Moses and cdec.
+
 \begin{figure}
@@
 
-Joshua's lower model score is likely due to the fact that in Moses and cdec, the calculation of language model state is delegated to the language model implementation~\cite{Heafield-left} which uses the internal data-structure of the language model to determine state.
-Joshua, on the other hand, uses the words in the left and right context to determine language model state.
-This results in more aggressive hypothesis recombination in Moses and cdec, leading to better model score.
-Joshua will soon be updated~\footnote{\url{https://github.com/hieuhoang/papers/commit/e417a6adcf4f1e49954836b14799509fecdea11c#mt-marathon.2013} } to incorporate better left-state minimisation to match the performance of Moses and cdec.
-
-\subsection{Glue Rules}
-
-The glue grammar can be constructed in a number of ways. In formulation (1), the grammar is:
-\begin{verbatim}
-    S -> <s>
-    S -> S X
-    GOAL -> S </s> 
-\end{verbatim}
- where <s> and </s> are first-class words like any other.
+% \subsection{Glue Rules}
+% 
+% The glue grammar can be constructed in a number of ways. In formulation (1), the grammar is:
+% \begin{verbatim}
+%     S -> <s>
+%     S -> S X
+%     GOAL -> S </s> 
+% \end{verbatim}
+%  where <s> and </s> are first-class words like any other.
 In this formulation, glued hypotheses know that they are bound to the beginning of sentence <s>.
 Therefore, their left language model state is empty and more hypotheses can recombine.
-
-The grammar for formulation (2) is:
-\begin{verbatim}
-    S -> X
-    S -> S X
-    GOAL -> <s> S </s> 
-\end{verbatim}
-Glued hypotheses are informed about <s> and </s> only after forming a complete sentence, so the left states are spuriously ambiguous.
+%  
+% The grammar for formulation (2) is:
+% \begin{verbatim}
+%     S -> X
+%     S -> S X
+%     GOAL -> <s> S </s> 
+% \end{verbatim}
+% Glued hypotheses are informed about <s> and </s> only after forming a complete sentence, so the left states are spuriously ambiguous.
 Moreover, hypothesis score estimates are more accurate in (1) as they know about <s> earlier.
-
-Figure~\ref{fig:glue} shows the improvement in cdec by changing the glue rule formulation from (2) to (1).
+% 
+% Figure~\ref{fig:glue} shows the improvement in cdec by changing the glue rule formulation from (2) to (1).
 The glue grammar for each decoder had varied since their inceptions.
@@
 (Result for cdec and Joshua in Figure~\ref{fig:model} were done with the improved glue grammar formulation.)
-
- \begin{figure}[h]
- \begin{center}
- \input{1111-8/glue_model.tex}
- \end{center}
- \caption{Affect on model score by changing glue-rule formulation}
- \label{fig:glue}
- \end{figure}
+% 
+%  \begin{figure}[h]
+%  \begin{center}
+%  \input{1111-8/glue_model.tex}
+%  \end{center}
+%  \caption{Affect on model score by changing glue-rule formulation}
+%  \label{fig:glue}
+%  \end{figure}
 

### 2013-07-18 mt-marathon.2013/PBML_article.tex
ken's suggestions

@@
 cdec also supports an arbitrary number of passes, feature functions can be assigned to a particular
-pass.
+pass. cdec's multi-pass strategy for feature function evaluation build on its course-to-fine framework.
 
@@
 \subsection{cdec}
-cdec eschews Moses' dense and sparse hybrid approach and implement only sparse maps. Dense features can still be modelled in the framework as sparse features, whose name are suffixed with indices.
+cdec eschews Moses' dense and sparse hybrid approach and implement only sparse maps.
+Dense features can still be modelled in the framework as sparse features, whose name are suffixed with indices.
+However, it optimizes sparse feature handling by using different data structures, depending on the number of feature scores.
 
-To improve speed and memory usage, hypotheses only store deltas of feature differences with antecedent hypotheses.
+Also, to improve speed and memory usage, hypotheses only store features deltas with respect to antecedent hypotheses.
 
@@
 
-Hypotheses only retain the weighted sum of the feature scores, .
+Hypotheses only retain the weighted sum of the feature scores.
 The individual score components are not stored unless required, for example, in the n-best list.

### 2013-07-19 mt-marathon.2013/PBML_article.tex
add appendix on how to create a new ff

@@
 \usepackage{tikz,gnuplot-lua-tikz} 
+\usepackage{nameref}
 
@@
 \end{figure}
+
+A detailed example of how to add a new feature function to the Moses decoder is contained in\emph{~\nameref{Appendix}}.
 
@@
 
+\section*{Appendix: Adding a New Feature Function to Moses, an Example}
+\label{Appendix}
+
+The section describes how to add a new feature function to the current (July 2013) version of Moses, and use it during decoding.
+
+\begin{enumerate}
+    \item Create a new class in the decoder source code. By convention, the files should be in the folder or subfolder of either
+\begin{verbatim}
+   moses/FF
+   moses/LM
+   moses/TranslationModel
+\end{verbatim}
+Inherit the class from either of
+\begin{verbatim}
+  StatelessFeatureFunction
+  StatefulFeatureFunction
+\end{verbatim}
+or any class that inherit from one of these classes. Other common classes to inherit from are:
+\begin{verbatim}
+  PhraseDictionary
+  LanguageModelImplementation
+\end{verbatim}
+
+    \item Create a constructor of the form
+\begin{verbatim}
+  ClassName(const std::string &line);
+\end{verbatim}
+
+    \item Implement the method
+\begin{verbatim}
+  bool IsUseable(const FactorMask &mask) const;
+\end{verbatim}
+The argument is a bit vector of which factors are available.
+The method should return \emph{true} if it can be evaluated with these factors, otherwise return \emph{false}.
+If the feature does not use any factors, always return  \emph{true}.
+
+The constructor, and this method, are the only mandatory function for feature functions. All others methods override default methods (that often do nothing).
+
+    \item Useful methods to override:
+      \begin{enumerate}
+        \item If the feature needs to load data, override the method
+	  \begin{verbatim}
+  void Load();
+	  \end{verbatim}
+        \item If the feature functions takes arguments
+	  \begin{verbatim}
+  void SetParameter(const std::string& key, const std::string& value);
+	  \end{verbatim}
+\emph{name}, \emph{num-features}, \emph{tuneable} are reserved keys that the new feature should not use.
+	\item If it requires initialisation and cleaning after each sentence (this is not particularly thread-safe):
+	  \begin{verbatim}
+  void InitializeForInput(InputType const& source);
+  void CleanUpAfterSentenceProcessing(const InputType& source);
+	  \end{verbatim}
+       \end{enumerate}
+    \item Override one or more of the Evaluate() method. 
+      \begin{verbatim}
+  void Evaluate(const Phrase &source
+	    , const TargetPhrase &targetPhrase
+	    , ScoreComponentCollection &scoreBreakdown
+	    , ScoreComponentCollection &estimatedFutureScore) const;
+  void Evaluate(const InputType &source
+            , ScoreComponentCollection &scoreBreakdown) const;
+       \end{verbatim}
+Also for stateless features:
+      \begin{verbatim}
+  void Evaluate(const PhraseBasedFeatureContext& context,
+                ScoreComponentCollection* accumulator) const;
+  void EvaluateChart(const ChartBasedFeatureContext& context,
+                ScoreComponentCollection* accumulator) const;
+       \end{verbatim}
+The method signature is similar for stateful features, but the methods return a state object:
+      \begin{verbatim}
+  FFState* Evaluate(const Hypothesis& cur_hypo,
+		    const FFState* prev_state,
+		    ScoreComponentCollection* accumulator) const;
+  FFState* EvaluateChart(const ChartHypothesis&,
+		      int featureID,
+		      ScoreComponentCollection* accumulator) const;
+      \end{verbatim}
+    \item Register the new feature function with the decoder by adding the following line to \emph{Factory.cpp}:
+      \begin{verbatim}
+  MOSES_FNAME(ClassName);
+      \end{verbatim}
+    \item To instantiate the feature during decoding, add the following line to the \emph{[feature]} section of the moses.ini file:
+      \begin{verbatim}
+  ClassName [key=value]*
+      \end{verbatim}
+If it has dense features, the weights should be in the \emph{[weight]} section, for example:
+      \begin{verbatim}
+  ClassName0= 0.1 0.2 0.3
+      \end{verbatim}
+
+\end{enumerate}
+
 \correspondingaddress

### 2013-08-14 amta.2012/paper.tex
start mt summit slides

%% Sample file $Id: PBML_article.tex 134 2010-07-01 19:28:04Z popel $
\documentclass{pbml}
%\documentclass[nofonts]{pbml} % for XeLaTeX without Pagella and DejaVu fonts

\usepackage{color,fancyvrb}
\definecolor{gray}{rgb}{0.9,0.9,0.9}
\usepackage{amsmath}
%\renewcommand*\ttdefault{txtt} % 20% tighter than courier

% This is a sample file for the PBML article.
% You can compile it with (ordered by preference)
%  1. XeLaTeX with installed fonts TeX Gyre Pagella and DejaVu
%  2. XeLaTeX without the fonts -> you must use \documentclass[nofonts]{pbml}
%  3. pdfLaTeX
%
% In all three methods, you can use Unicode (utf8) encoding for special letters
% (so instead of na\"{i}ve you can write directly naïve).
% Note that with methods 2 and 3, your output will be slightly DIFFERENT than
% our final print version (different line breaks and page breaks).
% See the end of this file for more information.

% Packages
% ========
% In this place you can load required packages by the \usepackage commands.
% The following packages are loaded automatically by the class:
% euler    (for math fonts)
% graphicx (for inclusion of images)
% multicol (for multicolumn typesetting)
% natbib   (for bibliography citations)
% amssymb  (for various symbols)
% If XeLaTeX is used, fontspec and xltxtra are loaded as well.

% Definitions
% ===========
% You can use your own macros defined by \newcommand, \providecommand,
% \DeclareRobustCommand (or even by \def) as well as environments declared by
% \newenvironment. These definitions will automatically be local to your
% article and will not clash with other parts of the journal. You can also
% use \newcounter in order to declare your own counters.

% This switches to color version and is explained below (sec. Color Images).
% \graphicspath{{./color/}{./grayscale/}}

\begin{document}

% Document title and authors
% ==========================
% Due to journal organization, the article title and authors must be specified
% AFTER \begin{document}. Note that \subtitle is not allowed anymore due to
% problems with indexing in science databases. If needed use colon in the title.

\title{An Experimental Management System}

% Now put the affiliated institutes first, then author names and the labels
% of their institutes in the "institute" field. The order of institutes and
% authors printed below the title will be the same as the order of your commands.
% Each author can be associated with more than one institute.
% One author must be chosen as the "corresponding author" (using attribute
% corresponding) and his or her email and full address must be provided.
% PLEASE, use Unicode (utf8) encoding (e.g ï instead of \"{i}).

\institute{label1}{University of Edinburgh}

\author{
  firstname=Philipp,
  surname=Koehn,
  institute=label1,
  corresponding=yes,
  email={pkoehn@inf.ed.ac.uk},
  address={School of Informatics\\ University of Edinburgh\\ Scotland, United Kingdom}
}

% If all authors belong to the same institute, you can use simpler syntax:
% \institute{}{Institute of Formal and Applied Linguistics, Charles University in Prague}
% \author{firstname=Humpty, surname=Dumpty}
% \author{firstname=Mock, surname=Turtle,
%   corresponding=yes,
%   email={turtle@seacoast.wl},
%   address={Institute of Formal and Applied Linguistics\\
%            Charles University in Prague\\
%            Malostranské náměstí 25\\
%            118 00 Praha 1, Czech Republic}}
% \author{firstname=Cheshire, surname=Cat}

% The title and authors' names are used in the running head. If they are
% long, you should define short versions. These definitions are optional. You
% define them only if they are needed. The example follows:
\shorttitle{Experimental Management System}
\shortauthor{P. Koehn}

% Now print the title by:
\maketitle

% Abstract
% ========
% The abstract is placed within the "abstract" environment. It is a mandatory
% part of the article.

\begin{abstract}
We describe Experiment.perl, an experimental management system, that allows the execution of the entire training and testing pipeline of a machine translation experiment with one configuration files. When carrying out multiple experimental runs with changed settings, Experiment.perl automatically detects which steps need to be re-run and which can be re-used.
\end{abstract}

% In addition to the \cite command, you can use natbib style of citations:
%\citet{PDT2} shows \ldots \\
%\ldots tree-based annotation (as e.g.~\citealp{PDT2}).

% For figures and tables always use "figure" resp. "table" floating environments
% and always supply a caption. See the paragraph about color images below.

\section{\label{experiment-perl.texntoc1}Introduction}

Running a machine translation experiment involves many steps: preparing training data, building language and translation models, tuning, testing, scoring and analysis of the results. For most of these steps, a different tool needs to be invoked, so this easily becomes very cumbersome.
The Experiment Management System (EMS), or Experiment.perl, for lack of a better name, makes it much easier to run experiments. It ships with the Moses machine translation toolkit \citep{koehn-EtAl:2007:PosterDemo}.

\begin{figure}
\begin{center}
\includegraphics[scale=0.45]{ems-agenda.png}
\end{center}
\caption{{\bf Workflow generated by Experiment.perl:} Steps (such as {\tt run-giza} are grouped into modules (such as {\tt TRAINING}). Dependencies are indicated by arrows. Skipped steps have a pale background.}
\label{ems-agenda}
\end{figure}

A typical example is given in Figure~\ref{ems-agenda}.
The graph was automatically generated by Experiment.perl. All that needed to be done was to specify one single configuration file that points to data files and settings for the experiment.
In the graph, each step is a small box. For each step, Experiment.perl builds a script file that gets either submitted to a compute cluster or executed on the same machine. Note that some steps are quite involved, for instance tuning: On a cluster, the tuning script runs on the head node a submits jobs to the queue itself.

Experiment.perl makes it easy for multiple experimental runs with different settings]. It automatically detects which steps do not have to be executed again.
Experiment.perl plays the same role as LoonyBin \citep{loonybin}, but there are significant differences in its usage. Experiment.perl uses a textual configuration file to set up an experiment and a meta configuration file to define all possible workflows, LoonyBin offers a graphical user interface that lets the user connect steps.

\section{Design}

Experiment.perl breaks up training, tuning, and evaluating of a statistical machine translation system into a number of steps, which are then scheduled to run in parallel or sequence depending on their inter-dependencies and available resources. The possible steps are defined in the file {\tt experiment.meta}. An experiment is defined by a configuration file.

\subsection{Experiment.Meta}

The actual steps, their dependencies and other salient information is to be found in the file {\tt experiment.meta}. Think of {\tt experiment.meta} as a "template" file.
Steps are grouped into modules, which are:

\begin{itemize}
\item {\tt CORPUS}: preparing a parallel corpus

\item {\tt INPUT-FACTOR} and {\tt OUTPUT-FACTOR}: commands to create factors

\item {\tt TRAINING}: training a translation model

\item {\tt LM}: training a language model

\item {\tt INTERPOLATED-LM}: interpolate language models

\item {\tt SPLITTER}: training a word splitting model

\item {\tt RECASING}: training a recaser

\item {\tt TRUECASING}: training a truecaser

\item {\tt TUNING}: running minumum error rate training to set component weights

\item {\tt TESTING}: translating and scoring a test set

\item {\tt REPORTING}: compile all scores in one file

\end{itemize}

To give an example of a step definition in {\tt experiment.meta}, here the parts of the definition for {\tt LM:get-corpus} and {\tt LM:tokenize}:

\vspace{2mm}
\begin{SaveVerbatim}{myverb}
 get-corpus
        in: get-corpus-script
        out: raw-corpus
        [...]

 tokenize
        in: raw-corpus
        out: tokenized-corpus
        [...]
\end{SaveVerbatim}
\colorbox{gray}{%
\BUseVerbatim{myverb}
}
\vspace{2mm}

Each step takes some input ({\tt in}) and provides some output ({\tt out}). This also establishes the dependencies between the steps. The step {\tt tokenize} requires the input {\tt raw-corpus}. This is provided by the step {\tt get-corpus}.

The file {\tt experiment.meta} provides a generic template for steps and their interaction. For an actual experiment, a configuration file determines which steps need to be run. This configuration file is specified when invoking {\tt experiment.perl}. It may contain for instance the following:

\vspace{2mm}
\begin{SaveVerbatim}{myverb}
 [LM:europarl]

 ### raw corpus file
 #
 raw-corpus = $europarl-v3/training/europarl-v3.en
\end{SaveVerbatim}
\colorbox{gray}{%
\BUseVerbatim{myverb}
}
\vspace{2mm}

Here, the corpus to be used for language modeling is named {\tt europarl} and it is provided in raw text format in the location {\tt \$europarl-v3/training/europarl-v3.en} (the variable {\tt \$europarl-v3} is defined elsewhere in the config file). The effect of this specification in the config file is that the step {\tt get-corpus} does not need to be run, since its output is given as a file. The workflow starts with the step {\tt tokenize}.

The entire definition of an experiment follows this logic, which is very similar to the principles of a Unix Makefile. The ultimate purpose of an experiment is to generate a result file at the end. If this is not given, then scoring scripts need to be called. Scoring scripts require the output of the decoder on the test sets. The decoder requires a tuned model. Tuning requires a trained model. Trained models require language models and model files, and so on. The workflow, as show in Figure~\ref{ems-agenda}, is generate bottom up, following the input/output dependencies of steps.

\subsection{Elements of Step Definitions}

Several parameters for step definitions are used in {\tt experiment.meta}:

\begin{itemize}
\item {\tt in} and {\tt out}: Established dependencies between steps; input may also be provided by files specified in the configuration.

\item {\tt default-name}: Name of the file in which the output of the step will be stored.

\item {\tt template}: Template for the command that is placed in the execution script for the step.

\item {\tt template-if}: Potential command for the execution script. Only used, if a specified setting exists.

\item {\tt error}: Experiment.perl detects if a step failed by scanning {\tt STDERR} for key words such as killed, error, died, not found, and so on. Additional key words and phrase are provided with this parameter.

\item {\tt not-error}: Declares default error key words as not indicating failures.

\item {\tt pass-unless}: Only if the given setting is used, this step is executed, otherwise the step is passed (illustrated by a yellow box in the graph).

\item {\tt ignore-unless}: If the given setting is used, this step is not executed. This overrides requirements of downstream steps.

\item {\tt rerun-on-change}: If similar experiment are runs, the output of steps may be used, if input and settings are the same. This specifies settings whose change disallows a re-use in different run.

\item {\tt parallelizable}: When running on a cluster or a multi-core machine, this step may be parallelized (only if {\tt generic-parallelizer} is set in the config file).

\item {\tt qsub-script}: If running on a cluster, this step is run on the head node, and not submitted to the queue (because it submits jobs itself).

\end{itemize}

To complete our example, the full definition of the step {\tt LM:tokenize} is below.

\vspace{2mm}
\begin{SaveVerbatim}{myverb}
 tokenize
         in: raw-corpus
         out: tokenized-corpus
         default-name: lm/tok
         pass-unless: output-tokenizer
         template: $output-tokenizer < IN > OUT
         parallelizable: yes
\end{SaveVerbatim}
\colorbox{gray}{%
\BUseVerbatim{myverb}
}
\vspace{2mm}

The step takes {\tt raw-corpus} and produces {\tt tokenized-corpus}. It is parallizable with the generic parallelizer. 
The output is stored in the file according to the definition {\tt corpus/tok}. Note that the actual file name also contains the corpus name, and the run number. In our example the tokenized corpus is stored in a file named  {\tt lm/europarl.tok.1}. 
The step is only executed, if {\tt output-tokenizer} is specified. The template indicate how the command lines in the execution script for the steps are formed.

\subsection{Multiple Corpora, One Translation Model}

We may use multiple parallel corpora for training a translation model or multiple monolingual corpora for training a language model (or use multiple language models). Each of these have their own instances of the {\tt CORPUS} and {\tt LM} module. There may be also multiple test sets in {\tt TESTING}). However, there is only one translation model and hence only one instance of the {\tt TRAINING} module. 
The definitions in {\tt experiment.meta} reflects the different nature of these modules. For instance {\tt CORPUS} is flagged as {\tt multiple}, while {\tt TRAINING} is flagged as {\tt single}.

When defining settings for the different modules, the singular module {\tt TRAIN\-ING} has only one section, while this one general section and specific {\tt LM} sections for each training corpus. In the specific section, the corpus is named, e.g. {\tt LM:europarl}.
When looking up the parameter settings for a step, first the set-specific section ({\tt LM:europarl}) is consulted. If there is no definition, then the module definition ({\tt LM}) and finally the general definition (in section {\tt GENERAL}) is consulted. In other words, local settings override global settings.

\subsection{Configuration File}

A configuration file for an experimental run consists of a collection of settings, one per line with empty lines and comment lines for better readability, organized in sections for each of the modules.

The start of each section is indicated by the section name in square brackets ({\tt [TRAINING]} or {\tt [CORPUS:europarl]}). If the word {\tt IGNORE} is appended to a section definition, then the entire section is ignored.

The syntax of setting definition is {\tt setting = value} (note: spaces around the equal sign). If the value contains spaces, it must be placed into quotes ({\tt setting = "the value"}), except when a vector of values is implied (only used when defining list of factors: {\tt output-factor = word pos}. 
Comments are indicated by a hash ({\tt \#}).

Settings can be used as variables to define other settings:

\vspace{2mm}
\begin{SaveVerbatim}{myverb}
 working-dir = /home/pkoehn/experiment
 wmt10-data = $working-dir/data
\end{SaveVerbatim}
\colorbox{gray}{%
\BUseVerbatim{myverb}
}
\vspace{2mm}

Variable names may be placed in curly brackets for clearer separation:

\vspace{2mm}
\begin{SaveVerbatim}{myverb}
 wmt10-data = ${working-dir}/data
\end{SaveVerbatim}
\colorbox{gray}{%
\BUseVerbatim{myverb}
}
\vspace{2mm}

Such variable references may also reach other modules:

\vspace{2mm}
\begin{SaveVerbatim}{myverb}
 [RECASING]
 tokenized = $LM:europarl:tokenized-corpus
\end{SaveVerbatim}
\colorbox{gray}{%
\BUseVerbatim{myverb}
}
\vspace{2mm}

Finally, reference can be made to settings that are not defined in the configuration file, but are the product of the defined sequence of steps. 
Say, in the above example, {\tt tokenized-corpus} is not defined in the section {\tt LM:europarl}, but instead {\tt raw-corpus}. Then, the tokenized corpus is produced by the normal processing pipeline. Such an intermediate file can be used elsewhere:

\vspace{2mm}
\begin{SaveVerbatim}{myverb}
 [RECASING]
 tokenized = [LM:europarl:tokenized-corpus]
\end{SaveVerbatim}
\colorbox{gray}{%
\BUseVerbatim{myverb}
}
\vspace{2mm}

Some error checking is done on the validity of the values in the configuration file before an experimental run is executed. All values that seem to be file paths trigger the existence check for such files. A file with the prefix of the value must exist.

\subsection{Step Files}
Let us follow our example of the tokenization step in the language model module in more detail. Recall that the {\tt LM:europarl} section has a specification of {\tt raw-corpus} to {\tt \$europarl-v3/train\-ing/europarl-v3.en}. Since only the raw corpus, but not a tokenized corpus is specified, Experiment.perl concludes that it needs to run the tokenization step.

The directory {\tt steps} contains the script that executes each step, its {\sc stderr} and {\sc stdout} output, and meta information:

\vspace{2mm}
\begin{SaveVerbatim}{myverb}
 steps/1/LM_europarl_tokenize.1
 steps/1/LM_europarl_tokenize.1.DONE
 steps/1/LM_europarl_tokenize.1.INFO
 steps/1/LM_europarl_tokenize.1.STDERR 
 steps/1/LM_europarl_tokenize.1.STDERR.digest
 steps/1/LM_europarl_tokenize.1.STDOUT
\end{SaveVerbatim}
\colorbox{gray}{%
\BUseVerbatim{myverb}
}
\vspace{2mm}

The file {\tt steps/1/LM\_europarl\_tokenize.1} is the shell script that is run to execute the step. The file with the extension {\tt DONE} is created when the step is finished - this communicates to the scheduler that subsequent steps can be executed. The file with the extension {\tt INFO} contains meta information - essential the settings and dependencies of the step. This file is checked to detect if a step can be re-used in subsequent experimental runs.

In case that the step crashed, we expect some indication of a fault in {\tt STDERR} (for instance the words {\tt core dumped} or {\tt killed}). This file is checked to see if the step was executed successfully, so subsequent steps can be scheduled or the step can be re-used in new experiments. Since the {\tt STDERR} file may be very large (some steps create Megabytes of such output), a digested version is created in {\tt STDERR.digest}. If the step was successful, it is empty. Otherwise it contains the error pattern that triggered the failure detection.

\subsection{Re-Use of Steps}

Let us now take a closer look at re-use. If we run the experiment again but change a settings, say, the order of the language model, then there is no need to re-run the tokenization, but only language model training.

Here is the definition of the language model training step in {\tt experiment.meta}:

\vspace{2mm}
\begin{SaveVerbatim}{myverb}
 train
   in: split-corpus
   out: lm
   default-name: lm/lm
   ignore-if: rlm-training
   rerun-on-change: lm-training order settings
   template: $lm-training -order $order $settings -text IN -lm OUT
   error: cannot execute binary file
\end{SaveVerbatim}
\colorbox{gray}{%
\BUseVerbatim{myverb}
}
\vspace{2mm}

The mention of {\tt order} in the list behind {\tt rerun-on-change} informs experiment.perl that this step does need to be re-run, if the order of the language model changes. Since none of the settings in the chain of steps leading up to the training have been changed, those can be re-used.

If the language model order is changed and the experiment.perl is run again in the same working directory, you will see the following files in the directory {\tt lm}:

\vspace{2mm}
\begin{SaveVerbatim}{myverb}
 % ls -tr lm/*
 lm/europarl.tok.1
 lm/europarl.truecased.1
 lm/europarl.lm.1
 lm/europarl.lm.2
\end{SaveVerbatim}
\colorbox{gray}{%
\BUseVerbatim{myverb}
}
\vspace{2mm}

Note that a new language model was trained for this second run ({\tt lm/europarl.lm.2}), but no new tokenized and truecased corpus files. These were re-used from run 1.

Steps are re-used from previous runs, unless settings listed under {\tt rerun-on\--change} are changed, one of its specified input files (if any) are changed, and if one of its previous steps are re-run. Note that if a filename is not changed, but its time stamp differs, this triggers re-running a step. This ensures that only a minimum number of steps are run to produce the exact same outcome as if all steps are run.

\subsection{Web Interface and Analysis}

Experiment.perl also offers a web interface to the experimental runs for easy access and comparison of experimental results. The web interface gives a listing of experiments and runs for each experiments, with a display of automatic metric scores, and links to configuration files and outputs.

You can include additional analysis for an experimental run in the web interface by specifying the setting {\tt analysis} in its configuration file.
This adds reports n-gram precision and recall statistics and color-coded n-gram correctness markup for the output sentences to the web interface. See Figure~\ref{ems-annotation} for an example.

\begin{figure}
\begin{center}
\includegraphics[scale=0.45]{ems-annotation.png}
\end{center}
\caption{Comparing outputs from two experimental runs}
\label{ems-annotation}
\end{figure}

The output is color-highlighted according to n-gram matches with the reference translation. The following colors are used. The darker the color of an output word, the higher n-gram match to the reference translation it is part of.

Additional reports are available when adding the  settings {\tt analyze-coverage} and {\tt report-segmentation}.
The setting {\tt analyze-coverage} include a coverage analysis: which words and phrases in the input occur in the training data or the translation table? This is reported in color coding and in a yellow report box when moving the mouse of the word or the phrase. Also, summary statistics for how many words occur how often are given, and a report on unknown or rare words is generated.

The setting {\tt report-segmentation} creates summary statistics about what kind of phrase mappings are used (one word to one word, 1-2, 2-2, etc.), as well as markup of the sentence pair with the phrase segmentation. The phrase segmentation is indicated with black boxes around the words, and the alignment is shown when moving the mouse on the phrases.

\section{Usage}

\subsection{Quick Start}

Experiment.perl is extremely simple to use:

\begin{itemize}
\item Find {\tt experiment.perl} in {\tt scripts/ems}

\item Get a sample configuration file from someplace (for instance {\tt scripts/ems/exam\-ple/config.toy}).

\item Set up a working directory for your experiments for this task ({\tt mkdir} does it).

\item Edit the following path settings in {\tt config.toy}
\begin{itemize}
\item {\tt working-dir}

\item {\tt data-dir}

\item {\tt moses-script-dir}

\item {\tt moses-src-dir}

\item {\tt srilm-dir}

\item {\tt decoder}

\end{itemize}

\item Run {\tt experiment.perl -config config.toy} from your working directory.

\item Marvel at the graphical plan of action.

\item Run {\tt experiment.perl -config config.toy -exec}.

\item Check the results of your experiment (in {\tt evaluation/report.1})

\end{itemize}

\subsection{More Examples}

The {\tt example} directory contains some additional examples. 
These require the training and tuning data released for the Shared Translation Task for WMT 2010.

The examples using these corpora are

\begin{itemize}
\item a basic phrase based model

\item a factored phrase based model

\item a hierarchical phrase based model

\item a target syntax model

\end{itemize}

The factored model using all the available corpora is identical to the Edinburgh submission \citep{koehn-EtAl:2010:WMT} to the WMT 2010 shared task for English-Spanish, Spanish-English, and English-German language pairs. The French language pairs also used the $\text{10}^\text{9}$ corpus, the Czech language pairs did not use the POS language model, and German-English used additional pre-processing steps.

\section{Outlook}
We have been using Experiment.perl for years and are satisfied with its core functionalities. In future work, we would like support job scheduling on Hadoop clusters and extend the analysis facility.

\section*{Acknowledgments}
This work was supported in part by the EuroMatrixPlus project funded by the 
 European Commission (7th Framework Programme) and in part under the GALE program of the 
 Defense Advanced Research Projects Agency, Contract No. HR0011-06-C-0022.

\bibliography{more,mt,WMT-2010}

\end{document}

### 2014-12-21 iit-bombay/experiment-perl.tex
hieu's slides

\label{experiment-perl.tex}

\section{
Experiment Management System}

      {\bf $\bullet$} Introduction (page \pageref{experiment-perl.texntoc1})

      {\bf $\bullet$} Requirements (page \pageref{experiment-perl.texntoc2})

      {\bf $\bullet$} Quick Start (page \pageref{experiment-perl.texntoc3})

      {\bf $\bullet$} More Examples (page \pageref{experiment-perl.texntoc4})

      {\bf $\bullet$} Phrase Model (page \pageref{experiment-perl.texntoc5})

      {\bf $\bullet$} Factored Phrase Model (page \pageref{experiment-perl.texntoc6})

      {\bf $\bullet$} Hierarchical model (page \pageref{experiment-perl.texntoc7})

      {\bf $\bullet$} Target syntax model (page \pageref{experiment-perl.texntoc8})

      {\bf $\bullet$} Try a Few More Things (page \pageref{experiment-perl.texntoc9})

      {\bf $\bullet$} Stemmed Word Alignment (page \pageref{experiment-perl.texntoc10})

      {\bf $\bullet$} Using Multi-Threaded GIZA++ (page \pageref{experiment-perl.texntoc11})

      {\bf $\bullet$} Using Berkeley Aligner (page \pageref{experiment-perl.texntoc12})

      {\bf $\bullet$} IRST Language Model (page \pageref{experiment-perl.texntoc13})

      {\bf $\bullet$} Randomized Language Model (page \pageref{experiment-perl.texntoc14})

      {\bf $\bullet$} Compound Splitting (page \pageref{experiment-perl.texntoc15})

      {\bf $\bullet$} A Short Manual (page \pageref{experiment-perl.texntoc16})

      {\bf $\bullet$} Experiment.Meta (page \pageref{experiment-perl.texntoc17})

      {\bf $\bullet$} Multiple Corpora, One Translation Model (page \pageref{experiment-perl.texntoc18})

      {\bf $\bullet$} Defining Settings (page \pageref{experiment-perl.texntoc19})

      {\bf $\bullet$} Working with Experiment.Perl (page \pageref{experiment-perl.texntoc20})

      {\bf $\bullet$} Continuing Crashed Experiments (page \pageref{experiment-perl.texntoc21})

      {\bf $\bullet$} Running on a Cluster (page \pageref{experiment-perl.texntoc22})

      {\bf $\bullet$} Running on a Multicore Machine (page \pageref{experiment-perl.texntoc23})

      {\bf $\bullet$} Web Interface (page \pageref{experiment-perl.texntoc24})

      {\bf $\bullet$} Analysis (page \pageref{experiment-perl.texntoc25})

\subsection{
\label{experiment-perl.texntoc1}Introduction}

The Experiment Management System (EMS), or Experiment.perl, for lack of a better name, makes it much easier to perform experiments with Moses.

There are many steps in running an experiment: the preparation of training data, building language and translation models, tuning, testing, scoring and analysis of the results. For most of these steps, a different tool needs to be invoked, so this easily becomes very messy.

Here a typical example:

\begin{center}
\includegraphics[scale=0.5]{images/ems-agenda.png}
\end{center}

This graph was automatically generated by Experiment.perl. All that needed to be done was to specify one single configuration file that points to data files and settings for the experiment.

In the graph, each step is a small box. For each step, Experiment.perl builds a script file that gets either submitted to the cluster or run on the same machine. Note that some steps are quite involved, for instance tuning: On a cluster, the tuning script runs on the head node a submits jobs to the queue itself.

Experiment.perl makes it easy to run multiple experimental runs with different settings or data resources. It automatically detects which steps do not have to be executed again but instead which results from an earlier run can be re-used.

Experiment.perl also offers a web interface to the experimental runs for easy access and comparison of experimental results.

\begin{center}
\includegraphics[scale=0.5]{images/ems-web.png}
\end{center}

The web interface also offers some basic analysis of results, such as comparing the n-gram matches between two different experimental runs:

\begin{center}
\includegraphics[scale=0.5]{images/ems-analysis.png}
\end{center}

\subsection{
\label{experiment-perl.texntoc2}Requirements}

In order to run properly, EMS will require:

\begin{itemize}
\item A version of Moses along with SRILM

\item The GraphViz toolkit\footnote{\sf http://www.graphviz.org/}.

\item The ImageMagick toolkit\footnote{\sf http://www.imagemagick.org/script/index.php}.

\item The GhostView tool\footnote{\sf http://www.gnu.org/software/gv/}.

\end{itemize}

\subsection{
\label{experiment-perl.texntoc3}Quick Start}

Experiment.perl is extremely simple to use:

\begin{itemize}
\item Find {\tt experiment.perl} in {\tt scripts/ems}

\item Get a sample configuration file from someplace (for instance {\tt scripts/ems/example/config.toy}).

\item Set up a working directory for your experiments for this task ({\tt mkdir} does it).

\item Edit the following path settings in {\tt config.toy}
\begin{itemize}
\item {\tt working-dir}

\item {\tt data-dir}

\item {\tt moses-script-dir}

\item {\tt moses-src-dir}

\item {\tt srilm-dir}

\item {\tt decoder}

\end{itemize}

\item Run {\tt experiment.perl -config config.toy} from your experiment working directory.

\item Marvel at the graphical plan of action.

\item Run {\tt experiment.perl -config config.toy -exec}.

\item Check the results of your experiment (in {\tt evaluation/report.1})

\end{itemize}

Let us take a closer look at what just happened.

The configuration file {\tt config.toy} consists of several sections. For instance there is a section for each language model corpus to be used. In our toy example, this section contains the following:

\begin{SaveVerbatim}{myverb}
 [LM:toy]

 ### raw corpus (untokenized)
 #
 raw-corpus = $toy-data/nc-5k.$output-extension
\end{SaveVerbatim}
\colorbox{gray}{%
\BUseVerbatim{myverb}
}

The setting {\tt raw-corpus} species the location of the corpus. The definition uses the variables {\tt \$toy-data} and {\tt \$output-extension}, which are also settings defined elsewhere in the configuration file. These variables are resolved, leading to the file path {\tt ems/examples/data/nc-5k.en} in your Moses scripts directory.

The authoritative definition of the steps and their interaction is in the file {\tt experiment.meta} (in the same directory as experiment.perl: {\tt scripts/ems}).

The logic of {\tt experiment.meta} is that it wants to create a report at the end. To generate the report it needs to evaluation scores, to get these it needs decoding output, to get these it needs to run the decoder, to be able to run the decoder it needs a trained model, to train a model it needs data. This process of defining the agenda of steps to be executed is very similar to the Make utility in Unix.

We can find the following step definitions for the language model module in {\tt experiment.meta}:

\begin{SaveVerbatim}{myverb}
 get-corpus
        in: get-corpus-script
        out: raw-corpus
        default-name: lm/txt
        template: IN > OUT
 tokenize
        in: raw-corpus
        out: tokenized-corpus
        default-name: lm/tok
        pass-unless: output-tokenizer
        template: $output-tokenizer < IN > OUT
        parallelizable: yes
\end{SaveVerbatim}
\colorbox{gray}{%
\BUseVerbatim{myverb}
}

The tokenization step {\tt tokenize} requires {\tt raw-corpus} as input. In our case, we specified the setting in the configuration file. We could have also specified an already tokenized corpus with {\tt tokenized-corpus}. This would allow us to skip the tokenization step. Or, to give another example, we could have not specified {\tt raw-corpus}, but rather specify a script that generates the corpus with the setting {\tt get-corpus-script}. This would have triggered the creation of the step {\tt get-corpus}.

The steps are linked with the definition of their input {\tt in} and output {\tt out}. Each step has also a default name for the output ({\tt efault-name}) and other settings.

The tokenization step has as default name {\tt lm/tok}. Let us look at the directory {\tt lm} to see which files it contains:

\begin{SaveVerbatim}{myverb}
 % ls -tr lm/*
 lm/toy.tok.1
 lm/toy.truecased.1
 lm/toy.lm.1
\end{SaveVerbatim}
\colorbox{gray}{%
\BUseVerbatim{myverb}
}

We find the output of the tokenization step in the file {\tt lm/toy.tok.1}. The {\tt toy} was added from the name definition of the language model (see {\tt [LM:toy]} in {\tt config.toy}). The {\tt 1} was added, because this is the first experimental run.

The directory {\tt steps} contains the script that executes each step, its STDERR and STDOUT output, and meta information. For instance:

\begin{SaveVerbatim}{myverb}
 % ls steps/1/LM_toy_tokenize.1* | cat
 steps/1/LM_toy_tokenize.1
 steps/1/LM_toy_tokenize.1.DONE
 steps/1/LM_toy_tokenize.1.INFO
 steps/1/LM_toy_tokenize.1.STDERR
 steps/1/LM_toy_tokenize.1.STDERR.digest
 steps/1/LM_toy_tokenize.1.STDOUT
\end{SaveVerbatim}
\colorbox{gray}{%
\BUseVerbatim{myverb}
}

The file {\tt steps/2/LM\_toy\_tokenize.2} is the script that is run to execute the step. The file with the extension {\tt DONE} is created when the step is finished - this communicates to the scheduler that subsequent steps can be executed. The file with the extension {\tt INFO} contains meta information - essential the settings and dependencies of the step. This file is checked to detect if a step can be re-used in new experimental runs.

In case that the step crashed, we expect some indication of a fault in {\tt STDERR} (for instance the words {\tt core dumped} or {\tt killed}). This file is checked to see if the step was executed successfully, so subsequent steps can be scheduled or the step can be re-used in new experiments. Since the {\tt STDERR} file may be very large (some steps create Megabytes of such output), a digested version is created in {\tt STDERR.digest}. If the step was successful, it is empty. Otherwise it contains the error pattern that triggered the failure detection.

Let us now take a closer look at re-use. If we run the experiment again but change some of the settings, say, the order of the language model, then there is no need to re-run the tokenization.

Here is the definition of the language model training step in {\tt experiment.meta}:

\begin{SaveVerbatim}{myverb}
 train
        in: split-corpus
        out: lm
        default-name: lm/lm
        ignore-if: rlm-training
        rerun-on-change: lm-training order settings
        template: $lm-training -order $order $settings -text IN -lm OUT
        error: cannot execute binary file
\end{SaveVerbatim}
\colorbox{gray}{%
\BUseVerbatim{myverb}
}

The mention of {\tt order} in the list behind {\tt rerun-on-change} informs experiment.perl that this step does need to be re-run, of the order of the language model changes. Since none of the settings in the chain of steps leading up to the training have been changed, the can be re-used.

Try changing the language model order ({\tt order = 5} in {\tt config.toy}), run experiment.perl again ({\tt experiment.perl -config config.toy}) in the working directory, and you will see the new language model in the directory {\tt lm}:

\begin{SaveVerbatim}{myverb}
 % ls -tr lm/*
 lm/toy.tok.1
 lm/toy.truecased.1
 lm/toy.lm.1
 lm/toy.lm.2
\end{SaveVerbatim}
\colorbox{gray}{%
\BUseVerbatim{myverb}
}

\subsection{
\label{experiment-perl.texntoc4}More Examples}

The {\tt example} directory contains some additional examples.

These require the training and tuning data released for the Shared Translation Task for WMT 2010.
Create a working directory, and change into it. Then execute the following steps:

\begin{SaveVerbatim}{myverb}
 mkdir data
 cd data
 wget://www.statmt.org/wmt10/training-parallel.tgz
 tar xzf training-parallel.tgz
 wget://www.statmt.org/wmt10/dev.tgz
 tar xzf dev.tgz
 cd ..
\end{SaveVerbatim}
\colorbox{gray}{%
\BUseVerbatim{myverb}
}

The examples using these corpora are

\begin{itemize}
\item {\tt config.basic} - a basic phrase based model

\item {\tt config.factored} - a factored phrase based model

\item {\tt config.hierarchical} - a hierarchical phrase based model

\item {\tt config.syntax} - a target syntax model

\end{itemize}

In all these example configuration files, most corpora are commented out. This is done by adding the word {\tt IGNORE} at the end of a corpus definition (also for the language models). This allows you to run a basic experiment with just the News Commentary corpus which finished relatively quickly. 
Remove the {\tt IGNORE} to include more training data. You may run into memory and disk space problems when using some of the larger corpora (especially the news language model), depending on your computing infrastructure.

If you decide to use multiple corpora for the language model, you may also want to try out interpolating the individual language models (instead of using them as separate feature functions). For this, you need to comment out the {\tt IGNORE} next to the {\tt [INTERPOLATED-LM]} section.

You may also specify different language pairs by changing the {\tt input-extension}, {\tt output-extension}, and {\tt pair-extension} settings.

Finally, you can run all the experiments with the different given configuration files and the data variations in the same working directory. The experimental management system figures out automatically which processing steps do not need to repeated because they can be re-used from prior experimental runs.

\vspace{2mm}

{\bf 
\label{experiment-perl.texntoc5}Phrase Model}

Phrase models are, compared to the following examples, the simplest models to be trained with Moses and the fastest models to run. You may prefer these models over the more sophisticated models whose added complexity may not justify the small (if any) gains.

The example {\tt config.basic} is similar to the toy example, except for a larger training and test corpora. Also, the tuning stage is not skipped. Thus, even with most of the corpora commented out, the entire experimental run will likely take a day, with most time taken up by word alignment ({\tt TRAINING\_run-giza} and {\tt TRAINING\_run-giza-inverse}) and tuning ({\tt TUNING\_tune}).

\vspace{2mm}

{\bf 
\label{experiment-perl.texntoc6}Factored Phrase Model}

Factored models allow for additional annotation at the word level which may be exploited in various models. The example in {\tt config.factored} uses part-of-speech tags on the English target side.

Annotation with part-of-speech tags is done with MXPOST, which needs to be installed first. Please read the installation instructions\footnote{\sf http://www.statmt.org/moses/?n=Moses.ExternalTools\#mxpost}. After this, you can run experiment.perl with the configuration file {\tt config.factored}.

If compare the factored example {\tt config.factored} with the phrase-based example {\tt config.basic} you will notice the definition of the factors used:

\begin{SaveVerbatim}{myverb}
 ### factored training: specify here which factors used
 # if none specified, single factor training is assumed
 # (one translation step, surface to surface)
 #
 input-factors = word
 output-factors = word pos
 alignment-factors = "word -> word"
 translation-factors = "word -> word+pos"
 reordering-factors = "word -> word"
 #generation-factors = 
 decoding-steps = "t0"
\end{SaveVerbatim}
\colorbox{gray}{%
\BUseVerbatim{myverb}
}

the factor definition:

\begin{SaveVerbatim}{myverb}
 #################################################################
 # FACTOR DEFINITION

 [INPUT-FACTOR]

 # also used for output factors
 temp-dir = $working-dir/training/factor

 [OUTPUT-FACTOR:pos]

 ### script that generates this factor
 #
 mxpost = /home/pkoehn/bin/mxpost
 factor-script = "$moses-script-dir/training/wrappers/make-factor-en-pos.mxpost.perl -mxpost $mxpost"
\end{SaveVerbatim}
\colorbox{gray}{%
\BUseVerbatim{myverb}
}

and the specification of a 7-gram language model over part of speech tags:

\begin{SaveVerbatim}{myverb}
 [LM:nc=pos]
 factors = "pos"
 order = 7
 settings = "-interpolate -unk"
 raw-corpus = $wmt10-data/training/news-commentary10.$pair-extension.$output-extension
\end{SaveVerbatim}
\colorbox{gray}{%
\BUseVerbatim{myverb}
}

This factored model using all the available corpora is identical to the Edinburgh submission to the WMT 2010 shared task for English-Spanish, Spanish-English, and English-German language pairs (the French language pairs also used the $\text{10}^\text{9}$ corpus, the Czech language pairs did not use the POS language model, and German-English used additional pre-processing steps).

\vspace{2mm}

{\bf 
\label{experiment-perl.texntoc7}Hierarchical model}

Hierarchical phrase models allow for rules with gaps. Since these are represented by non-terminals and such rules are best processed with a search algorithm that is similar to syntactic chart parsing, such models fall into the class of tree-based or grammar-based models. For more information, please check the <a class='createlinktext' rel='nofollow' 
    href='http://www.statmt.org/moses/?n=Moses.SyntaxTutoral?action=edit'>Syntax Tutorial</a><a rel='nofollow' 
    class='createlink' href='http://www.statmt.org/moses/?n=Moses.SyntaxTutoral?action=edit'>?</a>.

From the view of setting up hierarchical models with experiment.perl, very little has to be changed in comparison to the configuration file for phrase-based models:

\begin{SaveVerbatim}{myverb}
 % diff config.basic config.hierarchical 
 33c33
 < decoder = $moses-src-dir/moses-cmd/src/moses
 ---
 > decoder = $moses-src-dir/moses-chart-cmd/src/moses_chart
 36c36
 < ttable-binarizer = $moses-src-dir/misc/processPhraseTable
 ---
 > #ttable-binarizer = $moses-src-dir/misc/processPhraseTable
 39c39
 < #ttable-binarizer = "$moses-src-dir/CreateOnDisk/src/CreateOnDiskPt 1 1 5 100 2"
 ---
 > ttable-binarizer = "$moses-src-dir/CreateOnDisk/src/CreateOnDiskPt 1 1 5 100 2"
 280c280
 < lexicalized-reordering = msd-bidirectional-fe
 ---
 > #lexicalized-reordering = msd-bidirectional-fe
 284c284
 < #hierarchical-rule-set = true
 ---
 > hierarchical-rule-set = true
 413c413
 < decoder-settings = "-search-algorithm 1 -cube-pruning-pop-limit 5000 -s 5000"
 ---
 > #decoder-settings = ""
\end{SaveVerbatim}
\colorbox{gray}{%
\BUseVerbatim{myverb}
}

The changes are: a different decoder binary (by default compiled into {\tt moses-chart-cmd/src/moses\_chart}) and {\tt ttable-binarizer} are used. The decoder settings for phrasal cube pruning do not apply. 
Also, hierarchical models do not allow for lexicalized reordering (their rules fulfill the same purpose), and the setting for hierarchical rule sets has to be turned on. The use of hierarchical rules is indicated with the setting {\tt hierarchical-rule-set}.

\vspace{2mm}

{\bf 
\label{experiment-perl.texntoc8}Target syntax model}

Syntax models imply the use of linguistic annotation for the non-terminals of hierarchical models. This requires running a syntactic parser.

In our example {\tt config.syntax}, syntax is used only on the English target side. The syntactic constituents are labeled with Collins parser, which needs to be installed first. Please read the installation instructions\footnote{\sf http://www.statmt.org/moses/?n=Moses.ExternalTools\#collins}.

Compared to the hierarchical model, very little has to be changed in the configuration file:

\begin{SaveVerbatim}{myverb}
 % diff config.hierarchical config.syntax
 46a47,49
 > # syntactic parsers
 > collins = /home/pkoehn/bin/COLLINS-PARSER
 > output-parser = "$moses-script-dir/training/wrappers/parse-en-collins.perl"
 > 
 241c244
 < #extract-settings = ""
 ---
 > extract-settings = "--MinHoleSource 1 --NonTermConsecSource"
\end{SaveVerbatim}
\colorbox{gray}{%
\BUseVerbatim{myverb}
}

The parser needs to be specified, and the extraction settings may be adjusted. And you are ready to go.

\subsection{
\label{experiment-perl.texntoc9}Try a Few More Things}

\vspace{2mm}

{\bf 
\label{experiment-perl.texntoc10}Stemmed Word Alignment}

The factored translation model training makes it very easy to set up word alignment not based on the surface form of words, but any other property of a word. One relatively popular method is to use stemmed words for word alignment.

There are two reasons for this: For one, for morphologically rich languages, stemming overcome data sparsity problems. Secondly, GIZA++ may have difficulties with very large vocabulary sizes, and stemming reduces the number of unique words.

To set up stemmed word alignment in experiment.perl, you need to define a stem as a factor:

\begin{SaveVerbatim}{myverb}
 [OUTPUT-FACTOR:stem4]
 factor-script = "$moses-script-dir/training/wrappers/make-factor-stem.perl 4"

 [INPUT-FACTOR:stem4]
 factor-script = "$moses-script-dir/training/wrappers/make-factor-stem.perl 4"
\end{SaveVerbatim}
\colorbox{gray}{%
\BUseVerbatim{myverb}
}

and indicate the use of this factor in the {\tt TRAINING} section:

\begin{SaveVerbatim}{myverb}
 input-factors = word stem4
 output-factors = word stem4
 alignment-factors = "stem4 -> stem4"
 translation-factors = "word -> word"
 reordering-factors = "word -> word"
 #generation-factors = 
 decoding-steps = "t0"
\end{SaveVerbatim}
\colorbox{gray}{%
\BUseVerbatim{myverb}
}

\vspace{2mm}

{\bf 
\label{experiment-perl.texntoc11}Using Multi-Threaded GIZA++}

GIZA++ is one of the slowest steps in the training pipeline. Qin Gao implemented a multi-threaded version of GIZA++, called MGIZA, which speeds up word alignment on multi-core machines.

To use MGIZA, you will first need to install\footnote{\sf http://www.statmt.org/moses/?n=Moses.ExternalTools\#mgiza} it.

To use it, you simply need to add some training options in the section {\tt TRAINING}:

\begin{SaveVerbatim}{myverb}
 ### general options
 #
 training-options = "-mgiza -mgiza-cpus 8"
\end{SaveVerbatim}
\colorbox{gray}{%
\BUseVerbatim{myverb}
}

\vspace{2mm}

{\bf 
\label{experiment-perl.texntoc12}Using Berkeley Aligner}

The Berkeley aligner is a alternative to GIZA++ for word alignment. You may (or may not) get better results using this tool.

To use the Berkeley aligner, you will first need to install\footnote{\sf http://www.statmt.org/moses/?n=Moses.ExternalTools\#berkeley} it.

The example configuration file already has a section for the parameters for the tool. You need to un-comment them and adjust {\tt berkeley-jar} to your installation. You should comment out {\tt alignment-symmetrization-method}, since this is a GIZA++ setting.

\begin{SaveVerbatim}{myverb}
 ### symmetrization method to obtain word alignments from giza output
 # (commonly used: grow-diag-final-and)
 #
 #alignment-symmetrization-method = grow-diag-final-and

 ### use of berkeley aligner for word alignment
 #
 use-berkeley = true
 alignment-symmetrization-method = berkeley
 berkeley-train = $moses-script-dir/ems/support/berkeley-train.sh
 berkeley-process =  $moses-script-dir/ems/support/berkeley-process.sh
 berkeley-jar = /your/path/to/berkeleyaligner-2.1/berkeleyaligner.jar
 berkeley-java-options = "-server -mx30000m -ea"
 berkeley-training-options = "-Main.iters 5 5 -EMWordAligner.numThreads 8"
 berkeley-process-options = "-EMWordAligner.numThreads 8"
 berkeley-posterior = 0.5
\end{SaveVerbatim}
\colorbox{gray}{%
\BUseVerbatim{myverb}
}

The Berkeley aligner proceeds in two step: a training step to learn the alignment model from the data and a processing step to find the best alignment for the training data. This step has the parameter {\tt berkeley-posterior} to adjust a bias towards more or less alignment points. You can try different runs with different values for this parameter. Experiment.perl will not re-run the training step, just the processing step.

\vspace{2mm}

{\bf 
\label{experiment-perl.texntoc13}IRST Language Model}

The provided examples use the SRILM language model during decoding. When you want to use the IRSTLM instead, an additional processing step is required: the language model has to converted into a binary format.

This part of the {\tt LM} section defines the use of IRSTLM:

\begin{SaveVerbatim}{myverb}
 ### script to use for binary table format for irstlm
 # (default: no binarization)
 #
 #lm-binarizer = $moses-src-dir/irstlm/bin/compile-lm

 ### script to create quantized language model format (irstlm)
 # (default: no quantization)
 # 
 #lm-quantizer = $moses-src-dir/irstlm/bin/quantize-lm
\end{SaveVerbatim}
\colorbox{gray}{%
\BUseVerbatim{myverb}
}

If you un-comment {\tt lm-binarizer}, IRSTLM will be used. If you comment out in addition {\tt lm-quantizer}, the language model will be compressed into a more compact representation. Note that the values above assume that you installed the IRSTLM language model toolkit in the directory {\tt \$moses-src-dir/irstlm}.

\vspace{2mm}

{\bf 
\label{experiment-perl.texntoc14}Randomized Language Model}

Randomized language models allow a much more compact (but lossy) representation. Being able to use much larger corpora for the language model may be beneficial over the small chance of making mistakes.

First of all, you need to install\footnote{\sf http://www.statmt.org/moses/?n=FactoredTraining.BuildingLanguageModel\#randlm} the randomized language model toolkit.

There are two different ways to train a randomized language model. One is to train it from scratch. The other way is to convert a SRILM language model into randomized representation.

{\bf Training from scratch:} Find the following section in the example configuration files and un-comment the {\tt rlm-training} setting. Note that the section below assumes that you installed the randomized language model toolkit in the directory {\tt \$moses-src-dir/randlm}.

\begin{SaveVerbatim}{myverb}
 ### tool to be used for training randomized language model from scratch
 # (more commonly, a SRILM is trained)
 #
 rlm-training = "$moses-src-dir/randlm/bin/buildlm -falsepos 8 -values 8"
\end{SaveVerbatim}
\colorbox{gray}{%
\BUseVerbatim{myverb}
}

{\bf Converting SRILM language model:} Find the following section in the example configuration files and un-comment the {\tt lm-randomizer} setting.

\begin{SaveVerbatim}{myverb}
 ### script to use for converting into randomized table format
 # (default: no randomization)
 #
 lm-randomizer = "$moses-src-dir/randlm/bin/buildlm -falsepos 8 -values 8"
\end{SaveVerbatim}
\colorbox{gray}{%
\BUseVerbatim{myverb}
}

You may want to try other values for {\tt falsepos} and {\tt values}. Please see the language model section on the randomized language model\footnote{\sf http://www.statmt.org/moses/?n=FactoredTraining.BuildingLanguageModel\#randlm} for some more information about these parameters.

You can also randomize a interpolated language model by specifying the {\tt lm-randomizer} in the section {\tt INTERPOLATED-LM}.

\vspace{2mm}

{\bf 
\label{experiment-perl.texntoc15}Compound Splitting}

Compounding languages, such as German, allow the creation of long words such as {\em Neuwortgenerierung} ({\em new word generation}). This results in a lot of unknown words in any text, so splitting up these compounds is a common method when translating from such languages.

Moses offers an support tool that splits up words, if the geometric average of the frequency of its parts is higher than the frequency of a word. The method requires a model (the frequency statistics of words in a corpus), so there is a training and application step.

Such word splitting can be added to experiment.perl simply by specifying the splitter script in the {\tt GENERAL} section:

\begin{SaveVerbatim}{myverb}
 input-splitter = $moses-script-dir/generic/compound-splitter.perl
\end{SaveVerbatim}
\colorbox{gray}{%
\BUseVerbatim{myverb}
}

Splitting words on the output side is currently not supported.

\subsection{
\label{experiment-perl.texntoc16}A Short Manual}

The basic lay of the land is: experiment.perl is breaks up the training, tuning, and evaluating of a statistical machine translation system into a number of steps, which are then scheduled to run in parallel or sequence depending on their inter-dependencies and available resources. The possible steps are defined in the file {\tt experiment.meta}. An experiment is defined by a configuration file.

The main modules of running an experiment are:

\begin{itemize}
\item {\tt CORPUS}: preparing a parallel corpus

\item {\tt INPUT-FACTOR} and {\tt OUTPUT-FACTOR}: commands to create factors

\item {\tt TRAINING}: training a translation model

\item {\tt LM}: training a language model

\item {\tt INTERPOLATED-LM}: interpolate language models

\item {\tt SPLITTER}: training a word splitting model

\item {\tt RECASING}: training a recaser

\item {\tt TRUECASING}: training a truecaser

\item {\tt TUNING}: running minumum error rate training to set component weights

\item {\tt TESTING}: translating and scoring a test set

\item {\tt REPORTING}: compile all scores in one file

\end{itemize}

\vspace{2mm}

{\bf 
\label{experiment-perl.texntoc17}Experiment.Meta}

The actual steps, their dependencies and other salient information is to be found in the file {\tt experiment.meta}. Think of experiment.meta as a "template" file.

Here the parts of the step description for {\tt CORPUS:get-corpus} and {\tt CORPUS:tokenize}:

\begin{SaveVerbatim}{myverb}
 get-corpus
        in: get-corpus-script
        out: raw-stem
        [...]

 tokenize
        in: raw-stem
        out: tokenized-stem
        [...]
\end{SaveVerbatim}
\colorbox{gray}{%
\BUseVerbatim{myverb}
}

Each step takes some input ({\tt in}) and provides some output ({\tt out}). This also establishes the dependencies between the steps. The step {\tt tokenize} requires the input {\tt raw-stem}. This is provided by the step {\tt get-corpus}.

{\tt experiment.meta} provides a generic template for steps and their interaction. For an actual experiment, a configuration file determines which steps need to be run. This configuration file is the one that is specified when invocing {\tt experiment.perl}. It may contain for instance the following:

\begin{SaveVerbatim}{myverb}
 [CORPUS:europarl]

 ### raw corpus files (untokenized, but sentence aligned)
 #
 raw-stem = $europarl-v3/training/europarl-v3.fr-en
\end{SaveVerbatim}
\colorbox{gray}{%
\BUseVerbatim{myverb}
}

Here, the parallel corpus to be used is named {\tt europarl} and it is provided in raw text format in the location {\tt \$europarl-v3/training/europarl-v3.fr-en} (the variable {\tt \$europarl-v3} is defined elsewhere in the config file). The effect of this specification in the config file is that the step {\tt get-corpus} does not need to be run, since its output is given as a file. More on the configuration file below in the next section.

Several types of information are specified in {\tt experiment.meta}:

\begin{itemize}
\item {\tt in} and {\tt out}: Established dependencies between steps; input may also be provided by files specified in the configuration.

\item {\tt default-name}: Name of the file in which the output of the step will be stored.

\item {\tt template}: Template for the command that is placed in the execution script for the step.

\item {\tt template-if}: Potential command for the execution script. Only used, if the first parameter exists.

\item {\tt error}: {\tt experiment.perl} detects if a step failed by scanning STDERR for key words such as killed, error, died, not found, and so on. Additional key words and phrase are provided with this parameter.

\item {\tt not-error}: Declares default error key words as not indicating failures.

\item {\tt pass-unless}: Only if the given parameter is defined, this step is executed, otherwise the step is passed (illustrated by a yellow box in the graph).

\item {\tt ignore-unless}: If the given parameter is defined, this step is not executed. This overrides requirements of downstream steps.

\item {\tt rerun-on-change}: If similar experiment are runs, the output of steps may be used, if input and parameter settings are the same. This specifies a number of parameters whose change disallows a re-use in different run.

\item {\tt parallelizable}: When running on the cluster, this step may be parallelized (only if {\tt generic-parallelizer} is set in the config file, typically to {\tt \$edinburgh-script-dir/generic-parallelizer.perl}.

\item {\tt qsub-script}: If running on a cluster, this step is run on the head node, and not submitted to the queue (because it submits jobs itself).

\end{itemize}

Here now the full definition of the step {\tt CONFIG:tokenize}

\begin{SaveVerbatim}{myverb}
 tokenize
        in: raw-stem
        out: tokenized-stem
        default-name: corpus/tok
        pass-unless: input-tokenizer output-tokenizer
        template-if: input-tokenizer IN.$input-extension OUT.$input-extension
        template-if: output-tokenizer IN.$output-extension OUT.$output-extension
        parallelizable: yes
\end{SaveVerbatim}
\colorbox{gray}{%
\BUseVerbatim{myverb}
}

The step takes {\tt raw-stem} and produces {\tt tokenized-stem}. It is parallizable with the generic parallelizer.

That output is stored in the file {\tt corpus/tok}. Note that the actual file name also contains the corpus name, and the run number. Also, in this case, the parallel corpus is stored in two files, so file name may be something like {\tt corpus/europarl.tok.1.fr} and {\tt corpus/europarl.tok.1.en}.

The step is only executed, if either {\tt input-tokenizer} or {\tt output-tokenizer} are specified. The templates indicate how the command lines in the execution script for the steps look like.

\label{experiment-perl.texmultiple-sections}

\vspace{2mm}

{\bf 
\label{experiment-perl.texntoc18}Multiple Corpora, One Translation Model}

We may use multiple parallel corpora for training a translation model or multiple monolingual corpora for training a language model. Each of these have their own instances of the {\tt CORPUS} and {\tt LM} module. There may be also multiple test sets in {\tt TESTING}). However, there is only one translation model and hence only one instance of the {\tt TRAINING} module.

The definitions in {\tt experiment.meta} reflects the different nature of these modules. For instance {\tt CORPUS} is flagged as {\tt multiple}, while {\tt TRAINING} is flagged as {\tt single}.

When defining settings for the different modules, the singular module {\tt TRAINING} has only one section, while this one general section and specific {\tt LM} sections for each training corpus. In the specific section, the corpus is named, e.g. {\tt LM:europarl}.

As you may imagine, the tracking of dependencies between steps of different types of modules and the consolidation of corpus-specific instances of modules is a bit complex. But most of that is hidden from the user of the experimental management system.

When looking up the parameter settings for a step, first the set-specific section ({\tt LM:europarl}) is consulted. If there is no definition, then the module definition ({\tt LM}) and finally the general definition (in section {\tt GENERAL}) is consulted. In other words, local settings override global settings.

\label{experiment-perl.texdef-settings}

\vspace{2mm}

{\bf 
\label{experiment-perl.texntoc19}Defining Settings}

The configuration file for experimental runs is a collection of parameter settings, one per line with empty lines and comment lines for better readability, organized in sections for each of the modules.

The syntax of setting definition is {\tt setting = value} (note: spaces around the equal sign). If the value contains spaces, it must be placed into quotes ({\tt setting = "the value"}), except when a vector of values is implied (only used when defining list of factors: {\tt output-factor = word pos}.

Comments are indicated by a hash ({\tt \#}).

The start of sections is indicated by the section name in square brackets ({\tt [TRAINING]} or {\tt [CORPUS:europarl]}). If the word {\tt IGNORE} is appended to a section definition, then the entire section is ignored.

Settings can be used as variables to define other settings:

\begin{SaveVerbatim}{myverb}
 working-dir = /home/pkoehn/experiment
 wmt10-data = $working-dir/data
\end{SaveVerbatim}
\colorbox{gray}{%
\BUseVerbatim{myverb}
}

Variable names may be placed in curly brackets for clearer separation:

\begin{SaveVerbatim}{myverb}
 wmt10-data = ${working-dir}/data
\end{SaveVerbatim}
\colorbox{gray}{%
\BUseVerbatim{myverb}
}

Such variable references may also reach other modules:

\begin{SaveVerbatim}{myverb}
 [RECASING]
 tokenized = $LM:europarl:tokenized-corpus
\end{SaveVerbatim}
\colorbox{gray}{%
\BUseVerbatim{myverb}
}

Finally, reference can be made to settings that are not defined in the configuration file, but are the product of the defined sequence of steps.

Say, in the above example, {\tt tokenized-corpus} is not defined in the section {\tt LM:europarl}, but instead {\tt raw-corpus}. Then, the tokenized corpus is produced by the normal processing pipeline. Such an intermediate file can be used elsewhere:

\begin{SaveVerbatim}{myverb}
 [RECASING]
 tokenized = [LM:europarl:tokenized-corpus]
\end{SaveVerbatim}
\colorbox{gray}{%
\BUseVerbatim{myverb}
}

Some error checking is done on the validity of the values. All values that seem to be file paths trigger the existence check for such files. A file with the prefix of the value must exist.

There are a lot of settings reflecting the many steps, and explaining these would require explaining the entire training, tuning, and testing pipeline. Please find the required documentation for step elsewhere around here. Every effort has been made to include verbose descriptions in the example configuration files, which should be taken as starting point.

\label{experiment-perl.texworking-with-ems}

\vspace{2mm}

{\bf 
\label{experiment-perl.texntoc20}Working with Experiment.Perl}

You have to define an experiment in a configuration file and the experiment management system figures out which steps need to be run and schedules them either as jobs on a cluster or runs them serially on a single machine.

Other options:

\begin{itemize}
\item {\tt -no-graph} supresses the display of the graph

\item {\tt -continue RUN} continues the experiment {\tt RUN}, which crashed earlier. Make sure that crashed step and its output is deleted (see more below).

\item {\tt -max-active} specifies the number of steps that can be run in parallel when running on a single machine (default: 2, not used when run on cluster).

\item {\tt -sleep} sets the number of seconds to be waited in the scheduler before the completion of tasks is checked (default: 2).

\item {\tt -ignore-time} changes the re-use behavior. By default files cannot be re-used when their time stamp changed (typically a tool such as the tokenizer which was changed, thus requiring re-running all tokenization steps in new experiments). With this switch, files with changed time stamp can be re-used.

\item {\tt -meta} allows the specification of a custom {\tt experiment.meta} file, instead of using the one in the same directory as the {\tt experiment.perl} script.

\item {\tt -cluster} indicates that the current machine is a cluster head node. Step files are submitted as jobs to the cluster.

\item {\tt -multicore} indicates that the current machine is a multi-core machine. This allows for additional parallelization with the generic parallelizer setting.

\end{itemize}

The script may automatically detect if it is run on a compute cluster or a multi-core machine, if this is specified in the file {\tt experiment.machines}, for instance:

\begin{SaveVerbatim}{myverb}
 cluster: townhill seville
 multicore-8: tyr thor
 multicore-16: loki
\end{SaveVerbatim}
\colorbox{gray}{%
\BUseVerbatim{myverb}
}

defines the machines {\tt townhill} and {\tt seville} as GridEngine cluster machines, {\tt tyr} and {\tt thor} as 8-core machines and {\tt loki} as 16-core machines.

Typically, experiments are started with the command:

\begin{SaveVerbatim}{myverb}
 experiment.perl -config my-config -exec
\end{SaveVerbatim}
\colorbox{gray}{%
\BUseVerbatim{myverb}
}

Since experiments run for a long time, you may want to run this in the background and also set a nicer priority:

\begin{SaveVerbatim}{myverb}
 nice nohup -config my-config -exec >\& OUT.[RUN] \&
\end{SaveVerbatim}
\colorbox{gray}{%
\BUseVerbatim{myverb}
}

This keeps also a report (STDERR and STDOUT) on the execution in a file named, say, {\tt OUT.1}, with the number corresponding to the run number.

The meta-information for the run is stored in the directory {\tt steps}. Each run has a sub directory with its number ({\tt steps/1}, {\tt steps/2}, etc.). The sub directory {\tt steps/0} contains step specification when Experiment.perl is called without the {\tt -exec} switch.

The sub directories for each run contain the step definitions, as well as their meta information and output. The sub directories also contain a copy of the configuration file (e.g. {\tt steps/1/config.1}), the agenda graph (e.g. {\tt steps/1/graph.1.$\{$dot,ps,png$\}$}), a file containing all expanded parameter settings (e.g. {\tt steps/1/parameter.1}), and an empty file that is touched every minute as long as the experiment is still running (e.g. {\tt steps/1/running.1}).

\vspace{2mm}

{\bf 
\label{experiment-perl.texntoc21}Continuing Crashed Experiments}

Steps may crash. No, steps will crash, be it because faulty settings, faulty tools, problems with the computing resources, willful interruption of an experiment, or an act of God.

The first thing to continue a crashed experiment is to detect the crashed step. This is shown either by the red node in the displayed graph or reported on the command line in the last lines before crashing; though this may not be pretty obvious, if parallel steps kept running after that. However, the automatic error detection is not perfect and a step may have failed upstream without detection causes failure further down the road.

You should have a understanding of what each step does. Then, by looking at its {\tt STDERR} and {\tt STDOUT} file, and the output files it should have produced, you can track down what went wrong.

Fix the problem, and delete all files associated with the failed step (e.g., {\tt rm steps/13/TUNING\_tune.13*}, {\tt rm -r tuning/tmp.1}). To find what has been produced by the crashed step, you may need to consult where the output of this step is placed, by looking at  {\tt experiment.meta}.

You can continue a crashed experimental run (e.g., run number 13) with:

\begin{SaveVerbatim}{myverb}
 experiment.perl -continue 13 -exec
\end{SaveVerbatim}
\colorbox{gray}{%
\BUseVerbatim{myverb}
}

You may want to check what will be run by excluding the {\tt -exec} command at first. The graph indicates which steps will be re-used from the original crashed run.

If the mistake was a parameter setting, you can change that setting in the stored configuration file (e.g., {\tt steps/1/config.1}). Take care, however, to delete all steps (and their subsequent steps) that would have been run differently with that setting.

If an experimental run crashed early, or you do not want to repeat it, it may be easier to delete the entire step directory ({\tt rm -r steps/13}). Only do this with the latest experimental run (e.g., not when there is already a run 14), otherwise it may mess up the re-use of results.

You may also delete all output associated with a run with the command {\tt rm -r */*.13*}. However this requires some care, so you may want to check first what you are deleting ({\tt ls */*.13}).

\vspace{2mm}

{\bf 
\label{experiment-perl.texntoc22}Running on a Cluster}

Experiment.perl works with Sun GridEngine clusters. The script needs to be run on the head node and jobs are scheduled on the nodes.

There are two ways to tell experiment.perl that the current machine is a cluster computer. One is by using the switch {\tt -cluster}, or by adding the machine name into {\tt experiment.machines}.

The configuration file has a section that allows for the setting of cluster-specific settings. The setting {\tt jobs} is used to specify into how many jobs to split the decoding during tuning and testing. For more details on this, please see {\tt moses-parallel.pl}.

All other settings specify switches that are passed along with each submission of a job via qsub:

\begin{itemize}
\item {\tt qsub-memory}: number of memory slots ({\tt -pe memory NUMBER})

\item {\tt qsub-hours}: number of hours reserved for each job ({\tt -l h\_rt=NUMBER:0:0})

\item {\tt qsub-project}: name if the project for user accounting ({\tt -P PROJECT})

\item {\tt qsub-settings}: any other setting that is passed along verbatim

\end{itemize}

Note that the general settings can be overriden in each module definition - you may want to have different settings for different steps.

If the setting {\tt generic-parallelizer} is set (typically to the script {\tt \$moses-script-dir/ems/support/generic-parallelizer.perl}), then a number of additional steps are parallelized. For instance, tokenization is performed by breaking up the corpus into as many parts as specified with {\tt jobs}, jobs to process the parts are submitted in parallel to the cluster, and their output pieced together upon completion.

Be aware that there many different ways to configure a GridEngine cluster. Not all the options described here may be available, and it my not work out of the box, due to yout specific installation.

\vspace{2mm}

{\bf 
\label{experiment-perl.texntoc23}Running on a Multicore Machine}

Using a multicore machine means first of all that more steps can be scheduled in parallel. There is also a generic parallelizer ({\tt generic-multicore-parallelizer.perl}) that plays the same role as the generic parallelizer for clusters.

However, decoding is not broken up into several pieces. It is more sensible to use multi-threading in the decoder\footnote{\sf http://www.statmt.org/moses/?n=Moses.AdvancedFeatures\#multi-threaded}.

\vspace{2mm}

{\bf 
\label{experiment-perl.texntoc24}Web Interface}

The introduction included some screen shots of the web interface to the experimental management tool. You will need to have a running web server on a machine (LAMPP on Linux or MAMP on Mac does the trick) that has access to the file system where your working directory is stored.

Copy or link the {\tt web} directory (in {\tt scripts/ems}) on a web server. Make sure the web server user has the right write permissions on the web interface directory.

To add your experiments to this interface, add a line to the file {\tt /your/web/interface/dir/setup}. The format of the file is explained in the file.

\vspace{2mm}

{\bf 
\label{experiment-perl.texntoc25}Analysis}

You can include additional analysis for an experimental run into the web interface by specifying the setting {\tt analysis} in its configuration file.

\begin{SaveVerbatim}{myverb}
 analysis = $moses-script-dir/ems/support/analysis.perl
\end{SaveVerbatim}
\colorbox{gray}{%
\BUseVerbatim{myverb}
}

This currently reports n-gram precision and recall statistics and color-coded n-gram correctness markup for the output sentences.

The output is color-highlighted according to n-gram matches with the reference translation. The following colors are used:

\begin{itemize}
\item grey: word not in reference

\item light blue: word part of 1-gram match

\item blue: word part of 2-gram match

\item dark blue: word part of 3-gram match

\item very dark blue: word part of 4-gram match

\end{itemize}

<img src='http://www.statmt.org/moses/img/ems-annotation.png' alt='' title='' />

Additional reports are available when adding the following settings:

\begin{SaveVerbatim}{myverb}
 analyze-coverage = yes
 report-segmentation = yes
\end{SaveVerbatim}
\colorbox{gray}{%
\BUseVerbatim{myverb}
}

The setting {\tt analyze-coverage} include a coverage analysis: which words and phrases in the input occur in the training data or the translation table? This is reported in color coding and in a yellow report box when moving the mouse of the word or the phrase. Also, summary statistics for how many words occur how often are given, and a report on unknown or rare words is generated.

The setting {\tt report-segmentation} creates summary statistics about what kind of phrase mappings are used (one word to one word, 1-2, 2-2, etc.), as well as markup of the sentence pair with the phrase segmentation. The phrase segmentation is indicated with black boxes around the words, and the alignment is shown when moving the mouse on the phrases.

### 2016-01-22 acl.2016/acl2016.tex
start acl 2016 paper

%
% File ACL2016.tex
%

\documentclass[11pt]{article}
\usepackage{acl2016}
\usepackage{times}
\usepackage{latexsym}

%\aclfinalcopy % Uncomment this line for the final submission
%\def\aclpaperid{***} %  Enter the acl Paper ID here

% To expand the titlebox for more authors, uncomment
% below and set accordingly.
% \addtolength\titlebox{.5in}

\newcommand\BibTeX{B{\sc ib}\TeX}

\title{Instructions for ACL 2016 Proceedings\Thanks{This
    document has been adapted from the instructions for earlier ACL
    and NAACL proceedings, including those for
    NAACL HLT16 by Margaret Mitchell, 
    NAACL HLT15 by Matt Post and Adam Lopez,
    NAACL HLT12 by Nizar Habash and William Schuler,
    NAACL HLT10 by Claudia Leacock and Richard Wicentowski,
    NAACL HLT09 by Joakim Nivre and Noah Smith, 
    for ACL05 by Hwee Tou Ng and Kemal Oflazer,
    for ACL02 by Eugene Charniak and Dekang Lin, and earlier ACL and
    EACL formats.  Those versions were written by several people,
    including John Chen, Henry S. Thompson and Donald Walker.
    Additional elements were taken from the formatting instructions of
    the {\em International Joint Conference on Artificial Intelligence}
    and the {\em Conference on Computer Vision and Pattern Recognition}.}}

% Author information can be set in various styles:
% For several authors from the same institution:
% \author{Author 1 \and ... \and Author n \\
%         Address line \\ ... \\ Address line}
% if the names do not fit well on one line use
%         Author 1 \\ {\bf Author 2} \\ ... \\ {\bf Author n} \\
% For authors from different institutions:
% \author{Author 1 \\ Address line \\  ... \\ Address line
%         \And  ... \And
%         Author n \\ Address line \\ ... \\ Address line}
% To start a seperate ``row'' of authors use \AND, as in
% \author{Author 1 \\ Address line \\  ... \\ Address line
%         \AND
%         Author 2 \\ Address line \\ ... \\ Address line \And
%         Author 3 \\ Address line \\ ... \\ Address line}
% If the title and author information does not fit in the area allocated,
% place \setlength\titlebox{<new height>} right after
% at the top, where <new height> can be something larger than 2.25in
\author{Author 1\\
	    XYZ Company\\
	    111 Anywhere Street\\
	    Mytown, NY 10000, USA\\
	    {\tt author1@xyz.org}
	  \And
	Author 2\\
  	ABC University\\
  	900 Main Street\\
  	Ourcity, PQ, Canada A1A 1T2\\
  {\tt author2@abc.ca}}

\date{}

\begin{document}

\maketitle

\begin{abstract}
 This document contains instructions for preparing ACL 2016 submissions and camera-ready
  manuscripts.  The document itself conforms
  to its own specifications, and is therefore an example of what
  your manuscript should look like.  Papers are required to conform to
  all the directions reported in this document.  
By using the provided \LaTeX\ and \BibTeX\ styles ({\small\tt acl2016.sty}, {\small\tt acl2016.bst}), the required formatting will be enabled by default.

I hate latex - Hieu.

\end{abstract}

\section{Introduction}

The following instructions are directed to authors of papers submitted to and accepted
for publication in the ACL 2016 proceedings.  All authors are required
to adhere to these specifications. Authors are required to provide 
a Portable Document Format (PDF) version of
their papers.  The proceedings will be printed on A4 paper.
Authors from countries where access to word-processing systems is
limited should contact the publication chairs as soon as possible.
 Grayscale readability of all figures and
graphics will be encouraged for all accepted papers
(Section \ref{ssec:accessibility}).

Submitted and camera-ready formatting is similar,
  however, the submitted paper should have:
\begin{enumerate} 
\item Author-identifying information removed
\item A `ruler' on the left and right margins
\item Page numbers 
\item A confidentiality header.  
\end{enumerate}
In contrast, the camera-ready {\bf should  not have} a ruler, page numbers, nor a confidentiality header.  By uncommenting {\small\verb|\aclfinalcopy|} at the top of this 
 document, it will compile to produce an example of the camera-ready formatting; by leaving it commented out, the document will be anonymized for initial submission.  Authors should place this command after the {\small\verb|\usepackage|} declarations when preparing their camera-ready manuscript with the ACL 2016 style.

\section{General Instructions}

Manuscripts must be in two-column format.  Exceptions to the
two-column format include the title, as well as the 
authors' names and complete
addresses (only in the final version, not in the version submitted for review), 
which must be centered at the top of the first page (see
the guidelines in Subsection~\ref{ssec:first}), and any full-width
figures or tables.  Type single-spaced.  Do not number the pages 
in the camera-ready version.
Start all pages directly under the top margin.  See the guidelines
later regarding formatting the first page.

The maximum length of a manuscript is eight (8) pages for the main
conference, printed single-sided, plus two (2) pages for references
(see Section~\ref{sec:length} for additional information on the
maximum number of pages).

By uncommenting {\small\verb|\aclfinalcopy|} at the top of this 
 document, it will compile to produce an example of the camera-ready formatting; by leaving it commented out, the document will be anonymized for initial submission.  When you first create your submission on softconf, please fill in your submitted paper ID where {\small\verb|***|} appears in the {\small\verb|\def\aclpaperid{***}|} definition at the top.

The review process is double-blind, so do not include any author information (names, addresses) when submitting a paper for review.  
However, you should maintain space for names and addresses so that they will fit in the final (accepted) version.  The ACL 2016 \LaTeX\ style will create a titlebox space of 2.5in for you when {\small\verb|\aclfinalcopy|} is commented out.

\subsection{The Ruler}
The ACL 2016 style defines a printed ruler which should be presented in the
version submitted for review.  The ruler is provided in order that
reviewers may comment on particular lines in the paper without
circumlocution.  If you are preparing a document without the provided
style files, please arrange for an equivalent ruler to
appear on the final output pages.  The presence or absence of the ruler
should not change the appearance of any other content on the page.  The
camera ready copy should not contain a ruler. (\LaTeX\ users may uncomment
the {\small\verb|\aclfinalcopy|} command in the document preamble.)

Reviewers:
note that the ruler measurements do not align well with lines in the paper
--- this turns out to be very difficult to do well when the paper contains
many figures and equations, and, when done, looks ugly.  Just use fractional
references (e.g., the first line on this page is at mark $096.5$), although in most cases one would
expect that the approximate location will be adequate.

\subsection{Electronically-available resources}

ACL provides this description in \LaTeX2e{} ({\small\tt acl2016.tex}) and PDF
format ({\small\tt acl2016.pdf}), along with the \LaTeX2e{} style file used to
format it ({\small\tt acl2016.sty}) and an ACL bibliography style ({\small\tt acl2016.bst})
and example bibliography ({\small\tt acl2016.bib}).
These files are all available at
{\small\tt acl2016.org/index.php?article\_id=9}. We
strongly recommend the use of these style files, which have been
appropriately tailored for the ACL 2016 proceedings.

\subsection{Format of Electronic Manuscript}
\label{sect:pdf}

For the production of the electronic manuscript, you must use Adobe's
Portable Document Format (PDF). This format can be generated from
postscript files: on Unix systems, you can use {\small\tt ps2pdf} for this
purpose; under Microsoft Windows, you can use Adobe's Distiller, or
if you have cygwin installed, you can use {\small\tt dvipdf} or
{\small\tt ps2pdf}.  Note 
that some word processing programs generate PDF that may not include
all the necessary fonts (esp.\ tree diagrams, symbols). When you print
or create the PDF file, there is usually an option in your printer
setup to include none, all, or just non-standard fonts.  Please make
sure that you select the option of including ALL the fonts.  {\em
  Before sending it, test your {\/\em PDF} by printing it from a
  computer different from the one where it was created}. Moreover,
some word processors may generate very large postscript/PDF files,
where each page is rendered as an image. Such images may reproduce
poorly.  In this case, try alternative ways to obtain the postscript
and/or PDF.  One way on some systems is to install a driver for a
postscript printer, send your document to the printer specifying
``Output to a file'', then convert the file to PDF.

For reasons of uniformity, Adobe's {\bf Times Roman} font should be
used. In \LaTeX2e{} this is accomplished by putting
\small
\begin{verbatim}
\usepackage{times}
\usepackage{latexsym}
\end{verbatim}
\normalsize
in the preamble.

Print-outs of the PDF file on A4 paper should be identical to the
hardcopy version.  If you cannot meet the above requirements about the
production of your electronic submission, please contact the
publication chairs above  as soon as possible.

\subsection{The First Page}
\label{ssec:first}

Center the title, author name(s) and affiliation(s) across both
columns (or, in the case of initial submission, space for the names). 
Do not use footnotes for affiliations.  
Use the two-column format only when you begin the abstract.

{\bf Title}: Place the title centered at the top of the first page, in
a 15 point bold font.  (For a complete guide to font sizes and styles, see Table~\ref{font-table}.)
Long titles should be typed on two lines without
a blank line intervening. Approximately, put the title at 1in from the
top of the page, followed by a blank line, then the author name(s),
and the affiliation(s) on the following line.  Do not use only initials
for given names (middle initials are allowed). Do not format surnames
in all capitals (e.g., ``Mitchell,'' not ``MITCHELL'').  The affiliation should
contain the author's complete address, and if possible, an electronic
mail address. Leave about 0.75in between the affiliation and the body
of the first page.

{\bf Abstract}: Type the abstract at the beginning of the first
column.  The width of the abstract text should be smaller than the
width of the columns for the text in the body of the paper by about
0.25in on each side.  Center the word {\bf Abstract} in a 12 point
bold font above the body of the abstract. The abstract should be a
concise summary of the general thesis and conclusions of the paper.
It should be no longer than 200 words.  The abstract text should be in 10 point font.

\begin{table}
\centering
\small
\begin{tabular}{cc}
\begin{tabular}{|l|l|}
\hline
{\bf Command} & {\bf Output}\\\hline
\verb|{\"a}| & {\"a} \\
\verb|{\^e}| & {\^e} \\
\verb|{\`i}| & {\`i} \\ 
\verb|{\.I}| & {\.I} \\ 
\verb|{\o}| & {\o} \\
\verb|{\'u}| & {\'u}  \\ 
\verb|{\aa}| & {\aa}  \\\hline
\end{tabular} & 
\begin{tabular}{|l|l|}
\hline
{\bf Command} & {\bf  Output}\\\hline
\verb|{\c c}| & {\c c} \\ 
\verb|{\u g}| & {\u g} \\ 
\verb|{\l}| & {\l} \\ 
\verb|{\~n}| & {\~n} \\ 
\verb|{\H o}| & {\H o} \\ 
\verb|{\v r}| & {\v r} \\ 
\verb|{\ss}| & {\ss} \\\hline
\end{tabular}
\end{tabular}
\caption{Example commands for accented characters, to be used in, e.g., \BibTeX\ names.}\label{tab:accents}
\end{table}

{\bf Text}: Begin typing the main body of the text immediately after
the abstract, observing the two-column format as shown in 
the present document.  Do not include page numbers in the camera-ready manuscript.

{\bf Indent} when starting a new paragraph. For reasons of uniformity,
use Adobe's {\bf Times Roman} fonts, with 11 points for text and 
subsection headings, 12 points for section headings and 15 points for
the title.  If Times Roman is unavailable, use {\bf Computer Modern
  Roman} (\LaTeX2e{}'s default; see section \ref{sect:pdf} above).
Note that the latter is about 10\% less dense than Adobe's Times Roman
font.

\subsection{Sections}

{\bf Headings}: Type and label section and subsection headings in the
style shown on the present document.  Use numbered sections (Arabic
numerals) in order to facilitate cross references. Number subsections
with the section number and the subsection number separated by a dot,
in Arabic numerals.

{\bf Citations}: Citations within the text appear
in parentheses as~\cite{Gusfield:97} or, if the author's name appears in
the text itself, as Gusfield~\shortcite{Gusfield:97}.  Using the provided \LaTeX\ style, the former is accomplished using
{\small\verb|\cite|} and the latter with {\small\verb|\shortcite|} or {\small\verb|\newcite|}.  Collapse multiple citations as in~\cite{Gusfield:97,Aho:72}; this is accomplished with the provided style using commas within the {\small\verb|\cite|} command, e.g., {\small\verb|\cite{Gusfield:97,Aho:72}|}.  
Append lowercase letters to the year in cases of ambiguities.  
Treat double authors as in~\cite{Aho:72}, but write as 
in~\cite{Chandra:81} when more than two authors are involved.

\textbf{References}:  We recommend
including references in a separate~{\small\texttt .bib} file, and include an example file 
in this release ({\small\tt naalhlt2016.bib}).  Some commands
for names with accents are provided for convenience in Table \ref{tab:accents}.  
References stored in the separate~{\small\tt .bib} file are inserted into the document using the following commands:

\small
\begin{verbatim}
\bibliography{acl2016}
\bibliographystyle{acl2016}
\end{verbatim}
\normalsize

References should appear under the heading {\bf References} at the 
end of the document, but before any Appendices, unless the appendices contain references.  
Arrange the references alphabetically
by first author, rather than by order of occurrence in the text.  %This behavior is provided by default in the provided \BibTeX\ style ({\small\tt naaclhlt2016.bst}). 
Provide as complete a reference as possible, using a consistent format,
such as the one for {\em Computational Linguistics\/} or the one in the 
{\em Publication Manual of the American 
Psychological Association\/}~\cite{APA:83}.  Authors' full names rather than initials are preferred.  You may use
{\bf standard} abbreviations for conferences\footnote{\scriptsize {\tt https://en.wikipedia.org/wiki/ \\ \-\hspace{.75cm} List\_of\_computer\_science\_conference\_acronyms}} and journals\footnote{\tt http://www.abbreviations.com/jas.php}.

{\bf Appendices}: Appendices, if any, directly follow the text and the
references (but see above).  Letter them in sequence and provide an
informative title: {\bf Appendix A. Title of Appendix}.

\textbf{Acknowledgment} sections should go as a last (unnumbered) section immediately
before the references.

\subsection{Footnotes}

{\bf Footnotes}: Put footnotes at the bottom of the page. They may
be numbered or referred to by asterisks or other
symbols.\footnote{This is how a footnote should appear.} Footnotes
should be separated from the text by a line.\footnote{Note the
line separating the footnotes from the text.}  Footnotes should be in 9 point font.

\subsection{Graphics}

{\bf Illustrations}: Place figures, tables, and photographs in the
paper near where they are first discussed, rather than at the end, if
possible.  Wide illustrations may run across both columns and should be placed at
the top of a page. Color illustrations are discouraged, unless you have verified that 
they will be understandable when printed in black ink.

\begin{table}
\small
\centering
\begin{tabular}{|l|rl|}
\hline \bf Type of Text & \bf Font Size & \bf Style \\ \hline
paper title & 15 pt & bold \\
author names & 12 pt & bold \\
author affiliation & 12 pt & \\
the word ``Abstract'' & 12 pt & bold \\
section titles & 12 pt & bold \\
document text & 11 pt  &\\
abstract text & 10 pt & \\
captions & 9 pt & \\
caption label & 9 pt & bold \\
bibliography & 10 pt & \\
footnotes & 9 pt & \\
\hline
\end{tabular}
\caption{\label{font-table} Font guide.}
\end{table}

{\bf Captions}: Provide a caption for every illustration; number each one
sequentially in the form:  ``{\bf Figure 1:} Figure caption.'', ``{\bf Table 1:} Table caption.''  Type the captions of the figures and 
tables below the body, using 9 point text.  Table and Figure labels should be bold-faced.

\subsection{Accessibility}
\label{ssec:accessibility}

In an effort to accommodate the color-blind (as well as those printing
to paper), grayscale readability for all accepted papers will be
encouraged.  Color is not forbidden, but authors should ensure that
tables and figures do not rely solely on color to convey critical
distinctions.
Here we give a simple criterion on your colored figures, if your paper has to be printed in black and white, then you must assure that every curves or points in your figures can be still clearly distinguished.

\section{Length of Submission}
\label{sec:length}

The ACL 2016 main conference accepts submissions of long papers and short papers.  Long papers may consist of up to eight (8) pages of content, plus unlimited pages for references. Upon acceptance, final versions of long papers will be given one additional page (up to 9 pages with unlimited pages for references) so that reviewers' comments can be taken into account.  Short papers may consist of up to four (4) pages of content, plus unlimited pages for references. Upon acceptance, short papers will be given five (5) pages in the proceedings and unlimited pages for references.  For both long and short papers, all illustrations and appendices must be accommodated within these page limits, observing the formatting instructions given in the present document.  Papers that do not conform to the specified length and formatting requirements are subject to be rejected without review.

\section{Double-blind review process}
\label{sec:blind}

As the reviewing will be blind, the paper must not include the
authors' names and affiliations.  Furthermore, self-references that
reveal the author's identity, e.g., ``We previously showed (Smith,
1991) ...'' must be avoided. Instead, use citations such as ``Smith
previously showed (Smith, 1991) ...'' Papers that do not conform to
these requirements will be rejected without review. In addition,
please do not post your submissions on the web until after the
review process is complete (in special cases this is permitted: see 
the multiple submission policy below).

We will reject without review any papers that do not follow the
official style guidelines, anonymity conditions and page limits.

\section{Multiple Submission Policy}

Papers that have been or will be submitted to other meetings or
publications must indicate this at submission time. Authors of
papers accepted for presentation at ACL 2016 must notify the
program chairs by the camera-ready deadline as to whether the paper
will be presented. All accepted papers must be presented at the
conference to appear in the proceedings. We will not accept for
publication or presentation papers that overlap significantly in
content or results with papers that will be (or have been) published
elsewhere.

Preprint servers such as arXiv.org and ACL-related workshops that
do not have published proceedings in the ACL Anthology are not
considered archival for purposes of submission. Authors must state
in the online submission form the name of the workshop or preprint
server and title of the non-archival version.  The submitted version
should be suitably anonymized and not contain references to the
prior non-archival version. Reviewers will be told: ``The author(s)
have notified us that there exists a non-archival previous version
of this paper with significantly overlapping text. We have approved
submission under these circumstances, but to preserve the spirit
of blind review, the current submission does not reference the
non-archival version.'' Reviewers are free to do what they like with
this information.

Authors submitting more than one paper to ACL must ensure
that submissions do not overlap significantly ($>25\%$) with each other
in content or results. Authors should not submit short and long
versions of papers with substantial overlap in their original
contributions.

\section*{Acknowledgments}

Do not number the acknowledgment section.
This section should not be presented for the submission version.

\bibliography{acl2016}
\bibliographystyle{acl2016}

\end{document}

### 2016-01-22 acl.2016/acl2016.tex
start acl 2016 paper

@@
 
-\title{Instructions for ACL 2016 Proceedings\Thanks{This
-    document has been adapted from the instructions for earlier ACL
-    and NAACL proceedings, including those for
-    NAACL HLT16 by Margaret Mitchell, 
-    NAACL HLT15 by Matt Post and Adam Lopez,
-    NAACL HLT12 by Nizar Habash and William Schuler,
-    NAACL HLT10 by Claudia Leacock and Richard Wicentowski,
-    NAACL HLT09 by Joakim Nivre and Noah Smith, 
-    for ACL05 by Hwee Tou Ng and Kemal Oflazer,
-    for ACL02 by Eugene Charniak and Dekang Lin, and earlier ACL and
-    EACL formats.  Those versions were written by several people,
-    including John Chen, Henry S. Thompson and Donald Walker.
-    Additional elements were taken from the formatting instructions of
-    the {\em International Joint Conference on Artificial Intelligence}
-    and the {\em Conference on Computer Vision and Pattern Recognition}.}}
+\title{Fast, Scalable Phrase-Based SMT Decoding}
 
@@
 % at the top, where <new height> can be something larger than 2.25in
-\author{Author 1\\
-	    XYZ Company\\
-	    111 Anywhere Street\\
-	    Mytown, NY 10000, USA\\
-	    {\tt author1@xyz.org}
+\author{Hieu Hoang\\
+	    Moses Machine Translation CIC\\
+	    {\tt hieu@moses-mt.org}
 	  \And
@@
 \begin{abstract}
- This document contains instructions for preparing ACL 2016 submissions and camera-ready
-  manuscripts.  The document itself conforms
-  to its own specifications, and is therefore an example of what
-  your manuscript should look like.  Papers are required to conform to
-  all the directions reported in this document.  
-By using the provided \LaTeX\ and \BibTeX\ styles ({\small\tt acl2016.sty}, {\small\tt acl2016.bst}), the required formatting will be enabled by default.
+The utilization of statistical machine translation (SMT) has grown enormously over the last decade, many using open-source software developed by the NLP community.
+As commercial utilization has increased, there has been a pressing need that is optimized for their requirements.
+Specifically, faster phrase-based decoding, and more efficient utilization of modern multicore servers.
 
-I hate latex - Hieu.
+We present in this paper a re-assessment of the major components of phrase-based decoding and decoder implementation with particular emphasis on speed and scalability to multicore machines.
+The result is a drop-in replacement for the Moses decoder which is up to fifteen times faster and scales almost linearly with the number of cores.
+Furthermore, the decoder makes less search errors than the current Moses decoder.
 
@@
 \section{Introduction}
+
+SMT has been one of the outstanding success story from the NLP community in the last decade.
+It has transition from a mostly research discipline to services such as Google Translate, Microsoft Translator Hub, as well as services and products built around offline products such as the open-source Moses toolkit.
+The latter has spawned a cottage industry  encompassing a range of organizations and services from small language service providers that use SMT to reduce translation cost to large inter-govermental organizations such as the EU and the UN that provides high volume translation.
+
+For high volume users, decoding is a largest and most critical part of the translation process which needs to be fast and efficient.
+However, it has been noticed that the Moses decoder, amongst others, is unable to efficiently use multiple CPU cores that are now common on modern servers (reviewed paper, github discussion).
+That is, the time taken to decode a test set does not substantial decrease when more cores are used, in fact, decoding time may increase when more cores are added.
+The issue will only become more noticeable as the commercial use of SMT grows and the number of cores in servers increases.
+
+There have be speculation on the causes of the inefficiency as well as remedies.
+This paper is the first we know of that seeks to tackle this problem head on.
+We present an phrase-based decoder that is not only significantly faster than the Moses baseline for single-threaded operation, but is able to scale run multiple threads on multicore machines with only a slightly loss in linear speed.
+Model scores and functionality are compatible with Moses to aid comparison and ease of transition for users.
+All source code will be made available under an open-source license.
+
+\subsection{Prior Work}
+
+There are a number of open-source SMT projects, most includes a decoder.
+The most well known is Moses, which supports phrase-based models, hierarchical phrase-based as well as various syntax-based models.
+Joshua also supports hiearchical and syntax models and has recently supported phrase-based models.
+Phrasal supports a number of variants of the phrase-based model.
+CDEC supports hierarchical and syntactic models.
+
+A number of the decoders support multithreading whilst others use alternative methods such as Hadoop or external scripts to parallelize decoding.
+As we shall see in the Results Section, there are severe limitations to parallelizing using a external scripts to exectute multiple decoding processes.
+
+(mtplz) is perhaps closest in intent to this work. It describes a novel
 

### 2016-01-24 acl.2016/acl2016.tex
start acl 2016 paper

@@
 A number of the decoders support multithreading whilst others use alternative methods such as Hadoop or external scripts to parallelize decoding.
-As we shall see in the Results Section, there are severe limitations to parallelizing using a external scripts to exectute multiple decoding processes.
+We shall investigate the efficiency of using parallelizing decoding using the multi-processor approach.
+None of the decoder focus on multi-threads decoding.
 
-(mtplz) is perhaps closest in intent to this work. It describes a novel
+Other prior work look to optimizing specific components of decoding.
+(Liang and Chiang) describes the cube-pruning and cube-growing algorithm for decoding which allows the tradeoff between speed and translation quality to the adjusted with a simgle parameter.
+(KenLM) and (DALM) describes fast, efficient datastructures for language models.
+(Zen) describes an implementation of a phrase-table for an SMT decoder that is loaded on demand, reducing the initial loading time and memory requirements.
+(CompactPT) extends this by compressing the on-disk phrase table and lexicalized re-ordering model resulting in impressive speed gains over previous work.
+
+(mtplz) is perhaps closest in intent to this work.
+This takes a wholistic approach to decoding, describing a novel decoding algorithm which is fis focused on better decoding speed.
+It also describes a number of implementation details for faster decoding.
+However, the decoding algorithm is only able to incorporate one stateful feature function which precludes some of the useful decoding configurations which contains multiple stateful feature functions.
+It does not include a load-on-demand phrase table, therefore, cannot be used in a commercial environment where phrase-table has not be filtered with a know test set for any realistic size phrase-table.
+Neither is it multi-threaded.
+
+\section{BLAH BLAH}
 

### 2016-01-25 acl.2016/acl2016.tex
start acl 2016 paper

@@
 
+(Recently reviewed) describes running multiple processes of the Moses decoder for increased speed.
+
 Other prior work look to optimizing specific components of decoding.
@@
 It does not include a load-on-demand phrase table, therefore, cannot be used in a commercial environment where phrase-table has not be filtered with a know test set for any realistic size phrase-table.
-Neither is it multi-threaded.
+Neither did this paper analyze the scalability of their work to multicore servers.
+
+The rest of the paper will be broken up into the following sections.
+Next, we will describe the phrase-based model and the major implementation components, with particular emphasis on decoding time shortcomings.
+We will then describe modifications and present results.
+We conclude in the last section will discuss suggested improvements and future work.
+
+\section{Phrase-Based Model}
+
+The aim of decoding
 

### 2016-01-25 acl.2016/acl2016.tex
start acl 2016 paper

@@
 Next, we will describe the phrase-based model and the major implementation components, with particular emphasis on decoding time shortcomings.
-We will then describe modifications and present results.
-We conclude in the last section will discuss suggested improvements and future work.
+We will then describe modifications to improve decoding speed and present results.
+We conclude in the last section discuss suggested improvements and future work.
 
@@
 
-The aim of decoding
+The objective of decoding is to find the target translation with the maximum probability, given a source sentence.
+That is, for a source sentence $s$, the objective is to find a target translation $\hat{t}$ which has the highest conditional probability $p(t | s)$.
+Mathematically, this is written as:
+\begin{equation}
+\hat{t} = \arg \max_t p( t | s )
+\label{eq:argmax-factored-trans}
+\end{equation}
+where the \emph{arg max} function is the search.
+The log-linear model generalizes the noisy channel model to include more component models and weighting each model according to the contribution of each model to the total probability.
+\begin{equation}
+\label{eq:Log-Linear}
+p(t | s) 	=  \frac{1}{Z} \exp ( \sum_m \lambda_m h_m ( t, s)^{} )
+\end{equation}
+where $\lambda_m$ is the weight, and $h_m$ is the feature function, or `score', for model $m$.
+$Z$ is the partition function which can be ignored for optimization.
+The log-linear formulation in phrase-based SMT uses log probabilities as feature functions, in addition to features which do not have a probabilistic interpretation.
+Typical feature functions include the log transforms of the target language model probability $p(t)$, and translation model probabilities, $p_{TM}(t|s) $ and $p_{TM}(s|t)$, which we have suffixed with $_{TM} $ to avoid confusion with the overall model probability $ p(t|s) $ and $ p(s|t)$.
 

### 2016-01-25 acl.2016/acl2016.tex
start acl 2016 paper

@@
 Typical feature functions include the log transforms of the target language model probability $p(t)$, and translation model probabilities, $p_{TM}(t|s) $ and $p_{TM}(s|t)$, which we have suffixed with $_{TM} $ to avoid confusion with the overall model probability $ p(t|s) $ and $ p(s|t)$.
+
+\subsection{Beam Search}
+
+A translation of a source sentence is created by applying a series of translation rules which together translate each source word once, and only once.
+Each partial translation is called a \emph{hypothesis}, which is created by applying a rule to an existing hypothesis.
+This process is called \emph{hypothesis expansion} and starts with a hypothesis that has translated no source word and ends with a completed hypothesis that has translated all source words.
+The highest-scoring completed hypothesis, according to the model score, is returned as  most probable translation, $\hat{t} $.
+Incomplete hypotheses are referred to as partial hypotheses.
+
+Each rule translates a contiguous sequence of source words but successive translation options do not have to be adjacent on the source side, depending on the distortion limit.
+However, the target output is constructed strictly left-to-right from the target string of successive translation options.
+Therefore, successive translation options which are not adjacent and monotonic in the source causes translation reordering.
+
+A beam search algorithm is used to create the completed hypothesis set efficiently.
+Partial hypotheses are organized into stacks where each stack holds a number of comparable hypotheses.
+Hypotheses in the same stack have the same coverage cardinality $|C|$, where $C$ is  the coverage set, $C \subseteq \{1,2,...
+|s| \} $ of the number of  source words translated.
+Therefore, $|s| + 1$ number of stacks are created for the decoding of a sentence $s$.
+%There exist other stack layouts \citep{ortizmartinez-garciavarea-casacuberta:2006:WMT} but the use of coverage cardinality is the most common and the method we use.
 

### 2016-01-27 acl.2016/acl2016.tex
fixing segfault on thrud

@@
 where the \emph{arg max} function is the search.
-The log-linear model generalizes the noisy channel model to include more component models and weighting each model according to the contribution of each model to the total probability.
+The log-linear model generalizes Equation~\ref{eq:argmax-factored-trans} to include more component models and weighting each model according to the contribution of each model to the total probability.
 \begin{equation}
@@
 \end{equation}
-where $\lambda_m$ is the weight, and $h_m$ is the feature function, or `score', for model $m$.
-$Z$ is the partition function which can be ignored for optimization.
-The log-linear formulation in phrase-based SMT uses log probabilities as feature functions, in addition to features which do not have a probabilistic interpretation.
-Typical feature functions include the log transforms of the target language model probability $p(t)$, and translation model probabilities, $p_{TM}(t|s) $ and $p_{TM}(s|t)$, which we have suffixed with $_{TM} $ to avoid confusion with the overall model probability $ p(t|s) $ and $ p(s|t)$.
+where $\lambda_m$ is the weight, and $h_m$ is the feature function, or `score', for model $m$. $Z$ is the partition function which can be ignored for optimization.
 
@@
 %There exist other stack layouts \citep{ortizmartinez-garciavarea-casacuberta:2006:WMT} but the use of coverage cardinality is the most common and the method we use.
+
+\subsection{Feature Functions}
+
+Features functions are the $h_m$ in Equation~\ref{eq:Log-Linear}, calculating a score for each hypothesis which is then weighted and summed to give an overall score.
+
+Standard feature functions in the phrase-based model include:
+\begin{enumerate} 
+  \item log transforms of the target language model probability $p(t)$, 
+  \item log transforms translation model probabilities, $p_{TM}(t|s) $ and $p_{TM}(s|t)$, and word-based translation probabilities $p_w(t|s)$ and $p_w(s|t)$,
+  \item log transforms of the lexicalized re-ordering probabilities,
+  \item a distortion penalty
+  \item a phrase-penalty,
+  \item a word penalty,
+  \item an unknown word penalty.
+\end{enumerate}
+%which we have suffixed with $_{TM} $ to avoid confusion with the overall model probability $ p(t|s) $ and $ p(s|t)$.
+
+The first three feature functions frequently trained on data and require the feature to read the model from files. The other feature functions do not require model files.
+
+\section{Experimental Setup}
+
+We trained two phrase-based systems using the Moses toolkit, with standard settings.
+The first system was trained on most of the publicly available Arabic-English data from Opus (Jörg Tiedemann, 2012,) consisting of over 69 million parallel sentences, and tuned on a held out set.
+The second system was trained on the French-English Europarl corpus.
+The phrase-tables were then pruned, keeping only the top 100 entries per source phrase, according to $p(t|s)$.
+All models files were then binarized; thelanguage models were binaized using KenLM (???), the phrase table using Probing PT (???), lexicalized reordering model using the compact datastructure described in ???.
+These binary formats were choosen for their best-of-class multithreaded performance.
+Table~\ref{tab:model-files} gives details of the resultant sizes of the model files.
+\begin{table}[h]
+\begin{center}
+\begin{tabular}{|l|c|c|} \hline
+		& ar-en	& fr-en \\ \hline
+Phrase table  	& 17 	& 5.8 \\
+Language model 	& 3.1  	& 1.8 \\ 
+Lex-re model	& 2.3	& 637MB \\ \hline
+\end{tabular}
+\end{center}
+\caption{Model sizes in GB}
+\label{tab:model-files}
+\end{table}
+
+For testing decoding speed, we used a subset of the training data, Table~\ref{tab:test-sets}.
+The two test set have differing characterics that we are interested in analyzing, ar-en have short sentences while fr-en have overly long sentences.
+
+\begin{table}[h]
+\begin{center}
+\begin{tabular}{|l|c|c|} \hline
+		& ar-en	& fr-en \\ \hline
+\# sentences  	& 800k 	& 200k \\
+\# words 	& 5.8m 	& 5.9m \\ 
+\Avg words/sent	& 7.3	& 29.7 \\ \hline
+\end{tabular}
+\end{center}
+\caption{Model sizes in GB}
+\label{tab:test-sets}
+\end{table}
+
+Where we need to compare the model score of the algorithms, we used a held out set; ??? for ar-en and ??? for fr-en.
+
+Standard Moses phrase-based configurations are used, except that we use the cube-pruning algorithm (???) with a pop-limit of 400, rather than the basic phrase-based algorithm.
+The cube-pruning algorithm is often employ by users who are sensitive to decoding speed as it gives them the ability to trade speed with translation quality with a simple pop-limit parameter.
+
+\section{Results}
 

### 2016-01-28 acl.2016/acl2016.tex
fixing segfault on thrud

@@
 Standard Moses phrase-based configurations are used, except that we use the cube-pruning algorithm (???) with a pop-limit of 400, rather than the basic phrase-based algorithm.
-The cube-pruning algorithm is often employ by users who are sensitive to decoding speed as it gives them the ability to trade speed with translation quality with a simple pop-limit parameter.
+The cube-pruning algorithm is often employ by users who require fast decoding as it gives them the ability to trade speed with translation quality with a simple pop-limit parameter.
 

### 2016-01-28 acl.2016/acl2016.tex
fixing segfault on thrud

@@
 
+There are two main optimization to the search that we shall investigate.
+Firstly, the search creates and destroy a large number of hypothesis objects in memory which puts a heavy burden on the operating system.
+Also, the speed of memory access is dependent on whether the data is in the CPU cache.
+We shall optimize the search algorithm to use memory pools and object pools, replacing the operating system's general purpose memory management with our own application-aware management.
+We shall also seek to re-use recently accessed information to increase likelihood of the data being in the CPU cache.
+
+Secondly, we shall investigate different stack configurations other than coverage cardinality to see whether they can improve the speed / model score ratio.
+
 \subsection{Feature Functions}
@@
 \begin{enumerate}
-  \item log transforms of the target language model probability $p(t)$, 
   \item log transforms translation model probabilities, $p_{TM}(t|s) $ and $p_{TM}(s|t)$, and word-based translation probabilities $p_w(t|s)$ and $p_w(s|t)$,
   \item log transforms of the lexicalized re-ordering probabilities,
+  \item log transforms of the target language model probability $p(t)$, 
   \item a distortion penalty
@@
 
-The first three feature functions frequently trained on data and require the feature to read the model from files. The other feature functions do not require model files.
+The first three feature functions frequently trained on data and require the feature to read the model from files.
+The other feature functions do not require model files.
+We shall investigate the first two feature functions for optimization.
+
+%\subsection{Language Model}
+% TODO
+%
+\subsection{Translation Model}
+
+Load-on-demand 'binary' phrase-tables are often used for MT deployment due to the advantages of fast loading and fast querying speed, and because they can be used with large phrase-tables.
+We therefore focus on optimizing decoding speed with these phrase-tables, specifically the Probing PT.
+
+We shall look at the caching strategies to reduce the number of phrase-table lookups. We shall also investigate the datastructures used by the phrase-table and their impact on decoding speed.
+
+\subsection{Lexicalized Reordering Model}
 
@@
 \# words 	& 5.8m 	& 5.9m \\ 
-\Avg words/sent	& 7.3	& 29.7 \\ \hline
+Avg words/sent	& 7.3	& 29.7 \\ \hline
 \end{tabular}

### 2016-01-28 acl.2016/acl2016.tex
fixing segfault on thrud

@@
 
-Features functions are the $h_m$ in Equation~\ref{eq:Log-Linear}, calculating a score for each hypothesis which is then weighted and summed to give an overall score.
+Features functions are the $h_m$ in Equation~\ref{eq:Log-Linear}, calculating a score for each hypothesis.
 
-Standard feature functions in the phrase-based model include:
+The standard feature functions in the phrase-based model include:
 \begin{enumerate}
@@
 
-Load-on-demand 'binary' phrase-tables are often used for MT deployment due to the advantages of fast loading and fast querying speed, and because they can be used with large phrase-tables.
+Load-on-demand \emph{'binary'} phrase-tables are often used for MT deployment due to their fast loading and fast querying speed, and because they can be used with large phrase-tables.
 We therefore focus on optimizing decoding speed with these phrase-tables, specifically the Probing PT.

### 2016-01-28 acl.2016/acl2016.tex
acl paper

@@
 
-There are two main optimization to the search that we shall investigate.
+There are three main optimization to the search that we shall investigate.
 Firstly, the search creates and destroy a large number of hypothesis objects in memory which puts a heavy burden on the operating system.
-Also, the speed of memory access is dependent on whether the data is in the CPU cache.
 We shall optimize the search algorithm to use memory pools and object pools, replacing the operating system's general purpose memory management with our own application-aware management.
-We shall also seek to re-use recently accessed information to increase likelihood of the data being in the CPU cache.
 
-Secondly, we shall investigate different stack configurations other than coverage cardinality to see whether they can improve the speed / model score ratio.
+The speed of memory access is dependent on whether the data is in the CPU cache which is a constrained resource compared to memory size, typically 20MB in the latest processors.
+We shall seek to re-use recently accessed information to increase likelihood of the data being in the CPU cache.
+
+In multiprocessor servers, the CPU cache is attached to each processor and each core.
+If a sentences is being decoded on one CPU is switched to another, the CPU cache on the new CPU must be repopulated, slowing down decoding.
+We will therefore investigate binding threads to specific cores.
+
+Lastly, we shall investigate different stack configurations other than coverage cardinality to see whether they can improve the speed / model score ratio.
 
@@
 
-Load-on-demand \emph{'binary'} phrase-tables are often used for MT deployment due to their fast loading and fast querying speed, and because they can be used with large phrase-tables.
+Load-on-demand \emph{'binary'} phrase-tables are often used for MT deployment due to their fast loading and querying speed, and because they can be used with large phrase-tables.
 We therefore focus on optimizing decoding speed with these phrase-tables, specifically the Probing PT.
@@
 \end{center}
-\caption{Model sizes in GB}
+\caption{Test sets}
 \label{tab:test-sets}

### 2016-01-28 acl.2016/acl2016.tex
acl paper

@@
 Therefore, $|s| + 1$ number of stacks are created for the decoding of a sentence $s$.
-%There exist other stack layouts \citep{ortizmartinez-garciavarea-casacuberta:2006:WMT} but the use of coverage cardinality is the most common and the method we use.
 
@@
 
-Lastly, we shall investigate different stack configurations other than coverage cardinality to see whether they can improve the speed / model score ratio.
+Lastly, we shall investigate different stack configurations other than coverage cardinality to see whether they can improve speed and translation quality.
 
@@
 \subsection{Lexicalized Reordering Model}
+
+The lexicalized reordering model is trained on parallel data, usually requiring random lookups of the model file during decoding.
+However, the key to the lookup are the source and target phrase of each translation rule.
+We shall investigate the advantages of storing the mode data within the translation rule.
 
@@
 
+As a baseline, we use the latest version of the Moses decoder taken from the github repository.
+
+For all experiments, we used a Dell PowerEdge R620 server with 16 cores, 32 hyper-threads, split over 2 physical processors (Intel Xeon E5-2650 @ 2.00GHz).
+The server has 380GB RAM.
+The operating system was Ubuntu 14.04, the code was compiled with gcc 4.8.4 and Boost library 1.59.
+
 \section{Results}
+
+\subsection{Optimizing Memory}
+
+We create a dynamic memory pool which can grow as more memory is requested.
+The memory is not released, instead the pool can be reset in order for the memory to be re-used.
+We instantiate two pools for each thread, one which is never reset and another which is reset after the decoding each sentence.
+Objects are created in either pool according to their life cycle.
+
+For critical objects with high churn such as the hypotheses, thread-specific LIFO queues are used to recycle objects which are no longer used.
+This not only reduces memory wastage but re-uses recent objects which are likely to be in the CPU cache.
+
+Over 30\% of the Moses decoder running time is spent on memory management and this increases when more threads are used, Table~\ref{tab:optimizing-memory}.
+By contrast, our decoder spends ???\% on memory management.
+\begin{table}[h]
+\begin{center}
+\begin{tabular}{|l|c|c|} \hline
+		& 1 thread	& 30 threads \\ \hline
+Moses  		& ???\% 	& ???\% \\
+Our work 	& ???\% 	& ???\% \\ \hline
+\end{tabular}
+\end{center}
+\caption{\%age decoding time managing memory}
+\label{tab:optimizing-memory}
+\end{table}
+
+A potential disadvantage of using memory pool is that the decoder will use more memory, however, Table~\ref{tab:memory-usage} shows that peak memory consumption are comparable.
+\begin{table}[h]
+\begin{center}
+\begin{tabular}{|l|c|c|} \hline
+		& 1 thread	& 30 threads \\ \hline
+Moses  		& ???	 	& ??? \\
+Our work 	& ??? 		& ??? \\ \hline
+\end{tabular}
+\end{center}
+\caption{Peak Memory Usage in GB}
+\label{tab:memory-usage}
+\end{table}
+
+\subsection{Stack Configuration}
+
+The most popular stack configuration for phrase-based model, as implemented in Pharaoh, Moses and Joshua, has been by coverage cardinality, ie.
+hypotheses that have translated the same number of source words are stored in the same stack.
+There have been research into other stack layouts \citep{ortizmartinez-garciavarea-casacuberta:2006:WMT}, and it has been noted that the decoder described in (???IBM) uses coverage stacks, as opposed to coverage cardinality.
+We re-visit this issue by testing the advantages of stack coverage and cardinality layout.
+
+However, we also note that distortion limit which constrains hypothesis extension is dependent on the hypothesis' coverage vector \emph{and} the end position of most recent source word that has been translated.
+Therefore, if the hypotheses are separated by coverage vector \emph{and} end position, the distortion limit constraint can be moved to an outer loop, potentially speeding up decoding.
 

### 2016-01-28 acl.2016/acl2016.tex
acl paper

@@
 \usepackage{latexsym}
+\usepackage{algorithmic}
 
@@
 \begin{tabular}{|l|c|c|} \hline
-		& 1 thread	& 30 threads \\ \hline
+		& 1 thread	& 32 threads \\ \hline
 Moses  		& ???\% 	& ???\% \\
@@
 \begin{tabular}{|l|c|c|} \hline
-		& 1 thread	& 30 threads \\ \hline
+		& 1 thread	& 32 threads \\ \hline
 Moses  		& ???	 	& ??? \\
@@
 However, we also note that distortion limit which constrains hypothesis extension is dependent on the hypothesis' coverage vector \emph{and} the end position of most recent source word that has been translated.
-Therefore, if the hypotheses are separated by coverage vector \emph{and} end position, the distortion limit constraint can be moved to an outer loop, potentially speeding up decoding.
+Using a cardinality-based stack, the distortion limit must be checked for every instance, Figure~\ref{algo:Hypothesis Expansion with Unsorted Stack}.
+\begin{figure} [h]
+\begin{algorithmic}
+%\REQUIRE source sentence $s$, translation options
+\STATE insert empty hypothesis into $stack_0$
+\FOR{$i = 0$ to $|s|-1$} 
+  \STATE prune $stack_i$
+  \FORALL{ hypothesis in $stack_i$} 
+    \FORALL{ translation options}
+       \IF {hypothesis and translation option are compatible}
+         \STATE expand hypothesis with translation option $\rightarrow$ new hypothesis
+         \STATE coverage of  new hypothesis $\rightarrow C$
+         \STATE insert new hypothesis $\rightarrow stack_{|C|}$
+         \STATE recombine new hypothesis if possible
+         \STATE prune $stack_{|C|}$ if necessary
+       \ENDIF
+    \ENDFOR % { translation options}
+  \ENDFOR % { hypothesis in $stack_i$} 
+\ENDFOR% {$i = 0$ to $|s|-1$} 
+\RETURN highest scoring hypothesis in $stack_{|s|}$
+\end{algorithmic}
+\caption{Hypothesis Expansion with Unsorted Stack}
+\label{algo:Hypothesis Expansion with Unsorted Stack}
+\end{figure}
+
+Therefore, if the hypotheses are separated by coverage vector and end position, the distortion limit constraint can be moved to an outer loop, potentially speeding up decoding.
 

### 2016-01-28 acl.2016/acl2016.tex
acl paper

@@
 
-However, we also note that distortion limit which constrains hypothesis extension is dependent on the hypothesis' coverage vector \emph{and} the end position of most recent source word that has been translated.
-Using a cardinality-based stack, the distortion limit must be checked for every instance, Figure~\ref{algo:Hypothesis Expansion with Unsorted Stack}.
+We also note that distortion limit which constrains hypothesis extension is dependent on the hypothesis' coverage vector \emph{and} the end position of most recent source word that has been translated.
+The distortion limit must be checked for every instance of a hypothesis and translation rule, Figure~\ref{algo:Hypothesis Expansion with Unsorted Stack}.
 \begin{figure} [h]
@@
 %\REQUIRE source sentence $s$, translation options
-\STATE insert empty hypothesis into $stack_0$
-\FOR{$i = 0$ to $|s|-1$} 
-  \STATE prune $stack_i$
-  \FORALL{ hypothesis in $stack_i$} 
-    \FORALL{ translation options}
-       \IF {hypothesis and translation option are compatible}
-         \STATE expand hypothesis with translation option $\rightarrow$ new hypothesis
-         \STATE coverage of  new hypothesis $\rightarrow C$
-         \STATE insert new hypothesis $\rightarrow stack_{|C|}$
-         \STATE recombine new hypothesis if possible
-         \STATE prune $stack_{|C|}$ if necessary
+\FORALL{ hypo in $stack$} 
+  \FORALL{ translation rules}
+      \IF {CanExpand(hypo coverage, hypo end position, translation rule range)}
+	\STATE expand hypo with translation rule $\rightarrow$ new hypo
+	\STATE add new hypo into next stack
       \ENDIF
@@
 \ENDFOR % { hypothesis in $stack_i$} 
-\ENDFOR% {$i = 0$ to $|s|-1$} 
-\RETURN highest scoring hypothesis in $stack_{|s|}$
 \end{algorithmic}
@@
 \end{figure}
-
-Therefore, if the hypotheses are separated by coverage vector and end position, the distortion limit constraint can be moved to an outer loop, potentially speeding up decoding.
+However, by separating hypotheses according to coverage and end position, the distortion limit check only needs to be done for each stack, potentially speeding up decoding, Figure~\ref{algo:Hypothesis Expansion with Sorted Stack}.
+\begin{figure} [h]
+\begin{algorithmic}
+%\REQUIRE source sentence $s$, translation options
+\FORALL{ coverage and end position in $stack$}
+  \FORALL{ translation rules}
+      \IF {CanExpand(coverage, end position, translation rule range)}
+        \FORALL{hypo in [coverage and end position]}
+  	  \STATE expand hypo with translation rule $\rightarrow$ new hypo
+	  \STATE add new hypo into $next stack$
+	\ENDFOR
+      \ENDIF
+  \ENDFOR % { translation options}
+\ENDFOR % { hypothesis in $stack_i$} 
+\end{algorithmic}
+\caption{Hypothesis Expansion with Sorted Stack}
+\label{algo:Hypothesis Expansion with Sorted Stack}
+\end{figure}
 

### 2016-01-28 acl.2016/acl2016.tex
acl paper

@@
 
-We also note that distortion limit which constrains hypothesis extension is dependent on the hypothesis' coverage vector \emph{and} the end position of most recent source word that has been translated.
+We also note that distortion limit which constrains hypothesis extension is dependent on the hypothesis' coverage vector, $C$ \emph{and} the end position of most recent source word that has been translated, $e$.
 The distortion limit must be checked for every instance of a hypothesis and translation rule, Figure~\ref{algo:Hypothesis Expansion with Unsorted Stack}.
@@
 %\REQUIRE source sentence $s$, translation options
-\FORALL{ hypo in $stack$} 
+\FORALL{ $hypo$ in $stack_{|C|}$} 
   \FORALL{ translation rules}
-      \IF {CanExpand(hypo coverage, hypo end position, translation rule range)}
+      \IF {can-expand($C(hypo)$, $e(hypo)$, translation rule range)}
 	\STATE expand hypo with translation rule $\rightarrow$ new hypo
@@
 \end{figure}
-However, by separating hypotheses according to coverage and end position, the distortion limit check only needs to be done for each stack, potentially speeding up decoding, Figure~\ref{algo:Hypothesis Expansion with Sorted Stack}.
+However, by separating hypotheses into \emph{'ministack'} according to coverage and end position, the distortion limit only needs to be checked for each ministack, Figure~\ref{algo:Hypothesis Expansion with Sorted Stack}.
 \begin{figure} [h]
@@
 %\REQUIRE source sentence $s$, translation options
-\FORALL{ coverage and end position in $stack$}
+\FORALL{ $ministack_{C, e}$ in $stack_{|C|}$} 
   \FORALL{ translation rules}
-      \IF {CanExpand(coverage, end position, translation rule range)}
-        \FORALL{hypo in [coverage and end position]}
+    \IF {can-expand($C$, $e$, translation rule range)}
+      \FORALL{ $hypo$ in $ministack_{C, e}$} 
 	\STATE expand hypo with translation rule $\rightarrow$ new hypo
-	  \STATE add new hypo into $next stack$
+	\STATE add new hypo into next stack
       \ENDFOR
@@
 \end{figure}
+We therefore looked at the effects of three stack configurations.
+\begin{enumerate}
+  \item coverage cardinality,
+  \item coverage,
+  \item coverage and end position of most recent translated source word.
+\end{enumerate}
 

### 2016-01-29 acl.2016/acl2016.tex
acl paper

@@
 hypotheses that have translated the same number of source words are stored in the same stack.
-There have been research into other stack layouts \citep{ortizmartinez-garciavarea-casacuberta:2006:WMT}, and it has been noted that the decoder described in (???IBM) uses coverage stacks, as opposed to coverage cardinality.
+There have been research into other stack layouts ~\cite{ortizmartinez-garciavarea-casacuberta:2006:WMT}, and it has been noted that the decoder described in (???IBM) uses coverage stacks, as opposed to coverage cardinality.
 We re-visit this issue by testing the advantages of stack coverage and cardinality layout.
@@
 
-\bibliography{acl2016}
+\bibliography{acl2016,mt,more}
 \bibliographystyle{acl2016}

### 2016-01-29 acl.2016/acl2016.tex
acl paper

@@
 hypotheses that have translated the same number of source words are stored in the same stack.
-There have been research into other stack layouts ~\cite{ortizmartinez-garciavarea-casacuberta:2006:WMT}, and it has been noted that the decoder described in (???IBM) uses coverage stacks, as opposed to coverage cardinality.
-We re-visit this issue by testing the advantages of stack coverage and cardinality layout.
+There have been research into other stack layouts such as~\cite{ortizmartinez-garciavarea-casacuberta:2006:WMT}, and it has been noted that the decoder in~\cite{Brown:1993} uses coverage stacks, as opposed to coverage cardinality.
 

### 2016-01-29 acl.2016/acl2016.tex
acl paper

@@
 
-SMT has been one of the outstanding success story from the NLP community in the last decade.
-It has transition from a mostly research discipline to services such as Google Translate, Microsoft Translator Hub, as well as services and products built around offline products such as the open-source Moses toolkit.
-The latter has spawned a cottage industry  encompassing a range of organizations and services from small language service providers that use SMT to reduce translation cost to large inter-govermental organizations such as the EU and the UN that provides high volume translation.
+SMT has been one of the outstanding success story from the NLP community in the last decade, progressing from a mostly research discipline to public useability via services such as Google Translate, Microsoft Translator Hub, as well as services and products built around offline products such as Language Weaver and the open-source Moses toolkit.
+The latter has spawned a cottage industry  encompassing a range of organizations and services from small language service providers seeking to reduce translation cost, to large inter-govermental organizations such as the EU and the UN that require high volume, high quality translation.
 
@@
 That is, the time taken to decode a test set does not substantial decrease when more cores are used, in fact, decoding time may increase when more cores are added.
-The issue will only become more noticeable as the commercial use of SMT grows and the number of cores in servers increases.
+The problem will continue to grow as the commercial use of SMT increases and the number of CPU cores increases.
 
-There have be speculation on the causes of the inefficiency as well as remedies.
+There have be speculation on the causes of the inefficiency as well as potential remedies.
 This paper is the first we know of that seeks to tackle this problem head on.

### 2016-01-29 acl.2016/acl2016.tex
acl paper

@@
 \end{figure}
-However, by separating hypotheses into \emph{'ministack'} according to coverage and end position, the distortion limit only needs to be checked for each ministack, Figure~\ref{algo:Hypothesis Expansion with Sorted Stack}.
+However, by separating hypotheses into \emph{'ministacks'} according to coverage and end position, the distortion limit only needs to be checked for each ministack, Figure~\ref{algo:Hypothesis Expansion with Sorted Stack}.
 \begin{figure} [h]

### 2016-01-29 acl.2016/acl2016.tex
acl paper

@@
 
-We trained two phrase-based systems using the Moses toolkit, with standard settings.
-The first system was trained on most of the publicly available Arabic-English data from Opus (Jörg Tiedemann, 2012,) consisting of over 69 million parallel sentences, and tuned on a held out set.
-The second system was trained on the French-English Europarl corpus.
-The phrase-tables were then pruned, keeping only the top 100 entries per source phrase, according to $p(t|s)$.
-All models files were then binarized; thelanguage models were binaized using KenLM (???), the phrase table using Probing PT (???), lexicalized reordering model using the compact datastructure described in ???.
+We trained a phrase-based system using the Moses toolkit with standard settings.
+The training data consisted of most of the publicly available Arabic-English data from Opus (Jörg Tiedemann, 2012,) containing over 69 million parallel sentences, and tuned on a held out set.
+The phrase-table was then pruned, keeping only the top 100 entries per source phrase, according to $p(t|s)$.
+All models files were then binarized; the language models were binaized using KenLM (???), the phrase table using Probing PT (???), lexicalized reordering model using the compact datastructure described in ???.
 These binary formats were choosen for their best-of-class multithreaded performance.
 Table~\ref{tab:model-files} gives details of the resultant sizes of the model files.
+For verification with a different dataset, we also used a second system trained on the French-English Europarl corpus.
 \begin{table}[h]
@@
 \end{table}
-
 For testing decoding speed, we used a subset of the training data, Table~\ref{tab:test-sets}.
@@
 
-Over 30\% of the Moses decoder running time is spent on memory management and this increases when more threads are used, Table~\ref{tab:optimizing-memory}.
-By contrast, our decoder spends ???\% on memory management.
+Over 24\% of the Moses decoder running time is spent on memory management and this increases when more threads are used, Table~\ref{tab:optimizing-memory}, braking the scaling of the decoder to many cores.
+By contrast, our decoder spends 11\% on memory management and does not significantly increase with more cores.
 \begin{table}[h]
 \begin{center}
-\begin{tabular}{|l|c|c|} \hline
-		& 1 thread	& 32 threads \\ \hline
-Moses  		& ???\% 	& ???\% \\
-Our work 	& ???\% 	& ???\% \\ \hline
+\begin{tabular}{|l|c|c|c|c|} \hline
+		& \multicolumn{2}{c|} {Moses}	& \multicolumn{2}{c|} {Our Work} \\ \hline
+\# threads	& 1 		& 32	& 1 		& 32  \\ \hline
+Memory  	& 24		& 39 	& 11		& 13 \\
+LM 		& 12	 	& 2 	& 47		& 38 \\ 
+Phrase-table	& 9	 	& 5 	& 2		& 4 \\ 
+Lex RO 		& 8	 	& 2 	& 2		& 2 \\ 
+Search 		& 2	 	& 0 	& 14		& 19 \\ 
+Misc/Unknown	& 45	 	& 39 	& 24		& 29 \\ \hline
 \end{tabular}
 \end{center}
-\caption{\%age decoding time managing memory}
+\caption{Profile of \%age decoding time}
 \label{tab:optimizing-memory}
@@
 
-A potential disadvantage of using memory pool is that the decoder will use more memory, however, Table~\ref{tab:memory-usage} shows that peak memory consumption are comparable.
-\begin{table}[h]
-\begin{center}
-\begin{tabular}{|l|c|c|} \hline
-		& 1 thread	& 32 threads \\ \hline
-Moses  		& ???	 	& ??? \\
-Our work 	& ??? 		& ??? \\ \hline
-\end{tabular}
-\end{center}
-\caption{Peak Memory Usage in GB}
-\label{tab:memory-usage}
-\end{table}
+% A potential disadvantage of using memory pool is that the decoder will use more memory, however, Table~\ref{tab:memory-usage} shows that peak memory consumption are comparable.
+% \begin{table}[h]
+% \begin{center}
+% \begin{tabular}{|l|c|c|} \hline
+% 		& 1 thread	& 32 threads \\ \hline
+% Moses  		& ???	 	& ??? \\
+% Our work 	& ??? 		& ??? \\ \hline
+% \end{tabular}
+% \end{center}
+% \caption{Peak Memory Usage in GB}
+% \label{tab:memory-usage}
+% \end{table}
 

### 2016-01-29 acl.2016/acl2016.tex
acl paper

@@
 	\STATE expand hypo with translation rule $\rightarrow$ new hypo
-	\STATE add new hypo into next stack
+	\STATE add new hypo to next stack
       \ENDIF
@@
 	\STATE expand hypo with translation rule $\rightarrow$ new hypo
-	\STATE add new hypo into next stack
+	\STATE add new hypo to next ministack
       \ENDFOR
@@
 \end{enumerate}
+We therefore looked at the effects of three stack configurations.
+\begin{enumerate}
+  \item coverage cardinality,
+  \item coverage,
+  \item coverage and end position of most recent translated source word.
+\end{enumerate}
+Table~\ref{tab:stack-configuration} present the decoding time (with 32 threads) for each stack configuration, and the average model on the held out test set.
+\begin{table}[h]
+\begin{center}
+\begin{tabular}{|l|p{17mm}|p{17mm}|} \hline
+			& Decoding \newline Time(secs)	& Avg model \newline score \\ \hline
+Coverage cardinality  	& 186 	& -8.59922 \\
+Coverage 		& ??? 	& ??? \\ 
+Coverage \& end pos	& 196	& -8.59922 \\ \hline
+\end{tabular}
+\end{center}
+\caption{Decoding time and average model scores for different stack configurations}
+\label{tab:stack-configuration}
+\end{table}
+
+The algorithms in Figure~\ref{algo:Hypothesis Expansion with Unsorted Stack} and ~\ref{algo:Hypothesis Expansion with Sorted Stack} describes the standard phrase-based decoding algorithm, however, the stack configuration issues are the same in the cube-pruning algorithm.
+Furthermore, the cube-pruning algorithm also contain a priority queue which is attached to coverage cardinality in Moses, Figure~\ref{algo:Cube Pruning with Unsorted Stack}.
+\begin{figure} [h]
+\begin{algorithmic}
+%\REQUIRE source sentence $s$, translation options
+\STATE initialize $queue_{|C|}$
+\FOR{1 to pop-limit}
+  \STATE get best $item$ in $queue_{|C|}$
+  \STATE create new hypo from $item$
+  \STATE add new hypo to new $stack_{|C|}$
+  \STATE create next $items$
+  \STATE add new $items$ to $queue_{|C|}$
+\ENDFOR % { hypothesis in $stack_i$} 
+\end{algorithmic}
+\caption{Cube Pruning with Unsorted Stack}
+\label{algo:Cube Pruning with Unsorted Stack}
+\end{figure}
+We experiment with changing the queue, $queue_{|C|}$, by having separate queues for each (1) unique coverage $queue_{C}$, and each (2) unique coverage and end position $queue_{C, e}$.
+The stack configuration stayed constant.
 

### 2016-02-03 acl.2016/acl2016.tex
acl paper

@@
 All models files were then binarized; the language models were binaized using KenLM (???), the phrase table using Probing PT (???), lexicalized reordering model using the compact datastructure described in ???.
-These binary formats were choosen for their best-of-class multithreaded performance.
+These binary formats were choosen for their best-in-class multithreaded performance.
 Table~\ref{tab:model-files} gives details of the resultant sizes of the model files.
-For verification with a different dataset, we also used a second system trained on the French-English Europarl corpus.
+For verification with a different dataset, we also occasionally used a second system trained on the French-English Europarl corpus (2m parallel sentences).
 \begin{table}[h]
@@
 For testing decoding speed, we used a subset of the training data, Table~\ref{tab:test-sets}.
-The two test set have differing characterics that we are interested in analyzing, ar-en have short sentences while fr-en have overly long sentences.
+The two test scenarios have differing characterics that we are interested in analyzing, ar-en have short sentences with large models while fr-en have overly long sentences with smaller models.
 
@@
 
-Where we need to compare the model score of the algorithms, we used a held out set; ??? for ar-en and ??? for fr-en.
+Where we need to compare model scores, we used held out test sets; a 2,000 sentence subset of OpenSubtitles for ar-en and the 3,003 sentence newstest2011 test set for fr-en.
 
@@
 
-Over 24\% of the Moses decoder running time is spent on memory management and this increases when more threads are used, Table~\ref{tab:optimizing-memory}, braking the scaling of the decoder to many cores.
+Over 24\% of the Moses decoder running time is spent on memory management and this increases when more threads are used, Table~\ref{tab:optimizing-memory}, dampening the scalability ofthe decoder.
 By contrast, our decoder spends 11\% on memory management and does not significantly increase with more cores.
@@
 \end{enumerate}
-Table~\ref{tab:stack-configuration} present the decoding time (with 32 threads) for each stack configuration, and the average model on the held out test set.
+Table~\ref{tab:stack-configuration-model-score} present the average model on the held out test set, while Table~\ref{tab:stack-configuration-time} presents the decoding time.
 \begin{table}[h]
+\small
 \begin{center}
-\begin{tabular}{|l|p{17mm}|p{17mm}|} \hline
-			& Decoding \newline Time(secs)	& Avg model \newline score \\ \hline
-Coverage cardinality  	& 186 	& -8.59922 \\
-Coverage 		& ??? 	& ??? \\ 
-Coverage \& end pos	& 196	& -8.59922 \\ \hline
+\begin{tabular}{|r|p{12mm}|p{12mm}|p{13mm}|} \hline
+Pop-limit	& Coverage \newline cardinality	& Coverage	& Coverage \newline \& end pos \\ \hline
+100  		& -8.64513 			& ???	& -8.64513 \\
+500		& -8.59563			& ??? 	& -8.59612 \\
+1,000	 	& -8.5870			& ???	& -8.5870 \\ 
+5,000 		& -8.58165			& ???	& -8.58098 \\ 
+10,000 		& -8.58133			& ???	& -8.58015 \\ \hline
 \end{tabular}
 \end{center}
-\caption{Decoding time and average model scores for different stack configurations}
-\label{tab:stack-configuration}
+\caption{Average model scores for different stack configurations}
+\label{tab:stack-configuration-model-score}
+\end{table}
+
+\begin{table}[h]
+\small
+\begin{center}
+\begin{tabular}{|r|p{12mm}|p{12mm}|p{13mm}|} \hline
+Pop-limit	& Coverage \newline cardinality	& Coverage	& Coverage \newline \& end pos \\ \hline
+100  		& 73 			& ???	& 72 \\
+500		& 237			& ??? 	& 229 \\
+1,000	 	& 416			& ???	& 423 \\ 
+5,000 		& 1,930			& ???	& 2153 \\ 
+10,000 		& 			& ???	& 4576 \\ \hline
+\end{tabular}
+\end{center}
+\caption{Decoding time with 32-threads for different stack configurations (sec)}
+\label{tab:stack-configuration-time}
 \end{table}

### 2016-02-03 acl.2016/acl2016.tex
acl paper

@@
 Pop-limit	& Coverage \newline cardinality	& Coverage	& Coverage \newline \& end pos \\ \hline
-100  		& -8.64513 			& ???	& -8.64513 \\
-500		& -8.59563			& ??? 	& -8.59612 \\
-1,000	 	& -8.5870			& ???	& -8.5870 \\ 
-5,000 		& -8.58165			& ???	& -8.58098 \\ 
-10,000 		& -8.58133			& ???	& -8.58015 \\ \hline
+100  		& -8.64513 			& -8.64513	& -8.64513 \\
+500		& -8.59563			& -8.59563 	& -8.59612 \\
+1,000	 	& -8.58700			& -8.58700	& -8.5870 \\ 
+5,000 		& -8.58165			& -8.58165	& -8.58098 \\ 
+10,000 		& -8.58133			& -8.58133	& -8.58015 \\ \hline
 \end{tabular}
@@
 Pop-limit	& Coverage \newline cardinality	& Coverage	& Coverage \newline \& end pos \\ \hline
-100  		& 73 			& ???	& 72 \\
-500		& 237			& ??? 	& 229 \\
+100  		& 73 			& 75	& 72 \\
+500		& 237			& 225 	& 229 \\
 1,000	 	& 416			& ???	& 423 \\ 
 5,000 		& 1,930			& ???	& 2153 \\ 
-10,000 		& 			& ???	& 4576 \\ \hline
+10,000 		& 3,619			& ???	& 4576 \\ \hline
 \end{tabular}

### 2016-02-04 acl.2016/acl2016.tex
acl paper

--- 2016-02-03 acl.2016/acl2016.tex
@@
 \usepackage{algorithmic}
+\usepackage{graphicx}
 
@@
 \begin{table}[h]
-\small
+\tiny
 \begin{center}
@@
 Pop-limit	& Coverage \newline cardinality	& Coverage	& Coverage \newline \& end pos \\ \hline
-100  		& -8.64513 			& ???	& -8.64513 \\
-500		& -8.59563			& ??? 	& -8.59612 \\
-1,000	 	& -8.5870			& ???	& -8.5870 \\ 
-5,000 		& -8.58165			& ???	& -8.58098 \\ 
-10,000 		& -8.58133			& ???	& -8.58015 \\ \hline
+100  		& -8.64513 			& -8.64513	& -8.64513 \\
+500		& -8.59563			& -8.59563 	& -8.59612 \\
+1,000	 	& -8.58700			& -8.58700	& -8.5870 \\ 
+5,000 		& -8.58165			& -8.58165	& -8.58098 \\ 
+10,000 		& -8.58133			& -8.58133	& -8.58015 \\ 
+15,000 		& -8.58130			& -8.58130	& -8.57999 \\ 
+20,000 		& -8.58130			& -8.58130	& -8.57994 \\ \hline
 \end{tabular}
@@
 Pop-limit	& Coverage \newline cardinality	& Coverage	& Coverage \newline \& end pos \\ \hline
-100  		& 73 			& ???	& 72 \\
-500		& 237			& ??? 	& 229 \\
-1,000	 	& 416			& ???	& 423 \\ 
-5,000 		& 1,930			& ???	& 2153 \\ 
-10,000 		& 			& ???	& 4576 \\ \hline
+100  		& 73 			& 75	& 72 \\
+500		& 237			& 225 	& 229 \\
+1,000	 	& 416			& 397	& 423 \\ 
+5,000 		& 1,930			& 1031	& 2153 \\ 
+10,000 		& 3,619			& 3630	& 4576 \\ \hline
 \end{tabular}
@@
 \end{table}
+
+\begin{figure}[h]
+\centering
+\begin{tabular}{cc}
+{\includegraphics[scale=0.4]{stack-configuration.png}} 
+\end{tabular}
+\caption{Analysis and Generation Translation Model}
+\label{fig:analysisAndGen}
+\end{figure}
 

### 2016-02-04 acl.2016/acl2016.tex
acl paper

--- 2016-02-03 acl.2016/acl2016.tex
@@
 \end{enumerate}
-Table~\ref{tab:stack-configuration-model-score} present the average model on the held out test set, while Table~\ref{tab:stack-configuration-time} presents the decoding time.
+Table~\ref{tab:stack-configuration} and Figure~\ref{fig:stack-configuration} present the decoding time and average model on the held out test set.
 \begin{table}[h]
-\small
+\tiny
 \begin{center}
-\begin{tabular}{|r|p{12mm}|p{12mm}|p{13mm}|} \hline
-Pop-limit	& Coverage \newline cardinality	& Coverage	& Coverage \newline \& end pos \\ \hline
-100  		& -8.64513 			& ???	& -8.64513 \\
-500		& -8.59563			& ??? 	& -8.59612 \\
-1,000	 	& -8.5870			& ???	& -8.5870 \\ 
-5,000 		& -8.58165			& ???	& -8.58098 \\ 
-10,000 		& -8.58133			& ???	& -8.58015 \\ \hline
+\begin{tabular}{|r|r|r|r|r|r|r|} \hline
+Pop-limit	& \multicolumn{2}{c|}{Cardinality}	& \multicolumn{2}{c|}{Coverage}	& \multicolumn{2}{c|}{Coverage \& end pos} \\ \hline
+100  		& 73		& -8.64513 		& 75		& -8.64513	& 72	& -8.64513 \\
+500		& 237		& -8.59563		& 225		& -8.59563 	& 229	& -8.59612 \\
+1,000	 	& 416		& -8.58700		& 397		& -8.58700	& 423	& -8.5870 \\ 
+5,000 		& 1930		& -8.58165		& 1931		& -8.58165	& 2153	& -8.58098 \\ 
+10,000 		& 3619		& -8.58133		& 3630		& -8.58133	& 4576	& -8.58015 \\ 
+15,000 		& 4830		& -8.58130		& 5001		& -8.58130	& 7156	& -8.57999 \\ 
+20,000 		& 5849		& -8.58130		& 5916		& -8.58130	& 9583	& -8.57994 \\ \hline
 \end{tabular}
 \end{center}
-\caption{Average model scores for different stack configurations}
-\label{tab:stack-configuration-model-score}
+\caption{Decoding time (in secs with 32threads) and average model scores for different stack configurations}
+\label{tab:stack-configuration}
 \end{table}
 
-\begin{table}[h]
-\small
-\begin{center}
-\begin{tabular}{|r|p{12mm}|p{12mm}|p{13mm}|} \hline
-Pop-limit	& Coverage \newline cardinality	& Coverage	& Coverage \newline \& end pos \\ \hline
-100  		& 73 			& ???	& 72 \\
-500		& 237			& ??? 	& 229 \\
-1,000	 	& 416			& ???	& 423 \\ 
-5,000 		& 1,930			& ???	& 2153 \\ 
-10,000 		& 			& ???	& 4576 \\ \hline
+\begin{figure}[h]
+\centering
+\begin{tabular}{cc}
+{\includegraphics[scale=0.4]{stack-configuration.png}} 
 \end{tabular}
-\end{center}
-\caption{Decoding time with 32-threads for different stack configurations (sec)}
-\label{tab:stack-configuration-time}
-\end{table}
+\caption{Trade-off between decoding time average model scores for different stack configurations}
+\label{fig:stack-configuration}
+\end{figure}
 

### 2016-02-04 acl.2016/acl2016.tex
acl paper

@@
 Pop-limit	& \multicolumn{2}{c|}{Cardinality}	& \multicolumn{2}{c|}{Coverage}	& \multicolumn{2}{c|}{Coverage \& end pos} \\ \hline
+		& Time		& Score			& Time		& Score		& Time		& Score	\\ \hline
 100  		& 73		& -8.64513 		& 75		& -8.64513	& 72	& -8.64513 \\

### 2016-02-04 acl.2016/acl2016.tex
acl paper

@@
 \# threads	& 1 		& 32	& 1 		& 32  \\ \hline
-Memory  	& 24		& 39 	& 11		& 13 \\
-LM 		& 12	 	& 2 	& 47		& 38 \\ 
-Phrase-table	& 9	 	& 5 	& 2		& 4 \\ 
-Lex RO 		& 8	 	& 2 	& 2		& 2 \\ 
-Search 		& 2	 	& 0 	& 14		& 19 \\ 
-Misc/Unknown	& 45	 	& 39 	& 24		& 29 \\ \hline
+Memory  	& 24\%		& 39\% 	& 11\%		& 13\% \\
+LM 		& 12\%	 	& 2\% 	& 47\%		& 38\% \\ 
+Phrase-table	& 9\%	 	& 5\% 	& 2\%		& 4\% \\ 
+Lex RO 		& 8\%	 	& 2\% 	& 2\%		& 2\% \\ 
+Search 		& 2\%	 	& 0\% 	& 14\%		& 19\% \\ 
+Misc/Unknown	& 45\%	 	& 39\% 	& 24\%		& 29\% \\ \hline
 \end{tabular}
@@
 \end{center}
-\caption{Decoding time (in secs with 32threads) and average model scores for different stack configurations}
+\caption{Decoding time (in secs with 32 threads) and average model scores for different stack configurations}
 \label{tab:stack-configuration}

### 2016-02-04 acl.2016/acl2016.tex
acl paper

@@
 The two test scenarios have differing characterics that we are interested in analyzing, ar-en have short sentences with large models while fr-en have overly long sentences with smaller models.
+Where we need to compare model scores, we used held out test sets.
 
@@
 \begin{center}
-\begin{tabular}{|l|c|c|} \hline
+\begin{tabular}{|l|r|r|} \hline
 		& ar-en		& fr-en \\ \hline
+\multicolumn{3}{|l|}{For speed testing} \\ \hline
 \# sentences  	& 800k 		& 200k \\
@@
 Avg words/sent	& 7.3		& 29.7 \\ \hline
+\multicolumn{3}{|l|}{For model score testing} \\ \hline
+Set name	& OpenSubtitles	& newstest2011 \\
+\# sentences  	& 2000 		& 3003 \\
+\# words 	& 14,620 	& 86,162 \\ 
+Avg words/sent	& 7.3		& 28.7 \\ \hline
 \end{tabular}
@@
 \end{table}
-
-Where we need to compare model scores, we used held out test sets; a 2,000 sentence subset of OpenSubtitles for ar-en and the 3,003 sentence newstest2011 test set for fr-en.
 

### 2016-02-04 acl.2016/acl2016.tex
acl paper

@@
 \begin{center}
+\small
 \begin{tabular}{|l|r|r|} \hline
@@
 \multicolumn{3}{|l|}{For speed testing} \\ \hline
+Set name	& \multicolumn{2}{|c|}{Subset of training data} \\
 \# sentences  	& 800k 		& 200k \\
@@
 \begin{center}
+\small
 \begin{tabular}{|l|c|c|c|c|} \hline
@@
 500		& 237		& -8.59563		& 225		& -8.59563 	& 229	& -8.59612 \\
-1,000	 	& 416		& -8.58700		& 397		& -8.58700	& 423	& -8.5870 \\ 
+1,000	 	& 416		& -8.58700		& 397		& -8.58700	& 423	& -8.58700 \\ 
 5,000 		& 1930		& -8.58165		& 1931		& -8.58165	& 2153	& -8.58098 \\ 

### 2016-02-04 acl.2016/acl2016.tex
acl paper

@@
 \end{algorithmic}
-\caption{Hypothesis Expansion with Unsorted Stack}
+\caption{Hypothesis Expansion with Cardinality Stacks}
 \label{algo:Hypothesis Expansion with Unsorted Stack}
 \end{figure}
-However, by separating hypotheses into \emph{'ministacks'} according to coverage and end position, the distortion limit only needs to be checked for each ministack, Figure~\ref{algo:Hypothesis Expansion with Sorted Stack}.
+However, by separating hypotheses into set of hypotheses (\emph{'ministacks'}) according to coverage and end position, the distortion limit only needs to be checked for each ministack, Figure~\ref{algo:Hypothesis Expansion with Sorted Stack}.
+Furthermore, stack pruning is done on each of these hypotheses set therefore, changing how hypotheses are grouped can affect model scores.
 \begin{figure} [h]
@@
 \end{algorithmic}
-\caption{Hypothesis Expansion with Sorted Stack}
+\caption{Hypothesis Expansion with Coverage \& End Position Stacks}
 \label{algo:Hypothesis Expansion with Sorted Stack}
@@
 \end{enumerate}
-Table~\ref{tab:stack-configuration} and Figure~\ref{fig:stack-configuration} present the decoding time and average model on the held out test set.
+Table~\ref{tab:stack-configuration} and Figure~\ref{fig:stack-configuration} present the tradeoff between decoding time and average model at various pop-limits.
 \begin{table}[h]
@@
 \end{figure} 
+As can be seen, the model scores for all stack configurations are identical for low pop-limits parameters but grouping hypotheses into coverage \& end position produces higher model scores for higher pop-limits.
+It is also slower but the time/quality tradeoff is better overall with this stack configuration.
+For lower pop-limits this configuration is slightly slower, but not by much, therefore, we shall stick with this Configuration for the remainder of the paper.
 
-The algorithms in Figure~\ref{algo:Hypothesis Expansion with Unsorted Stack} and ~\ref{algo:Hypothesis Expansion with Sorted Stack} describes the standard phrase-based decoding algorithm, however, the stack configuration issues are the same in the cube-pruning algorithm.
-Furthermore, the cube-pruning algorithm also contain a priority queue which is attached to coverage cardinality in Moses, Figure~\ref{algo:Cube Pruning with Unsorted Stack}.
+The cube-pruning algorithm contain a further priority queue which is attached to hypotheses sets which can be independent of how hypotheses are grouped.
+In the Moses implementation, the priority queue is also attached to the coverage cardinality, Figure~\ref{algo:Cube Pruning with Unsorted Stack}.
 \begin{figure} [h]
@@
 \end{figure}
-We experiment with changing the queue, $queue_{|C|}$, by having separate queues for each (1) unique coverage $queue_{C}$, and each (2) unique coverage and end position $queue_{C, e}$.
-The stack configuration stayed constant.
+Again, we experiment with diffirent queue configurations, having separate queues for each cardinality, coverage, and coverage \& end position.
+The stack configuration remained constant (coverage \& end position with pop-limit of 400).
+\begin{table}[h]
+\small
+\begin{center}
+\begin{tabular}{|r|r|r|} \hline
+Queue configuration		& Time		& Score \\ \hline
+Cardinality			& 192		& -8.59922 \\
+Coverage			&		& -8.58635 \\
+Coverage \& end position	&		& -8.58263 \\ \hline
+\end{tabular}
+\end{center}
+\caption{Decoding time (in secs with 32 threads) and average model scores for different queue configurations}
+\label{tab:queue-configuration}
+\end{table}
+From the results in Table~\ref{tab:queue-configuration}, using finer grain queues results in better model scores, it is significantly slower to decode.
+
+\subsection{Translation Model}
+
+\subsection{Lexicalized Reordering Model}
 

### 2016-02-04 acl.2016/acl2016.tex
acl paper

@@
 
-Over 24\% of the Moses decoder running time is spent on memory management and this increases when more threads are used, Table~\ref{tab:optimizing-memory}, dampening the scalability ofthe decoder.
+Over 24\% of the Moses decoder running time is spent on memory management and this increases when more threads are used, Table~\ref{tab:optimizing-memory}, dampening the scalability of the decoder.
 By contrast, our decoder spends 11\% on memory management and does not significantly increase with more cores.
@@
 Cardinality			& 192		& -8.59922 \\
-Coverage			&		& -8.58635 \\
-Coverage \& end position	&		& -8.58263 \\ \hline
+Coverage			& 2413		& -8.58635 \\
+Coverage \& end position	& 7472		& -8.58263 \\ \hline
 \end{tabular}
@@
 \end{table}
-From the results in Table~\ref{tab:queue-configuration}, using finer grain queues results in better model scores, it is significantly slower to decode.
+From the results in Table~\ref{tab:queue-configuration}, using finer grain queues results in better model scores but it is significantly slower to decode.
 

### 2016-02-04 acl.2016/acl2016.tex
acl paper

@@
 \end{figure}
-We therefore looked at the effects of three stack configurations.
-\begin{enumerate}
-  \item coverage cardinality,
-  \item coverage,
-  \item coverage and end position of most recent translated source word.
-\end{enumerate}
-We therefore looked at the effects of three stack configurations.
+We therefore looked at the effects of three stack configurations:
 \begin{enumerate}
@@
 Table~\ref{tab:stack-configuration} and Figure~\ref{fig:stack-configuration} present the tradeoff between decoding time and average model at various pop-limits.
-\begin{table}[h]
-\tiny
+\begin{table*}[t]
+\small
 \begin{center}
@@
 \label{tab:stack-configuration}
-\end{table}
+\end{table*}
 
@@
   \STATE get best $item$ in $queue_{|C|}$
-  \STATE create new hypo from $item$
-  \STATE add new hypo to new $stack_{|C|}$
+  \STATE create hypo from $item$, coverage $C_{new}$
+  \STATE add hypo to $stack_{|C_{new}|}$
   \STATE create next $items$
@@
 \end{algorithmic}
-\caption{Cube Pruning with Unsorted Stack}
+\caption{Cube Pruning with Cardinality Stacks}
 \label{algo:Cube Pruning with Unsorted Stack}
@@
 \end{table}
-From the results in Table~\ref{tab:queue-configuration}, using finer grain queues results in better model scores but it is significantly slower to decode.
+From the results in Table~\ref{tab:queue-configuration}, using finer grain queues does results in better model scores but it is significantly slower to decode.
 

### 2016-02-05 acl.2016/acl2016.tex
acl paper

@@
 Cardinality			& 192		& -8.59922 \\
-Coverage			& 2413		& -8.58635 \\
-Coverage \& end position	& 7472		& -8.58263 \\ \hline
+Coverage			& 2,413		& -8.58635 \\
+Coverage \& end position	& 7,472		& -8.58263 \\ \hline
 \end{tabular}
@@
 \subsection{Translation Model}
+
+For any realistic sized phrase-based model to be used in an online situation, memory and loading time constraints requires us to use load-on-demand phrase-table implementations.
+Moses contains a number of such implementations with different performance characterics, Figure~\ref{fig:moses-phrase-tables-time}.
+\begin{figure}[h]
+\centering
+\begin{tabular}{cc}
+{\includegraphics[scale=0.4]{moses-phrase-table.png}} 
+\end{tabular}
+\caption{Moses decoding time with different phrase-tables implementations}
+\label{fig:moses-phrase-tables-time}
+\end{figure} 
+We will use the Probing phrase-table (???) exclusively in our work as this appears to have the fastest translation rule lookup, especially with large number of cores.
+
+There are two main areas
 

### 2016-02-05 acl.2016/acl2016.tex
acl paper

@@
 We will use the Probing phrase-table (???) exclusively in our work as this appears to have the fastest translation rule lookup, especially with large number of cores.
+We will look at two main areas for phrase-table optimization specifically in relation to the Probing implementation.
 
-There are two main areas
+Firstly, we will look at the \textbf{caching} of often used translation rules.
+
+Secondly, the Probing phrase-table use a simple \textbf{compression} algorithm to compress the target side of the translation rule. We will look at the affect this has on decoding speed.
 

### 2016-02-06 acl.2016/acl2016.tex
acl paper

@@
 
-The speed of memory access is dependent on whether the data is in the CPU cache which is a constrained resource compared to memory size, typically 20MB in the latest processors.
+%The speed of memory access is dependent on whether the data is in the CPU cache which is a constrained resource compared to memory size, typically 20MB in the latest processors.
 We shall seek to re-use recently accessed information to increase likelihood of the data being in the CPU cache.
@@
 
-Load-on-demand \emph{'binary'} phrase-tables are often used for MT deployment due to their fast loading and querying speed, and because they can be used with large phrase-tables.
-We therefore focus on optimizing decoding speed with these phrase-tables, specifically the Probing PT.
+For any realistic sized phrase-based model to be used in an online situation, memory and loading time constraints requires us to use load-on-demand phrase-table implementations.
+Moses contains a number of such implementations with different performance characterics, we show the time take to decode 800,000 sentences for the fastest two in Figure~\ref{fig:moses-phrase-tables-time}.
+\begin{figure}[h]
+\centering
+\begin{tabular}{cc}
+{\includegraphics[scale=0.4]{moses-phrase-table.png}} 
+\end{tabular}
+\caption{Moses decoding time (in sec) with two different phrase-table implementations}
+\label{fig:moses-phrase-tables-time}
+\end{figure} 
+From this, it appears that the Probing phrase-table (???) has the fastest translation rule lookup, especially with large number of cores, therefore, we will concentrate exclusively on this implementation in our work as this .
+We will look at two main areas for phrase-table optimization specifically in relation to the Probing implementation.
 
-We shall look at the caching strategies to reduce the number of phrase-table lookups. We shall also investigate the datastructures used by the phrase-table and their impact on decoding speed.
+We will focus on two main optimizations that differs from the current Probing phrase-table implementation and that have application to other implementations.
+Firstly, we will look at the \textbf{caching} of often used translation rules.
+The default caching framework save the most recently looked up rules.
+However, this caching mechanism is actually slower than re-querying the phrase-table, Table~\ref{???}.
+We shall explore a simpler caching mechanism that creates the cache at the start of decoding and remain static thereafter.
+
+Secondly, the Probing phrase-table use a simple \textbf{compression} algorithm to compress the target side of the translation rule.
+Compression was championed being the main factor for the speed of the Compact phrase-table but as we saw in Figure~\ref{fig:moses-phrase-tables-time}, it may come at a cost when used with higher number of cores.
+We shall therefore take the opposite approach to (???) and explore the decoding speed gain by disabling compression.
 
@@
 
-For any realistic sized phrase-based model to be used in an online situation, memory and loading time constraints requires us to use load-on-demand phrase-table implementations.
-Moses contains a number of such implementations with different performance characterics, Figure~\ref{fig:moses-phrase-tables-time}.
-\begin{figure}[h]
-\centering
-\begin{tabular}{cc}
-{\includegraphics[scale=0.4]{moses-phrase-table.png}} 
+The Moses translation model caches the most recently queried translation rules for later re-use.
+This has been shown to perform badly for fast phrase-tables such as the Probing phrase-table.
+We explore a simple caching mechanism that populates the cache during loading with rules that translates the most common source phrases.
+\begin{table}[h]
+\small
+\begin{center}
+\begin{tabular}{|r|r|r|} \hline
+Cache size	& Decoding Time & Cache Hit \%age	\\ \hline
+Before caching	& 229	& N/A	 \\
+0		& 239 (+4.4\%)	& 0\% \\
+1,000		& 213 (-7.0\%)	& 11\% \\
+2,000		& 204 (-10.9\%)	& 13\% \\
+4,000		& 205 (-10.5\%)	& 14\% \\
+10,000		& 207 (-9.7\%)	& 17\% \\ \hline
 \end{tabular}
-\caption{Moses decoding time with different phrase-tables implementations}
-\label{fig:moses-phrase-tables-time}
-\end{figure} 
-We will use the Probing phrase-table (???) exclusively in our work as this appears to have the fastest translation rule lookup, especially with large number of cores.
-We will look at two main areas for phrase-table optimization specifically in relation to the Probing implementation.
+\end{center}
+\caption{Decoding time (in secs with 32 threads) for varying cache sizes}
+\label{tab:cache}
+\end{table}
+The static cache does not require the overhead of managing the most recently queried lookups but there is still some overhead in using a cache.
+Overall however, there was over a 10\% decrease in decoding time using the optimum cache size, Table~\ref{tab:cache}.
 
-Firstly, we will look at the \textbf{caching} of often used translation rules.
-
-Secondly, the Probing phrase-table use a simple \textbf{compression} algorithm to compress the target side of the translation rule. We will look at the affect this has on decoding speed.
+In the second optimatization, we disable the compression.
+This increase the size of the binary files from 17GB to 23GB but the time saved not needing to decompress the data resulted in a 1.5\% decrease in decoding time with 1 thread and nearly 7\% when the CPU is saturated, Table~\ref{tab:compression}.
+\begin{table}[h]
+\small
+\begin{center}
+\begin{tabular}{|r|r|r|} \hline
+\# threads	& Compressed pt & Non-compressed pt \\ \hline
+1		& 3052	& 3006 (-1.5\%) \\
+5		& 756	& 644 (-14.8\%) \\
+10		& 372	& 362 (-2.7\%) \\
+15		& 284	& 250 (-12.0\%) \\
+20		& 244	& 227 (-7.0\%) \\
+25		& 218	& 209 (-4.1\%) \\
+30		& 206	& 192 (-6.8\%) \\
+35		& 203	& 189 (-6.9\%) \\ \hline
+\end{tabular}
+\end{center}
+\caption{Decoding time (in secs with 32 threads) for compressed and non-compressed phrase-tables}
+\label{tab:compression}
+\end{table}
 

### 2016-02-06 acl.2016/acl2016.tex
acl paper

@@
 \end{figure} 
-From this, it appears that the Probing phrase-table (???) has the fastest translation rule lookup, especially with large number of cores, therefore, we will concentrate exclusively on this implementation in our work as this .
+From this, it appears that the Probing phrase-table~\cite{Bogochev:Thesis:2013} has the fastest translation rule lookup, especially with large number of cores, therefore, we will concentrate exclusively on this implementation in our work as this .
 We will look at two main areas for phrase-table optimization specifically in relation to the Probing implementation.
@@
 Compression was championed being the main factor for the speed of the Compact phrase-table but as we saw in Figure~\ref{fig:moses-phrase-tables-time}, it may come at a cost when used with higher number of cores.
-We shall therefore take the opposite approach to (???) and explore the decoding speed gain by disabling compression.
+We shall therefore take the opposite approach to~\newcite{junczys_tsd_2012b} and explore the decoding speed gain by disabling compression.
 
@@
 The phrase-table was then pruned, keeping only the top 100 entries per source phrase, according to $p(t|s)$.
-All models files were then binarized; the language models were binaized using KenLM (???), the phrase table using Probing PT (???), lexicalized reordering model using the compact datastructure described in ???.
+All models files were then binarized; the language models were binaized using KenLM~\cite{Heafield-kenlm}, the phrase table using the Probing phrase-table, lexicalized reordering model using the compact datastructure.
 These binary formats were choosen for their best-in-class multithreaded performance.
@@
 
-Standard Moses phrase-based configurations are used, except that we use the cube-pruning algorithm (???) with a pop-limit of 400, rather than the basic phrase-based algorithm.
+Standard Moses phrase-based configurations are used, except that we use the cube-pruning algorithm~\cite{Chiang:2007:cl} with a pop-limit of 400, rather than the basic phrase-based algorithm.
 The cube-pruning algorithm is often employ by users who require fast decoding as it gives them the ability to trade speed with translation quality with a simple pop-limit parameter.
@@
 \subsection{Lexicalized Reordering Model}
+
+The lexicalized reordering model assign a probability to a translation rule, given the relative ordering of the rule in the hypothesis.
+Using the model can significantly increase translation quality at the cost of significantly longer decoding time.
+Efforts such as~\newcite{junczys_tsd_2012b} have been made to increase the speed of the model, with some success, Figure~\ref{???}.
+
+We take a different approach by integrating the lexicalized model file into the translation model, reducing the need to query another model file.
+This resulted in a large decrease in decoding time, Figure~\ref{fig:lex-ro}.
+
+\begin{figure}[h]
+\centering
+\begin{tabular}{cc}
+{\includegraphics[scale=0.4]{lex-ro.png}} 
+\end{tabular}
+\caption{Decoding time (in sec) with Compact Lexicalized Reordering, and integrated into a model the phrase-table}
+\label{fig:lex-ro}
+\end{figure}
 

### 2016-02-07 acl.2016/acl2016.tex
acl paper

@@
 We take a different approach by integrating the lexicalized model file into the translation model, reducing the need to query another model file.
-This resulted in a large decrease in decoding time, Figure~\ref{fig:lex-ro}.
+This resulted in a large decrease in decoding time, especially with high number of cores, Figure~\ref{fig:lex-ro}.
+Integrating the lexicalized reordering model into the translation model decreases decoding time by 29\% with a single core but it is over 5 times faster using all cores.
+In fact, it enables the decoder to continue to scale, making efficient use of all hyperthreads, while using a separate lexicalized reordering model performance flatten out and actually worsens atfer approximately 10 threads.
 

### 2016-02-07 acl.2016/acl2016.tex
acl paper

--- 2016-02-06 acl.2016/acl2016.tex
@@
 
+Figure~\ref{fig:mempool} compares the decoding time for Moses and our decoder, using the same models, parameters and test set.
+Our decoder is over 3 times faster with one thread, and 4.7\% faster using all cores.
+Like Moses, performance actually worsens after approximately 15 cores, however, the problem is not as pronounced.
+This gives us a better foundation on which to build further innovations for fast, multi-core decoding.
+\begin{figure}[h]
+\centering
+\begin{tabular}{cc}
+{\includegraphics[scale=0.4]{mempool.png}} 
+\end{tabular}
+\caption{Decoding time of Moses and our decoder, using the same models}
+\label{fig:mempool}
+\end{figure}
+
 % A potential disadvantage of using memory pool is that the decoder will use more memory, however, Table~\ref{tab:memory-usage} shows that peak memory consumption are comparable.
@@
 We take a different approach by integrating the lexicalized model file into the translation model, reducing the need to query another model file.
-This resulted in a large decrease in decoding time, Figure~\ref{fig:lex-ro}.
-
+This resulted in a large decrease in decoding time, especially with high number of cores, Figure~\ref{fig:lex-ro}.
 \begin{figure}[h]
@@
 \end{figure} 
+Integrating the lexicalized reordering model into the translation model decreases decoding time by 29\% with a single core but it is over 5 times faster using all cores.
+In fact, the decoding time with the integrated model is similar to that \emph{without} a lexilized reordering model.
+Critically for systems with large number of cores, it enables the decoder to continue to scale, making efficient use of all available cores, while using a separate lexicalized reordering model decoding time flatten out and actually worsens after approximately 10 threads.
 

### 2016-02-07 acl.2016/acl2016.tex
acl paper

@@
 \end{tabular}
-\caption{Moses decoding time (in sec) with two different phrase-table implementations}
+\caption{Moses decoding time with two different phrase-table implementations}
 \label{fig:moses-phrase-tables-time}
@@
 The lexicalized reordering model assign a probability to a translation rule, given the relative ordering of the rule in the hypothesis.
-Using the model can significantly increase translation quality at the cost of significantly longer decoding time.
-Efforts such as~\newcite{junczys_tsd_2012b} have been made to increase the speed of the model, with some success, Figure~\ref{???}.
+Using the model can increase translation quality at the cost of longer decoding time.
+Efforts such as~\newcite{junczys_tsd_2012b} have been made to increase the speed of the model, with some success at low core counts, Figure~\ref{fig:moses-lex-ro}.
+\begin{figure}[h]
+\centering
+\begin{tabular}{cc}
+{\includegraphics[scale=0.4]{moses-lex-ro.png}} 
+\end{tabular}
+\caption{Comparing Moses decoding time with different Lexicalized Reordering implementations}
+\label{fig:moses-lex-ro}
+\end{figure}
 
 We take a different approach by integrating the lexicalized model file into the translation model, reducing the need to query another model file.
-This resulted in a large decrease in decoding time, especially with high number of cores, Figure~\ref{fig:lex-ro}.
+This resulted in a significant decrease in decoding time, especially with high number of cores, Figure~\ref{fig:lex-ro}.
 \begin{figure}[h]
@@
 \end{tabular}
-\caption{Decoding time (in sec) with Compact Lexicalized Reordering, and integrated into a model the phrase-table}
+\caption{Decoding time with Compact Lexicalized Reordering, and integrated into a model the phrase-table}
 \label{fig:lex-ro}
@@
 In fact, the decoding time with the integrated model is similar to that \emph{without} a lexilized reordering model.
-Critically for systems with large number of cores, it enables the decoder to continue to scale, making efficient use of all available cores, while using a separate lexicalized reordering model decoding time flatten out and actually worsens after approximately 10 threads.
+Critically for systems with large number of cores, it enables the decoder to continue to scale, making efficient use of all available cores.
+This is in contrast to using a separate lexicalized reordering model where decoding time flatten out and actually worsens after approximately 10 threads.
+
+\subsection{Core Affinity}
+
+The phrase-based decoding algorithm requires fast access to memory in order to 
+Core affinity enables the binding of specific threads to specific CPU cores.
+
+\section{Other Models and Even More Cores}
+
+more cores
+fr-en results
 

### 2016-02-07 acl.2016/acl2016.tex
acl paper

@@
 For high volume users, decoding is a largest and most critical part of the translation process which needs to be fast and efficient.
-However, it has been noticed that the Moses decoder, amongst others, is unable to efficiently use multiple CPU cores that are now common on modern servers (reviewed paper, github discussion).
+However, it has been noticed that the Moses decoder, amongst others, is unable to efficiently use multiple CPU cores that are now common on modern servers~\cite{mfernandez2016boosting}.
 That is, the time taken to decode a test set does not substantial decrease when more cores are used, in fact, decoding time may increase when more cores are added.
@@
 
-(Recently reviewed) describes running multiple processes of the Moses decoder for increased speed.
+~\newcite{mfernandez2016boosting} describes running multiple processes of the Moses decoder to increase speed, treating it as a black box within a parallelization framework.
 
 Other prior work look to optimizing specific components of decoding.
-(Liang and Chiang) describes the cube-pruning and cube-growing algorithm for decoding which allows the tradeoff between speed and translation quality to the adjusted with a simgle parameter.
-(KenLM) and (DALM) describes fast, efficient datastructures for language models.
-(Zen) describes an implementation of a phrase-table for an SMT decoder that is loaded on demand, reducing the initial loading time and memory requirements.
-(CompactPT) extends this by compressing the on-disk phrase table and lexicalized re-ordering model resulting in impressive speed gains over previous work.
+~\newcite{Chiang:2007:cl} describes the cube-pruning and cube-growing algorithm for decoding which allows the tradeoff between speed and translation quality to the adjusted with a simgle parameter.
+~\newcite{Heafield-kenlm} and ~\newcite{tanaka2013efficient} describes fast, efficient datastructures for language models.
+~\newcite{zens2007efficient} describes an implementation of a phrase-table for an SMT decoder that is loaded on demand, reducing the initial loading time and memory requirements.
+~\newcite{junczys_tsd_2012b} extends this by compressing the on-disk phrase table and lexicalized re-ordering model resulting in impressive speed gains over previous work.
 
-(mtplz) is perhaps closest in intent to this work.
+~\newcite{Heafield-mtplz} is perhaps closest in intent to this work.
 This takes a wholistic approach to decoding, describing a novel decoding algorithm which is fis focused on better decoding speed.
@@
 We trained a phrase-based system using the Moses toolkit with standard settings.
-The training data consisted of most of the publicly available Arabic-English data from Opus (Jörg Tiedemann, 2012,) containing over 69 million parallel sentences, and tuned on a held out set.
+The training data consisted of most of the publicly available Arabic-English data from Opus~\cite{tiedemann2012parallel} containing over 69 million parallel sentences, and tuned on a held out set.
 The phrase-table was then pruned, keeping only the top 100 entries per source phrase, according to $p(t|s)$.

### 2016-02-07 acl.2016/acl2016.tex
acl paper

@@
 The lexicalized reordering model is trained on parallel data, usually requiring random lookups of the model file during decoding.
-However, the key to the lookup are the source and target phrase of each translation rule.
-We shall investigate the advantages of storing the mode data within the translation rule.
+Using the model can increase translation quality at the cost of longer decoding time.
+Efforts such as~\newcite{junczys_tsd_2012b} have been made to increase the speed of the model, with some success at low core counts, Figure~\ref{fig:moses-lex-ro}.
+\begin{figure}[h]
+\centering
+\begin{tabular}{cc}
+{\includegraphics[scale=0.4]{moses-lex-ro.png}} 
+\end{tabular}
+\caption{Comparing Moses decoding time with different Lexicalized Reordering implementations}
+\label{fig:moses-lex-ro}
+\end{figure}
+
+However, the querying key to the lookup are the source and target phrase of each translation rule.
+Rather than storing the model separately, we shall investigate integrating the lexicalized reordering model into the translation model.
 
@@
 The lexicalized reordering model assign a probability to a translation rule, given the relative ordering of the rule in the hypothesis.
-Using the model can increase translation quality at the cost of longer decoding time.
-Efforts such as~\newcite{junczys_tsd_2012b} have been made to increase the speed of the model, with some success at low core counts, Figure~\ref{fig:moses-lex-ro}.
-\begin{figure}[h]
-\centering
-\begin{tabular}{cc}
-{\includegraphics[scale=0.4]{moses-lex-ro.png}} 
-\end{tabular}
-\caption{Comparing Moses decoding time with different Lexicalized Reordering implementations}
-\label{fig:moses-lex-ro}
-\end{figure}
 

### 2016-02-08 acl.2016/acl2016.tex
acl paper

@@
 
-\subsection{Core Affinity}
+\subsection{Scalability}
 
-The phrase-based decoding algorithm requires fast access to memory in order to 
-Core affinity enables the binding of specific threads to specific CPU cores.
+\begin{figure}[h]
+\centering
+\begin{tabular}{cc}
+{\includegraphics[scale=0.4]{scalability.png}} 
+\end{tabular}
+\caption{Decoding speed}
+\label{fig:speed}
+\end{figure}
+
+Figure~\ref{fig:speed} shows decoding speed against the number of threads, measured in words translated per second.
+Speed increases linearly with 2 threads as each thread is run on separate physical processors.
+The translation rate only decrease slightly with every additional threads until 16 threads are used which produces a 12.5 times faster than single-threaded decoding.
+The need to share resources such as CPU cache and internal bus bandwidth is probably responsible for the non-linear increase in speed.
+
+Since the test server only has 16 real cores, additional threads must be run on the same physical cores as existing threads.
+This further constrains the resources that each thread has, therefore we see a markedly lower speed increase with each additional thread once there are more than 16 concurrent threads.
+
+Overall though, the scalability is remarkably good, the decoder is able to make full use of all real and virtual cores.
+When all cores are saturated, the decoder is 16 times faster than single-threaded decoding, on this 16 real core server.
+
+This contrast with Moses which scales to approximately 16 threads but then actually become slower if more are used, Figure~\ref{fig:moses-speed}.
+\begin{figure}[h]
+\centering
+\begin{tabular}{cc}
+{\includegraphics[scale=0.4]{moses-scalability.png}} 
+\end{tabular}
+\caption{Moses decoding speed}
+\label{fig:moses-speed}
+\end{figure}
 

### 2016-02-08 acl.2016/acl2016.tex
acl paper

@@
 
-This contrast with Moses which scales to approximately 16 threads but then actually become slower if more are used, Figure~\ref{fig:moses-speed}.
+This contrast with Moses which scales to approximately 16 threads but then actually become slower if more threads are used, Figure~\ref{fig:moses-speed}.
 \begin{figure}[h]
@@
 \section{Other Models and Even More Cores}
+
+We verify the results we the smaller French-English phrase-based model and test set container longer sentences.
+The speed gain is even greater than our main testing scenario, Figure~\ref{fig:fr-en-speed}.
+\begin{figure}[h]
+\centering
+\begin{tabular}{cc}
+{\includegraphics[scale=0.4]{fr-en-speed.png}} 
+\end{tabular}
+\caption{Decoding speed for fr-en model}
+\label{fig:fr-en-speed}
+\end{figure}
 

### 2016-02-08 acl.2016/acl2016.tex
acl paper

@@
 
-\section{BLAH BLAH}
-
-The following instructions are directed to authors of papers submitted to and accepted
-for publication in the ACL 2016 proceedings.  All authors are required
-to adhere to these specifications. Authors are required to provide 
-a Portable Document Format (PDF) version of
-their papers.  The proceedings will be printed on A4 paper.
-Authors from countries where access to word-processing systems is
-limited should contact the publication chairs as soon as possible.
- Grayscale readability of all figures and
-graphics will be encouraged for all accepted papers
-(Section \ref{ssec:accessibility}).
-
-Submitted and camera-ready formatting is similar,
-  however, the submitted paper should have:
-\begin{enumerate} 
-\item Author-identifying information removed
-\item A `ruler' on the left and right margins
-\item Page numbers 
-\item A confidentiality header.  
-\end{enumerate}
-In contrast, the camera-ready {\bf should  not have} a ruler, page numbers, nor a confidentiality header.  By uncommenting {\small\verb|\aclfinalcopy|} at the top of this 
- document, it will compile to produce an example of the camera-ready formatting; by leaving it commented out, the document will be anonymized for initial submission.
-Authors should place this command after the {\small\verb|\usepackage|} declarations when preparing their camera-ready manuscript with the ACL 2016 style.
-
-\section{General Instructions}
-
-Manuscripts must be in two-column format.  Exceptions to the
-two-column format include the title, as well as the 
-authors' names and complete
-addresses (only in the final version, not in the version submitted for review), 
-which must be centered at the top of the first page (see
-the guidelines in Subsection~\ref{ssec:first}), and any full-width
-figures or tables.  Type single-spaced.  Do not number the pages 
-in the camera-ready version.
-Start all pages directly under the top margin.  See the guidelines
-later regarding formatting the first page.
-
-The maximum length of a manuscript is eight (8) pages for the main
-conference, printed single-sided, plus two (2) pages for references
-(see Section~\ref{sec:length} for additional information on the
-maximum number of pages).
-
-By uncommenting {\small\verb|\aclfinalcopy|} at the top of this 
- document, it will compile to produce an example of the camera-ready formatting; by leaving it commented out, the document will be anonymized for initial submission.
-When you first create your submission on softconf, please fill in your submitted paper ID where {\small\verb|***|} appears in the {\small\verb|\def\aclpaperid{***}|} definition at the top.
-
-The review process is double-blind, so do not include any author information (names, addresses) when submitting a paper for review.  
-However, you should maintain space for names and addresses so that they will fit in the final (accepted) version.
-The ACL 2016 \LaTeX\ style will create a titlebox space of 2.5in for you when {\small\verb|\aclfinalcopy|} is commented out.
-
-\subsection{The Ruler}
-The ACL 2016 style defines a printed ruler which should be presented in the
-version submitted for review.  The ruler is provided in order that
-reviewers may comment on particular lines in the paper without
-circumlocution.  If you are preparing a document without the provided
-style files, please arrange for an equivalent ruler to
-appear on the final output pages.  The presence or absence of the ruler
-should not change the appearance of any other content on the page.  The
-camera ready copy should not contain a ruler. (\LaTeX\ users may uncomment
-the {\small\verb|\aclfinalcopy|} command in the document preamble.)
-
-Reviewers:
-note that the ruler measurements do not align well with lines in the paper
---- this turns out to be very difficult to do well when the paper contains
-many figures and equations, and, when done, looks ugly.  Just use fractional
-references (e.g., the first line on this page is at mark $096.5$), although in most cases one would
-expect that the approximate location will be adequate.
-
-\subsection{Electronically-available resources}
-
-ACL provides this description in \LaTeX2e{} ({\small\tt acl2016.tex}) and PDF
-format ({\small\tt acl2016.pdf}), along with the \LaTeX2e{} style file used to
-format it ({\small\tt acl2016.sty}) and an ACL bibliography style ({\small\tt acl2016.bst})
-and example bibliography ({\small\tt acl2016.bib}).
-These files are all available at
-{\small\tt acl2016.org/index.php?article\_id=9}. We
-strongly recommend the use of these style files, which have been
-appropriately tailored for the ACL 2016 proceedings.
-
-\subsection{Format of Electronic Manuscript}
-\label{sect:pdf}
-
-For the production of the electronic manuscript, you must use Adobe's
-Portable Document Format (PDF). This format can be generated from
-postscript files: on Unix systems, you can use {\small\tt ps2pdf} for this
-purpose; under Microsoft Windows, you can use Adobe's Distiller, or
-if you have cygwin installed, you can use {\small\tt dvipdf} or
-{\small\tt ps2pdf}.  Note 
-that some word processing programs generate PDF that may not include
-all the necessary fonts (esp.\ tree diagrams, symbols). When you print
-or create the PDF file, there is usually an option in your printer
-setup to include none, all, or just non-standard fonts.  Please make
-sure that you select the option of including ALL the fonts.  {\em
-  Before sending it, test your {\/\em PDF} by printing it from a
-  computer different from the one where it was created}. Moreover,
-some word processors may generate very large postscript/PDF files,
-where each page is rendered as an image. Such images may reproduce
-poorly.  In this case, try alternative ways to obtain the postscript
-and/or PDF.  One way on some systems is to install a driver for a
-postscript printer, send your document to the printer specifying
-``Output to a file'', then convert the file to PDF.
-
-For reasons of uniformity, Adobe's {\bf Times Roman} font should be
-used. In \LaTeX2e{} this is accomplished by putting
-\small
-\begin{verbatim}
-\usepackage{times}
-\usepackage{latexsym}
-\end{verbatim}
-\normalsize
-in the preamble.
-
-Print-outs of the PDF file on A4 paper should be identical to the
-hardcopy version.  If you cannot meet the above requirements about the
-production of your electronic submission, please contact the
-publication chairs above  as soon as possible.
-
-\subsection{The First Page}
-\label{ssec:first}
-
-Center the title, author name(s) and affiliation(s) across both
-columns (or, in the case of initial submission, space for the names). 
-Do not use footnotes for affiliations.  
-Use the two-column format only when you begin the abstract.
-
-{\bf Title}: Place the title centered at the top of the first page, in
-a 15 point bold font.  (For a complete guide to font sizes and styles, see Table~\ref{font-table}.)
-Long titles should be typed on two lines without
-a blank line intervening. Approximately, put the title at 1in from the
-top of the page, followed by a blank line, then the author name(s),
-and the affiliation(s) on the following line.  Do not use only initials
-for given names (middle initials are allowed). Do not format surnames
-in all capitals (e.g., ``Mitchell,'' not ``MITCHELL'').  The affiliation should
-contain the author's complete address, and if possible, an electronic
-mail address. Leave about 0.75in between the affiliation and the body
-of the first page.
-
-{\bf Abstract}: Type the abstract at the beginning of the first
-column.  The width of the abstract text should be smaller than the
-width of the columns for the text in the body of the paper by about
-0.25in on each side.  Center the word {\bf Abstract} in a 12 point
-bold font above the body of the abstract. The abstract should be a
-concise summary of the general thesis and conclusions of the paper.
-It should be no longer than 200 words.  The abstract text should be in 10 point font.
-
-\begin{table}
-\centering
-\small
-\begin{tabular}{cc}
-\begin{tabular}{|l|l|}
-\hline
-{\bf Command} & {\bf Output}\\\hline
-\verb|{\"a}| & {\"a} \\
-\verb|{\^e}| & {\^e} \\
-\verb|{\`i}| & {\`i} \\ 
-\verb|{\.I}| & {\.I} \\ 
-\verb|{\o}| & {\o} \\
-\verb|{\'u}| & {\'u}  \\ 
-\verb|{\aa}| & {\aa}  \\\hline
-\end{tabular} & 
-\begin{tabular}{|l|l|}
-\hline
-{\bf Command} & {\bf  Output}\\\hline
-\verb|{\c c}| & {\c c} \\ 
-\verb|{\u g}| & {\u g} \\ 
-\verb|{\l}| & {\l} \\ 
-\verb|{\~n}| & {\~n} \\ 
-\verb|{\H o}| & {\H o} \\ 
-\verb|{\v r}| & {\v r} \\ 
-\verb|{\ss}| & {\ss} \\\hline
-\end{tabular}
-\end{tabular}
-\caption{Example commands for accented characters, to be used in, e.g., \BibTeX\ names.}\label{tab:accents}
-\end{table}
-
-{\bf Text}: Begin typing the main body of the text immediately after
-the abstract, observing the two-column format as shown in 
-the present document.  Do not include page numbers in the camera-ready manuscript.
-
-{\bf Indent} when starting a new paragraph. For reasons of uniformity,
-use Adobe's {\bf Times Roman} fonts, with 11 points for text and 
-subsection headings, 12 points for section headings and 15 points for
-the title.  If Times Roman is unavailable, use {\bf Computer Modern
-  Roman} (\LaTeX2e{}'s default; see section \ref{sect:pdf} above).
-Note that the latter is about 10\% less dense than Adobe's Times Roman
-font.
-
-\subsection{Sections}
-
-{\bf Headings}: Type and label section and subsection headings in the
-style shown on the present document.  Use numbered sections (Arabic
-numerals) in order to facilitate cross references. Number subsections
-with the section number and the subsection number separated by a dot,
-in Arabic numerals.
-
-{\bf Citations}: Citations within the text appear
-in parentheses as~\cite{Gusfield:97} or, if the author's name appears in
-the text itself, as Gusfield~\shortcite{Gusfield:97}.  Using the provided \LaTeX\ style, the former is accomplished using
-{\small\verb|\cite|} and the latter with {\small\verb|\shortcite|} or {\small\verb|\newcite|}.
-Collapse multiple citations as in~\cite{Gusfield:97,Aho:72}; this is accomplished with the provided style using commas within the {\small\verb|\cite|} command, e.g., {\small\verb|\cite{Gusfield:97,Aho:72}|}.
-Append lowercase letters to the year in cases of ambiguities.  
-Treat double authors as in~\cite{Aho:72}, but write as 
-in~\cite{Chandra:81} when more than two authors are involved.
-
-\textbf{References}:  We recommend
-including references in a separate~{\small\texttt .bib} file, and include an example file 
-in this release ({\small\tt naalhlt2016.bib}).  Some commands
-for names with accents are provided for convenience in Table \ref{tab:accents}.  
-References stored in the separate~{\small\tt .bib} file are inserted into the document using the following commands:
-
-\small
-\begin{verbatim}
-\bibliography{acl2016}
-\bibliographystyle{acl2016}
-\end{verbatim}
-\normalsize
-
-References should appear under the heading {\bf References} at the 
-end of the document, but before any Appendices, unless the appendices contain references.  
-Arrange the references alphabetically
-by first author, rather than by order of occurrence in the text.  %This behavior is provided by default in the provided \BibTeX\ style ({\small\tt naaclhlt2016.bst}). 
-Provide as complete a reference as possible, using a consistent format,
-such as the one for {\em Computational Linguistics\/} or the one in the 
-{\em Publication Manual of the American 
-Psychological Association\/}~\cite{APA:83}.  Authors' full names rather than initials are preferred.  You may use
-{\bf standard} abbreviations for conferences\footnote{\scriptsize {\tt https://en.wikipedia.org/wiki/ \\ \-\hspace{.75cm} List\_of\_computer\_science\_conference\_acronyms}} and journals\footnote{\tt http://www.abbreviations.com/jas.php}.
-
-{\bf Appendices}: Appendices, if any, directly follow the text and the
-references (but see above).  Letter them in sequence and provide an
-informative title: {\bf Appendix A. Title of Appendix}.
-
-\textbf{Acknowledgment} sections should go as a last (unnumbered) section immediately
-before the references.
-
-\subsection{Footnotes}
-
-{\bf Footnotes}: Put footnotes at the bottom of the page. They may
-be numbered or referred to by asterisks or other
-symbols.\footnote{This is how a footnote should appear.} Footnotes
-should be separated from the text by a line.\footnote{Note the
-line separating the footnotes from the text.}  Footnotes should be in 9 point font.
-
-\subsection{Graphics}
-
-{\bf Illustrations}: Place figures, tables, and photographs in the
-paper near where they are first discussed, rather than at the end, if
-possible.  Wide illustrations may run across both columns and should be placed at
-the top of a page. Color illustrations are discouraged, unless you have verified that 
-they will be understandable when printed in black ink.
-
-\begin{table}
-\small
-\centering
-\begin{tabular}{|l|rl|}
-\hline \bf Type of Text & \bf Font Size & \bf Style \\ \hline
-paper title & 15 pt & bold \\
-author names & 12 pt & bold \\
-author affiliation & 12 pt & \\
-the word ``Abstract'' & 12 pt & bold \\
-section titles & 12 pt & bold \\
-document text & 11 pt  &\\
-abstract text & 10 pt & \\
-captions & 9 pt & \\
-caption label & 9 pt & bold \\
-bibliography & 10 pt & \\
-footnotes & 9 pt & \\
-\hline
-\end{tabular}
-\caption{\label{font-table} Font guide.}
-\end{table}
-
-{\bf Captions}: Provide a caption for every illustration; number each one
-sequentially in the form:  ``{\bf Figure 1:} Figure caption.'', ``{\bf Table 1:} Table caption.''  Type the captions of the figures and 
-tables below the body, using 9 point text.  Table and Figure labels should be bold-faced.
-
-\subsection{Accessibility}
-\label{ssec:accessibility}
-
-In an effort to accommodate the color-blind (as well as those printing
-to paper), grayscale readability for all accepted papers will be
-encouraged.  Color is not forbidden, but authors should ensure that
-tables and figures do not rely solely on color to convey critical
-distinctions.
-Here we give a simple criterion on your colored figures, if your paper has to be printed in black and white, then you must assure that every curves or points in your figures can be still clearly distinguished.
-
-\section{Length of Submission}
-\label{sec:length}
-
-The ACL 2016 main conference accepts submissions of long papers and short papers.
-Long papers may consist of up to eight (8) pages of content, plus unlimited pages for references.
-Upon acceptance, final versions of long papers will be given one additional page (up to 9 pages with unlimited pages for references) so that reviewers' comments can be taken into account.
-Short papers may consist of up to four (4) pages of content, plus unlimited pages for references.
-Upon acceptance, short papers will be given five (5) pages in the proceedings and unlimited pages for references.
-For both long and short papers, all illustrations and appendices must be accommodated within these page limits, observing the formatting instructions given in the present document.
-Papers that do not conform to the specified length and formatting requirements are subject to be rejected without review.
-
-\section{Double-blind review process}
-\label{sec:blind}
-
-As the reviewing will be blind, the paper must not include the
-authors' names and affiliations.  Furthermore, self-references that
-reveal the author's identity, e.g., ``We previously showed (Smith,
-1991) ...'' must be avoided. Instead, use citations such as ``Smith
-previously showed (Smith, 1991) ...'' Papers that do not conform to
-these requirements will be rejected without review. In addition,
-please do not post your submissions on the web until after the
-review process is complete (in special cases this is permitted: see 
-the multiple submission policy below).
-
-We will reject without review any papers that do not follow the
-official style guidelines, anonymity conditions and page limits.
-
-\section{Multiple Submission Policy}
-
-Papers that have been or will be submitted to other meetings or
-publications must indicate this at submission time. Authors of
-papers accepted for presentation at ACL 2016 must notify the
-program chairs by the camera-ready deadline as to whether the paper
-will be presented. All accepted papers must be presented at the
-conference to appear in the proceedings. We will not accept for
-publication or presentation papers that overlap significantly in
-content or results with papers that will be (or have been) published
-elsewhere.
-
-Preprint servers such as arXiv.org and ACL-related workshops that
-do not have published proceedings in the ACL Anthology are not
-considered archival for purposes of submission. Authors must state
-in the online submission form the name of the workshop or preprint
-server and title of the non-archival version.  The submitted version
-should be suitably anonymized and not contain references to the
-prior non-archival version. Reviewers will be told: ``The author(s)
-have notified us that there exists a non-archival previous version
-of this paper with significantly overlapping text. We have approved
-submission under these circumstances, but to preserve the spirit
-of blind review, the current submission does not reference the
-non-archival version.'' Reviewers are free to do what they like with
-this information.
-
-Authors submitting more than one paper to ACL must ensure
-that submissions do not overlap significantly ($>25\%$) with each other
-in content or results. Authors should not submit short and long
-versions of papers with substantial overlap in their original
-contributions.
-
-\section*{Acknowledgments}
-
-Do not number the acknowledgment section.
-This section should not be presented for the submission version.
+% \section*{Acknowledgments}
+% 
+% Do not number the acknowledgment section.
+% This section should not be presented for the submission version.
 

### 2016-02-08 acl.2016/acl2016.tex
acl paper

@@
 
-This contrast with Moses which scales to approximately 16 threads but then actually become slower if more threads are used, Figure~\ref{fig:moses-speed}.
+This contrast with Moses where speed increases to approximately 16 threads but then actually become slower if more threads are used, Figure~\ref{fig:moses-speed}.
 \begin{figure}[h]
@@
 
-We verify the results we the smaller French-English phrase-based model and test set container longer sentences.
+more cores
+
+We verify the results on a smaller French-English phrase-based model and test set container longer sentences.
 The speed gain is even greater than our main testing scenario, Figure~\ref{fig:fr-en-speed}.
@@
 
-more cores
-fr-en results
+\section{Conclusion}
 

### 2016-02-08 acl.2016/acl2016.tex
acl paper

@@
 	  \And
-	Author 2\\
-  	ABC University\\
-  	900 Main Street\\
-  	Ourcity, PQ, Canada A1A 1T2\\
-  {\tt author2@abc.ca}}
+	Nicolay Bogoychev\\
+  	University of Edinburgh\\
+  	10 Crichton Street\\
+  	Edinburgh, Scotland, EH10 9AB\\
+  {\tt s1031254@sms.ed.ac.uk}
+	  \And
+	Lane Schwartz\\
+  	University of Illinois\\
+  {\tt lanes@illinois.edu}
+	  \AND
+	Kenneth Heafield\\
+  	University of Edinburgh\\
+  {\tt kheafiel@inf.ed.ac.uk}
+	  \And
+	Marcin Junczys-Dowmunt\\
+	Faculty of Mathematics and Computer Science \\
+	Adam Mickiewicz University\\
+  {\tt junczys@amu.edu.pl}
+	}
 
@@
 When all cores are saturated, the decoder is 16 times faster than single-threaded decoding, on this 16 real core server.
+It is also 4.5 times faster than Moses with a single-thread and 14.5 faster when all cores are used.
 
@@
 
-more cores
+??? more cores
 
@@
 The speed gain is even greater than our main testing scenario, Figure~\ref{fig:fr-en-speed}.
+Our decoder is 5.4 times faster than Moses with a single-thread and 14.5 faster when all cores are saturated.
 \begin{figure}[h]
@@
 
+We have presented a new decoder that is compatible with Moses.
+By studying the shortcomings of the current implementations, we are able to optimize for speed, especially for multicore operation.
+This has resulted in double digit gains compared to Moses on the same hardware.
+Our implementation is also unaffected by scalability issues that has afflicted Moses after 10-15 threads and continue to scale well for all possible cores we have tested it on.
+
+In future, we shall work on exploring the underlying reasons for the scalability issues in Moses to get a better understanding where potential performance issues can arise.
+We shall also investigate other major components of the decoding algorithm, particulaly the language model which has not been touched in this paper.
+
 % \section*{Acknowledgments}

### 2016-02-08 acl.2016/acl2016.tex
acl paper

@@
 The result is a drop-in replacement for the Moses decoder which is up to fifteen times faster and scales almost linearly with the number of cores.
-Furthermore, the decoder makes less search errors than the current Moses decoder.
 
@@
 ~\newcite{Heafield-mtplz} is perhaps closest in intent to this work.
-This takes a wholistic approach to decoding, describing a novel decoding algorithm which is fis focused on better decoding speed.
+This takes a holistic approach to decoding, describing a novel decoding algorithm which is fis focused on better decoding speed.
 It also describes a number of implementation details for faster decoding.
@@
 The rest of the paper will be broken up into the following sections.
-Next, we will describe the phrase-based model and the major implementation components, with particular emphasis on decoding time shortcomings.
+Next, we will describe the phrase-based model and the major implementation components, with particular emphasis on decoding speed.
 We will then describe modifications to improve decoding speed and present results.
-We conclude in the last section discuss suggested improvements and future work.
+We conclude in the last section and discuss future work.
 
@@
 The first three feature functions frequently trained on data and require the feature to read the model from files.
-The other feature functions do not require model files.
-We shall investigate the first two feature functions for optimization.
+The other feature functions are simple and do not leave much room for optimization.
+The language model consumes a large part of decoding time, and it is for this reason that it has already been heavily optmized.
+We shall therefore investigate the first two feature functions for optimization.
 
@@
 When all cores are saturated, the decoder is 16 times faster than single-threaded decoding, on this 16 real core server.
-It is also 4.5 times faster than Moses with a single-thread and 14.5 faster when all cores are used.
+It is also 4.5 times faster than Moses with a single-thread and 9.6 faster when all cores are used.
 
@@
 
-In future, we shall work on exploring the underlying reasons for the scalability issues in Moses to get a better understanding where potential performance issues can arise.
-We shall also investigate other major components of the decoding algorithm, particulaly the language model which has not been touched in this paper.
+In future, we shall investigate other major components of the decoding algorithm, particulaly the language model which has not been touched in this paper.
+We shall also explore the underlying reasons for the scalability issues in Moses to get a better understanding where potential performance issues can arise.
+This has application to other algorithms beside MT decoding.
 

### 2016-02-08 acl.2016/acl2016.tex
acl paper

@@
 The default caching framework save the most recently looked up rules.
-However, this caching mechanism is actually slower than re-querying the phrase-table, Table~\ref{???}.
+However, this caching mechanism is actually slower than re-querying the phrase-table, Table~\ref{tab:phrase-table-cache}.
+\begin{table}[h]
+\small
+\begin{center}
+\begin{tabular}{|l|c|c|} \hline
+		& No cache	& Caching \\ \hline
+Decoding time  	& 2032 		& 2302 \\ \hline
+\end{tabular}
+\end{center}
+\caption{Decoding time (in sec with 32 threads) when using phrase-table cache}
+\label{tab:phrase-table-cache}
+\end{table}
 We shall explore a simpler caching mechanism that creates the cache at the start of decoding and remain static thereafter.

### 2016-02-08 acl.2016/acl2016.tex
acl paper

@@
 \begin{enumerate}
-  \item log transforms translation model probabilities, $p_{TM}(t|s) $ and $p_{TM}(s|t)$, and word-based translation probabilities $p_w(t|s)$ and $p_w(s|t)$,
-  \item log transforms of the lexicalized re-ordering probabilities,
-  \item log transforms of the target language model probability $p(t)$, 
-  \item a distortion penalty
-  \item a phrase-penalty,
-  \item a word penalty,
-  \item an unknown word penalty.
+  \item \vspace{-2 mm} log transforms translation model probabilities, $p_{TM}(t|s) $ and $p_{TM}(s|t)$, and word-based translation probabilities $p_w(t|s)$ and $p_w(s|t)$,
+  \item \vspace{-2 mm} log transforms of the lexicalized re-ordering probabilities,
+  \item \vspace{-2 mm} log transforms of the target language model probability $p(t)$, 
+  \item \vspace{-2 mm} a distortion penalty
+  \item \vspace{-2 mm} a phrase-penalty,
+  \item \vspace{-2 mm} a word penalty,
+  \item \vspace{-2 mm} an unknown word penalty.
 \end{enumerate}

### 2016-02-08 acl.2016/acl2016.tex
acl paper

@@
 
-For critical objects with high churn such as the hypotheses, thread-specific LIFO queues are used to recycle objects which are no longer used.
-This not only reduces memory wastage but re-uses recent objects which are likely to be in the CPU cache.
+For critical objects with high churn such as hypotheses, thread-specific LIFO queues are used to recycle objects which are no longer used.
+This not only reduces memory wastage but re-uses recent objects which are likely to be in the CPU cache memory.
 

### 2016-02-09 acl.2016/acl2016.tex
acl paper

@@
 
-We present in this paper a re-assessment of the major components of phrase-based decoding and decoder implementation with particular emphasis on speed and scalability to multicore machines.
+In this paper we re-assess the major components of phrase-based decoding and decoder implementation with particular emphasis on speed and scalability on multicore machines.
 The result is a drop-in replacement for the Moses decoder which is up to fifteen times faster and scales almost linearly with the number of cores.
@@
 This paper is the first we know of that seeks to tackle this problem head on.
-We present an phrase-based decoder that is not only significantly faster than the Moses baseline for single-threaded operation, but is able to scale run multiple threads on multicore machines with only a slightly loss in linear speed.
+We present an phrase-based decoder that is not only significantly faster than the Moses baseline for single-threaded operation, but is able to run multiple threads on multicore machines with only a slightly loss in linear speed.
 Model scores and functionality are compatible with Moses to aid comparison and ease of transition for users.

### 2016-02-09 acl.2016/acl2016.tex
acl paper

@@
 Language model 	& 3.1  	& 1.8 \\ 
-Lex-re model	& 2.3	& 637MB \\ \hline
+Lex re. model	& 2.3	& 637MB \\ \hline
 \end{tabular}
@@
 We instantiate two pools for each thread, one which is never reset and another which is reset after the decoding each sentence.
-Objects are created in either pool according to their life cycle.
+Datastructures are created in either pool according to their life cycle.
 
-For critical objects with high churn such as hypotheses, thread-specific LIFO queues are used to recycle objects which are no longer used.
+For critical datastructures with high churn such as hypotheses, thread-specific LIFO queues are used to recycle objects which are no longer used.
 This not only reduces memory wastage but re-uses recent objects which are likely to be in the CPU cache memory.
@@
 \end{tabular}
-\caption{Decoding speed}
+\caption{Our decoder's decoding speed}
 \label{fig:speed}
@@
 \end{tabular}
-\caption{Moses decoding speed}
+\caption{Moses' decoding speed}
 \label{fig:moses-speed}

### 2016-02-09 acl.2016/acl2016.tex
acl paper

@@
 		& No cache	& Caching \\ \hline
-Decoding time  	& 2032 		& 2302 \\ \hline
+Decoding time  	& 2032 		& 2302 (+13\%) \\ \hline
 \end{tabular}
@@
 Secondly, the Probing phrase-table use a simple \textbf{compression} algorithm to compress the target side of the translation rule.
-Compression was championed being the main factor for the speed of the Compact phrase-table but as we saw in Figure~\ref{fig:moses-phrase-tables-time}, it may come at a cost when used with higher number of cores.
-We shall therefore take the opposite approach to~\newcite{junczys_tsd_2012b} and explore the decoding speed gain by disabling compression.
+Compression was championed as the main reason behind the speed of the Compact phrase-table but as we saw in Figure~\ref{fig:moses-phrase-tables-time}, this comes at the cost of scalability to large number of threads.
+We shall therefore take the opposite approach to~\newcite{junczys_tsd_2012b} and explore optimizing decoding speed by disabling compression.
 
@@
 Overall though, the scalability is remarkably good, the decoder is able to make full use of all real and virtual cores.
-When all cores are saturated, the decoder is 16 times faster than single-threaded decoding, on this 16 real core server.
+When all 16 cores are saturated with 2 hyper-thread each, the decoder is 16 times faster than single-threaded decoding.
 It is also 4.5 times faster than Moses with a single-thread and 9.6 faster when all cores are used.
@@
 
-We verify the results on a smaller French-English phrase-based model and test set container longer sentences.
+We verify the results on a smaller French-English phrase-based model and test set containng longer sentences.
 The speed gain is even greater than our main testing scenario, Figure~\ref{fig:fr-en-speed}.

### 2016-02-10 acl.2016/acl2016.tex
acl paper

@@
 \begin{enumerate}
-  \item coverage cardinality,
-  \item coverage,
-  \item coverage and end position of most recent translated source word.
+  \item \vspace{-2 mm} coverage cardinality,
+  \item \vspace{-2 mm} coverage,
+  \item \vspace{-2 mm} coverage and end position of most recent translated source word.
 \end{enumerate}
@@
 
-??? more cores
+We tested with the same model and tested set on bigger servers, our decoder shows no sign of scalability problems, Figure~\ref{fig:more-cores}.
+\begin{figure}[h]
+\centering
+\begin{tabular}{cc}
+{\includegraphics[scale=0.4]{more-cores.png}} 
+\end{tabular}
+\caption{Decoding speed for with bigger servers}
+\label{fig:more-cores}
+\end{figure}
 

### 2016-02-10 acl.2016/acl2016.tex
acl paper

@@
 There are a number of open-source SMT projects, most includes a decoder.
-The most well known is Moses, which supports phrase-based models, hierarchical phrase-based as well as various syntax-based models.
-Joshua also supports hiearchical and syntax models and has recently supported phrase-based models.
+The most well known is Moses~\cite{koehn-EtAl:2007:PosterDemo}, which supports phrase-based models, hierarchical phrase-based as well as various syntax-based models.
+Joshua~\cite{Joshua-Decoder} also supports hiearchical and syntax models and has recently supported phrase-based models.
 Phrasal supports a number of variants of the phrase-based model.
-CDEC supports hierarchical and syntactic models.
+CDEC~\cite{dyer2010cdec} supports hierarchical and syntactic models.
 

### 2016-03-03 acl.2016/acl2016.tex
acl paper
Couple typos

@@
 
-% Author information can be set in various styles:
-% For several authors from the same institution:
-% \author{Author 1 \and ... \and Author n \\
-%         Address line \\ ... \\ Address line}
-% if the names do not fit well on one line use
-%         Author 1 \\ {\bf Author 2} \\ ... \\ {\bf Author n} \\
-% For authors from different institutions:
-% \author{Author 1 \\ Address line \\  ... \\ Address line
-%         \And  ... \And
-%         Author n \\ Address line \\ ... \\ Address line}
-% To start a seperate ``row'' of authors use \AND, as in
-% \author{Author 1 \\ Address line \\  ... \\ Address line
-%         \AND
-%         Author 2 \\ Address line \\ ... \\ Address line \And
-%         Author 3 \\ Address line \\ ... \\ Address line}
-% If the title and author information does not fit in the area allocated,
-% place \setlength\titlebox{<new height>} right after
-% at the top, where <new height> can be something larger than 2.25in
 \author{Hieu Hoang\\
@@
 ~\newcite{Heafield-mtplz} is perhaps closest in intent to this work.
-This takes a holistic approach to decoding, describing a novel decoding algorithm which is fis focused on better decoding speed.
+This takes a holistic approach to decoding, describing a novel decoding algorithm which is focused on better decoding speed.
 It also describes a number of implementation details for faster decoding.
@@
 The phrase-table was then pruned, keeping only the top 100 entries per source phrase, according to $p(t|s)$.
-All models files were then binarized; the language models were binaized using KenLM~\cite{Heafield-kenlm}, the phrase table using the Probing phrase-table, lexicalized reordering model using the compact datastructure.
+All models files were then binarized; the language models were binarized using KenLM~\cite{Heafield-kenlm}, the phrase table using the Probing phrase-table, lexicalized reordering model using the compact datastructure.
 These binary formats were choosen for their best-in-class multithreaded performance.

### 2016-03-03 acl.2016/acl2016.tex
Minor clarity

@@
 Standard Moses phrase-based configurations are used, except that we use the cube-pruning algorithm~\cite{Chiang:2007:cl} with a pop-limit of 400, rather than the basic phrase-based algorithm.
-The cube-pruning algorithm is often employ by users who require fast decoding as it gives them the ability to trade speed with translation quality with a simple pop-limit parameter.
+The cube-pruning algorithm is often employed by users who require fast decoding as it gives them the ability to trade speed with translation quality via a simple pop-limit parameter.
 

### 2016-03-03 acl.2016/acl2016.tex
typo

@@
 However, the decoding algorithm is only able to incorporate one stateful feature function which precludes some of the useful decoding configurations which contains multiple stateful feature functions.
-It does not include a load-on-demand phrase table, therefore, cannot be used in a commercial environment where phrase-table has not be filtered with a know test set for any realistic size phrase-table.
+It does not include a load-on-demand phrase table, therefore, cannot be used in a commercial environment where phrase-table has not be filtered with a known test set for any realistic size phrase-table.
 Neither did this paper analyze the scalability of their work to multicore servers.

### 2016-03-03 acl.2016/acl2016.tex
typo

--- 2016-03-03 acl.2016/acl2016.tex
@@
 However, the decoding algorithm is only able to incorporate one stateful feature function which precludes some of the useful decoding configurations which contains multiple stateful feature functions.
-It does not include a load-on-demand phrase table, therefore, cannot be used in a commercial environment where phrase-table has not be filtered with a know test set for any realistic size phrase-table.
+It does not include a load-on-demand phrase table, therefore, cannot be used in a commercial environment where the phrase-table has not be filtered with a known test set for any realistic size phrase-table.
 Neither did this paper analyze the scalability of their work to multicore servers.

### 2016-03-04 acl.2016/acl2016.tex
acl paper

@@
 Using the model can increase translation quality at the cost of longer decoding time.
-Efforts such as~\newcite{junczys_tsd_2012b} have been made to increase the speed of the model, with some success at low core counts, Figure~\ref{fig:moses-lex-ro}.
-\begin{figure}[h]
-\centering
-\begin{tabular}{cc}
-{\includegraphics[scale=0.4]{moses-lex-ro.png}} 
-\end{tabular}
-\caption{Comparing Moses decoding time with different Lexicalized Reordering implementations}
-\label{fig:moses-lex-ro}
-\end{figure}
+Efforts such as~\newcite{junczys_tsd_2012b} have been made to increase the speed of the model.
 

### 2016-03-10 acl.2016/acl2016.tex
nikolay ref

@@
 \end{figure} 
-From this, it appears that the Probing phrase-table~\cite{Bogochev:Thesis:2013} has the fastest translation rule lookup, especially with large number of cores, therefore, we will concentrate exclusively on this implementation in our work as this .
+From this, it appears that the Probing phrase-table~\cite{Bogoychev:Thesis:2013} has the fastest translation rule lookup, especially with large number of cores, therefore, we will concentrate exclusively on this implementation in our work as this .
 We will look at two main areas for phrase-table optimization specifically in relation to the Probing implementation.

### 2016-03-18 acl.2016/acl2016.tex
acl

@@
 The utilization of statistical machine translation (SMT) has grown enormously over the last decade, many using open-source software developed by the NLP community.
-As commercial utilization has increased, there has been a pressing need that is optimized for their requirements.
-Specifically, faster phrase-based decoding, and more efficient utilization of modern multicore servers.
+As commercial use has increased, there is need that for the software to be optimized for their requirements, in particular, faster phrase-based decoding and more efficient utilization of modern multicore servers.
 
@@
 
-SMT has been one of the outstanding success story from the NLP community in the last decade, progressing from a mostly research discipline to public useability via services such as Google Translate, Microsoft Translator Hub, as well as services and products built around offline products such as Language Weaver and the open-source Moses toolkit.
-The latter has spawned a cottage industry  encompassing a range of organizations and services from small language service providers seeking to reduce translation cost, to large inter-govermental organizations such as the EU and the UN that require high volume, high quality translation.
+SMT has steadily progressed from a research discipline to public useability during the past decade via services such as Google Translate, Microsoft Translator Hub, as well as products and services for customized translation such as those from Language Weaver and the open-source Moses toolkit.
+The latter has spawned a cottage industry ranging from small language service providers seeking to reduce translation cost, to large inter-govermental agencies such as the EU and the UN that require high volume, in-house machine translation.
 
-For high volume users, decoding is a largest and most critical part of the translation process which needs to be fast and efficient.
-However, it has been noticed that the Moses decoder, amongst others, is unable to efficiently use multiple CPU cores that are now common on modern servers~\cite{mfernandez2016boosting}.
-That is, the time taken to decode a test set does not substantial decrease when more cores are used, in fact, decoding time may increase when more cores are added.
-The problem will continue to grow as the commercial use of SMT increases and the number of CPU cores increases.
+For high volume users, decoding is a most time critical part of the translation process.
+A common strategy to increase decoding throughput is to make use of the multiple cores that are now ubiquitous in todays servers.
+However, it has been noticed that the Moses decoder, amongst others, is unable to efficiently use multiple~\cite{mfernandez2016boosting}.
+That is, decoding speed does not substantial increase when more cores are used, in fact, it may actually \emph{decrease} when more cores are added.
 
@@
 We present an phrase-based decoder that is not only significantly faster than the Moses baseline for single-threaded operation, but is able to run multiple threads on multicore machines with only a slightly loss in linear speed.
-Model scores and functionality are compatible with Moses to aid comparison and ease of transition for users.
+Model scores and functionality are compatible with Moses to aid comparison and ease transition for users.
 All source code will be made available under an open-source license.
@@
 The most well known is Moses~\cite{koehn-EtAl:2007:PosterDemo}, which supports phrase-based models, hierarchical phrase-based as well as various syntax-based models.
-Joshua~\cite{Joshua-Decoder} also supports hiearchical and syntax models and has recently supported phrase-based models.
+Joshua~\cite{Joshua-Decoder} also supports hierarchical and syntax models and has recently supported phrase-based models.
 Phrasal supports a number of variants of the phrase-based model.
@@
 A number of the decoders support multithreading whilst others use alternative methods such as Hadoop or external scripts to parallelize decoding.
-We shall investigate the efficiency of using parallelizing decoding using the multi-processor approach.
-None of the decoder focus on multi-threads decoding.
 
-~\newcite{mfernandez2016boosting} describes running multiple processes of the Moses decoder to increase speed, treating it as a black box within a parallelization framework.
-
-Other prior work look to optimizing specific components of decoding.
+Most prior work on increasing decoding speed look to optimizing specific components of decoding.
 ~\newcite{Chiang:2007:cl} describes the cube-pruning and cube-growing algorithm for decoding which allows the tradeoff between speed and translation quality to the adjusted with a simgle parameter.
@@
 ~\newcite{junczys_tsd_2012b} extends this by compressing the on-disk phrase table and lexicalized re-ordering model resulting in impressive speed gains over previous work.
+~\newcite{mfernandez2016boosting} describes running multiple processes of the Moses decoder to increase speed, treating it as a black box within a parallelization framework.
 
-~\newcite{Heafield-mtplz} is perhaps closest in intent to this work.
+~\newcite{Heafield-mtplz} is perhaps closest in intent to our work.
 This takes a holistic approach to decoding, describing a novel decoding algorithm which is focused on better decoding speed.
-It also describes a number of implementation details for faster decoding.
 However, the decoding algorithm is only able to incorporate one stateful feature function which precludes some of the useful decoding configurations which contains multiple stateful feature functions.
-It does not include a load-on-demand phrase table, therefore, cannot be used in a commercial environment where the phrase-table has not be filtered with a known test set for any realistic size phrase-table.
+It does not include a load-on-demand phrase table, therefore, cannot be used in a commercial environment with any realistic size models.
 Neither did this paper analyze the scalability of their work to multicore servers.

### 2016-03-18 acl.2016/acl2016.tex
acl

@@
 That is, for a source sentence $s$, the objective is to find a target translation $\hat{t}$ which has the highest conditional probability $p(t | s)$.
-Mathematically, this is written as:
+Formally, this is written as:
 \begin{equation}
@@
 A translation of a source sentence is created by applying a series of translation rules which together translate each source word once, and only once.
-Each partial translation is called a \emph{hypothesis}, which is created by applying a rule to an existing hypothesis.
-This process is called \emph{hypothesis expansion} and starts with a hypothesis that has translated no source word and ends with a completed hypothesis that has translated all source words.
+Each partial translation is known as a \emph{hypothesis}, which is created by applying a rule to an existing hypothesis.
+This process is called \emph{hypothesis expansion} and starts with a hypothesis that has translated no source word and ends with a completed hypotheses that have translated all source words.
 The highest-scoring completed hypothesis, according to the model score, is returned as  most probable translation, $\hat{t} $.
-Incomplete hypotheses are referred to as partial hypotheses.
 
-Each rule translates a contiguous sequence of source words but successive translation options do not have to be adjacent on the source side, depending on the distortion limit.
-However, the target output is constructed strictly left-to-right from the target string of successive translation options.
-Therefore, successive translation options which are not adjacent and monotonic in the source causes translation reordering.
+In the phrase-based model, each rule translates a contiguous sequence of source words but successive translation options do not have to be adjacent on the source side, depending on the distortion limit.
+The target output is constructed strictly left-to-right from the target side of successive translation rules.
 
 A beam search algorithm is used to create the completed hypothesis set efficiently.
-Partial hypotheses are organized into stacks where each stack holds a number of comparable hypotheses.
-Hypotheses in the same stack have the same coverage cardinality $|C|$, where $C$ is  the coverage set, $C \subseteq \{1,2,...
+Partial translations are organized into stacks where each stack holds a number of comparable hypotheses.
+Most phrase-based implementations place hypotheses in the same stack that have the same coverage cardinality $|C|$, where $C$ is  the coverage set, $C \subseteq \{1,2,...
 |s| \} $ of the number of  source words translated.
-Therefore, $|s| + 1$ number of stacks are created for the decoding of a sentence $s$.
-
-There are three main optimization to the search that we shall investigate.
-Firstly, the search creates and destroy a large number of hypothesis objects in memory which puts a heavy burden on the operating system.
-We shall optimize the search algorithm to use memory pools and object pools, replacing the operating system's general purpose memory management with our own application-aware management.
-
-%The speed of memory access is dependent on whether the data is in the CPU cache which is a constrained resource compared to memory size, typically 20MB in the latest processors.
-We shall seek to re-use recently accessed information to increase likelihood of the data being in the CPU cache.
-
-In multiprocessor servers, the CPU cache is attached to each processor and each core.
-If a sentences is being decoded on one CPU is switched to another, the CPU cache on the new CPU must be repopulated, slowing down decoding.
-We will therefore investigate binding threads to specific cores.
-
-Lastly, we shall investigate different stack configurations other than coverage cardinality to see whether they can improve speed and translation quality.
 
@@
 Rather than storing the model separately, we shall investigate integrating the lexicalized reordering model into the translation model.
+
+\section{Proposed Improvements}
+
+There are four main optimizations that we shall investigate.
+Firstly, the search creates and destroy a large number of hypothesis objects in memory which puts a burden on the operating system.
+We shall optimize the search algorithm to use memory pools and object pools, replacing the operating system's general purpose memory management with our own application-aware management.
+
+%The speed of memory access is dependent on whether the data is in the CPU cache which is a constrained resource compared to memory size, typically 20MB in the latest processors.
+We shall seek to re-use recently accessed information to increase likelihood of the data being in the CPU cache.
+
+%In multiprocessor servers, the CPU cache is attached to each processor and each core.
+If a sentences is being decoded on one CPU is switched to another, the CPU cache on the new CPU must be repopulated, slowing down decoding.
+We will therefore investigate binding threads to specific cores.
+
+Secondly, we shall investigate different stack configurations other than coverage cardinality to see whether they can improve speed and translation quality.
 

### 2016-03-18 acl.2016/acl2016.tex
acl

@@
 The other feature functions are simple and do not leave much room for optimization.
-The language model consumes a large part of decoding time, and it is for this reason that it has already been heavily optmized.
-We shall therefore investigate the first two feature functions for optimization.
+The language model consumes a large part of decoding time, and it is for this reason that it has already been heavily optimized.
+%We shall therefore investigate the first two feature functions for optimization.
 
@@
 \end{figure} 
-From this, it appears that the Probing phrase-table~\cite{Bogoychev:Thesis:2013} has the fastest translation rule lookup, especially with large number of cores, therefore, we will concentrate exclusively on this implementation in our work as this .
-We will look at two main areas for phrase-table optimization specifically in relation to the Probing implementation.
-
-We will focus on two main optimizations that differs from the current Probing phrase-table implementation and that have application to other implementations.
-Firstly, we will look at the \textbf{caching} of often used translation rules.
-The default caching framework save the most recently looked up rules.
-However, this caching mechanism is actually slower than re-querying the phrase-table, Table~\ref{tab:phrase-table-cache}.
-\begin{table}[h]
-\small
-\begin{center}
-\begin{tabular}{|l|c|c|} \hline
-		& No cache	& Caching \\ \hline
-Decoding time  	& 2032 		& 2302 (+13\%) \\ \hline
-\end{tabular}
-\end{center}
-\caption{Decoding time (in sec with 32 threads) when using phrase-table cache}
-\label{tab:phrase-table-cache}
-\end{table}
-We shall explore a simpler caching mechanism that creates the cache at the start of decoding and remain static thereafter.
-
-Secondly, the Probing phrase-table use a simple \textbf{compression} algorithm to compress the target side of the translation rule.
-Compression was championed as the main reason behind the speed of the Compact phrase-table but as we saw in Figure~\ref{fig:moses-phrase-tables-time}, this comes at the cost of scalability to large number of threads.
-We shall therefore take the opposite approach to~\newcite{junczys_tsd_2012b} and explore optimizing decoding speed by disabling compression.
+From this, it appears that the Probing phrase-table~\cite{Bogoychev:Thesis:2013} has the fastest translation rule lookup, especially with large number of cores, therefore, we will concentrate exclusively on this implementation in our work as this.
 
@@
 Efforts such as~\newcite{junczys_tsd_2012b} have been made to increase the speed of the model.
-
-However, the querying key to the lookup are the source and target phrase of each translation rule.
-Rather than storing the model separately, we shall investigate integrating the lexicalized reordering model into the translation model.
 
@@
 
+Thirdly, we will look at for optimizating the phrase-table, specifically in relation to the Probing implementation, focusing on two techniques.
+One, we will investigate caching of often used translation rules.
+The default caching framework save the most recently looked up rules.
+However, this caching mechanism actually makes decoding slower, Table~\ref{tab:phrase-table-cache}.
+\begin{table}[h]
+\small
+\begin{center}
+\begin{tabular}{|l|c|c|} \hline
+		& No cache	& Caching \\ \hline
+Decoding time  	& 2032 		& 2302 (+13\%) \\ \hline
+\end{tabular}
+\end{center}
+\caption{Decoding time (in sec with 32 threads) when using phrase-table cache}
+\label{tab:phrase-table-cache}
+\end{table}
+We shall explore a simpler caching mechanism that creates a static cache at the start of decoding.
+
+Two, the Probing phrase-table use a simple compression algorithm to compress the target side of the translation rule.
+Compression was championed by ~\newcite{junczys_tsd_2012b} as the main reason behind the speed of their phrase-table but as we saw in Figure~\ref{fig:moses-phrase-tables-time}, this comes at the cost of scalability to large number of threads.
+We shall therefore take the opposite approach to and explore optimizing decoding speed by disabling compression.
+
+Lastly, the lexicalized reordering model require random lookups but query key are the source and target phrase of each translation rule.
+Rather than storing this model separately, we shall investigate integrating it into the translation model.
+
 \section{Experimental Setup}
@@
 Table~\ref{tab:model-files} gives details of the resultant sizes of the model files.
-For verification with a different dataset, we also occasionally used a second system trained on the French-English Europarl corpus (2m parallel sentences).
+For testing decoding speed, we used a subset of the training data, Table~\ref{tab:test-sets}.
 \begin{table}[h]
@@
 \end{table}
-For testing decoding speed, we used a subset of the training data, Table~\ref{tab:test-sets}.
-The two test scenarios have differing characterics that we are interested in analyzing, ar-en have short sentences with large models while fr-en have overly long sentences with smaller models.
+
+For verification with a different dataset, we also used a second system trained on the French-English Europarl corpus (2m parallel sentences).
+The two different systems have characterics that we are interested in analyzing; ar-en have short sentences with large models while fr-en have overly long sentences with smaller models.
 Where we need to compare model scores, we used held out test sets.
@@
 The server has 380GB RAM.
-The operating system was Ubuntu 14.04, the code was compiled with gcc 4.8.4 and Boost library 1.59.
+The operating system was Ubuntu 14.04, the code was compiled with gcc 4.8.4 and Boost 1.59\footnote{http://boost.org/} and the tcmalloc library\footnote{http://goog-perftools.sourceforge.net/doc/tcmalloc.html}.
 
@@
 We create a dynamic memory pool which can grow as more memory is requested.
-The memory is not released, instead the pool can be reset in order for the memory to be re-used.
-We instantiate two pools for each thread, one which is never reset and another which is reset after the decoding each sentence.
+The memory is not released, instead the pools are reset and re-used.
+We instantiate two pools for each thread, one which is never reset and another which is reset after the decoding of each sentence.
 Datastructures are created in either pool according to their life cycle.
@@
 
-Over 24\% of the Moses decoder running time is spent on memory management and this increases when more threads are used, Table~\ref{tab:optimizing-memory}, dampening the scalability of the decoder.
-By contrast, our decoder spends 11\% on memory management and does not significantly increase with more cores.
+Over 24\% of the Moses decoder running time is spent on memory management and this increases to 39\% when 32 threads are used, Table~\ref{tab:optimizing-memory}, dampening the scalability of the decoder.
+By contrast, our decoder spends 11\% on memory management and does not significantly increase with more threads.
 \begin{table}[h]
@@
 Figure~\ref{fig:mempool} compares the decoding time for Moses and our decoder, using the same models, parameters and test set.
-Our decoder is over 3 times faster with one thread, and 4.7\% faster using all cores.
-Like Moses, performance actually worsens after approximately 15 cores, however, the problem is not as pronounced.
+Our decoder is over 3 times faster with one thread, and 4.7 times faster using all cores.
+Like Moses, performance actually worsens after approximately 15 threads, however, the problem is not as pronounced.
 This gives us a better foundation on which to build further innovations for fast, multi-core decoding.
@@
 
-The most popular stack configuration for phrase-based model, as implemented in Pharaoh, Moses and Joshua, has been by coverage cardinality, ie.
+The most popular stack configuration for phrase-based model, as implemented in Pharaoh, Moses and Joshua, is coverage cardinality, ie.
 hypotheses that have translated the same number of source words are stored in the same stack.
@@
 
-We also note that distortion limit which constrains hypothesis extension is dependent on the hypothesis' coverage vector, $C$ \emph{and} the end position of most recent source word that has been translated, $e$.
+We also note that the distortion limit which constrains hypothesis extension is dependent on the hypothesis' coverage vector, $C$ \emph{and} the end position of most recent source word that has been translated, $e$.
 The distortion limit must be checked for every instance of a hypothesis and translation rule, Figure~\ref{algo:Hypothesis Expansion with Unsorted Stack}.
@@
 It is also slower but the time/quality tradeoff is better overall with this stack configuration.
-For lower pop-limits this configuration is slightly slower, but not by much, therefore, we shall stick with this Configuration for the remainder of the paper.
+For lower pop-limits this configuration is slightly slower but not by much.
+%, therefore, we shall stick with this Configuration for the remainder of the paper.
 
-The cube-pruning algorithm contain a further priority queue which is attached to hypotheses sets which can be independent of how hypotheses are grouped.
+%Also, the cube-pruning algorithm contain a priority queue per set of hypotheses sets.
+This can be independent of how hypotheses are grouped.
 In the Moses implementation, the priority queue is also attached to the coverage cardinality, Figure~\ref{algo:Cube Pruning with Unsorted Stack}.
-\begin{figure} [h]
-\begin{algorithmic}
-%\REQUIRE source sentence $s$, translation options
-\STATE initialize $queue_{|C|}$
-\FOR{1 to pop-limit}
-  \STATE get best $item$ in $queue_{|C|}$
-  \STATE create hypo from $item$, coverage $C_{new}$
-  \STATE add hypo to $stack_{|C_{new}|}$
-  \STATE create next $items$
-  \STATE add new $items$ to $queue_{|C|}$
-\ENDFOR % { hypothesis in $stack_i$} 
-\end{algorithmic}
-\caption{Cube Pruning with Cardinality Stacks}
-\label{algo:Cube Pruning with Unsorted Stack}
-\end{figure}
-Again, we experiment with diffirent queue configurations, having separate queues for each cardinality, coverage, and coverage \& end position.
+%\begin{figure} [h]
+%\begin{algorithmic}
+%%\REQUIRE source sentence $s$, translation options
+%\STATE initialize $queue_{|C|}$
+%\FOR{1 to pop-limit}
+%  \STATE get best $item$ in $queue_{|C|}$
+%  \STATE create hypo from $item$, coverage $C_{new}$
+%  \STATE add hypo to $stack_{|C_{new}|}$
+%  \STATE create next $items$
+%  \STATE add new $items$ to $queue_{|C|}$
+%\ENDFOR % { hypothesis in $stack_i$} 
+%\end{algorithmic}
+%\caption{Cube Pruning with Cardinality Stacks}
+%\label{algo:Cube Pruning with Unsorted Stack}
+%\end{figure}
+%Again, we experiment with diffirent queue configurations, having separate queues for each cardinality, coverage, and coverage \& end position.
 The stack configuration remained constant (coverage \& end position with pop-limit of 400).
-\begin{table}[h]
-\small
-\begin{center}
-\begin{tabular}{|r|r|r|} \hline
-Queue configuration		& Time		& Score \\ \hline
-Cardinality			& 192		& -8.59922 \\
-Coverage			& 2,413		& -8.58635 \\
-Coverage \& end position	& 7,472		& -8.58263 \\ \hline
-\end{tabular}
-\end{center}
-\caption{Decoding time (in secs with 32 threads) and average model scores for different queue configurations}
-\label{tab:queue-configuration}
-\end{table}
-From the results in Table~\ref{tab:queue-configuration}, using finer grain queues does results in better model scores but it is significantly slower to decode.
+%\begin{table}[h]
+%\small
+%\begin{center}
+%\begin{tabular}{|r|r|r|} \hline
+%Queue configuration		& Time		& Score \\ \hline
+%Cardinality			& 192		& -8.59922 \\
+%Coverage			& 2,413		& -8.58635 \\
+%Coverage \& end position	& 7,472		& -8.58263 \\ \hline
+%\end{tabular}
+%\end{center}
+%\caption{Decoding time (in secs with 32 threads) and average model scores for different queue configurations}
+%\label{tab:queue-configuration}
+%\end{table}
+%From the results in Table~\ref{tab:queue-configuration}, using finer grain queues does results in better model scores but it is significantly slower to decode.
+%
 
@@
 The Moses translation model caches the most recently queried translation rules for later re-use.
-This has been shown to perform badly for fast phrase-tables such as the Probing phrase-table.
+This has been shown to perform badly for fast phrase-tables such as those described by~\cite{Bogoychev:Thesis:2013}.
 We explore a simple caching mechanism that populates the cache during loading with rules that translates the most common source phrases.
@@
 In the second optimatization, we disable the compression.
-This increase the size of the binary files from 17GB to 23GB but the time saved not needing to decompress the data resulted in a 1.5\% decrease in decoding time with 1 thread and nearly 7\% when the CPU is saturated, Table~\ref{tab:compression}.
+This increase the size of the binary files from 17GB to 23GB but the time saved not needing to decompress the data resulted in a 1.5\% decrease in decoding time with 1 thread and nearly 7\% when the CPUs are saturated, Table~\ref{tab:compression}.
 \begin{table}[h]
@@
 
-We take a different approach by integrating the lexicalized model file into the translation model, reducing the need to query another model file.
+We integrate the lexicalized model file into the translation model but storing the model's probabilities in the phrase-table.
 This resulted in a significant decrease in decoding time, especially with high number of cores, Figure~\ref{fig:lex-ro}.
@@
 In fact, the decoding time with the integrated model is similar to that \emph{without} a lexilized reordering model.
-Critically for systems with large number of cores, it enables the decoder to continue to scale, making efficient use of all available cores.
-This is in contrast to using a separate lexicalized reordering model where decoding time flatten out and actually worsens after approximately 10 threads.
+Critically for systems with multicores servers, it enables the decoder to continue to scale, making efficient use of all available cores.
+This is in contrast to using a separate lexicalized reordering model where decoding time flatten out and actually worsens after approximately 15 threads.
 
@@
 Figure~\ref{fig:speed} shows decoding speed against the number of threads, measured in words translated per second.
-Speed increases linearly with 2 threads as each thread is run on separate physical processors.
-The translation rate only decrease slightly with every additional threads until 16 threads are used which produces a 12.5 times faster than single-threaded decoding.
-The need to share resources such as CPU cache and internal bus bandwidth is probably responsible for the non-linear increase in speed.
+There is a constant increase in decoding speed when mrore threads are used, only decreasing slightly after 16 threads when hyperthreads are used.
+Overall, decoding is 12.5 times faster than single-threaded decoding when all 16 cores (32 hyperthreads) are fully utilized.
 
-Since the test server only has 16 real cores, additional threads must be run on the same physical cores as existing threads.
+%Since the test server only has 16 real cores, additional threads must be run on the same physical cores as existing threads.
 This further constrains the resources that each thread has, therefore we see a markedly lower speed increase with each additional thread once there are more than 16 concurrent threads.
@@
 
-This contrast with Moses where speed increases to approximately 16 threads but then actually become slower if more threads are used, Figure~\ref{fig:moses-speed}.
+This contrast with Moses where speed increases to approximately 16 threads but then actually become slower thereafter, Figure~\ref{fig:moses-speed}.
 \begin{figure}[h]
@@
 
-We tested with the same model and tested set on bigger servers, our decoder shows no sign of scalability problems, Figure~\ref{fig:more-cores}.
+Our decoder show no scalability issues when we tested with the same model and tested set on a larger server, Figure~\ref{fig:more-cores}.
 \begin{figure}[h]
@@
 
-We verify the results on a smaller French-English phrase-based model and test set containng longer sentences.
-The speed gain is even greater than our main testing scenario, Figure~\ref{fig:fr-en-speed}.
+We verify the results with the French-English phrase-based system and test set.
+The speed gains are even greater than the Arabic-English test scenario, Figure~\ref{fig:fr-en-speed}.
 Our decoder is 5.4 times faster than Moses with a single-thread and 14.5 faster when all cores are saturated.
@@
 We have presented a new decoder that is compatible with Moses.
-By studying the shortcomings of the current implementations, we are able to optimize for speed, especially for multicore operation.
-This has resulted in double digit gains compared to Moses on the same hardware.
-Our implementation is also unaffected by scalability issues that has afflicted Moses after 10-15 threads and continue to scale well for all possible cores we have tested it on.
+By studying the shortcomings of the current implementation, we are able to optimize for speed, particularly for multicore operation.
+This resulted in double digit gains compared to Moses on the same hardware.
+Our implementation is also unaffected by scalability issues that has afflicted Moses.
+% and continue to scale well for all possible cores we have tested it on.
 
-In future, we shall investigate other major components of the decoding algorithm, particulaly the language model which has not been touched in this paper.
+In future, we shall investigate other major components of the decoding algorithm, particularly the language model which has not been touched in this paper.
 We shall also explore the underlying reasons for the scalability issues in Moses to get a better understanding where potential performance issues can arise.

### 2016-05-02 acl.2016/response.txt
changes after review

Thank you for reviewing our paper, the comments are very useful for us in correcting our paper. I'll try and respond to most of questions.

1. We will fix the numerous typos and difficult-to-read sentences all reviewers have pointed out.

2. We will add references to (Zens & Ney, 2008) and (Wuebker et al, 2012), and look at references to the Jane decoder.

3. Strangely, using lexicalised reordering model results in slightly faster decoding than without the model. We double checked with other datasets. We suspect this may be an artifact of the algorithm or datastructures, for example, more work is needed to resolve hypothesis recombination. We will endeavour to find the cause of this issue.

4. Two of the coverage curves in Figure 5 exactly overlap, we will make that clear in the paper.

5. All models in all experiments are read on-demand. The text phrase-table is 8.1GB zipped (40GB unzipped), the Lex RO is 3.5GB zipped (20GB unzipped), the LM is 3.1GB zipped (5.3GB unzipped). Furthermore, the decoder often requires multiple times the file sizes in RAM to hold the datastructures in memory.

For example, trying to load the phrase-table into memory took 67 minutes to load and decode and used over 200GB of virtual memory. Trying to load the phrase-table and re-ordering model into memory exhausted our 340GB server.

This compares to 3-4 minutes decoding time and approximately 30GB of virtual memory when on-demand models are used.

Therefore, even though it may be possible to load some models into RAM on our specific server, it is suboptimal and not advisable for most scenarios due to memory and/or time constraints. We will add these remarks to the paper.

6. We cache the 'most common source phrase' according to the count statistics obtained from the training data. We will make that clearer in the paper.

7. We will add a more detailed description of memory pools.

8. We will add a more detailed explanation of the different stack configurations and discussions. We also have further results which will include as it will be appropriate if the paper contains more discussion on the subject.

9. The speedup in our work was such that we could not accurately evaluate decoding speeds when measured on typical test sets which contains 1000-2000 sentences. Since decoding speed is not affected by whether it is decoding the training data, or held out data, we decoded the training data for convenience. We will test this assumption and add a remark in the paper. Where reporting model scores was necessary, we did, of course, used a held-out test set.

### 2016-05-02 acl.2016/acl2016.tex
changes after review

@@
+% Jane - FastSCSS single-best output, but is considerably faster by not maintaining separate model costs and by deleting recombined hypotheses
+% Jane - not multithreaded
+% Zens & Ney - don't compare different coverage config. Not focused on speed.
 %
@@
 
-In this paper we re-assess the major components of phrase-based decoding and decoder implementation with particular emphasis on speed and scalability on multicore machines.
-The result is a drop-in replacement for the Moses decoder which is up to fifteen times faster and scales almost linearly with the number of cores.
+In this paper we re-examining the major components of phrase-based decoding and decoder implementation with particular emphasis on speed and scalability on multicore machines.
+The result is a drop-in replacement for the Moses decoder which is up to fifteen times faster and scales monotonically with the number of cores.
 
@@
 
-SMT has steadily progressed from a research discipline to public useability during the past decade via services such as Google Translate, Microsoft Translator Hub, as well as products and services for customized translation such as those from Language Weaver and the open-source Moses toolkit.
-The latter has spawned a cottage industry ranging from small language service providers seeking to reduce translation cost, to large inter-govermental agencies such as the EU and the UN that require high volume, in-house machine translation.
+SMT has steadily progressed from a research discipline to commercial viability during the past decade as can be seen from services such as the Google and Microsoft Translation services.
+As well as general purpose services such as these, there is a large number of companies that offer customized translation systems, as well as  companies and organization that implement in-house solutions.
+Many of these customized solutions use Moses as their SMT engine.
 
-For high volume users, decoding is a most time critical part of the translation process.
-A common strategy to increase decoding throughput is to make use of the multiple cores that are now ubiquitous in todays servers.
-However, it has been noticed that the Moses decoder, amongst others, is unable to efficiently use multiple~\cite{mfernandez2016boosting}.
-That is, decoding speed does not substantial increase when more cores are used, in fact, it may actually \emph{decrease} when more cores are added.
+For many users, decoding is the most time-critical part of the translation process.
+Making use of the multiple cores that are now ubiquitous in todays servers is a common strategy to ameliorate this issue.
+However, it has been noticed that the Moses decoder, amongst others, is unable to efficiently use multiple cores~\cite{mfernandez2016boosting}.
+That is, decoding speed does not substantially increase when more cores are used, in fact, it may actually \emph{decrease} when using more cores.
+There have be speculation on the causes of the inefficiency as well as potential remedies.
 
-There have be speculation on the causes of the inefficiency as well as potential remedies.
-This paper is the first we know of that seeks to tackle this problem head on.
-We present an phrase-based decoder that is not only significantly faster than the Moses baseline for single-threaded operation, but is able to run multiple threads on multicore machines with only a slightly loss in linear speed.
-Model scores and functionality are compatible with Moses to aid comparison and ease transition for users.
+This paper is the first we know of that focuses on improving decoding speed on multicore servers.
+We will take a holistic approach to solving this issue, re-implementing the decoder's structure to optimize for speed, re-assessing some of the optimization assumptions made in the implementation of significant parts of the decoder, as well as exploring changes to the core search algorithm.
+We will compare and contrast decoder implementations, not just decoding algorithms or individual components.
+
+We will present a phrase-based decoder that is significantly faster than the Moses baseline for single-threaded operation, and scales monotonically with the number of cores.
+
+As far as possible, model scores and functionality are compatible with Moses to aid comparison and ease transition for existing users.
 All source code will be made available under an open-source license.
@@
 
-There are a number of open-source SMT projects, most includes a decoder.
-The most well known is Moses~\cite{koehn-EtAl:2007:PosterDemo}, which supports phrase-based models, hierarchical phrase-based as well as various syntax-based models.
+There are a number of phrase-based decoding implementations, we shall detail some of the open-source implementations.
+
+The most well known is Moses~\cite{koehn-EtAl:2007:PosterDemo} which supports phrase-based models, hierarchical phrase-based as well as various syntax-based models.
+It is widely used for MT research and commercial use.
+
 Joshua~\cite{Joshua-Decoder} also supports hierarchical and syntax models and has recently supported phrase-based models.
-Phrasal supports a number of variants of the phrase-based model.
-CDEC~\cite{dyer2010cdec} supports hierarchical and syntactic models.
+Phrasal~\cite{spence2014phrasal} supports a number of variants of the phrase-based model.
+Jane~\cite{peitz2012jane} supports hierarchical and standard phrase-based models.
+mtplz~\newcite{Heafield-mtplz} implements an speed-optimized variant of the phrase-based search algorithm.
 
-A number of the decoders support multithreading whilst others use alternative methods such as Hadoop or external scripts to parallelize decoding.
+The Moses, Joshua and Phrasal decoders implement multithreading, while Jane uses an external application to parallelize operation.
+All the multi-threaded implementations report scalability problems, either in the paper (Phrasal) or via social media (Moses\footnote{https://github.com/moses-smt/mosesdecoder/issues/39} and Joshua\footnote{https://twitter.com/ApacheJoshua/status/342022794097340416}).
 

### 2016-05-03 acl.2016/acl2016.tex
add reviews

@@
 
-The Moses, Joshua and Phrasal decoders implement multithreading, while Jane uses an external application to parallelize operation.
-All the multi-threaded implementations report scalability problems, either in the paper (Phrasal) or via social media (Moses\footnote{https://github.com/moses-smt/mosesdecoder/issues/39} and Joshua\footnote{https://twitter.com/ApacheJoshua/status/342022794097340416}).
+The Moses, Joshua and Phrasal decoders implement multithreading, however, they all report scalability problems, either in the paper (Phrasal) or via social media (Moses\footnote{https://github.com/moses-smt/mosesdecoder/issues/39} and Joshua\footnote{https://twitter.com/ApacheJoshua/status/342022794097340416}).
+
+Jane and mtplz are single-threaded decoders, relying on external applications to parallelize operations.
+This has the advantage it is easier to parallelize processing across multiple servers, not just multiple cores.
+However, within one server, this method does not scale beyond a small number of cores, as we shall see.
 

### 2016-05-03 acl.2016/acl2016.tex
changes after review

@@
 
-There are a number of phrase-based decoding implementations, we shall detail some of the open-source implementations.
+Most prior work on increasing decoding speed look to optimizing specific components of the decoder or the decoding algorithm.
 
-The most well known is Moses~\cite{koehn-EtAl:2007:PosterDemo} which supports phrase-based models, hierarchical phrase-based as well as various syntax-based models.
-It is widely used for MT research and commercial use.
+~\newcite{Heafield-kenlm} and ~\newcite{yasuhara-EtAl:2013:EMNLP} describes fast, efficient datastructures for language models.
+~\newcite{zens2007efficient} describes an implementation of a phrase-table for an SMT decoder that is loaded on demand, reducing the initial loading time and memory requirements.
+~\newcite{junczys_tsd_2012b} extends this by compressing the on-disk phrase table and lexicalized re-ordering model.
 
-Joshua~\cite{Joshua-Decoder} also supports hierarchical and syntax models and has recently supported phrase-based models.
+~\newcite{Chiang:2007:cl} describes the cube-pruning and cube-growing algorithm which allows the tradeoff between speed and translation quality to the adjusted with a single parameter.
+~\newcite{wuebker2012fast} note that language model querying is amongst the most expensive operation in decoding.
+They sought to improved decoding speed by caching score computations early pruning of translation options.
+This work is similar to ~\newcite{Heafield-mtplz} which group hypotheses with identical language model context and incrementally expand them, reducing LM querying.
+
+~\newcite{mfernandez2016boosting} was concerned with multi-core speed but treat the decoding process as a black box within a parallelization framework.
+
+There are a number of phrase-based decoding implementations, many of which implements the phrase-based extensions above.
+
+The most well known is Moses~\cite{koehn-EtAl:2007:PosterDemo} which implements a number of speed optimzations, including cube-pruning. It is widely used for MT research and commercial use.
+
+Joshua~\cite{Joshua-Decoder} also supports cube-pruning for phrase-based models.
 Phrasal~\cite{spence2014phrasal} supports a number of variants of the phrase-based model.
-Jane~\cite{peitz2012jane} supports hierarchical and standard phrase-based models.
-mtplz~\newcite{Heafield-mtplz} implements an speed-optimized variant of the phrase-based search algorithm.
+Jane~\cite{peitz2012jane} supports the language model look-ahead described in ~\newcite{wuebker2012fast} in addition to other tools to speed up decoding such as having a separate fast, lightweight decoder.
+mtplz is a specialized decoder developed to implement the incremental decoding described in ~\newcite{Heafield-mtplz}.
 
@@
 This has the advantage it is easier to parallelize processing across multiple servers, not just multiple cores.
-However, within one server, this method does not scale beyond a small number of cores, as we shall see.
+As we shall see, this method does not scale beyond a small number of cores when attempting to scale to multiple cores within one server.
 
-Most prior work on increasing decoding speed look to optimizing specific components of decoding.
-~\newcite{Chiang:2007:cl} describes the cube-pruning and cube-growing algorithm for decoding which allows the tradeoff between speed and translation quality to the adjusted with a simgle parameter.
-~\newcite{Heafield-kenlm} and ~\newcite{tanaka2013efficient} describes fast, efficient datastructures for language models.
-~\newcite{zens2007efficient} describes an implementation of a phrase-table for an SMT decoder that is loaded on demand, reducing the initial loading time and memory requirements.
-~\newcite{junczys_tsd_2012b} extends this by compressing the on-disk phrase table and lexicalized re-ordering model resulting in impressive speed gains over previous work.
-~\newcite{mfernandez2016boosting} describes running multiple processes of the Moses decoder to increase speed, treating it as a black box within a parallelization framework.
-
-~\newcite{Heafield-mtplz} is perhaps closest in intent to our work.
-This takes a holistic approach to decoding, describing a novel decoding algorithm which is focused on better decoding speed.
-However, the decoding algorithm is only able to incorporate one stateful feature function which precludes some of the useful decoding configurations which contains multiple stateful feature functions.
-It does not include a load-on-demand phrase table, therefore, cannot be used in a commercial environment with any realistic size models.
-Neither did this paper analyze the scalability of their work to multicore servers.
+This paper focuses on overcoming the shortcomings of existing decoding implementations on multicore servers.
 

### 2016-05-03 acl.2016/acl2016.tex
changes

@@
 
-The most well known is Moses~\cite{koehn-EtAl:2007:PosterDemo} which implements a number of speed optimzations, including cube-pruning. It is widely used for MT research and commercial use.
+The most well known is Moses~\cite{koehn-EtAl:2007:PosterDemo} which implements a number of speed optimizations, including cube-pruning. It is widely used for MT research and commercial use.
 
@@
 
-The Moses, Joshua and Phrasal decoders implement multithreading, however, they all report scalability problems, either in the paper (Phrasal) or via social media (Moses\footnote{https://github.com/moses-smt/mosesdecoder/issues/39} and Joshua\footnote{https://twitter.com/ApacheJoshua/status/342022794097340416}).
+The Moses, Joshua and Phrasal decoders implement multithreading, however, they all report scalability problems, either in the paper (Phrasal) or via social media (Moses\footnote{\scriptsize https://github.com/moses-smt/mosesdecoder/issues/39} and Joshua\footnote{\scriptsize https://twitter.com/ApacheJoshua/status/342022794097340416}).
 
@@
 Each partial translation is known as a \emph{hypothesis}, which is created by applying a rule to an existing hypothesis.
-This process is called \emph{hypothesis expansion} and starts with a hypothesis that has translated no source word and ends with a completed hypotheses that have translated all source words.
-The highest-scoring completed hypothesis, according to the model score, is returned as  most probable translation, $\hat{t} $.
+This \emph{hypothesis expansion} process starts with a hypothesis that has translated no source word and ends with completed hypotheses that have translated all source words.
+The highest-scoring completed hypothesis, according to the model score, is considered the best translation, $\hat{t} $.
 
-In the phrase-based model, each rule translates a contiguous sequence of source words but successive translation options do not have to be adjacent on the source side, depending on the distortion limit.
-The target output is constructed strictly left-to-right from the target side of successive translation rules.
+In the phrase-based model, each rule translates a contiguous sequence of source words.
+Successive applications of translation rules do not have to be adjacent on the source side, depending on the distortion limit.
+The target output is constructed strictly left-to-right from the target side from the series of translation rules.
 
@@
 
-The first three feature functions frequently trained on data and require the feature to read the model from files.
-The other feature functions are simple and do not leave much room for optimization.
-The language model consumes a large part of decoding time, and it is for this reason that it has already been heavily optimized.
-%We shall therefore investigate the first two feature functions for optimization.
+Of these feature functions, the last four are simple and do not leave much room for optimization.
+As have already been mentioned, the language model consumes a large part of decoding time but it has already been heavily optimized in previous works such as ~\newcite{Heafield-kenlm} and ~\newcite{yasuhara-EtAl:2013:EMNLP}.
+
+Therefore, we will focus on optimizing the functionality of the phrase-table and lexicalized re-ordering model.
 
@@
 
-For any realistic sized phrase-based model to be used in an online situation, memory and loading time constraints requires us to use load-on-demand phrase-table implementations.
-Moses contains a number of such implementations with different performance characterics, we show the time take to decode 800,000 sentences for the fastest two in Figure~\ref{fig:moses-phrase-tables-time}.
+For any phrase-table table of a realistic size, memory and loading time constraints requires us to use load-on-demand implementations of which Moses can make use of, each with differing performance characteristics.
+Figure~\ref{fig:moses-phrase-tables-time} shows the time taken to decode 800,000 sentences for the fastest two implementations.
 \begin{figure}[h]
@@
 There are four main optimizations that we shall investigate.
-Firstly, the search creates and destroy a large number of hypothesis objects in memory which puts a burden on the operating system.
-We shall optimize the search algorithm to use memory pools and object pools, replacing the operating system's general purpose memory management with our own application-aware management.
+
+\subsection{Efficient Memory Allocation}
+
+Firstly, the search algorithm creates and destroy a large number of hypothesis and related objects.
+The memory allocation required puts a burden on the operating system due to the need to synchronize  memory access.
+This problem worsens with multi-threaded applications such as Moses.
+
+Libraries such as tcmalloc~\cite{ghemawat2009tcmalloc} are designed to reduce locking contention for multi-threaded application but in our case, this is still not enough.
+
+We shall seek to improve decoding speed by replacing the operating system's general purpose memory management with our own custom memory management scheme.
+In time-critical parts of the decoder, memory will be allocated from a memory pool.
+
+A memory pool is a large block of memory that has been pre-allocated by the operating system from which smaller blocks can be give out to the application.
+We will use a thread-specific memory pools to avoid locking contention during memory access.
+Our memory pool will be dynamic, growing in size when required but never reducing in size.
+
+To further increase memory management speed, objects are not deleted, instead the memory pool is simply reset and reused for each input sentence.
+This can result in high memory usage so object queues are available for high-churn objects which allows the decoder to re-use portion of the memory pool before it is reset.
+The object queues are LIFO so that recently accessed memory are used, increasing CPU cache hits.
 
@@
 
+\subsection{Stack Configurations}
+
 Secondly, we shall investigate different stack configurations other than coverage cardinality to see whether they can improve speed and translation quality.
 
+\subsection{Phrase-Table Optimizations}
+
 Thirdly, we will look at for optimizating the phrase-table, specifically in relation to the Probing implementation, focusing on two techniques.
-One, we will investigate caching of often used translation rules.
-The default caching framework save the most recently looked up rules.
-However, this caching mechanism actually makes decoding slower, Table~\ref{tab:phrase-table-cache}.
+
+One, the default translation rule caching mechanism in Moses saves the most recently used rules.
+However, this is a complex mechanism which require locking when timestamping datastructures and periodic clearing of old rules in order in order to reduce memory consumption.
+It actually makes decoding slower, Table~\ref{tab:phrase-table-cache}.
 \begin{table}[h]
@@
 \end{table}
-We shall explore a simpler caching mechanism that creates a static cache at the start of decoding.
+
+We shall explore a simpler caching mechanism that creates a static cache at the start of decoding of the most likely translation rules to be used.
 
@@
 We shall therefore take the opposite approach to and explore optimizing decoding speed by disabling compression.
+
+\subsection{Lexicalized Reordering Model Optimizations}
 

### 2016-05-03 acl.2016/acl2016.tex
changes

@@
 That is, decoding speed does not substantially increase when more cores are used, in fact, it may actually \emph{decrease} when using more cores.
-There have be speculation on the causes of the inefficiency as well as potential remedies.
+There have been speculation on the causes of the inefficiency as well as potential remedies.
 
@@
 
-There are a number of phrase-based decoding implementations, many of which implements the phrase-based extensions above.
+There are a number of phrase-based decoding implementations, many of which implements the extensions to the phrase-based model described above.
 
@@
 
-This paper focuses on overcoming the shortcomings of existing decoding implementations on multicore servers.
+This paper not only focuses faster single-threaded decoding but also on overcoming the shortcomings of existing decoding implementations on multicore servers.
 
@@
 \begin{enumerate}
-  \item \vspace{-2 mm} log transforms translation model probabilities, $p_{TM}(t|s) $ and $p_{TM}(s|t)$, and word-based translation probabilities $p_w(t|s)$ and $p_w(s|t)$,
-  \item \vspace{-2 mm} log transforms of the lexicalized re-ordering probabilities,
-  \item \vspace{-2 mm} log transforms of the target language model probability $p(t)$, 
   \item \vspace{-2 mm} a distortion penalty
@@
   \item \vspace{-2 mm} an unknown word penalty.
+  \item \vspace{-2 mm} log transforms of the target language model probability $p(t)$, 
+  \item \vspace{-2 mm} log transforms translation model probabilities, $p_{TM}(t|s) $ and $p_{TM}(s|t)$, and word-based translation probabilities $p_w(t|s)$ and $p_w(s|t)$,
+  \item \vspace{-2 mm} log transforms of the lexicalized re-ordering probabilities,
 \end{enumerate}
@@
 
-Of these feature functions, the last four are simple and do not leave much room for optimization.
+Of these feature functions, the first four are simple and do not leave much room for optimization.
 As have already been mentioned, the language model consumes a large part of decoding time but it has already been heavily optimized in previous works such as ~\newcite{Heafield-kenlm} and ~\newcite{yasuhara-EtAl:2013:EMNLP}.
 
-Therefore, we will focus on optimizing the functionality of the phrase-table and lexicalized re-ordering model.
-
-%\subsection{Language Model}
-% TODO
-%
-\subsection{Translation Model}
-
-For any phrase-table table of a realistic size, memory and loading time constraints requires us to use load-on-demand implementations of which Moses can make use of, each with differing performance characteristics.
-Figure~\ref{fig:moses-phrase-tables-time} shows the time taken to decode 800,000 sentences for the fastest two implementations.
-\begin{figure}[h]
-\centering
-\begin{tabular}{cc}
-{\includegraphics[scale=0.4]{moses-phrase-table.png}} 
-\end{tabular}
-\caption{Moses decoding time with two different phrase-table implementations}
-\label{fig:moses-phrase-tables-time}
-\end{figure} 
-From this, it appears that the Probing phrase-table~\cite{Bogoychev:Thesis:2013} has the fastest translation rule lookup, especially with large number of cores, therefore, we will concentrate exclusively on this implementation in our work as this.
-
-\subsection{Lexicalized Reordering Model}
-
-The lexicalized reordering model is trained on parallel data, usually requiring random lookups of the model file during decoding.
-Using the model can increase translation quality at the cost of longer decoding time.
-Efforts such as~\newcite{junczys_tsd_2012b} have been made to increase the speed of the model.
+Therefore, we will focus on optimizing only the phrase-table and lexicalized reordering model.
 
@@
 
-There are four main optimizations that we shall investigate.
+Apart from the two feature function optimization detailed in the last section, we will investigate two other optimizations, all described below.
 
@@
 
-Thirdly, we will look at for optimizating the phrase-table, specifically in relation to the Probing implementation, focusing on two techniques.
+Thirdly, we will look at optimizing the phrase-table.
 
+For any phrase-table table of a realistic size, memory and loading time constraints requires us to use load-on-demand implementations of which Moses can make use of, each with differing performance characteristics.
+Figure~\ref{fig:moses-phrase-tables-time} shows the time taken to decode 800,000 sentences for the fastest two implementations.
+\begin{figure}[h]
+\centering
+\begin{tabular}{cc}
+{\includegraphics[scale=0.4]{moses-phrase-table.png}} 
+\end{tabular}
+\caption{Moses decoding time with two different phrase-table implementations}
+\label{fig:moses-phrase-tables-time}
+\end{figure} 
+From this, it appears that the Probing phrase-table~\cite{Bogoychev:Thesis:2013} has the fastest translation rule lookup, especially with large number of cores, therefore, we will concentrate exclusively on this implementation in our work as this.
+
+We will focus on two techniques.
 One, the default translation rule caching mechanism in Moses saves the most recently used rules.
-However, this is a complex mechanism which require locking when timestamping datastructures and periodic clearing of old rules in order in order to reduce memory consumption.
-It actually makes decoding slower, Table~\ref{tab:phrase-table-cache}.
+However, this require locking and periodic clearing of old rules in order to reduce memory consumption resulting in \emph{slower} decoding, Table~\ref{tab:phrase-table-cache}.
 \begin{table}[h]
@@
 
-Lastly, the lexicalized reordering model require random lookups but query key are the source and target phrase of each translation rule.
-Rather than storing this model separately, we shall investigate integrating it into the translation model.
+Similar to the phrase-table, the lexicalized reordering model is trained on parallel data, a model file is created during training which is then queried during decoding.
+Inevitably, the querying results in some loss of decoding speed.
+Efforts such as~\newcite{junczys_tsd_2012b} have been made to increase the speed of the query.
+
+However, the model's query key are the source and target phrase of each translation rule.
+Rather than storing the lexicalized reordering model separately, we shall integrating it into the translation model, eliminating the need to query a separate file.
 
@@
 The phrase-table was then pruned, keeping only the top 100 entries per source phrase, according to $p(t|s)$.
-All models files were then binarized; the language models were binarized using KenLM~\cite{Heafield-kenlm}, the phrase table using the Probing phrase-table, lexicalized reordering model using the compact datastructure.
-These binary formats were choosen for their best-in-class multithreaded performance.
+All model files were then binarized; the language models were binarized using KenLM~\cite{Heafield-kenlm}, the phrase table using the Probing phrase-table, lexicalized reordering model using the compact datastructure.
+These binary formats were chosen for their best-in-class multithreaded performance.
 Table~\ref{tab:model-files} gives details of the resultant sizes of the model files.
@@
 
-In the second optimatization, we disable the compression.
+In the second optimization, we disable the compression.
 This increase the size of the binary files from 17GB to 23GB but the time saved not needing to decompress the data resulted in a 1.5\% decrease in decoding time with 1 thread and nearly 7\% when the CPUs are saturated, Table~\ref{tab:compression}.
@@
 Figure~\ref{fig:speed} shows decoding speed against the number of threads, measured in words translated per second.
-There is a constant increase in decoding speed when mrore threads are used, only decreasing slightly after 16 threads when hyperthreads are used.
+There is a constant increase in decoding speed when more threads are used, only decreasing slightly after 16 threads when hyperthreads are used.
 Overall, decoding is 12.5 times faster than single-threaded decoding when all 16 cores (32 hyperthreads) are fully utilized.

### 2016-05-03 acl.2016/acl2016.tex
changes after review

@@
 
-Similar to the phrase-table, the lexicalized reordering model is trained on parallel data, a model file is created during training which is then queried during decoding.
-Inevitably, the querying results in some loss of decoding speed.
-Efforts such as~\newcite{junczys_tsd_2012b} have been made to increase the speed of the query.
+Similar to the phrase-table, the lexicalized reordering model is trained on parallel data where a model file is created which is then queried during decoding.
+Inevitably, the querying results in loss of decoding speed.
+Efforts such as~\newcite{junczys_tsd_2012b} have been made to increase querying speed.
 
@@
 Rather than storing the lexicalized reordering model separately, we shall integrating it into the translation model, eliminating the need to query a separate file.
+This has precedent in ~\newcite{peitz2012jane} but the effect on decoding speed were not published.
+We will compare results with using a separate model in this paper.
 
@@
 We instantiate two pools for each thread, one which is never reset and another which is reset after the decoding of each sentence.
-Datastructures are created in either pool according to their life cycle.
+Data structures are created in either pool according to their life cycle.
 
-For critical datastructures with high churn such as hypotheses, thread-specific LIFO queues are used to recycle objects which are no longer used.
+For critical data structures with high churn such as hypotheses, thread-specific LIFO queues are used to recycle objects which are no longer used.
 This not only reduces memory wastage but re-uses recent objects which are likely to be in the CPU cache memory.

### 2016-05-04 acl.2016/acl2016.tex
changes

@@
 \end{figure} 
-From this, it appears that the Probing phrase-table~\cite{Bogoychev:Thesis:2013} has the fastest translation rule lookup, especially with large number of cores, therefore, we will concentrate exclusively on this implementation in our work as this.
+From this, it appears that the Probing phrase-table~\cite{Bogoychev:Thesis:2013} has the fastest translation rule lookup, especially with large number of cores, therefore, we will concentrate exclusively on this implementation from hereon.
 
-We will focus on two techniques.
+We propose two optimizations.
 One, the default translation rule caching mechanism in Moses saves the most recently used rules.
@@
 
-Similar to the phrase-table, the lexicalized reordering model is trained on parallel data where a model file is created which is then queried during decoding.
+Similar to the phrase-table, the lexicalized reordering model is trained on parallel data.
+A resultant model file is created in training which is then queried during decoding.
 Inevitably, the querying results in loss of decoding speed.
-Efforts such as~\newcite{junczys_tsd_2012b} have been made to increase querying speed.
+Previous work such as~\newcite{junczys_tsd_2012b} improve querying speed with more efficient data structures.
 
-However, the model's query key are the source and target phrase of each translation rule.
+However, the model's query keys are the source and target phrase of each translation rule.
 Rather than storing the lexicalized reordering model separately, we shall integrating it into the translation model, eliminating the need to query a separate file.
@@
 The phrase-table was then pruned, keeping only the top 100 entries per source phrase, according to $p(t|s)$.
-All model files were then binarized; the language models were binarized using KenLM~\cite{Heafield-kenlm}, the phrase table using the Probing phrase-table, lexicalized reordering model using the compact datastructure.
+All model files were then binarized; the language models were binarized using KenLM~\cite{Heafield-kenlm}, the phrase table using the Probing phrase-table, lexicalized reordering model using the compact data structure.
 These binary formats were chosen for their best-in-class multithreaded performance.

### 2016-05-04 acl.2016/acl2016.tex
changes after review

@@
 
-Firstly, the search algorithm creates and destroy a large number of hypothesis and related objects.
+The search algorithm creates and destroy a large number of hypothesis and related objects.
 The memory allocation required puts a burden on the operating system due to the need to synchronize  memory access.
@@
 
-Secondly, we shall investigate different stack configurations other than coverage cardinality to see whether they can improve speed and translation quality.
+The most popular stack configuration for phrase-based model, as implemented in Pharaoh, Moses and Joshua, is coverage cardinality, that is, hypotheses that have translated the same number of source words are stored in the same stack.
+Since stack pruning occurs in each stack independently of each other, this affect the decoder's search errors.
+
+However, there are alternatives to the coverage cardinality configuration.
+~\newcite{Och:2001a} uses a single stack for all hypotheses, ~\newcite{Brown:1993} uses coverage stacks (ie.
+one stack per unique coverage vector) while ~\newcite{peitz2012jane} and ~\newcite{Zens+Ney:2008:iwslt} apply both coverage and cardinality pruning.
+While useful, these prior works present only one particular stack configuration each.
+
+~\newcite{ortizmartinez-garciavarea-casacuberta:2006:WMT} explore a range of stack configurations by defining a granularity parameter which controls the maximum number of stacks required to decode a sentence.
+
+We shall re-visit the question of stack configuration with a particular emphasis on how they can help improve the tradeoff between speed and translation quality.
+We will do so in the context of the cube-pruning algorithm, the algorithm that we will be using and which was not available to many of the earlier work.
 
 \subsection{Phrase-Table Optimizations}
-
-Thirdly, we will look at optimizing the phrase-table.
 
@@
 We propose two optimizations.
-One, the default translation rule caching mechanism in Moses saves the most recently used rules.
+Firstly, the default translation rule caching mechanism in Moses saves the most recently used rules.
 However, this require locking and periodic clearing of old rules in order to reduce memory consumption resulting in \emph{slower} decoding, Table~\ref{tab:phrase-table-cache}.
@@
 
-Two, the Probing phrase-table use a simple compression algorithm to compress the target side of the translation rule.
+Secondly, the Probing phrase-table use a simple compression algorithm to compress the target side of the translation rule.
 Compression was championed by ~\newcite{junczys_tsd_2012b} as the main reason behind the speed of their phrase-table but as we saw in Figure~\ref{fig:moses-phrase-tables-time}, this comes at the cost of scalability to large number of threads.
@@
 
-As a baseline, we use the latest version of the Moses decoder taken from the github repository.
+As a baseline, we use a recent\footnote{between January and May 2016} version of the Moses decoder taken from the github repository.
 
@@
 The server has 380GB RAM.
-The operating system was Ubuntu 14.04, the code was compiled with gcc 4.8.4 and Boost 1.59\footnote{http://boost.org/} and the tcmalloc library\footnote{http://goog-perftools.sourceforge.net/doc/tcmalloc.html}.
+The operating system was Ubuntu 14.04, the code was compiled with gcc 4.8.4 and Boost 1.59\footnote{http://boost.org/} and the tcmalloc library.
 
@@
 
-We create a dynamic memory pool which can grow as more memory is requested.
-The memory is not released, instead the pools are reset and re-used.
-We instantiate two pools for each thread, one which is never reset and another which is reset after the decoding of each sentence.
+We instantiate two memory pools per decoding thread, one which is never reset and another which is reset after the decoding of each sentence.
 Data structures are created in either pool according to their life cycle.
 
-For critical data structures with high churn such as hypotheses, thread-specific LIFO queues are used to recycle objects which are no longer used.
-This not only reduces memory wastage but re-uses recent objects which are likely to be in the CPU cache memory.
+Hypothesis objects and the data structures it contains are re-used by maintaining a thread-specific LIFO queue of unused, recombined and pruned objects from which new hypotheses are allocated in preference to create new hypotheses in the memory pool.
 
-Over 24\% of the Moses decoder running time is spent on memory management and this increases to 39\% when 32 threads are used, Table~\ref{tab:optimizing-memory}, dampening the scalability of the decoder.
+Each sentence is decoded from start to finish using one worker thread. The CPU affinity of each thread is set to a specific core to minimize memory page faults as a result of threads switching cores.
+
+As can be seen in Table~\ref{tab:optimizing-memory}, over 24\% of the Moses decoder running time is spent on memory management and this increases to 39\% when 32 threads are used, dampening the scalability of the decoder.
 By contrast, our decoder spends 11\% on memory management and does not significantly increase with more threads.
@@
 \subsection{Stack Configuration}
-
-The most popular stack configuration for phrase-based model, as implemented in Pharaoh, Moses and Joshua, is coverage cardinality, ie.
-hypotheses that have translated the same number of source words are stored in the same stack.
-There have been research into other stack layouts such as~\cite{ortizmartinez-garciavarea-casacuberta:2006:WMT}, and it has been noted that the decoder in~\cite{Brown:1993} uses coverage stacks, as opposed to coverage cardinality.
 

### 2016-05-04 acl.2016/acl2016.tex
changes

@@
 
-We also note that the distortion limit which constrains hypothesis extension is dependent on the hypothesis' coverage vector, $C$ \emph{and} the end position of most recent source word that has been translated, $e$.
-The distortion limit must be checked for every instance of a hypothesis and translation rule, Figure~\ref{algo:Hypothesis Expansion with Unsorted Stack}.
+We shall investigate the effects of the following three stack configurations on model score and decoding speed:
+\begin{enumerate}
+  \item \vspace{-2 mm} coverage cardinality,
+  \item \vspace{-2 mm} coverage,
+  \item \vspace{-2 mm} coverage and end position of most recently translated source word.
+\end{enumerate}
+Coverage cardinality is the same as that in Moses and Joshua. Coverage configuration uses one stack per unique coverage vector.
+
+\emph{Coverage and end position of most recently translated source word} extends the coverage configuration by separating hypotheses where the last translate word are different.
+This is an optimization to reduce the number of checks on the distortion limit, which is dependent on the last word position.
+
+The check is a binary function $d(C_{h}, e_{hypo}, range_r)$, where $C_{h}$ is the coverage vector of hypothesis $h$, $e_{h}$ is the end position of most recent source word that has been translated, and $range_r$ is the coverage of the rule to be applied.
+Figure~\ref{algo:Hypothesis Expansion with Unsorted Stack} shows the standard hypothesis expansion algorithm.
 \begin{figure} [h]
@@
 %\REQUIRE source sentence $s$, translation options
-\FORALL{ $hypo$ in $stack_{|C|}$} 
-  \FORALL{ translation rules}
-      \IF {can-expand($C(hypo)$, $e(hypo)$, translation rule range)}
-	\STATE expand hypo with translation rule $\rightarrow$ new hypo
-	\STATE add new hypo to next stack
+\FORALL{ $h$ in $stack_{|C|}$} 
+  \FORALL{$r$ in translation rules}
+      \IF {$d(C_{h}, e_{hypo}, range_r)$}
+	\STATE expand $h$ with $r$ $\rightarrow$ $h'$
+	\STATE add $h'$ to stack $stack_{C'}$
       \ENDIF
@@
 \end{figure}
-However, by separating hypotheses into set of hypotheses (\emph{'ministacks'}) according to coverage and end position, the distortion limit only needs to be checked for each ministack, Figure~\ref{algo:Hypothesis Expansion with Sorted Stack}.
-Furthermore, stack pruning is done on each of these hypotheses set therefore, changing how hypotheses are grouped can affect model scores.
+
+By grouping hypotheses according to coverage \emph{and} end position into groups we call \emph{'ministack'}, the distortion limit only needs to be checked for each group, Figure~\ref{algo:Hypothesis Expansion with Sorted Stack}.
 \begin{figure} [h]
@@
 \FORALL{ $ministack_{C, e}$ in $stack_{|C|}$} 
-  \FORALL{ translation rules}
-    \IF {can-expand($C$, $e$, translation rule range)}
-      \FORALL{ $hypo$ in $ministack_{C, e}$} 
-	\STATE expand hypo with translation rule $\rightarrow$ new hypo
-	\STATE add new hypo to next ministack
+  \FORALL{$r$ in translation rules}
+    \IF {$d(C_{h}, e_{hypo}, range_r)$}
+      \FORALL{ $h$ in $ministack_{C, e}$} 
+	\STATE expand $h$ with $r$ $\rightarrow$ $h'$
+	\STATE add $h'$ to $ministack_{C',e'}$
       \ENDFOR
@@
 \end{figure}
-We therefore looked at the effects of three stack configurations:
-\begin{enumerate}
-  \item \vspace{-2 mm} coverage cardinality,
-  \item \vspace{-2 mm} coverage,
-  \item \vspace{-2 mm} coverage and end position of most recent translated source word.
-\end{enumerate}
+
+Also, stack pruning occurs on each hypothesis group independently, therefore, potentially affect search errors and model scores.
+
 Table~\ref{tab:stack-configuration} and Figure~\ref{fig:stack-configuration} present the tradeoff between decoding time and average model at various pop-limits.
+The model scores for all stack configurations are identical for low pop-limits but grouping hypotheses into coverage \& end position produces higher model scores for higher pop-limits.
+It is also slower but the time/quality tradeoff is better overall with this stack configuration.
+Model scores for coverage and cardinality configurations are identical, which differs from the results in ~\newcite{ortizmartinez-garciavarea-casacuberta:2006:WMT}.
+This may be due to differing test set characteristics, or interactions with cube-pruning algorithm we used.
 \begin{table*}[t]
@@
 \end{figure} 
-As can be seen, the model scores for all stack configurations are identical for low pop-limits parameters but grouping hypotheses into coverage \& end position produces higher model scores for higher pop-limits.
-It is also slower but the time/quality tradeoff is better overall with this stack configuration.
-For lower pop-limits this configuration is slightly slower but not by much.
  %, therefore, we shall stick with this Configuration for the remainder of the paper.

### 2016-05-04 acl.2016/acl2016.tex
changes

@@
 We propose two optimizations.
-Firstly, the default translation rule caching mechanism in Moses saves the most recently used rules.
+Firstly, the translation rule caching mechanism in Moses saves the most recently used rules.
 However, this require locking and periodic clearing of old rules in order to reduce memory consumption resulting in \emph{slower} decoding, Table~\ref{tab:phrase-table-cache}.
@@
 
-The Moses translation model caches the most recently queried translation rules for later re-use.
-This has been shown to perform badly for fast phrase-tables such as those described by~\cite{Bogoychev:Thesis:2013}.
-We explore a simple caching mechanism that populates the cache during loading with rules that translates the most common source phrases.
+In the first optimization, we create a static translation model cache containing translation rules that translates the most common source phrases.
+This is constructed during phrase-table training based on the source counts.
+The cache is then loaded when the decoder is started.
+It does not require the overhead of managing an active cache but there is still some overhead in using a cache.
 \begin{table}[h]
@@
 \end{table}
-The static cache does not require the overhead of managing the most recently queried lookups but there is still some overhead in using a cache.
-Overall however, there was over a 10\% decrease in decoding time using the optimum cache size, Table~\ref{tab:cache}.
+Overall however, using a static cache result in a 10\% decrease in decoding time if the optimum cache size is used, Table~\ref{tab:cache}.
 
-In the second optimization, we disable the compression.
+For the second optimization, we disable the compression of the target side of the translation rules.
 This increase the size of the binary files from 17GB to 23GB but the time saved not needing to decompress the data resulted in a 1.5\% decrease in decoding time with 1 thread and nearly 7\% when the CPUs are saturated, Table~\ref{tab:compression}.

### 2016-05-05 acl.2016/acl2016.tex
changes after review

@@
 
-~\newcite{mfernandez2016boosting} was concerned with multi-core speed but treat the decoding process as a black box within a parallelization framework.
+~\newcite{mfernandez2016boosting} was concerned with multi-core speed but treated the decoding process as a black box within a parallelization framework.
 
@@
 
-The lexicalized reordering model assign a probability to a translation rule, given the relative ordering of the rule in the hypothesis.
+The lexicalized reordering model requires a probability distribution of the reordering behaviour of each translation rule learnt from the training data.
+This is represented in the model file as a fixed number of probabilities for each rule, exactly how many probabilities is dependant on the model's parameterization during training.
+During decoding, a probability from this distribution is assigned to each hypothesis according to the reordering of the translation rule.
 
-We integrate the lexicalized model file into the translation model but storing the model's probabilities in the phrase-table.
+Rather than holding the model probability distributions in the separate file, we pre-process the translation model file to include the lexicalized reordering model distributions for each rule.
+During decoding, the probability distribution is then taken from the translation model instead of querying a separate file.
+
 This resulted in a significant decrease in decoding time, especially with high number of cores, Figure~\ref{fig:lex-ro}. 
@@
 \end{figure} 
-Integrating the lexicalized reordering model into the translation model decreases decoding time by 29\% with a single core but it is over 5 times faster using all cores.
-In fact, the decoding time with the integrated model is similar to that \emph{without} a lexilized reordering model.
+Decoding time decreased by 29\% with running with a single thread but it is over 5 times faster using 32 threads.
+In fact, the decoding time with the integrated model is similar to that \emph{without} a lexicalized reordering model\footnote{Strangely, using lexicalized reordering model results in slightly faster decoding than without the model.
+We double checked with other datasets.
+We suspect this may be an artefact of the algorithm or data structures, for example, more work is needed to resolve hypothesis recombination.}.
 Critically for systems with multicores servers, it enables the decoder to continue to scale, making efficient use of all available cores.
-This is in contrast to using a separate lexicalized reordering model where decoding time flatten out and actually worsens after approximately 15 threads.
 
@@
 
+Showing decoding \emph{speed} rather than decoding time better illustrate the scalability of our approach.
 Figure~\ref{fig:speed} shows decoding speed against the number of threads, measured in words translated per second.
-There is a constant increase in decoding speed when more threads are used, only decreasing slightly after 16 threads when hyperthreads are used.
+There is a constant increase in decoding speed when more threads are used, decreasing slightly after 16 threads when virtual cores are employed by the CPU.
 Overall, decoding is 12.5 times faster than single-threaded decoding when all 16 cores (32 hyperthreads) are fully utilized.
@@
 This further constrains the resources that each thread has, therefore we see a markedly lower speed increase with each additional thread once there are more than 16 concurrent threads.
-
-Overall though, the scalability is remarkably good, the decoder is able to make full use of all real and virtual cores.
-When all 16 cores are saturated with 2 hyper-thread each, the decoder is 16 times faster than single-threaded decoding.
-It is also 4.5 times faster than Moses with a single-thread and 9.6 faster when all cores are used.
 
@@
 \end{figure}
+
+%Overall though, the scalability is remarkably good, the decoder is able to make full use of all real and virtual cores.
+When all 16 cores are saturated with 2 hyper-thread each, the decoder is 16 times faster than single-threaded decoding.
+
+Our work is 4.5 times faster than Moses with a single-thread and 9.6 faster when all cores are used.
 

### 2016-05-05 acl.2016/acl2016.tex
changes after review

@@
 
-Table~\ref{tab:stack-configuration} and Figure~\ref{fig:stack-configuration} present the tradeoff between decoding time and average model at various pop-limits.
+%Table~\ref{tab:stack-configuration} and 
+Figure~\ref{fig:stack-configuration} present the tradeoff between decoding time and average model at various pop-limits.
 The model scores for all stack configurations are identical for low pop-limits but grouping hypotheses into coverage \& end position produces higher model scores for higher pop-limits.
@@
 This may be due to differing test set characteristics, or interactions with cube-pruning algorithm we used.
-\begin{table*}[t]
-\small
-\begin{center}
-\begin{tabular}{|r|r|r|r|r|r|r|} \hline
-Pop-limit	& \multicolumn{2}{c|}{Cardinality}	& \multicolumn{2}{c|}{Coverage}	& \multicolumn{2}{c|}{Coverage \& end pos} \\ \hline
-		& Time		& Score			& Time		& Score		& Time		& Score	\\ \hline
-100  		& 73		& -8.64513 		& 75		& -8.64513	& 72	& -8.64513 \\
-500		& 237		& -8.59563		& 225		& -8.59563 	& 229	& -8.59612 \\
-1,000	 	& 416		& -8.58700		& 397		& -8.58700	& 423	& -8.58700 \\ 
-5,000 		& 1930		& -8.58165		& 1931		& -8.58165	& 2153	& -8.58098 \\ 
-10,000 		& 3619		& -8.58133		& 3630		& -8.58133	& 4576	& -8.58015 \\ 
-15,000 		& 4830		& -8.58130		& 5001		& -8.58130	& 7156	& -8.57999 \\ 
-20,000 		& 5849		& -8.58130		& 5916		& -8.58130	& 9583	& -8.57994 \\ \hline
-\end{tabular}
-\end{center}
-\caption{Decoding time (in secs with 32 threads) and average model scores for different stack configurations}
-\label{tab:stack-configuration}
-\end{table*}
+%\begin{table*}[t]
+%\small
+%\begin{center}
+%\begin{tabular}{|r|r|r|r|r|r|r|} \hline
+%Pop-limit	& \multicolumn{2}{c|}{Cardinality}	& \multicolumn{2}{c|}{Coverage}	& \multicolumn{2}{c|}{Coverage \& end pos} \\ \hline
+%		& Time		& Score			& Time		& Score		& Time		& Score	\\ \hline
+%100  		& 73		& -8.64513 		& 75		& -8.64513	& 72	& -8.64513 \\
+%500		& 237		& -8.59563		& 225		& -8.59563 	& 229	& -8.59612 \\
+%1,000	 	& 416		& -8.58700		& 397		& -8.58700	& 423	& -8.58700 \\ 
+%5,000 		& 1930		& -8.58165		& 1931		& -8.58165	& 2153	& -8.58098 \\ 
+%10,000 		& 3619		& -8.58133		& 3630		& -8.58133	& 4576	& -8.58015 \\ 
+%15,000 		& 4830		& -8.58130		& 5001		& -8.58130	& 7156	& -8.57999 \\ 
+%20,000 		& 5849		& -8.58130		& 5916		& -8.58130	& 9583	& -8.57994 \\ \hline
+%\end{tabular}
+%\end{center}
+%\caption{Decoding time (in secs with 32 threads) and average model scores for different stack configurations}
+%\label{tab:stack-configuration}
+%\end{table*}
 
@@
 
+We mimic the use of multi-core servers by single-threaded decoders such as mtplz~\cite{Heafield-mtplz} and Jane~\cite{peitz2012jane} by running multiple processes of the Moses decoder, each running one thread.
+The input is distributed to the processes using a external script.
+
+The results in ???
+shows that not only does the parallelization method scales badly even compared to Moses, it is unable to run at all when more that a certain number of threads, possibly due to higher memory requirements of running multiple memory-hungry processes, or the inefficiency of distributing work to many processes.
+Based on this result, single-thread decoding applications are not a viable option for efficiently utilizing the multicore servers of today.
+
 \section{Other Models and Even More Cores}

### 2016-05-05 acl.2016/acl2016.tex
changes

@@
 
-This contrast with Moses where speed increases to approximately 16 threads but then actually become slower thereafter, Figure~\ref{fig:moses-speed}.
+This contrast with Moses where speed increases to approximately 16 threads but then actually become slower thereafter, Figure~\ref{fig:moses-speed} (solid line).
 \begin{figure}[h]
@@
 \begin{tabular}{cc}
-{\includegraphics[scale=0.4]{moses-scalability.png}} 
+{\includegraphics[scale=0.5]{moses-scalability.png}} 
 \end{tabular}
@@
 The input is distributed to the processes using a external script.
-
-The results in ???
-shows that not only does the parallelization method scales badly even compared to Moses, it is unable to run at all when more that a certain number of threads, possibly due to higher memory requirements of running multiple memory-hungry processes, or the inefficiency of distributing work to many processes.
-Based on this result, single-thread decoding applications are not a viable option for efficiently utilizing the multicore servers of today.
+The result in Figure~\ref{fig:moses-speed} (dashed line) shows that this method scales much better than multithreading the application.
 

### 2016-05-06 acl.2016/acl2016.tex
changes

--- 2016-05-05 acl.2016/acl2016.tex
@@
 Jane and mtplz are single-threaded decoders, relying on external applications to parallelize operations.
-This has the advantage it is easier to parallelize processing across multiple servers, not just multiple cores.
+%This has the advantage it is easier to parallelize processing across multiple servers, not just multiple cores.
 As we shall see, this method does not scale beyond a small number of cores when attempting to scale to multiple cores within one server.
@@
 
-We mimic the use of multi-core servers by single-threaded decoders such as mtplz~\cite{Heafield-mtplz} and Jane~\cite{peitz2012jane} by running multiple processes of the Moses decoder, each running one thread.
+%We mimic the use of multi-core servers by single-threaded decoders such as mtplz~\cite{Heafield-mtplz} and Jane~\cite{peitz2012jane} by running multiple processes of the Moses decoder, each running one thread.
 The input is distributed to the processes using a external script.
-
-The results in ???
-shows that not only does the parallelization method scales badly even compared to Moses, it is unable to run at all when more that a certain number of threads, possibly due to higher memory requirements of running multiple memory-hungry processes, or the inefficiency of distributing work to many processes.
-Based on this result, single-thread decoding applications are not a viable option for efficiently utilizing the multicore servers of today.
+The result in Figure~\ref{fig:moses-speed} (dashed line) shows that this method scales much better than multithreading the application.
 

### 2016-05-06 acl.2016/acl2016.tex
changes

@@
 
-The search algorithm creates and destroy a large number of hypothesis and related objects.
-The memory allocation required puts a burden on the operating system due to the need to synchronize  memory access.
+The search algorithm creates and destroy a large number of intermediate objects such as hypotheses and feature function states.
+The memory management this requires puts a burden on the operating system due to the need to synchronize memory access.
 This problem worsens with multi-threaded applications such as Moses.

### 2016-05-06 acl.2016/acl2016.tex
changes

@@
 We shall seek to improve decoding speed by replacing the operating system's general purpose memory management with our own custom memory management scheme.
-In time-critical parts of the decoder, memory will be allocated from a memory pool.
+Memory will be allocated from a memory pool rather than use the operating system's general purpose allocation functions.
 
@@
 To further increase memory management speed, objects are not deleted, instead the memory pool is simply reset and reused for each input sentence.
-This can result in high memory usage so object queues are available for high-churn objects which allows the decoder to re-use portion of the memory pool before it is reset.
+Not deleting objects can result in unacceptably high memory usage so object queues are available for high-churn objects which allows the decoder to re-use portion of the memory pool before it is reset.
 The object queues are LIFO so that recently accessed memory are used, increasing CPU cache hits.

### 2016-05-08 acl.2016/acl2016.tex
changes

@@
 This contrast with Moses where speed increases to approximately 16 threads but then actually become slower thereafter, Figure~\ref{fig:moses-speed}.
+Using the tcmalloc library has a small positive effect on decoding speed but does little to improve scalability
 \begin{figure}[h]
@@
 
+It has been suggested that using a larger language model would overpower the improvements in decoding speed.
+We tested this posit by replacing the language model in our ar-en experiment with a 96GB language model.
+Results show that the decoding speed compared to Moses is less than with a smaller language model, Figure~\ref{fig:large-lm}, however, our work still scales monotonically until all CPUs are saturated, unlike Moses.
+\begin{figure}[h]
+\centering
+\begin{tabular}{cc}
+{\includegraphics[scale=0.4]{large-lm.png}} 
+\end{tabular}
+\caption{Decoding speed with a large language model}
+\label{fig:large-lm}
+\end{figure}
+
 \section{Conclusion}

### 2016-05-08 acl.2016/acl2016.tex
timer loaded

@@
 It has been suggested that using a larger language model would overpower the improvements in decoding speed.
-We tested this posit by replacing the language model in our ar-en experiment with a 96GB language model.
-Results show that the decoding speed compared to Moses is less than with a smaller language model, Figure~\ref{fig:large-lm}, however, our work still scales monotonically until all CPUs are saturated, unlike Moses.
+We tested this conjecture by replacing the language model in the ar-en experiment with a 96GB language model.
+The time to load of language model is significant (394 sec) and was excluded from the translation speed.
+Results show that our decoder is 7 times faster than Moses and still scales monotonically until all CPUs are saturated, Figure~\ref{fig:large-lm}.
 \begin{figure}[h]

### 2016-05-09 acl.2016/acl2016.tex
changes

@@
 
-Therefore, we will focus on optimizing only the phrase-table and lexicalized reordering model.
+Therefore, we will focus on optimizing only the phrase-table and lexicalized re-ordering model.
 
@@
 
-For any phrase-table table of a realistic size, memory and loading time constraints requires us to use load-on-demand implementations of which Moses can make use of, each with differing performance characteristics.
+For any phrase-table table of a realistic size, memory and loading time constraints requires us to use a load-on-demand implementation.
+Moses has several which we can make use of, each with differing performance characteristics.
 Figure~\ref{fig:moses-phrase-tables-time} shows the time taken to decode 800,000 sentences for the fastest two implementations.
@@
 
-\subsection{Lexicalized Reordering Model Optimizations}
+\subsection{Lexicalized Re-ordering Model Optimizations}
 
-Similar to the phrase-table, the lexicalized reordering model is trained on parallel data.
+Similar to the phrase-table, the lexicalized re-ordering model is trained on parallel data.
 A resultant model file is created in training which is then queried during decoding.
@@
 However, the model's query keys are the source and target phrase of each translation rule.
-Rather than storing the lexicalized reordering model separately, we shall integrating it into the translation model, eliminating the need to query a separate file.
+Rather than storing the lexicalized re-ordering model separately, we shall integrating it into the translation model, eliminating the need to query a separate file.
 This has precedent in ~\newcite{peitz2012jane} but the effect on decoding speed were not published.
@@
 The phrase-table was then pruned, keeping only the top 100 entries per source phrase, according to $p(t|s)$.
-All model files were then binarized; the language models were binarized using KenLM~\cite{Heafield-kenlm}, the phrase table using the Probing phrase-table, lexicalized reordering model using the compact data structure.
+All model files were then binarized; the language models were binarized using KenLM~\cite{Heafield-kenlm}, the phrase table using the Probing phrase-table, lexicalized re-ordering model using the compact data structure.
 These binary formats were chosen for their best-in-class multithreaded performance.
@@
 
-\subsection{Lexicalized Reordering Model}
+\subsection{Lexicalized Re-ordering Model}
 
-The lexicalized reordering model requires a probability distribution of the reordering behaviour of each translation rule learnt from the training data.
+The lexicalized re-ordering model requires a probability distribution of the re-ordering behaviour of each translation rule learnt from the training data.
 This is represented in the model file as a fixed number of probabilities for each rule, exactly how many probabilities is dependant on the model's parameterization during training.
-During decoding, a probability from this distribution is assigned to each hypothesis according to the reordering of the translation rule.
+During decoding, a probability from this distribution is assigned to each hypothesis according to the re-ordering of the translation rule.
 
-Rather than holding the model probability distributions in the separate file, we pre-process the translation model file to include the lexicalized reordering model distributions for each rule.
+Rather than holding the model probability distributions in the separate file, we pre-process the translation model file to include the lexicalized re-ordering model distributions for each rule.
 During decoding, the probability distribution is then taken from the translation model instead of querying a separate file.
@@
 \end{tabular}
-\caption{Decoding time with Compact Lexicalized Reordering, and integrated into a model the phrase-table}
+\caption{Decoding time with Compact Lexicalized Re-ordering, and integrated into a model the phrase-table}
 \label{fig:lex-ro}
@@
 Decoding time decreased by 29\% with running with a single thread but it is over 5 times faster using 32 threads.
-In fact, the decoding time with the integrated model is similar to that \emph{without} a lexicalized reordering model\footnote{Strangely, using lexicalized reordering model results in slightly faster decoding than without the model.
+In fact, the decoding time with the integrated model is similar to that \emph{without} a lexicalized re-ordering model\footnote{Strangely, using lexicalized re-ordering model results in slightly faster decoding than without the model.
 We double checked with other datasets.

### 2016-06-09 acl.2016/acl2016.tex
bool -> size_t. Prevent segfault when using mempool

@@
 This paper is the first we know of that focuses on improving decoding speed on multicore servers.
-We will take a holistic approach to solving this issue, re-implementing the decoder's structure to optimize for speed, re-assessing some of the optimization assumptions made in the implementation of significant parts of the decoder, as well as exploring changes to the core search algorithm.
-We will compare and contrast decoder implementations, not just decoding algorithms or individual components.
-
-We will present a phrase-based decoder that is significantly faster than the Moses baseline for single-threaded operation, and scales monotonically with the number of cores.
+We will take a holistic approach to solving this issue.
+We create a decoder that is optimized for multi-core processing speed by concentrating on four main areas.
+\begin{enumerate}
+  \item \vspace{-2 mm} Faster memory management of data-structures through the use of customized memory pools
+  \item \vspace{-2 mm} Exploring alternatives to cardinality-based hypothesis stack configuration
+  \item \vspace{-2 mm} Re-examining the efficiency of phrase-table lookup using translation rule caching and data compression
+  \item \vspace{-2 mm} Integrating the lexicalized re-ordering model into the phrase-table, thus eliminating the need for independent random lookup this model 
+\end{enumerate}
+The result is a decoder that is significantly faster than the Moses baseline for single-threaded operation, and scales with the number of cores.
 
@@
 
-Of these feature functions, the first four are simple and do not leave much room for optimization.
+Of these feature functions, we will focus on optimizing only the phrase-table and lexicalized re-ordering model.
+
+The first four feature functions are simple and do not leave much room for optimization.
 As have already been mentioned, the language model consumes a large part of decoding time but it has already been heavily optimized in previous works such as ~\newcite{Heafield-kenlm} and ~\newcite{yasuhara-EtAl:2013:EMNLP}.
-
-Therefore, we will focus on optimizing only the phrase-table and lexicalized re-ordering model.
 
@@
 
-Apart from the two feature function optimization detailed in the last section, we will investigate two other optimizations, all described below.
+We will also investigate two other optimizations described below.
 

### 2016-06-09 acl.2016/acl2016.tex
bool -> size_t. Prevent segfault when using mempool

@@
 
-\section{Phrase-Based Model}
+\subsection{Feature Functions}
 
@@
 \end{equation}
-where $\lambda_m$ is the weight, and $h_m$ is the feature function, or `score', for model $m$. $Z$ is the partition function which can be ignored for optimization.
-
-\subsection{Beam Search}
-
-A translation of a source sentence is created by applying a series of translation rules which together translate each source word once, and only once.
-Each partial translation is known as a \emph{hypothesis}, which is created by applying a rule to an existing hypothesis.
-This \emph{hypothesis expansion} process starts with a hypothesis that has translated no source word and ends with completed hypotheses that have translated all source words.
-The highest-scoring completed hypothesis, according to the model score, is considered the best translation, $\hat{t} $.
-
-In the phrase-based model, each rule translates a contiguous sequence of source words.
-Successive applications of translation rules do not have to be adjacent on the source side, depending on the distortion limit.
-The target output is constructed strictly left-to-right from the target side from the series of translation rules.
-
-A beam search algorithm is used to create the completed hypothesis set efficiently.
-Partial translations are organized into stacks where each stack holds a number of comparable hypotheses.
-Most phrase-based implementations place hypotheses in the same stack that have the same coverage cardinality $|C|$, where $C$ is  the coverage set, $C \subseteq \{1,2,...
-|s| \} $ of the number of  source words translated.
-
-\subsection{Feature Functions}
-
-Features functions are the $h_m$ in Equation~\ref{eq:Log-Linear}, calculating a score for each hypothesis.
+where $\lambda_m$ is the weight, and $h_m$ is the \emph{feature function}, or `score', for model $m$. $Z$ is the partition function which can be ignored for optimization.
 
@@
 
-Of these feature functions, we will focus on optimizing only the phrase-table and lexicalized re-ordering model.
+Of these feature functions, we will focus on optimizing the speed of the phrase-table and lexicalized re-ordering models.
 
-The first four feature functions are simple and do not leave much room for optimization.
+%The first four feature functions are simple and do not leave much room for optimization.
 As have already been mentioned, the language model consumes a large part of decoding time but it has already been heavily optimized in previous works such as ~\newcite{Heafield-kenlm} and ~\newcite{yasuhara-EtAl:2013:EMNLP}.
+
+\subsection{Beam Search}
+
+A translation of a source sentence is created by applying a series of translation rules which together translate each source word once, and only once.
+Each partial translation is known as a \emph{hypothesis}, which is created by applying a rule to an existing hypothesis.
+This \emph{hypothesis expansion} process starts with a hypothesis that has translated no source word and ends with completed hypotheses that has translated all source words.
+The highest-scoring completed hypothesis, according to the model score, is considered the best translation, $\hat{t} $.
+
+In the phrase-based model, each rule translates a contiguous sequence of source words.
+Successive applications of translation rules do not have to be adjacent on the source side, depending on the distortion limit.
+The target output is constructed strictly left-to-right from the target side from the series of translation rules.
+
+A beam search algorithm is used to create the completed hypothesis set efficiently.
+Hypotheses are grouped into stacks where each stack holds a number of comparable hypotheses.
+Most phrase-based implementations group hypotheses according to coverage cardinality.
+% in the same stack that have the same coverage cardinality $|C|$, where $C$ is  the coverage set, $C \subseteq \{1,2,...
+|s| \} $ of the number of  source words translated.
 
@@
 
-We will also investigate two other optimizations described below.
+We will also concentrate on four main areas for optimization.
 
@@
 
-A memory pool is a large block of memory that has been pre-allocated by the operating system from which smaller blocks can be give out to the application.
-We will use a thread-specific memory pools to avoid locking contention during memory access.
+A memory pool is a large block of memory that has been pre-allocated by the operating system from which smaller blocks can be given out to the application.
+We will use thread-specific memory pools to avoid locking contention during memory access.
 Our memory pool will be dynamic, growing in size when required but never reducing in size.
 
-To further increase memory management speed, objects are not deleted, instead the memory pool is simply reset and reused for each input sentence.
-Not deleting objects can result in unacceptably high memory usage so object queues are available for high-churn objects which allows the decoder to re-use portion of the memory pool before it is reset.
-The object queues are LIFO so that recently accessed memory are used, increasing CPU cache hits.
+To further increase memory management speed, objects in the memory pool are not deleted.
+Unused objects accumulates in the pool until the decoding of a sentence completes and the pool is simply reused for the decoding of the next sentence.
+Accumulating unused objects can result in unacceptably high memory usage so object queues are available for high-churn objects which allows the decoder to re-cycle some unused objects during the decoding of a particular sentence.
+LIFO queues are used so that the most recently accessed memory are used, increasing CPU cache hits.
 

### 2016-06-09 acl.2016/acl2016.tex
bool -> size_t. Prevent segfault when using mempool

@@
 
-A memory pool is a large block of memory that has been pre-allocated by the operating system from which smaller blocks can be given out to the application.
-We will use thread-specific memory pools to avoid locking contention during memory access.
-Our memory pool will be dynamic, growing in size when required but never reducing in size.
+A memory pool is a large block of memory that has been given to the application by the operating system.
+The application is then responsible for allocating portions of this memory to its components as required.
+We will use thread-specific memory pools to increase speed by avoiding locking contention during memory access.
+Our memory pools will be dynamic, growing in size when required but never reducing in size.
+The pools are not deleted until the decoder application ends.
 
 To further increase memory management speed, objects in the memory pool are not deleted.
-Unused objects accumulates in the pool until the decoding of a sentence completes and the pool is simply reused for the decoding of the next sentence.
-Accumulating unused objects can result in unacceptably high memory usage so object queues are available for high-churn objects which allows the decoder to re-cycle some unused objects during the decoding of a particular sentence.
-LIFO queues are used so that the most recently accessed memory are used, increasing CPU cache hits.
+Unused data structures accumulates in the pool until a reset event.
+After the event, the pool is assumed to be empty and simply reused.
+
+We instantiate two memory pools per decoding thread, one which is never reset and another which is reset after the decoding of each sentence.
+Data structures are created in either pool according to their life cycle.
+
+Accumulating unused objects in the memory pools can result in unacceptably high memory usage so object queues are available for high-churn objects which allows the decoder to re-cycle unused objects before the reset event.
+This also increase speed as LIFO queues are used so that the most recently accessed memory are used, increasing CPU cache hits.
 
@@
 
-The most popular stack configuration for phrase-based model, as implemented in Pharaoh, Moses and Joshua, is coverage cardinality, that is, hypotheses that have translated the same number of source words are stored in the same stack.
-Since stack pruning occurs in each stack independently of each other, this affect the decoder's search errors.
+The most popular stack configuration for phrase-based models is coverage cardinality, that is, hypotheses that have translated the same number of source words are stored in the same stack.
+This is implemented in Pharaoh, Moses and Joshua.
 
-However, there are alternatives to the coverage cardinality configuration.
+%Since stack pruning occurs in each stack independently of each other, this affect the decoder's search errors.
+
+However, there are alternatives to this configuration.
 ~\newcite{Och:2001a} uses a single stack for all hypotheses, ~\newcite{Brown:1993} uses coverage stacks (ie.
@@
 Firstly, the translation rule caching mechanism in Moses saves the most recently used rules.
-However, this require locking and periodic clearing of old rules in order to reduce memory consumption resulting in \emph{slower} decoding, Table~\ref{tab:phrase-table-cache}.
+However, this require locking and active management in clearing of old rules.
+The result is \emph{slower} decoding, Table~\ref{tab:phrase-table-cache}.
 \begin{table}[h]
@@
 Compression was championed by ~\newcite{junczys_tsd_2012b} as the main reason behind the speed of their phrase-table but as we saw in Figure~\ref{fig:moses-phrase-tables-time}, this comes at the cost of scalability to large number of threads.
-We shall therefore take the opposite approach to and explore optimizing decoding speed by disabling compression.
+We shall therefore take the opposite approach to and improve decoding speed by disabling compression.
 
@@
 Similar to the phrase-table, the lexicalized re-ordering model is trained on parallel data.
-A resultant model file is created in training which is then queried during decoding.
-Inevitably, the querying results in loss of decoding speed.
+A resultant model file is then queried during decoding.
+The need for random lookup during querying inevitably results in slower decoding speed.
 Previous work such as~\newcite{junczys_tsd_2012b} improve querying speed with more efficient data structures.
@@
 
-Standard Moses phrase-based configurations are used, except that we use the cube-pruning algorithm~\cite{Chiang:2007:cl} with a pop-limit of 400, rather than the basic phrase-based algorithm.
+Standard Moses phrase-based configurations are used, except that we use the cube-pruning algorithm~\cite{Chiang:2007:cl} with a pop-limit of 400\footnote{\scriptsize the pop-limit was chosen following public discussion on finding a balance between decoding speed and translation quality with Moses https://www.mail-archive.com/moses-support\%40mit.edu/msg13181.html }, rather than the basic phrase-based algorithm.
 The cube-pruning algorithm is often employed by users who require fast decoding as it gives them the ability to trade speed with translation quality via a simple pop-limit parameter.
 
-As a baseline, we use a recent\footnote{between January and May 2016} version of the Moses decoder taken from the github repository.
+As a baseline, we use a recent\footnote{between January and June 2016} version of the Moses decoder taken from the github repository.
 
@@
 
-We instantiate two memory pools per decoding thread, one which is never reset and another which is reset after the decoding of each sentence.
-Data structures are created in either pool according to their life cycle.
-
-Hypothesis objects and the data structures it contains are re-used by maintaining a thread-specific LIFO queue of unused, recombined and pruned objects from which new hypotheses are allocated in preference to create new hypotheses in the memory pool.
-
-Each sentence is decoded from start to finish using one worker thread. The CPU affinity of each thread is set to a specific core to minimize memory page faults as a result of threads switching cores.
+Each sentence is decoded from start to finish using one worker thread. %The CPU affinity of each thread is set to a specific core to minimize memory page faults as a result of threads switching cores.
 
@@
 Decoding time decreased by 29\% with running with a single thread but it is over 5 times faster using 32 threads.
-In fact, the decoding time with the integrated model is similar to that \emph{without} a lexicalized re-ordering model\footnote{Strangely, using lexicalized re-ordering model results in slightly faster decoding than without the model.
+%In fact, the decoding time with the integrated model is similar to that \emph{without} a lexicalized re-ordering model\footnote{Strangely, using lexicalized re-ordering model results in slightly faster decoding than without the model.
 We double checked with other datasets.

### 2016-06-09 acl.2016/acl2016.tex
bool -> size_t. Prevent segfault when using mempool

@@
 
-Standard Moses phrase-based configurations are used, except that we use the cube-pruning algorithm~\cite{Chiang:2007:cl} with a pop-limit of 400\footnote{\scriptsize the pop-limit was chosen following public discussion on finding a balance between decoding speed and translation quality with Moses https://www.mail-archive.com/moses-support\%40mit.edu/msg13181.html }, rather than the basic phrase-based algorithm.
+Standard Moses phrase-based configurations are used, except that we use the cube-pruning algorithm~\cite{Chiang:2007:cl} with a pop-limit of 400\footnote{\scriptsize the pop-limit was chosen following public discussion on finding a balance between decoding speed and translation quality with Moses for commercial use https://www.mail-archive.com/moses-support\%40mit.edu/msg13181.html }, rather than the basic phrase-based algorithm.
 The cube-pruning algorithm is often employed by users who require fast decoding as it gives them the ability to trade speed with translation quality via a simple pop-limit parameter.

### 2016-06-09 acl.2016/acl2016.tex
bool -> size_t. Prevent segfault when using mempool

@@
 Rather than storing the lexicalized re-ordering model separately, we shall integrating it into the translation model, eliminating the need to query a separate file.
-This has precedent in ~\newcite{peitz2012jane} but the effect on decoding speed were not published.
-We will compare results with using a separate model in this paper.
+The lexicalized reordering model retains its own weights in the log-linear framework.
+The total weighted score is expected to be identical to the Moses baseline apart from rare training configurations where the lexicalized reordering model is trained on a different dataset than the translation model.
+
+This optimization has precedent in ~\newcite{peitz2012jane} but the effect on decoding speed were not published. We will compare results with using a separate model in this paper.
 
@@
 
-As a baseline, we use a recent\footnote{between January and June 2016} version of the Moses decoder taken from the github repository.
+As a baseline, we use a recent\footnote{as at 8th June, 2016, commit hash 482457bf2bbb5673fbab5f663da5fd3d78ed132f } version of the Moses decoder taken from the github repository.
 

### 2016-06-09 acl.2016/acl2016.tex
fucking rewrite

@@
 The utilization of statistical machine translation (SMT) has grown enormously over the last decade, many using open-source software developed by the NLP community.
-As commercial use has increased, there is need that for the software to be optimized for their requirements, in particular, faster phrase-based decoding and more efficient utilization of modern multicore servers.
+As commercial use has increased, there is need for  software that is optimized for commercial requirements, in particular, faster phrase-based decoding and more efficient utilization of modern multicore servers.
 
-In this paper we re-examining the major components of phrase-based decoding and decoder implementation with particular emphasis on speed and scalability on multicore machines.
+In this paper we re-examine the major components of phrase-based decoding and decoder implementation with particular emphasis on speed and scalability on multicore machines.
 The result is a drop-in replacement for the Moses decoder which is up to fifteen times faster and scales monotonically with the number of cores.
@@
 
-Standard Moses phrase-based configurations are used, except that we use the cube-pruning algorithm~\cite{Chiang:2007:cl} with a pop-limit of 400\footnote{\scriptsize the pop-limit was chosen following public discussion on finding a balance between decoding speed and translation quality with Moses for commercial use https://www.mail-archive.com/moses-support\%40mit.edu/msg13181.html }, rather than the basic phrase-based algorithm.
+Standard Moses phrase-based configurations are used, except that we use the cube-pruning algorithm~\cite{Chiang:2007:cl} with a pop-limit of 400\footnote{the pop-limit was chosen from public discussion on the Moses mailing list on an acceptable balance between decoding speed and translation quality with Moses for commercial use %https://www.mail-archive.com/moses-support\%40mit.edu/msg13181.html
+}, rather than the basic phrase-based algorithm.
 The cube-pruning algorithm is often employed by users who require fast decoding as it gives them the ability to trade speed with translation quality via a simple pop-limit parameter.
 
-As a baseline, we use a recent\footnote{as at 8th June, 2016, commit hash 482457bf2bbb5673fbab5f663da5fd3d78ed132f } version of the Moses decoder taken from the github repository.
+As a baseline, we use a recent\footnote{as at 8th June, 2016, commit hash \scriptsize{482457bf2bbb5673fbab5f663da5fd3d78ed132f} } version of the Moses decoder taken from the github repository.
 

### 2016-06-09 acl.2016/acl2016.tex
fucking rewrite

@@
 
-This paper not only focuses faster single-threaded decoding but also on overcoming the shortcomings of existing decoding implementations on multicore servers.
+This paper not only focuses on faster single-threaded decoding but also on overcoming the shortcomings of existing decoding implementations on multicore servers.
 
@@
 \end{figure} 
-From this, it appears that the Probing phrase-table~\cite{Bogoychev:Thesis:2013} has the fastest translation rule lookup, especially with large number of cores, therefore, we will concentrate exclusively on this implementation from hereon.
+From this, it appears that the Probing phrase-table~\cite{Bogoychev.probing.2016} has the fastest translation rule lookup, especially with large number of cores, therefore, we will concentrate exclusively on this implementation from hereon.
 

### 2016-06-09 acl.2016/acl2016.tex
fucking rewrite

@@
 The utilization of statistical machine translation (SMT) has grown enormously over the last decade, many using open-source software developed by the NLP community.
-As commercial use has increased, there is need for  software that is optimized for commercial requirements, in particular, faster phrase-based decoding and more efficient utilization of modern multicore servers.
+As commercial use has increased, there is need for  software that is optimized for commercial requirements, in particular, fast phrase-based decoding and more efficient utilization of modern multicore servers.
 

### 2016-06-09 acl.2016/acl2016.tex
fucking rewrite

@@
 
+The rest of the paper will be broken up into the following sections.
+The rest of this section will discuss prior work and an outline of the phrase-based model.
+Section~\ref{sec:Proposed Improvements} will then describe the modifications to improve decoding speed.
+We describe the experiment setup in Section~\ref{sec:Experimental Setup} and present results Section~\ref{sec:Results}.
+We conclude in the last section and discuss possible future work.
+
 \subsection{Prior Work}
@@
 
-The rest of the paper will be broken up into the following sections.
-Next, we will describe the phrase-based model and the major implementation components, with particular emphasis on decoding speed.
-We will then describe modifications to improve decoding speed and present results.
-We conclude in the last section and discuss future work.
-
-\subsection{Feature Functions}
+\subsection{Phrase-Based Model}
 
@@
 \section{Proposed Improvements}
+\label{sec:Proposed Improvements}
 
@@
 \section{Experimental Setup}
+\label{sec:Experimental Setup}
 
@@
 \section{Results}
+\label{sec:Results}
 
@@
 In future, we shall investigate other major components of the decoding algorithm, particularly the language model which has not been touched in this paper.
-We shall also explore the underlying reasons for the scalability issues in Moses to get a better understanding where potential performance issues can arise.
-This has application to other algorithms beside MT decoding.
+We are also keen to explore the underlying reasons for the scalability issues in Moses to get a better understanding where potential performance issues can arise.
 

### 2016-06-09 acl.2016/acl2016.tex
fucking rewrite

@@
 
-We shall explore a simpler caching mechanism that creates a static cache at the start of decoding of the most likely translation rules to be used.
+We shall explore a simpler caching mechanism by creating a static cache of the most likely translation rules to be used at the start of decoding.
 
@@
 The lexicalized reordering model retains its own weights in the log-linear framework.
-The total weighted score is expected to be identical to the Moses baseline apart from rare training configurations where the lexicalized reordering model is trained on a different dataset than the translation model.
+The total weighted score is expected to be identical to the Moses baseline apart from edge case training configurations, for example, where the lexicalized reordering model is trained on a different dataset than the translation model.
 

### 2016-06-09 acl.2016/acl2016.tex
fucking rewrite

@@
 
-As a baseline, we use a recent\footnote{as at 8th June, 2016, commit hash \scriptsize{482457bf2bbb5673fbab5f663da5fd3d78ed132f} } version of the Moses decoder taken from the github repository.
+As a baseline, we use a recent\footnote{as at 8th June, 2016, commit hash \scriptsize{bc5f8d15c6ce4bc678ba992860bfd4be6719cee8} } version of the Moses decoder taken from the github repository.
 

### 2016-06-10 acl.2016/acl2016.tex
oov penalty

@@
 Rather than storing the lexicalized re-ordering model separately, we shall integrating it into the translation model, eliminating the need to query a separate file.
-The lexicalized reordering model retains its own weights in the log-linear framework.
-The total weighted score is expected to be identical to the Moses baseline apart from edge case training configurations, for example, where the lexicalized reordering model is trained on a different dataset than the translation model.
+However, the model remain the same under the log-linear framework, including having its own weights.
+%The search algorithm will produce identical results to the Moses baseline apart from edge case training configurations, for example, where the lexicalized reordering model is trained on a different dataset than the translation model.
 

### 2016-06-10 acl.2016/acl2016.tex
oov penalty

@@
 Moses has several which we can make use of, each with differing performance characteristics.
-Figure~\ref{fig:moses-phrase-tables-time} shows the time taken to decode 800,000 sentences for the fastest two implementations.
+Figure~\ref{fig:moses-phrase-tables-time} shows the decoding speed for the fastest two implementations.
 \begin{figure}[h]
@@
 \end{tabular}
-\caption{Moses decoding time with two different phrase-table implementations}
+\caption{Moses decoding speed with two different phrase-table implementations}
 \label{fig:moses-phrase-tables-time}
@@
 		& No cache	& Caching \\ \hline
-Decoding time  	& 2032 		& 2302 (+13\%) \\ \hline
+Decoding time  	& 2877	& 2540 (-12\%) \\ \hline
 \end{tabular}
 \end{center}
-\caption{Decoding time (in sec with 32 threads) when using phrase-table cache}
+\caption{Decoding speed (in words / sec with 32 threads) when using phrase-table cache}
 \label{tab:phrase-table-cache}
@@
 
-Figure~\ref{fig:mempool} compares the decoding time for Moses and our decoder, using the same models, parameters and test set.
+Figure~\ref{fig:mempool} compares the decoding speed for Moses and our decoder, using the same models, parameters and test set.
 Our decoder is over 3 times faster with one thread, and 4.7 times faster using all cores.

### 2016-06-10 acl.2016/acl2016.tex
oov penalty

@@
 \end{tabular}
-\caption{Our decoder's decoding speed}
+\caption{Comparison of decoding speed of our work and Moses (with \& without the tcmalloc library}
 \label{fig:speed}
@@
 
-Showing decoding \emph{speed} rather than decoding time better illustrate the scalability of our approach.
-Figure~\ref{fig:speed} shows decoding speed against the number of threads, measured in words translated per second.
-There is a constant increase in decoding speed when more threads are used, decreasing slightly after 16 threads when virtual cores are employed by the CPU.
-Overall, decoding is 12.5 times faster than single-threaded decoding when all 16 cores (32 hyperthreads) are fully utilized.
+Figure~\ref{fig:speed} shows decoding speed against the number of threads used.
+In our work, there is a constant increase in decoding speed when more threads are used, decreasing slightly after 16 threads when virtual cores are employed by the CPU.
+Overall, decoding is 16 times faster than single-threaded decoding when all 16 cores (32 hyperthreads) are fully utilized.
 
@@
 
-This contrast with Moses where speed increases to approximately 16 threads but then actually become slower thereafter, Figure~\ref{fig:moses-speed}.
+This contrast with Moses where speed increases to approximately 16 threads but then actually become slower thereafter.
 Using the tcmalloc library has a small positive effect on decoding speed but does little to improve scalability
-\begin{figure}[h]
-\centering
-\begin{tabular}{cc}
-{\includegraphics[scale=0.5]{moses-scalability.png}} 
-\end{tabular}
-\caption{Moses' decoding speed}
-\label{fig:moses-speed}
-\end{figure}
+%\begin{figure}[h]
+%\centering
+%\begin{tabular}{cc}
+%{\includegraphics[scale=0.5]{moses-scalability.png}} 
+%\end{tabular}
+%\caption{Moses' decoding speed}
+%\label{fig:moses-speed}
+%\end{figure}
 
@@
 
-Our work is 4.5 times faster than Moses with a single-thread and 9.6 faster when all cores are used.
+Our work is 4.3 times faster than Moses with a single-thread and 10.4 faster when all cores are used.
 

### 2016-06-10 acl.2016/acl2016.tex
oov penalty

@@
 
-As a baseline, we use a recent\footnote{as at 8th June, 2016, commit hash \scriptsize{bc5f8d15c6ce4bc678ba992860bfd4be6719cee8} } version of the Moses decoder taken from the github repository.
+As a baseline, we use a recent\footnote{The experiments were performed between January and May 2016 with the latest github code to hand.
+The main ar-en experiments were rerun with the source code as at 8th June, 2016 to ensure there were no material difference.
+The commit hash was \scriptsize{bc5f8d15c6ce4bc678ba992860bfd4be6719cee8} } version of the Moses decoder taken from the github repository.
 

### 2016-06-10 acl.2016/amta2016.tex
add amta style

\documentclass[]{article}
\usepackage[letterpaper]{geometry}
\usepackage{amta2016}
\usepackage{times}
\usepackage{url}
\usepackage{latexsym}
\usepackage{natbib}
\usepackage{layout}

\newcommand{\confname}{AMTA 2016}
\newcommand{\website}{\protect\url{http://www.amtaweb.org/}}
\newcommand{\contactname}{research track co-chair Lane Schwartz}
\newcommand{\contactemail}{lanes@illinois.edu} 
\newcommand{\conffilename}{amta2016}
\newcommand{\downloadsite}{\protect\url{http://www.amtaweb.org/}}
\newcommand{\paperlength}{$12$ (twelve)}
\newcommand{\shortpaperlength}{$6$ (six)}

%% do not add any other page- or text-size instruction here

\parskip=0.00in

\begin{document}

% \mtsummitHeader{x}{x}{xxx-xxx}{2016}{45-character paper description goes here}{Author(s) initials and last name go here}
\title{\bf Formatting Your Paper \\
  for the \confname~Conference}

\author{\name{\bf First Author} \hfill  \addr{author1@abc.university.country}\\ 
        \name{\bf Second Author} \hfill \addr{author2@abc.university.country}\\ 
        \addr{Department of Science, My University, 
        MyTown, Zip, Country}
\AND
       \name{\bf Third Author} \hfill \addr{author3@abc.university.country}\\
        \addr{Department of Science, My University, 
        MyTown, Zip, Country}
}

\maketitle
\pagestyle{empty}

\begin{abstract}
  This document contains the instructions for preparing a camera-ready
  manuscript for the proceedings of \confname . The document itself
  conforms to its own specifications, and is therefore an example of
  what your manuscript should look like. Authors are asked to conform
  to all the directions reported in this document.
\end{abstract}

\section{Credits}

This text format is derived from that of the Journal of Evolutionary
Computation.\footnote{Originally written by Darrell Whitley, and only
  later modified by Marc Schoenauer, especially regarding the
  bibliography style} The instructions in this document are derived
from those of the MT Summit 2013, % update this list at next change
which was the latest in a long line of adaptations: EAMT 2011, 2010
and 2009, Coling 2008, ACL-07, Coling/ACL-06, EACL-06, ACL-05 and
EACL-03.  All these were based on the formats of earlier ACL and EACL
Conference proceedings.  Those versions were written by several
people, including John Chen, Henry S. Thompson and Donald Walker.

\section{Introduction}

The following formatting instructions are directed to authors of
papers accepted for publication in \confname{}~proceedings, including
the main conference, workshops, posters and demos. See also the
conference website \website~for additional advice and information
regarding submission.  All authors are required to adhere to these
specifications. Authors are required to submit their papers in PDF
(Portable Document Format). Papers are designed to be printed on
\textbf{US letter paper}. Authors from countries in which access to
word processing systems is limited should contact the appropriate
program chair(s) as soon as possible.

\section{General Instructions}

Manuscripts must be in single-column format. The {\nobreak{title}}
must be centered at the top of the first page. Authors' names, email
addresses and complete mailing addresses must appear below the title,
as in the current document. {\bf Type single-spaced.}  Start all pages
directly under the top margin. See the guidelines later regarding
formatting the first page.

Unless otherwise specified, the maximum length of a manuscript is
\paperlength~pages, plus 2 (two) pages for references
%%% for regular and workshop papers and \shortpaperlength~pages for posters and demos, 
(see Section~\ref{sec:length} for additional information on the
maximum number of pages).

\subsection{Electronically-available resources}

This description is provided in \LaTeX2e (\nobreak{\conffilename.tex})
along with the \LaTeX2e style file used to format it
(\nobreak{\conffilename.sty}) and bibliography
(\textsc{Bib}\TeX) example file (\nobreak{\conffilename.bib}); and in
PDF format (\nobreak{\conffilename.pdf}). These files are all
available at \downloadsite. There is also a Microsoft Word document
template (\nobreak{\conffilename.dot}) available at the same URL. We
strongly re\-commend the use of these style files, which have been
appropriately tailored for the \confname~proceedings.

\subsection{Format of Electronic Manuscript}
\label{sect:pdf}

For the production of the electronic manuscript you must use Adobe's
Portable Document Format (PDF). This format can be generated directly
from \LaTeX source files, using program pdf\LaTeX, or it can be
generated from postscript files. On Unix systems, you can use {\tt
  ps2pdf} for this purpose.  Recent versions of Microsoft Windows can
produce PDF directly (File$>$Save As$>$Save As Type: PDF). Otherwise,
you can use Adobe's Distiller or GSview (File$>$Convert$>$pdfwrite);
if you have \textit{cygwin} installed, you can use
\textit{ps2pdf}. Note that some word processing programs generate PDF
which may not include all the necessary fonts (esp. tree diagrams,
symbols). When you print or create the PDF file, there is usually an
option in your printer setup to include none, all or just non-standard
fonts.  Please make sure that you select the option of including ALL
the fonts. {\em Before sending it, test your PDF by printing it from a
  computer different from the one where it was created.} Moreover,
some word processors may generate very large postscript / PDF files,
where each page is rendered as an image. Such images may reproduce
poorly. In this case, try alternative ways to obtain the postscript
and / or PDF. One way on some systems is to install a driver for a
postscript printer, send your document to the printer specifying
``Output to a file'', then convert the file to PDF.

% US Letter
It is important to specify the \textbf{US Letter format} (21.6 cm x
27.9 cm) / (8.5 in x 11 in) when formatting the paper. When working with {\tt
dvips}, for instance, one should specify {\tt -t letter}.

% A4
%It is of utmost importance to specify the \textbf{A4 format} (21.0 cm
%x 29.7 cm) / (8.3 in x 11.7 in) when formatting the paper. When working with
%{\tt dvips}, for instance, one should specify {\tt -t a4}.

Print-outs of the PDF file on US Letter paper should look like the
present document, which conforms to the formatting requirements. {\em
  Note that in order for your paper to print correctly, you should
  disable centering and scale-to-fit options on your printer.} If you
cannot meet the above requirements about the production of your
camera-ready paper, please contact the program chairs as soon as
possible.

\subsection{Layout}
\label{ssec:layout}

Format manuscripts in a single column, in the manner these
instructions are formatted. The exact dimensions for a page on
US letter paper are:

% US letter
\begin{itemize}
\item Left margin: 4.45 cm (1.75 in)
\item Right margin: 3.8 cm (1.5 in)
\item Top margin: 3.8 cm (1.5 in)
\item Bottom margin: 3.2 cm (1.25 in)
\item Text width: 13.3 cm (5.25 in)
\item Text height: 21.0 cm (8.25 in)
\end{itemize}

Pages should not be numbered. In LaTeX, this can be achieved by
inserting  
\begin{quote}
\begin{verbatim}
\pagestyle{empty}
\end{verbatim}
\end{quote}
after the \verb=\maketitle= command.
% \layout{}

\subsection{Fonts}

For uniformity, Adobe's {\bf Times Roman} font should be
used. In \LaTeX2e{} this is accomplished by putting

\begin{quote}
\begin{verbatim}
\usepackage{times}
\usepackage{latexsym}
\end{verbatim}
\end{quote}
in the preamble. If Times Roman is unavailable, use {\bf Computer
  Modern Roman} (\LaTeX2e{}'s default).  Note that the latter is about
  10\% less dense than Adobe's Times Roman font.

\subsection{The First Page}
\label{ssec:first}

Draw a horizontal line the full width of the text above the title, and
another one below the authors' names and affiliations.  Center the
title across the top of the page. Authors' names appear under the
title, one per line, with the author name in boldface, flush to the
left and corresponding email flush to the right.  Affiliations and
mail addresses appear under each corresponding author name, flush to
the left. Do not use footnotes for affiliations. Do not include the
paper ID number assigned during the submission process. Use the
single-column format throughout the text.

{\bf Title}: Place the title centered at the top of the first page, in
a 18-point bold font. Long titles should be typed on two lines without
a blank line intervening. Approximately, put the title at 3.8 cm (1.5
in) from the top of the page, followed by a blank line, then the
authors' names in 11-point bold font, and the affiliations in 11-point
regular font on the following line. Do not use only initials for given
names (middle initials are allowed). Avoid capitalizing last
names. The affiliation should contain the author's complete address,
and if possible an electronic mail address. Leave about 1 cm (0.4 in)
between the affiliation and the body of the first page.

{\bf Abstract}: Type the abstract at the beginning of the first
column. The width of the abstract text should be smaller than the
width of the columns for the text in the body of the paper by about 1
cm (0.4 in) on each side. The word {\bf Abstract} should be in 9-point
bold font above the body of the abstract, which should be in 9-point
font. The abstract should be a concise summary of the general thesis
and conclusions of the paper. It should be no longer than 200 words.

{\bf Text}: Begin typing the main body of the text immediately after
the abstract. Use 10 point font for text. {\bf Indent} when starting a
new paragraph, except for the first paragraph following each heading.

\subsection{Sections}

{\bf Headings}: Type and label section and subsection headings in the
style shown on the present document.  Use numbered sections (Arabic
numerals) in order to facilitate cross references. Number subsections
with the section number and the subsection number separated by a dot,
in Arabic numerals. Do not number subsubsections. Use 9-point bold
font for subsection headings and 11-point bold font for section
headings.

{\bf Citations}: Citations within the text appear in parentheses
as~\citep{Smith} or, if the author's name appears in the text itself,
as \cite{Smith}. Citations in parentheses should not be used as
linguistic phrases; for example, instead of ``\citep{Smith}
\nobreak{argues} that \ldots'' say ``\cite{Smith} \nobreak{argues}
that \ldots''.  Treat double authors as in~\citep{DD15}, but write as
in~\citep{PDC} when more than two authors are involved.
\nobreak{Append} lowercase letters to the year in cases of ambiguity
as in \citep{JonesFirst}.  Collapse multiple citations in parenthesis
as in~\citep{Smith,JonesFirst} and like this for multiple citations
with the same-named \nobreak{author:}
\citep{Smith,SmithConc,JonesFirst,JonesSecond}
%(Tam and Schultz, 2006, 2007; Gledson and Keane, 2008a,b).

\textbf{References}: Gather the full set of references together under the
heading {\bf References}; place the section before any Appendices, unless they
contain references. Arrange the references alphabetically by the first
author's last-name, rather than by order of occurrence in the text, and invert
the first-name and last-name of the first author (only). Provide as complete
a citation as possible, using a consistent format, such as the one for {\em
North American Opera\/}.  Use of full names for
authors rather than initials is preferred.  Use full names for journals and
conferences, not abbreviations (for example ``45th Meeting of the Association
for Computational Linguistics'', not ``ACL07'').

The \LaTeX2e{} and Bib\TeX{} style files provided roughly fit the
American Psychological Association format, allowing regular citations,
short citations and multiple citations as described above.

{\bf Appendices}: Appendices, if any, directly follow the text and the
references (but see above).  Letter them in sequence and provide an
informative title: {\bf Appendix A. Title of Appendix}.

\textbf{Acknowledgement} section should go as a last section immediately
\textit{before the references}.  Do not number the acknowledgement section.

\subsection{Footnotes}

{\bf Footnotes}: Put footnotes at the bottom of the page and use
9-point font. They may be numbered or referred to by asterisks or
other symbols.\footnote{This is how a footnote should appear.}
Footnotes should be separated from the main text by a
line.\footnote{Note the line separating the footnotes from the text.}

\subsection{Copyright}

% \confname{} requires a copyright license statement. This should be inserted as
% an unnumbered footnote on the first column of the first page. The \LaTeX{} style file
% (and the Word template) adds copyright statement automatically. Refer to the style file for instructions to
% change this if required.

\confname{} lets copyright stay with the authors and allows for free verbatim distribution of papers for optimal dissemination. A footer with the text
\begin{quote}
\begin{small}
\begin{it}
  Proceedings of the Twelfth Conference of the Association for Machine Translation in the Americas, Austin, Texas, 29 October - 3 November 2016.\\
  Beregovaya, O., Doyon, J., Green, S., Langlois, L., Richardson, S., Schwartz, L., eds.\\
  \copyright{} 2016 The authors.\\
  This article is licensed under a Creative Commons 3.0 license, no derivative works, attribution, CC-BY-ND.
\end{it}
\end{small}
\end{quote}
will be added to the bottom of all papers when proceedings are
edited. If, for some reason, this copyright notice is not adequate for
your paper, please contact the program chairs as soon as possible.

\subsection{Graphics}

{\bf Illustrations}: Place figures, tables, and photo\-graphs in the
paper near where they are first discussed, rather than at the end, if
possible.

{\bf Captions}: Provide a caption for every illustration; number
each one sequentially in the form:  ``Figure 1. Caption of the
Figure.'', ``Table 1. Caption of the Table.''  Type the captions of
the figures and tables below the body, using 10-point text.

\section{Length of Submission}
\label{sec:length}

Unless otherwise specified, the maximum length is \paperlength~pages,
plus 2 (two) pages for references. 
%%%%% for regular and workshop papers. and \shortpaperlength pages for posters and demos. 
The page limit should be observed strictly. All illustrations and
appendices must be accommodated within these page limits, following
the formatting instructions given in the present document.

\small

\bibliographystyle{apalike}
\bibliography{amta2016}

\end{document}

### 2016-06-10 acl.2016/amta2016.tex
add amta style

@@
 \usepackage{layout}
+\usepackage{algorithmic}
+\usepackage{graphicx}
 
@@
 \begin{abstract}
-  This document contains the instructions for preparing a camera-ready
-  manuscript for the proceedings of \confname . The document itself
-  conforms to its own specifications, and is therefore an example of
-  what your manuscript should look like. Authors are asked to conform
-  to all the directions reported in this document.
+The utilization of statistical machine translation (SMT) has grown enormously over the last decade, many using open-source software developed by the NLP community.
+As commercial use has increased, there is need for  software that is optimized for commercial requirements, in particular, fast phrase-based decoding and more efficient utilization of modern multicore servers.
+
+In this paper we re-examine the major components of phrase-based decoding and decoder implementation with particular emphasis on speed and scalability on multicore machines.
+The result is a drop-in replacement for the Moses decoder which is up to fifteen times faster and scales monotonically with the number of cores.
 \end{abstract}
-
-\section{Credits}
-
-This text format is derived from that of the Journal of Evolutionary
-Computation.\footnote{Originally written by Darrell Whitley, and only
-  later modified by Marc Schoenauer, especially regarding the
-  bibliography style} The instructions in this document are derived
-from those of the MT Summit 2013, % update this list at next change
-which was the latest in a long line of adaptations: EAMT 2011, 2010
-and 2009, Coling 2008, ACL-07, Coling/ACL-06, EACL-06, ACL-05 and
-EACL-03.  All these were based on the formats of earlier ACL and EACL
-Conference proceedings.  Those versions were written by several
-people, including John Chen, Henry S. Thompson and Donald Walker.
 
@@
 
-The following formatting instructions are directed to authors of
-papers accepted for publication in \confname{}~proceedings, including
-the main conference, workshops, posters and demos. See also the
-conference website \website~for additional advice and information
-regarding submission.  All authors are required to adhere to these
-specifications. Authors are required to submit their papers in PDF
-(Portable Document Format). Papers are designed to be printed on
-\textbf{US letter paper}. Authors from countries in which access to
-word processing systems is limited should contact the appropriate
-program chair(s) as soon as possible.
+SMT has steadily progressed from a research discipline to commercial viability during the past decade as can be seen from services such as the Google and Microsoft Translation services.
+As well as general purpose services such as these, there is a large number of companies that offer customized translation systems, as well as  companies and organization that implement in-house solutions.
+Many of these customized solutions use Moses as their SMT engine.
 
-\section{General Instructions}
+For many users, decoding is the most time-critical part of the translation process.
+Making use of the multiple cores that are now ubiquitous in todays servers is a common strategy to ameliorate this issue.
+However, it has been noticed that the Moses decoder, amongst others, is unable to efficiently use multiple cores~\cite{mfernandez2016boosting}.
+That is, decoding speed does not substantially increase when more cores are used, in fact, it may actually \emph{decrease} when using more cores.
+There have been speculation on the causes of the inefficiency as well as potential remedies.
 
-Manuscripts must be in single-column format. The {\nobreak{title}}
-must be centered at the top of the first page. Authors' names, email
-addresses and complete mailing addresses must appear below the title,
-as in the current document. {\bf Type single-spaced.}  Start all pages
-directly under the top margin. See the guidelines later regarding
-formatting the first page.
+This paper is the first we know of that focuses on improving decoding speed on multicore servers.
+We will take a holistic approach to solving this issue.
+We create a decoder that is optimized for multi-core processing speed by concentrating on four main areas.
+\begin{enumerate}
+  \item \vspace{-2 mm} Faster memory management of data-structures through the use of customized memory pools
+  \item \vspace{-2 mm} Exploring alternatives to cardinality-based hypothesis stack configuration
+  \item \vspace{-2 mm} Re-examining the efficiency of phrase-table lookup using translation rule caching and data compression
+  \item \vspace{-2 mm} Integrating the lexicalized re-ordering model into the phrase-table, thus eliminating the need for independent random lookup this model 
+\end{enumerate}
+The result is a decoder that is significantly faster than the Moses baseline for single-threaded operation, and scales with the number of cores.
 
-Unless otherwise specified, the maximum length of a manuscript is
-\paperlength~pages, plus 2 (two) pages for references
-%%% for regular and workshop papers and \shortpaperlength~pages for posters and demos, 
-(see Section~\ref{sec:length} for additional information on the
-maximum number of pages).
+As far as possible, model scores and functionality are compatible with Moses to aid comparison and ease transition for existing users.
+All source code will be made available under an open-source license.
 
-\subsection{Electronically-available resources}
+The rest of the paper will be broken up into the following sections.
+The rest of this section will discuss prior work and an outline of the phrase-based model.
+Section~\ref{sec:Proposed Improvements} will then describe the modifications to improve decoding speed.
+We describe the experiment setup in Section~\ref{sec:Experimental Setup} and present results Section~\ref{sec:Results}.
+We conclude in the last section and discuss possible future work.
 
-This description is provided in \LaTeX2e (\nobreak{\conffilename.tex})
-along with the \LaTeX2e style file used to format it
-(\nobreak{\conffilename.sty}) and bibliography
-(\textsc{Bib}\TeX) example file (\nobreak{\conffilename.bib}); and in
-PDF format (\nobreak{\conffilename.pdf}). These files are all
-available at \downloadsite. There is also a Microsoft Word document
-template (\nobreak{\conffilename.dot}) available at the same URL. We
-strongly re\-commend the use of these style files, which have been
-appropriately tailored for the \confname~proceedings.
+\subsection{Prior Work}
 
-\subsection{Format of Electronic Manuscript}
-\label{sect:pdf}
+Most prior work on increasing decoding speed look to optimizing specific components of the decoder or the decoding algorithm.
 
-For the production of the electronic manuscript you must use Adobe's
-Portable Document Format (PDF). This format can be generated directly
-from \LaTeX source files, using program pdf\LaTeX, or it can be
-generated from postscript files. On Unix systems, you can use {\tt
-  ps2pdf} for this purpose.  Recent versions of Microsoft Windows can
-produce PDF directly (File$>$Save As$>$Save As Type: PDF). Otherwise,
-you can use Adobe's Distiller or GSview (File$>$Convert$>$pdfwrite);
-if you have \textit{cygwin} installed, you can use
-\textit{ps2pdf}. Note that some word processing programs generate PDF
-which may not include all the necessary fonts (esp. tree diagrams,
-symbols). When you print or create the PDF file, there is usually an
-option in your printer setup to include none, all or just non-standard
-fonts.  Please make sure that you select the option of including ALL
-the fonts. {\em Before sending it, test your PDF by printing it from a
-  computer different from the one where it was created.} Moreover,
-some word processors may generate very large postscript / PDF files,
-where each page is rendered as an image. Such images may reproduce
-poorly. In this case, try alternative ways to obtain the postscript
-and / or PDF. One way on some systems is to install a driver for a
-postscript printer, send your document to the printer specifying
-``Output to a file'', then convert the file to PDF.
+~\citep{Heafield-kenlm} and ~\citep{yasuhara-EtAl:2013:EMNLP} describes fast, efficient datastructures for language models.
+~\citep{zens2007efficient} describes an implementation of a phrase-table for an SMT decoder that is loaded on demand, reducing the initial loading time and memory requirements.
+~\citep{junczys_tsd_2012b} extends this by compressing the on-disk phrase table and lexicalized re-ordering model.
 
-% US Letter
-It is important to specify the \textbf{US Letter format} (21.6 cm x
-27.9 cm) / (8.5 in x 11 in) when formatting the paper. When working with {\tt
-dvips}, for instance, one should specify {\tt -t letter}.
+~\citep{Chiang:2007:cl} describes the cube-pruning and cube-growing algorithm which allows the tradeoff between speed and translation quality to the adjusted with a single parameter.
+~\citep{wuebker2012fast} note that language model querying is amongst the most expensive operation in decoding.
+They sought to improved decoding speed by caching score computations early pruning of translation options.
+This work is similar to ~\citep{Heafield-mtplz} which group hypotheses with identical language model context and incrementally expand them, reducing LM querying.
 
-% A4
-%It is of utmost importance to specify the \textbf{A4 format} (21.0 cm
-%x 29.7 cm) / (8.3 in x 11.7 in) when formatting the paper. When working with
-%{\tt dvips}, for instance, one should specify {\tt -t a4}.
+~\citep{mfernandez2016boosting} was concerned with multi-core speed but treated the decoding process as a black box within a parallelization framework.
 
-Print-outs of the PDF file on US Letter paper should look like the
-present document, which conforms to the formatting requirements. {\em
-  Note that in order for your paper to print correctly, you should
-  disable centering and scale-to-fit options on your printer.} If you
-cannot meet the above requirements about the production of your
-camera-ready paper, please contact the program chairs as soon as
-possible.
+There are a number of phrase-based decoding implementations, many of which implements the extensions to the phrase-based model described above.
 
-\subsection{Layout}
-\label{ssec:layout}
+The most well known is Moses~\cite{koehn-EtAl:2007:PosterDemo} which implements a number of speed optimizations, including cube-pruning. It is widely used for MT research and commercial use.
 
-Format manuscripts in a single column, in the manner these
-instructions are formatted. The exact dimensions for a page on
-US letter paper are:
+Joshua~\cite{Joshua-Decoder} also supports cube-pruning for phrase-based models.
+Phrasal~\cite{spence2014phrasal} supports a number of variants of the phrase-based model.
+Jane~\cite{peitz2012jane} supports the language model look-ahead described in ~\citep{wuebker2012fast} in addition to other tools to speed up decoding such as having a separate fast, lightweight decoder.
+mtplz is a specialized decoder developed to implement the incremental decoding described in ~\citep{Heafield-mtplz}.
 
-% US letter
-\begin{itemize}
-\item Left margin: 4.45 cm (1.75 in)
-\item Right margin: 3.8 cm (1.5 in)
-\item Top margin: 3.8 cm (1.5 in)
-\item Bottom margin: 3.2 cm (1.25 in)
-\item Text width: 13.3 cm (5.25 in)
-\item Text height: 21.0 cm (8.25 in)
-\end{itemize}
+The Moses, Joshua and Phrasal decoders implement multithreading, however, they all report scalability problems, either in the paper (Phrasal) or via social media (Moses\footnote{\scriptsize https://github.com/moses-smt/mosesdecoder/issues/39} and Joshua\footnote{\scriptsize https://twitter.com/ApacheJoshua/status/342022794097340416}).
 
-Pages should not be numbered. In LaTeX, this can be achieved by
-inserting  
-\begin{quote}
-\begin{verbatim}
-\pagestyle{empty}
-\end{verbatim}
-\end{quote}
-after the \verb=\maketitle= command.
-% \layout{}
+Jane and mtplz are single-threaded decoders, relying on external applications to parallelize operations.
+%This has the advantage it is easier to parallelize processing across multiple servers, not just multiple cores.
+As we shall see, this method does not scale beyond a small number of cores when attempting to scale to multiple cores within one server.
 
-\subsection{Fonts}
+This paper not only focuses on faster single-threaded decoding but also on overcoming the shortcomings of existing decoding implementations on multicore servers.
 
-For uniformity, Adobe's {\bf Times Roman} font should be
-used. In \LaTeX2e{} this is accomplished by putting
+\subsection{Phrase-Based Model}
 
-\begin{quote}
-\begin{verbatim}
-\usepackage{times}
-\usepackage{latexsym}
-\end{verbatim}
-\end{quote}
-in the preamble. If Times Roman is unavailable, use {\bf Computer
-  Modern Roman} (\LaTeX2e{}'s default).  Note that the latter is about
-  10\% less dense than Adobe's Times Roman font.
+The objective of decoding is to find the target translation with the maximum probability, given a source sentence.
+That is, for a source sentence $s$, the objective is to find a target translation $\hat{t}$ which has the highest conditional probability $p(t | s)$.
+Formally, this is written as:
+\begin{equation}
+\hat{t} = \arg \max_t p( t | s )
+\label{eq:argmax-factored-trans}
+\end{equation}
+where the \emph{arg max} function is the search.
+The log-linear model generalizes Equation~\ref{eq:argmax-factored-trans} to include more component models and weighting each model according to the contribution of each model to the total probability.
+\begin{equation}
+\label{eq:Log-Linear}
+p(t | s) 	=  \frac{1}{Z} \exp ( \sum_m \lambda_m h_m ( t, s)^{} )
+\end{equation}
+where $\lambda_m$ is the weight, and $h_m$ is the \emph{feature function}, or `score', for model $m$. $Z$ is the partition function which can be ignored for optimization.
 
-\subsection{The First Page}
-\label{ssec:first}
+The standard feature functions in the phrase-based model include:
+\begin{enumerate}
+  \item \vspace{-2 mm} a distortion penalty
+  \item \vspace{-2 mm} a phrase-penalty,
+  \item \vspace{-2 mm} a word penalty,
+  \item \vspace{-2 mm} an unknown word penalty.
+  \item \vspace{-2 mm} log transforms of the target language model probability $p(t)$, 
+  \item \vspace{-2 mm} log transforms translation model probabilities, $p_{TM}(t|s) $ and $p_{TM}(s|t)$, and word-based translation probabilities $p_w(t|s)$ and $p_w(s|t)$,
+  \item \vspace{-2 mm} log transforms of the lexicalized re-ordering probabilities,
+\end{enumerate}
+%which we have suffixed with $_{TM} $ to avoid confusion with the overall model probability $ p(t|s) $ and $ p(s|t)$.
 
-Draw a horizontal line the full width of the text above the title, and
-another one below the authors' names and affiliations.  Center the
-title across the top of the page. Authors' names appear under the
-title, one per line, with the author name in boldface, flush to the
-left and corresponding email flush to the right.  Affiliations and
-mail addresses appear under each corresponding author name, flush to
-the left. Do not use footnotes for affiliations. Do not include the
-paper ID number assigned during the submission process. Use the
-single-column format throughout the text.
+Of these feature functions, we will focus on optimizing the speed of the phrase-table and lexicalized re-ordering models.
 
-{\bf Title}: Place the title centered at the top of the first page, in
-a 18-point bold font. Long titles should be typed on two lines without
-a blank line intervening. Approximately, put the title at 3.8 cm (1.5
-in) from the top of the page, followed by a blank line, then the
-authors' names in 11-point bold font, and the affiliations in 11-point
-regular font on the following line. Do not use only initials for given
-names (middle initials are allowed). Avoid capitalizing last
-names. The affiliation should contain the author's complete address,
-and if possible an electronic mail address. Leave about 1 cm (0.4 in)
-between the affiliation and the body of the first page.
+%The first four feature functions are simple and do not leave much room for optimization.
+As have already been mentioned, the language model consumes a large part of decoding time but it has already been heavily optimized in previous works such as ~\citep{Heafield-kenlm} and ~\citep{yasuhara-EtAl:2013:EMNLP}.
 
-{\bf Abstract}: Type the abstract at the beginning of the first
-column. The width of the abstract text should be smaller than the
-width of the columns for the text in the body of the paper by about 1
-cm (0.4 in) on each side. The word {\bf Abstract} should be in 9-point
-bold font above the body of the abstract, which should be in 9-point
-font. The abstract should be a concise summary of the general thesis
-and conclusions of the paper. It should be no longer than 200 words.
+\subsection{Beam Search}
 
-{\bf Text}: Begin typing the main body of the text immediately after
-the abstract. Use 10 point font for text. {\bf Indent} when starting a
-new paragraph, except for the first paragraph following each heading.
+A translation of a source sentence is created by applying a series of translation rules which together translate each source word once, and only once.
+Each partial translation is known as a \emph{hypothesis}, which is created by applying a rule to an existing hypothesis.
+This \emph{hypothesis expansion} process starts with a hypothesis that has translated no source word and ends with completed hypotheses that has translated all source words.
+The highest-scoring completed hypothesis, according to the model score, is considered the best translation, $\hat{t} $.
 
-\subsection{Sections}
+In the phrase-based model, each rule translates a contiguous sequence of source words.
+Successive applications of translation rules do not have to be adjacent on the source side, depending on the distortion limit.
+The target output is constructed strictly left-to-right from the target side from the series of translation rules.
 
-{\bf Headings}: Type and label section and subsection headings in the
-style shown on the present document.  Use numbered sections (Arabic
-numerals) in order to facilitate cross references. Number subsections
-with the section number and the subsection number separated by a dot,
-in Arabic numerals. Do not number subsubsections. Use 9-point bold
-font for subsection headings and 11-point bold font for section
-headings.
+A beam search algorithm is used to create the completed hypothesis set efficiently.
+Hypotheses are grouped into stacks where each stack holds a number of comparable hypotheses.
+Most phrase-based implementations group hypotheses according to coverage cardinality.
+% in the same stack that have the same coverage cardinality $|C|$, where $C$ is  the coverage set, $C \subseteq \{1,2,...
+|s| \} $ of the number of  source words translated.
 
-{\bf Citations}: Citations within the text appear in parentheses
-as~\citep{Smith} or, if the author's name appears in the text itself,
-as \cite{Smith}. Citations in parentheses should not be used as
-linguistic phrases; for example, instead of ``\citep{Smith}
-\nobreak{argues} that \ldots'' say ``\cite{Smith} \nobreak{argues}
-that \ldots''.  Treat double authors as in~\citep{DD15}, but write as
-in~\citep{PDC} when more than two authors are involved.
-\nobreak{Append} lowercase letters to the year in cases of ambiguity
-as in \citep{JonesFirst}.  Collapse multiple citations in parenthesis
-as in~\citep{Smith,JonesFirst} and like this for multiple citations
-with the same-named \nobreak{author:}
-\citep{Smith,SmithConc,JonesFirst,JonesSecond}
-%(Tam and Schultz, 2006, 2007; Gledson and Keane, 2008a,b).
+\section{Proposed Improvements}
+\label{sec:Proposed Improvements}
 
-\textbf{References}: Gather the full set of references together under the
-heading {\bf References}; place the section before any Appendices, unless they
-contain references. Arrange the references alphabetically by the first
-author's last-name, rather than by order of occurrence in the text, and invert
-the first-name and last-name of the first author (only). Provide as complete
-a citation as possible, using a consistent format, such as the one for {\em
-North American Opera\/}.  Use of full names for
-authors rather than initials is preferred.  Use full names for journals and
-conferences, not abbreviations (for example ``45th Meeting of the Association
-for Computational Linguistics'', not ``ACL07'').
+We will also concentrate on four main areas for optimization.
 
-The \LaTeX2e{} and Bib\TeX{} style files provided roughly fit the
-American Psychological Association format, allowing regular citations,
-short citations and multiple citations as described above.
+\subsection{Efficient Memory Allocation}
 
-{\bf Appendices}: Appendices, if any, directly follow the text and the
-references (but see above).  Letter them in sequence and provide an
-informative title: {\bf Appendix A. Title of Appendix}.
+The search algorithm creates and destroy a large number of intermediate objects such as hypotheses and feature function states.
+The memory management this requires puts a burden on the operating system due to the need to synchronize memory access.
+This problem worsens with multi-threaded applications such as Moses.
 
-\textbf{Acknowledgement} section should go as a last section immediately
-\textit{before the references}.  Do not number the acknowledgement section.
+Libraries such as tcmalloc~\cite{ghemawat2009tcmalloc} are designed to reduce locking contention for multi-threaded application but in our case, this is still not enough.
 
-\subsection{Footnotes}
+We shall seek to improve decoding speed by replacing the operating system's general purpose memory management with our own custom memory management scheme.
+Memory will be allocated from a memory pool rather than use the operating system's general purpose allocation functions.
 
-{\bf Footnotes}: Put footnotes at the bottom of the page and use
-9-point font. They may be numbered or referred to by asterisks or
-other symbols.\footnote{This is how a footnote should appear.}
-Footnotes should be separated from the main text by a
-line.\footnote{Note the line separating the footnotes from the text.}
+A memory pool is a large block of memory that has been given to the application by the operating system.
+The application is then responsible for allocating portions of this memory to its components as required.
+We will use thread-specific memory pools to increase speed by avoiding locking contention during memory access.
+Our memory pools will be dynamic, growing in size when required but never reducing in size.
+The pools are not deleted until the decoder application ends.
 
-\subsection{Copyright}
+To further increase memory management speed, objects in the memory pool are not deleted.
+Unused data structures accumulates in the pool until a reset event.
+After the event, the pool is assumed to be empty and simply reused.
 
-% \confname{} requires a copyright license statement. This should be inserted as
-% an unnumbered footnote on the first column of the first page. The \LaTeX{} style file
-% (and the Word template) adds copyright statement automatically. Refer to the style file for instructions to
-% change this if required.
+We instantiate two memory pools per decoding thread, one which is never reset and another which is reset after the decoding of each sentence.
+Data structures are created in either pool according to their life cycle.
 
-\confname{} lets copyright stay with the authors and allows for free verbatim distribution of papers for optimal dissemination. A footer with the text
-\begin{quote}
-\begin{small}
-\begin{it}
-  Proceedings of the Twelfth Conference of the Association for Machine Translation in the Americas, Austin, Texas, 29 October - 3 November 2016.\\
-  Beregovaya, O., Doyon, J., Green, S., Langlois, L., Richardson, S., Schwartz, L., eds.\\
-  \copyright{} 2016 The authors.\\
-  This article is licensed under a Creative Commons 3.0 license, no derivative works, attribution, CC-BY-ND.
-\end{it}
-\end{small}
-\end{quote}
-will be added to the bottom of all papers when proceedings are
-edited. If, for some reason, this copyright notice is not adequate for
-your paper, please contact the program chairs as soon as possible.
+Accumulating unused objects in the memory pools can result in unacceptably high memory usage so object queues are available for high-churn objects which allows the decoder to re-cycle unused objects before the reset event.
+This also increase speed as LIFO queues are used so that the most recently accessed memory are used, increasing CPU cache hits.
 
-\subsection{Graphics}
+%The speed of memory access is dependent on whether the data is in the CPU cache which is a constrained resource compared to memory size, typically 20MB in the latest processors.
+We shall seek to re-use recently accessed information to increase likelihood of the data being in the CPU cache.
 
-{\bf Illustrations}: Place figures, tables, and photo\-graphs in the
-paper near where they are first discussed, rather than at the end, if
-possible.
+%In multiprocessor servers, the CPU cache is attached to each processor and each core.
+If a sentences is being decoded on one CPU is switched to another, the CPU cache on the new CPU must be repopulated, slowing down decoding.
+We will therefore investigate binding threads to specific cores.
 
-{\bf Captions}: Provide a caption for every illustration; number
-each one sequentially in the form:  ``Figure 1. Caption of the
-Figure.'', ``Table 1. Caption of the Table.''  Type the captions of
-the figures and tables below the body, using 10-point text.
+\subsection{Stack Configurations}
 
-\section{Length of Submission}
-\label{sec:length}
+The most popular stack configuration for phrase-based models is coverage cardinality, that is, hypotheses that have translated the same number of source words are stored in the same stack.
+This is implemented in Pharaoh, Moses and Joshua.
 
-Unless otherwise specified, the maximum length is \paperlength~pages,
-plus 2 (two) pages for references. 
-%%%%% for regular and workshop papers. and \shortpaperlength pages for posters and demos. 
-The page limit should be observed strictly. All illustrations and
-appendices must be accommodated within these page limits, following
-the formatting instructions given in the present document.
+%Since stack pruning occurs in each stack independently of each other, this affect the decoder's search errors.
+
+However, there are alternatives to this configuration.
+~\citep{Och:2001a} uses a single stack for all hypotheses, ~\citep{Brown:1993} uses coverage stacks (ie.
+one stack per unique coverage vector) while ~\citep{peitz2012jane} and ~\citep{Zens+Ney:2008:iwslt} apply both coverage and cardinality pruning.
+While useful, these prior works present only one particular stack configuration each.
+
+~\citep{ortizmartinez-garciavarea-casacuberta:2006:WMT} explore a range of stack configurations by defining a granularity parameter which controls the maximum number of stacks required to decode a sentence.
+
+We shall re-visit the question of stack configuration with a particular emphasis on how they can help improve the tradeoff between speed and translation quality.
+We will do so in the context of the cube-pruning algorithm, the algorithm that we will be using and which was not available to many of the earlier work.
+
+\subsection{Phrase-Table Optimizations}
+
+For any phrase-table table of a realistic size, memory and loading time constraints requires us to use a load-on-demand implementation.
+Moses has several which we can make use of, each with differing performance characteristics.
+Figure~\ref{fig:moses-phrase-tables-time} shows the decoding speed for the fastest two implementations.
+\begin{figure}[h]
+\centering
+\begin{tabular}{cc}
+{\includegraphics[scale=0.4]{moses-phrase-table.png}} 
+\end{tabular}
+\caption{Moses decoding speed with two different phrase-table implementations}
+\label{fig:moses-phrase-tables-time}
+\end{figure} 
+From this, it appears that the Probing phrase-table~\cite{Bogoychev.probing.2016} has the fastest translation rule lookup, especially with large number of cores, therefore, we will concentrate exclusively on this implementation from hereon.
+
+We propose two optimizations.
+Firstly, the translation rule caching mechanism in Moses saves the most recently used rules.
+However, this require locking and active management in clearing of old rules.
+The result is \emph{slower} decoding, Table~\ref{tab:phrase-table-cache}.
+\begin{table}[h]
+\small
+\begin{center}
+\begin{tabular}{|l|c|c|} \hline
+		& No cache	& Caching \\ \hline
+Decoding time  	& 2877	& 2540 (-12\%) \\ \hline
+\end{tabular}
+\end{center}
+\caption{Decoding speed (in words / sec with 32 threads) when using phrase-table cache}
+\label{tab:phrase-table-cache}
+\end{table}
+
+We shall explore a simpler caching mechanism by creating a static cache of the most likely translation rules to be used at the start of decoding.
+
+Secondly, the Probing phrase-table use a simple compression algorithm to compress the target side of the translation rule.
+Compression was championed by ~\citep{junczys_tsd_2012b} as the main reason behind the speed of their phrase-table but as we saw in Figure~\ref{fig:moses-phrase-tables-time}, this comes at the cost of scalability to large number of threads.
+We shall therefore take the opposite approach to and improve decoding speed by disabling compression.
+
+\subsection{Lexicalized Re-ordering Model Optimizations}
+
+Similar to the phrase-table, the lexicalized re-ordering model is trained on parallel data.
+A resultant model file is then queried during decoding.
+The need for random lookup during querying inevitably results in slower decoding speed.
+Previous work such as~\citep{junczys_tsd_2012b} improve querying speed with more efficient data structures.
+
+However, the model's query keys are the source and target phrase of each translation rule.
+Rather than storing the lexicalized re-ordering model separately, we shall integrating it into the translation model, eliminating the need to query a separate file.
+However, the model remain the same under the log-linear framework, including having its own weights.
+%The search algorithm will produce identical results to the Moses baseline apart from edge case training configurations, for example, where the lexicalized reordering model is trained on a different dataset than the translation model.
+
+This optimization has precedent in ~\citep{peitz2012jane} but the effect on decoding speed were not published. We will compare results with using a separate model in this paper.
+
+\section{Experimental Setup}
+\label{sec:Experimental Setup}
+
+We trained a phrase-based system using the Moses toolkit with standard settings.
+The training data consisted of most of the publicly available Arabic-English data from Opus~\cite{tiedemann2012parallel} containing over 69 million parallel sentences, and tuned on a held out set.
+The phrase-table was then pruned, keeping only the top 100 entries per source phrase, according to $p(t|s)$.
+All model files were then binarized; the language models were binarized using KenLM~\cite{Heafield-kenlm}, the phrase table using the Probing phrase-table, lexicalized re-ordering model using the compact data structure.
+These binary formats were chosen for their best-in-class multithreaded performance.
+Table~\ref{tab:model-files} gives details of the resultant sizes of the model files.
+For testing decoding speed, we used a subset of the training data, Table~\ref{tab:test-sets}.
+\begin{table}[h]
+\begin{center}
+\begin{tabular}{|l|c|c|} \hline
+		& ar-en	& fr-en \\ \hline
+Phrase table  	& 17 	& 5.8 \\
+Language model 	& 3.1  	& 1.8 \\ 
+Lex re. model	& 2.3	& 637MB \\ \hline
+\end{tabular}
+\end{center}
+\caption{Model sizes in GB}
+\label{tab:model-files}
+\end{table}
+
+For verification with a different dataset, we also used a second system trained on the French-English Europarl corpus (2m parallel sentences).
+The two different systems have characterics that we are interested in analyzing; ar-en have short sentences with large models while fr-en have overly long sentences with smaller models.
+Where we need to compare model scores, we used held out test sets.
+
+\begin{table}[h]
+\begin{center}
+\small
+\begin{tabular}{|l|r|r|} \hline
+		& ar-en		& fr-en \\ \hline
+\multicolumn{3}{|l|}{For speed testing} \\ \hline
+Set name	& \multicolumn{2}{|c|}{Subset of training data} \\
+\# sentences  	& 800k 		& 200k \\
+\# words 	& 5.8m 		& 5.9m \\ 
+Avg words/sent	& 7.3		& 29.7 \\ \hline
+\multicolumn{3}{|l|}{For model score testing} \\ \hline
+Set name	& OpenSubtitles	& newstest2011 \\
+\# sentences  	& 2000 		& 3003 \\
+\# words 	& 14,620 	& 86,162 \\ 
+Avg words/sent	& 7.3		& 28.7 \\ \hline
+\end{tabular}
+\end{center}
+\caption{Test sets}
+\label{tab:test-sets}
+\end{table}
+
+Standard Moses phrase-based configurations are used, except that we use the cube-pruning algorithm~\cite{Chiang:2007:cl} with a pop-limit of 400\footnote{the pop-limit was chosen from public discussion on the Moses mailing list on an acceptable balance between decoding speed and translation quality with Moses for commercial use %https://www.mail-archive.com/moses-support\%40mit.edu/msg13181.html
+}, rather than the basic phrase-based algorithm.
+The cube-pruning algorithm is often employed by users who require fast decoding as it gives them the ability to trade speed with translation quality via a simple pop-limit parameter.
+
+As a baseline, we use a recent\footnote{The experiments were performed between January and May 2016 with the latest github code to hand.
+The main ar-en experiments were rerun with the source code as of 8th June, 2016 to ensure there were no material difference.
+The commit hash was \scriptsize{bc5f8d15c6ce4bc678ba992860bfd4be6719cee8} } version of the Moses decoder taken from the github repository.
+
+For all experiments, we used a Dell PowerEdge R620 server with 16 cores, 32 hyper-threads, split over 2 physical processors (Intel Xeon E5-2650 @ 2.00GHz).
+The server has 380GB RAM.
+The operating system was Ubuntu 14.04, the code was compiled with gcc 4.8.4 and Boost 1.59\footnote{http://boost.org/} and the tcmalloc library.
+
+\section{Results}
+\label{sec:Results}
+
+\subsection{Optimizing Memory}
+
+Each sentence is decoded from start to finish using one worker thread. %The CPU affinity of each thread is set to a specific core to minimize memory page faults as a result of threads switching cores.
+
+As can be seen in Table~\ref{tab:optimizing-memory}, over 24\% of the Moses decoder running time is spent on memory management and this increases to 39\% when 32 threads are used, dampening the scalability of the decoder.
+By contrast, our decoder spends 11\% on memory management and does not significantly increase with more threads.
+\begin{table}[h]
+\begin{center}
+\small
+\begin{tabular}{|l|c|c|c|c|} \hline
+		& \multicolumn{2}{c|} {Moses}	& \multicolumn{2}{c|} {Our Work} \\ \hline
+\# threads	& 1 		& 32	& 1 		& 32  \\ \hline
+Memory  	& 24\%		& 39\% 	& 11\%		& 13\% \\
+LM 		& 12\%	 	& 2\% 	& 47\%		& 38\% \\ 
+Phrase-table	& 9\%	 	& 5\% 	& 2\%		& 4\% \\ 
+Lex RO 		& 8\%	 	& 2\% 	& 2\%		& 2\% \\ 
+Search 		& 2\%	 	& 0\% 	& 14\%		& 19\% \\ 
+Misc/Unknown	& 45\%	 	& 39\% 	& 24\%		& 29\% \\ \hline
+\end{tabular}
+\end{center}
+\caption{Profile of \%age decoding time}
+\label{tab:optimizing-memory}
+\end{table}
+
+Figure~\ref{fig:mempool} compares the decoding speed for Moses and our decoder, using the same models, parameters and test set.
+Our decoder is over 3 times faster with one thread, and 4.7 times faster using all cores.
+Like Moses, performance actually worsens after approximately 15 threads, however, the problem is not as pronounced.
+This gives us a better foundation on which to build further innovations for fast, multi-core decoding.
+\begin{figure}[h]
+\centering
+\begin{tabular}{cc}
+{\includegraphics[scale=0.4]{mempool.png}} 
+\end{tabular}
+\caption{Decoding time of Moses and our decoder, using the same models}
+\label{fig:mempool}
+\end{figure}
+
+% A potential disadvantage of using memory pool is that the decoder will use more memory, however, Table~\ref{tab:memory-usage} shows that peak memory consumption are comparable.
+% \begin{table}[h]
+% \begin{center}
+% \begin{tabular}{|l|c|c|} \hline
+% 		& 1 thread	& 32 threads \\ \hline
+% Moses  		& ???	 	& ??? \\
+% Our work 	& ??? 		& ??? \\ \hline
+% \end{tabular}
+% \end{center}
+% \caption{Peak Memory Usage in GB}
+% \label{tab:memory-usage}
+% \end{table}
+
+\subsection{Stack Configuration}
+
+We shall investigate the effects of the following three stack configurations on model score and decoding speed:
+\begin{enumerate}
+  \item \vspace{-2 mm} coverage cardinality,
+  \item \vspace{-2 mm} coverage,
+  \item \vspace{-2 mm} coverage and end position of most recently translated source word.
+\end{enumerate}
+Coverage cardinality is the same as that in Moses and Joshua. Coverage configuration uses one stack per unique coverage vector.
+
+\emph{Coverage and end position of most recently translated source word} extends the coverage configuration by separating hypotheses where the last translate word are different.
+This is an optimization to reduce the number of checks on the distortion limit, which is dependent on the last word position.
+
+The check is a binary function $d(C_{h}, e_{hypo}, range_r)$, where $C_{h}$ is the coverage vector of hypothesis $h$, $e_{h}$ is the end position of most recent source word that has been translated, and $range_r$ is the coverage of the rule to be applied.
+Figure~\ref{algo:Hypothesis Expansion with Unsorted Stack} shows the standard hypothesis expansion algorithm.
+\begin{figure} [h]
+\begin{algorithmic}
+%\REQUIRE source sentence $s$, translation options
+\FORALL{ $h$ in $stack_{|C|}$} 
+  \FORALL{$r$ in translation rules}
+      \IF {$d(C_{h}, e_{hypo}, range_r)$}
+	\STATE expand $h$ with $r$ $\rightarrow$ $h'$
+	\STATE add $h'$ to stack $stack_{C'}$
+      \ENDIF
+  \ENDFOR % { translation options}
+\ENDFOR % { hypothesis in $stack_i$} 
+\end{algorithmic}
+\caption{Hypothesis Expansion with Cardinality Stacks}
+\label{algo:Hypothesis Expansion with Unsorted Stack}
+\end{figure}
+
+By grouping hypotheses according to coverage \emph{and} end position into groups we call \emph{'ministack'}, the distortion limit only needs to be checked for each group, Figure~\ref{algo:Hypothesis Expansion with Sorted Stack}.
+\begin{figure} [h]
+\begin{algorithmic}
+%\REQUIRE source sentence $s$, translation options
+\FORALL{ $ministack_{C, e}$ in $stack_{|C|}$} 
+  \FORALL{$r$ in translation rules}
+    \IF {$d(C_{h}, e_{hypo}, range_r)$}
+      \FORALL{ $h$ in $ministack_{C, e}$} 
+	\STATE expand $h$ with $r$ $\rightarrow$ $h'$
+	\STATE add $h'$ to $ministack_{C',e'}$
+      \ENDFOR
+    \ENDIF
+  \ENDFOR % { translation options}
+\ENDFOR % { hypothesis in $stack_i$} 
+\end{algorithmic}
+\caption{Hypothesis Expansion with Coverage \& End Position Stacks}
+\label{algo:Hypothesis Expansion with Sorted Stack}
+\end{figure}
+
+Also, stack pruning occurs on each hypothesis group independently, therefore, potentially affect search errors and model scores.
+
+%Table~\ref{tab:stack-configuration} and 
+Figure~\ref{fig:stack-configuration} present the tradeoff between decoding time and average model at various pop-limits.
+The model scores for all stack configurations are identical for low pop-limits but grouping hypotheses into coverage \& end position produces higher model scores for higher pop-limits.
+It is also slower but the time/quality tradeoff is better overall with this stack configuration.
+Model scores for coverage and cardinality configurations are identical, which differs from the results in ~\citep{ortizmartinez-garciavarea-casacuberta:2006:WMT}.
+This may be due to differing test set characteristics, or interactions with cube-pruning algorithm we used.
+%\begin{table*}[t]
+%\small
+%\begin{center}
+%\begin{tabular}{|r|r|r|r|r|r|r|} \hline
+%Pop-limit	& \multicolumn{2}{c|}{Cardinality}	& \multicolumn{2}{c|}{Coverage}	& \multicolumn{2}{c|}{Coverage \& end pos} \\ \hline
+%		& Time		& Score			& Time		& Score		& Time		& Score	\\ \hline
+%100  		& 73		& -8.64513 		& 75		& -8.64513	& 72	& -8.64513 \\
+%500		& 237		& -8.59563		& 225		& -8.59563 	& 229	& -8.59612 \\
+%1,000	 	& 416		& -8.58700		& 397		& -8.58700	& 423	& -8.58700 \\ 
+%5,000 		& 1930		& -8.58165		& 1931		& -8.58165	& 2153	& -8.58098 \\ 
+%10,000 		& 3619		& -8.58133		& 3630		& -8.58133	& 4576	& -8.58015 \\ 
+%15,000 		& 4830		& -8.58130		& 5001		& -8.58130	& 7156	& -8.57999 \\ 
+%20,000 		& 5849		& -8.58130		& 5916		& -8.58130	& 9583	& -8.57994 \\ \hline
+%\end{tabular}
+%\end{center}
+%\caption{Decoding time (in secs with 32 threads) and average model scores for different stack configurations}
+%\label{tab:stack-configuration}
+%\end{table*}
+
+\begin{figure}[h]
+\centering
+\begin{tabular}{cc}
+{\includegraphics[scale=0.4]{stack-configuration.png}} 
+\end{tabular}
+\caption{Trade-off between decoding time average model scores for different stack configurations}
+\label{fig:stack-configuration}
+\end{figure} 
+ %, therefore, we shall stick with this Configuration for the remainder of the paper.
+
+%Also, the cube-pruning algorithm contain a priority queue per set of hypotheses sets.
+This can be independent of how hypotheses are grouped.
+In the Moses implementation, the priority queue is also attached to the coverage cardinality, Figure~\ref{algo:Cube Pruning with Unsorted Stack}.
+%\begin{figure} [h]
+%\begin{algorithmic}
+%%\REQUIRE source sentence $s$, translation options
+%\STATE initialize $queue_{|C|}$
+%\FOR{1 to pop-limit}
+%  \STATE get best $item$ in $queue_{|C|}$
+%  \STATE create hypo from $item$, coverage $C_{new}$
+%  \STATE add hypo to $stack_{|C_{new}|}$
+%  \STATE create next $items$
+%  \STATE add new $items$ to $queue_{|C|}$
+%\ENDFOR % { hypothesis in $stack_i$} 
+%\end{algorithmic}
+%\caption{Cube Pruning with Cardinality Stacks}
+%\label{algo:Cube Pruning with Unsorted Stack}
+%\end{figure}
+%Again, we experiment with diffirent queue configurations, having separate queues for each cardinality, coverage, and coverage \& end position.
+The stack configuration remained constant (coverage \& end position with pop-limit of 400).
+%\begin{table}[h]
+%\small
+%\begin{center}
+%\begin{tabular}{|r|r|r|} \hline
+%Queue configuration		& Time		& Score \\ \hline
+%Cardinality			& 192		& -8.59922 \\
+%Coverage			& 2,413		& -8.58635 \\
+%Coverage \& end position	& 7,472		& -8.58263 \\ \hline
+%\end{tabular}
+%\end{center}
+%\caption{Decoding time (in secs with 32 threads) and average model scores for different queue configurations}
+%\label{tab:queue-configuration}
+%\end{table}
+%From the results in Table~\ref{tab:queue-configuration}, using finer grain queues does results in better model scores but it is significantly slower to decode.
+%
+
+\subsection{Translation Model}
+
+In the first optimization, we create a static translation model cache containing translation rules that translates the most common source phrases.
+This is constructed during phrase-table training based on the source counts.
+The cache is then loaded when the decoder is started.
+It does not require the overhead of managing an active cache but there is still some overhead in using a cache.
+\begin{table}[h]
+\small
+\begin{center}
+\begin{tabular}{|r|r|r|} \hline
+Cache size	& Decoding Time & Cache Hit \%age	\\ \hline
+Before caching	& 229	& N/A	 \\
+0		& 239 (+4.4\%)	& 0\% \\
+1,000		& 213 (-7.0\%)	& 11\% \\
+2,000		& 204 (-10.9\%)	& 13\% \\
+4,000		& 205 (-10.5\%)	& 14\% \\
+10,000		& 207 (-9.7\%)	& 17\% \\ \hline
+\end{tabular}
+\end{center}
+\caption{Decoding time (in secs with 32 threads) for varying cache sizes}
+\label{tab:cache}
+\end{table}
+Overall however, using a static cache result in a 10\% decrease in decoding time if the optimum cache size is used, Table~\ref{tab:cache}.
+
+For the second optimization, we disable the compression of the target side of the translation rules.
+This increase the size of the binary files from 17GB to 23GB but the time saved not needing to decompress the data resulted in a 1.5\% decrease in decoding time with 1 thread and nearly 7\% when the CPUs are saturated, Table~\ref{tab:compression}.
+\begin{table}[h]
+\small
+\begin{center}
+\begin{tabular}{|r|r|r|} \hline
+\# threads	& Compressed pt & Non-compressed pt \\ \hline
+1		& 3052	& 3006 (-1.5\%) \\
+5		& 756	& 644 (-14.8\%) \\
+10		& 372	& 362 (-2.7\%) \\
+15		& 284	& 250 (-12.0\%) \\
+20		& 244	& 227 (-7.0\%) \\
+25		& 218	& 209 (-4.1\%) \\
+30		& 206	& 192 (-6.8\%) \\
+35		& 203	& 189 (-6.9\%) \\ \hline
+\end{tabular}
+\end{center}
+\caption{Decoding time (in secs with 32 threads) for compressed and non-compressed phrase-tables}
+\label{tab:compression}
+\end{table}
+
+\subsection{Lexicalized Re-ordering Model}
+
+The lexicalized re-ordering model requires a probability distribution of the re-ordering behaviour of each translation rule learnt from the training data.
+This is represented in the model file as a fixed number of probabilities for each rule, exactly how many probabilities is dependant on the model's parameterization during training.
+During decoding, a probability from this distribution is assigned to each hypothesis according to the re-ordering of the translation rule.
+
+Rather than holding the model probability distributions in the separate file, we pre-process the translation model file to include the lexicalized re-ordering model distributions for each rule.
+During decoding, the probability distribution is then taken from the translation model instead of querying a separate file.
+
+This resulted in a significant decrease in decoding time, especially with high number of cores, Figure~\ref{fig:lex-ro}. 
+\begin{figure}[h]
+\centering
+\begin{tabular}{cc}
+{\includegraphics[scale=0.4]{lex-ro.png}} 
+\end{tabular}
+\caption{Decoding time with Compact Lexicalized Re-ordering, and integrated into a model the phrase-table}
+\label{fig:lex-ro}
+\end{figure} 
+Decoding time decreased by 29\% with running with a single thread but it is over 5 times faster using 32 threads.
+%In fact, the decoding time with the integrated model is similar to that \emph{without} a lexicalized re-ordering model\footnote{Strangely, using lexicalized re-ordering model results in slightly faster decoding than without the model.
+We double checked with other datasets.
+We suspect this may be an artefact of the algorithm or data structures, for example, more work is needed to resolve hypothesis recombination.}.
+Critically for systems with multicores servers, it enables the decoder to continue to scale, making efficient use of all available cores.
+
+\subsection{Scalability}
+
+\begin{figure}[h]
+\centering
+\begin{tabular}{cc}
+{\includegraphics[scale=0.4]{scalability.png}} 
+\end{tabular}
+\caption{Comparison of decoding speed of our work and Moses (with \& without the tcmalloc library}
+\label{fig:speed}
+\end{figure}
+
+Figure~\ref{fig:speed} shows decoding speed against the number of threads used.
+In our work, there is a constant increase in decoding speed when more threads are used, decreasing slightly after 16 threads when virtual cores are employed by the CPU.
+Overall, decoding is 16 times faster than single-threaded decoding when all 16 cores (32 hyperthreads) are fully utilized.
+
+%Since the test server only has 16 real cores, additional threads must be run on the same physical cores as existing threads.
+This further constrains the resources that each thread has, therefore we see a markedly lower speed increase with each additional thread once there are more than 16 concurrent threads.
+
+This contrast with Moses where speed increases to approximately 16 threads but then actually become slower thereafter.
+Using the tcmalloc library has a small positive effect on decoding speed but does little to improve scalability
+%\begin{figure}[h]
+%\centering
+%\begin{tabular}{cc}
+%{\includegraphics[scale=0.5]{moses-scalability.png}} 
+%\end{tabular}
+%\caption{Moses' decoding speed}
+%\label{fig:moses-speed}
+%\end{figure}
+
+%Overall though, the scalability is remarkably good, the decoder is able to make full use of all real and virtual cores.
+When all 16 cores are saturated with 2 hyper-thread each, the decoder is 16 times faster than single-threaded decoding.
+
+Our work is 4.3 times faster than Moses with a single-thread and 10.4 faster when all cores are used.
+
+%We mimic the use of multi-core servers by single-threaded decoders such as mtplz~\cite{Heafield-mtplz} and Jane~\cite{peitz2012jane} by running multiple processes of the Moses decoder, each running one thread.
+The input is distributed to the processes using a external script.
+The result in Figure~\ref{fig:moses-speed} (dashed line) shows that this method scales much better than multithreading the application.
+
+\section{Other Models and Even More Cores}
+
+Our decoder show no scalability issues when we tested with the same model and tested set on a larger server, Figure~\ref{fig:more-cores}.
+\begin{figure}[h]
+\centering
+\begin{tabular}{cc}
+{\includegraphics[scale=0.4]{more-cores.png}} 
+\end{tabular}
+\caption{Decoding speed for with bigger servers}
+\label{fig:more-cores}
+\end{figure}
+
+We verify the results with the French-English phrase-based system and test set.
+The speed gains are even greater than the Arabic-English test scenario, Figure~\ref{fig:fr-en-speed}.
+Our decoder is 5.4 times faster than Moses with a single-thread and 14.5 faster when all cores are saturated.
+\begin{figure}[h]
+\centering
+\begin{tabular}{cc}
+{\includegraphics[scale=0.4]{fr-en-speed.png}} 
+\end{tabular}
+\caption{Decoding speed for fr-en model}
+\label{fig:fr-en-speed}
+\end{figure}
+
+It has been suggested that using a larger language model would overpower the improvements in decoding speed.
+We tested this conjecture by replacing the language model in the ar-en experiment with a 96GB language model.
+The time to load of language model is significant (394 sec) and was excluded from the translation speed.
+Results show that our decoder is 7 times faster than Moses and still scales monotonically until all CPUs are saturated, Figure~\ref{fig:large-lm}.
+\begin{figure}[h]
+\centering
+\begin{tabular}{cc}
+{\includegraphics[scale=0.4]{large-lm.png}} 
+\end{tabular}
+\caption{Decoding speed with a large language model}
+\label{fig:large-lm}
+\end{figure}
+
+\section{Conclusion}
+
+We have presented a new decoder that is compatible with Moses.
+By studying the shortcomings of the current implementation, we are able to optimize for speed, particularly for multicore operation.
+This resulted in double digit gains compared to Moses on the same hardware.
+Our implementation is also unaffected by scalability issues that has afflicted Moses.
+% and continue to scale well for all possible cores we have tested it on.
+
+In future, we shall investigate other major components of the decoding algorithm, particularly the language model which has not been touched in this paper.
+We are also keen to explore the underlying reasons for the scalability issues in Moses to get a better understanding where potential performance issues can arise.
+
+% \section*{Acknowledgments}
+% 
+% Do not number the acknowledgment section.
+% This section should not be presented for the submission version.
 
@@
 \bibliographystyle{apalike}
-\bibliography{amta2016}
+\bibliography{amta2016,mt,more}
 

### 2016-06-10 acl.2016/amta2016.tex
cite format

@@
 Making use of the multiple cores that are now ubiquitous in todays servers is a common strategy to ameliorate this issue.
-However, it has been noticed that the Moses decoder, amongst others, is unable to efficiently use multiple cores~\cite{mfernandez2016boosting}.
+However, it has been noticed that the Moses decoder, amongst others, is unable to efficiently use multiple cores~\citep{mfernandez2016boosting}.
 That is, decoding speed does not substantially increase when more cores are used, in fact, it may actually \emph{decrease} when using more cores.
@@
 
-~\citep{Heafield-kenlm} and ~\citep{yasuhara-EtAl:2013:EMNLP} describes fast, efficient datastructures for language models.
-~\citep{zens2007efficient} describes an implementation of a phrase-table for an SMT decoder that is loaded on demand, reducing the initial loading time and memory requirements.
-~\citep{junczys_tsd_2012b} extends this by compressing the on-disk phrase table and lexicalized re-ordering model.
+~\cite{Heafield-kenlm} and ~\cite{yasuhara-EtAl:2013:EMNLP} describes fast, efficient datastructures for language models.
+~\cite{zens2007efficient} describes an implementation of a phrase-table for an SMT decoder that is loaded on demand, reducing the initial loading time and memory requirements.
+~\cite{junczys_tsd_2012b} extends this by compressing the on-disk phrase table and lexicalized re-ordering model.
 
-~\citep{Chiang:2007:cl} describes the cube-pruning and cube-growing algorithm which allows the tradeoff between speed and translation quality to the adjusted with a single parameter.
-~\citep{wuebker2012fast} note that language model querying is amongst the most expensive operation in decoding.
+~\cite{Chiang:2007:cl} describes the cube-pruning and cube-growing algorithm which allows the tradeoff between speed and translation quality to the adjusted with a single parameter.
+~\cite{wuebker2012fast} note that language model querying is amongst the most expensive operation in decoding.
 They sought to improved decoding speed by caching score computations early pruning of translation options.
-This work is similar to ~\citep{Heafield-mtplz} which group hypotheses with identical language model context and incrementally expand them, reducing LM querying.
+This work is similar to ~\cite{Heafield-mtplz} which group hypotheses with identical language model context and incrementally expand them, reducing LM querying.
 
-~\citep{mfernandez2016boosting} was concerned with multi-core speed but treated the decoding process as a black box within a parallelization framework.
+~\cite{mfernandez2016boosting} was concerned with multi-core speed but treated the decoding process as a black box within a parallelization framework.
 
@@
 
-The most well known is Moses~\cite{koehn-EtAl:2007:PosterDemo} which implements a number of speed optimizations, including cube-pruning. It is widely used for MT research and commercial use.
+The most well known is Moses~\citep{koehn-EtAl:2007:PosterDemo} which implements a number of speed optimizations, including cube-pruning. It is widely used for MT research and commercial use.
 
-Joshua~\cite{Joshua-Decoder} also supports cube-pruning for phrase-based models.
-Phrasal~\cite{spence2014phrasal} supports a number of variants of the phrase-based model.
-Jane~\cite{peitz2012jane} supports the language model look-ahead described in ~\citep{wuebker2012fast} in addition to other tools to speed up decoding such as having a separate fast, lightweight decoder.
-mtplz is a specialized decoder developed to implement the incremental decoding described in ~\citep{Heafield-mtplz}.
+Joshua~\citep{Joshua-Decoder} also supports cube-pruning for phrase-based models.
+Phrasal~\citep{spence2014phrasal} supports a number of variants of the phrase-based model.
+Jane~\citep{peitz2012jane} supports the language model look-ahead described in ~\cite{wuebker2012fast} in addition to other tools to speed up decoding such as having a separate fast, lightweight decoder.
+mtplz is a specialized decoder developed to implement the incremental decoding described in ~\cite{Heafield-mtplz}.
 
@@
 %The first four feature functions are simple and do not leave much room for optimization.
-As have already been mentioned, the language model consumes a large part of decoding time but it has already been heavily optimized in previous works such as ~\citep{Heafield-kenlm} and ~\citep{yasuhara-EtAl:2013:EMNLP}.
+As have already been mentioned, the language model consumes a large part of decoding time but it has already been heavily optimized in previous works such as ~\cite{Heafield-kenlm} and ~\cite{yasuhara-EtAl:2013:EMNLP}.
 
@@
 
-Libraries such as tcmalloc~\cite{ghemawat2009tcmalloc} are designed to reduce locking contention for multi-threaded application but in our case, this is still not enough.
+Libraries such as tcmalloc~\citep{ghemawat2009tcmalloc} are designed to reduce locking contention for multi-threaded application but in our case, this is still not enough.
 
@@
 However, there are alternatives to this configuration.
-~\citep{Och:2001a} uses a single stack for all hypotheses, ~\citep{Brown:1993} uses coverage stacks (ie.
-one stack per unique coverage vector) while ~\citep{peitz2012jane} and ~\citep{Zens+Ney:2008:iwslt} apply both coverage and cardinality pruning.
+~\cite{Och:2001a} uses a single stack for all hypotheses, ~\cite{Brown:1993} uses coverage stacks (ie.
+one stack per unique coverage vector) while ~\cite{peitz2012jane} and ~\cite{Zens+Ney:2008:iwslt} apply both coverage and cardinality pruning.
 While useful, these prior works present only one particular stack configuration each.
 
-~\citep{ortizmartinez-garciavarea-casacuberta:2006:WMT} explore a range of stack configurations by defining a granularity parameter which controls the maximum number of stacks required to decode a sentence.
+~\cite{ortizmartinez-garciavarea-casacuberta:2006:WMT} explore a range of stack configurations by defining a granularity parameter which controls the maximum number of stacks required to decode a sentence.
 
@@
 \end{figure} 
-From this, it appears that the Probing phrase-table~\cite{Bogoychev.probing.2016} has the fastest translation rule lookup, especially with large number of cores, therefore, we will concentrate exclusively on this implementation from hereon.
+From this, it appears that the Probing phrase-table~\citep{Bogoychev.probing.2016} has the fastest translation rule lookup, especially with large number of cores, therefore, we will concentrate exclusively on this implementation from hereon.
 
@@
 Secondly, the Probing phrase-table use a simple compression algorithm to compress the target side of the translation rule.
-Compression was championed by ~\citep{junczys_tsd_2012b} as the main reason behind the speed of their phrase-table but as we saw in Figure~\ref{fig:moses-phrase-tables-time}, this comes at the cost of scalability to large number of threads.
+Compression was championed by ~\cite{junczys_tsd_2012b} as the main reason behind the speed of their phrase-table but as we saw in Figure~\ref{fig:moses-phrase-tables-time}, this comes at the cost of scalability to large number of threads.
 We shall therefore take the opposite approach to and improve decoding speed by disabling compression.
@@
 The need for random lookup during querying inevitably results in slower decoding speed.
-Previous work such as~\citep{junczys_tsd_2012b} improve querying speed with more efficient data structures.
+Previous work such as~\cite{junczys_tsd_2012b} improve querying speed with more efficient data structures.
 
@@
 
-This optimization has precedent in ~\citep{peitz2012jane} but the effect on decoding speed were not published. We will compare results with using a separate model in this paper.
+This optimization has precedent in ~\cite{peitz2012jane} but the effect on decoding speed were not published. We will compare results with using a separate model in this paper.
 
@@
 We trained a phrase-based system using the Moses toolkit with standard settings.
-The training data consisted of most of the publicly available Arabic-English data from Opus~\cite{tiedemann2012parallel} containing over 69 million parallel sentences, and tuned on a held out set.
+The training data consisted of most of the publicly available Arabic-English data from Opus~\citep{tiedemann2012parallel} containing over 69 million parallel sentences, and tuned on a held out set.
 The phrase-table was then pruned, keeping only the top 100 entries per source phrase, according to $p(t|s)$.
-All model files were then binarized; the language models were binarized using KenLM~\cite{Heafield-kenlm}, the phrase table using the Probing phrase-table, lexicalized re-ordering model using the compact data structure.
+All model files were then binarized; the language models were binarized using KenLM~\citep{Heafield-kenlm}, the phrase table using the Probing phrase-table, lexicalized re-ordering model using the compact data structure.
 These binary formats were chosen for their best-in-class multithreaded performance.
@@
 
-Standard Moses phrase-based configurations are used, except that we use the cube-pruning algorithm~\cite{Chiang:2007:cl} with a pop-limit of 400\footnote{the pop-limit was chosen from public discussion on the Moses mailing list on an acceptable balance between decoding speed and translation quality with Moses for commercial use %https://www.mail-archive.com/moses-support\%40mit.edu/msg13181.html
+Standard Moses phrase-based configurations are used, except that we use the cube-pruning algorithm~\citep{Chiang:2007:cl} with a pop-limit of 400\footnote{the pop-limit was chosen from public discussion on the Moses mailing list on an acceptable balance between decoding speed and translation quality with Moses for commercial use %https://www.mail-archive.com/moses-support\%40mit.edu/msg13181.html
 }, rather than the basic phrase-based algorithm.
@@
 It is also slower but the time/quality tradeoff is better overall with this stack configuration.
-Model scores for coverage and cardinality configurations are identical, which differs from the results in ~\citep{ortizmartinez-garciavarea-casacuberta:2006:WMT}.
+Model scores for coverage and cardinality configurations are identical, which differs from the results in ~\cite{ortizmartinez-garciavarea-casacuberta:2006:WMT}.
 This may be due to differing test set characteristics, or interactions with cube-pruning algorithm we used.
@@
 
-%We mimic the use of multi-core servers by single-threaded decoders such as mtplz~\cite{Heafield-mtplz} and Jane~\cite{peitz2012jane} by running multiple processes of the Moses decoder, each running one thread.
+%We mimic the use of multi-core servers by single-threaded decoders such as mtplz~\citep{Heafield-mtplz} and Jane~\citep{peitz2012jane} by running multiple processes of the Moses decoder, each running one thread.
 The input is distributed to the processes using a external script.

### 2016-06-10 acl.2016/amta2016.tex
authors

@@
 % \mtsummitHeader{x}{x}{xxx-xxx}{2016}{45-character paper description goes here}{Author(s) initials and last name go here}
-\title{\bf Formatting Your Paper \\
-  for the \confname~Conference}
+\title{\bf Fast, Scalable Phrase-Based SMT Decoding}
 
-\author{\name{\bf First Author} \hfill  \addr{author1@abc.university.country}\\ 
-        \name{\bf Second Author} \hfill \addr{author2@abc.university.country}\\ 
-        \addr{Department of Science, My University, 
-        MyTown, Zip, Country}
+\author{\name{\bf Hieu Hoang} \hfill  \addr{hieu@moses-mt.org}\\ 
+        \addr{Moses Machine Translation CIC, UK}
 \AND
-       \name{\bf Third Author} \hfill \addr{author3@abc.university.country}\\
-        \addr{Department of Science, My University, 
-        MyTown, Zip, Country}
+       \name{\bf Nicolay Bogoychev} \hfill \addr{s1031254@sms.ed.ac.uk}\\
+       \name{\bf Kenneth Heafield} \hfill \addr{kheafiel@inf.ed.ac.uk}\\
+        \addr{University of Edinburgh, Scotland}
+\AND
+       \name{\bf Lane Schwartz} \hfill \addr{lanes@illinois.edu}\\
+        \addr{University of Illinois, USA}
+\AND
+       \name{\bf Marcin Junczys-Dowmunt} \hfill \addr{junczys@amu.edu.pl}\\
+        \addr{Adam Mickiewicz University}
 }

### 2016-06-10 acl.2016/amta2016.tex
authors

--- 2016-06-10 acl.2016/amta2016.tex
@@
 % \mtsummitHeader{x}{x}{xxx-xxx}{2016}{45-character paper description goes here}{Author(s) initials and last name go here}
-\title{\bf Formatting Your Paper \\
-  for the \confname~Conference}
+\title{\bf Fast, Scalable Phrase-Based SMT Decoding}
 
-\author{\name{\bf First Author} \hfill  \addr{author1@abc.university.country}\\ 
-        \name{\bf Second Author} \hfill \addr{author2@abc.university.country}\\ 
-        \addr{Department of Science, My University, 
-        MyTown, Zip, Country}
-\AND
-       \name{\bf Third Author} \hfill \addr{author3@abc.university.country}\\
-        \addr{Department of Science, My University, 
-        MyTown, Zip, Country}
-}
+%\author{\name{\bf Hieu Hoang} \hfill  \addr{hieu@moses-mt.org}\\ 
+%        \addr{Moses Machine Translation CIC, UK}
+%\AND
+%       \name{\bf Nicolay Bogoychev} \hfill \addr{s1031254@sms.ed.ac.uk}\\
+%       \name{\bf Kenneth Heafield} \hfill \addr{kheafiel@inf.ed.ac.uk}\\
+%        \addr{University of Edinburgh, Scotland}
+%\AND
+%       \name{\bf Lane Schwartz} \hfill \addr{lanes@illinois.edu}\\
+%        \addr{University of Illinois, USA}
+%\AND
+%       \name{\bf Marcin Junczys-Dowmunt} \hfill \addr{junczys@amu.edu.pl}\\
+%        \addr{Adam Mickiewicz University}
+%}
 

### 2016-06-10 acl.2016/amta2016.tex
whatever

@@
 Figure~\ref{fig:mempool} compares the decoding speed for Moses and our decoder, using the same models, parameters and test set.
-Our decoder is over 3 times faster with one thread, and 4.7 times faster using all cores.
-Like Moses, performance actually worsens after approximately 15 threads, however, the problem is not as pronounced.
+Our decoder is over 3 times faster with one thread, and 5.0 times faster using all cores.
+However, like Moses, performance actually worsens after approximately 15 threads.
+%, however, the problem is not as pronounced.
 This gives us a better foundation on which to build further innovations for fast, multi-core decoding.
@@
 \end{tabular}
-\caption{Decoding time of Moses and our decoder, using the same models}
+\caption{Decoding speed of Moses and our decoder, using the same models}
 \label{fig:mempool}

### 2016-06-10 acl.2016/amta2016.tex
whatever

@@
 The search algorithm creates and destroy a large number of intermediate objects such as hypotheses and feature function states.
-The memory management this requires puts a burden on the operating system due to the need to synchronize memory access.
-This problem worsens with multi-threaded applications such as Moses.
-
+This puts a burden on the operating system due to the need to synchronize memory access, especially when using a large number of threads.
 Libraries such as tcmalloc~\citep{ghemawat2009tcmalloc} are designed to reduce locking contention for multi-threaded application but in our case, this is still not enough.
@@
 A memory pool is a large block of memory that has been given to the application by the operating system.
-The application is then responsible for allocating portions of this memory to its components as required.
+The application is then responsible for allocating portions of this memory to its components when requested.
 We will use thread-specific memory pools to increase speed by avoiding locking contention during memory access.
-Our memory pools will be dynamic, growing in size when required but never reducing in size.
-The pools are not deleted until the decoder application ends.
+Our memory pools will be dynamic.
+That is, the memory requirement does not have to be known or specified before running the application, the pool can grow when required but they will never reduce in size.
+The pools are deleted only when the application ends.
 
@@
 Unused data structures accumulates in the pool until a reset event.
-After the event, the pool is assumed to be empty and simply reused.
-
+The pool is assumed to be empty and simply reused after the event.
 We instantiate two memory pools per decoding thread, one which is never reset and another which is reset after the decoding of each sentence.

### 2016-06-10 acl.2016/amta2016.tex
whatever

@@
 
-\section{Other Models and Even More Cores}
+\subsection{Other Models and Even More Cores}
 
@@
 
+\subsection{Translation Quality}
+
+Our work is intended as a drop-in replacement for the Moses decoder.
+Given the same parameters and models, the model scores are comparable and the search space should be identical (or very nearly identical).
+This should lead to almost translation quality, which we verify with Figure~\ref{???}.
+
 \section{Conclusion}

### 2016-06-10 acl.2016/amta2016.tex
whatever

@@
 Figure~\ref{fig:mempool} compares the decoding speed for Moses and our decoder, using the same models, parameters and test set.
-Our decoder is over 3 times faster with one thread, and 5.0 times faster using all cores.
-However, like Moses, performance actually worsens after approximately 15 threads.
+Our decoder is 4.4 times faster with one thread, and 5.0 times faster using all cores.
+Like Moses, however, performance actually worsens after approximately 15 threads.
 %, however, the problem is not as pronounced.
@@
 \end{tabular}
-\caption{Decoding speed for with bigger servers}
+\caption{Decoding speed with more cores}
 \label{fig:more-cores}

### 2016-06-10 acl.2016/amta2016.tex
whatever

@@
 \end{tabular}
-\caption{Decoding time with Compact Lexicalized Re-ordering, and integrated into a model the phrase-table}
+\caption{Decoding speed with Compact Lexicalized Re-ordering, and integrated into a model the phrase-table}
 \label{fig:lex-ro}
 \end{figure} 
-Decoding time decreased by 29\% with running with a single thread but it is over 5 times faster using 32 threads.
+Decoding speed increased by 40\% when using one thread but is 5 times faster when using 32 threads.
 %In fact, the decoding time with the integrated model is similar to that \emph{without} a lexicalized re-ordering model\footnote{Strangely, using lexicalized re-ordering model results in slightly faster decoding than without the model.
@@
 Given the same parameters and models, the model scores are comparable and the search space should be identical (or very nearly identical).
-This should lead to almost translation quality, which we verify with Figure~\ref{???}.
+This should lead to almost identical translation quality, which we verify with Figure~\ref{fig:bleu}.
+\begin{figure}[h]
+\centering
+\begin{tabular}{cc}
+{\includegraphics[scale=0.5]{bleu.png}} 
+\end{tabular}
+\caption{Translation quality (BLEU) for different pop-limits}
+\label{fig:bleu}
+\end{figure}
 

### 2016-06-11 acl.2016/amta2016.tex
whatever

@@
 
-Also, stack pruning occurs on each hypothesis group independently, therefore, potentially affect search errors and model scores.
+Also, stack pruning occurs on each hypothesis group independently, therefore, potentially affecting search errors and model scores.
 
 %Table~\ref{tab:stack-configuration} and 
-Figure~\ref{fig:stack-configuration} present the tradeoff between decoding time and average model at various pop-limits.
+Figure~\ref{fig:stack-configuration} present the tradeoff between decoding time and average model, created by varying the cube-pruning pop-limit.
 The model scores for all stack configurations are identical for low pop-limits but grouping hypotheses into coverage \& end position produces higher model scores for higher pop-limits.
@@
 % 
+We verified that the translation quality of our decoder is comparable to that of Moses in Figure~\ref{fig:bleu}.
+This fits in with our intention of creating a drop-in replacement for the Moses decoder.
+\begin{figure}[h]
+\centering
+\begin{tabular}{cc}
+{\includegraphics[scale=0.5]{bleu.png}} 
+\end{tabular}
+\caption{Translation quality (BLEU) for different pop-limits}
+\label{fig:bleu}
+\end{figure}
 
@@
 
-This contrast with Moses where speed increases to approximately 16 threads but then actually become slower thereafter.
+This contrast with Moses where speed increases up to approximately 16 threads but then become slower thereafter.
 Using the tcmalloc library has a small positive effect on decoding speed but does little to improve scalability
@@
 
-\subsection{Translation Quality}
-
-Our work is intended as a drop-in replacement for the Moses decoder.
-Given the same parameters and models, the model scores are comparable and the search space should be identical (or very nearly identical).
-This should lead to almost identical translation quality, which we verify with Figure~\ref{fig:bleu}.
-\begin{figure}[h]
-\centering
-\begin{tabular}{cc}
-{\includegraphics[scale=0.5]{bleu.png}} 
-\end{tabular}
-\caption{Translation quality (BLEU) for different pop-limits}
-\label{fig:bleu}
-\end{figure}
-
 \section{Conclusion}

### 2016-06-11 acl.2016/amta2016.tex
whatever

@@
 % 
-We verified that the translation quality of our decoder is comparable to that of Moses in Figure~\ref{fig:bleu}.
+We verified that the translation quality of our decoder is comparable to that of Moses in Figure~\ref{fig:bleu}, given the same parameters and models.
 This fits in with our intention of creating a drop-in replacement for the Moses decoder.
@@
 \end{tabular}
-\caption{Translation quality (BLEU) for different pop-limits}
+\caption{Translation quality for different pop-limits}
 \label{fig:bleu}

### 2016-09-10 amta.2016/amta2016.tex
move to amta.2016

--- 2016-06-11 acl.2016/amta2016.tex
@@
 This paper is the first we know of that focuses on improving decoding speed on multicore servers.
-We will take a holistic approach to solving this issue.
-We create a decoder that is optimized for multi-core processing speed by concentrating on four main areas.
+We take a holistic approach to solving this issue, creating a decoder that is optimized for multi-core processing speed by concentrating on four main areas:
 \begin{enumerate}

### 2016-09-10 amta.2016/amta2016.tex
move to amta.2016

@@
 
-%\author{\name{\bf Hieu Hoang} \hfill  \addr{hieu@moses-mt.org}\\ 
-%        \addr{Moses Machine Translation CIC, UK}
-%\AND
-%       \name{\bf Nicolay Bogoychev} \hfill \addr{s1031254@sms.ed.ac.uk}\\
-%       \name{\bf Kenneth Heafield} \hfill \addr{kheafiel@inf.ed.ac.uk}\\
-%        \addr{University of Edinburgh, Scotland}
-%\AND
-%       \name{\bf Lane Schwartz} \hfill \addr{lanes@illinois.edu}\\
-%        \addr{University of Illinois, USA}
-%\AND
-%       \name{\bf Marcin Junczys-Dowmunt} \hfill \addr{junczys@amu.edu.pl}\\
-%        \addr{Adam Mickiewicz University}
-%}
+\author{\name{\bf Hieu Hoang} \hfill  \addr{hieu@moses-mt.org}\\ 
+        \addr{Moses Machine Translation CIC, UK}
+\AND
+       \name{\bf Nicolay Bogoychev} \hfill \addr{s1031254@sms.ed.ac.uk}\\
+       %\name{\bf Kenneth Heafield} \hfill \addr{kheafiel@inf.ed.ac.uk}\\
+        \addr{University of Edinburgh, Scotland}
+\AND
+       \name{\bf Lane Schwartz} \hfill \addr{lanes@illinois.edu}\\
+        \addr{University of Illinois, USA}
+\AND
+       \name{\bf Marcin Junczys-Dowmunt} \hfill \addr{junczys@amu.edu.pl}\\
+        \addr{Adam Mickiewicz University}
+}
 
@@
 \end{enumerate}
+
 The result is a decoder that is significantly faster than the Moses baseline for single-threaded operation, and scales with the number of cores.
 
+We will maintain the Moses decoder's embarrassingly parallel, one sentence-per-thread decoding framework.
 As far as possible, model scores and functionality are compatible with Moses to aid comparison and ease transition for existing users.
-All source code will be made available under an open-source license.
+The source code is available in the existing Moses repository\footnote{\scriptsize https://github.com/moses-smt/mosesdecoder/tree/master/contrib/moses2}.
 

### 2016-09-10 amta.2016/amta2016.tex
move to amta.2016
delete all [h]

@@
 
-~\cite{mfernandez2016boosting} was concerned with multi-core speed but treated the decoding process as a black box within a parallelization framework.
+~\cite{mfernandez2016boosting} was concerned with multi-core speed but treated decoding as a black box within a parallelization framework.
 
 There are a number of phrase-based decoding implementations, many of which implements the extensions to the phrase-based model described above.
-
-The most well known is Moses~\citep{koehn-EtAl:2007:PosterDemo} which implements a number of speed optimizations, including cube-pruning. It is widely used for MT research and commercial use.
+The most well known is Moses~\citep{koehn-EtAl:2007:PosterDemo} which implements a number of speed optimizations, including cube-pruning.
+It is widely used for MT research and commercial use.
 
@@
 This paper not only focuses on faster single-threaded decoding but also on overcoming the shortcomings of existing decoding implementations on multicore servers.
+We will compare multicore performance the best-of-breed phrase-table described in ~\cite{junczys_tsd_2012b} with our own implementation.
+We will use the cube-pruning algorithm, however, the standard phrase-based decoding algorithm is also available.
 
@@
 Figure~\ref{fig:moses-phrase-tables-time} shows the decoding speed for the fastest two implementations.
-\begin{figure}[h]
+\begin{figure}
 \centering
@@
 The result is \emph{slower} decoding, Table~\ref{tab:phrase-table-cache}.
-\begin{table}[h]
+\begin{table}
 \small
@@
 For testing decoding speed, we used a subset of the training data, Table~\ref{tab:test-sets}.
-\begin{table}[h]
+\begin{table}
 \begin{center}
@@
 Language model 	& 3.1  	& 1.8 \\ 
-Lex re. model	& 2.3	& 637MB \\ \hline
+Lex re. model	& 2.3	& 0.6 \\ \hline
 \end{tabular}
@@
 
-\begin{table}[h]
+\begin{table}
 \begin{center}
@@
 By contrast, our decoder spends 11\% on memory management and does not significantly increase with more threads.
-\begin{table}[h]
+\begin{table}
 \begin{center}
@@
 This gives us a better foundation on which to build further innovations for fast, multi-core decoding.
-\begin{figure}[h]
+\begin{figure}
 \centering
@@
 
-\begin{figure}[h]
+\begin{figure}
 \centering
@@
 This fits in with our intention of creating a drop-in replacement for the Moses decoder.
-\begin{figure}[h]
+\begin{figure}
 \centering
@@
 It does not require the overhead of managing an active cache but there is still some overhead in using a cache.
-\begin{table}[h]
+\begin{table}
 \small
@@
 This increase the size of the binary files from 17GB to 23GB but the time saved not needing to decompress the data resulted in a 1.5\% decrease in decoding time with 1 thread and nearly 7\% when the CPUs are saturated, Table~\ref{tab:compression}.
-\begin{table}[h]
+\begin{table}
 \small
@@
 This resulted in a significant decrease in decoding time, especially with high number of cores, Figure~\ref{fig:lex-ro}. 
-\begin{figure}[h]
+\begin{figure}
 \centering
@@
 Our decoder show no scalability issues when we tested with the same model and tested set on a larger server, Figure~\ref{fig:more-cores}.
-\begin{figure}[h]
+\begin{figure}
 \centering
@@
 Our decoder is 5.4 times faster than Moses with a single-thread and 14.5 faster when all cores are saturated.
-\begin{figure}[h]
+\begin{figure}
 \centering
@@
 Results show that our decoder is 7 times faster than Moses and still scales monotonically until all CPUs are saturated, Figure~\ref{fig:large-lm}.
-\begin{figure}[h]
+\begin{figure}
 \centering

### 2016-09-10 amta.2016/amta2016.tex
more

@@
 This paper not only focuses on faster single-threaded decoding but also on overcoming the shortcomings of existing decoding implementations on multicore servers.
+Unlike ~\cite{mfernandez2016boosting}, we will optimize decoding speed by looking inside the black box.
 We will compare multicore performance the best-of-breed phrase-table described in ~\cite{junczys_tsd_2012b} with our own implementation.
-We will use the cube-pruning algorithm, however, the standard phrase-based decoding algorithm is also available.
+We will use the cube-pruning algorithm, however, the standard phrase-based decoding algorithm is also available and a framework exists to accommodate other decoding algorithms in future.
+We use KenLM~\citep{Heafield-kenlm} due to it's popularity and consistent performance, but as with Moses, other language model implementations can be added later.
 
@@
 The need for random lookup during querying inevitably results in slower decoding speed.
-Previous work such as~\cite{junczys_tsd_2012b} improve querying speed with more efficient data structures.
+Previous work such as~\cite{junczys_tsd_2012b} improve querying speed with more compact data structures.
 
@@
 The phrase-table was then pruned, keeping only the top 100 entries per source phrase, according to $p(t|s)$.
-All model files were then binarized; the language models were binarized using KenLM~\citep{Heafield-kenlm}, the phrase table using the Probing phrase-table, lexicalized re-ordering model using the compact data structure.
+All model files were then binarized; the language models were binarized using KenLM~\citep{Heafield-kenlm}, the phrase table using the Probing phrase-table, lexicalized re-ordering model using the compact data structure~\citep{junczys_tsd_2012b}.
 These binary formats were chosen for their best-in-class multithreaded performance.

### 2016-09-10 amta.2016/amta2016.tex
more

@@
 Phrase table  	& 17 	& 5.8 \\
-Language model 	& 3.1  	& 1.8 \\ 
+Language model (5-gram) 	& 3.1  	& 1.8 \\ 
 Lex re. model	& 2.3	& 0.6 \\ \hline
@@
 
-Each sentence is decoded from start to finish using one worker thread. %The CPU affinity of each thread is set to a specific core to minimize memory page faults as a result of threads switching cores.
+%Each sentence is decoded from start to finish using one worker thread.
+%The CPU affinity of each thread is set to a specific core to minimize memory page faults as a result of threads switching cores.
 
-As can be seen in Table~\ref{tab:optimizing-memory}, over 24\% of the Moses decoder running time is spent on memory management and this increases to 39\% when 32 threads are used, dampening the scalability of the decoder.
+Over 24\% of the Moses decoder running time is spent on memory management, Table~\ref{tab:optimizing-memory}.
+This increases to 39\% when 32 threads are used, dampening the scalability of the decoder.
 By contrast, our decoder spends 11\% on memory management and does not significantly increase with more threads.
@@
 
-We shall investigate the effects of the following three stack configurations on model score and decoding speed:
+We investigated the effects of the following three stack configurations on model score and decoding speed:
 \begin{enumerate}
@@
 \end{figure}
-
-Also, stack pruning occurs on each hypothesis group independently, therefore, potentially affecting search errors and model scores.
+However, stack pruning occurs on each hypothesis group independently, therefore, potentially affecting search errors and model scores.
 
@@
 Figure~\ref{fig:stack-configuration} present the tradeoff between decoding time and average model, created by varying the cube-pruning pop-limit.
-The model scores for all stack configurations are identical for low pop-limits but grouping hypotheses into coverage \& end position produces higher model scores for higher pop-limits.
+None of the different stack configurations significantly outperform the others in either quality or decoding speed.
+However, the coverage \& end position produces slightly higher model scores at higher pop-limits, therefore, we continue to use this configuration throughout the rest of this paper.
+
+%The model scores for all stack configurations are identical for low pop-limits but grouping hypotheses into coverage \& end position produces higher model scores for higher pop-limits.
 It is also slower but the time/quality tradeoff is better overall with this stack configuration.
@@
 This may be due to differing test set characteristics, or interactions with cube-pruning algorithm we used.
+
 %\begin{table*}[t]

### 2016-09-10 amta.2016/amta2016.tex
more

@@
 While useful, these prior works present only one particular stack configuration each.
-
 ~\cite{ortizmartinez-garciavarea-casacuberta:2006:WMT} explore a range of stack configurations by defining a granularity parameter which controls the maximum number of stacks required to decode a sentence.
@@
 \end{enumerate}
-Coverage cardinality is the same as that in Moses and Joshua. Coverage configuration uses one stack per unique coverage vector.
+Coverage cardinality is the same as that in Moses and Joshua.
+Coverage configuration uses one stack per unique coverage vector.
+\emph{Coverage and end position} of most recently translated source word extends the coverage configuration by separating hypotheses where the position of the last translate word are different, even if the coverages are identical.
 
-\emph{Coverage and end position of most recently translated source word} extends the coverage configuration by separating hypotheses where the last translate word are different.
 This is an optimization to reduce the number of checks on the distortion limit, which is dependent on the last word position.
+The check is a binary function $d(C_{h}, e_{hypo}, range_r)$, where $C_{h}$ is the coverage vector of hypothesis $h$, $e_{h}$ is the end position of most recent source word that has been translated, and $range_r$ is the coverage of the rule to be applied.
+%Figure~\ref{algo:Hypothesis Expansion with Unsorted Stack} shows the standard hypothesis expansion algorithm.
+%\begin{figure} [h]
+%\begin{algorithmic}
+%%\REQUIRE source sentence $s$, translation options
+%\FORALL{ $h$ in $stack_{|C|}$} 
+%  \FORALL{$r$ in translation rules}
+%      \IF {$d(C_{h}, e_{hypo}, range_r)$}
+%	\STATE expand $h$ with $r$ $\rightarrow$ $h'$
+%	\STATE add $h'$ to stack $stack_{C'}$
+%      \ENDIF
+%  \ENDFOR % { translation options}
+%\ENDFOR % { hypothesis in $stack_i$} 
+%\end{algorithmic}
+%\caption{Hypothesis Expansion with Cardinality Stacks}
+%\label{algo:Hypothesis Expansion with Unsorted Stack}
+%\end{figure}
 
-The check is a binary function $d(C_{h}, e_{hypo}, range_r)$, where $C_{h}$ is the coverage vector of hypothesis $h$, $e_{h}$ is the end position of most recent source word that has been translated, and $range_r$ is the coverage of the rule to be applied.
-Figure~\ref{algo:Hypothesis Expansion with Unsorted Stack} shows the standard hypothesis expansion algorithm.
-\begin{figure} [h]
-\begin{algorithmic}
-%\REQUIRE source sentence $s$, translation options
-\FORALL{ $h$ in $stack_{|C|}$} 
-  \FORALL{$r$ in translation rules}
-      \IF {$d(C_{h}, e_{hypo}, range_r)$}
-	\STATE expand $h$ with $r$ $\rightarrow$ $h'$
-	\STATE add $h'$ to stack $stack_{C'}$
-      \ENDIF
-  \ENDFOR % { translation options}
-\ENDFOR % { hypothesis in $stack_i$} 
-\end{algorithmic}
-\caption{Hypothesis Expansion with Cardinality Stacks}
-\label{algo:Hypothesis Expansion with Unsorted Stack}
-\end{figure}
-
-By grouping hypotheses according to coverage \emph{and} end position into groups we call \emph{'ministack'}, the distortion limit only needs to be checked for each group, Figure~\ref{algo:Hypothesis Expansion with Sorted Stack}.
-\begin{figure} [h]
-\begin{algorithmic}
-%\REQUIRE source sentence $s$, translation options
-\FORALL{ $ministack_{C, e}$ in $stack_{|C|}$} 
-  \FORALL{$r$ in translation rules}
-    \IF {$d(C_{h}, e_{hypo}, range_r)$}
-      \FORALL{ $h$ in $ministack_{C, e}$} 
-	\STATE expand $h$ with $r$ $\rightarrow$ $h'$
-	\STATE add $h'$ to $ministack_{C',e'}$
-      \ENDFOR
-    \ENDIF
-  \ENDFOR % { translation options}
-\ENDFOR % { hypothesis in $stack_i$} 
-\end{algorithmic}
-\caption{Hypothesis Expansion with Coverage \& End Position Stacks}
-\label{algo:Hypothesis Expansion with Sorted Stack}
-\end{figure}
+By grouping hypotheses according to coverage \emph{and} end position, the distortion limit only needs to be checked for each group. %, Figure~\ref{algo:Hypothesis Expansion with Sorted Stack}. 
+%\begin{figure} [h]
+%\begin{algorithmic}
+%%\REQUIRE source sentence $s$, translation options
+%\FORALL{ $ministack_{C, e}$ in $stack_{|C|}$} 
+%  \FORALL{$r$ in translation rules}
+%    \IF {$d(C_{h}, e_{hypo}, range_r)$}
+%      \FORALL{ $h$ in $ministack_{C, e}$} 
+%	\STATE expand $h$ with $r$ $\rightarrow$ $h'$
+%	\STATE add $h'$ to $ministack_{C',e'}$
+%      \ENDFOR
+%    \ENDIF
+%  \ENDFOR % { translation options}
+%\ENDFOR % { hypothesis in $stack_i$} 
+%\end{algorithmic}
+%\caption{Hypothesis Expansion with Coverage \& End Position Stacks}
+%\label{algo:Hypothesis Expansion with Sorted Stack}
+%\end{figure}
 However, stack pruning occurs on each hypothesis group independently, therefore, potentially affecting search errors and model scores.

### 2016-09-10 amta.2016/amta2016.tex
ack

@@
 
-% \section*{Acknowledgments}
-% 
-% Do not number the acknowledgment section.
-% This section should not be presented for the submission version.
-
+ \section*{Acknowledgments}
+This work is sponsored by the Air Force Research Laboratory, prime contract FA8650-11-C-6160.
+The views and conclusions contained in this document are those of the authors and should not be interpreted as representative of the official policies, either expressed or implied, of the Air Force Research Laboratory or the U.S.
+Government.
 \small

### 2016-09-10 amta.2016/amta2016.tex
Minor fixes, mostly typos and grammatical errors

@@
 For many users, decoding is the most time-critical part of the translation process.
-Making use of the multiple cores that are now ubiquitous in todays servers is a common strategy to ameliorate this issue.
+Making use of the multiple cores that are now ubiquitous in today's servers is a common strategy to ameliorate this issue.
 However, it has been noticed that the Moses decoder, amongst others, is unable to efficiently use multiple cores~\citep{mfernandez2016boosting}.
 That is, decoding speed does not substantially increase when more cores are used, in fact, it may actually \emph{decrease} when using more cores.
-There have been speculation on the causes of the inefficiency as well as potential remedies.
+There has been speculation on the causes of the inefficiency as well as potential remedies.
 
@@
 
-~\cite{Heafield-kenlm} and ~\cite{yasuhara-EtAl:2013:EMNLP} describes fast, efficient datastructures for language models.
+~\cite{Heafield-kenlm} and ~\cite{yasuhara-EtAl:2013:EMNLP} describe fast, efficient data structures for language models.
 ~\cite{zens2007efficient} describes an implementation of a phrase-table for an SMT decoder that is loaded on demand, reducing the initial loading time and memory requirements.
@@
 We will use the cube-pruning algorithm, however, the standard phrase-based decoding algorithm is also available and a framework exists to accommodate other decoding algorithms in future.
-We use KenLM~\citep{Heafield-kenlm} due to it's popularity and consistent performance, but as with Moses, other language model implementations can be added later.
+We use KenLM~\citep{Heafield-kenlm} due to its popularity and consistent performance, but as with Moses, other language model implementations can be added later.
 
@@
 Successive applications of translation rules do not have to be adjacent on the source side, depending on the distortion limit.
-The target output is constructed strictly left-to-right from the target side from the series of translation rules.
+The target output is constructed strictly left-to-right from the target side using this series of translation rules.
 
@@
 Accumulating unused objects in the memory pools can result in unacceptably high memory usage so object queues are available for high-churn objects which allows the decoder to re-cycle unused objects before the reset event.
-This also increase speed as LIFO queues are used so that the most recently accessed memory are used, increasing CPU cache hits.
+This also increases speed as LIFO queues are used so that the most recently accessed memory are used, increasing CPU cache hits.
 
@@
 Rather than storing the lexicalized re-ordering model separately, we shall integrating it into the translation model, eliminating the need to query a separate file.
-However, the model remain the same under the log-linear framework, including having its own weights.
+However, the model remains the same under the log-linear framework, including having its own weights.
 %The search algorithm will produce identical results to the Moses baseline apart from edge case training configurations, for example, where the lexicalized reordering model is trained on a different dataset than the translation model.
@@
 
-Over 24\% of the Moses decoder running time is spent on memory management, Table~\ref{tab:optimizing-memory}.
+Over 24\% of the Moses decoder running time is spent on memory management (Table~\ref{tab:optimizing-memory}).
 This increases to 39\% when 32 threads are used, dampening the scalability of the decoder.
@@
 %Table~\ref{tab:stack-configuration} and 
-Figure~\ref{fig:stack-configuration} present the tradeoff between decoding time and average model, created by varying the cube-pruning pop-limit.
+Figure~\ref{fig:stack-configuration} presents the tradeoff between decoding time and average model scores, created by varying the cube-pruning pop-limit.
 None of the different stack configurations significantly outperform the others in either quality or decoding speed.
@@
 \end{tabular}
-\caption{Decoding speed with more cores}
+\caption{Decoding speed of our decoder with more cores}
 \label{fig:more-cores}
@@
 This resulted in double digit gains compared to Moses on the same hardware.
-Our implementation is also unaffected by scalability issues that has afflicted Moses.
+Our implementation is also unaffected by scalability issues that have afflicted Moses.
 % and continue to scale well for all possible cores we have tested it on.

### 2016-10-18 amta.2016/acl2016.tex
move to amta.2016
spelt nikolay name wrong

--- 2016-06-10 acl.2016/acl2016.tex
@@
 	  \And
-	Nicolay Bogoychev\\
+	Nikolay Bogoychev\\
   	University of Edinburgh\\
@@
 As a baseline, we use a recent\footnote{The experiments were performed between January and May 2016 with the latest github code to hand.
-The main ar-en experiments were rerun with the source code as at 8th June, 2016 to ensure there were no material difference.
+The main ar-en experiments were rerun with the source code as of 8th June, 2016 to ensure there were no material difference.
 The commit hash was \scriptsize{bc5f8d15c6ce4bc678ba992860bfd4be6719cee8} } version of the Moses decoder taken from the github repository.

### 2018-01-09 fast-beam/fast-beam.tex
start fast-beam paper

--- 2016-09-10 amta.2016/amta2016.tex
@@
 
-\author{\name{\bf Hieu Hoang} \hfill  \addr{hieu@moses-mt.org}\\ 
+\author{\name{\bf Hieu Hoang} \hfill  \addr{hieu@hoang.co.uk}\\ 
         \addr{Moses Machine Translation CIC, UK}
 \AND
-       \name{\bf Nicolay Bogoychev} \hfill \addr{s1031254@sms.ed.ac.uk}\\
+       \name{\bf Nikolay Bogoychev} \hfill \addr{s1031254@sms.ed.ac.uk}\\
        %\name{\bf Kenneth Heafield} \hfill \addr{kheafiel@inf.ed.ac.uk}\\
@@
 
-SMT has steadily progressed from a research discipline to commercial viability during the past decade as can be seen from services such as the Google and Microsoft Translation services.
-As well as general purpose services such as these, there is a large number of companies that offer customized translation systems, as well as  companies and organization that implement in-house solutions.
-Many of these customized solutions use Moses as their SMT engine.
-
-For many users, decoding is the most time-critical part of the translation process.
-Making use of the multiple cores that are now ubiquitous in today's servers is a common strategy to ameliorate this issue.
-However, it has been noticed that the Moses decoder, amongst others, is unable to efficiently use multiple cores~\citep{mfernandez2016boosting}.
-That is, decoding speed does not substantially increase when more cores are used, in fact, it may actually \emph{decrease} when using more cores.
-There has been speculation on the causes of the inefficiency as well as potential remedies.
-
-This paper is the first we know of that focuses on improving decoding speed on multicore servers.
-We take a holistic approach to solving this issue, creating a decoder that is optimized for multi-core processing speed by concentrating on four main areas:
-\begin{enumerate}
-  \item \vspace{-2 mm} Faster memory management of data-structures through the use of customized memory pools
-  \item \vspace{-2 mm} Exploring alternatives to cardinality-based hypothesis stack configuration
-  \item \vspace{-2 mm} Re-examining the efficiency of phrase-table lookup using translation rule caching and data compression
-  \item \vspace{-2 mm} Integrating the lexicalized re-ordering model into the phrase-table, thus eliminating the need for independent random lookup this model 
-\end{enumerate}
-
-The result is a decoder that is significantly faster than the Moses baseline for single-threaded operation, and scales with the number of cores.
-
-We will maintain the Moses decoder's embarrassingly parallel, one sentence-per-thread decoding framework.
-As far as possible, model scores and functionality are compatible with Moses to aid comparison and ease transition for existing users.
-The source code is available in the existing Moses repository\footnote{\scriptsize https://github.com/moses-smt/mosesdecoder/tree/master/contrib/moses2}.
-
-The rest of the paper will be broken up into the following sections.
-The rest of this section will discuss prior work and an outline of the phrase-based model.
-Section~\ref{sec:Proposed Improvements} will then describe the modifications to improve decoding speed.
-We describe the experiment setup in Section~\ref{sec:Experimental Setup} and present results Section~\ref{sec:Results}.
-We conclude in the last section and discuss possible future work.
-
 \subsection{Prior Work}
 
-Most prior work on increasing decoding speed look to optimizing specific components of the decoder or the decoding algorithm.
-
-~\cite{Heafield-kenlm} and ~\cite{yasuhara-EtAl:2013:EMNLP} describe fast, efficient data structures for language models.
-~\cite{zens2007efficient} describes an implementation of a phrase-table for an SMT decoder that is loaded on demand, reducing the initial loading time and memory requirements.
-~\cite{junczys_tsd_2012b} extends this by compressing the on-disk phrase table and lexicalized re-ordering model.
-
-~\cite{Chiang:2007:cl} describes the cube-pruning and cube-growing algorithm which allows the tradeoff between speed and translation quality to the adjusted with a single parameter.
-~\cite{wuebker2012fast} note that language model querying is amongst the most expensive operation in decoding.
-They sought to improved decoding speed by caching score computations early pruning of translation options.
-This work is similar to ~\cite{Heafield-mtplz} which group hypotheses with identical language model context and incrementally expand them, reducing LM querying.
-
-~\cite{mfernandez2016boosting} was concerned with multi-core speed but treated decoding as a black box within a parallelization framework.
-
-There are a number of phrase-based decoding implementations, many of which implements the extensions to the phrase-based model described above.
-The most well known is Moses~\citep{koehn-EtAl:2007:PosterDemo} which implements a number of speed optimizations, including cube-pruning.
-It is widely used for MT research and commercial use.
-
-Joshua~\citep{Joshua-Decoder} also supports cube-pruning for phrase-based models.
-Phrasal~\citep{spence2014phrasal} supports a number of variants of the phrase-based model.
-Jane~\citep{peitz2012jane} supports the language model look-ahead described in ~\cite{wuebker2012fast} in addition to other tools to speed up decoding such as having a separate fast, lightweight decoder.
-mtplz is a specialized decoder developed to implement the incremental decoding described in ~\cite{Heafield-mtplz}.
-
-The Moses, Joshua and Phrasal decoders implement multithreading, however, they all report scalability problems, either in the paper (Phrasal) or via social media (Moses\footnote{\scriptsize https://github.com/moses-smt/mosesdecoder/issues/39} and Joshua\footnote{\scriptsize https://twitter.com/ApacheJoshua/status/342022794097340416}).
-
-Jane and mtplz are single-threaded decoders, relying on external applications to parallelize operations.
-%This has the advantage it is easier to parallelize processing across multiple servers, not just multiple cores.
-As we shall see, this method does not scale beyond a small number of cores when attempting to scale to multiple cores within one server.
-
-This paper not only focuses on faster single-threaded decoding but also on overcoming the shortcomings of existing decoding implementations on multicore servers.
-Unlike ~\cite{mfernandez2016boosting}, we will optimize decoding speed by looking inside the black box.
-We will compare multicore performance the best-of-breed phrase-table described in ~\cite{junczys_tsd_2012b} with our own implementation.
-We will use the cube-pruning algorithm, however, the standard phrase-based decoding algorithm is also available and a framework exists to accommodate other decoding algorithms in future.
-We use KenLM~\citep{Heafield-kenlm} due to its popularity and consistent performance, but as with Moses, other language model implementations can be added later.
-
-\subsection{Phrase-Based Model}
-
-The objective of decoding is to find the target translation with the maximum probability, given a source sentence.
-That is, for a source sentence $s$, the objective is to find a target translation $\hat{t}$ which has the highest conditional probability $p(t | s)$.
-Formally, this is written as:
-\begin{equation}
-\hat{t} = \arg \max_t p( t | s )
-\label{eq:argmax-factored-trans}
-\end{equation}
-where the \emph{arg max} function is the search.
-The log-linear model generalizes Equation~\ref{eq:argmax-factored-trans} to include more component models and weighting each model according to the contribution of each model to the total probability.
-\begin{equation}
-\label{eq:Log-Linear}
-p(t | s) 	=  \frac{1}{Z} \exp ( \sum_m \lambda_m h_m ( t, s)^{} )
-\end{equation}
-where $\lambda_m$ is the weight, and $h_m$ is the \emph{feature function}, or `score', for model $m$. $Z$ is the partition function which can be ignored for optimization.
-
-The standard feature functions in the phrase-based model include:
-\begin{enumerate}
-  \item \vspace{-2 mm} a distortion penalty
-  \item \vspace{-2 mm} a phrase-penalty,
-  \item \vspace{-2 mm} a word penalty,
-  \item \vspace{-2 mm} an unknown word penalty.
-  \item \vspace{-2 mm} log transforms of the target language model probability $p(t)$, 
-  \item \vspace{-2 mm} log transforms translation model probabilities, $p_{TM}(t|s) $ and $p_{TM}(s|t)$, and word-based translation probabilities $p_w(t|s)$ and $p_w(s|t)$,
-  \item \vspace{-2 mm} log transforms of the lexicalized re-ordering probabilities,
-\end{enumerate}
-%which we have suffixed with $_{TM} $ to avoid confusion with the overall model probability $ p(t|s) $ and $ p(s|t)$.
-
-Of these feature functions, we will focus on optimizing the speed of the phrase-table and lexicalized re-ordering models.
-
-%The first four feature functions are simple and do not leave much room for optimization.
-As have already been mentioned, the language model consumes a large part of decoding time but it has already been heavily optimized in previous works such as ~\cite{Heafield-kenlm} and ~\cite{yasuhara-EtAl:2013:EMNLP}.
-
 \subsection{Beam Search}
-
-A translation of a source sentence is created by applying a series of translation rules which together translate each source word once, and only once.
-Each partial translation is known as a \emph{hypothesis}, which is created by applying a rule to an existing hypothesis.
-This \emph{hypothesis expansion} process starts with a hypothesis that has translated no source word and ends with completed hypotheses that has translated all source words.
-The highest-scoring completed hypothesis, according to the model score, is considered the best translation, $\hat{t} $.
-
-In the phrase-based model, each rule translates a contiguous sequence of source words.
-Successive applications of translation rules do not have to be adjacent on the source side, depending on the distortion limit.
-The target output is constructed strictly left-to-right from the target side using this series of translation rules.
-
-A beam search algorithm is used to create the completed hypothesis set efficiently.
-Hypotheses are grouped into stacks where each stack holds a number of comparable hypotheses.
-Most phrase-based implementations group hypotheses according to coverage cardinality.
-% in the same stack that have the same coverage cardinality $|C|$, where $C$ is  the coverage set, $C \subseteq \{1,2,...
-|s| \} $ of the number of  source words translated.
 
@@
 
-\subsection{Efficient Memory Allocation}
-
-The search algorithm creates and destroy a large number of intermediate objects such as hypotheses and feature function states.
-This puts a burden on the operating system due to the need to synchronize memory access, especially when using a large number of threads.
-Libraries such as tcmalloc~\citep{ghemawat2009tcmalloc} are designed to reduce locking contention for multi-threaded application but in our case, this is still not enough.
-
-We shall seek to improve decoding speed by replacing the operating system's general purpose memory management with our own custom memory management scheme.
-Memory will be allocated from a memory pool rather than use the operating system's general purpose allocation functions.
-
-A memory pool is a large block of memory that has been given to the application by the operating system.
-The application is then responsible for allocating portions of this memory to its components when requested.
-We will use thread-specific memory pools to increase speed by avoiding locking contention during memory access.
-Our memory pools will be dynamic.
-That is, the memory requirement does not have to be known or specified before running the application, the pool can grow when required but they will never reduce in size.
-The pools are deleted only when the application ends.
-
-To further increase memory management speed, objects in the memory pool are not deleted.
-Unused data structures accumulates in the pool until a reset event.
-The pool is assumed to be empty and simply reused after the event.
-We instantiate two memory pools per decoding thread, one which is never reset and another which is reset after the decoding of each sentence.
-Data structures are created in either pool according to their life cycle.
-
-Accumulating unused objects in the memory pools can result in unacceptably high memory usage so object queues are available for high-churn objects which allows the decoder to re-cycle unused objects before the reset event.
-This also increases speed as LIFO queues are used so that the most recently accessed memory are used, increasing CPU cache hits.
-
-%The speed of memory access is dependent on whether the data is in the CPU cache which is a constrained resource compared to memory size, typically 20MB in the latest processors.
-We shall seek to re-use recently accessed information to increase likelihood of the data being in the CPU cache.
-
-%In multiprocessor servers, the CPU cache is attached to each processor and each core.
-If a sentences is being decoded on one CPU is switched to another, the CPU cache on the new CPU must be repopulated, slowing down decoding.
-We will therefore investigate binding threads to specific cores.
-
-\subsection{Stack Configurations}
-
-The most popular stack configuration for phrase-based models is coverage cardinality, that is, hypotheses that have translated the same number of source words are stored in the same stack.
-This is implemented in Pharaoh, Moses and Joshua.
-
-%Since stack pruning occurs in each stack independently of each other, this affect the decoder's search errors.
-
-However, there are alternatives to this configuration.
-~\cite{Och:2001a} uses a single stack for all hypotheses, ~\cite{Brown:1993} uses coverage stacks (ie.
-one stack per unique coverage vector) while ~\cite{peitz2012jane} and ~\cite{Zens+Ney:2008:iwslt} apply both coverage and cardinality pruning.
-While useful, these prior works present only one particular stack configuration each.
-~\cite{ortizmartinez-garciavarea-casacuberta:2006:WMT} explore a range of stack configurations by defining a granularity parameter which controls the maximum number of stacks required to decode a sentence.
-
-We shall re-visit the question of stack configuration with a particular emphasis on how they can help improve the tradeoff between speed and translation quality.
-We will do so in the context of the cube-pruning algorithm, the algorithm that we will be using and which was not available to many of the earlier work.
-
-\subsection{Phrase-Table Optimizations}
-
-For any phrase-table table of a realistic size, memory and loading time constraints requires us to use a load-on-demand implementation.
-Moses has several which we can make use of, each with differing performance characteristics.
-Figure~\ref{fig:moses-phrase-tables-time} shows the decoding speed for the fastest two implementations.
-\begin{figure}
-\centering
-\begin{tabular}{cc}
-{\includegraphics[scale=0.4]{moses-phrase-table.png}} 
-\end{tabular}
-\caption{Moses decoding speed with two different phrase-table implementations}
-\label{fig:moses-phrase-tables-time}
-\end{figure} 
-From this, it appears that the Probing phrase-table~\citep{Bogoychev.probing.2016} has the fastest translation rule lookup, especially with large number of cores, therefore, we will concentrate exclusively on this implementation from hereon.
-
-We propose two optimizations.
-Firstly, the translation rule caching mechanism in Moses saves the most recently used rules.
-However, this require locking and active management in clearing of old rules.
-The result is \emph{slower} decoding, Table~\ref{tab:phrase-table-cache}.
-\begin{table}
-\small
-\begin{center}
-\begin{tabular}{|l|c|c|} \hline
-		& No cache	& Caching \\ \hline
-Decoding time  	& 2877	& 2540 (-12\%) \\ \hline
-\end{tabular}
-\end{center}
-\caption{Decoding speed (in words / sec with 32 threads) when using phrase-table cache}
-\label{tab:phrase-table-cache}
-\end{table}
-
-We shall explore a simpler caching mechanism by creating a static cache of the most likely translation rules to be used at the start of decoding.
-
-Secondly, the Probing phrase-table use a simple compression algorithm to compress the target side of the translation rule.
-Compression was championed by ~\cite{junczys_tsd_2012b} as the main reason behind the speed of their phrase-table but as we saw in Figure~\ref{fig:moses-phrase-tables-time}, this comes at the cost of scalability to large number of threads.
-We shall therefore take the opposite approach to and improve decoding speed by disabling compression.
-
-\subsection{Lexicalized Re-ordering Model Optimizations}
-
-Similar to the phrase-table, the lexicalized re-ordering model is trained on parallel data.
-A resultant model file is then queried during decoding.
-The need for random lookup during querying inevitably results in slower decoding speed.
-Previous work such as~\cite{junczys_tsd_2012b} improve querying speed with more compact data structures.
-
-However, the model's query keys are the source and target phrase of each translation rule.
-Rather than storing the lexicalized re-ordering model separately, we shall integrating it into the translation model, eliminating the need to query a separate file.
-However, the model remains the same under the log-linear framework, including having its own weights.
-%The search algorithm will produce identical results to the Moses baseline apart from edge case training configurations, for example, where the lexicalized reordering model is trained on a different dataset than the translation model.
-
-This optimization has precedent in ~\cite{peitz2012jane} but the effect on decoding speed were not published. We will compare results with using a separate model in this paper.
-
 \section{Experimental Setup}
@@
 
-We trained a phrase-based system using the Moses toolkit with standard settings.
-The training data consisted of most of the publicly available Arabic-English data from Opus~\citep{tiedemann2012parallel} containing over 69 million parallel sentences, and tuned on a held out set.
-The phrase-table was then pruned, keeping only the top 100 entries per source phrase, according to $p(t|s)$.
-All model files were then binarized; the language models were binarized using KenLM~\citep{Heafield-kenlm}, the phrase table using the Probing phrase-table, lexicalized re-ordering model using the compact data structure~\citep{junczys_tsd_2012b}.
-These binary formats were chosen for their best-in-class multithreaded performance.
-Table~\ref{tab:model-files} gives details of the resultant sizes of the model files.
-For testing decoding speed, we used a subset of the training data, Table~\ref{tab:test-sets}.
-\begin{table}
-\begin{center}
-\begin{tabular}{|l|c|c|} \hline
-		& ar-en	& fr-en \\ \hline
-Phrase table  	& 17 	& 5.8 \\
-Language model (5-gram) 	& 3.1  	& 1.8 \\ 
-Lex re. model	& 2.3	& 0.6 \\ \hline
-\end{tabular}
-\end{center}
-\caption{Model sizes in GB}
-\label{tab:model-files}
-\end{table}
-
-For verification with a different dataset, we also used a second system trained on the French-English Europarl corpus (2m parallel sentences).
-The two different systems have characterics that we are interested in analyzing; ar-en have short sentences with large models while fr-en have overly long sentences with smaller models.
-Where we need to compare model scores, we used held out test sets.
-
-\begin{table}
-\begin{center}
-\small
-\begin{tabular}{|l|r|r|} \hline
-		& ar-en		& fr-en \\ \hline
-\multicolumn{3}{|l|}{For speed testing} \\ \hline
-Set name	& \multicolumn{2}{|c|}{Subset of training data} \\
-\# sentences  	& 800k 		& 200k \\
-\# words 	& 5.8m 		& 5.9m \\ 
-Avg words/sent	& 7.3		& 29.7 \\ \hline
-\multicolumn{3}{|l|}{For model score testing} \\ \hline
-Set name	& OpenSubtitles	& newstest2011 \\
-\# sentences  	& 2000 		& 3003 \\
-\# words 	& 14,620 	& 86,162 \\ 
-Avg words/sent	& 7.3		& 28.7 \\ \hline
-\end{tabular}
-\end{center}
-\caption{Test sets}
-\label{tab:test-sets}
-\end{table}
-
-Standard Moses phrase-based configurations are used, except that we use the cube-pruning algorithm~\citep{Chiang:2007:cl} with a pop-limit of 400\footnote{the pop-limit was chosen from public discussion on the Moses mailing list on an acceptable balance between decoding speed and translation quality with Moses for commercial use %https://www.mail-archive.com/moses-support\%40mit.edu/msg13181.html
-}, rather than the basic phrase-based algorithm.
-The cube-pruning algorithm is often employed by users who require fast decoding as it gives them the ability to trade speed with translation quality via a simple pop-limit parameter.
-
-As a baseline, we use a recent\footnote{The experiments were performed between January and May 2016 with the latest github code to hand.
-The main ar-en experiments were rerun with the source code as of 8th June, 2016 to ensure there were no material difference.
-The commit hash was \scriptsize{bc5f8d15c6ce4bc678ba992860bfd4be6719cee8} } version of the Moses decoder taken from the github repository.
-
-For all experiments, we used a Dell PowerEdge R620 server with 16 cores, 32 hyper-threads, split over 2 physical processors (Intel Xeon E5-2650 @ 2.00GHz).
-The server has 380GB RAM.
-The operating system was Ubuntu 14.04, the code was compiled with gcc 4.8.4 and Boost 1.59\footnote{http://boost.org/} and the tcmalloc library.
-
 \section{Results}
 \label{sec:Results}
-
-\subsection{Optimizing Memory}
-
-%Each sentence is decoded from start to finish using one worker thread.
-%The CPU affinity of each thread is set to a specific core to minimize memory page faults as a result of threads switching cores.
-
-Over 24\% of the Moses decoder running time is spent on memory management (Table~\ref{tab:optimizing-memory}).
-This increases to 39\% when 32 threads are used, dampening the scalability of the decoder.
-By contrast, our decoder spends 11\% on memory management and does not significantly increase with more threads.
-\begin{table}
-\begin{center}
-\small
-\begin{tabular}{|l|c|c|c|c|} \hline
-		& \multicolumn{2}{c|} {Moses}	& \multicolumn{2}{c|} {Our Work} \\ \hline
-\# threads	& 1 		& 32	& 1 		& 32  \\ \hline
-Memory  	& 24\%		& 39\% 	& 11\%		& 13\% \\
-LM 		& 12\%	 	& 2\% 	& 47\%		& 38\% \\ 
-Phrase-table	& 9\%	 	& 5\% 	& 2\%		& 4\% \\ 
-Lex RO 		& 8\%	 	& 2\% 	& 2\%		& 2\% \\ 
-Search 		& 2\%	 	& 0\% 	& 14\%		& 19\% \\ 
-Misc/Unknown	& 45\%	 	& 39\% 	& 24\%		& 29\% \\ \hline
-\end{tabular}
-\end{center}
-\caption{Profile of \%age decoding time}
-\label{tab:optimizing-memory}
-\end{table}
-
-Figure~\ref{fig:mempool} compares the decoding speed for Moses and our decoder, using the same models, parameters and test set.
-Our decoder is 4.4 times faster with one thread, and 5.0 times faster using all cores.
-Like Moses, however, performance actually worsens after approximately 15 threads.
-%, however, the problem is not as pronounced.
-This gives us a better foundation on which to build further innovations for fast, multi-core decoding.
-\begin{figure}
-\centering
-\begin{tabular}{cc}
-{\includegraphics[scale=0.4]{mempool.png}} 
-\end{tabular}
-\caption{Decoding speed of Moses and our decoder, using the same models}
-\label{fig:mempool}
-\end{figure}
-
-% A potential disadvantage of using memory pool is that the decoder will use more memory, however, Table~\ref{tab:memory-usage} shows that peak memory consumption are comparable.
-% \begin{table}[h]
-% \begin{center}
-% \begin{tabular}{|l|c|c|} \hline
-% 		& 1 thread	& 32 threads \\ \hline
-% Moses  		& ???	 	& ??? \\
-% Our work 	& ??? 		& ??? \\ \hline
-% \end{tabular}
-% \end{center}
-% \caption{Peak Memory Usage in GB}
-% \label{tab:memory-usage}
-% \end{table}
-
-\subsection{Stack Configuration}
-
-We investigated the effects of the following three stack configurations on model score and decoding speed:
-\begin{enumerate}
-  \item \vspace{-2 mm} coverage cardinality,
-  \item \vspace{-2 mm} coverage,
-  \item \vspace{-2 mm} coverage and end position of most recently translated source word.
-\end{enumerate}
-Coverage cardinality is the same as that in Moses and Joshua.
-Coverage configuration uses one stack per unique coverage vector.
-\emph{Coverage and end position} of most recently translated source word extends the coverage configuration by separating hypotheses where the position of the last translate word are different, even if the coverages are identical.
-
-This is an optimization to reduce the number of checks on the distortion limit, which is dependent on the last word position.
-The check is a binary function $d(C_{h}, e_{hypo}, range_r)$, where $C_{h}$ is the coverage vector of hypothesis $h$, $e_{h}$ is the end position of most recent source word that has been translated, and $range_r$ is the coverage of the rule to be applied.
-%Figure~\ref{algo:Hypothesis Expansion with Unsorted Stack} shows the standard hypothesis expansion algorithm.
-%\begin{figure} [h]
-%\begin{algorithmic}
-%%\REQUIRE source sentence $s$, translation options
-%\FORALL{ $h$ in $stack_{|C|}$} 
-%  \FORALL{$r$ in translation rules}
-%      \IF {$d(C_{h}, e_{hypo}, range_r)$}
-%	\STATE expand $h$ with $r$ $\rightarrow$ $h'$
-%	\STATE add $h'$ to stack $stack_{C'}$
-%      \ENDIF
-%  \ENDFOR % { translation options}
-%\ENDFOR % { hypothesis in $stack_i$} 
-%\end{algorithmic}
-%\caption{Hypothesis Expansion with Cardinality Stacks}
-%\label{algo:Hypothesis Expansion with Unsorted Stack}
-%\end{figure}
-
-By grouping hypotheses according to coverage \emph{and} end position, the distortion limit only needs to be checked for each group. %, Figure~\ref{algo:Hypothesis Expansion with Sorted Stack}. 
-%\begin{figure} [h]
-%\begin{algorithmic}
-%%\REQUIRE source sentence $s$, translation options
-%\FORALL{ $ministack_{C, e}$ in $stack_{|C|}$} 
-%  \FORALL{$r$ in translation rules}
-%    \IF {$d(C_{h}, e_{hypo}, range_r)$}
-%      \FORALL{ $h$ in $ministack_{C, e}$} 
-%	\STATE expand $h$ with $r$ $\rightarrow$ $h'$
-%	\STATE add $h'$ to $ministack_{C',e'}$
-%      \ENDFOR
-%    \ENDIF
-%  \ENDFOR % { translation options}
-%\ENDFOR % { hypothesis in $stack_i$} 
-%\end{algorithmic}
-%\caption{Hypothesis Expansion with Coverage \& End Position Stacks}
-%\label{algo:Hypothesis Expansion with Sorted Stack}
-%\end{figure}
-However, stack pruning occurs on each hypothesis group independently, therefore, potentially affecting search errors and model scores.
-
-%Table~\ref{tab:stack-configuration} and 
-Figure~\ref{fig:stack-configuration} presents the tradeoff between decoding time and average model scores, created by varying the cube-pruning pop-limit.
-None of the different stack configurations significantly outperform the others in either quality or decoding speed.
-However, the coverage \& end position produces slightly higher model scores at higher pop-limits, therefore, we continue to use this configuration throughout the rest of this paper.
-
-%The model scores for all stack configurations are identical for low pop-limits but grouping hypotheses into coverage \& end position produces higher model scores for higher pop-limits.
-It is also slower but the time/quality tradeoff is better overall with this stack configuration.
-Model scores for coverage and cardinality configurations are identical, which differs from the results in ~\cite{ortizmartinez-garciavarea-casacuberta:2006:WMT}.
-This may be due to differing test set characteristics, or interactions with cube-pruning algorithm we used.
-
-%\begin{table*}[t]
-%\small
-%\begin{center}
-%\begin{tabular}{|r|r|r|r|r|r|r|} \hline
-%Pop-limit	& \multicolumn{2}{c|}{Cardinality}	& \multicolumn{2}{c|}{Coverage}	& \multicolumn{2}{c|}{Coverage \& end pos} \\ \hline
-%		& Time		& Score			& Time		& Score		& Time		& Score	\\ \hline
-%100  		& 73		& -8.64513 		& 75		& -8.64513	& 72	& -8.64513 \\
-%500		& 237		& -8.59563		& 225		& -8.59563 	& 229	& -8.59612 \\
-%1,000	 	& 416		& -8.58700		& 397		& -8.58700	& 423	& -8.58700 \\ 
-%5,000 		& 1930		& -8.58165		& 1931		& -8.58165	& 2153	& -8.58098 \\ 
-%10,000 		& 3619		& -8.58133		& 3630		& -8.58133	& 4576	& -8.58015 \\ 
-%15,000 		& 4830		& -8.58130		& 5001		& -8.58130	& 7156	& -8.57999 \\ 
-%20,000 		& 5849		& -8.58130		& 5916		& -8.58130	& 9583	& -8.57994 \\ \hline
-%\end{tabular}
-%\end{center}
-%\caption{Decoding time (in secs with 32 threads) and average model scores for different stack configurations}
-%\label{tab:stack-configuration}
-%\end{table*}
-
-\begin{figure}
-\centering
-\begin{tabular}{cc}
-{\includegraphics[scale=0.4]{stack-configuration.png}} 
-\end{tabular}
-\caption{Trade-off between decoding time average model scores for different stack configurations}
-\label{fig:stack-configuration}
-\end{figure} 
- %, therefore, we shall stick with this Configuration for the remainder of the paper.
-
-%Also, the cube-pruning algorithm contain a priority queue per set of hypotheses sets.
-This can be independent of how hypotheses are grouped.
-In the Moses implementation, the priority queue is also attached to the coverage cardinality, Figure~\ref{algo:Cube Pruning with Unsorted Stack}.
-%\begin{figure} [h]
-%\begin{algorithmic}
-%%\REQUIRE source sentence $s$, translation options
-%\STATE initialize $queue_{|C|}$
-%\FOR{1 to pop-limit}
-%  \STATE get best $item$ in $queue_{|C|}$
-%  \STATE create hypo from $item$, coverage $C_{new}$
-%  \STATE add hypo to $stack_{|C_{new}|}$
-%  \STATE create next $items$
-%  \STATE add new $items$ to $queue_{|C|}$
-%\ENDFOR % { hypothesis in $stack_i$} 
-%\end{algorithmic}
-%\caption{Cube Pruning with Cardinality Stacks}
-%\label{algo:Cube Pruning with Unsorted Stack}
-%\end{figure}
-%Again, we experiment with diffirent queue configurations, having separate queues for each cardinality, coverage, and coverage \& end position.
-The stack configuration remained constant (coverage \& end position with pop-limit of 400).
-%\begin{table}[h]
-%\small
-%\begin{center}
-%\begin{tabular}{|r|r|r|} \hline
-%Queue configuration		& Time		& Score \\ \hline
-%Cardinality			& 192		& -8.59922 \\
-%Coverage			& 2,413		& -8.58635 \\
-%Coverage \& end position	& 7,472		& -8.58263 \\ \hline
-%\end{tabular}
-%\end{center}
-%\caption{Decoding time (in secs with 32 threads) and average model scores for different queue configurations}
-%\label{tab:queue-configuration}
-%\end{table}
-%From the results in Table~\ref{tab:queue-configuration}, using finer grain queues does results in better model scores but it is significantly slower to decode.
-% 
-We verified that the translation quality of our decoder is comparable to that of Moses in Figure~\ref{fig:bleu}, given the same parameters and models.
-This fits in with our intention of creating a drop-in replacement for the Moses decoder.
-\begin{figure}
-\centering
-\begin{tabular}{cc}
-{\includegraphics[scale=0.5]{bleu.png}} 
-\end{tabular}
-\caption{Translation quality for different pop-limits}
-\label{fig:bleu}
-\end{figure}
-
-\subsection{Translation Model}
-
-In the first optimization, we create a static translation model cache containing translation rules that translates the most common source phrases.
-This is constructed during phrase-table training based on the source counts.
-The cache is then loaded when the decoder is started.
-It does not require the overhead of managing an active cache but there is still some overhead in using a cache.
-\begin{table}
-\small
-\begin{center}
-\begin{tabular}{|r|r|r|} \hline
-Cache size	& Decoding Time & Cache Hit \%age	\\ \hline
-Before caching	& 229	& N/A	 \\
-0		& 239 (+4.4\%)	& 0\% \\
-1,000		& 213 (-7.0\%)	& 11\% \\
-2,000		& 204 (-10.9\%)	& 13\% \\
-4,000		& 205 (-10.5\%)	& 14\% \\
-10,000		& 207 (-9.7\%)	& 17\% \\ \hline
-\end{tabular}
-\end{center}
-\caption{Decoding time (in secs with 32 threads) for varying cache sizes}
-\label{tab:cache}
-\end{table}
-Overall however, using a static cache result in a 10\% decrease in decoding time if the optimum cache size is used, Table~\ref{tab:cache}.
-
-For the second optimization, we disable the compression of the target side of the translation rules.
-This increase the size of the binary files from 17GB to 23GB but the time saved not needing to decompress the data resulted in a 1.5\% decrease in decoding time with 1 thread and nearly 7\% when the CPUs are saturated, Table~\ref{tab:compression}.
-\begin{table}
-\small
-\begin{center}
-\begin{tabular}{|r|r|r|} \hline
-\# threads	& Compressed pt & Non-compressed pt \\ \hline
-1		& 3052	& 3006 (-1.5\%) \\
-5		& 756	& 644 (-14.8\%) \\
-10		& 372	& 362 (-2.7\%) \\
-15		& 284	& 250 (-12.0\%) \\
-20		& 244	& 227 (-7.0\%) \\
-25		& 218	& 209 (-4.1\%) \\
-30		& 206	& 192 (-6.8\%) \\
-35		& 203	& 189 (-6.9\%) \\ \hline
-\end{tabular}
-\end{center}
-\caption{Decoding time (in secs with 32 threads) for compressed and non-compressed phrase-tables}
-\label{tab:compression}
-\end{table}
-
-\subsection{Lexicalized Re-ordering Model}
-
-The lexicalized re-ordering model requires a probability distribution of the re-ordering behaviour of each translation rule learnt from the training data.
-This is represented in the model file as a fixed number of probabilities for each rule, exactly how many probabilities is dependant on the model's parameterization during training.
-During decoding, a probability from this distribution is assigned to each hypothesis according to the re-ordering of the translation rule.
-
-Rather than holding the model probability distributions in the separate file, we pre-process the translation model file to include the lexicalized re-ordering model distributions for each rule.
-During decoding, the probability distribution is then taken from the translation model instead of querying a separate file.
-
-This resulted in a significant decrease in decoding time, especially with high number of cores, Figure~\ref{fig:lex-ro}. 
-\begin{figure}
-\centering
-\begin{tabular}{cc}
-{\includegraphics[scale=0.4]{lex-ro.png}} 
-\end{tabular}
-\caption{Decoding speed with Compact Lexicalized Re-ordering, and integrated into a model the phrase-table}
-\label{fig:lex-ro}
-\end{figure} 
-Decoding speed increased by 40\% when using one thread but is 5 times faster when using 32 threads.
-%In fact, the decoding time with the integrated model is similar to that \emph{without} a lexicalized re-ordering model\footnote{Strangely, using lexicalized re-ordering model results in slightly faster decoding than without the model.
-We double checked with other datasets.
-We suspect this may be an artefact of the algorithm or data structures, for example, more work is needed to resolve hypothesis recombination.}.
-Critically for systems with multicores servers, it enables the decoder to continue to scale, making efficient use of all available cores.
-
-\subsection{Scalability}
-
-\begin{figure}
-\centering
-\begin{tabular}{cc}
-{\includegraphics[scale=0.4]{scalability.png}} 
-\end{tabular}
-\caption{Comparison of decoding speed of our work and Moses (with \& without the tcmalloc library}
-\label{fig:speed}
-\end{figure}
-
-Figure~\ref{fig:speed} shows decoding speed against the number of threads used.
-In our work, there is a constant increase in decoding speed when more threads are used, decreasing slightly after 16 threads when virtual cores are employed by the CPU.
-Overall, decoding is 16 times faster than single-threaded decoding when all 16 cores (32 hyperthreads) are fully utilized.
-
-%Since the test server only has 16 real cores, additional threads must be run on the same physical cores as existing threads.
-This further constrains the resources that each thread has, therefore we see a markedly lower speed increase with each additional thread once there are more than 16 concurrent threads.
-
-This contrast with Moses where speed increases up to approximately 16 threads but then become slower thereafter.
-Using the tcmalloc library has a small positive effect on decoding speed but does little to improve scalability
-%\begin{figure}[h]
-%\centering
-%\begin{tabular}{cc}
-%{\includegraphics[scale=0.5]{moses-scalability.png}} 
-%\end{tabular}
-%\caption{Moses' decoding speed}
-%\label{fig:moses-speed}
-%\end{figure}
-
-%Overall though, the scalability is remarkably good, the decoder is able to make full use of all real and virtual cores.
-When all 16 cores are saturated with 2 hyper-thread each, the decoder is 16 times faster than single-threaded decoding.
-
-Our work is 4.3 times faster than Moses with a single-thread and 10.4 faster when all cores are used.
-
-%We mimic the use of multi-core servers by single-threaded decoders such as mtplz~\citep{Heafield-mtplz} and Jane~\citep{peitz2012jane} by running multiple processes of the Moses decoder, each running one thread.
-The input is distributed to the processes using a external script.
-The result in Figure~\ref{fig:moses-speed} (dashed line) shows that this method scales much better than multithreading the application.
-
-\subsection{Other Models and Even More Cores}
-
-Our decoder show no scalability issues when we tested with the same model and tested set on a larger server, Figure~\ref{fig:more-cores}.
-\begin{figure}
-\centering
-\begin{tabular}{cc}
-{\includegraphics[scale=0.4]{more-cores.png}} 
-\end{tabular}
-\caption{Decoding speed of our decoder with more cores}
-\label{fig:more-cores}
-\end{figure}
-
-We verify the results with the French-English phrase-based system and test set.
-The speed gains are even greater than the Arabic-English test scenario, Figure~\ref{fig:fr-en-speed}.
-Our decoder is 5.4 times faster than Moses with a single-thread and 14.5 faster when all cores are saturated.
-\begin{figure}
-\centering
-\begin{tabular}{cc}
-{\includegraphics[scale=0.4]{fr-en-speed.png}} 
-\end{tabular}
-\caption{Decoding speed for fr-en model}
-\label{fig:fr-en-speed}
-\end{figure}
-
-It has been suggested that using a larger language model would overpower the improvements in decoding speed.
-We tested this conjecture by replacing the language model in the ar-en experiment with a 96GB language model.
-The time to load of language model is significant (394 sec) and was excluded from the translation speed.
-Results show that our decoder is 7 times faster than Moses and still scales monotonically until all CPUs are saturated, Figure~\ref{fig:large-lm}.
-\begin{figure}
-\centering
-\begin{tabular}{cc}
-{\includegraphics[scale=0.4]{large-lm.png}} 
-\end{tabular}
-\caption{Decoding speed with a large language model}
-\label{fig:large-lm}
-\end{figure}
 
@@
 Government.
+
+Thanks to Kenneth Heafield for advice and code.
 \small

### 2018-01-09 fast-beam/fast-beam.tex
fast-beam

@@
 % \mtsummitHeader{x}{x}{xxx-xxx}{2016}{45-character paper description goes here}{Author(s) initials and last name go here}
-\title{\bf Fast, Scalable Phrase-Based SMT Decoding}
+\title{\bf Fast Beam Search for Large Multiclass RNN}
 
 \author{\name{\bf Hieu Hoang} \hfill  \addr{hieu@hoang.co.uk}\\ 
-        \addr{Moses Machine Translation CIC, UK}
+        \addr{}
 \AND
-       \name{\bf Nikolay Bogoychev} \hfill \addr{s1031254@sms.ed.ac.uk}\\
+       \name{\bf Tomasz Dwojak} \hfill \addr{???}\\
+        \addr{Adam Mickiewicz University}
+\AND
+       \name{\bf Kenneth Heafield} \hfill \addr{???}\\
        %\name{\bf Kenneth Heafield} \hfill \addr{kheafiel@inf.ed.ac.uk}\\
         \addr{University of Edinburgh, Scotland}
-\AND
-       \name{\bf Lane Schwartz} \hfill \addr{lanes@illinois.edu}\\
-        \addr{University of Illinois, USA}
-\AND
-       \name{\bf Marcin Junczys-Dowmunt} \hfill \addr{junczys@amu.edu.pl}\\
-        \addr{Adam Mickiewicz University}
 }
@@
 \begin{abstract}
-The utilization of statistical machine translation (SMT) has grown enormously over the last decade, many using open-source software developed by the NLP community.
-As commercial use has increased, there is need for  software that is optimized for commercial requirements, in particular, fast phrase-based decoding and more efficient utilization of modern multicore servers.
 
-In this paper we re-examine the major components of phrase-based decoding and decoder implementation with particular emphasis on speed and scalability on multicore machines.
-The result is a drop-in replacement for the Moses decoder which is up to fifteen times faster and scales monotonically with the number of cores.
+Recurrent neural networks (RNN) has been popular in machine translation and other tasks language generation tasks.
+However, there are two characterisitics of these applications which causes ineffieciencies, leading to slow inference speed.
+Firstly, the RNN is used to discriminate between a large number of classes corresponding to the output vocabulary.
+Secondly, the output sentence lengths are unpredictable, leading to ineffieciencies during batch processing.
+We provide solutions for each of these issues which, together, increase batched inference speed by up to ???
+on modern GPUs without affecting model quality.
+For applications where the use of maxi-batching introduces unacceptable delays in response time, our work speed up inference speed by up to ???.
+
 \end{abstract}
@@
 
-We have presented a new decoder that is compatible with Moses.
-By studying the shortcomings of the current implementation, we are able to optimize for speed, particularly for multicore operation.
-This resulted in double digit gains compared to Moses on the same hardware.
-Our implementation is also unaffected by scalability issues that have afflicted Moses.
-% and continue to scale well for all possible cores we have tested it on.
-
-In future, we shall investigate other major components of the decoding algorithm, particularly the language model which has not been touched in this paper.
-We are also keen to explore the underlying reasons for the scalability issues in Moses to get a better understanding where potential performance issues can arise.
-
  \section*{Acknowledgments}

### 2018-01-09 fast-beam/fast-beam.tex
fast-beam

@@
 % \mtsummitHeader{x}{x}{xxx-xxx}{2016}{45-character paper description goes here}{Author(s) initials and last name go here}
-\title{\bf Fast Beam Search for Large Multiclass RNN}
+\title{\bf Faster Beam Search for Neural Machine Translation}
 
@@
 
-Recurrent neural networks (RNN) has been popular in machine translation and other tasks language generation tasks.
-However, there are two characterisitics of these applications which causes ineffieciencies, leading to slow inference speed.
-Firstly, the RNN is used to discriminate between a large number of classes corresponding to the output vocabulary.
-Secondly, the output sentence lengths are unpredictable, leading to ineffieciencies during batch processing.
-We provide solutions for each of these issues which, together, increase batched inference speed by up to ???
+Deep learning models are widely deployed for machine translation.
+In comparison on many other deep learning models, there are two characterisitics of machine translation which have not been adequately addressed, leading to slow inference speed.
+Firstly, as opposed to standard binary class models, machine translation models are used to discriminate between a large number of classes corresponding to the output vocabulary.
+Secondly, output sentence lengths are unpredictable, leading to ineffieciencies during batch processing.
+We provide solutions for these issues which, together, increase batched inference speed by up to ???
 on modern GPUs without affecting model quality.
 For applications where the use of maxi-batching introduces unacceptable delays in response time, our work speed up inference speed by up to ???.
+Our work has application to other language generation tasks and beyond.
 

### 2018-01-10 fast-beam/fast-beam.tex
fast-beam

@@
 Deep learning models are widely deployed for machine translation.
-In comparison on many other deep learning models, there are two characterisitics of machine translation which have not been adequately addressed, leading to slow inference speed.
+In comparison on many other deep learning models, there are two characterisitics of machine translation which have not been adequately addressed by most software implementations, leading to slow inference speed.
 Firstly, as opposed to standard binary class models, machine translation models are used to discriminate between a large number of classes corresponding to the output vocabulary.
-Secondly, output sentence lengths are unpredictable, leading to ineffieciencies during batch processing.
-We provide solutions for these issues which, together, increase batched inference speed by up to ???
+Secondly, rather than a single label, the output from machine translation models is a sequence of class labels that makes up the words in the target sentence.
+The sentence lengths are unpredictable, leading to inefficiencies during batch processing.
+We provide solutions for these issues which together, increase batched inference speed by up to ???
 on modern GPUs without affecting model quality.
 For applications where the use of maxi-batching introduces unacceptable delays in response time, our work speed up inference speed by up to ???.
-Our work has application to other language generation tasks and beyond.
+Our work is applicable to other language generation tasks and beyond.
 
@@
 
-\section{Introduction}
+\section{Introduction and Prior Work}
 
-\subsection{Prior Work}
+Deep learning models have been successfully used on many machine learning task in recent years in areas as diverse as computer vision and natural language processing.
+This success have been followed by the rush to put these model into production.
+However, the computational resources required in order to run these models have ben chanllenging.
+One possible solution involves creating faster models that can approximate the original model (???
+Student-Teacher model).
+Another solution is tackling specific computationally intensive functions by approximating them (???
+softmax).
+(???
+Cho) created the GRU that has the required properties of LSTM but is computationally cheaper.
+(???
+Attention is all you need) and (???
+course-to-fine) has been proposed as a fast alternatives to RNN-based models as it is possible to process more of the units in parallel.
 
-\subsection{Beam Search}
+It has been noticed that half precision arithmetic can be used for deep learning model without significan loss of model quality (???).
+Other solutions include specialized hardware, the most popular being graphical processing units (GPU) but other hardware such as custom processors (???
+TPU), FPGA (???
+Intel) have been used.
 
-\section{Proposed Improvements}
-\label{sec:Proposed Improvements}
+Many of these optimization are general purpose improvements for deep learning models, regardless of the task it is being used in.
+(???
+Devlin) is a noticeable machhine translation-specific optimization which describe the speed improvements that can be obtained by techniques such as the use of half-precision, model changes and pre-computation.
 
-We will also concentrate on four main areas for optimization.
+\section{Proposal}
+\label{sec:Proposal}
+
+We choose to focus on the use of GPU, rather than CPU as pursued in (??? Devlin).
+
+We based our proposal by reasoning what are the relevant differences between machine translation models and those for other tasks that deep learning has been used.
+But unlike (???
+Devlin), we have avoided proposing improvements for a specific model.
+Therefore, our solution should be applicable to other task which have the similar characteristics such as text summarization, chatbot or image captioning.
 
@@
 
-Thanks to Kenneth Heafield for advice and code.
-\small
-
 \bibliographystyle{apalike}

### 2018-01-10 fast-beam/fast-beam.tex
fast-beam

@@
 softmax).
+
+Most current neural MT models follow an encoder-decoder architecture.
+The encoder calculates the word and sentence embeddings while the decoder generates the output sentence.
+The architecture within the encoder is still a subject of much research.
 (???
-Cho) created the GRU that has the required properties of LSTM but is computationally cheaper.
+Kalshammer) used a convolution to encode the input and a recurrent neural network (RNN) for the decoder.
+RNNs was used for both encoding and decoding in (???
+who used RNN 1st).
+(???) significantly improved translation quality by using LSTM was used in each RNN node.
+(???
+Cho) used GRU that has the required properties of LSTM but is computationally cheaper.
+(???
+Badhannau) added attention model which improved translation at a cost of slower speed.
+
+There has been recent attempts to move away from the sequence-to-sequence models of RNN encoder-decoder.
+Some justify their architecture on faster inference as well as better translation results.
 (???
@@
 
-We choose to focus on the use of GPU, rather than CPU as pursued in (??? Devlin).
+We based our model on the sequence-to-sequence model of (???
+Cho) for machine translation models, but unlike (???
+Devlin), we avoid solutions for the specific model.
+Therefore, our solution should be applicable to other models, architectures and task which have the similar characteristics.
+We envisage that our solutions would be of value to models used in text summarization, chatbot or image captioning.
 
-We based our proposal by reasoning what are the relevant differences between machine translation models and those for other tasks that deep learning has been used.
-But unlike (???
-Devlin), we have avoided proposing improvements for a specific model.
-Therefore, our solution should be applicable to other task which have the similar characteristics such as text summarization, chatbot or image captioning.
+We also choose to focus on the use of GPU, rather than CPU as pursued in (??? Devlin).
+
+\subsection{Softmax and Beam Search Fusion}
+
+The output layer of most deep learning models consist of
+   1. activation - multiplication of the input with the weight matrix for that layer
+   2. addition of a bias term to the resulting scores
+   3. activation function - in the output layers, this is commonly a softmax or a variant, eg. log-softmax.
+   4. a search for the argmax output class and probability.
 
@@
 \end{document}
+\grid

### 2018-01-10 fast-beam/fast-beam.tex
fast-beam

@@
 
-The output layer of most deep learning models consist of
-   1. activation - multiplication of the input with the weight matrix for that layer
-   2. addition of a bias term to the resulting scores
-   3. activation function - in the output layers, this is commonly a softmax or a variant, eg. log-softmax.
-   4. a search for the argmax output class and probability.
+The output layer of most deep learning models consist of the following steps
+\begin{enumerate}
+   \item \vspace{-2 mm} activation - multiplication of the input with the weight matrix for that layer
+   \item \vspace{-2 mm} addition of a bias term to the resulting scores
+   \item \vspace{-2 mm} activation function - in the output layers, this is commonly softmax or a variant, eg. log-softmax.
+   \item \vspace{-2 mm} a search for the argmax output class, and probability is necessary.
+\end{enumerate}
+
+In models with a small number of classes such as binary classification, the calculation of softmax and argmax is trivial and fast.
+The activation and bias term steps are also fast unless the dimensionality is large.
+
+However, the output layer of most machine translation models frequently contains tens of thousand, if not hundred of thousands, classes corresponding to the vocabulary of the output language.
+For example, the best German-English system for WMT 2014 (???
+Rico) contains 85,000 subword units in the target vocabulary.
+In these scenarios, the trivial procedure used to calculate the output layer takes up a significant amount of time.
+Table ????
+shows the percentage of the total translationt time each step takes in our baseline system (described in Section ???).
+
+??? Table of percentage each step takes
+
+In addition, models with large number of classes frequently desire a list of the n-best classes rather than just the argmax, further complicating problem.
+
+However, there are algorithmic similarites between each step.
 

### 2018-01-10 fast-beam/fast-beam.tex
fast-beam

@@
 
-In addition, models with large number of classes frequently desire a list of the n-best classes rather than just the argmax, further complicating problem.
+In addition, we are interested not just in the argmax value but of beam search for the n-best classes and their probabilities, further complicating the problem.
 
-However, there are algorithmic similarites between each step.
+We shall leave the matrix multiplication for the next proposal and future work, and concentrate on the last three steps. There are algorithmic similarites between these steps, which we outline in ???.
+
+??? alogirthm of each step
 

### 2018-01-23 fast-beam/fast-beam.tex
fast nmt paper

@@
 % \mtsummitHeader{x}{x}{xxx-xxx}{2016}{45-character paper description goes here}{Author(s) initials and last name go here}
-\title{\bf Faster Beam Search for Neural Machine Translation}
+\title{\bf Faster Neural Machine Translation Inference}
 
@@
 
+We will look at two areas that are critical to fast NMT inference where the models in NMT differ significantly from those in other applications.
+These areas have been overlooked by the general deep-learning community, we aim to improve their efficient for NMT-specific tasks.
+
+Firstly, the number of classes in many deep-learning applications is small.
+However, the number of classes in NMT models is typically in the tens or hundreds of thousands, corresponding to the vocabulary size of the output language.
+For example, the best German-English NMT model at WMT14 contains 85,000 classes (Rico ???).
+This makes the output layer of NMT models very computationally expensive.
+???
+shows the breakdown of amount of time during translation using RNN model similar to (Rico ???); over 40\% of the time is involved in calculating the activation, softmax and the beam search.
+We will look at optimizations which explicitly target the output layer.
+
+Secondly, mini-batching is often used to increase the speed of deep-learning applications, including NMT systems.
+(???
+graph of batch v.
+speed) shows using batching can increase speed by 17 times.
+However, mini-batching does not take into acccount the variable length of NMT inputs and outputs, creating computational issues and efficiency challenges.
+The computational issues are often solved by masking unwanted calculations.
+Translation speed can often be improved by employing max-batching whereby the input set is pre-sorted by sentence length before mini-batching, creating mini-batches with similar length input sentences.
+However, the target sentence lengths will still differ even for similar length inputs but the standard mini-batchin algorithm must continue to process the batch until all target sentences have been completed.
+(???
+batchsize size v.
+iterations) shows the number of sentences still being processed for each decoding iteration, for a specific batch.
+The number of sentences still to be processed decreases, reducing the effectiveness of mini-batching.
+We will propose an alternative to the mini-batching algorithm.
+
 We based our model on the sequence-to-sequence model of (???
@@
 We envisage that our solutions would be of value to models used in text summarization, chatbot or image captioning.
-
-We also choose to focus on the use of GPU, rather than CPU as pursued in (??? Devlin).
+We also choose to focus on the use of GPU, rather than CPU as pursued in (???
+Devlin).
 
@@
 
-In models with a small number of classes such as binary classification, the calculation of softmax and argmax is trivial and fast.
-The activation and bias term steps are also fast unless the dimensionality is large.
-
-However, the output layer of most machine translation models frequently contains tens of thousand, if not hundred of thousands, classes corresponding to the vocabulary of the output language.
-For example, the best German-English system for WMT 2014 (???
-Rico) contains 85,000 subword units in the target vocabulary.
-In these scenarios, the trivial procedure used to calculate the output layer takes up a significant amount of time.
-Table ????
-shows the percentage of the total translationt time each step takes in our baseline system (described in Section ???).
-
-??? Table of percentage each step takes
-
-In addition, we are interested not just in the argmax value but of beam search for the n-best classes and their probabilities, further complicating the problem.
+In models with a small number of classes such as binary classification, the computational effort required is trivial and fast.
+However, this is not the case for large number of classes such as those found in NMT models.
 

### 2018-01-23 fast-beam/fast-beam.tex
fast nmt paper

@@
 
-We shall leave the matrix multiplication for the next proposal and future work, and concentrate on the last three steps. There are algorithmic similarites between these steps, which we outline in ???.
-
-??? alogirthm of each step
+We shall leave the matrix multiplication for the next proposal and future work, and concentrate on the last three steps, the outline of which are shown in ???.
 

### 2018-01-23 fast-beam/fast-beam.tex
fast nmt paper

@@
 We shall leave the matrix multiplication for the next proposal and future work, and concentrate on the last three steps, the outline of which are shown in ???.
+As can be seen, the operations iterate over the matrix p five times - once to add the bias, three times to calculate the softmax, and once to search for the best class.
+We shall fused the three functions and combine the matrix access into two iterations, reducing memory bandwidth.
+Kernel fusion is a well practised optimization technique (???).
+
+Secondly, we make use of the fact that softmax and exp are monotonic functions therefore we can avoid the search for the argmax in ??? as it will be the same as that in ???.
+
+Thirdly, we are only interested in the probability of the best class.
+Since we have precalculated the best class in ???, we shall only calculate softmax for this one class, avoiding the computational and memory bandwith of computing the probability for all classes.
+
+In fact, we are only interested in the argmax class during inference, not the probability.
+Therefore, by using the argmax calculated in ???
+rather than ???, we can avoid computing the softmax altogether.
+The outline of the our function, combining the work of ???-???
+is shown in ???.
 

### 2018-01-23 fast-beam/fast-beam.tex
fast nmt paper

@@
 
-\section{Introduction and Prior Work}
+\section{Introduction}
+
+We will look at two areas that are critical to fast NMT inference where the models in NMT differ significantly from those in other applications.
+These areas have been overlooked by the general deep-learning community, we aim to improve their efficient for NMT-specific tasks.
+
+Firstly, the number of classes in many deep-learning applications is small.
+However, the number of classes in NMT models is typically in the tens or hundreds of thousands, corresponding to the vocabulary size of the output language.
+For example, the best German-English NMT model at WMT14 contains 85,000 classes (Rico ???).
+This makes the output layer of NMT models very computationally expensive.
+???
+shows the breakdown of amount of time during translation using RNN model similar to (Rico ???); over 40\% of the time is involved in calculating the activation, softmax and the beam search.
+We will look at optimizations which explicitly target the output layer.
+
+Secondly, mini-batching is often used to increase the speed of deep-learning applications, including NMT systems.
+(???
+graph of batch v.
+speed) shows using batching can increase speed by 17 times.
+However, mini-batching does not take into acccount the variable length of NMT inputs and outputs, creating computational issues and efficiency challenges.
+The computational issues are often solved by masking unwanted calculations.
+Translation speed can often be improved by employing max-batching whereby the input set is pre-sorted by sentence length before mini-batching, creating mini-batches with similar length input sentences.
+However, the target sentence lengths will still differ even for similar length inputs but the standard mini-batchin algorithm must continue to process the batch until all target sentences have been completed.
+(???
+batchsize size v.
+iterations) shows the number of sentences still being processed for each decoding iteration, for a specific batch.
+The number of sentences still to be processed decreases, reducing the effectiveness of mini-batching.
+We will propose an alternative to the mini-batching algorithm.
+
+We based our model on the sequence-to-sequence model of (???
+Cho) for machine translation models, but unlike (???
+Devlin), we avoid solutions for the specific model.
+Therefore, our solution should be applicable to other models, architectures and task which have the similar characteristics.
+We envisage that our solutions would be of value to models used in text summarization, chatbot or image captioning.
+We also choose to focus on the use of GPU, rather than CPU as pursued in (???
+Devlin).
+
+\section{Prior Work}
 
@@
 
-We will look at two areas that are critical to fast NMT inference where the models in NMT differ significantly from those in other applications.
-These areas have been overlooked by the general deep-learning community, we aim to improve their efficient for NMT-specific tasks.
-
-Firstly, the number of classes in many deep-learning applications is small.
-However, the number of classes in NMT models is typically in the tens or hundreds of thousands, corresponding to the vocabulary size of the output language.
-For example, the best German-English NMT model at WMT14 contains 85,000 classes (Rico ???).
-This makes the output layer of NMT models very computationally expensive.
-???
-shows the breakdown of amount of time during translation using RNN model similar to (Rico ???); over 40\% of the time is involved in calculating the activation, softmax and the beam search.
-We will look at optimizations which explicitly target the output layer.
-
-Secondly, mini-batching is often used to increase the speed of deep-learning applications, including NMT systems.
-(???
-graph of batch v.
-speed) shows using batching can increase speed by 17 times.
-However, mini-batching does not take into acccount the variable length of NMT inputs and outputs, creating computational issues and efficiency challenges.
-The computational issues are often solved by masking unwanted calculations.
-Translation speed can often be improved by employing max-batching whereby the input set is pre-sorted by sentence length before mini-batching, creating mini-batches with similar length input sentences.
-However, the target sentence lengths will still differ even for similar length inputs but the standard mini-batchin algorithm must continue to process the batch until all target sentences have been completed.
-(???
-batchsize size v.
-iterations) shows the number of sentences still being processed for each decoding iteration, for a specific batch.
-The number of sentences still to be processed decreases, reducing the effectiveness of mini-batching.
-We will propose an alternative to the mini-batching algorithm.
-
-We based our model on the sequence-to-sequence model of (???
-Cho) for machine translation models, but unlike (???
-Devlin), we avoid solutions for the specific model.
-Therefore, our solution should be applicable to other models, architectures and task which have the similar characteristics.
-We envisage that our solutions would be of value to models used in text summarization, chatbot or image captioning.
-We also choose to focus on the use of GPU, rather than CPU as pursued in (???
-Devlin).
-
 \subsection{Softmax and Beam Search Fusion}
@@
 
-In fact, we are only interested in the argmax class during inference, not the probability.
+In fact, we are usually only interested in the argmax class during inference, not the probability.
 Therefore, by using the argmax calculated in ???
@@
 is shown in ???.
+
+The algorithm was extended from argmax to beam search for n-best classes.
+
+\subsection{Top-up Batching}
+
+The standard mini-batching algorithm is outlined in Figure~\ref{algo:Mini-batching}.
+
+\begin{figure} [h]
+\begin{algorithmic}
+%\REQUIRE source sentence $s$, translation options
+\WHILE{more input} 
+  \STATE Create batch
+  \STATE Encode
+  \WHILE{batch is not empty} 
+    \STATE Decode batch
+    \FOR{each sentence in batch}
+      \IF{translation is complete}
+        \STATE Remove sentence from batch
+      \ENDIF
+    \ENDFOR
+  \ENDWHILE
+\ENDWHILE 
+\end{algorithmic}
+\caption{Mini-batching}
+\label{algo:Mini-batching}
+\end{figure}
+
+This algorithm encode the sentences for a batch, followed by decoding the batch.
+The decoding stop once all sentences in the batch are completed.
+This is a potential inefficiency as the number of remaining sentences may not be optimal.
+
+We will focus on decoding as this is the more compute-intensive step, and issues with differing sentence sizes in encoding can partly be ameriorated by maxi-batching.
+
+Our proposed top-up batching algorithm encode and decode asynchronously.
+The encoding step, Figure~\ref{algo:Encoding for top-up batching}, is similar to the main loop of the standard algorithm but the results are added to a queue to be consumed by the decoding step later.
+
+\begin{figure} [h]
+\begin{algorithmic}
+%\REQUIRE source sentence $s$, translation options
+\WHILE{more input} 
+  \STATE Create encoding batch
+  \STATE Encode
+  \STATE Add to queue
+\ENDWHILE 
+\end{algorithmic}
+\caption{Encoding for top-up batching}
+\label{algo:Encoding for top-up batching}
+\end{figure}
+
+Rather than decoding the same batch until all sentences in the batch are completed, the decoding step processing the same batch continuously.
+New sentences are added to the batch as old sentences completes, Figure~\ref{algo:Decoding for top-up batching}.
+
+\begin{figure} [h]
+\begin{algorithmic}
+%\REQUIRE source sentence $s$, translation options
+\STATE create decoding batch $b$ from queue
+\WHILE{$b$ is not empty}
+  \STATE Decode $b$
+  \STATE Replace completed sentences with new sentence from queue
+\ENDWHILE 
+\end{algorithmic}
+\caption{Decoding for top-up batching}
+\label{algo:Decoding for top-up batching}
+\end{figure}
 

### 2018-01-24 fast-beam/fast-beam.tex
fast nmt paper

@@
 
-We shall leave the matrix multiplication for the next proposal and future work, and concentrate on the last three steps, the outline of which are shown in ???.
+We shall leave the matrix multiplication for the next proposal and future work, and concentrate on the last three steps, the outline of which are shown in Figures~\ref{algo:Add Bias Term} to ~\ref{algo:Find best}.
+
+\begin{figure} [h]
+\begin{algorithmic}
+\WHILE{more input} 
+  \STATE Create encoding batch
+\ENDWHILE 
+\end{algorithmic}
+\caption{Add Bias Term}
+\label{algo:Add Bias Term}
+\end{figure}
+
+\begin{figure} [h]
+\begin{algorithmic}
+\WHILE{more input} 
+  \STATE Create encoding batch
+\ENDWHILE 
+\end{algorithmic}
+\caption{Calculate softmax}
+\label{algo:Calculate softmax}
+\end{figure}
+
+\begin{figure} [h]
+\begin{algorithmic}
+\WHILE{more input} 
+  \STATE Create encoding batch
+\ENDWHILE 
+\end{algorithmic}
+\caption{Find best}
+\label{algo:Find best}
+\end{figure}
+
 As can be seen, the operations iterate over the matrix p five times - once to add the bias, three times to calculate the softmax, and once to search for the best class.
@@
 
-Secondly, we make use of the fact that softmax and exp are monotonic functions therefore we can avoid the search for the argmax in ??? as it will be the same as that in ???.
+Secondly, we make use of the fact that softmax and exp are monotonic functions therefore we can avoid the search for the argmax in Figure~\ref{algo:Find best} as it will be the same as that in Figure~\ref{algo:Calculate softmax}.
 
 Thirdly, we are only interested in the probability of the best class.
-Since we have precalculated the best class in ???, we shall only calculate softmax for this one class, avoiding the computational and memory bandwith of computing the probability for all classes.
+Since we have precalculated the best class in Figure~\ref{algo:Calculate softmax}, we shall only calculate softmax for this one class, avoiding the computational and memory bandwith of computing the probability for all classes.
 
 In fact, we are usually only interested in the argmax class during inference, not the probability.
-Therefore, by using the argmax calculated in ???
-rather than ???, we can avoid computing the softmax altogether.
-The outline of the our function, combining the work of ???-???
-is shown in ???.
+Therefore, by using the argmax calculated in Figure~\ref{algo:Calculate softmax}, we can avoid computing the softmax altogether.
+The outline of the our function is shown in Figure~\ref{algo:Fused Kernel}.
+
+\begin{figure} [h]
+\begin{algorithmic}
+\WHILE{more input} 
+  \STATE Create encoding batch
+\ENDWHILE 
+\end{algorithmic}
+\caption{Fused Kernel}
+\label{algo:Fused Kernel}
+\end{figure}
 

### 2018-01-24 fast-beam/fast-beam.tex
fast nmt paper

@@
 \usepackage{algorithmic}
+\usepackage{algorithm}
 \usepackage{graphicx}
@@
 \begin{algorithmic}
-\WHILE{more input} 
-  \STATE Create encoding batch
-\ENDWHILE 
+\REQUIRE activation vector $p$, bias vector $b$
+\FORALL{$p_i$ in $p$}
+  \STATE $p_i \gets p_i + b_i$
+\ENDFOR 
 \end{algorithmic}
@@
 \begin{algorithmic}
-\WHILE{more input} 
-  \STATE Create encoding batch
-\ENDWHILE 
+\REQUIRE activation vector $p$
+
+\COMMENT{calculate max for softmax stability}
+
+\STATE $max \gets - \infty$ 
+\FORALL{$p_i$ in $p$}
+  \IF{$p_i > max$}
+    \STATE $max \gets p_i$
+  \ENDIF
+\ENDFOR
+
+\COMMENT{calculate denominator}
+
+\STATE $sum \gets 0$ 
+\FORALL{$p_i$ in $p$}
+  \STATE $sum \gets sum + e^{p_i}$
+\ENDFOR
+
+\COMMENT{calculate softmax}
+
+\FORALL{$p_i$ in $p$}
+  \STATE $p_i \gets \frac{e^{p_i}}{sum} $
+\ENDFOR
+
+\RETURN $p$
+
 \end{algorithmic}
@@
 \begin{algorithmic}
-\WHILE{more input} 
-  \STATE Create encoding batch
-\ENDWHILE 
+\REQUIRE softmax vector $p$
+
+\STATE $max \gets - \infty$ 
+\FORALL{$p_i$ in $p$}
+  \IF{$p_i > max$}
+    \STATE $max \gets p_i$
+    \STATE $argmax \gets i$
+  \ENDIF
+\ENDFOR
+
+\RETURN $max$, $argmax$
+
 \end{algorithmic}
@@
 Since we have precalculated the best class in Figure~\ref{algo:Calculate softmax}, we shall only calculate softmax for this one class, avoiding the computational and memory bandwith of computing the probability for all classes.
+The outline of the our function is shown in Figure~\ref{algo:Fused Kernel}.
 
 In fact, we are usually only interested in the argmax class during inference, not the probability.
-Therefore, by using the argmax calculated in Figure~\ref{algo:Calculate softmax}, we can avoid computing the softmax altogether.
-The outline of the our function is shown in Figure~\ref{algo:Fused Kernel}.
+Therefore, we can skip the second iteration over $p$ in Figure~\ref{algo:Fused Kernel} and avoid computing the softmax altogether.
 
@@
 \begin{algorithmic}
-\WHILE{more input} 
-  \STATE Create encoding batch
-\ENDWHILE 
+\REQUIRE activation vector $p$, bias vector $b$
+
+\COMMENT{add bias, calculate $max$ \& $argmax$}
+
+\STATE $max \gets - \infty$ 
+\FORALL{$p_i$ in $p$}
+  \IF{$p_i + b_i > max$}
+    \STATE $max \gets p_i + b_i$
+    \STATE $argmax \gets i$
+  \ENDIF
+\ENDFOR
+
+\COMMENT{calculate denominator}
+
+\STATE $sum \gets 0$ 
+\FORALL{$p_i$ in $p$}
+  \IF{$p_i > max$}
+    \STATE $sum \gets sum + e^{p_i - max}$
+  \ENDIF
+\ENDFOR
+
+\COMMENT{calculate softmax($p_i$)}
+\RETURN $\frac{e^{max}}{sum}$, $argmax$
+
 \end{algorithmic}
@@
 
+We trained a sequence-to-sequence, encoder-decoder NMT system similar to that described in (Rico ???).
+This uses recurrent neural network with gated recurrent units.
+The input and output vocabulary size were both set to 85,000 using BPE ???
+to adjust the vocabulary to the desired size.
+The hidden layer dimensions were kept at the default of 512.
+We used the Marian toolkit to train our models.
+
+For inference, we used Amun (???), the current fastest open-source inference engine for the RNN/GRU model we aware of.
+
+Our training data consisted of the German-English parallel sentences from the Europarl corpus (???). To test inference speed, we used two test set with differing characteristics:
+\begin{enumerate}
+   \item \vspace{-2 mm} a subset of the training data, which contains mostly long sentences, and is, of course, in the same domain as the training data
+   \item \vspace{-2 mm} a subset of the German-English data from the Open-Subtitles corpus, consisting of mostly short, out-of-domain sentences.
+\end{enumerate}
+Table~\ref{tab:corpora} gives details of the training and test corpora.
+
+\begin{table}
+\begin{center}
+\small
+\begin{tabular}{|l|r|r|} \hline
+		& ar-en		& fr-en \\ \hline
+\multicolumn{3}{|l|}{For speed testing} \\ \hline
+Set name	& \multicolumn{2}{|c|}{Subset of training data} \\
+\# sentences  	& 800k 		& 200k \\
+\# words 	& 5.8m 		& 5.9m \\ 
+Avg words/sent	& 7.3		& 29.7 \\ \hline
+\multicolumn{3}{|l|}{For model score testing} \\ \hline
+Set name	& OpenSubtitles	& newstest2011 \\
+\# sentences  	& 2000 		& 3003 \\
+\# words 	& 14,620 	& 86,162 \\ 
+Avg words/sent	& 7.3		& 28.7 \\ \hline
+\end{tabular}
+\end{center}
+\caption{Corpora}
+\label{tab:corpora}
+\end{table}
+
 \section{Results}

### 2018-01-24 fast-beam/fast-beam.tex
fast nmt paper

@@
 This uses recurrent neural network with gated recurrent units.
-The input and output vocabulary size were both set to 85,000 using BPE ???
+The input and output vocabulary size were both set to 85,000 subwords using BPE ???
 to adjust the vocabulary to the desired size.
@@
 
-For inference, we used Amun (???), the current fastest open-source inference engine for the RNN/GRU model we aware of.
+For inference, we used and extended Amun (???), the current fastest open-source inference engine for the RNN/GRU model we aware of.
+All experiments was performed on an Nvidia GTX 1060 GPU.
+The host contains 8 Intel cores running at 2.8Ghz, 16GB RAM and SSD hard drive.
 
-Our training data consisted of the German-English parallel sentences from the Europarl corpus (???). To test inference speed, we used two test set with differing characteristics:
+Our training data consisted of the German-English parallel sentences from the Europarl corpus (???). To test inference speed, we used two test sets with differing characteristics:
 \begin{enumerate}
@@
 \end{enumerate}
-Table~\ref{tab:corpora} gives details of the training and test corpora.
+Table~\ref{tab:corpora} gives further details of the test sets.
 
@@
 \begin{tabular}{|l|r|r|} \hline
-		& ar-en		& fr-en \\ \hline
-\multicolumn{3}{|l|}{For speed testing} \\ \hline
-Set name	& \multicolumn{2}{|c|}{Subset of training data} \\
+		& Subset of training data		& OpenSubtitles \\ \hline
 \# sentences  	& 800k 		& 200k \\
-\# words 	& 5.8m 		& 5.9m \\ 
-Avg words/sent	& 7.3		& 29.7 \\ \hline
-\multicolumn{3}{|l|}{For model score testing} \\ \hline
-Set name	& OpenSubtitles	& newstest2011 \\
-\# sentences  	& 2000 		& 3003 \\
-\# words 	& 14,620 	& 86,162 \\ 
-Avg words/sent	& 7.3		& 28.7 \\ \hline
+\# subwords 	& 5.8m 		& 5.9m \\ 
+Avg subwords/sent	& 7.3		& 29.7 \\ \hline
 \end{tabular}

### 2018-01-24 fast-beam/fast-beam.tex
fast nmt paper

@@
 We trained a sequence-to-sequence, encoder-decoder NMT system similar to that described in (Rico ???).
-This uses recurrent neural network with gated recurrent units.
+This uses recurrent neural networks with gated recurrent units.
 The input and output vocabulary size were both set to 85,000 subwords using BPE ???
@@
 The hidden layer dimensions were kept at the default of 512.
-We used the Marian toolkit to train our models.
+%We used the Marian toolkit to train our models.
 
-For inference, we used and extended Amun (???), the current fastest open-source inference engine for the RNN/GRU model we aware of.
-All experiments was performed on an Nvidia GTX 1060 GPU.
-The host contains 8 Intel cores running at 2.8Ghz, 16GB RAM and SSD hard drive.
+For inference, we used and extend Amun (???), the fastest open-source inference engine we are aware of for RNN/GRU model used in this paper.
+Mini-batch was set to 128 sentences, unless explicitly stated, and a maxi-batch was ten times that of mini-batch.
+
+All experiments was performed on an Nvidia GTX 1060 GPU. The host contains 8 Intel hypercores running at 2.8Ghz, 16GB RAM and SSD hard drive.
 

### 2018-01-26 fast-beam/fast-beam.tex
fast beam paper

@@
 \FORALL{$p_i$ in $p$}
-  \STATE $sum \gets sum + e^{p_i}$
+  \STATE $sum \gets sum + e^{p_i - max}$
 \ENDFOR
@@
 \FORALL{$p_i$ in $p$}
-  \STATE $p_i \gets \frac{e^{p_i}}{sum} $
+  \STATE $p_i \gets \frac{e^{p_i} - max}{sum} $
 \ENDFOR
@@
 \COMMENT{calculate softmax($p_i$)}
-\RETURN $\frac{e^{max}}{sum}$, $argmax$
+\RETURN $\frac{1}{sum}$, $argmax$
 
 \end{algorithmic}
-\caption{Fused Kernel}
+\caption{Combined softmax and argmax}
 \label{algo:Fused Kernel}
@@
 
-For inference, we used and extend Amun (???), the fastest open-source inference engine we are aware of for RNN/GRU model used in this paper.
-Mini-batch was set to 128 sentences, unless explicitly stated, and a maxi-batch was ten times that of mini-batch.
+For inference, we used and extend Amun (???), the fastest open-source inference engine we are aware of for the model used in this paper.
+We uused a mini-batch of 128 sentences and maxi-batch of 1280 sentences, unless otherwise stated.
 
-All experiments was performed on an Nvidia GTX 1060 GPU. The host contains 8 Intel hypercores running at 2.8Ghz, 16GB RAM and SSD hard drive.
+The hardware used in all experiments was an Nvidia GTX 1060 GPU on a host containing 8 Intel hypercores running at 2.8Ghz, 16GB RAM and SSD hard drive.
 

### 2018-01-26 fast-beam/fast-beam.tex
fast beam paper

@@
 
-In fact, we are usually only interested in the argmax class during inference, not the probability.
-Therefore, we can skip the second iteration over $p$ in Figure~\ref{algo:Fused Kernel} and avoid computing the softmax altogether.
-
 \begin{figure} [h]
@@
 
-The algorithm was extended from argmax to beam search for n-best classes.
+In fact, we are usually only interested in the argmax class during inference, not the probability.
+Therefore, we can skip the second iteration over $p$ and avoid computing the softmax altogether.
+%, Figure~\ref{algo:Argmax only}.
+
+% \begin{figure} [h]
+% \begin{algorithmic}
+% \REQUIRE activation vector $p$, bias vector $b$
+% 
+% \COMMENT{add bias, calculate $argmax$}t
+% \FORALL{$p_i$ in $p$}
+%   \IF{$p_i + b_i > max$}
+%     \STATE $max \gets p_i + b_i$
+%     \STATE $argmax \gets i$
+%   \ENDIF
+% \ENDFOR 
+% 
+% 
+% \RETURN $argmax$ 
+% 
+% \end{algorithmic}
+% \caption{Argmax only}
+% \label{algo:Argmax only}
+% \end{figure}
+
+It has been shown ???
+that using the n-best number of classes, rather than just the best class, improves translation quality.
+This is a simple extension to the algorithm of Figure~\ref{algo:Fused Kernel}.
+Unlike the 1-best case, however, the softmax calculation cannot be skipped as the denominator differs for each input.
 

### 2018-01-26 fast-beam/fast-beam.tex
fast beam paper

@@
 However, the number of classes in NMT models is typically in the tens or hundreds of thousands, corresponding to the vocabulary size of the output language.
-For example, the best German-English NMT model at WMT14 contains 85,000 classes (Rico ???).
+For example, ~\cite{sennrich-haddow-birch:2016:P16-12} experimented with target vocabulary size of 60,000 and 90,000 subword units.
 This makes the output layer of NMT models very computationally expensive.
 ???
-shows the breakdown of amount of time during translation using RNN model similar to (Rico ???); over 40\% of the time is involved in calculating the activation, softmax and the beam search.
+shows the breakdown of amount of time during translation our NMT system; over 40\% of the time is involved in calculating the activation, softmax and the beam search.
 We will look at optimizations which explicitly target the output layer.
@@
 
-We based our model on the sequence-to-sequence model of (???
-Cho) for machine translation models, but unlike (???
-Devlin), we avoid solutions for the specific model.
+We based our model on the sequence-to-sequence model of ~\cite{D14-1179} for machine translation models, but unlike ~\cite{DBLP:conf/emnlp/Devlin17}, we avoid solutions for the specific model.
 Therefore, our solution should be applicable to other models, architectures and task which have the similar characteristics.
 We envisage that our solutions would be of value to models used in text summarization, chatbot or image captioning.
-We also choose to focus on the use of GPU, rather than CPU as pursued in (???
-Devlin).
+We also choose to focus on the use of GPU, rather than CPU as pursued in ~\cite{DBLP:conf/emnlp/Devlin17}.
 
@@
 However, the computational resources required in order to run these models have ben chanllenging.
-One possible solution involves creating faster models that can approximate the original model (???
-Student-Teacher model).
+One possible solution involves creating faster models that can approximate the original model~\citep{DBLP:conf/emnlp/KimR16}.
 Another solution is tackling specific computationally intensive functions by approximating them (???
@@
 The architecture within the encoder is still a subject of much research.
-(???
-Kalshammer) used a convolution to encode the input and a recurrent neural network (RNN) for the decoder.
+~\cite{kalchbrenner13emnlp} used a convolution to encode the input and a recurrent neural network (RNN) for the decoder.
 RNNs was used for both encoding and decoding in (???
@@
 (???) significantly improved translation quality by using LSTM was used in each RNN node.
-(???
-Cho) used GRU that has the required properties of LSTM but is computationally cheaper.
-(???
-Badhannau) added attention model which improved translation at a cost of slower speed.
+~\cite{D14-1179} used GRU that has the required properties of LSTM but is computationally cheaper.
+~\cite{DBLP:journals/corr/BahdanauCB14} added attention model which improved translation at a cost of slower speed.
 
@@
 
-It has been noticed that half precision arithmetic can be used for deep learning model without significan loss of model quality (???).
-Other solutions include specialized hardware, the most popular being graphical processing units (GPU) but other hardware such as custom processors (???
-TPU), FPGA (???
-Intel) have been used.
+It has been noticed that half precision arithmetic can be used for deep learning model without significan loss of model quality~\citep{DBLP:journals/corr/abs-1710-03740}.
+Other solutions include specialized hardware, the most popular being graphical processing units (GPU) but other hardware such as custom processors TPU~\citep{DBLP:journals/corr/JouppiYPPABBBBB17}, FPGA~\citep{DBLP:journals/corr/LaceyTA16} have been used.
 
 Many of these optimization are general purpose improvements for deep learning models, regardless of the task it is being used in.
-(???
-Devlin) is a noticeable machhine translation-specific optimization which describe the speed improvements that can be obtained by techniques such as the use of half-precision, model changes and pre-computation.
+~\cite{DBLP:conf/emnlp/Devlin17} is a noticeable machhine translation-specific optimization which describe the speed improvements that can be obtained by techniques such as the use of half-precision, model changes and pre-computation.
 
@@
 We shall fused the three functions and combine the matrix access into two iterations, reducing memory bandwidth.
-Kernel fusion is a well practised optimization technique (???).
+Kernel fusion~\citep{Guevara2009EnablingTP} is a well practised optimization technique.
 
@@
 
-It has been shown ???
-that using the n-best number of classes, rather than just the best class, improves translation quality.
+It has been shown ~\citep{koehn-knowles:2017:NMT} that using beam search to use the n-best number of classes, rather than just the best class, improves translation quality.
 This is a simple extension to the algorithm of Figure~\ref{algo:Fused Kernel}.
@@
 
-We trained a sequence-to-sequence, encoder-decoder NMT system similar to that described in (Rico ???).
+We trained a sequence-to-sequence, encoder-decoder NMT system similar to that described in ~\cite{sennrich-haddow-birch:2016:P16-12}.
 This uses recurrent neural networks with gated recurrent units.
-The input and output vocabulary size were both set to 85,000 subwords using BPE ???
-to adjust the vocabulary to the desired size.
-The hidden layer dimensions were kept at the default of 512.
+The input and output vocabulary size were both set to 85,000 subwords using byte-pair encoding (BPE) to adjust the vocabulary to the desired size.
+The hidden layer dimensions was set to 512.
 %We used the Marian toolkit to train our models.
 
-For inference, we used and extend Amun (???), the fastest open-source inference engine we are aware of for the model used in this paper.
+For inference, we used and extend Amun~\citep{junczys2016neural}, the fastest open-source inference engine we are aware of for the model used in this paper.
 We uused a mini-batch of 128 sentences and maxi-batch of 1280 sentences, unless otherwise stated.
@@
 
-Our training data consisted of the German-English parallel sentences from the Europarl corpus (???). To test inference speed, we used two test sets with differing characteristics:
+Our training data consisted of the German-English parallel sentences from the Europarl corpus~\citep{Koehn:2005:MTS}. To test inference speed, we used two test sets with differing characteristics:
 \begin{enumerate}

### 2018-01-26 fast-beam/fast-beam.tex
fast beam paper

@@
-\documentclass[]{article}
+\documentclass[11pt,a4paper]{article}
+\usepackage[hyperref]{acl2018}
+\usepackage{times}
+\usepackage{latexsym}
+
 \usepackage[letterpaper]{geometry}
-\usepackage{amta2016}
-\usepackage{times}
+%\usepackage{amta2016}
 \usepackage{url}
-\usepackage{latexsym}
 \usepackage{natbib}
@@
 \usepackage{graphicx}
+\usepackage{amsmath}
+\DeclareMathOperator*{\argmax}{argmax}
 
@@
 However, the number of classes in NMT models is typically in the tens or hundreds of thousands, corresponding to the vocabulary size of the output language.
-For example, ~\cite{sennrich-haddow-birch:2016:P16-12} experimented with target vocabulary size of 60,000 and 90,000 subword units.
+For example, ~\cite{sennrich-haddow-birch:2016:P16-12} experimented with target vocabulary sizes of 60,000 and 90,000 subword units.
 This makes the output layer of NMT models very computationally expensive.
@@
 who used RNN 1st).
-(???) significantly improved translation quality by using LSTM was used in each RNN node.
+~\cite{Sutskever:2014:SSL:2969033.2969173} significantly improved translation quality by using LSTM was used in each RNN node.
 ~\cite{D14-1179} used GRU that has the required properties of LSTM but is computationally cheaper.
@@
 Some justify their architecture on faster inference as well as better translation results.
-(???
-Attention is all you need) and (???
+~\cite{DBLP:journals/corr/VaswaniSPUJGKP17} and (???
 course-to-fine) has been proposed as a fast alternatives to RNN-based models as it is possible to process more of the units in parallel.
@@
 \begin{enumerate}
-   \item \vspace{-2 mm} activation - multiplication of the input with the weight matrix for that layer
-   \item \vspace{-2 mm} addition of a bias term to the resulting scores
-   \item \vspace{-2 mm} activation function - in the output layers, this is commonly softmax or a variant, eg. log-softmax.
-   \item \vspace{-2 mm} a search for the argmax output class, and probability is necessary.
+   \item \vspace{-2 mm} multiplication of the input with the weight matrix $p = w x$
+   \item \vspace{-2 mm} addition of a bias term to the resulting scores $p = p + b$
+   \item \vspace{-2 mm} applying the activation function, most commonly softmax $ p_i = \exp(p_i) / \sum \exp(p_i) $
+   \item \vspace{-2 mm} a search for the best output class, and probability if necessary $\argmax_i p_i$
 \end{enumerate}
@@
 \begin{algorithmic}
-\REQUIRE activation vector $p$, bias vector $b$
+\REQUIRE vector $p$, bias vector $b$
 \FORALL{$p_i$ in $p$}
@@
 \begin{algorithmic}
-\REQUIRE activation vector $p$
+\REQUIRE vector $p$
 
@@
 \FORALL{$p_i$ in $p$}
-  \STATE $sum \gets sum + e^{p_i - max}$
+  \STATE $sum \gets sum + \exp(p_i - max)$
 \ENDFOR
@@
 \FORALL{$p_i$ in $p$}
-  \STATE $p_i \gets \frac{e^{p_i} - max}{sum} $
+  \STATE $p_i \gets \frac{\exp(p_i) - max}{sum} $
 \ENDFOR
@@
     \STATE $max \gets p_i$
-    \STATE $argmax \gets i$
+    \STATE $best \gets i$
   \ENDIF
@@
 
-\RETURN $max$, $argmax$
+\RETURN $max$, $best$
 
@@
     \STATE $max \gets p_i + b_i$
-    \STATE $argmax \gets i$
+    \STATE $best \gets i$
   \ENDIF
@@
   \IF{$p_i > max$}
-    \STATE $sum \gets sum + e^{p_i - max}$
+    \STATE $sum \gets sum + \exp(p_i - max)$
   \ENDIF
@@
 \COMMENT{calculate softmax($p_i$)}
-\RETURN $\frac{1}{sum}$, $argmax$
+\RETURN $\frac{1}{sum}$, $best$
 
 \end{algorithmic}
-\caption{Combined softmax and argmax}
+\caption{Fused softmax and argmax}
 \label{algo:Fused Kernel}
@@
 
- \section*{Acknowledgments}
-This work is sponsored by the Air Force Research Laboratory, prime contract FA8650-11-C-6160.
+%\section*{Acknowledgments}
+%This work is sponsored by the Air Force Research Laboratory, prime contract FA8650-11-C-6160.
 The views and conclusions contained in this document are those of the authors and should not be interpreted as representative of the official policies, either expressed or implied, of the Air Force Research Laboratory or the U.S.

### 2018-01-27 fast-beam/fast-beam.tex
fast beam paper

@@
 
-Thirdly, we are only interested in the probability of the best class.
-Since we have precalculated the best class in Figure~\ref{algo:Calculate softmax}, we shall only calculate softmax for this one class, avoiding the computational and memory bandwith of computing the probability for all classes.
+Thirdly, we are only interested in the best class.
+Since the best class is known, we can avoid calculating softmax for all classes.
 The outline of the our function is shown in Figure~\ref{algo:Fused Kernel}.
@@
 
-\COMMENT{calculate softmax($p_i$)}
 \RETURN $\frac{1}{sum}$, $best$
@@
 
-In fact, we are usually only interested in the argmax class during inference, not the probability.
-Therefore, we can skip the second iteration over $p$ and avoid computing the softmax altogether.
+In fact, we are usually only interested in the best class during inference, not the probability.
+Therefore, we can skip the second iteration over $p$ in Figure~\ref{algo:Fused Kernel} and avoid computing the softmax altogether.
 %, Figure~\ref{algo:Argmax only}.

### 2018-01-29 fast-beam/fast-beam.tex
fast beam paper

@@
 This makes the output layer of NMT models very computationally expensive.
-???
-shows the breakdown of amount of time during translation our NMT system; over 40\% of the time is involved in calculating the activation, softmax and the beam search.
+Figure~\ref{fig:pie-time} shows the breakdown of amount of time during translation our NMT system; nearly 70\% of the time is involved in the output layer.
 We will look at optimizations which explicitly target the output layer.
+\begin{figure}
+\centering
+\begin{tabular}{cc}
+{\includegraphics[scale=0.4]{pie-time.png}} 
+\end{tabular}
+\caption{Time spent during translation (beam size 5)}
+\label{fig:pie-time}
+\end{figure}
 
@@
 
-We shall leave the matrix multiplication for the next proposal and future work, and concentrate on the last three steps, the outline of which are shown in Figures~\ref{algo:Add Bias Term} to ~\ref{algo:Find best}.
+We shall leave step 1 for future work and focus on the last three steps, the outline for which are shown in Figures~\ref{algo:Add Bias Term} to ~\ref{algo:Find best}.
 
@@
 As can be seen, the operations iterate over the matrix p five times - once to add the bias, three times to calculate the softmax, and once to search for the best class.
-We shall fused the three functions and combine the matrix access into two iterations, reducing memory bandwidth.
-Kernel fusion~\citep{Guevara2009EnablingTP} is a well practised optimization technique.
+We shall use a popular method, kernel fusion~\citep{Guevara2009EnablingTP}, to optimize the output layer.
 
-Secondly, we make use of the fact that softmax and exp are monotonic functions therefore we can avoid the search for the argmax in Figure~\ref{algo:Find best} as it will be the same as that in Figure~\ref{algo:Calculate softmax}.
+Secondly, we make use of the fact that softmax and $\exp$ are monotonic functions therefore we can move the search for the best class from Figure~\ref{algo:Find best} to Figure~\ref{algo:Calculate softmax} while $max$ is being sought.
 
@@
 \begin{algorithmic}
-\REQUIRE activation vector $p$, bias vector $b$
+\REQUIRE vector $p$, bias vector $b$
 

### 2018-01-29 fast-beam/fast-beam.tex
fast beam paper

@@
 \begin{tabular}{cc}
-{\includegraphics[scale=0.4]{pie-time.png}} 
+{\includegraphics[scale=0.3]{pie-time.png}} 
 \end{tabular}
@@
 \begin{enumerate}
-   \item \vspace{-2 mm} a subset of the training data, which contains mostly long sentences, and is, of course, in the same domain as the training data
+   \item \vspace{-2 mm} a subset of the Europarl training data, which contains mostly long sentences, and is, of course, in the same domain as the training data
    \item \vspace{-2 mm} a subset of the German-English data from the Open-Subtitles corpus, consisting of mostly short, out-of-domain sentences.
@@
 \begin{tabular}{|l|r|r|} \hline
-		& Subset of training data		& OpenSubtitles \\ \hline
-\# sentences  	& 800k 		& 200k \\
-\# subwords 	& 5.8m 		& 5.9m \\ 
-Avg subwords/sent	& 7.3		& 29.7 \\ \hline
+		  & Europarl		& OpenSubtitles \\ \hline
+\# sentences  	  & 30,000 		& 50,000 \\
+\# subwords 	  & 787,908 		& 467,654 \\ 
+Avg subwords/sent & 26.3		& 9.4 \\ \hline
 \end{tabular}
 \end{center}
-\caption{Corpora}
+\caption{Test sets}
 \label{tab:corpora}
@@
 \label{sec:Results}
+
+Fusing the softmax and beam search increase the speed of the output layer by 43\% for beam size 1, decreasing to 25\% for a beam of 10 when translating the Europarl dataset.
+This led to an overall increase in translation speed of up to 23\%, Figure~\ref{fig:europarl-beam}.
+
+\begin{figure}
+\centering
+\begin{tabular}{cc}
+{\includegraphics[scale=0.5]{europarl-beam.png}} 
+\end{tabular}
+\caption{Translation speed for Europarl dataset}
+\label{fig:europarl-beam}
+\end{figure}
 

### 2018-01-29 fast-beam/fast-beam.tex
fast beam paper

@@
 \begin{tabular}{cc}
-{\includegraphics[scale=0.3]{pie-time.png}} 
+{\includegraphics[scale=0.3]{pie-time-europarl.png}} 
 \end{tabular}
@@
 Fusing the softmax and beam search increase the speed of the output layer by 43\% for beam size 1, decreasing to 25\% for a beam of 10 when translating the Europarl dataset.
-This led to an overall increase in translation speed of up to 23\%, Figure~\ref{fig:europarl-beam}.
+This led to an overall increase in translation speed of up to 23\%, Figure~\ref{fig:beam-europarl}.
+Translation speed improved by up to 41\% when translating the Opensubtitles datset, Figure~\ref{fig:beam-opensubtitles}.
 
@@
 \begin{tabular}{cc}
-{\includegraphics[scale=0.5]{europarl-beam.png}} 
+{\includegraphics[scale=0.5]{beam-europarl.png}} 
 \end{tabular}
-\caption{Translation speed for Europarl dataset}
-\label{fig:europarl-beam}
+\caption{Translation speed for Europarl test set}
+\label{fig:beam-europarl}
+\end{figure}
+
+\begin{figure}
+\centering
+\begin{tabular}{cc}
+{\includegraphics[scale=0.5]{beam-opensubtitles.png}} 
+\end{tabular}
+\caption{Translation speed for Opensubtitles test set}
+\label{fig:beam-opensubtitles}
 \end{figure}

### 2018-01-29 fast-beam/fast-beam.tex
fast beam paper

@@
 \end{tabular}
-\caption{Time spent during translation (beam size 5)}
+\caption{Time spent during translation (beam size 5, Europarl test set)}
 \label{fig:pie-time}
@@
 
+\subsection{Softmax and Beam Search Fusion}
+
 Fusing the softmax and beam search increase the speed of the output layer by 43\% for beam size 1, decreasing to 25\% for a beam of 10 when translating the Europarl dataset.
@@
 Translation speed improved by up to 41\% when translating the Opensubtitles datset, Figure~\ref{fig:beam-opensubtitles}.
+The amount of time taken to calculate the output layer dominates trannslation time for the short OpenSubtitles test set, Figure~\ref{fig:pie-time-opensubtitles}.
 
@@
 
+\begin{figure}
+\centering
+\begin{tabular}{cc}
+{\includegraphics[scale=0.3]{pie-time-opensubtitles.png}} 
+\end{tabular}
+\caption{Time spent during translation (beam size 5, Opensubtitles test set)}
+\label{fig:pie-time-opensubtitles}
+\end{figure}
+
+\subsection{Top-up Batching}
+
 \section{Conclusion}

### 2018-01-30 fast-beam/fast-beam.tex
begin matrix wrapper

@@
 
-Secondly, mini-batching is often used to increase the speed of deep-learning applications, including NMT systems.
-(???
-graph of batch v.
-speed) shows using batching can increase speed by 17 times.
-However, mini-batching does not take into acccount the variable length of NMT inputs and outputs, creating computational issues and efficiency challenges.
-The computational issues are often solved by masking unwanted calculations.
-Translation speed can often be improved by employing max-batching whereby the input set is pre-sorted by sentence length before mini-batching, creating mini-batches with similar length input sentences.
-However, the target sentence lengths will still differ even for similar length inputs but the standard mini-batchin algorithm must continue to process the batch until all target sentences have been completed.
-(???
-batchsize size v.
-iterations) shows the number of sentences still being processed for each decoding iteration, for a specific batch.
-The number of sentences still to be processed decreases, reducing the effectiveness of mini-batching.
-We will propose an alternative to the mini-batching algorithm.
+Secondly, the use of mini-batching is critical for fast model inference.
+However, mini-batching does not take into acccount variable sentence lengths which decrease the actual number of input or output sentences that are processed in parallel, negating the benefit of mini-batching.
+This can be partially reduced in the encoding with \emph{maxi-batching}, i.e.
+pre-sorting sentences by length before creating mini-batches with similar length source sentences.
+However, target sentence lengths will still differ even for similar length inputs, compromising decoding performance.
+Figure~\ref{fig:batch-size} shows the actual batch size during decoding with a maximum batch size of 128; in only 42\% of decoding iteration was batch full.
 
-We based our model on the sequence-to-sequence model of ~\cite{D14-1179} for machine translation models, but unlike ~\cite{DBLP:conf/emnlp/Devlin17}, we avoid solutions for the specific model.
-Therefore, our solution should be applicable to other models, architectures and task which have the similar characteristics.
-We envisage that our solutions would be of value to models used in text summarization, chatbot or image captioning.
-We also choose to focus on the use of GPU, rather than CPU as pursued in ~\cite{DBLP:conf/emnlp/Devlin17}.
+\begin{figure}
+\centering
+\begin{tabular}{cc}
+{\includegraphics[scale=0.3]{batch-size.png}} 
+\end{tabular}
+\caption{Actual batchsize during decoding (max batchsize 128, beam size 5, Europarl test set)}
+\label{fig:batch-size}
+\end{figure}
+
+We will propose an alternative batching algorithm to increase the actual batch size during decoding.
+
+We base our model on the sequence-to-sequence model of ~\cite{D14-1179} for machine translation models and also choose to focus on the using GPUs rather than CPUs as pursued by ~\cite{DBLP:conf/emnlp/Devlin17}.
 

### 2018-01-30 fast-beam/fast-beam.tex
begin matrix wrapper

@@
 
-We will look at two areas that are critical to fast NMT inference where the models in NMT differ significantly from those in other applications.
-These areas have been overlooked by the general deep-learning community, we aim to improve their efficient for NMT-specific tasks.
+We will examine two areas that are critical to fast NMT inference where the models differ significantly from other deep-learning models.
+We believe the optimizations of these models have been overlooked by the general deep-learning community.
 
@@
 However, target sentence lengths will still differ even for similar length inputs, compromising decoding performance.
-Figure~\ref{fig:batch-size} shows the actual batch size during decoding with a maximum batch size of 128; in only 42\% of decoding iteration was batch full.
+Figure~\ref{fig:batch-size} shows the actual batch size during decoding with a maximum batch size of 128; the batch was full in only 42\% of decoding iterations.
 
@@
 
-We base our model on the sequence-to-sequence model of ~\cite{D14-1179} for machine translation models and also choose to focus on the using GPUs rather than CPUs as pursued by ~\cite{DBLP:conf/emnlp/Devlin17}.
+We base our work on the sequence-to-sequence model of ~\cite{D14-1179}. We also choose to focus on the using NMT inference on GPUs. % rather than CPUs as pursued by ~\cite{DBLP:conf/emnlp/Devlin17}.
 
 \section{Prior Work}
+
+REDO
 

### 2018-01-31 fast-beam/fast-beam.tex
fast beam paper

@@
 pre-sorting sentences by length before creating mini-batches with similar length source sentences.
-However, target sentence lengths will still differ even for similar length inputs, compromising decoding performance.
+However, maxi-batching can introduce unacceptable delay in response for online applications as the input are not processed in the order they came.
+It is also not appropriate for applications where only the output lengths varies, for example, image captioning.
+
+Even with sequence-to-sequence NMT models, target sentence lengths will still differ even for similar length inputs, compromising decoding performance.
 Figure~\ref{fig:batch-size} shows the actual batch size during decoding with a maximum batch size of 128; the batch was full in only 42\% of decoding iterations.
@@
 
-We will propose an alternative batching algorithm to increase the actual batch size during decoding.
+We will propose an alternative batching algorithm to increase the actual batch size during decoding without the problems associated with mini-batching and maxi-batching.
 

### 2018-02-02 fast-beam/fast-beam.tex
fast beam paper

@@
 \section{Introduction}
+\label{sec:Introduction}
 
@@
 This makes the output layer of NMT models very computationally expensive.
-Figure~\ref{fig:pie-time} shows the breakdown of amount of time during translation our NMT system; nearly 70\% of the time is involved in the output layer.
+Figure~\ref{fig:pie-time-europarl} shows the breakdown of amount of time during translation our NMT system; nearly 70\% of the time is involved in the output layer.
 We will look at optimizations which explicitly target the output layer.
@@
 \end{tabular}
-\caption{Time spent during translation (beam size 5, Europarl test set)}
-\label{fig:pie-time}
+\caption{Proportion of time spent during translation (beam size 5, Europarl test set)}
+\label{fig:pie-time-europarl}
 \end{figure}
@@
 For inference, we used and extend Amun~\citep{junczys2016neural}, the fastest open-source inference engine we are aware of for the model used in this paper.
-We uused a mini-batch of 128 sentences and maxi-batch of 1280 sentences, unless otherwise stated.
+We uused a mini-batch of 128 sentences and maxi-batch 10 times the size of the mini-batch, unless otherwise stated.
 
@@
 Translation speed improved by up to 41\% when translating the Opensubtitles datset, Figure~\ref{fig:beam-opensubtitles}.
-The amount of time taken to calculate the output layer dominates trannslation time for the short OpenSubtitles test set, Figure~\ref{fig:pie-time-opensubtitles}.
+Compared to translating the Europarl test st, Figure~\ref{fig:pie-time-europarl} in Section~\ref{sec:Introduction}, the amount of time taken by the output layer dominates translation time for the OpenSubtitles test set, Figure~\ref{fig:pie-time-opensubtitles}.
 
@@
 \end{tabular}
-\caption{Translation speed for Europarl test set}
+\caption{Translation speed for Europarl test set when using the fused kernel}
 \label{fig:beam-europarl}
@@
 \end{tabular}
-\caption{Translation speed for Opensubtitles test set}
+\caption{Translation speed for Opensubtitles test set when using the fused kernel}
 \label{fig:beam-opensubtitles}
@@
 \end{tabular}
-\caption{Time spent during translation (beam size 5, Opensubtitles test set)}
+\caption{Proportion of time spent during translation (beam size 5, Opensubtitles test set)}
 \label{fig:pie-time-opensubtitles}
@@
 \subsection{Top-up Batching}
+
+After some experimentation, we decided fully fill the decoding batch only when it is at least half empty, rather than whenever a sentence has completed.
+
+The top-up batching and maxi-batching have similar goals of maximizing efficiency when translating batches of different lengths.
+Therefore, using both methods together gives limited gains, in fact, using top-up batching with maxi-batch slows of between 2\% to 18\% when translating the Europarl test set due to the overhead of using the algorithm, Figure~\ref{fig:topup-europarl}.
+When maxi-batching is inappropriate, using top-up batching alone matches the performance of maxi-batching, even being slightly faster when a small beam is used.
+
+\begin{figure}
+\centering
+\begin{tabular}{cc}
+{\includegraphics[scale=0.3]{topup-europarl.png}} 
+\end{tabular}
+\caption{Translation speed for Europarl test set using top-up batching)}
+\label{fig:topup-europarl}
+\end{figure}
+
+The results are better when translating the OpenSubtitles test set, Figure~\ref{fig:topup-opensubtitles}.
+The top-up batching does not harm performance when used with maxi-batching, even helping a little for small beams.
+However, it increases translation speed by up to 12\% when used alone.
+
+\begin{figure}
+\centering
+\begin{tabular}{cc}
+{\includegraphics[scale=0.3]{topup-opensubtitles.png}} 
+\end{tabular}
+\caption{Translation speed for OpenSubtitles test set using top-up batching)}
+\label{fig:topup-opensubtitles}
+\end{figure}
 

### 2018-02-02 fast-beam/fast-beam.tex
fast beam paper

@@
 \end{tabular}
-\caption{Proportion of time spent during translation (beam size 5, Europarl test set)}
+\caption{Proportion of time spent during translation (Europarl test set)}
 \label{fig:pie-time-europarl}
@@
 \end{tabular}
-\caption{Actual batchsize during decoding (max batchsize 128, beam size 5, Europarl test set)}
+\caption{Actual batchsize during decoding (Europarl test set)}
 \label{fig:batch-size}
@@
 For inference, we used and extend Amun~\citep{junczys2016neural}, the fastest open-source inference engine we are aware of for the model used in this paper.
-We uused a mini-batch of 128 sentences and maxi-batch 10 times the size of the mini-batch, unless otherwise stated.
+We used a beam size of 5, mini-batch of 128 sentences and maxi-batch 1280, unless otherwise stated.
 
@@
 \end{tabular}
-\caption{Proportion of time spent during translation (beam size 5, Opensubtitles test set)}
+\caption{Proportion of time spent during translation (Opensubtitles test set)}
 \label{fig:pie-time-opensubtitles}
@@
 
-After some experimentation, we decided fully fill the decoding batch only when it is at least half empty, rather than whenever a sentence has completed.
+After some experimentation, we decided to fully fill the decoding batch only when it is at least half empty, rather than whenever a sentence has completed.
 
@@
 \end{tabular}
-\caption{Translation speed for Europarl test set using top-up batching)}
+\caption{Translation speed for Europarl test set using top-up batching}
 \label{fig:topup-europarl}
@@
 \end{tabular}
-\caption{Translation speed for OpenSubtitles test set using top-up batching)}
+\caption{Translation speed for OpenSubtitles test set using top-up batching}
 \label{fig:topup-opensubtitles}

### 2018-02-02 fast-beam/fast-beam.tex
fast beam paper

@@
 \# subwords 	  & 787,908 		& 467,654 \\ 
-Avg subwords/sent & 26.3		& 9.4 \\ \hline
+Avg subwords/sent & 26.3		& 9.4 \\
+Std dev subwords/sent & 14.9		& 6.1 \\ \hline
 \end{tabular}
@@
 The top-up batching does not harm performance when used with maxi-batching, even helping a little for small beams.
-However, it increases translation speed by up to 12\% when used alone.
+However, top-up batching increases translation speed by up to 12\% when used alone.
 
@@
 \label{fig:topup-opensubtitles}
+\end{figure}
+
+The gain from top-up batching is partially dependent on how much time the standard mini-batching algorithm spend decoding with a very small number of sentences as this is does not fully utilize the GPU cores.
+From Figure~\ref{fig:batch-size-small}, 20\% of the decoding iterations have less than 8 sentences remaining in the batch when translating the Europarl test test with maxi-batching.
+This figure is 50\% for the OpenSubtitles test set with no maxi-batching.
+The OpenSubtitles test set has a higher variance of sentence lengths relative to its average sentence length, which forces the mini-batch algorithm to continue decoding with a small number of sentences while the other sentences have already completed.
+
+\begin{figure}
+\centering
+\begin{tabular}{cc}
+{\includegraphics[scale=0.3]{batch-size-small.png}} 
+\end{tabular}
+\caption{Actual batchsize during decoding}
+\label{fig:batch-size-small}
 \end{figure}

### 2018-02-02 fast-beam/fast-beam.tex
fast beam paper

@@
 
-After some experimentation, we decided to fully fill the decoding batch only when it is at least half empty, rather than whenever a sentence has completed.
+After some experimentation, we decided to top-up the decoding batch only when it is at least half empty, rather than whenever a sentence has completed.
 
@@
 
-\section{Conclusion}
+\section{Conclusion and Future Work}
+
+We have presented two methods for faster deep-learning inference, targeted at neural machine translation.
+
+The first method focused on output layer of the neural network which accounts for a large part of the running time of the NMT model.
+By fusing the output layer with the beam search, we are able to increase translation speed by up to 41\%.
+
+The second method replaces the mini-batching algorithm with one that avoids decoding with a small number of sentences, maximizing the parallel processing potential of GPUs.
+For certain scenarios, this increases translating  speed by up to 12\%.
+
+For future work, we would like to apply our optimization for other NLP and deep-learning tasks.
+We are also interested in further optimization of the output layer in NMT, specifically the matrix multiplication which currrently take up ???
+of the translation time.
 

### 2018-02-02 fast-beam/fast-beam.tex
fast beam paper

@@
 
-Deep learning models are widely deployed for machine translation.
-In comparison on many other deep learning models, there are two characterisitics of machine translation which have not been adequately addressed by most software implementations, leading to slow inference speed.
-Firstly, as opposed to standard binary class models, machine translation models are used to discriminate between a large number of classes corresponding to the output vocabulary.
-Secondly, rather than a single label, the output from machine translation models is a sequence of class labels that makes up the words in the target sentence.
+As neural machine translation (NMT) models have become the new state-of-the-art, the challenge is to make their deployment efficient and economical.
+We have identified two areas where typical NMT models differ significantly from other deep-learning models that impact on their inference speed.
+Firstly, NMT models usualy contains a large number of output classes, corresponding to the output vocabulary.
+Secondly, rather than a single label, the output from machine translation models is a sequence of class labels that makes up the words or subwords in the target sentence.
 The sentence lengths are unpredictable, leading to inefficiencies during batch processing.
-We provide solutions for these issues which together, increase batched inference speed by up to ???
-on modern GPUs without affecting model quality.
-For applications where the use of maxi-batching introduces unacceptable delays in response time, our work speed up inference speed by up to ???.
-Our work is applicable to other language generation tasks and beyond.
+We provide solutions for each of these issues which can increase batched inference speed by up to 41\% on modern GPUs without affecting model quality.
 

### 2018-02-02 fast-beam/fast-beam.tex
fast beam paper

@@
 
-REDO
+Deep-learning models has been successfully used on many machine learning task in recent years in areas as diverse as computer vision and natural language processing.
+This success have been followed by the rush to put these model into production.
+The computational resources required in order to run these models has been challenging but, fortunately, there has been a lot work to meet this challenge.
 
-Deep learning models have been successfully used on many machine learning task in recent years in areas as diverse as computer vision and natural language processing.
-This success have been followed by the rush to put these model into production.
-However, the computational resources required in order to run these models have ben chanllenging.
-One possible solution involves creating faster models that can approximate the original model~\citep{DBLP:conf/emnlp/KimR16}.
-Another solution is tackling specific computationally intensive functions by approximating them (???
-softmax).
+Hardware accelerators such as GPUs are very popular but other specialized hardware such as custom processors~\citep{DBLP:journals/corr/JouppiYPPABBBBB17} and FPGA~\citep{DBLP:journals/corr/LaceyTA16} have been used.
+Hardware-supported reduced precision is also an ideal way to speed up model inference~\citep{DBLP:journals/corr/abs-1710-03740}.
 
-Most current neural MT models follow an encoder-decoder architecture.
-The encoder calculates the word and sentence embeddings while the decoder generates the output sentence.
-The architecture within the encoder is still a subject of much research.
-~\cite{kalchbrenner13emnlp} used a convolution to encode the input and a recurrent neural network (RNN) for the decoder.
-RNNs was used for both encoding and decoding in (???
-who used RNN 1st).
-~\cite{Sutskever:2014:SSL:2969033.2969173} significantly improved translation quality by using LSTM was used in each RNN node.
-~\cite{D14-1179} used GRU that has the required properties of LSTM but is computationally cheaper.
-~\cite{DBLP:journals/corr/BahdanauCB14} added attention model which improved translation at a cost of slower speed.
+Simpler, faster models have been created that are as good, or almost as good, as more slower, more complex models~\citep{DBLP:journals/corr/BahdanauCB14}.
+Research have also gone into smaller, faster models that can approximate the original, slower, bigger model~\citep{DBLP:conf/emnlp/KimR16}.
+Some models have been justified as more suited for the parallel architecture of GPUs~\cite{DBLP:journals/corr/VaswaniSPUJGKP17}.
 
-There has been recent attempts to move away from the sequence-to-sequence models of RNN encoder-decoder.
-Some justify their architecture on faster inference as well as better translation results.
-~\cite{DBLP:journals/corr/VaswaniSPUJGKP17} and (???
-course-to-fine) has been proposed as a fast alternatives to RNN-based models as it is possible to process more of the units in parallel.
+The speed of the softmax layer when used with large vocabularies have been looked at in~\citep{DBLP:journals/corr/GraveJCGJ16} but there have been many other attempts at faster softmax, for example~\cite{DBLP:journals/corr/abs-1301-3781}, ~\cite{Zoph-2016}.
 
-It has been noticed that half precision arithmetic can be used for deep learning model without significan loss of model quality~\citep{DBLP:journals/corr/abs-1710-03740}.
-Other solutions include specialized hardware, the most popular being graphical processing units (GPU) but other hardware such as custom processors TPU~\citep{DBLP:journals/corr/JouppiYPPABBBBB17}, FPGA~\citep{DBLP:journals/corr/LaceyTA16} have been used.
+There has been little work on batching algorithms, which is suprising considering its critical importance for fast model inference.
+~\cite{Neubig-autobatching} describe an novel batching algorithm, however, its aim is to alleviate the burden on developers in batching of new models, not faster batching.
 
-Many of these optimization are general purpose improvements for deep learning models, regardless of the task it is being used in.
-~\cite{DBLP:conf/emnlp/Devlin17} is a noticeable machhine translation-specific optimization which describe the speed improvements that can be obtained by techniques such as the use of half-precision, model changes and pre-computation.
+Our paper follows most closely on from ~\cite{DBLP:conf/emnlp/Devlin17} which achieved faster NMT inference mainly by novel implementation of an existing model.
+However, we differ by focusing on GPU implementation, and the output layer and batching algorithm which were not touched on by the previous work.
 
@@
 For future work, we would like to apply our optimization for other NLP and deep-learning tasks.
-We are also interested in further optimization of the output layer in NMT, specifically the matrix multiplication which currrently take up ???
-of the translation time.
+We are also interested in further optimization of the output layer in NMT, specifically the matrix multiplication which still takes up a significant proportion of the translation time.
 

### 2018-02-02 fast-beam/fast-beam.tex
fast beam paper

@@
 We have identified two areas where typical NMT models differ significantly from other deep-learning models that impact on their inference speed.
-Firstly, NMT models usualy contains a large number of output classes, corresponding to the output vocabulary.
-Secondly, rather than a single label, the output from machine translation models is a sequence of class labels that makes up the words or subwords in the target sentence.
+Firstly, NMT models usually contains a large number of output classes, corresponding to the output vocabulary.
+Secondly, rather than a single label, the output from machine translation models is a sequence of class labels that makes up the words or sub-words in the target sentence.
 The sentence lengths are unpredictable, leading to inefficiencies during batch processing.
@@
 However, the number of classes in NMT models is typically in the tens or hundreds of thousands, corresponding to the vocabulary size of the output language.
-For example, ~\cite{sennrich-haddow-birch:2016:P16-12} experimented with target vocabulary sizes of 60,000 and 90,000 subword units.
+For example, ~\cite{sennrich-haddow-birch:2016:P16-12} experimented with target vocabulary sizes of 60,000 and 90,000 sub-word units.
 This makes the output layer of NMT models very computationally expensive.
@@
 Secondly, the use of mini-batching is critical for fast model inference.
-However, mini-batching does not take into acccount variable sentence lengths which decrease the actual number of input or output sentences that are processed in parallel, negating the benefit of mini-batching.
+However, mini-batching does not take into account variable sentence lengths which decrease the actual number of input or output sentences that are processed in parallel, negating the benefit of mini-batching.
 This can be partially reduced in the encoding with \emph{maxi-batching}, i.e.
@@
 \end{tabular}
-\caption{Actual batchsize during decoding (Europarl test set)}
+\caption{Actual batch size during decoding (Europarl test set)}
 \label{fig:batch-size}
@@
 
-There has been little work on batching algorithms, which is suprising considering its critical importance for fast model inference.
+There has been little work on batching algorithms, which is surprising considering its critical importance for fast model inference.
 ~\cite{Neubig-autobatching} describe an novel batching algorithm, however, its aim is to alleviate the burden on developers in batching of new models, not faster batching.
@@
 
-We will focus on decoding as this is the more compute-intensive step, and issues with differing sentence sizes in encoding can partly be ameriorated by maxi-batching.
+We will focus on decoding as this is the more compute-intensive step, and issues with differing sentence sizes in encoding can partly be ameliorated by maxi-batching.
 
@@
 This uses recurrent neural networks with gated recurrent units.
-The input and output vocabulary size were both set to 85,000 subwords using byte-pair encoding (BPE) to adjust the vocabulary to the desired size.
+The input and output vocabulary size were both set to 85,000 sub-words using byte-pair encoding (BPE) to adjust the vocabulary to the desired size.
 The hidden layer dimensions was set to 512.
@@
 \# sentences  	  & 30,000 		& 50,000 \\
-\# subwords 	  & 787,908 		& 467,654 \\ 
-Avg subwords/sent & 26.3		& 9.4 \\
+\# sub-words 	  & 787,908 		& 467,654 \\ 
+Avg sub-words/sent & 26.3		& 9.4 \\
 Std dev subwords/sent & 14.9		& 6.1 \\ \hline
@@
 This led to an overall increase in translation speed of up to 23\%, Figure~\ref{fig:beam-europarl}.
-Translation speed improved by up to 41\% when translating the Opensubtitles datset, Figure~\ref{fig:beam-opensubtitles}.
+Translation speed improved by up to 41\% when translating the Opensubtitles dataset, Figure~\ref{fig:beam-opensubtitles}.
 Compared to translating the Europarl test st, Figure~\ref{fig:pie-time-europarl} in Section~\ref{sec:Introduction}, the amount of time taken by the output layer dominates translation time for the OpenSubtitles test set, Figure~\ref{fig:pie-time-opensubtitles}.
@@
 \end{tabular}
-\caption{Actual batchsize during decoding}
+\caption{Actual batch size during decoding}
 \label{fig:batch-size-small}

### 2018-02-03 fast-beam/fast-beam.tex
fast beam paper

@@
 The sentence lengths are unpredictable, leading to inefficiencies during batch processing.
-We provide solutions for each of these issues which can increase batched inference speed by up to 41\% on modern GPUs without affecting model quality.
+We provide solutions for each of these issues which can increase batched inference speed by up to 57\% on modern GPUs without affecting model quality.
 
@@
 
-There has been little work on batching algorithms, which is surprising considering its critical importance for fast model inference.
+There has been surprisingly little work on batching algorithms, considering its critical importance for efficient deep-learning training and inference.
 ~\cite{Neubig-autobatching} describe an novel batching algorithm, however, its aim is to alleviate the burden on developers in batching of new models, not faster batching.
@@
 \end{tabular}
-\caption{Translation speed for Europarl test set when using the fused kernel}
+\caption{Speed using the fused kernel (Europarl)}
 \label{fig:beam-europarl}
@@
 \end{tabular}
-\caption{Translation speed for Opensubtitles test set when using the fused kernel}
+\caption{Speed using the fused kernel (Opensubtitles)}
 \label{fig:beam-opensubtitles}
@@
 \end{tabular}
-\caption{Proportion of time spent during translation (Opensubtitles test set)}
+\caption{Proportion of time spent during translation (Opensubtitles)}
 \label{fig:pie-time-opensubtitles}
@@
 \end{tabular}
-\caption{Translation speed for Europarl test set using top-up batching}
+\caption{Speed using top-up batching (Europarl)}
 \label{fig:topup-europarl}
@@
 \end{tabular}
-\caption{Translation speed for OpenSubtitles test set using top-up batching}
+\caption{Speed using top-up batching (OpenSubtitles)}
 \label{fig:topup-opensubtitles}
@@
 \end{figure}
+
+\subsection{Cummulative Results}
+
+Using both the fused kernel and top-up batching to translate
 

### 2018-02-03 fast-beam/fast-beam.tex
fast beam paper

@@
 This led to an overall increase in translation speed of up to 23\%, Figure~\ref{fig:beam-europarl}.
-Translation speed improved by up to 41\% when translating the Opensubtitles dataset, Figure~\ref{fig:beam-opensubtitles}.
+Translation speed improved by up to 41\% when translating the OpenSubtitles dataset, Figure~\ref{fig:beam-opensubtitles}.
 Compared to translating the Europarl test st, Figure~\ref{fig:pie-time-europarl} in Section~\ref{sec:Introduction}, the amount of time taken by the output layer dominates translation time for the OpenSubtitles test set, Figure~\ref{fig:pie-time-opensubtitles}.
@@
 \end{tabular}
-\caption{Speed using the fused kernel (Opensubtitles)}
+\caption{Speed using the fused kernel (OpenSubtitles)}
 \label{fig:beam-opensubtitles}
@@
 \end{tabular}
-\caption{Proportion of time spent during translation (Opensubtitles)}
+\caption{Proportion of time spent during translation (OpenSubtitles)}
 \label{fig:pie-time-opensubtitles}

### 2018-02-04 fast-beam/fast-beam.tex
fast beam paper

@@
 
-Using both the fused kernel and top-up batching to translate
+Using both the fused kernel and top-up batching to translate led to a cummalative speed improvement of up to 57\% and 34\% for the OpenSubtitles and Europarl test set, respectively, when no maxi-batching is used, Figure~\ref{fig:cummulative-europarl} and Figure~\ref{fig:cummulative-opensubtitles}.
+With maxi-batching, the speed was up to 41\% and 21\%.
+
+\begin{figure}
+\centering
+\begin{tabular}{cc}
+{\includegraphics[scale=0.3]{cummulative-europarl.png}} 
+\end{tabular}
+\caption{Cummulative results (Europarl}
+\label{fig:cummulative-europarl}
+\end{figure}
+
+\begin{figure}
+\centering
+\begin{tabular}{cc}
+{\includegraphics[scale=0.3]{cummulative-opensubtitles.png}} 
+\end{tabular}
+\caption{Cummulative results (OpenSubtitles}
+\label{fig:cummulative-opensubtitles}
+\end{figure}
 

### 2018-02-05 fast-beam/fast-beam.tex
fast beam paper

@@
 \usepackage{layout}
-\usepackage{algorithmic}
-\usepackage{algorithm}
+
+\usepackage{algorithm}% http://ctan.org/pkg/algorithms
+\usepackage{algpseudocode}% http://ctan.org/pkg/algorithmicx
+
 \usepackage{graphicx}
@@
 
-We shall leave step 1 for future work and focus on the last three steps, the outline for which are shown in Figures~\ref{algo:Add Bias Term} to ~\ref{algo:Find best}.
+We shall leave step 1 for future work and focus on the last three steps, the outline for which are shown in Figures~\ref{algo:original-softmax-beamsearch}.
 
-\begin{figure} [h]
+\begin{figure} 
 \begin{algorithmic}
-\REQUIRE vector $p$, bias vector $b$
-\FORALL{$p_i$ in $p$}
-  \STATE $p_i \gets p_i + b_i$
-\ENDFOR 
-\end{algorithmic}
-\caption{Add Bias Term}
-\label{algo:Add Bias Term}
-\end{figure}
 
-\begin{figure} [h]
-\begin{algorithmic}
-\REQUIRE vector $p$
+\Procedure{AddBias}{vector $p$, bias vector $b$}
+\ForAll{$p_i$ in $p$}
+  \State $p_i \gets p_i + b_i$
+\EndFor 
+\EndProcedure
 
-\COMMENT{calculate max for softmax stability}
+\State
 
-\STATE $max \gets - \infty$ 
-\FORALL{$p_i$ in $p$}
-  \IF{$p_i > max$}
-    \STATE $max \gets p_i$
-  \ENDIF
-\ENDFOR
+\Procedure{Softmax}{vector $p$}
 
-\COMMENT{calculate denominator}
+\Comment{calculate max for softmax stability}
+\State $max \gets - \infty$ 
+\ForAll{$p_i$ in $p$}
+  \If{$p_i > max$}
+    \State $max \gets p_i$
+  \EndIf
+\EndFor
 
-\STATE $sum \gets 0$ 
-\FORALL{$p_i$ in $p$}
-  \STATE $sum \gets sum + \exp(p_i - max)$
-\ENDFOR
+\Comment{calculate denominator} 
+\State $sum \gets 0$ 
+\ForAll{$p_i$ in $p$}
+  \State $sum \gets sum + \exp(p_i - max)$
+\EndFor
 
-\COMMENT{calculate softmax}
+\Comment{calculate softmax}
+\ForAll{$p_i$ in $p$}
+  \State $p_i \gets \frac{\exp(p_i) - max}{sum} $
+\EndFor
 
-\FORALL{$p_i$ in $p$}
-  \STATE $p_i \gets \frac{\exp(p_i) - max}{sum} $
-\ENDFOR
+\EndProcedure
 
-\RETURN $p$
+\State
+
+\Procedure{FindBest}{softmax vector $p$}
+
+\State $max \gets - \infty$ 
+\ForAll{$p_i$ in $p$}
+  \If{$p_i > max$}
+    \State $max \gets p_i$
+    \State $best \gets i$
+  \EndIf
+\EndFor
+
+\EndProcedure
 
 \end{algorithmic}
-\caption{Calculate softmax}
-\label{algo:Calculate softmax}
-\end{figure}
-
-\begin{figure} [h]
-\begin{algorithmic}
-\REQUIRE softmax vector $p$
-
-\STATE $max \gets - \infty$ 
-\FORALL{$p_i$ in $p$}
-  \IF{$p_i > max$}
-    \STATE $max \gets p_i$
-    \STATE $best \gets i$
-  \ENDIF
-\ENDFOR
-
-\RETURN $max$, $best$
-
-\end{algorithmic}
-\caption{Find best}
-\label{algo:Find best}
+\caption{Original Softmax and Beam Search Algorithm}
+\label{algo:original-softmax-beamsearch}
 \end{figure}
@@
 
-\begin{figure} [h]
-\begin{algorithmic}
-\REQUIRE vector $p$, bias vector $b$
-
-\COMMENT{add bias, calculate $max$ \& $argmax$}
-
-\STATE $max \gets - \infty$ 
-\FORALL{$p_i$ in $p$}
-  \IF{$p_i + b_i > max$}
-    \STATE $max \gets p_i + b_i$
-    \STATE $best \gets i$
-  \ENDIF
-\ENDFOR
-
-\COMMENT{calculate denominator}
-
-\STATE $sum \gets 0$ 
-\FORALL{$p_i$ in $p$}
-  \IF{$p_i > max$}
-    \STATE $sum \gets sum + \exp(p_i - max)$
-  \ENDIF
-\ENDFOR
-
-\RETURN $\frac{1}{sum}$, $best$
-
-\end{algorithmic}
-\caption{Fused softmax and argmax}
-\label{algo:Fused Kernel}
-\end{figure}
+% \begin{figure} [h]
+% \begin{algorithmic}
+% \REQUIRE vector $p$, bias vector $b$
+% 
+% \COMMENT{add bias, calculate $max$ \& $argmax$}
+% 
+% \STATE $max \gets - \infty$ 
+% \FORALL{$p_i$ in $p$}
+%   \IF{$p_i + b_i > max$}
+%     \STATE $max \gets p_i + b_i$
+%     \STATE $best \gets i$
+%   \ENDIF
+% \ENDFOR 
+% 
+% \COMMENT{calculate denominator}
+% 
+% \STATE $sum \gets 0$ 
+% \FORALL{$p_i$ in $p$}
+%   \IF{$p_i > max$}
+%     \STATE $sum \gets sum + \exp(p_i - max)$
+%   \ENDIF
+% \ENDFOR 
+% 
+% \RETURN $\frac{1}{sum}$, $best$ 
+% 
+% 
+% \end{algorithmic}
+% \caption{Fused softmax and argmax}
+% \label{algo:Fused Kernel}
+% \end{figure}
 
@@
 
-\begin{figure} [h]
-\begin{algorithmic}
-%\REQUIRE source sentence $s$, translation options
-\WHILE{more input} 
-  \STATE Create batch
-  \STATE Encode
-  \WHILE{batch is not empty} 
-    \STATE Decode batch
-    \FOR{each sentence in batch}
-      \IF{translation is complete}
-        \STATE Remove sentence from batch
-      \ENDIF
-    \ENDFOR
-  \ENDWHILE
-\ENDWHILE 
-\end{algorithmic}
-\caption{Mini-batching}
-\label{algo:Mini-batching}
-\end{figure}
+% \begin{figure} [h]
+% \begin{algorithmic}
+% %\REQUIRE source sentence $s$, translation options
+% \WHILE{more input} 
+%   \STATE Create batch
+%   \STATE Encode
+%   \WHILE{batch is not empty} 
+%     \STATE Decode batch
+%     \FOR{each sentence in batch}
+%       \IF{translation is complete}
+%         \STATE Remove sentence from batch
+%       \ENDIF
+%     \ENDFOR
+%   \ENDWHILE
+% \ENDWHILE 
+% \end{algorithmic}
+% \caption{Mini-batching}
+% \label{algo:Mini-batching}
+% \end{figure}
 
@@
 
-\begin{figure} [h]
-\begin{algorithmic}
-%\REQUIRE source sentence $s$, translation options
-\WHILE{more input} 
-  \STATE Create encoding batch
-  \STATE Encode
-  \STATE Add to queue
-\ENDWHILE 
-\end{algorithmic}
-\caption{Encoding for top-up batching}
-\label{algo:Encoding for top-up batching}
-\end{figure}
+% \begin{figure} [h]
+% \begin{algorithmic}
+% %\REQUIRE source sentence $s$, translation options
+% \WHILE{more input} 
+%   \STATE Create encoding batch
+%   \STATE Encode
+%   \STATE Add to queue
+% \ENDWHILE 
+% \end{algorithmic}
+% \caption{Encoding for top-up batching}
+% \label{algo:Encoding for top-up batching}
+% \end{figure}
 
@@
 
-\begin{figure} [h]
-\begin{algorithmic}
-%\REQUIRE source sentence $s$, translation options
-\STATE create decoding batch $b$ from queue
-\WHILE{$b$ is not empty}
-  \STATE Decode $b$
-  \STATE Replace completed sentences with new sentence from queue
-\ENDWHILE 
-\end{algorithmic}
-\caption{Decoding for top-up batching}
-\label{algo:Decoding for top-up batching}
-\end{figure}
+% \begin{figure} [h]
+% \begin{algorithmic}
+% %\REQUIRE source sentence $s$, translation options
+% \STATE create decoding batch $b$ from queue
+% \WHILE{$b$ is not empty}
+%   \STATE Decode $b$
+%   \STATE Replace completed sentences with new sentence from queue
+% \ENDWHILE 
+% \end{algorithmic}
+% \caption{Decoding for top-up batching}
+% \label{algo:Decoding for top-up batching}
+% \end{figure}
 

### 2018-02-05 fast-beam/fast-beam.tex
fast beam paper

@@
 \begin{enumerate}
-   \item \vspace{-2 mm} multiplication of the input with the weight matrix $p = w x$
+   \item \vspace{-2 mm} multiplication of the input with the weight vector $p = w x$
    \item \vspace{-2 mm} addition of a bias term to the resulting scores $p = p + b$
@@
 
-\Procedure{FindBest}{softmax vector $p$}
+\Procedure{Find-Best}{softmax vector $p$}
 
@@
 
-Secondly, we make use of the fact that softmax and $\exp$ are monotonic functions therefore we can move the search for the best class from Figure~\ref{algo:Find best} to Figure~\ref{algo:Calculate softmax} while $max$ is being sought.
+Secondly, we make use of the fact that softmax and $\exp$ are monotonic functions therefore we can move the search for the best class from FIND-BEST to SOFTMAX.
 
@@
 
-% \begin{figure} [h]
-% \begin{algorithmic}
-% \REQUIRE vector $p$, bias vector $b$
-% 
-% \COMMENT{add bias, calculate $max$ \& $argmax$}
-% 
-% \STATE $max \gets - \infty$ 
-% \FORALL{$p_i$ in $p$}
-%   \IF{$p_i + b_i > max$}
-%     \STATE $max \gets p_i + b_i$
-%     \STATE $best \gets i$
-%   \ENDIF
-% \ENDFOR 
-% 
-% \COMMENT{calculate denominator}
-% 
-% \STATE $sum \gets 0$ 
-% \FORALL{$p_i$ in $p$}
-%   \IF{$p_i > max$}
-%     \STATE $sum \gets sum + \exp(p_i - max)$
-%   \ENDIF
-% \ENDFOR 
-% 
-% \RETURN $\frac{1}{sum}$, $best$ 
-% 
-% 
-% \end{algorithmic}
-% \caption{Fused softmax and argmax}
-% \label{algo:Fused Kernel}
-% \end{figure}
+\begin{figure} [h]
+\begin{algorithmic}
+\Procedure{Fused Kernel}{vector $p$, bias vector $b$}
+
+\Comment{add bias, calculate $max$ \& $argmax$}
+
+\State $max \gets - \infty$ 
+\ForAll{$p_i$ in $p$}
+  \If{$p_i + b_i > max$}
+    \State $max \gets p_i + b_i$
+    \State $best \gets i$
+  \EndIf
+\EndFor
+
+\Comment{calculate denominator}
+
+\State $sum \gets 0$ 
+\ForAll{$p_i$ in $p$}
+  \If{$p_i > max$}
+    \State $sum \gets sum + \exp(p_i - max)$
+  \EndIf
+\EndFor
+
+\Return $\frac{1}{sum}$, $best$
+
+\EndProcedure
+\end{algorithmic}
+
+\caption{Fused softmax and argmax}
+\label{algo:Fused Kernel}
+\end{figure}
 
 In fact, we are usually only interested in the best class during inference, not the probability.
-Therefore, we can skip the second iteration over $p$ in Figure~\ref{algo:Fused Kernel} and avoid computing the softmax altogether.
-%, Figure~\ref{algo:Argmax only}.
+Therefore, we can skip the second iteration over $p$ in Figure~\ref{algo:Fused Kernel} and avoid computing the softmax altogether, Figure~\ref{algo:Argmax only}.
 
-% \begin{figure} [h]
-% \begin{algorithmic}
-% \REQUIRE activation vector $p$, bias vector $b$
-% 
-% \COMMENT{add bias, calculate $argmax$}t
-% \FORALL{$p_i$ in $p$}
-%   \IF{$p_i + b_i > max$}
-%     \STATE $max \gets p_i + b_i$
-%     \STATE $argmax \gets i$
-%   \ENDIF
-% \ENDFOR 
-% 
-% 
-% \RETURN $argmax$ 
-% 
-% \end{algorithmic}
-% \caption{Argmax only}
-% \label{algo:Argmax only}
-% \end{figure}
+\begin{figure} [h]
+\begin{algorithmic}
+\Procedure{Find-1-Best}{vector $p$, bias vector $b$}
+
+\State $max \gets - \infty$ 
+\ForAll{$p_i$ in $p$}
+  \If{$p_i + b_i > max$}
+    \State $max \gets p_i + b_i$
+    \State $best \gets i$
+  \EndIf
+\EndFor
+
+\Return $best$
+
+\EndProcedure
+
+\end{algorithmic}
+\caption{Find 1 best only}
+\label{algo:Argmax only}
+\end{figure}
 
@@
 
-% \begin{figure} [h]
-% \begin{algorithmic}
-% %\REQUIRE source sentence $s$, translation options
-% \WHILE{more input} 
-%   \STATE Create batch
-%   \STATE Encode
-%   \WHILE{batch is not empty} 
-%     \STATE Decode batch
-%     \FOR{each sentence in batch}
-%       \IF{translation is complete}
-%         \STATE Remove sentence from batch
-%       \ENDIF
-%     \ENDFOR
-%   \ENDWHILE
-% \ENDWHILE 
-% \end{algorithmic}
-% \caption{Mini-batching}
-% \label{algo:Mini-batching}
-% \end{figure}
+\begin{figure} [h]
+\begin{algorithmic}
+\Procedure{Mini-Batching}{}
+%\REQUIRE source sentence $s$, translation options
+\While{more input} 
+  \State Create batch
+  \State Encode(batch)
+  \While{batch is not empty} 
+    \State Decode(batch)
+    \For{each sentence in batch}
+      \If{translation is complete}
+        \State Remove sentence %from batch
+      \EndIf
+    \EndFor
+  \EndWhile
+\EndWhile
+
+\EndProcedure
+
+\end{algorithmic}
+\caption{Mini-batching}
+\label{algo:Mini-batching}
+\end{figure}
 
@@
 
-% \begin{figure} [h]
-% \begin{algorithmic}
-% %\REQUIRE source sentence $s$, translation options
-% \WHILE{more input} 
-%   \STATE Create encoding batch
-%   \STATE Encode
-%   \STATE Add to queue
-% \ENDWHILE 
-% \end{algorithmic}
-% \caption{Encoding for top-up batching}
-% \label{algo:Encoding for top-up batching}
-% \end{figure}
+\begin{figure} [h]
+\begin{algorithmic}
+
+\Procedure{Encode}{}
+\While{more input} 
+  \State Create encoding batch $b$
+  \State Encode($b$)
+  \State Add $b$ to queue
+\EndWhile
+
+\EndProcedure
+
+\end{algorithmic}
+\caption{Encoding for top-up batching}
+\label{algo:Encoding for top-up batching}
+\end{figure}
 
@@
 
-% \begin{figure} [h]
-% \begin{algorithmic}
-% %\REQUIRE source sentence $s$, translation options
-% \STATE create decoding batch $b$ from queue
-% \WHILE{$b$ is not empty}
-%   \STATE Decode $b$
-%   \STATE Replace completed sentences with new sentence from queue
-% \ENDWHILE 
-% \end{algorithmic}
-% \caption{Decoding for top-up batching}
-% \label{algo:Decoding for top-up batching}
-% \end{figure}
+\begin{figure} [h]
+\begin{algorithmic}
+
+\Procedure{Decode}{}
+
+\State create batch $b$ from queue
+\While{$b$ is not empty}
+  \State Decode($b$)
+  \State Replace completed sentences with new sentence from queue
+\EndWhile
+
+\EndProcedure
+
+\end{algorithmic}
+\caption{Decoding for top-up batching}
+\label{algo:Decoding for top-up batching}
+\end{figure}
 

### 2018-02-05 fast-beam/fast-beam.tex
fast beam paper

@@
 
-We shall leave step 1 for future work and focus on the last three steps, the outline for which are shown in Figures~\ref{algo:original-softmax-beamsearch}.
+We shall leave step 1 for future work and focus on the last three steps, the outline for which are shown in Algorithm~\ref{algo:original-softmax-beamsearch}.
 
-\begin{figure} 
+\begin{algorithm} 
 \begin{algorithmic}
@@
 \label{algo:original-softmax-beamsearch}
-\end{figure}
+\end{algorithm}
 
@@
 Since the best class is known, we can avoid calculating softmax for all classes.
-The outline of the our function is shown in Figure~\ref{algo:Fused Kernel}.
+The outline of the our function is shown in Algorithm~\ref{algo:Fused Kernel}.
 
-\begin{figure} [h]
+\begin{algorithm}
 \begin{algorithmic}
@@
 \label{algo:Fused Kernel}
-\end{figure}
+\end{algorithm}
 
 In fact, we are usually only interested in the best class during inference, not the probability.
-Therefore, we can skip the second iteration over $p$ in Figure~\ref{algo:Fused Kernel} and avoid computing the softmax altogether, Figure~\ref{algo:Argmax only}.
+Therefore, we can skip the second iteration over $p$ in Algorithm~\ref{algo:Fused Kernel} and avoid computing the softmax altogether, Algorithm~\ref{algo:Argmax only}.
 
-\begin{figure} [h]
+\begin{algorithm}
 \begin{algorithmic}
@@
 \label{algo:Argmax only}
-\end{figure}
+\end{algorithm}
 
 It has been shown ~\citep{koehn-knowles:2017:NMT} that using beam search to use the n-best number of classes, rather than just the best class, improves translation quality.
-This is a simple extension to the algorithm of Figure~\ref{algo:Fused Kernel}.
+This is a simple extension to the algorithm of Algorithm~\ref{algo:Fused Kernel}.
 Unlike the 1-best case, however, the softmax calculation cannot be skipped as the denominator differs for each input.
@@
 
-The standard mini-batching algorithm is outlined in Figure~\ref{algo:Mini-batching}.
+The standard mini-batching algorithm is outlined in Algorithm~\ref{algo:Mini-batching}.
 
-\begin{figure} [h]
+\begin{algorithm}
 \begin{algorithmic}
@@
 \label{algo:Mini-batching}
-\end{figure}
+\end{algorithm}
 
@@
 Our proposed top-up batching algorithm encode and decode asynchronously.
-The encoding step, Figure~\ref{algo:Encoding for top-up batching}, is similar to the main loop of the standard algorithm but the results are added to a queue to be consumed by the decoding step later.
+The encoding step, Algorithm~\ref{algo:Encoding for top-up batching}, is similar to the main loop of the standard algorithm but the results are added to a queue to be consumed by the decoding step later.
 
-\begin{figure} [h]
+\begin{algorithm}
 \begin{algorithmic}
@@
 \label{algo:Encoding for top-up batching}
-\end{figure}
+\end{algorithm}
 
 Rather than decoding the same batch until all sentences in the batch are completed, the decoding step processing the same batch continuously.
-New sentences are added to the batch as old sentences completes, Figure~\ref{algo:Decoding for top-up batching}.
+New sentences are added to the batch as old sentences completes, Algorithm~\ref{algo:Decoding for top-up batching}.
 
-\begin{figure} [h]
+\begin{algorithm}
 \begin{algorithmic}
@@
 \label{algo:Decoding for top-up batching}
-\end{figure}
+\end{algorithm}
 

### 2018-02-06 fast-beam/fast-beam.tex
fast beam paper

@@
    \item \vspace{-2 mm} applying the activation function, most commonly softmax $ p_i = \exp(p_i) / \sum \exp(p_i) $
-   \item \vspace{-2 mm} a search for the best output class, and probability if necessary $\argmax_i p_i$
+   \item \vspace{-2 mm} a beam search for the best (or n-best) output classes $\argmax_i p_i$
 \end{enumerate}
@@
 \While{more input} 
-  \State Create batch
-  \State Encode(batch)
+  \State Create batch $b$
+  \State Encode($b$)
   \While{batch is not empty} 
-    \State Decode(batch)
-    \For{each sentence in batch}
+    \State Decode($b$)
+    \ForAll{sentence $s$ in $b$}
       \If{translation is complete}
-        \State Remove sentence %from batch
+        \State Remove $s$ from $b$
       \EndIf
@@
   \State Encode($b$)
-  \State Add $b$ to queue
+  \State Add $b$ to queue $q$
 \EndWhile
@@
 
-\State create batch $b$ from queue
+\State create batch $b$ from queue $q$
 \While{$b$ is not empty}
   \State Decode($b$)
-  \State Replace completed sentences with new sentence from queue
+  \ForAll{sentence $s$ in $b$}
+    \If{translation is complete}
+      \State Replace $s$ with $s'$ from $q$
+    \EndIf
+  \EndFor
 \EndWhile
@@
 \end{figure}
+
+\begin{table}
+\begin{center}
+\small
+\begin{tabular}{|l|r|r|} \hline
+		& Baseline	& Fused \\ \hline
+\multicolumn{3}{|c|}{\emph{Beam size 1}}	\\ \hline	
+Multiplication 	& 25.62 	& 26.47 (+3.3\%) \\ \hline
+Add bias 	& 4.98		&  \\ 
+Softmax 	& 12.81		& 7.94 (-76.9\%)\\
+Beam search	& 16.64		&  \\ \hline
+\multicolumn{3}{|c|}{\emph{Beam size 5}}	\\ \hline	
+Multiplication 	& 112.9 	& 115.04 (+1.9\%) \\ \hline
+Add bias 	& 23.66		&  \\ 
+Softmax 	& 56.61		& 67.58 (-56.5\%)\\
+Beam search	& 75.12		&  \\ \hline
+\end{tabular}
+\end{center}
+\caption{Time taken in sec (Europarl)}
+\label{tab:fused-breakdown}
+\end{table}
 

### 2018-02-06 fast-beam/fast-beam.tex
fast beam paper

@@
 \caption{Time taken in sec (Europarl)}
-\label{tab:fused-breakdown}
+\label{tab:fused-breakdown-europarl}
+\end{table}
+
+\begin{table}
+\begin{center}
+\small
+\begin{tabular}{|l|r|r|} \hline
+		& Baseline	& Fused \\ \hline
+\multicolumn{3}{|c|}{\emph{Beam size 1}}	\\ \hline	
+Multiplication 	& 21.47 	& 21.94 (+2.2\%) \\ \hline
+Add bias 	& 33.59		&  \\ 
+Softmax 	& 8.83		& 5.70 (-89.9\%)\\
+Beam search	& 13.91		&  \\ \hline
+\multicolumn{3}{|c|}{\emph{Beam size 5}}	\\ \hline	
+Multiplication 	& 79.66 	& 80.01 (+4.4\%) \\ \hline
+Add bias 	& 14.95		&  \\ 
+Softmax 	& 36.86		& 42.91 (-64.4\%)\\
+Beam search	& 68.80		&  \\ \hline
+\end{tabular}
+\end{center}
+\caption{Time taken in sec (OpenSubtitles)}
+\label{tab:fused-breakdown-opensubtitles}
 \end{table}

### 2018-02-06 fast-beam/fast-beam.tex
fast beam paper

@@
 
-Fusing the softmax and beam search increase the speed of the output layer by 43\% for beam size 1, decreasing to 25\% for a beam of 10 when translating the Europarl dataset.
-This led to an overall increase in translation speed of up to 23\%, Figure~\ref{fig:beam-europarl}.
-Translation speed improved by up to 41\% when translating the OpenSubtitles dataset, Figure~\ref{fig:beam-opensubtitles}.
-Compared to translating the Europarl test st, Figure~\ref{fig:pie-time-europarl} in Section~\ref{sec:Introduction}, the amount of time taken by the output layer dominates translation time for the OpenSubtitles test set, Figure~\ref{fig:pie-time-opensubtitles}.
+Fusing the last 3 steps led to a substantial reduction in the time taken, especially for beam size of 1 where the softmax does not actually have to calculated, Table~\ref{tab:fused-breakdown-europarl} and ~\ref{tab:fused-breakdown-opensubtitles}.
+This led to an overall increase in translation speed of up to 23\% for the Europarl test set, Figure~\ref{fig:beam-europarl}, and up to 41\% for the OpenSubtitles test set, Figure~\ref{fig:beam-opensubtitles}.
 
-\begin{figure}
-\centering
-\begin{tabular}{cc}
-{\includegraphics[scale=0.5]{beam-europarl.png}} 
-\end{tabular}
-\caption{Speed using the fused kernel (Europarl)}
-\label{fig:beam-europarl}
-\end{figure}
-
-\begin{figure}
-\centering
-\begin{tabular}{cc}
-{\includegraphics[scale=0.5]{beam-opensubtitles.png}} 
-\end{tabular}
-\caption{Speed using the fused kernel (OpenSubtitles)}
-\label{fig:beam-opensubtitles}
-\end{figure}
-
-\begin{figure}
-\centering
-\begin{tabular}{cc}
-{\includegraphics[scale=0.3]{pie-time-opensubtitles.png}} 
-\end{tabular}
-\caption{Proportion of time spent during translation (OpenSubtitles)}
-\label{fig:pie-time-opensubtitles}
-\end{figure}
+The amount of time taken by the output layer dominates translation time for the OpenSubtitles test set, Figure~\ref{fig:pie-time-opensubtitles}.
 
@@
 \end{table}
+
+\begin{figure}
+\centering
+\begin{tabular}{cc}
+{\includegraphics[scale=0.5]{beam-europarl.png}} 
+\end{tabular}
+\caption{Speed using the fused kernel (Europarl)}
+\label{fig:beam-europarl}
+\end{figure}
+
+\begin{figure}
+\centering
+\begin{tabular}{cc}
+{\includegraphics[scale=0.5]{beam-opensubtitles.png}} 
+\end{tabular}
+\caption{Speed using the fused kernel (OpenSubtitles)}
+\label{fig:beam-opensubtitles}
+\end{figure}
+
+\begin{figure}
+\centering
+\begin{tabular}{cc}
+{\includegraphics[scale=0.3]{pie-time-opensubtitles.png}} 
+\end{tabular}
+\caption{Proportion of time spent during translation (OpenSubtitles)}
+\label{fig:pie-time-opensubtitles}
+\end{figure}
 

### 2018-02-06 fast-beam/fast-beam.tex
fast beam paper

@@
 
-The amount of time taken by the output layer dominates translation time for the OpenSubtitles test set, Figure~\ref{fig:pie-time-opensubtitles}.
+The amount of time taken by the output layer dominates translation time for the OpenSubtitles test set, Figure~\ref{fig:pie-time-opensubtitles}, explaining the bigger overall translation speed with the fused kernel.
 

### 2018-02-06 fast-beam/fast-beam.tex
fast beam paper

@@
 
-Fusing the last 3 steps led to a substantial reduction in the time taken, especially for beam size of 1 where the softmax does not actually have to calculated, Table~\ref{tab:fused-breakdown-europarl} and ~\ref{tab:fused-breakdown-opensubtitles}.
+Fusing the last 3 steps led to a substantial reduction in the time taken, especially for beam size of 1 where the softmax probability does not actually have to calculated, Table~\ref{tab:fused-breakdown-europarl} and ~\ref{tab:fused-breakdown-opensubtitles}.
 This led to an overall increase in translation speed of up to 23\% for the Europarl test set, Figure~\ref{fig:beam-europarl}, and up to 41\% for the OpenSubtitles test set, Figure~\ref{fig:beam-opensubtitles}.

### 2018-02-06 fast-beam/fast-beam.tex
fast beam paper

@@
 Secondly, rather than a single label, the output from machine translation models is a sequence of class labels that makes up the words or sub-words in the target sentence.
-The sentence lengths are unpredictable, leading to inefficiencies during batch processing.
-We provide solutions for each of these issues which can increase batched inference speed by up to 57\% on modern GPUs without affecting model quality.
+Target sentence lengths are unpredictable, leading to inefficiencies during batch processing.
+We provide solutions for each of these issues which together increase batched inference speed by up to 57\% on modern GPUs without affecting model quality.
 
@@
 \begin{algorithmic}
-\Procedure{Fused Kernel}{vector $p$, bias vector $b$}
+\Procedure{Fused-Kernel}{vector $p$, bias vector $b$}
 

### 2018-02-06 fast-beam/fast-beam.tex
fast beam paper

@@
 We shall leave step 1 for future work and focus on the last three steps, the outline for which are shown in Algorithm~\ref{algo:original-softmax-beamsearch}.
+For brevity, we show the algorithm for 1-best, a beam search of the n-best is a simple extension of this.
 
@@
 \end{algorithmic}
-\caption{Original Softmax and Beam Search Algorithm}
+\caption{Original softmax and beam Search Algorithm}
 \label{algo:original-softmax-beamsearch}
@@
 
-Thirdly, we are only interested in the best class.
-Since the best class is known, we can avoid calculating softmax for all classes.
+Thirdly, we are only interested in the probabilities of the best classes.
+Since these classes are known, we can avoid calculating softmax for all classes.
 The outline of the our function is shown in Algorithm~\ref{algo:Fused Kernel}.
@@
 
-\caption{Fused softmax and argmax}
+\caption{Fused softmax and beam search}
 \label{algo:Fused Kernel}

### 2018-02-06 fast-beam/fast-beam.tex
fast beam paper

@@
 We shall leave step 1 for future work and focus on the last three steps, the outline for which are shown in Algorithm~\ref{algo:original-softmax-beamsearch}.
-For brevity, we show the algorithm for 1-best, a beam search of the n-best is a simple extension of this.
+For brevity, we show the algorithm for 1-best, a beam search of the n-bests is a simple extension of this.
 
@@
 
-It has been shown ~\citep{koehn-knowles:2017:NMT} that using beam search to use the n-best number of classes, rather than just the best class, improves translation quality.
-This is a simple extension to the algorithm of Algorithm~\ref{algo:Fused Kernel}.
+It has been shown ~\citep{koehn-knowles:2017:NMT} that using beam search to use the n-best number of classes, rather than just the best class, improves translation quality, up to a point.
 Unlike the 1-best case, however, the softmax calculation cannot be skipped as the denominator differs for each input.

### 2018-02-06 fast-beam/fast-beam.tex
fast beam paper

@@
 
-It has been shown ~\citep{koehn-knowles:2017:NMT} that using beam search to use the n-best number of classes, rather than just the best class, improves translation quality, up to a point.
+It has been noted by ~\citep{koehn-knowles:2017:NMT} and others that using n-best number of classes, rather than just the best class, improves translation quality.
 Unlike the 1-best case, however, the softmax calculation cannot be skipped as the denominator differs for each input.

### 2018-02-06 fast-beam/fast-beam.tex
fast beam paper

@@
 \begin{enumerate}
-   \item \vspace{-2 mm} multiplication of the input with the weight vector $p = w x$
+   \item \vspace{-2 mm} multiplication of the weight matrix with the input vector $p = w x$
    \item \vspace{-2 mm} addition of a bias term to the resulting scores $p = p + b$
@@
 As can be seen, the operations iterate over the matrix p five times - once to add the bias, three times to calculate the softmax, and once to search for the best class.
-We shall use a popular method, kernel fusion~\citep{Guevara2009EnablingTP}, to optimize the output layer.
+We fuse the three operation into one kernel, a popular optimization technique~\citep{Guevara2009EnablingTP}.
 
@@
 \begin{algorithmic}
-\Procedure{Find-1-Best}{vector $p$, bias vector $b$}
+\Procedure{Fused-kernel-1-best}{vector $p$, bias vector $b$}
 
@@
 \end{algorithmic}
-\caption{Find 1 best only}
+\caption{Find 1-best only}
 \label{algo:Argmax only}
@@
 \Procedure{Mini-Batching}{}
-%\REQUIRE source sentence $s$, translation options
 \While{more input} 

### 2018-02-06 fast-beam/fast-beam.tex
fast beam paper

@@
 
-We shall leave step 1 for future work and focus on the last three steps, the outline for which are shown in Algorithm~\ref{algo:original-softmax-beamsearch}.
+We leave step 1 for future work and focus on the last three steps, the outline for which are shown in Algorithm~\ref{algo:original-softmax-beamsearch}.
 For brevity, we show the algorithm for 1-best, a beam search of the n-bests is a simple extension of this.

### 2018-02-08 fast-beam/fast-beam.tex
fast beam paper

@@
 Our paper follows most closely on from ~\cite{DBLP:conf/emnlp/Devlin17} which achieved faster NMT inference mainly by novel implementation of an existing model.
-However, we differ by focusing on GPU implementation, and the output layer and batching algorithm which were not touched on by the previous work.
+However, we differ by focusing on GPU implementation, specifically the output layer, and the batching algorithm which were not touched on by the previous work.
 
@@
 In fact, we are usually only interested in the best class during inference, not the probability.
-Therefore, we can skip the second iteration over $p$ in Algorithm~\ref{algo:Fused Kernel} and avoid computing the softmax altogether, Algorithm~\ref{algo:Argmax only}.
+Since we now calculate the best class before the softmax, we can skip softmax altogether, Algorithm~\ref{algo:Argmax only}.
+However, is this only possible for beam size 1; the comparison between softmax probobilities is required for larger beam sizes as the denominator is different in different beams.
 
@@
 \end{algorithm}
-
-It has been noted by ~\citep{koehn-knowles:2017:NMT} and others that using n-best number of classes, rather than just the best class, improves translation quality.
-Unlike the 1-best case, however, the softmax calculation cannot be skipped as the denominator differs for each input.
 

### 2018-02-08 fast-beam/fast-beam.tex
fast beam paper

@@
 However, the number of classes in NMT models is typically in the tens or hundreds of thousands, corresponding to the vocabulary size of the output language.
-For example, ~\cite{sennrich-haddow-birch:2016:P16-12} experimented with target vocabulary sizes of 60,000 and 90,000 sub-word units.
+For example, ~\citet{sennrich-haddow-birch:2016:P16-12} experimented with target vocabulary sizes of 60,000 and 90,000 sub-word units.
 This makes the output layer of NMT models very computationally expensive.
@@
 
-We base our work on the sequence-to-sequence model of ~\cite{D14-1179}. We also choose to focus on the using NMT inference on GPUs. % rather than CPUs as pursued by ~\cite{DBLP:conf/emnlp/Devlin17}.
+We base our work on the sequence-to-sequence model of ~\citet{D14-1179}. We also choose to focus on the using NMT inference on GPUs. % rather than CPUs as pursued by ~\cite{DBLP:conf/emnlp/Devlin17}.
 
@@
 
-The speed of the softmax layer when used with large vocabularies have been looked at in~\citep{DBLP:journals/corr/GraveJCGJ16} but there have been many other attempts at faster softmax, for example~\cite{DBLP:journals/corr/abs-1301-3781}, ~\cite{Zoph-2016}.
+The speed of the softmax layer when used with large vocabularies have been looked at in~\citet{DBLP:journals/corr/GraveJCGJ16} but there have been many other attempts at faster softmax, for example~\citet{DBLP:journals/corr/abs-1301-3781}, ~\citet{Zoph-2016}.
 
 There has been surprisingly little work on batching algorithms, considering its critical importance for efficient deep-learning training and inference.
-~\cite{Neubig-autobatching} describe an novel batching algorithm, however, its aim is to alleviate the burden on developers in batching of new models, not faster batching.
+~\citet{Neubig-autobatching} describe an novel batching algorithm, however, its aim is to alleviate the burden on developers of batching models rather than faster batching.
 
-Our paper follows most closely on from ~\cite{DBLP:conf/emnlp/Devlin17} which achieved faster NMT inference mainly by novel implementation of an existing model.
+Our paper follows most closely on from ~\citet{DBLP:conf/emnlp/Devlin17} which achieved faster NMT inference mainly by novel implementation of an existing model.
 However, we differ by focusing on GPU implementation, specifically the output layer, and the batching algorithm which were not touched on by the previous work.
@@
     \ForAll{sentence $s$ in $b$}
-      \If{translation is complete}
+      \If{trans($s$) is complete}
         \State Remove $s$ from $b$
@@
   \ForAll{sentence $s$ in $b$}
-    \If{translation is complete}
+    \If{trans($s$) is complete}
       \State Replace $s$ with $s'$ from $q$
@@
 
-We trained a sequence-to-sequence, encoder-decoder NMT system similar to that described in ~\cite{sennrich-haddow-birch:2016:P16-12}.
+We trained a sequence-to-sequence, encoder-decoder NMT system similar to that described in ~\citet{sennrich-haddow-birch:2016:P16-12}.
 This uses recurrent neural networks with gated recurrent units.

### 2018-02-08 fast-beam/fast-beam.tex
fast beam paper

@@
 Since we now calculate the best class before the softmax, we can skip softmax altogether, Algorithm~\ref{algo:Argmax only}.
-However, is this only possible for beam size 1; the comparison between softmax probobilities is required for larger beam sizes as the denominator is different in different beams.
+However, this is only possible for beam size 1; the comparison between softmax probobilities is required for larger beam sizes as the denominator is different in different beams.
 

### 2018-02-09 fast-beam/fast-beam.tex
fast beam paper

@@
 Simpler, faster models have been created that are as good, or almost as good, as more slower, more complex models~\citep{DBLP:journals/corr/BahdanauCB14}.
-Research have also gone into smaller, faster models that can approximate the original, slower, bigger model~\citep{DBLP:conf/emnlp/KimR16}.
-Some models have been justified as more suited for the parallel architecture of GPUs~\cite{DBLP:journals/corr/VaswaniSPUJGKP17}.
+Research have also gone into smaller, faster models that can approximate slower, bigger models~\citep{DBLP:conf/emnlp/KimR16}.
+Some models have been justified as more suited for the parallel architecture of GPUs~\cite{DBLP:journals/corr/VaswaniSPUJGKP17}~\cite{gehring2017convs2s}.
 

### 2018-02-16 fast-beam/fast-beam.tex
fast beam paper
fast paper

@@
 \ForAll{$p_i$ in $p$}
-  \State $p_i \gets \frac{\exp(p_i) - max}{sum} $
+  \State $p_i \gets \frac{\exp(p_i - max)}{sum} $
 \EndFor
@@
 
-\Procedure{Find-Best}{softmax vector $p$}
+\Procedure{Find-Best}{vector $p$}
 
@@
 
+\Return $max$, $best$
+
 \EndProcedure
@@
 
-As can be seen, the operations iterate over the matrix p five times - once to add the bias, three times to calculate the softmax, and once to search for the best class.
-We fuse the three operation into one kernel, a popular optimization technique~\citep{Guevara2009EnablingTP}.
+As can be seen, the matrix p is iterated over five times - once to add the bias, three times to calculate the softmax, and once to search for the best class.
+We fuse the three functions into one kernel, a popular optimization technique~\citep{Guevara2009EnablingTP}, making use of the following observations.
 
-Secondly, we make use of the fact that softmax and $\exp$ are monotonic functions therefore we can move the search for the best class from FIND-BEST to SOFTMAX.
+Firstly, softmax and $\exp$ are monotonic functions, therefore, we can move the search for the best class from FIND-BEST to SOFTMAX, at the start of the kernel.
 
-Thirdly, we are only interested in the probabilities of the best classes.
-Since these classes are known, we can avoid calculating softmax for all classes.
+Secondly, we are only interested in the probabilities of the best classes.
+Since they are now known at the start of the kernel, we compute softmax only for those classes.
 The outline of the our function is shown in Algorithm~\ref{algo:Fused Kernel}.

### 2018-02-22 fast-beam/fast-beam.tex
fast

@@
 
-We will examine two areas that are critical to fast NMT inference where the models differ significantly from other deep-learning models.
+We examine two areas that are critical to fast NMT inference where the models differ significantly from other deep-learning models.
 We believe the optimizations of these models have been overlooked by the general deep-learning community.
@@
 
-We base our work on the sequence-to-sequence model of ~\citet{D14-1179}. We also choose to focus on the using NMT inference on GPUs. % rather than CPUs as pursued by ~\cite{DBLP:conf/emnlp/Devlin17}.
-
 \section{Prior Work}
@@
 Our paper follows most closely on from ~\citet{DBLP:conf/emnlp/Devlin17} which achieved faster NMT inference mainly by novel implementation of an existing model.
-However, we differ by focusing on GPU implementation, specifically the output layer, and the batching algorithm which were not touched on by the previous work.
+However, we differ by focusing on GPU implementation, specifically the output layer, and the batching algorithm which was not touched on by the previous work.
+Our model is based on that of ~\citet{D14-1179} but our work is applicable to other models and applications.
+% rather than CPUs as pursued by ~\cite{DBLP:conf/emnlp/Devlin17}.
 
@@
 
-In models with a small number of classes such as binary classification, the computational effort required is trivial and fast.
-However, this is not the case for large number of classes such as those found in NMT models.
+In models with a small number of classes such as binary classification, the computational effort required is trivial and fast but this is not the case for the large number of classes found in typical NMT models.
 

### 2018-02-22 fast-beam/fast-beam.tex
fast

@@
 However, we differ by focusing on GPU implementation, specifically the output layer, and the batching algorithm which was not touched on by the previous work.
-Our model is based on that of ~\citet{D14-1179} but our work is applicable to other models and applications.
+Our model is based on that of ~\citet{D14-1179} and ~\citet{sennrich-haddow-birch:2016:P16-12} but our work is applicable to other models and applications.
 % rather than CPUs as pursued by ~\cite{DBLP:conf/emnlp/Devlin17}.
@@
 
-As can be seen, the matrix p is iterated over five times - once to add the bias, three times to calculate the softmax, and once to search for the best class.
-We fuse the three functions into one kernel, a popular optimization technique~\citep{Guevara2009EnablingTP}, making use of the following observations.
+As can be seen, the matrix p is iterated over five times - once to add the bias, three times to calculate the softmax, and once to search for the best classes.
+We propose fusing the three functions into one kernel, a popular optimization technique~\citep{Guevara2009EnablingTP}, making use of the following observations.
 
@@
 Our proposed top-up batching algorithm encode and decode asynchronously.
-The encoding step, Algorithm~\ref{algo:Encoding for top-up batching}, is similar to the main loop of the standard algorithm but the results are added to a queue to be consumed by the decoding step later.
+The encoding step, Algorithm~\ref{algo:Encoding for top-up batching}, is similar to the main loop of the standard algorithm but the results are added to a queue to be consumed by the decoding step.
 

### 2018-02-22 fast-beam/fast-beam.tex
fast

@@
 Multiplication 	& 21.47 	& 21.94 (+2.2\%) \\ \hline
-Add bias 	& 33.59		&  \\ 
-Softmax 	& 8.83		& 5.70 (-89.9\%)\\
+Add bias 	& 3.29		&  \\ 
+Softmax 	& 8.83		& 5.70 (-78.1\%)\\
 Beam search	& 13.91		&  \\ \hline

### 2018-02-22 fast-beam/fast-beam-short.tex
fast

--- 2018-02-22 fast-beam/fast-beam.tex
@@
 
-As neural machine translation (NMT) models have become the new state-of-the-art, the challenge is to make their deployment efficient and economical.
-We have identified two areas where typical NMT models differ significantly from other deep-learning models that impact on their inference speed.
-Firstly, NMT models usually contains a large number of output classes, corresponding to the output vocabulary.
-Secondly, rather than a single label, the output from machine translation models is a sequence of class labels that makes up the words or sub-words in the target sentence.
-Target sentence lengths are unpredictable, leading to inefficiencies during batch processing.
-We provide solutions for each of these issues which together increase batched inference speed by up to 57\% on modern GPUs without affecting model quality.
+We present two optimizations in areas where typical NMT models differ significantly from other deep-learning models.
+Firstly, we optimize the output layer which takes a significant amount of time for NMT models due to the large number of output classes, corresponding to the output vocabulary.
+Secondly, we present a novel batching algorithm which takes into account the differing output sentence lengths which leads to inefficiencies in current batch algorithms.
+Together, we increase inference speed by up to 57\% on modern GPUs without affecting model quality.
 
@@
 
-We examine two areas that are critical to fast NMT inference where the models differ significantly from other deep-learning models.
-We believe the optimizations of these models have been overlooked by the general deep-learning community.
-
-Firstly, the number of classes in many deep-learning applications is small.
-However, the number of classes in NMT models is typically in the tens or hundreds of thousands, corresponding to the vocabulary size of the output language.
-For example, ~\citet{sennrich-haddow-birch:2016:P16-12} experimented with target vocabulary sizes of 60,000 and 90,000 sub-word units.
-This makes the output layer of NMT models very computationally expensive.
-Figure~\ref{fig:pie-time-europarl} shows the breakdown of amount of time during translation our NMT system; nearly 70\% of the time is involved in the output layer.
-We will look at optimizations which explicitly target the output layer.
+The number of classes in NMT models is typically in the tens or hundreds of thousands, for example, ~\citet{sennrich-haddow-birch:2016:P16-12} experimented with target vocabulary sizes of 60,000 and 90,000 sub-word units.
+This makes the output layer of NMT models computationally expensive.
+Figure~\ref{fig:pie-time-opensubtitles} shows the breakdown of amount of time during translation our NMT system; 73\% of the time is involved in the output layer or beam search.
+Our first proposal will explicitly target this computation.
 \begin{figure}
@@
 \begin{tabular}{cc}
-{\includegraphics[scale=0.3]{pie-time-europarl.png}} 
+{\includegraphics[scale=0.3]{pie-time-opensubtitles.png}} 
 \end{tabular}
-\caption{Proportion of time spent during translation (Europarl test set)}
-\label{fig:pie-time-europarl}
+\caption{Proportion of time spent during translation}
+\label{fig:pie-time-opensubtitles}
 \end{figure}
@@
 pre-sorting sentences by length before creating mini-batches with similar length source sentences.
-However, maxi-batching can introduce unacceptable delay in response for online applications as the input are not processed in the order they came.
-It is also not appropriate for applications where only the output lengths varies, for example, image captioning.
+However, maxi-batching can introduce unacceptable delays in response time as inputs are not processed in the order they came, a particular concern for online applications.
 
-Even with sequence-to-sequence NMT models, target sentence lengths will still differ even for similar length inputs, compromising decoding performance.
+Even with sequence-to-sequence NMT models, target sentence lengths will still differ even for similar length inputs.
 Figure~\ref{fig:batch-size} shows the actual batch size during decoding with a maximum batch size of 128; the batch was full in only 42\% of decoding iterations.
+We will propose an alternative batching algorithm to increase the actual batch size during decoding without the problems associated with mini-batching and maxi-batching.
 
@@
 
-We will propose an alternative batching algorithm to increase the actual batch size during decoding without the problems associated with mini-batching and maxi-batching.
-
-\section{Prior Work}
-
-Deep-learning models has been successfully used on many machine learning task in recent years in areas as diverse as computer vision and natural language processing.
+% \section{Prior Work}
+% 
+% Deep-learning models has been successfully used on many machine learning task in recent years in areas as diverse as computer vision and natural language processing.
 This success have been followed by the rush to put these model into production.
 The computational resources required in order to run these models has been challenging but, fortunately, there has been a lot work to meet this challenge.
-
-Hardware accelerators such as GPUs are very popular but other specialized hardware such as custom processors~\citep{DBLP:journals/corr/JouppiYPPABBBBB17} and FPGA~\citep{DBLP:journals/corr/LaceyTA16} have been used.
+% 
+% Hardware accelerators such as GPUs are very popular but other specialized hardware such as custom processors~\citep{DBLP:journals/corr/JouppiYPPABBBBB17} and FPGA~\citep{DBLP:journals/corr/LaceyTA16} have been used.
 Hardware-supported reduced precision is also an ideal way to speed up model inference~\citep{DBLP:journals/corr/abs-1710-03740}.
-
-Simpler, faster models have been created that are as good, or almost as good, as more slower, more complex models~\citep{DBLP:journals/corr/BahdanauCB14}.
+% 
+% Simpler, faster models have been created that are as good, or almost as good, as more slower, more complex models~\citep{DBLP:journals/corr/BahdanauCB14}.
 Research have also gone into smaller, faster models that can approximate slower, bigger models~\citep{DBLP:conf/emnlp/KimR16}.
 Some models have been justified as more suited for the parallel architecture of GPUs~\cite{DBLP:journals/corr/VaswaniSPUJGKP17}~\cite{gehring2017convs2s}.
-
-The speed of the softmax layer when used with large vocabularies have been looked at in~\citet{DBLP:journals/corr/GraveJCGJ16} but there have been many other attempts at faster softmax, for example~\citet{DBLP:journals/corr/abs-1301-3781}, ~\citet{Zoph-2016}.
-
-There has been surprisingly little work on batching algorithms, considering its critical importance for efficient deep-learning training and inference.
+% 
+% The speed of the softmax layer when used with large vocabularies have been looked at in~\citet{DBLP:journals/corr/GraveJCGJ16} but there have been many other attempts at faster softmax, for example~\citet{DBLP:journals/corr/abs-1301-3781}, ~\citet{Zoph-2016}.
+% 
+% There has been surprisingly little work on batching algorithms, considering its critical importance for efficient deep-learning training and inference.
 ~\citet{Neubig-autobatching} describe an novel batching algorithm, however, its aim is to alleviate the burden on developers of batching models rather than faster batching.
-
-Our paper follows most closely on from ~\citet{DBLP:conf/emnlp/Devlin17} which achieved faster NMT inference mainly by novel implementation of an existing model.
+% 
+% Our paper follows most closely on from ~\citet{DBLP:conf/emnlp/Devlin17} which achieved faster NMT inference mainly by novel implementation of an existing model.
 However, we differ by focusing on GPU implementation, specifically the output layer, and the batching algorithm which was not touched on by the previous work.
@@
 
-We leave step 1 for future work and focus on the last three steps, the outline for which are shown in Algorithm~\ref{algo:original-softmax-beamsearch}.
+We focus on the last three steps, the outline for which are shown in Algorithm~\ref{algo:original-softmax-beamsearch}.
 For brevity, we show the algorithm for 1-best, a beam search of the n-bests is a simple extension of this.
@@
 In fact, we are usually only interested in the best class during inference, not the probability.
-Since we now calculate the best class before the softmax, we can skip softmax altogether, Algorithm~\ref{algo:Argmax only}.
-However, this is only possible for beam size 1; the comparison between softmax probobilities is required for larger beam sizes as the denominator is different in different beams.
-
-\begin{algorithm}
-\begin{algorithmic}
-\Procedure{Fused-kernel-1-best}{vector $p$, bias vector $b$}
-
-\State $max \gets - \infty$ 
-\ForAll{$p_i$ in $p$}
-  \If{$p_i + b_i > max$}
-    \State $max \gets p_i + b_i$
-    \State $best \gets i$
-  \EndIf
-\EndFor
-
-\Return $best$
-
-\EndProcedure
-
-\end{algorithmic}
-\caption{Find 1-best only}
-\label{algo:Argmax only}
-\end{algorithm}
+Since we now compute the best class before the softmax, we can skip softmax altogether.
+This is only possible for beam size 1; the comparison between softmax probabilities is required for larger beam sizes as the denominators are different for different hypotheses.
 
@@
 
-The standard mini-batching algorithm is outlined in Algorithm~\ref{algo:Mini-batching}.
+The standard mini-batching algorithm, Algorithm~\ref{algo:Mini-batching}, encode the sentences for a batch, followed by decoding the batch.
+Decoding stop only once all sentences in the batch are completed.
 
@@
 
-This algorithm encode the sentences for a batch, followed by decoding the batch.
-The decoding stop once all sentences in the batch are completed.
-This is a potential inefficiency as the number of remaining sentences may not be optimal.
-
-We will focus on decoding as this is the more compute-intensive step, and issues with differing sentence sizes in encoding can partly be ameliorated by maxi-batching.
-
-Our proposed top-up batching algorithm encode and decode asynchronously.
+Our proposal focuses on decoding as this is the more compute-intensive task, encoding and decoding asynchronously.
 The encoding step, Algorithm~\ref{algo:Encoding for top-up batching}, is similar to the main loop of the standard algorithm but the results are added to a queue to be consumed by the decoding step.
@@
 
-Rather than decoding the same batch until all sentences in the batch are completed, the decoding step processing the same batch continuously.
-New sentences are added to the batch as old sentences completes, Algorithm~\ref{algo:Decoding for top-up batching}.
+Rather than decoding a batch until completion, new sentences are added to the batch as old sentences completes, Algorithm~\ref{algo:Decoding for top-up batching}.
 
@@
 
-We trained a sequence-to-sequence, encoder-decoder NMT system similar to that described in ~\citet{sennrich-haddow-birch:2016:P16-12}.
+We train a sequence-to-sequence, encoder-decoder NMT system similar to that described in ~\citet{sennrich-haddow-birch:2016:P16-12}.
 This uses recurrent neural networks with gated recurrent units.
-The input and output vocabulary size were both set to 85,000 sub-words using byte-pair encoding (BPE) to adjust the vocabulary to the desired size.
-The hidden layer dimensions was set to 512.
+The input and output vocabulary size are both set to 85,000 sub-words using byte-pair encoding (BPE), the hidden layer dimensions is 512.
 %We used the Marian toolkit to train our models.
@@
 
-Our training data consisted of the German-English parallel sentences from the Europarl corpus~\citep{Koehn:2005:MTS}. To test inference speed, we used two test sets with differing characteristics:
-\begin{enumerate}
-   \item \vspace{-2 mm} a subset of the Europarl training data, which contains mostly long sentences, and is, of course, in the same domain as the training data
-   \item \vspace{-2 mm} a subset of the German-English data from the Open-Subtitles corpus, consisting of mostly short, out-of-domain sentences.
-\end{enumerate}
-Table~\ref{tab:corpora} gives further details of the test sets.
+We trained a German-English with data from the Europarl corpus~\citep{Koehn:2005:MTS}, our test set consisted of a subset of the Open-Subtitles corpus, Table~\ref{tab:corpora}.
 
@@
 \begin{tabular}{|l|r|r|} \hline
-		  & Europarl		& OpenSubtitles \\ \hline
-\# sentences  	  & 30,000 		& 50,000 \\
-\# sub-words 	  & 787,908 		& 467,654 \\ 
-Avg sub-words/sent & 26.3		& 9.4 \\
-Std dev subwords/sent & 14.9		& 6.1 \\ \hline
+		 & OpenSubtitles \\ \hline
+\# sentences  	 & 50,000 \\
+\# sub-words 	 & 467,654 \\ 
+Avg sub-words/sent & 9.4 \\
+Std dev subwords/sent & 6.1 \\ \hline
 \end{tabular}
@@
 
-Fusing the last 3 steps led to a substantial reduction in the time taken, especially for beam size of 1 where the softmax probability does not actually have to calculated, Table~\ref{tab:fused-breakdown-europarl} and ~\ref{tab:fused-breakdown-opensubtitles}.
-This led to an overall increase in translation speed of up to 23\% for the Europarl test set, Figure~\ref{fig:beam-europarl}, and up to 41\% for the OpenSubtitles test set, Figure~\ref{fig:beam-opensubtitles}.
-
-The amount of time taken by the output layer dominates translation time for the OpenSubtitles test set, Figure~\ref{fig:pie-time-opensubtitles}, explaining the bigger overall translation speed with the fused kernel.
-
-\begin{table}
-\begin{center}
-\small
-\begin{tabular}{|l|r|r|} \hline
-		& Baseline	& Fused \\ \hline
-\multicolumn{3}{|c|}{\emph{Beam size 1}}	\\ \hline	
-Multiplication 	& 25.62 	& 26.47 (+3.3\%) \\ \hline
-Add bias 	& 4.98		&  \\ 
-Softmax 	& 12.81		& 7.94 (-76.9\%)\\
-Beam search	& 16.64		&  \\ \hline
-\multicolumn{3}{|c|}{\emph{Beam size 5}}	\\ \hline	
-Multiplication 	& 112.9 	& 115.04 (+1.9\%) \\ \hline
-Add bias 	& 23.66		&  \\ 
-Softmax 	& 56.61		& 67.58 (-56.5\%)\\
-Beam search	& 75.12		&  \\ \hline
-\end{tabular}
-\end{center}
-\caption{Time taken in sec (Europarl)}
-\label{tab:fused-breakdown-europarl}
-\end{table}
+Fusing the last 3 steps led to a substantial reduction in the time taken, especially for beam size of 1 where the softmax probability does not actually have to calculated, Table~\ref{tab:fused-breakdown-opensubtitles}.
+This led to an overall increase in translation speed of up to 41\%, Figure~\ref{fig:beam-opensubtitles}.
 
@@
 \begin{tabular}{cc}
-{\includegraphics[scale=0.5]{beam-europarl.png}} 
-\end{tabular}
-\caption{Speed using the fused kernel (Europarl)}
-\label{fig:beam-europarl}
-\end{figure}
-
-\begin{figure}
-\centering
-\begin{tabular}{cc}
 {\includegraphics[scale=0.5]{beam-opensubtitles.png}} 
 \end{tabular}
-\caption{Speed using the fused kernel (OpenSubtitles)}
+\caption{Speed using the fused kernel}
 \label{fig:beam-opensubtitles}
-\end{figure}
-
-\begin{figure}
-\centering
-\begin{tabular}{cc}
-{\includegraphics[scale=0.3]{pie-time-opensubtitles.png}} 
-\end{tabular}
-\caption{Proportion of time spent during translation (OpenSubtitles)}
-\label{fig:pie-time-opensubtitles}
 \end{figure}
@@
 The top-up batching and maxi-batching have similar goals of maximizing efficiency when translating batches of different lengths.
-Therefore, using both methods together gives limited gains, in fact, using top-up batching with maxi-batch slows of between 2\% to 18\% when translating the Europarl test set due to the overhead of using the algorithm, Figure~\ref{fig:topup-europarl}.
-When maxi-batching is inappropriate, using top-up batching alone matches the performance of maxi-batching, even being slightly faster when a small beam is used.
-
-\begin{figure}
-\centering
-\begin{tabular}{cc}
-{\includegraphics[scale=0.3]{topup-europarl.png}} 
-\end{tabular}
-\caption{Speed using top-up batching (Europarl)}
-\label{fig:topup-europarl}
-\end{figure}
-
-The results are better when translating the OpenSubtitles test set, Figure~\ref{fig:topup-opensubtitles}.
-The top-up batching does not harm performance when used with maxi-batching, even helping a little for small beams.
-However, top-up batching increases translation speed by up to 12\% when used alone.
+The top-up batching is of limited utility when used with maxi-batching but it increases translation speed by up to 12\% when used alone, Figure~\ref{fig:topup-opensubtitles}.
 
@@
 \end{tabular}
-\caption{Speed using top-up batching (OpenSubtitles)}
+\caption{Speed using top-up batching}
 \label{fig:topup-opensubtitles}
-\end{figure}
-
-The gain from top-up batching is partially dependent on how much time the standard mini-batching algorithm spend decoding with a very small number of sentences as this is does not fully utilize the GPU cores.
-From Figure~\ref{fig:batch-size-small}, 20\% of the decoding iterations have less than 8 sentences remaining in the batch when translating the Europarl test test with maxi-batching.
-This figure is 50\% for the OpenSubtitles test set with no maxi-batching.
-The OpenSubtitles test set has a higher variance of sentence lengths relative to its average sentence length, which forces the mini-batch algorithm to continue decoding with a small number of sentences while the other sentences have already completed.
-
-\begin{figure}
-\centering
-\begin{tabular}{cc}
-{\includegraphics[scale=0.3]{batch-size-small.png}} 
-\end{tabular}
-\caption{Actual batch size during decoding}
-\label{fig:batch-size-small}
 \end{figure}
@@
 
-Using both the fused kernel and top-up batching to translate led to a cummalative speed improvement of up to 57\% and 34\% for the OpenSubtitles and Europarl test set, respectively, when no maxi-batching is used, Figure~\ref{fig:cummulative-europarl} and Figure~\ref{fig:cummulative-opensubtitles}.
-With maxi-batching, the speed was up to 41\% and 21\%.
-
-\begin{figure}
-\centering
-\begin{tabular}{cc}
-{\includegraphics[scale=0.3]{cummulative-europarl.png}} 
-\end{tabular}
-\caption{Cummulative results (Europarl}
-\label{fig:cummulative-europarl}
-\end{figure}
+Using both the fused kernel and top-up batching to translate led to a cummalative speed improvement of up to 57\% when no maxi-batching is used, Figure~\ref{fig:cummulative-opensubtitles}.
+With maxi-batching, the speed was up to 41\% faster.
 
@@
 \end{tabular}
-\caption{Cummulative results (OpenSubtitles}
+\caption{Cummulative results}
 \label{fig:cummulative-opensubtitles}

### 2018-02-22 fast-beam/fast-beam.tex
fast

@@
 In fact, we are usually only interested in the best class during inference, not the probability.
-Since we now calculate the best class before the softmax, we can skip softmax altogether, Algorithm~\ref{algo:Argmax only}.
-However, this is only possible for beam size 1; the comparison between softmax probobilities is required for larger beam sizes as the denominator is different in different beams.
+Since we now compute the best class before the softmax, we can skip softmax altogether, Algorithm~\ref{algo:Argmax only}.
+This is only possible for beam size 1; the comparison between softmax probabilities is required for larger beam sizes as the denominators are different for different hypotheses.
 

### 2018-02-22 fast-beam/fast-beam-short.tex
fast

@@
 Our first proposal will explicitly target this computation.
-\begin{figure}
+\begin{figure}[h]
 \centering
@@
 
-Even with sequence-to-sequence NMT models, target sentence lengths will still differ even for similar length inputs.
+Even with maxi-batching, target sentence lengths will still differ even for similar length inputs.
 Figure~\ref{fig:batch-size} shows the actual batch size during decoding with a maximum batch size of 128; the batch was full in only 42\% of decoding iterations.
@@
 
-\begin{figure}
+\begin{figure}[h]
 \centering
@@
 \begin{enumerate}
-   \item \vspace{-2 mm} multiplication of the weight matrix with the input vector $p = w x$
-   \item \vspace{-2 mm} addition of a bias term to the resulting scores $p = p + b$
-   \item \vspace{-2 mm} applying the activation function, most commonly softmax $ p_i = \exp(p_i) / \sum \exp(p_i) $
-   \item \vspace{-2 mm} a beam search for the best (or n-best) output classes $\argmax_i p_i$
+   \item \vspace{-3 mm} multiplication of the weight matrix with the input vector $p = w x$
+   \item \vspace{-3 mm} addition of a bias term to the resulting scores $p = p + b$
+   \item \vspace{-3 mm} applying the activation function, most commonly softmax $ p_i = \exp(p_i) / \sum \exp(p_i) $
+   \item \vspace{-3 mm} a beam search for the best (or n-best) output classes $\argmax_i p_i$
 \end{enumerate}
-
 In models with a small number of classes such as binary classification, the computational effort required is trivial and fast but this is not the case for the large number of classes found in typical NMT models.

### 2018-02-22 fast-beam/fast-beam-short.tex
fast

@@
 \end{tabular}
-\caption{Actual batch size during decoding (Europarl test set)}
+\caption{Actual batch size during decoding}
 \label{fig:batch-size}
@@
 \end{enumerate}
-In models with a small number of classes such as binary classification, the computational effort required is trivial and fast but this is not the case for the large number of classes found in typical NMT models.
 
@@
 
+As can be seen, the matrix p is iterated over five times - once to add the bias, three times to calculate the softmax, and once to search for the best classes.
+We propose fusing the three functions into one kernel, a popular optimization technique~\citep{Guevara2009EnablingTP}, making use of the following observations.
+
+Firstly, softmax and $\exp$ are monotonic functions, therefore, we can move the search for the best class from FIND-BEST to SOFTMAX, at the start of the kernel.
+
 \begin{algorithm}
 \begin{algorithmic}
+\small
 
@@
 
-As can be seen, the matrix p is iterated over five times - once to add the bias, three times to calculate the softmax, and once to search for the best classes.
-We propose fusing the three functions into one kernel, a popular optimization technique~\citep{Guevara2009EnablingTP}, making use of the following observations.
-
-Firstly, softmax and $\exp$ are monotonic functions, therefore, we can move the search for the best class from FIND-BEST to SOFTMAX, at the start of the kernel.
-
 Secondly, we are only interested in the probabilities of the best classes.
 Since they are now known at the start of the kernel, we compute softmax only for those classes.
-The outline of the our function is shown in Algorithm~\ref{algo:Fused Kernel}.
+The outline of the our kernel is shown in Algorithm~\ref{algo:Fused Kernel}.
 
@@
 \begin{algorithmic}
+\small
+
 \Procedure{Fused-Kernel}{vector $p$, bias vector $b$}
@@
 
-The standard mini-batching algorithm, Algorithm~\ref{algo:Mini-batching}, encode the sentences for a batch, followed by decoding the batch.
-Decoding stop only once all sentences in the batch are completed.
+The standard mini-batching algorithm encode the sentences for a batch, followed by decoding the batch. Decoding stop only once all sentences in the batch are completed.
 
@@
 \begin{algorithmic}
-\Procedure{Mini-Batching}{}
-\While{more input} 
-  \State Create batch $b$
-  \State Encode($b$)
-  \While{batch is not empty} 
-    \State Decode($b$)
-    \ForAll{sentence $s$ in $b$}
-      \If{trans($s$) is complete}
-        \State Remove $s$ from $b$
-      \EndIf
-    \EndFor
-  \EndWhile
-\EndWhile
-
-\EndProcedure
-
-\end{algorithmic}
-\caption{Mini-batching}
-\label{algo:Mini-batching}
-\end{algorithm}
-
-Our proposal focuses on decoding as this is the more compute-intensive task, encoding and decoding asynchronously.
-The encoding step, Algorithm~\ref{algo:Encoding for top-up batching}, is similar to the main loop of the standard algorithm but the results are added to a queue to be consumed by the decoding step.
-
-\begin{algorithm}
-\begin{algorithmic}
+\small
 
@@
 
-Rather than decoding a batch until completion, new sentences are added to the batch as old sentences completes, Algorithm~\ref{algo:Decoding for top-up batching}.
+Our proposal focuses on decoding as this is the more compute-intensive task, encoding and decoding asynchronously.
+The encoding step, Algorithm~\ref{algo:Encoding for top-up batching}, is similar to that in mini-batching except that the results are added to a queue.
+
+The decoding continuously process the same batch, adding new sentences as old sentences completes, Algorithm~\ref{algo:Decoding for top-up batching}.
 
@@
 \begin{algorithmic}
+\small
 
@@
 
-For inference, we used and extend Amun~\citep{junczys2016neural}, the fastest open-source inference engine we are aware of for the model used in this paper.
-We used a beam size of 5, mini-batch of 128 sentences and maxi-batch 1280, unless otherwise stated.
+For inference, we used and extend Amun~\citep{junczys2016neural}. We used a beam size of 5, mini-batch of 128 sentences and maxi-batch 1280, unless otherwise stated.
 
@@
 
-We trained a German-English with data from the Europarl corpus~\citep{Koehn:2005:MTS}, our test set consisted of a subset of the Open-Subtitles corpus, Table~\ref{tab:corpora}.
-
-\begin{table}
-\begin{center}
-\small
-\begin{tabular}{|l|r|r|} \hline
-		 & OpenSubtitles \\ \hline
-\# sentences  	 & 50,000 \\
-\# sub-words 	 & 467,654 \\ 
-Avg sub-words/sent & 9.4 \\
-Std dev subwords/sent & 6.1 \\ \hline
-\end{tabular}
-\end{center}
-\caption{Test sets}
-\label{tab:corpora}
-\end{table}
+The German-English version of the Europarl corpus~\citep{Koehn:2005:MTS} was used to train the model, our test set consisted of the first 50,000 sentences of the Open-Subtitles corpus~\citep{TIEDEMANN12.463}.
 
@@
 
-\section{Conclusion and Future Work}
-
-We have presented two methods for faster deep-learning inference, targeted at neural machine translation.
-
-The first method focused on output layer of the neural network which accounts for a large part of the running time of the NMT model.
-By fusing the output layer with the beam search, we are able to increase translation speed by up to 41\%.
-
-The second method replaces the mini-batching algorithm with one that avoids decoding with a small number of sentences, maximizing the parallel processing potential of GPUs.
-For certain scenarios, this increases translating  speed by up to 12\%.
-
-For future work, we would like to apply our optimization for other NLP and deep-learning tasks.
-We are also interested in further optimization of the output layer in NMT, specifically the matrix multiplication which still takes up a significant proportion of the translation time.
-
 %\section*{Acknowledgments}

### 2018-05-11 fast-beam/fast-beam.tex
use requireProb

@@
 
-In fact, we are usually only interested in the best class during inference, not the probability.
-Since we now compute the best class before the softmax, we can skip softmax altogether, Algorithm~\ref{algo:Argmax only}.
-This is only possible for beam size 1; the comparison between softmax probabilities is required for larger beam sizes as the denominators are different for different hypotheses.
+Thirdly, calculation of $max$ and $sum$ can be done in one loop by adjusting the sum whenever a higher $max$ is found, Equation~`.
+\begin{eqnarray*}
+sum & = & e^{a - max_b} + e^{b - max_b} \\
+    & = & e^{a - max_a + \Delta} + e^{b - max_b} \\
+    & = & e^{\Delta} e^{a - max_a} + e^{b - max_b}
+\end{eqnarray*}
+where $\Delta = max_b - max_a$ and $max_b > max_a$.
+
+In fact, a well known optimization is to skip the softmax operation altogether and calculate the argmax over the input vector, Algorithm~\ref{algo:Argmax only}.
+This is only possible for beam size 1 and when we are not interested in returning the softmax probabilities.
 

### 2018-05-11 fast-beam/fast-beam.tex
use requireProb

@@
 
-Secondly, we are only interested in the probabilities of the best classes.
-Since they are now known at the start of the kernel, we compute softmax only for those classes.
-The outline of the our function is shown in Algorithm~\ref{algo:Fused Kernel}.
+Secondly, we are only interested in the probabilities of the best classes. Since they are now known at the start of the kernel, we compute softmax only for those classes.
 
@@
 \State $max \gets - \infty$ 
+\State $sum \gets 0$
+
 \ForAll{$p_i$ in $p$}
-  \If{$p_i + b_i > max$}
-    \State $max \gets p_i + b_i$
+  \State $p_i' \gets p_i + b_i$  
+  \If{$p_i' > max$}
+    \State $\Delta \gets max - p_i'$
+    \State $sum \gets \Delta \times sum + 1 $
+    \State $max \gets p_i'$
     \State $best \gets i$
-  \EndIf
-\EndFor
-
-\Comment{calculate denominator}
-
-\State $sum \gets 0$ 
-\ForAll{$p_i$ in $p$}
-  \If{$p_i > max$}
-    \State $sum \gets sum + \exp(p_i - max)$
+  \Else
+    \State $sum \gets sum + \exp(p_i' - max)$
   \EndIf
@@
 
-Thirdly, calculation of $max$ and $sum$ can be done in one loop by adjusting the sum whenever a higher $max$ is found, Equation~`.
+Thirdly, the calculation of $max$ and $sum$ can be done in one loop by adjusting $sum$ whenever a higher $max$ is found, Equation~`.
 \begin{eqnarray*}
@@
 \end{eqnarray*}
-where $\Delta = max_b - max_a$ and $max_b > max_a$.
+where $\Delta = max_a - max_b$ and $max_b > max_a$. The outline of the our function is shown in Algorithm~\ref{algo:Fused Kernel}.
 
@@
 \end{algorithm}
+
+This changes does not reduce the asymtotic runtime of our algorithm which remains at $\mathcal{O}(|p|)$, it brings the number of iterations over $p$ down from 5 to 1.
+In practise, this has a significant affect on inference speed.
 

### 2018-05-11 fast-beam/fast-beam.tex
use requireProb

@@
 
-Thirdly, the calculation of $max$ and $sum$ can be done in one loop by adjusting $sum$ whenever a higher $max$ is found, Equation~`.
+Thirdly, the calculation of $max$ and $sum$ can be done in one loop by adjusting $sum$ whenever a higher $max$ is found:
 \begin{eqnarray*}
@@
 \end{eqnarray*}
-where $\Delta = max_a - max_b$ and $max_b > max_a$. The outline of the our function is shown in Algorithm~\ref{algo:Fused Kernel}.
+where $\Delta = max_a - max_b$ and $max_b > max_a$. The outline of our function is shown in Algorithm~\ref{algo:Fused Kernel}.
 
-In fact, a well known optimization is to skip the softmax operation altogether and calculate the argmax over the input vector, Algorithm~\ref{algo:Argmax only}.
+In fact, a well known optimization is to skip softmax altogether and calculate the argmax over the input vector, Algorithm~\ref{algo:Argmax only}.
 This is only possible for beam size 1 and when we are not interested in returning the softmax probabilities.
@@
 
-This changes does not reduce the asymtotic runtime of our algorithm which remains at $\mathcal{O}(|p|)$, it brings the number of iterations over $p$ down from 5 to 1.
+These changes do not reduce the asymtotic runtime of our algorithm, which remains at $\mathcal{O}(|p|)$, but it does bring the number of iterations over $p$ down from 5 to 1.
 In practise, this has a significant affect on inference speed.

### 2018-05-11 fast-beam/fast-beam.tex
use requireProb

@@
 
-These changes do not reduce the asymtotic runtime of our algorithm, which remains at $\mathcal{O}(|p|)$, but it does bring the number of iterations over $p$ down from 5 to 1.
+These changes do not reduce the asymptotic runtime of our algorithm, which remains at $\mathcal{O}(|p|)$, but it does bring the number of iterations over $p$ down from 5 to 1.
 In practise, this has a significant affect on inference speed.

### 2018-05-11 wnmt.2018/wnmt.tex
use requireProb

--- 2018-05-11 fast-beam/fast-beam.tex
@@
 % \mtsummitHeader{x}{x}{xxx-xxx}{2016}{45-character paper description goes here}{Author(s) initials and last name go here}
-\title{\bf Faster Neural Machine Translation Inference}
+\title{\bf Focusing on Fast Neural Network Inference}
 
@@
 
+This paper describe the submissions based on Amun, the fast inference engine for neural machine translation.
+Through efficient implementation, novel changes to the algorithms, our fastest submission is by far the fastest in the shared task.
+By using more complex models and with model distillation, our other submissions combine the inherent speed of our inference engine with better quality translation.
+
+%The team is composed of members from the University of Edinburgh, Adam Mickiewicz University, Microsoft, Tilde, and the Internation Maritime Organization.
+
+\end{abstract}
+
+\section{Introduction}
+\label{sec:Introduction}
+
 As neural machine translation (NMT) models have become the new state-of-the-art, the challenge is to make their deployment efficient and economical.
+This is the challenge that this shared task is shining a spotlight on.
+
+It is tempting to use a general deep-learning toolkit such as Tensorflow or Pytorch to train and do the the inference.
+Faster translation could be obtained by tuning parameters within the toolkit, and choosing fast models.
+We believe this is the approach most other submissions have taken.
+
+Our inference engine, Amun, is developed on the premise that fast neural network inference is an issue that deserves dedicated tools which are not hobbled by other objectives such as training or support for multiple models.
+A dedicated tool that can deliver on the useful goal of fast inference, but also serves as a test bed for research into neural network inference without having to be concerned with maintaining a more complete toolkit.
+Amun is also useful as a means to explore the upper limit of what is possible for a particular model and hardware.
+
+In summary, Amun is an inference-only engine that only supports a limited number of NMT models and put fast inference on modern GPU above all other priorities.
+The consequence of this focused approach can be seen in the result of the shared task.
+
+//////////////////////////////////////////////
+
 We have identified two areas where typical NMT models differ significantly from other deep-learning models that impact on their inference speed.
@@
 We provide solutions for each of these issues which together increase batched inference speed by up to 57\% on modern GPUs without affecting model quality.
-
-\end{abstract}
-
-\section{Introduction}
-\label{sec:Introduction}
 

### 2018-05-11 wnmt.2018/wnmt.tex
use requireProb

@@
 Our inference engine, Amun, is developed on the premise that fast neural network inference is an issue that deserves dedicated tools which are not hobbled by other objectives such as training or support for multiple models.
-A dedicated tool that can deliver on the useful goal of fast inference, but also serves as a test bed for research into neural network inference without having to be concerned with maintaining a more complete toolkit.
-Amun is also useful as a means to explore the upper limit of what is possible for a particular model and hardware.
+A tool that can deliver on the useful goal of fast inference, but also serves as a test bed for research into neural network inference without having to be concerned with maintaining a more complete toolkit.
+A tool that is also useful as a means to explore the upper limit of the possible speed for a particular model and hardware.
 
-In summary, Amun is an inference-only engine that only supports a limited number of NMT models and put fast inference on modern GPU above all other priorities.
+In summary, Amun is an inference-only engine that supports a limited number of NMT models and put fast inference on modern GPU above all other priorities.
 The consequence of this focused approach can be seen in the result of the shared task.
+
+We submitted three systems to this years shared task.
+Our first submission was tailored to be as fast as possible while being above the baseline BLEU score.
+Our second submission trade some of the speed of first submission to return good quality translation.
+Both of these systems were trained with standard training regimes.
+
+Our last submission uses model distillation to improve translation quality and increase speed.
 

### 2018-05-15 wnmt.2018/wnmt.tex
cast

--- 2018-05-11 fast-beam/fast-beam.tex
@@
 
+This paper describe the submissions by members of the University of Edinburgh, Adam Mickiewicz University, Tilde and University of Alicante.
+We submit two systems based on Amun, the fast inference engine for neural machine translation.
+We partipated
+
+Through efficient implementation and novel enhancements to the algorithms, our fastest submission is by far the fastest in the shared task.
+By using more complex models and with model distillation, our other submissions combine the inherent speed of our inference engine with better quality translation.
+
+%The team is composed of members from the University of Edinburgh, Adam Mickiewicz University, Microsoft, Tilde, and the Internation Maritime Organization.
+
+\end{abstract}
+
+\section{Introduction}
+\label{sec:Introduction}
+
 As neural machine translation (NMT) models have become the new state-of-the-art, the challenge is to make their deployment efficient and economical.
+This is the challenge that this shared task is shining a spotlight on.
+
+It is tempting to use a general purpose deep-learning toolkit such as Tensorflow or Pytorch to train and do the the inference where faster translation could be obtained by tuning parameters within the toolkit, and choosing fast models.
+We believe this is the approach most other submissions have taken.
+
+We take an opposing approach by using and enhancing a custom inference engine, Amun, which is developed on the premise that fast neural network inference is an issue that deserves dedicated tools which are not compromised by other objectives such as training or support for multiple models.
+As well as delivering on the useful goal of fast inference, it can serve as a test-bed for novel ideas on neural network inference, and is useful as a means to explore the upper limit of the possible speed for a particular model and hardware.
+
+That is, Amun is an inference-only engine that supports a limited number of NMT models that put fast inference on modern GPU above all other consideration.
+The consequence of this focused approach can be seen in the results of the shared task ???.
+
+We submitted two systems to this year's shared task for the efficient translation on GPU.
+Our first submission was tailored to be as fast as possible while being above the baseline BLEU score.
+Our second submission trade some of the speed of first submission to return good quality translation.
+
+We will describe below our experimental setup as well as the enhancements to Amun that we feel enabled it to achieve its outstanding speed.
+
+Exper
+
+//////////////////////////////////////////////
+
 We have identified two areas where typical NMT models differ significantly from other deep-learning models that impact on their inference speed.

### 2018-05-15 wnmt.2018/wnmt.tex
cast

@@
 
-This paper describe the submissions by members of the University of Edinburgh, Adam Mickiewicz University, Tilde and University of Alicante.
-We submit two systems based on Amun, the fast inference engine for neural machine translation.
-We partipated
-
-Through efficient implementation and novel enhancements to the algorithms, our fastest submission is by far the fastest in the shared task.
-By using more complex models and with model distillation, our other submissions combine the inherent speed of our inference engine with better quality translation.
-
-%The team is composed of members from the University of Edinburgh, Adam Mickiewicz University, Microsoft, Tilde, and the Internation Maritime Organization.
+This paper describe the submissions to the efficiency track for GPUs by members of the University of Edinburgh, Adam Mickiewicz University, Tilde and University of Alicante.
+We focus on efficient implementation of the recurrent deep-learning model as implemented in Amun, the fast inference engine for neural machine translation.
+We improve the performance with an efficient mini-batching and maxi-batching algorithm.
+Translation speed was also reduced by fusing the softmax operation with the beam seach algorithm.
 
@@
 
-Exper
+\section{Improvements}
 
-//////////////////////////////////////////////
+We desribe the main enhancements to Amun since the original public in 2016.
 
-We have identified two areas where typical NMT models differ significantly from other deep-learning models that impact on their inference speed.
-Firstly, NMT models usually contains a large number of output classes, corresponding to the output vocabulary.
-Secondly, rather than a single label, the output from machine translation models is a sequence of class labels that makes up the words or sub-words in the target sentence.
-Target sentence lengths are unpredictable, leading to inefficiencies during batch processing.
-We provide solutions for each of these issues which together increase batched inference speed by up to 57\% on modern GPUs without affecting model quality.
+\subsection{Batching}
 
-We examine two areas that are critical to fast NMT inference where the models differ significantly from other deep-learning models.
-We believe the optimizations of these models have been overlooked by the general deep-learning community.
-
-Firstly, the number of classes in many deep-learning applications is small.
-However, the number of classes in NMT models is typically in the tens or hundreds of thousands, corresponding to the vocabulary size of the output language.
-For example, ~\citet{sennrich-haddow-birch:2016:P16-12} experimented with target vocabulary sizes of 60,000 and 90,000 sub-word units.
-This makes the output layer of NMT models very computationally expensive.
-Figure~\ref{fig:pie-time-europarl} shows the breakdown of amount of time during translation our NMT system; nearly 70\% of the time is involved in the output layer.
-We will look at optimizations which explicitly target the output layer.
-\begin{figure}
-\centering
-\begin{tabular}{cc}
-{\includegraphics[scale=0.3]{pie-time-europarl.png}} 
-\end{tabular}
-\caption{Proportion of time spent during translation (Europarl test set)}
-\label{fig:pie-time-europarl}
-\end{figure}
-
-Secondly, the use of mini-batching is critical for fast model inference.
-However, mini-batching does not take into account variable sentence lengths which decrease the actual number of input or output sentences that are processed in parallel, negating the benefit of mini-batching.
-This can be partially reduced in the encoding with \emph{maxi-batching}, i.e.
-pre-sorting sentences by length before creating mini-batches with similar length source sentences.
-However, maxi-batching can introduce unacceptable delay in response for online applications as the input are not processed in the order they came.
-It is also not appropriate for applications where only the output lengths varies, for example, image captioning.
-
-Even with sequence-to-sequence NMT models, target sentence lengths will still differ even for similar length inputs, compromising decoding performance.
-Figure~\ref{fig:batch-size} shows the actual batch size during decoding with a maximum batch size of 128; the batch was full in only 42\% of decoding iterations.
-
-\begin{figure}
-\centering
-\begin{tabular}{cc}
-{\includegraphics[scale=0.3]{batch-size.png}} 
-\end{tabular}
-\caption{Actual batch size during decoding (Europarl test set)}
-\label{fig:batch-size}
-\end{figure}
-
-We will propose an alternative batching algorithm to increase the actual batch size during decoding without the problems associated with mini-batching and maxi-batching.
-
-\section{Prior Work}
-
-Deep-learning models has been successfully used on many machine learning task in recent years in areas as diverse as computer vision and natural language processing.
-This success have been followed by the rush to put these model into production.
-The computational resources required in order to run these models has been challenging but, fortunately, there has been a lot work to meet this challenge.
-
-Hardware accelerators such as GPUs are very popular but other specialized hardware such as custom processors~\citep{DBLP:journals/corr/JouppiYPPABBBBB17} and FPGA~\citep{DBLP:journals/corr/LaceyTA16} have been used.
-Hardware-supported reduced precision is also an ideal way to speed up model inference~\citep{DBLP:journals/corr/abs-1710-03740}.
-
-Simpler, faster models have been created that are as good, or almost as good, as more slower, more complex models~\citep{DBLP:journals/corr/BahdanauCB14}.
-Research have also gone into smaller, faster models that can approximate slower, bigger models~\citep{DBLP:conf/emnlp/KimR16}.
-Some models have been justified as more suited for the parallel architecture of GPUs~\cite{DBLP:journals/corr/VaswaniSPUJGKP17}~\cite{gehring2017convs2s}.
-
-The speed of the softmax layer when used with large vocabularies have been looked at in~\citet{DBLP:journals/corr/GraveJCGJ16} but there have been many other attempts at faster softmax, for example~\citet{DBLP:journals/corr/abs-1301-3781}, ~\citet{Zoph-2016}.
-
-There has been surprisingly little work on batching algorithms, considering its critical importance for efficient deep-learning training and inference.
-~\citet{Neubig-autobatching} describe an novel batching algorithm, however, its aim is to alleviate the burden on developers of batching models rather than faster batching.
-
-Our paper follows most closely on from ~\citet{DBLP:conf/emnlp/Devlin17} which achieved faster NMT inference mainly by novel implementation of an existing model.
-However, we differ by focusing on GPU implementation, specifically the output layer, and the batching algorithm which was not touched on by the previous work.
-Our model is based on that of ~\citet{D14-1179} and ~\citet{sennrich-haddow-birch:2016:P16-12} but our work is applicable to other models and applications.
-% rather than CPUs as pursued by ~\cite{DBLP:conf/emnlp/Devlin17}.
-
-\section{Proposal}
-\label{sec:Proposal}
-
-\subsection{Softmax and Beam Search Fusion}
+\subsection{Softmax and Beam Search Kernel Fusion}
 
@@
 In practise, this has a significant affect on inference speed.
+
+\subsection{Half-Precision}
+
+/////////////////////////////////////////////////////////////////////////////////////
+
+We have identified two areas where typical NMT models differ significantly from other deep-learning models that impact on their inference speed.
+Firstly, NMT models usually contains a large number of output classes, corresponding to the output vocabulary.
+Secondly, rather than a single label, the output from machine translation models is a sequence of class labels that makes up the words or sub-words in the target sentence.
+Target sentence lengths are unpredictable, leading to inefficiencies during batch processing.
+We provide solutions for each of these issues which together increase batched inference speed by up to 57\% on modern GPUs without affecting model quality.
+
+We examine two areas that are critical to fast NMT inference where the models differ significantly from other deep-learning models.
+We believe the optimizations of these models have been overlooked by the general deep-learning community.
+
+Firstly, the number of classes in many deep-learning applications is small.
+However, the number of classes in NMT models is typically in the tens or hundreds of thousands, corresponding to the vocabulary size of the output language.
+For example, ~\citet{sennrich-haddow-birch:2016:P16-12} experimented with target vocabulary sizes of 60,000 and 90,000 sub-word units.
+This makes the output layer of NMT models very computationally expensive.
+Figure~\ref{fig:pie-time-europarl} shows the breakdown of amount of time during translation our NMT system; nearly 70\% of the time is involved in the output layer.
+We will look at optimizations which explicitly target the output layer.
+\begin{figure}
+\centering
+\begin{tabular}{cc}
+{\includegraphics[scale=0.3]{pie-time-europarl.png}} 
+\end{tabular}
+\caption{Proportion of time spent during translation (Europarl test set)}
+\label{fig:pie-time-europarl}
+\end{figure}
+
+Secondly, the use of mini-batching is critical for fast model inference.
+However, mini-batching does not take into account variable sentence lengths which decrease the actual number of input or output sentences that are processed in parallel, negating the benefit of mini-batching.
+This can be partially reduced in the encoding with \emph{maxi-batching}, i.e.
+pre-sorting sentences by length before creating mini-batches with similar length source sentences.
+However, maxi-batching can introduce unacceptable delay in response for online applications as the input are not processed in the order they came.
+It is also not appropriate for applications where only the output lengths varies, for example, image captioning.
+
+Even with sequence-to-sequence NMT models, target sentence lengths will still differ even for similar length inputs, compromising decoding performance.
+Figure~\ref{fig:batch-size} shows the actual batch size during decoding with a maximum batch size of 128; the batch was full in only 42\% of decoding iterations.
+
+\begin{figure}
+\centering
+\begin{tabular}{cc}
+{\includegraphics[scale=0.3]{batch-size.png}} 
+\end{tabular}
+\caption{Actual batch size during decoding (Europarl test set)}
+\label{fig:batch-size}
+\end{figure}
+
+We will propose an alternative batching algorithm to increase the actual batch size during decoding without the problems associated with mini-batching and maxi-batching.
+
+\section{Prior Work}
+
+Deep-learning models has been successfully used on many machine learning task in recent years in areas as diverse as computer vision and natural language processing.
+This success have been followed by the rush to put these model into production.
+The computational resources required in order to run these models has been challenging but, fortunately, there has been a lot work to meet this challenge.
+
+Hardware accelerators such as GPUs are very popular but other specialized hardware such as custom processors~\citep{DBLP:journals/corr/JouppiYPPABBBBB17} and FPGA~\citep{DBLP:journals/corr/LaceyTA16} have been used.
+Hardware-supported reduced precision is also an ideal way to speed up model inference~\citep{DBLP:journals/corr/abs-1710-03740}.
+
+Simpler, faster models have been created that are as good, or almost as good, as more slower, more complex models~\citep{DBLP:journals/corr/BahdanauCB14}.
+Research have also gone into smaller, faster models that can approximate slower, bigger models~\citep{DBLP:conf/emnlp/KimR16}.
+Some models have been justified as more suited for the parallel architecture of GPUs~\cite{DBLP:journals/corr/VaswaniSPUJGKP17}~\cite{gehring2017convs2s}.
+
+The speed of the softmax layer when used with large vocabularies have been looked at in~\citet{DBLP:journals/corr/GraveJCGJ16} but there have been many other attempts at faster softmax, for example~\citet{DBLP:journals/corr/abs-1301-3781}, ~\citet{Zoph-2016}.
+
+There has been surprisingly little work on batching algorithms, considering its critical importance for efficient deep-learning training and inference.
+~\citet{Neubig-autobatching} describe an novel batching algorithm, however, its aim is to alleviate the burden on developers of batching models rather than faster batching.
+
+Our paper follows most closely on from ~\citet{DBLP:conf/emnlp/Devlin17} which achieved faster NMT inference mainly by novel implementation of an existing model.
+However, we differ by focusing on GPU implementation, specifically the output layer, and the batching algorithm which was not touched on by the previous work.
+Our model is based on that of ~\citet{D14-1179} and ~\citet{sennrich-haddow-birch:2016:P16-12} but our work is applicable to other models and applications.
+% rather than CPUs as pursued by ~\cite{DBLP:conf/emnlp/Devlin17}.
+
+\section{Proposal}
+\label{sec:Proposal}
 

### 2018-05-15 wnmt.2018/wnmt.tex
cast

@@
 
-We desribe the main enhancements to Amun since the original public in 2016.
+We desribe the main enhancements to Amun since the original 2016 publication.
 
@@
 
-\subsection{Softmax and Beam Search Kernel Fusion}
+The use of mini-batching is critical for fast model inference. However, a percularity
+
+However, mini-batching often does not take into account variable sentence lengths which decrease the actual number of input or output sentences that are processed in parallel, negating the benefit of mini-batching.
+This can be partially reduced in the encoding with \emph{maxi-batching}, i.e.
+pre-sorting sentences by length before creating mini-batches with similar length source sentences.
+However, maxi-batching can introduce unacceptable delay in response for online applications as the input are not processed in the order they came.
+It is also not appropriate for applications where only the output lengths varies, for example, image captioning.
+
+\subsection{Softmax and Beam Search Fusion}
+
+The number of classes in many deep-learning applications is small but this is often not the case with NMT models.
+For example, ~\citet{sennrich-haddow-birch:2016:P16-12} experimented with target vocabulary sizes of 60,000 and 90,000 sub-word units.
+This makes the output layer of NMT models very computationally expensive.
+Figure~\ref{fig:pie-time-europarl} shows the breakdown of amount of time during translation our NMT system; nearly 70\% of the time is involved in the output layer.
+We describe the steps to fuse the operations to improve efficiency.
+\begin{figure}
+\centering
+\begin{tabular}{cc}
+{\includegraphics[scale=0.3]{pie-time-europarl.png}} 
+\end{tabular}
+\caption{Proportion of time spent during translation (Europarl test set)}
+\label{fig:pie-time-europarl}
+\end{figure}
 
@@
 
-/////////////////////////////////////////////////////////////////////////////////////
+////////////////////////////////////////////////////////////////
 
@@
 We believe the optimizations of these models have been overlooked by the general deep-learning community.
-
-Firstly, the number of classes in many deep-learning applications is small.
-However, the number of classes in NMT models is typically in the tens or hundreds of thousands, corresponding to the vocabulary size of the output language.
-For example, ~\citet{sennrich-haddow-birch:2016:P16-12} experimented with target vocabulary sizes of 60,000 and 90,000 sub-word units.
-This makes the output layer of NMT models very computationally expensive.
-Figure~\ref{fig:pie-time-europarl} shows the breakdown of amount of time during translation our NMT system; nearly 70\% of the time is involved in the output layer.
-We will look at optimizations which explicitly target the output layer.
-\begin{figure}
-\centering
-\begin{tabular}{cc}
-{\includegraphics[scale=0.3]{pie-time-europarl.png}} 
-\end{tabular}
-\caption{Proportion of time spent during translation (Europarl test set)}
-\label{fig:pie-time-europarl}
-\end{figure}
-
-Secondly, the use of mini-batching is critical for fast model inference.
-However, mini-batching does not take into account variable sentence lengths which decrease the actual number of input or output sentences that are processed in parallel, negating the benefit of mini-batching.
-This can be partially reduced in the encoding with \emph{maxi-batching}, i.e.
-pre-sorting sentences by length before creating mini-batches with similar length source sentences.
-However, maxi-batching can introduce unacceptable delay in response for online applications as the input are not processed in the order they came.
-It is also not appropriate for applications where only the output lengths varies, for example, image captioning.
 

### 2018-05-16 wnmt.2018/wnmt.tex
small numerical differences

@@
 As well as delivering on the useful goal of fast inference, it can serve as a test-bed for novel ideas on neural network inference, and is useful as a means to explore the upper limit of the possible speed for a particular model and hardware.
-
-That is, Amun is an inference-only engine that supports a limited number of NMT models that put fast inference on modern GPU above all other consideration.
-The consequence of this focused approach can be seen in the results of the shared task ???.
+That is, Amun is an inference-only engine that supports a limited number of NMT models that put fast inference on modern GPU above all other considerations.
+%The consequence of this focused approach can be seen in the results of the shared task ???.
 
@@
 
-The use of mini-batching is critical for fast model inference. However, a percularity
+The use of mini-batching is critical for fast model inference.
+The size of the batch is determined by the number of inputs sentences to the encoder in an encoder-decoder model.
+However, the number of batches during decoding can vary as some sentences have completed translating or the beam search add more hypotheses to the batch.
 
-However, mini-batching often does not take into account variable sentence lengths which decrease the actual number of input or output sentences that are processed in parallel, negating the benefit of mini-batching.
-This can be partially reduced in the encoding with \emph{maxi-batching}, i.e.
-pre-sorting sentences by length before creating mini-batches with similar length source sentences.
-However, maxi-batching can introduce unacceptable delay in response for online applications as the input are not processed in the order they came.
-It is also not appropriate for applications where only the output lengths varies, for example, image captioning.
+It is tempting to ignore these consideration, for example, by always decoding with a constant batch and beam size and ignoring hypotheses which are not needed.
+This is the strategy taken by other inference engines such as Marian, at the cost of lower translation speed due to wasteful processing.
+
+The Amun engine implements an efficient batching algorithm that takes into account the actual number of hypotheses that needs to be decoded at each decoding step, Figure~\ref{algo:Mini-batching}.
+
+\begin{algorithm}
+\begin{algorithmic}
+\Procedure{Batching}{encoded records $i$}
+
+\State Create batch $b$ from $i$
+
+\While{$b$ is not empty} 
+  \State $b' \gets$ DecodeAndBeamSearch($b$)
+  \State $b \gets \emptyset$
+
+  \ForAll{hypo $h$ in $b'$}
+    \If{$h \neq \textless/s \textgreater$  }
+      \State Add $h$ to $b$
+    \EndIf
+  \EndFor
+\EndWhile
+
+\EndProcedure
+
+\end{algorithmic}
+\caption{Mini-batching}
+\label{algo:Mini-batching}
+\end{algorithm}
 
@@
 The standard mini-batching algorithm is outlined in Algorithm~\ref{algo:Mini-batching}.
-
-\begin{algorithm}
-\begin{algorithmic}
-\Procedure{Mini-Batching}{}
-\While{more input} 
-  \State Create batch $b$
-  \State Encode($b$)
-  \While{batch is not empty} 
-    \State Decode($b$)
-    \ForAll{sentence $s$ in $b$}
-      \If{trans($s$) is complete}
-        \State Remove $s$ from $b$
-      \EndIf
-    \EndFor
-  \EndWhile
-\EndWhile
-
-\EndProcedure
-
-\end{algorithmic}
-\caption{Mini-batching}
-\label{algo:Mini-batching}
-\end{algorithm}
 
@@
 \grid
+\grid

### 2018-05-16 wnmt.2018/wnmt.tex
small numerical differences

@@
 
-It is tempting to ignore these consideration, for example, by always decoding with a constant batch and beam size and ignoring hypotheses which are not needed.
-This is the strategy taken by other inference engines such as Marian, at the cost of lower translation speed due to wasteful processing.
+It is tempting to ignore these considerations, for example, by always decoding with a constant batch and beam size and ignoring hypotheses which are not needed but this comes at a cost of lower translation speed due to wasteful processing.
 
@@
 
-The number of classes in many deep-learning applications is small but this is often not the case with NMT models.
+Most NMT more predict a large number of classes in their output layer, corresponding to the number of words or subword units in their target language.
 For example, ~\citet{sennrich-haddow-birch:2016:P16-12} experimented with target vocabulary sizes of 60,000 and 90,000 sub-word units.
@@
 
-Thirdly, the calculation of $max$ and $sum$ can be done in one loop by adjusting $sum$ whenever a higher $max$ is found:
+Thirdly, the calculation of $max$ and $sum$ can be accomplished in one loop by adjusting $sum$ whenever a higher $max$ is found:
 \begin{eqnarray*}
@@
 \end{eqnarray*}
-where $\Delta = max_a - max_b$ and $max_b > max_a$. The outline of our function is shown in Algorithm~\ref{algo:Fused Kernel}.
+where $max_a$ is the original maximum value, $max_b$ is the new, higher maximum value, ie.
+$max_b > max_a$, and $\Delta = max_a - max_b$.
+The outline of our function is shown in Algorithm~\ref{algo:Fused Kernel}.
 

### 2018-05-16 wnmt.2018/wnmt.tex
small numerical differences

@@
 \subsection{Half-Precision}
+
+Reducing the number of bits needed to store floating point values from 32-bits to 16-bits promises to increase translation speed through faster calculations and reduced bandwidth usage.
+16-bit floating point operations are supported by the GPU hardware and software available in the shared task.
+
+In practise, however, efficiently using half-precision value requires a comprehensive redvelopment of the GPU code.
+We therefore make do with using the GPU's Tensor Core fast matrix multiplication routines which transparently converts 32-bit float point input matrices to 16-bit values.
 

### 2018-05-16 wnmt.2018/wnmt.tex
cast

@@
 It is tempting to use a general purpose deep-learning toolkit such as Tensorflow or Pytorch to train and do the the inference where faster translation could be obtained by tuning parameters within the toolkit, and choosing fast models.
-We believe this is the approach most other submissions have taken.
+%We believe this is the approach most other submissions have taken.
 
-We take an opposing approach by using and enhancing a custom inference engine, Amun, which is developed on the premise that fast neural network inference is an issue that deserves dedicated tools which are not compromised by other objectives such as training or support for multiple models.
+We take an opposing approach by using and enhancing a custom inference engine, Amun~\citep{junczys2016neural}, which we developed on the premise that fast deep-learning inference is an issue that deserves dedicated tools which are not compromised by other objectives such as training or support for multiple models.
 As well as delivering on the useful goal of fast inference, it can serve as a test-bed for novel ideas on neural network inference, and is useful as a means to explore the upper limit of the possible speed for a particular model and hardware.
@@
 In practise, however, efficiently using half-precision value requires a comprehensive redvelopment of the GPU code.
-We therefore make do with using the GPU's Tensor Core fast matrix multiplication routines which transparently converts 32-bit float point input matrices to 16-bit values.
+We therefore make do with using the GPU's Tensor Core fast matrix multiplication routines which transparently converts 32-bit float point input matrices to 16-bit values and output a 32-bit float point product of the inputs.
+
+\section{Experimental Setup}
+\label{sec:Experimental Setup}
+
+Both of our submitted systems uses the sequence-to-sequence model similar to that described in ~\citet{sennrich-haddow-birch:2016:P16-12}, containing a bidirectional RNN in the encoder and a two-layer RNN in the decoder.
+We use BPE to adjust the vocabulary size.
+
+\subsection{GRU-based system}
+
+Our first system uses gated recurrent units (GRU) throughout, trained with Marian, but submitted both systems using the Amun inference engine.
+
+We experimented with varying the vocabulary size and the RNN state size before settling for a vocabulary size of 30,000 (for both source and target language) and 256 for the state size, Figure~\ref{tab:BLEU for newstest2014}.
+
+\begin{table}
+\begin{center}
+\small
+\begin{tabular}{|l|r|r|r|} \hline
+		& \multicolumn{3}{|c|}{\emph{State dim}}	\\ \hline	
+Vocab size	& 256	& 512	& 1024 \\ \hline
+1000 		& 10.9 	& 	& 11.8 \\ 
+5000		& 15.9	&  	& 16.4 \\ 
+10000		& 17.2	& 	& 17.9 \\
+20000		& -	&	& 18.4 \\ 
+30000		& 17.8	& 18.5	& 18.7 \\ \hline
+\end{tabular}
+\end{center}
+\caption{Test set BLEU (newstest2014) for hyper-parameters}
+\label{tab:BLEU for newstest2014}
+\end{table}
+
+After further experimentation, we decided to use sentence length normalization and NVidia's Tensor Core matrix multiplication which increased translation quality as well as translation speed.
+The beam was kept at 1 throughout for the fastet possible inference.
+
+\subsection{mLSTM-based system}
+
+Our second system uses multiplicative-LSTM in the encoder and the first layer of a decder, and a GRU in the second layer.
+We trained 2 systems with differing vocabulary sizes using an extension of the Nematus toolkit which supports such model.
 
@@
 \end{algorithm}
-
-\section{Experimental Setup}
-\label{sec:Experimental Setup}
-
-We trained a sequence-to-sequence, encoder-decoder NMT system similar to that described in ~\citet{sennrich-haddow-birch:2016:P16-12}.
-This uses recurrent neural networks with gated recurrent units.
-The input and output vocabulary size were both set to 85,000 sub-words using byte-pair encoding (BPE) to adjust the vocabulary to the desired size.
-The hidden layer dimensions was set to 512.
-%We used the Marian toolkit to train our models.
-
-For inference, we used and extend Amun~\citep{junczys2016neural}, the fastest open-source inference engine we are aware of for the model used in this paper.
-We used a beam size of 5, mini-batch of 128 sentences and maxi-batch 1280, unless otherwise stated.
-
-The hardware used in all experiments was an Nvidia GTX 1060 GPU on a host containing 8 Intel hypercores running at 2.8Ghz, 16GB RAM and SSD hard drive.
-
-Our training data consisted of the German-English parallel sentences from the Europarl corpus~\citep{Koehn:2005:MTS}. To test inference speed, we used two test sets with differing characteristics:
-\begin{enumerate}
-   \item \vspace{-2 mm} a subset of the Europarl training data, which contains mostly long sentences, and is, of course, in the same domain as the training data
-   \item \vspace{-2 mm} a subset of the German-English data from the Open-Subtitles corpus, consisting of mostly short, out-of-domain sentences.
-\end{enumerate}
-Table~\ref{tab:corpora} gives further details of the test sets.
-
-\begin{table}
-\begin{center}
-\small
-\begin{tabular}{|l|r|r|} \hline
-		  & Europarl		& OpenSubtitles \\ \hline
-\# sentences  	  & 30,000 		& 50,000 \\
-\# sub-words 	  & 787,908 		& 467,654 \\ 
-Avg sub-words/sent & 26.3		& 9.4 \\
-Std dev subwords/sent & 14.9		& 6.1 \\ \hline
-\end{tabular}
-\end{center}
-\caption{Test sets}
-\label{tab:corpora}
-\end{table}
 

### 2018-05-16 wnmt.2018/wnmt.tex
small numerical differences

@@
 
-We experimented with varying the vocabulary size and the RNN state size before settling for a vocabulary size of 30,000 (for both source and target language) and 256 for the state size, Figure~\ref{tab:BLEU for newstest2014}.
+We experimented with varying the vocabulary size and the RNN state size before settling for a vocabulary size of 30,000 (for both source and target language) and 256 for the state size, Table~\ref{tab:BLEU for newstest2014}.
 
@@
 \begin{center}
-\small
 \begin{tabular}{|l|r|r|r|} \hline
@@
 Vocab size	& 256	& 512	& 1024 \\ \hline
-1000 		& 10.9 	& 	& 11.8 \\ 
-5000		& 15.9	&  	& 16.4 \\ 
-10000		& 17.2	& 	& 17.9 \\
-20000		& -	&	& 18.4 \\ 
-30000		& 17.8	& 18.5	& 18.7 \\ \hline
+1000 		& 12.23 &	& 12.77 \\ 
+5000		& 16.79	&  	& 17.16 \\ 
+10,000		& 18.00	& 	& 18.19 \\
+20,000		& -	&	& 19.52 \\ 
+30,000		& 18.51	& 19.17	& 19.64 \\ \hline
 \end{tabular}
 \end{center}
-\caption{Test set BLEU (newstest2014) for hyper-parameters}
+\caption{Validation set BLEU (newstest2014) for GRU-based model}
 \label{tab:BLEU for newstest2014}
@@
 
-Our second system uses multiplicative-LSTM in the encoder and the first layer of a decder, and a GRU in the second layer.
-We trained 2 systems with differing vocabulary sizes using an extension of the Nematus toolkit which supports such model.
+Our second system uses multiplicative-LSTM in the encoder and the first layer of a decder, and a GRU in the second layer, using an extension of the Nematus toolkit which supports such model.
+We trained 2 systems with differing vocabulary sizes and varied the beam sizes before chose the the settings that produces good translation quality on the validation set, Table~\ref{tab:mLSTM BLEU - vocab sizes}.
+
+\begin{table}
+\small
+\begin{center}
+\begin{tabular}{|l|r|r|r|r|r|} \hline
+		& \multicolumn{5}{|l|}{\emph{Beam size}}	\\ \hline	
+Vocab size	& 1	& 2	& 2	& 4	& 5 \\ \hline
+40,000 		& 23.45	& 24.15	& 24.48	& 24.42	& 24.48 \\ 
+50,000		& 23.32	& 24.04	& 	&	& \\ \hline
+\end{tabular}
+\end{center}
+\caption{Validation set BLEU for mLSTM-based model}
+\label{tab:mLSTM BLEU - vocab sizes}
+\end{table}
+
+\section{Result}
+\label{sec:Result}
 

### 2018-05-16 wnmt.2018/wnmt.tex
small numerical differences

@@
 
+We used a variety of GPUs to train the models but all testing was done on an Nvidia V100.
+Translation quality was measured using BLEU, specifically multi-bleu as found in the Moses toolkit.
+The validation and test sets provided by the shared task organisers were used to measure translation quality, but a 50,000 sentence subset of the training data was used to measure translation speed to obtain longer, more accurate measuremenrs.
+
 \subsection{GRU-based system}
@@
 After further experimentation, we decided to use sentence length normalization and NVidia's Tensor Core matrix multiplication which increased translation quality as well as translation speed.
-The beam was kept at 1 throughout for the fastet possible inference.
+The beam was kept at 1 throughout for the fastest possible inference.
 
@@
 Our second system uses multiplicative-LSTM in the encoder and the first layer of a decder, and a GRU in the second layer, using an extension of the Nematus toolkit which supports such model.
-We trained 2 systems with differing vocabulary sizes and varied the beam sizes before chose the the settings that produces good translation quality on the validation set, Table~\ref{tab:mLSTM BLEU - vocab sizes}.
+We trained 2 systems with differing vocabulary sizes and varied the beam sizes before choosing the the settings that produces good translation quality on the validation set, Table~\ref{tab:mLSTM BLEU - vocab sizes}.
 
 \begin{table}
-\small
 \begin{center}
-\begin{tabular}{|l|r|r|r|r|r|} \hline
-		& \multicolumn{5}{|l|}{\emph{Beam size}}	\\ \hline	
-Vocab size	& 1	& 2	& 2	& 4	& 5 \\ \hline
-40,000 		& 23.45	& 24.15	& 24.48	& 24.42	& 24.48 \\ 
-50,000		& 23.32	& 24.04	& 	&	& \\ \hline
+\begin{tabular}{|l|r|r|} \hline
+		& \multicolumn{2}{|l|}{\emph{Vocab size}}	\\ \hline	
+Beam size	& 40,000	& 50,000 \\ \hline
+1 		& 23.45		& 23.32	\\ 
+2		& 24.15		& 24.04	\\
+3		& 24.48		&  	\\
+4		& 24.42		& 	\\
+5		& 24.48		& 	\\ \hline
 \end{tabular}

### 2018-05-16 wnmt.2018/wnmt.tex
small numerical differences

@@
 Translation quality was measured using BLEU, specifically multi-bleu as found in the Moses toolkit.
-The validation and test sets provided by the shared task organisers were used to measure translation quality, but a 50,000 sentence subset of the training data was used to measure translation speed to obtain longer, more accurate measuremenrs.
+The validation and test sets provided by the shared task organisers were used to measure translation quality, but a 50,000 sentence subset of the training data was used to measure translation speed to obtain longer, more accurate measurements.
 

### 2018-05-16 wnmt.2018/wnmt.tex
small numerical differences

@@
 
-Thirdly, the calculation of $max$ and $sum$ can be accomplished in one loop by adjusting $sum$ whenever a higher $max$ is found:
+Thirdly, the calculation of $max$ and $sum$ can be accomplished in one loop by adjusting $sum$ whenever a higher $max$ is found during the looping:
 \begin{eqnarray*}
-sum & = & e^{a - max_b} + e^{b - max_b} \\
-    & = & e^{a - max_a + \Delta} + e^{b - max_b} \\
-    & = & e^{\Delta} e^{a - max_a} + e^{b - max_b}
+sum & = & e^{x_t - max_b} + \sum_{i=0...t-1}{e^{x_i - max_b}} \\
+    & = & e^{x_t - max_b} + \sum_{i=0...t-1}{e^{x_i - max_a + \Delta}}  \\
+    & = & e^{x_t - max_b} + e^{\Delta} \times \sum_{i=0...t-1}{e^{x_i - max_a}} 
 \end{eqnarray*}
@@
 
-These changes do not reduce the asymptotic runtime of our algorithm, which remains at $\mathcal{O}(|p|)$, but it does bring the number of iterations over $p$ down from 5 to 1.
+%These changes do not reduce the asymptotic runtime of our algorithm, which remains at $\mathcal{O}(|p|)$, but it does bring the number of iterations over $p$ down from 5 to 1.
 In practise, this has a significant affect on inference speed.
@@
 \label{sec:Result}
+
+\subsection{Batching}
+
+Using batching can increase the translation speed by over 20 times in Amun and shows little degradation with larger batch sizes, Figure~\ref{fig:batch-size}.
+This contrast with using the Marian, which uses a simpler batching algorithm, where the speedup is approximately 16 times and inference speed slows down when the batch size is above 1000.
+
+\begin{figure}
+\centering
+\begin{tabular}{cc}
+{\includegraphics[scale=0.3]{batch-size.png}} 
+\end{tabular}
+\caption{Actual batch size during decoding}
+\label{fig:batch-size}
+\end{figure}
+
+\subsection{Softmax and Beam Search Fusion}
 

### 2018-05-16 wnmt.2018/wnmt.tex
cast

@@
 
+The efficiency of Amun's batching algorithm can be seen by observing the time taken for each decoding step in a batch of sentence, Figure~\ref{fig:decode-step}.
+Amun's decoding becomes faster as sentences finish translating.
+By contrast, Marian's decoding time stays relatively constant.
+
+\begin{figure}
+\centering
+\begin{tabular}{cc}
+{\includegraphics[scale=0.3]{decode-step.png}} 
+\end{tabular}
+\caption{Time taken for decoding each for a batch of 1280 sentences}
+\label{fig:decode-step}
+\end{figure}
+
 \subsection{Softmax and Beam Search Fusion}

### 2018-05-17 wnmt.2018/wnmt.tex
cast

@@
 \begin{algorithmic}
-\Procedure{Batching}{encoded records $i$}
+\Procedure{Batching}{encoded sentences $i$}
 
@@
 
-\While{$b$ is not empty} 
+\While{$b \neq \emptyset$} 
   \State $b' \gets$ DecodeAndBeamSearch($b$)
@@
 Vocab size	& 256	& 512	& 1024 \\ \hline
-1000 		& 12.23 &	& 12.77 \\ 
-5000		& 16.79	&  	& 17.16 \\ 
+1.000 		& 12.23 &	& 12.77 \\ 
+5,000		& 16.79	&  	& 17.16 \\ 
 10,000		& 18.00	& 	& 18.19 \\
@@
 
+The efficiency of Amun's batching algorithm can be seen by observing the time taken for each decoding step in a batch of sentence, Figure~\ref{fig:decode-step}.
+Amun's decoding becomes faster as sentences finish translating.
+This contrast with the Marian inference engine, which uses a simpler batching algorithm, where the speed stays relatively constant throughout the decoding of the batch.
+
+\begin{figure}
+\centering
+\begin{tabular}{cc}
+{\includegraphics[scale=0.3]{decode-step.png}} 
+\end{tabular}
+\caption{Time taken for each decoding step for a batch of 1280 sentences}
+\label{fig:decode-step}
+\end{figure}
+
 Using batching can increase the translation speed by over 20 times in Amun and shows little degradation with larger batch sizes, Figure~\ref{fig:batch-size}.
-This contrast with using the Marian, which uses a simpler batching algorithm, where the speedup is approximately 16 times and inference speed slows down when the batch size is above 1000.
+Just as importantly, however, it doesn't suffer degradation with large batch sizes unlike the simpler algorithm which slows down when the batch size is above 1000.
+This scalability issue is likely to increase with ever increasing core counts.
 
@@
 \end{tabular}
-\caption{Actual batch size during decoding}
+\caption{Speed v. batch size}
 \label{fig:batch-size}
-\end{figure}
-
-The efficiency of Amun's batching algorithm can be seen by observing the time taken for each decoding step in a batch of sentence, Figure~\ref{fig:decode-step}.
-Amun's decoding becomes faster as sentences finish translating.
-By contrast, Marian's decoding time stays relatively constant.
-
-\begin{figure}
-\centering
-\begin{tabular}{cc}
-{\includegraphics[scale=0.3]{decode-step.png}} 
-\end{tabular}
-\caption{Time taken for decoding each for a batch of 1280 sentences}
-\label{fig:decode-step}
 \end{figure}
@@
 Figure~\ref{fig:batch-size} shows the actual batch size during decoding with a maximum batch size of 128; the batch was full in only 42\% of decoding iterations.
-
-\begin{figure}
-\centering
-\begin{tabular}{cc}
-{\includegraphics[scale=0.3]{batch-size.png}} 
-\end{tabular}
-\caption{Actual batch size during decoding (Europarl test set)}
-\label{fig:batch-size}
-\end{figure}
 

### 2018-05-17 wnmt.2018/wnmt.tex
cast

@@
 \begin{tabular}{cc}
-{\includegraphics[scale=0.3]{pie-time-europarl.png}} 
+{\includegraphics[scale=0.3]{pie-time.png}} 
 \end{tabular}
-\caption{Proportion of time spent during translation (Europarl test set)}
-\label{fig:pie-time-europarl}
+\caption{Proportion of time spent during translation}
+\label{fig:pie-time.png}
 \end{figure}
@@
 
-\begin{figure}
-\centering
-\begin{tabular}{cc}
-{\includegraphics[scale=0.3]{pie-time-opensubtitles.png}} 
-\end{tabular}
-\caption{Proportion of time spent during translation (OpenSubtitles)}
-\label{fig:pie-time-opensubtitles}
-\end{figure}
-
 \subsection{Top-up Batching}

### 2018-05-17 wnmt.2018/wnmt.tex
cast

@@
 
-////////////////////////////////////////////////////////////////
-
-We have identified two areas where typical NMT models differ significantly from other deep-learning models that impact on their inference speed.
-Firstly, NMT models usually contains a large number of output classes, corresponding to the output vocabulary.
-Secondly, rather than a single label, the output from machine translation models is a sequence of class labels that makes up the words or sub-words in the target sentence.
-Target sentence lengths are unpredictable, leading to inefficiencies during batch processing.
-We provide solutions for each of these issues which together increase batched inference speed by up to 57\% on modern GPUs without affecting model quality.
-
-We examine two areas that are critical to fast NMT inference where the models differ significantly from other deep-learning models.
-We believe the optimizations of these models have been overlooked by the general deep-learning community.
-
-Even with sequence-to-sequence NMT models, target sentence lengths will still differ even for similar length inputs, compromising decoding performance.
-Figure~\ref{fig:batch-size} shows the actual batch size during decoding with a maximum batch size of 128; the batch was full in only 42\% of decoding iterations.
-
-We will propose an alternative batching algorithm to increase the actual batch size during decoding without the problems associated with mini-batching and maxi-batching.
-
-\section{Prior Work}
-
-Deep-learning models has been successfully used on many machine learning task in recent years in areas as diverse as computer vision and natural language processing.
-This success have been followed by the rush to put these model into production.
-The computational resources required in order to run these models has been challenging but, fortunately, there has been a lot work to meet this challenge.
-
-Hardware accelerators such as GPUs are very popular but other specialized hardware such as custom processors~\citep{DBLP:journals/corr/JouppiYPPABBBBB17} and FPGA~\citep{DBLP:journals/corr/LaceyTA16} have been used.
-Hardware-supported reduced precision is also an ideal way to speed up model inference~\citep{DBLP:journals/corr/abs-1710-03740}.
-
-Simpler, faster models have been created that are as good, or almost as good, as more slower, more complex models~\citep{DBLP:journals/corr/BahdanauCB14}.
-Research have also gone into smaller, faster models that can approximate slower, bigger models~\citep{DBLP:conf/emnlp/KimR16}.
-Some models have been justified as more suited for the parallel architecture of GPUs~\cite{DBLP:journals/corr/VaswaniSPUJGKP17}~\cite{gehring2017convs2s}.
-
-The speed of the softmax layer when used with large vocabularies have been looked at in~\citet{DBLP:journals/corr/GraveJCGJ16} but there have been many other attempts at faster softmax, for example~\citet{DBLP:journals/corr/abs-1301-3781}, ~\citet{Zoph-2016}.
-
-There has been surprisingly little work on batching algorithms, considering its critical importance for efficient deep-learning training and inference.
-~\citet{Neubig-autobatching} describe an novel batching algorithm, however, its aim is to alleviate the burden on developers of batching models rather than faster batching.
-
-Our paper follows most closely on from ~\citet{DBLP:conf/emnlp/Devlin17} which achieved faster NMT inference mainly by novel implementation of an existing model.
-However, we differ by focusing on GPU implementation, specifically the output layer, and the batching algorithm which was not touched on by the previous work.
-Our model is based on that of ~\citet{D14-1179} and ~\citet{sennrich-haddow-birch:2016:P16-12} but our work is applicable to other models and applications.
-% rather than CPUs as pursued by ~\cite{DBLP:conf/emnlp/Devlin17}.
-
-\section{Proposal}
-\label{sec:Proposal}
-
-\subsection{Top-up Batching}
-
-The standard mini-batching algorithm is outlined in Algorithm~\ref{algo:Mini-batching}.
-
-This algorithm encode the sentences for a batch, followed by decoding the batch.
-The decoding stop once all sentences in the batch are completed.
-This is a potential inefficiency as the number of remaining sentences may not be optimal.
-
-We will focus on decoding as this is the more compute-intensive step, and issues with differing sentence sizes in encoding can partly be ameliorated by maxi-batching.
-
-Our proposed top-up batching algorithm encode and decode asynchronously.
-The encoding step, Algorithm~\ref{algo:Encoding for top-up batching}, is similar to the main loop of the standard algorithm but the results are added to a queue to be consumed by the decoding step.
-
-\begin{algorithm}
-\begin{algorithmic}
-
-\Procedure{Encode}{}
-\While{more input} 
-  \State Create encoding batch $b$
-  \State Encode($b$)
-  \State Add $b$ to queue $q$
-\EndWhile
-
-\EndProcedure
-
-\end{algorithmic}
-\caption{Encoding for top-up batching}
-\label{algo:Encoding for top-up batching}
-\end{algorithm}
-
-Rather than decoding the same batch until all sentences in the batch are completed, the decoding step processing the same batch continuously.
-New sentences are added to the batch as old sentences completes, Algorithm~\ref{algo:Decoding for top-up batching}.
-
-\begin{algorithm}
-\begin{algorithmic}
-
-\Procedure{Decode}{}
-
-\State create batch $b$ from queue $q$
-\While{$b$ is not empty}
-  \State Decode($b$)
-  \ForAll{sentence $s$ in $b$}
-    \If{trans($s$) is complete}
-      \State Replace $s$ with $s'$ from $q$
-    \EndIf
-  \EndFor
-\EndWhile
-
-\EndProcedure
-
-\end{algorithmic}
-\caption{Decoding for top-up batching}
-\label{algo:Decoding for top-up batching}
-\end{algorithm}
-
-\section{Results}
-\label{sec:Results}
-
-\subsection{Softmax and Beam Search Fusion}
-
-Fusing the last 3 steps led to a substantial reduction in the time taken, especially for beam size of 1 where the softmax probability does not actually have to calculated, Table~\ref{tab:fused-breakdown-europarl} and ~\ref{tab:fused-breakdown-opensubtitles}.
-This led to an overall increase in translation speed of up to 23\% for the Europarl test set, Figure~\ref{fig:beam-europarl}, and up to 41\% for the OpenSubtitles test set, Figure~\ref{fig:beam-opensubtitles}.
-
-The amount of time taken by the output layer dominates translation time for the OpenSubtitles test set, Figure~\ref{fig:pie-time-opensubtitles}, explaining the bigger overall translation speed with the fused kernel.
-
-\begin{table}
-\begin{center}
-\small
-\begin{tabular}{|l|r|r|} \hline
-		& Baseline	& Fused \\ \hline
-\multicolumn{3}{|c|}{\emph{Beam size 1}}	\\ \hline	
-Multiplication 	& 25.62 	& 26.47 (+3.3\%) \\ \hline
-Add bias 	& 4.98		&  \\ 
-Softmax 	& 12.81		& 7.94 (-76.9\%)\\
-Beam search	& 16.64		&  \\ \hline
-\multicolumn{3}{|c|}{\emph{Beam size 5}}	\\ \hline	
-Multiplication 	& 112.9 	& 115.04 (+1.9\%) \\ \hline
-Add bias 	& 23.66		&  \\ 
-Softmax 	& 56.61		& 67.58 (-56.5\%)\\
-Beam search	& 75.12		&  \\ \hline
-\end{tabular}
-\end{center}
-\caption{Time taken in sec (Europarl)}
-\label{tab:fused-breakdown-europarl}
-\end{table}
-
-\begin{table}
-\begin{center}
-\small
-\begin{tabular}{|l|r|r|} \hline
-		& Baseline	& Fused \\ \hline
-\multicolumn{3}{|c|}{\emph{Beam size 1}}	\\ \hline	
-Multiplication 	& 21.47 	& 21.94 (+2.2\%) \\ \hline
-Add bias 	& 3.29		&  \\ 
-Softmax 	& 8.83		& 5.70 (-78.1\%)\\
-Beam search	& 13.91		&  \\ \hline
-\multicolumn{3}{|c|}{\emph{Beam size 5}}	\\ \hline	
-Multiplication 	& 79.66 	& 80.01 (+4.4\%) \\ \hline
-Add bias 	& 14.95		&  \\ 
-Softmax 	& 36.86		& 42.91 (-64.4\%)\\
-Beam search	& 68.80		&  \\ \hline
-\end{tabular}
-\end{center}
-\caption{Time taken in sec (OpenSubtitles)}
-\label{tab:fused-breakdown-opensubtitles}
-\end{table}
+Fusing the bias and softmax operations in the output layer with the beam search results in a speed improvement by 25\%, Figure~\ref{fig:fused}.
+Its relative improvement decrease marginally with increasing beam size.
 
@@
 \begin{tabular}{cc}
-{\includegraphics[scale=0.5]{beam-europarl.png}} 
+{\includegraphics[scale=0.3]{fused.png}} 
 \end{tabular}
-\caption{Speed using the fused kernel (Europarl)}
-\label{fig:beam-europarl}
+\caption{Using fused operation}
+\label{fig:fused}
 \end{figure}
 
-\begin{figure}
-\centering
-\begin{tabular}{cc}
-{\includegraphics[scale=0.5]{beam-opensubtitles.png}} 
+Further insight can be gained by examining the time taken for each step in the output layer and beam search, Table~\ref{tab:fused-breakdown}.
+The fused operation only has to loop through the large cost matrix once, therefore, for low beam sizes its is comparable in speed to the simple kernel to add the bias.
+For higher beam sizes, the cost of maintaining the n-best list is felt.
+
+\begin{table}
+\begin{center}
+\begin{tabular}{|l|r|r|} \hline
+		& Baseline	& Fused \\ \hline
+\multicolumn{3}{|c|}{\emph{Beam size 1}}	\\ \hline	
+Multiplication 	& 5.39 		& 5.38 (+0\%) \\ \hline
+Add bias 	& 1.26		&  \\ 
+Softmax 	& 1.69		& 2.07 (-86.6\%)\\
+Beam search	& 12.53		&  \\ \hline
+\multicolumn{3}{|c|}{\emph{Beam size 3}}	\\ \hline	
+Multiplication 	& 14.18 	& 14.16 (+0\%) \\ \hline
+Add bias 	& 3.76		&  \\ 
+Softmax 	& 4.75		& 3.43 (-87.1\%)\\
+Beam search	& 18.23		&  \\ \hline
+\multicolumn{3}{|c|}{\emph{Beam size 9}}	\\ \hline	
+Multiplication 	& 38.35 	& 38.42 (+0\%) \\ \hline
+Add bias 	& 11.64		&  \\ 
+Softmax 	& 14.4		& 17.5 (-72.1\%)\\
+Beam search	& 36.7		&  \\ \hline
 \end{tabular}
-\caption{Speed using the fused kernel (OpenSubtitles)}
-\label{fig:beam-opensubtitles}
-\end{figure}
+\end{center}
+\caption{Time taken (sec) breakdown}
+\label{tab:fused-breakdown}
+\end{table}
 
-\subsection{Top-up Batching}
+\subsection{Tensor Cores}
 
-After some experimentation, we decided to top-up the decoding batch only when it is at least half empty, rather than whenever a sentence has completed.
+By taking advantage of the GPU's hardware accelerated matrix multiplication, we can gain up to 20\% in speed, Table~\ref{tab:tensor-cores}.
 
-The top-up batching and maxi-batching have similar goals of maximizing efficiency when translating batches of different lengths.
-Therefore, using both methods together gives limited gains, in fact, using top-up batching with maxi-batch slows of between 2\% to 18\% when translating the Europarl test set due to the overhead of using the algorithm, Figure~\ref{fig:topup-europarl}.
-When maxi-batching is inappropriate, using top-up batching alone matches the performance of maxi-batching, even being slightly faster when a small beam is used.
-
-\begin{figure}
-\centering
-\begin{tabular}{cc}
-{\includegraphics[scale=0.3]{topup-europarl.png}} 
+\begin{table}[ht]
+\begin{center}
+\begin{tabular}{|l|r|r|} \hline
+Beam size	& Baseline	& Tensor Cores \\ \hline
+1	 	& 39.97 	& 34.54 (-13.6\%) \\ 
+9 		& 145.8		& 116.8 (-20.0\%) \\ \hline
 \end{tabular}
-\caption{Speed using top-up batching (Europarl)}
-\label{fig:topup-europarl}
-\end{figure}
-
-The results are better when translating the OpenSubtitles test set, Figure~\ref{fig:topup-opensubtitles}.
-The top-up batching does not harm performance when used with maxi-batching, even helping a little for small beams.
-However, top-up batching increases translation speed by up to 12\% when used alone.
-
-\begin{figure}
-\centering
-\begin{tabular}{cc}
-{\includegraphics[scale=0.3]{topup-opensubtitles.png}} 
-\end{tabular}
-\caption{Speed using top-up batching (OpenSubtitles)}
-\label{fig:topup-opensubtitles}
-\end{figure}
-
-The gain from top-up batching is partially dependent on how much time the standard mini-batching algorithm spend decoding with a very small number of sentences as this is does not fully utilize the GPU cores.
-From Figure~\ref{fig:batch-size-small}, 20\% of the decoding iterations have less than 8 sentences remaining in the batch when translating the Europarl test test with maxi-batching.
-This figure is 50\% for the OpenSubtitles test set with no maxi-batching.
-The OpenSubtitles test set has a higher variance of sentence lengths relative to its average sentence length, which forces the mini-batch algorithm to continue decoding with a small number of sentences while the other sentences have already completed.
-
-\begin{figure}
-\centering
-\begin{tabular}{cc}
-{\includegraphics[scale=0.3]{batch-size-small.png}} 
-\end{tabular}
-\caption{Actual batch size during decoding}
-\label{fig:batch-size-small}
-\end{figure}
-
-\subsection{Cummulative Results}
-
-Using both the fused kernel and top-up batching to translate led to a cummalative speed improvement of up to 57\% and 34\% for the OpenSubtitles and Europarl test set, respectively, when no maxi-batching is used, Figure~\ref{fig:cummulative-europarl} and Figure~\ref{fig:cummulative-opensubtitles}.
-With maxi-batching, the speed was up to 41\% and 21\%.
-
-\begin{figure}
-\centering
-\begin{tabular}{cc}
-{\includegraphics[scale=0.3]{cummulative-europarl.png}} 
-\end{tabular}
-\caption{Cummulative results (Europarl}
-\label{fig:cummulative-europarl}
-\end{figure}
-
-\begin{figure}
-\centering
-\begin{tabular}{cc}
-{\includegraphics[scale=0.3]{cummulative-opensubtitles.png}} 
-\end{tabular}
-\caption{Cummulative results (OpenSubtitles}
-\label{fig:cummulative-opensubtitles}
-\end{figure}
+\end{center}
+\caption{Time taken (sec) using Tensor Cores}
+\label{tab:tensor-cores}
+\end{table}
 
@@
 
-We have presented two methods for faster deep-learning inference, targeted at neural machine translation.
+We have presented some of the improvement to Amun which are focused to improvement NMT inference.
 
-The first method focused on output layer of the neural network which accounts for a large part of the running time of the NMT model.
-By fusing the output layer with the beam search, we are able to increase translation speed by up to 41\%.
-
-The second method replaces the mini-batching algorithm with one that avoids decoding with a small number of sentences, maximizing the parallel processing potential of GPUs.
-For certain scenarios, this increases translating  speed by up to 12\%.
-
-For future work, we would like to apply our optimization for other NLP and deep-learning tasks.
-We are also interested in further optimization of the output layer in NMT, specifically the matrix multiplication which still takes up a significant proportion of the translation time.
+We are also working to make deep-learning faster using more specialised hardware such as FPGAs.
+It would be interesting as future work to bring our focused approach to fast deep-learning inference to a more general training toolkit.
 

### 2018-05-17 wnmt.2018/wnmt.tex
cast

@@
 
-Our first system uses gated recurrent units (GRU) throughout, trained with Marian, but submitted both systems using the Amun inference engine.
+Our first system uses gated recurrent units (GRU) throughout, trained with Marian~\citep{junczys2018marian}, but submitted both systems using the Amun inference engine.
 

### 2018-05-17 wnmt.2018/wnmt.tex
cast

@@
 In practise, however, efficiently using half-precision value requires a comprehensive redvelopment of the GPU code.
-We therefore make do with using the GPU's Tensor Core fast matrix multiplication routines which transparently converts 32-bit float point input matrices to 16-bit values and output a 32-bit float point product of the inputs.
+We therefore make do with using the GPU's Tensor Core\footnote{https://devblogs.nvidia.com/programming-tensor-cores-cuda-9/} fast matrix multiplication routines which transparently converts 32-bit float point input matrices to 16-bit values and output a 32-bit float point product of the inputs.
 
@@
 We used a variety of GPUs to train the models but all testing was done on an Nvidia V100.
-Translation quality was measured using BLEU, specifically multi-bleu as found in the Moses toolkit.
+Translation quality was measured using BLEU, specifically multi-bleu as found in the Moses toolkit\footnote{https://github.com/moses-smt/mosesdecoder}.
 The validation and test sets provided by the shared task organisers were used to measure translation quality, but a 50,000 sentence subset of the training data was used to measure translation speed to obtain longer, more accurate measurements.

### 2018-05-17 wnmt.2018/wnmt.tex
cast

@@
 
-Our second system uses multiplicative-LSTM in the encoder and the first layer of a decder, and a GRU in the second layer, using an extension of the Nematus toolkit which supports such model.
+Our second system uses multiplicative-LSTM in the encoder and the first layer of a decder, and a GRU in the second layer, using an extension of the Nematus~\citep{sennrich-EtAl:2017:EACLDemo} toolkit which supports such models.
 We trained 2 systems with differing vocabulary sizes and varied the beam sizes before choosing the the settings that produces good translation quality on the validation set, Table~\ref{tab:mLSTM BLEU - vocab sizes}.
@@
 We are also working to make deep-learning faster using more specialised hardware such as FPGAs.
-It would be interesting as future work to bring our focused approach to fast deep-learning inference to a more general training toolkit.
+It would be interesting as future work to bring our focused approach to fast deep-learning inference to a more general toolkit.
 

### 2018-05-17 wnmt.2018/wnmt.tex
turn on timings

@@
 
-It is tempting to use a general purpose deep-learning toolkit such as Tensorflow or Pytorch to train and do the the inference where faster translation could be obtained by tuning parameters within the toolkit, and choosing fast models.
-%We believe this is the approach most other submissions have taken.
+One approach is to use an off-the-shelf deep-learning toolkit such as Tensorflow\footnote{https://www.tensorflow.org/} or Pytorch\footnote{https://pytorch.org/} to train and do the inference where the novelty comes from selecting the appropriate models and tuning parameters within the toolkit for optimal performance.
 
-We take an opposing approach by using and enhancing a custom inference engine, Amun~\citep{junczys2016neural}, which we developed on the premise that fast deep-learning inference is an issue that deserves dedicated tools which are not compromised by other objectives such as training or support for multiple models.
+We take an opposing approach eschewing model selection in favour efficient implementation.
+We use a custom inference engine, Amun~\citep{junczys2016neural}, which we developed on the premise that fast deep-learning inference is an issue that deserves dedicated tools that are not compromised by competing objectives such as training or support for multiple models.
 As well as delivering on the useful goal of fast inference, it can serve as a test-bed for novel ideas on neural network inference, and is useful as a means to explore the upper limit of the possible speed for a particular model and hardware.
@@
 Our first submission was tailored to be as fast as possible while being above the baseline BLEU score.
-Our second submission trade some of the speed of first submission to return good quality translation.
+Our second submission trade some of the speed of first submission to return better quality translations.
 
-We will describe below our experimental setup as well as the enhancements to Amun that we feel enabled it to achieve its outstanding speed.
+%We will describe below our experimental setup as well as the enhancements to Amun that we feel enabled it to achieve its speed.
 
@@
 
-We desribe the main enhancements to Amun since the original 2016 publication.
+We describe the main enhancements to Amun since the original 2016 publication.
 

### 2018-05-17 wnmt.2018/wnmt.tex
turn on timings

@@
 
-One approach is to use an off-the-shelf deep-learning toolkit such as Tensorflow\footnote{https://www.tensorflow.org/} or Pytorch\footnote{https://pytorch.org/} to train and do the inference where the novelty comes from selecting the appropriate models and tuning parameters within the toolkit for optimal performance.
+One approach is to use an off-the-shelf deep-learning toolkit to train and do the inference where the novelty comes from selecting the appropriate models and tuning parameters within the toolkit for optimal performance.
+% such as Tensorflow\footnote{https://www.tensorflow.org/} or Pytorch\footnote{https://pytorch.org/}
 

### 2018-05-17 wnmt.2018/wnmt.tex
turn on timings

@@
 We take an opposing approach eschewing model selection in favour efficient implementation.
-We use a custom inference engine, Amun~\citep{junczys2016neural}, which we developed on the premise that fast deep-learning inference is an issue that deserves dedicated tools that are not compromised by competing objectives such as training or support for multiple models.
-As well as delivering on the useful goal of fast inference, it can serve as a test-bed for novel ideas on neural network inference, and is useful as a means to explore the upper limit of the possible speed for a particular model and hardware.
+We use and enhanced a custom inference engine, Amun~\citep{junczys2016neural}, which we developed on the premise that fast deep-learning inference is an issue that deserves dedicated tools that are not compromised by competing objectives such as training or support for multiple models.
+As well as delivering on the practical goal of fast inference, it can serve as a test-bed for novel ideas on neural network inference, and is useful as a means to explore the upper limit of the possible speed for a particular model and hardware.
 That is, Amun is an inference-only engine that supports a limited number of NMT models that put fast inference on modern GPU above all other considerations.
@@
 
-\Comment{add bias, calculate $max$ \& $argmax$}
+%\Comment{add bias, calculate $max$ \& $argmax$}
 

### 2018-05-17 wnmt.2018/wnmt.tex
turn on timings

@@
 For example, ~\citet{sennrich-haddow-birch:2016:P16-12} experimented with target vocabulary sizes of 60,000 and 90,000 sub-word units.
-This makes the output layer of NMT models very computationally expensive.
-Figure~\ref{fig:pie-time-europarl} shows the breakdown of amount of time during translation our NMT system; nearly 70\% of the time is involved in the output layer.
-We describe the steps to fuse the operations to improve efficiency.
-\begin{figure}
-\centering
-\begin{tabular}{cc}
-{\includegraphics[scale=0.3]{pie-time.png}} 
-\end{tabular}
-\caption{Proportion of time spent during translation}
-\label{fig:pie-time.png}
-\end{figure}
 
@@
 
-In models with a small number of classes such as binary classification, the computational effort required is trivial and fast but this is not the case for the large number of classes found in typical NMT models.
+Figure~\ref{fig:pie-time} shows the amount of time spent in each step during translation.
+Clearly, the output layer of NMT models are very computationally expensive, accounting for over 60\% of the translation time.
 
-We leave step 1 for future work and focus on the last three steps, the outline for which are shown in Algorithm~\ref{algo:original-softmax-beamsearch}.
+\begin{figure}
+\centering
+\begin{tabular}{cc}
+{\includegraphics[scale=0.3]{pie-time.png}} 
+\end{tabular}
+\caption{Proportion of time spent during translation}
+\label{fig:pie-time}
+\end{figure}
+
+We focus on the last three steps, the outline for which are shown in Algorithm~\ref{algo:original-softmax-beamsearch}.
 For brevity, we show the algorithm for 1-best, a beam search of the n-bests is a simple extension of this.

### 2018-05-17 wnmt.2018/wnmt.tex
turn on timings

@@
 
-\begin{algorithm}
+\begin{algorithm}[ht]
 \begin{algorithmic}
@@
 
-Both of our submitted systems uses the sequence-to-sequence model similar to that described in ~\citet{sennrich-haddow-birch:2016:P16-12}, containing a bidirectional RNN in the encoder and a two-layer RNN in the decoder.
-We use BPE to adjust the vocabulary size.
+Both of our submitted systems uses the sequence-to-sequence model similar to that described in ~\citet{DBLP:journals/corr/BahdanauCB14}, containing a bidirectional RNN in the encoder and a two-layer RNN in the decoder.
+We use byte pair encoding~\citep{sennrich-haddow-birch:2016:P16-12} to adjust the vocabulary size.
 
@@
 
-Our second system uses multiplicative-LSTM in the encoder and the first layer of a decder, and a GRU in the second layer, using an extension of the Nematus~\citep{sennrich-EtAl:2017:EACLDemo} toolkit which supports such models.
+Our second system uses multiplicative-LSTM~\citep{krause2017} in the encoder and the first layer of a decder, and a GRU in the second layer, using an extension of the Nematus~\citep{sennrich-EtAl:2017:EACLDemo} toolkit which supports such models.
 We trained 2 systems with differing vocabulary sizes and varied the beam sizes before choosing the the settings that produces good translation quality on the validation set, Table~\ref{tab:mLSTM BLEU - vocab sizes}.
@@
 
-\begin{figure}
+\begin{figure}[ht]
 \centering
@@
 
-\begin{table}
+\begin{table}[ht]
 \begin{center}

### 2018-05-17 wnmt.2018/wnmt.tex
turn on timings

@@
 % \mtsummitHeader{x}{x}{xxx-xxx}{2016}{45-character paper description goes here}{Author(s) initials and last name go here}
-\title{\bf Focusing on Fast Neural Network Inference}
+\title{\bf Fast Neural Machine Translation Implementation}
 
@@
 \AND
-       \name{\bf Tomasz Dwojak} \hfill \addr{???}\\
+       \name{\bf Tomasz Dwojak} \hfill \addr{t.dwojak@amu.edu.pl}\\
         \addr{Adam Mickiewicz University}
 \AND
-       \name{\bf Kenneth Heafield} \hfill \addr{???}\\
+       \name{\bf Rihards Krislauks} \hfill \addr{rihards.krislauks@tilde.lv}\\
+        \addr{Tilde}
+\AND
+       \name{\bf Daniel Torregrosa} \hfill \addr{dtorregrosa@dlsi.ua.es}\\
+        \addr{Universitat d’Alacant}
+\AND
+       \name{\bf Kenneth Heafield} \hfill \addr{kheafiel@inf.ed.ac.uk}\\
        %\name{\bf Kenneth Heafield} \hfill \addr{kheafiel@inf.ed.ac.uk}\\

### 2018-05-17 wnmt.2018/wnmt.tex
turn on timings

@@
 
-We take an opposing approach eschewing model selection in favour efficient implementation.
+We take an opposing approach by eschewing model selection in favour of efficient implementation.
 We use and enhanced a custom inference engine, Amun~\citep{junczys2016neural}, which we developed on the premise that fast deep-learning inference is an issue that deserves dedicated tools that are not compromised by competing objectives such as training or support for multiple models.
@@
 
-Our second system uses multiplicative-LSTM~\citep{krause2017} in the encoder and the first layer of a decder, and a GRU in the second layer, using an extension of the Nematus~\citep{sennrich-EtAl:2017:EACLDemo} toolkit which supports such models.
+Our second system uses multiplicative-LSTM~\citep{krause2017} in the encoder and the first layer of a decder, and a GRU in the second layer, trained with an extension of the Nematus~\citep{sennrich-EtAl:2017:EACLDemo} toolkit which supports such models.
 We trained 2 systems with differing vocabulary sizes and varied the beam sizes before choosing the the settings that produces good translation quality on the validation set, Table~\ref{tab:mLSTM BLEU - vocab sizes}.

### 2018-05-17 wnmt.2018/wnmt.tex
turn on timings

@@
 
-Using batching can increase the translation speed by over 20 times in Amun and shows little degradation with larger batch sizes, Figure~\ref{fig:batch-size}.
-Just as importantly, however, it doesn't suffer degradation with large batch sizes unlike the simpler algorithm which slows down when the batch size is above 1000.
-This scalability issue is likely to increase with ever increasing core counts.
+Using batching can increase the translation speed by over 20 times in Amun, Figure~\ref{fig:batch-size}.
+Just as important, it doesn't suffer degradation with large batch sizes unlike the simpler algorithm which slows down when batch size over 1000 is used.
+This scalability issue is likely to rise in importance with new GPUs with ever increasing core counts.
 

### 2018-05-17 wnmt.2018/wnmt.tex
turn on timings

@@
+%
+% File acl2018.tex
+%
+%% Based on the style files for ACL-2017, with some changes, which were, in turn,
+%% Based on the style files for ACL-2015, with some improvements
+%%  taken from the NAACL-2016 style
+%% Based on the style files for ACL-2014, which were, in turn,
+%% based on ACL-2013, ACL-2012, ACL-2011, ACL-2010, ACL-IJCNLP-2009,
+%% EACL-2009, IJCNLP-2008...
+%% Based on the style files for EACL 2006 by 
+%%e.agirre@ehu.es or Sergi.Balari@uab.es
+%% and that of ACL 08 by Joakim Nivre and Noah Smith
+
 \documentclass[11pt,a4paper]{article}
@@
 
-\usepackage[letterpaper]{geometry}
-%\usepackage{amta2016}
 \usepackage{url}
-\usepackage{natbib}
-\usepackage{layout}
 
@@
 
-\newcommand{\confname}{AMTA 2016}
-\newcommand{\website}{\protect\url{http://www.amtaweb.org/}}
-\newcommand{\contactname}{research track co-chair Lane Schwartz}
-\newcommand{\contactemail}{lanes@illinois.edu} 
-\newcommand{\conffilename}{amta2016}
-\newcommand{\downloadsite}{\protect\url{http://www.amtaweb.org/}}
-\newcommand{\paperlength}{$12$ (twelve)}
-\newcommand{\shortpaperlength}{$6$ (six)}
+\aclfinalcopy % Uncomment this line for the final submission
+\def\aclpaperid{***} %  Enter the acl Paper ID here
 
-%% do not add any other page- or text-size instruction here
+%\setlength\titlebox{5cm}
+% You can expand the titlebox if you need extra space
+% to show all the authors. Please do not make the titlebox
+% smaller than 5cm (the original size); we will check this
+% in the camera-ready version and ask you to change it back.
 
-\parskip=0.00in
+\newcommand\BibTeX{B{\sc ib}\TeX}
+
+% \mtsummitHeader{x}{x}{xxx-xxx}{2016}{45-character paper description goes here}{Author(s) initials and last name go here}
+\title{Fast Neural Machine Translation Implementation}
+
+\author{Hieu Hoang \\
+  University of Edinburgh, Scotland \\
+  {\tt hieu@hoang.co.uk} \\
+  \And
+  Tomasz Dwojak \\
+  Adam Mickiewicz University \\
+  {\tt t.dwojak@amu.edu.pl} \\
+  \And
+  Rihards Krislauks \\
+  Tilde \\
+  {\tt rihards.krislauks@tilde.lv} \\
+  \And
+  Daniel Torregrosa \\
+  Universitat d’Alacant \\
+  {\tt dtorregrosa@dlsi.ua.es} \\
+  \And
+  Kenneth Heafield \\
+  University of Edinburgh, Scotland \\
+  {\tt kheafiel@inf.ed.ac.uk} \\
+  }
+
+\date{}
 
 \begin{document}
-
-% \mtsummitHeader{x}{x}{xxx-xxx}{2016}{45-character paper description goes here}{Author(s) initials and last name go here}
-\title{\bf Fast Neural Machine Translation Implementation}
-
-\author{\name{\bf Hieu Hoang} \hfill  \addr{hieu@hoang.co.uk}\\ 
-        \addr{}
-\AND
-       \name{\bf Tomasz Dwojak} \hfill \addr{t.dwojak@amu.edu.pl}\\
-        \addr{Adam Mickiewicz University}
-\AND
-       \name{\bf Rihards Krislauks} \hfill \addr{rihards.krislauks@tilde.lv}\\
-        \addr{Tilde}
-\AND
-       \name{\bf Daniel Torregrosa} \hfill \addr{dtorregrosa@dlsi.ua.es}\\
-        \addr{Universitat d’Alacant}
-\AND
-       \name{\bf Kenneth Heafield} \hfill \addr{kheafiel@inf.ed.ac.uk}\\
-       %\name{\bf Kenneth Heafield} \hfill \addr{kheafiel@inf.ed.ac.uk}\\
-        \addr{University of Edinburgh, Scotland}
-}
-
 \maketitle
-\pagestyle{empty}
-
 \begin{abstract}
@@
 \end{document}
-\grid
-\grid

### 2018-05-17 wnmt.2018/wnmt.tex
turn on timings

@@
 We focus on efficient implementation of the recurrent deep-learning model as implemented in Amun, the fast inference engine for neural machine translation.
-We improve the performance with an efficient mini-batching and maxi-batching algorithm.
-Translation speed was also reduced by fusing the softmax operation with the beam seach algorithm.
+We improve the performance with an efficient mini-batching algorithm and by fusing the softmax operation with the beam seach algorithm.
 
@@
 
-\begin{algorithm}[ht]
+\begin{algorithm}
 \begin{algorithmic}
@@
 
-\begin{figure}[ht]
+\begin{figure}
 \centering
@@
 
-\begin{table}[ht]
+\begin{table}
 \begin{center}

### 2018-05-17 wnmt.2018/wnmt.tex
turn on timings

@@
 
-One approach is to use an off-the-shelf deep-learning toolkit to train and do the inference where the novelty comes from selecting the appropriate models and tuning parameters within the toolkit for optimal performance.
+One approach is to use an off-the-shelf deep-learning toolkit to complete the shared task where the novelty comes from selecting the appropriate models and tuning parameters within the toolkit for optimal performance.
 % such as Tensorflow\footnote{https://www.tensorflow.org/} or Pytorch\footnote{https://pytorch.org/}

### 2018-05-17 wnmt.2018/wnmt.tex
turn on timings

@@
 
-\author{Hieu Hoang \\
-  University of Edinburgh, Scotland \\
-  {\tt hieu@hoang.co.uk} \\
-  \And
-  Tomasz Dwojak \\
-  Adam Mickiewicz University \\
-  {\tt t.dwojak@amu.edu.pl} \\
-  \And
-  Rihards Krislauks \\
-  Tilde \\
-  {\tt rihards.krislauks@tilde.lv} \\
-  \And
-  Daniel Torregrosa \\
-  Universitat d’Alacant \\
-  {\tt dtorregrosa@dlsi.ua.es} \\
-  \And
-  Kenneth Heafield \\
-  University of Edinburgh, Scotland \\
-  {\tt kheafiel@inf.ed.ac.uk} \\
-  }
+  \author{Hieu Hoang$^{\dagger}$ \, Tomasz Dwojak$^{*}$ \, Rihards Krislauks$^{\ddagger}$ \\ {\bf Daniel Torregrosa$^{\mathparagraph}$ \, Kenneth Heafield$^{\dagger}$ } \\[2mm]
+  $^{\dagger}$University of Edinburgh \, $^{*}$Adam Mickiewicz University in Pozna\'{n} \\ $^{\ddagger}$Tilde \,  $^{\mathparagraph}$Universitat d’Alacant }
 

### 2018-05-18 wnmt.2018/wnmt.tex
small numerical differences

@@
 
-It is tempting to ignore these considerations, for example, by always decoding with a constant batch and beam size and ignoring hypotheses which are not needed but this comes at a cost of lower translation speed due to wasteful processing.
+It is tempting to ignore these considerations, for example, by always decoding with a constant batch and beam size and ignoring hypotheses which are not needed.
+Figure~\ref{algo:simple-mini-batching} illustrates a na\"ive mini-batching with a constant size batch.
+The downside to this algorithm is lower translation speed due to wasteful processing.
+
+\begin{algorithm}
+\begin{algorithmic}
+\Procedure{Batching}{encoded sentences $i$, beam size $s$}
+
+\State Create batch $b$ from $i$, $s$
+
+\While{hypo $h \neq \textless/s \textgreater, \forall h \in b $ } 
+  \State $b$ = DecodeAndBeamSearch($b$)
+\EndWhile
+
+\EndProcedure
+
+\end{algorithmic}
+\caption{Na\"ive mini-batching}
+\label{algo:simple-mini-batching}
+\end{algorithm}
 
@@
 
-  \ForAll{hypo $h$ in $b'$}
+  \ForAll{hypo $h \in b'$}
     \If{$h \neq \textless/s \textgreater$  }
@@
 Amun's decoding becomes faster as sentences finish translating.
-This contrast with the Marian inference engine, which uses a simpler batching algorithm, where the speed stays relatively constant throughout the decoding of the batch.
+This contrast with the Marian inference engine, which uses a na\"ive batching algorithm, where the speed stays relatively constant throughout the decoding of the batch.
 
@@
 Using batching can increase the translation speed by over 20 times in Amun, Figure~\ref{fig:batch-size}.
-Just as important, it doesn't suffer degradation with large batch sizes unlike the simpler algorithm which slows down when batch size over 1000 is used.
-This scalability issue is likely to rise in importance with new GPUs with ever increasing core counts.
+Just as important, it doesn't suffer degradation with large batch sizes unlike the na\"ive algorithm which slows down when batch size over 1000 is used.
+This scalability issue is likely to increase in importance with new GPUs with ever increasing core counts.
 

### 2018-05-18 wnmt.2018/wnmt.tex
Minor corrections, rephrasing, comments
Search for %!dtr! to find my comments.

@@
   \author{Hieu Hoang$^{\dagger}$ \, Tomasz Dwojak$^{*}$ \, Rihards Krislauks$^{\ddagger}$ \\ {\bf Daniel Torregrosa$^{\mathparagraph}$ \, Kenneth Heafield$^{\dagger}$ } \\[2mm]
-  $^{\dagger}$University of Edinburgh \, $^{*}$Adam Mickiewicz University in Pozna\'{n} \\ $^{\ddagger}$Tilde \,  $^{\mathparagraph}$Universitat d’Alacant }
+  $^{\dagger}$University of Edinburgh \, $^{*}$Adam Mickiewicz University in Pozna\'{n} \\ $^{\ddagger}$Tilde \,  $^{\mathparagraph}$Universitat d'Alacant }
 
@@
 
-This paper describe the submissions to the efficiency track for GPUs by members of the University of Edinburgh, Adam Mickiewicz University, Tilde and University of Alicante.
+This paper describes the submissions to the efficiency track for GPUs by members of the University of Edinburgh, Adam Mickiewicz University, Tilde and University of Alicante.
 We focus on efficient implementation of the recurrent deep-learning model as implemented in Amun, the fast inference engine for neural machine translation.
@@
 We use and enhanced a custom inference engine, Amun~\citep{junczys2016neural}, which we developed on the premise that fast deep-learning inference is an issue that deserves dedicated tools that are not compromised by competing objectives such as training or support for multiple models.
-As well as delivering on the practical goal of fast inference, it can serve as a test-bed for novel ideas on neural network inference, and is useful as a means to explore the upper limit of the possible speed for a particular model and hardware.
+As well as delivering on the practical goal of fast inference, it can serve as a test-bed for novel ideas on neural network inference, and is useful as a mean to explore the upper bound of the possible speed for a particular model and hardware.
 That is, Amun is an inference-only engine that supports a limited number of NMT models that put fast inference on modern GPU above all other considerations.
@@
 Our first submission was tailored to be as fast as possible while being above the baseline BLEU score.
-Our second submission trade some of the speed of first submission to return better quality translations.
+Our second submission trades some of the speed of first submission to return better quality translations.
 
@@
 
-Most NMT more predict a large number of classes in their output layer, corresponding to the number of words or subword units in their target language.
+Most NMT models predict a large number of classes in their output layer, corresponding to the number of words or subword units in their target language.
 For example, ~\citet{sennrich-haddow-birch:2016:P16-12} experimented with target vocabulary sizes of 60,000 and 90,000 sub-word units.
@@
 \end{eqnarray*}
-where $max_a$ is the original maximum value, $max_b$ is the new, higher maximum value, ie.
-$max_b > max_a$, and $\Delta = max_a - max_b$.
-The outline of our function is shown in Algorithm~\ref{algo:Fused Kernel}.
+where $max_a$ is the previous
+%!dtr! original as in the one at the start of this iteration, or original as the one at the start of the process? 
+maximum value, $max_b$ is the now higher maximum value, ie. $max_b > max_a$, and $\Delta = max_a - max_b$. The outline of our function is shown in Algorithm~\ref{algo:Fused Kernel}.
 
@@
 
-Both of our submitted systems uses the sequence-to-sequence model similar to that described in ~\citet{DBLP:journals/corr/BahdanauCB14}, containing a bidirectional RNN in the encoder and a two-layer RNN in the decoder.
+Both of our submitted systems use the sequence-to-sequence model similar to that described in ~\citet{DBLP:journals/corr/BahdanauCB14}, containing a bidirectional RNN in the encoder and a two-layer RNN in the decoder.
 We use byte pair encoding~\citep{sennrich-haddow-birch:2016:P16-12} to adjust the vocabulary size.
@@
 
-Our first system uses gated recurrent units (GRU) throughout, trained with Marian~\citep{junczys2018marian}, but submitted both systems using the Amun inference engine.
+%Our first system uses gated recurrent units (GRU) throughout, trained with Marian~\citep{junczys2018marian}, but submitted both systems using the Amun inference engine.
+
+Our first submitted system uses gated recurred units (GRU) throughout. It was trained using Marian~\citep{junczys2018marian}, but Amun was chosen as inference engine.
 
@@
 Our second system uses multiplicative-LSTM~\citep{krause2017} in the encoder and the first layer of a decder, and a GRU in the second layer, trained with an extension of the Nematus~\citep{sennrich-EtAl:2017:EACLDemo} toolkit which supports such models.
-We trained 2 systems with differing vocabulary sizes and varied the beam sizes before choosing the the settings that produces good translation quality on the validation set, Table~\ref{tab:mLSTM BLEU - vocab sizes}.
+As with our first submission, Amun is used as inference engine.
+We trained 2 systems with differing vocabulary sizes and varied the beam sizes, and chose the configuration that produced the best results for translation quality on the validation set, Table~\ref{tab:mLSTM BLEU - vocab sizes}.
 
@@
 
-The efficiency of Amun's batching algorithm can be seen by observing the time taken for each decoding step in a batch of sentence, Figure~\ref{fig:decode-step}.
-Amun's decoding becomes faster as sentences finish translating.
-This contrast with the Marian inference engine, which uses a na\"ive batching algorithm, where the speed stays relatively constant throughout the decoding of the batch.
+The efficiency of Amun's batching algorithm can be seen by observing the time taken for each decoding step in a batch of sentences, Figure~\ref{fig:decode-step}.
+Amun's decoding becomes faster as sentences are completely translated.
+This contrasts with the Marian inference engine, which uses a na\"ive batching algorithm, where the speed stays relatively constant throughout the decoding of the batch.
 
@@
 Using batching can increase the translation speed by over 20 times in Amun, Figure~\ref{fig:batch-size}.
-Just as important, it doesn't suffer degradation with large batch sizes unlike the na\"ive algorithm which slows down when batch size over 1000 is used.
-This scalability issue is likely to increase in importance with new GPUs with ever increasing core counts.
+Just as important, it doesn't suffer degradation with large batch sizes, unlike the na\"ive algorithm which slows down when batch sizes over 1000 is used.
+This scalability issue is likely to become more relevant as newer GPUs with ever increasing core counts are released.
+%increase in importance with new GPUs with ever increasing core counts.
 
@@
 Fusing the bias and softmax operations in the output layer with the beam search results in a speed improvement by 25\%, Figure~\ref{fig:fused}.
-Its relative improvement decrease marginally with increasing beam size.
+Its relative improvement decreases marginally as the beam size increases.
 
@@
 For higher beam sizes, the cost of maintaining the n-best list is felt.
+%!dtr! felt?
 
@@
 
-We have presented some of the improvement to Amun which are focused to improvement NMT inference.
+We have presented some of the improvement to Amun which are focused on improving NMT inference.
+%to improvement NMT inference.
 

### 2018-05-18 wnmt.2018/wnmt.tex
bah

@@
   \author{Hieu Hoang$^{\dagger}$ \, Tomasz Dwojak$^{*}$ \, Rihards Krislauks$^{\ddagger}$ \\ {\bf Daniel Torregrosa$^{\mathparagraph}$ \, Kenneth Heafield$^{\dagger}$ } \\[2mm]
-  $^{\dagger}$University of Edinburgh \, $^{*}$Adam Mickiewicz University in Pozna\'{n} \\ $^{\ddagger}$Tilde \,  $^{\mathparagraph}$Universitat d'Alacant }
+  $^{\dagger}$University of Edinburgh \, $^{*}$Adam Mickiewicz University \\ $^{\ddagger}$Tilde \,  $^{\mathparagraph}$Universitat d'Alacant }
 
@@
 
-We take an opposing approach by eschewing model selection in favour of efficient implementation.
+We take an opposing approach by eschewing model selection and parameter tuning in favour of efficient implementation.
 We use and enhanced a custom inference engine, Amun~\citep{junczys2016neural}, which we developed on the premise that fast deep-learning inference is an issue that deserves dedicated tools that are not compromised by competing objectives such as training or support for multiple models.
@@
 \begin{algorithmic}
-\Procedure{Batching}{encoded sentences $i$, beam size $s$}
+\Procedure{Batching}{encoded sentences $i$}
 
-\State Create batch $b$ from $i$, $s$
+\State Create batch $b$ from $i$
 
-\While{hypo $h \neq \textless/s \textgreater, \forall h \in b $ } 
-  \State $b$ = DecodeAndBeamSearch($b$)
+\While{hypo $h \neq EOS, \forall h \in b $ } 
+  \State Decode($b$)
 \EndWhile
@@
 \While{$b \neq \emptyset$} 
-  \State $b' \gets$ DecodeAndBeamSearch($b$)
-  \State $b \gets \emptyset$
+  \State Decode($b$)
 
-  \ForAll{hypo $h \in b'$}
-    \If{$h \neq \textless/s \textgreater$  }
-      \State Add $h$ to $b$
+  \ForAll{hypo $h \in b$}
+    \If{$h = EOS$  }
+      \State Remove $h$ from $b$
     \EndIf

### 2018-05-18 wnmt.2018/wnmt.tex
bah

@@
 
-The Amun engine implements an efficient batching algorithm that takes into account the actual number of hypotheses that needs to be decoded at each decoding step, Figure~\ref{algo:Mini-batching}.
+Amun implements an efficient batching algorithm that takes into account the actual number of hypotheses that needs to be decoded at each decoding step, Figure~\ref{algo:Mini-batching}.
 
@@
 
-As can be seen, the matrix p is iterated over five times - once to add the bias, three times to calculate the softmax, and once to search for the best classes.
+As can be seen, the vector p is iterated over five times - once to add the bias, three times to calculate the softmax, and once to search for the best classes.
 We propose fusing the three functions into one kernel, a popular optimization technique~\citep{Guevara2009EnablingTP}, making use of the following observations.
@@
 
-Secondly, we are only interested in the probabilities of the best classes. Since they are now known at the start of the kernel, we compute softmax only for those classes.
+Secondly, we are only interested in the probabilities of the best classes during inference, not of all classes.
+Since they are now known at the start of the softmax kernel, we compute softmax only for those classes.
 
@@
 \end{eqnarray*}
-where $max_a$ is the previous
-%!dtr! original as in the one at the start of this iteration, or original as the one at the start of the process? 
-maximum value, $max_b$ is the now higher maximum value, ie. $max_b > max_a$, and $\Delta = max_a - max_b$. The outline of our function is shown in Algorithm~\ref{algo:Fused Kernel}.
+where $max_a$ is the previous maximum value, $max_b$ is the now higher maximum value, ie.
+$max_b > max_a$, and $\Delta = max_a - max_b$.
+The outline of our function is shown in Algorithm~\ref{algo:Fused Kernel}.
 
@@
 The fused operation only has to loop through the large cost matrix once, therefore, for low beam sizes its is comparable in speed to the simple kernel to add the bias.
-For higher beam sizes, the cost of maintaining the n-best list is felt.
-%!dtr! felt?
+For higher beam sizes, the cost of maintaining the n-best list is begins to impact on speed.
 

### 2018-05-18 wnmt.2018/wnmt.tex
Fix some grammar issues; add a reference to Tildes WMT17 paper

--- 2018-05-18 wnmt.2018/wnmt.tex
@@
 We focus on efficient implementation of the recurrent deep-learning model as implemented in Amun, the fast inference engine for neural machine translation.
-We improve the performance with an efficient mini-batching algorithm and by fusing the softmax operation with the beam seach algorithm.
+We improve the performance with an efficient mini-batching algorithm and by fusing the softmax operation with the beam search algorithm.
 
@@
 We use and enhanced a custom inference engine, Amun~\citep{junczys2016neural}, which we developed on the premise that fast deep-learning inference is an issue that deserves dedicated tools that are not compromised by competing objectives such as training or support for multiple models.
-As well as delivering on the practical goal of fast inference, it can serve as a test-bed for novel ideas on neural network inference, and is useful as a mean to explore the upper bound of the possible speed for a particular model and hardware.
+As well as delivering on the practical goal of fast inference, it can serve as a test-bed for novel ideas on neural network inference, and it is useful as a means to explore the upper bound of the possible speed for a particular model and hardware.
 That is, Amun is an inference-only engine that supports a limited number of NMT models that put fast inference on modern GPU above all other considerations.
@@
 Our first submission was tailored to be as fast as possible while being above the baseline BLEU score.
-Our second submission trades some of the speed of first submission to return better quality translations.
+Our second submission trades some of the speed of the first submission to return better quality translations.
 
@@
 
-The Amun engine implements an efficient batching algorithm that takes into account the actual number of hypotheses that needs to be decoded at each decoding step, Figure~\ref{algo:Mini-batching}.
+The Amun engine implements an efficient batching algorithm that takes into account the actual number of hypotheses that need to be decoded at each decoding step, Figure~\ref{algo:Mini-batching}.
 
@@
 
-We focus on the last three steps, the outline for which are shown in Algorithm~\ref{algo:original-softmax-beamsearch}.
+We focus on the last three steps; their outline is shown in Algorithm~\ref{algo:original-softmax-beamsearch}.
 For brevity, we show the algorithm for 1-best, a beam search of the n-bests is a simple extension of this.
@@
 %!dtr! original as in the one at the start of this iteration, or original as the one at the start of the process? 
-maximum value, $max_b$ is the now higher maximum value, ie. $max_b > max_a$, and $\Delta = max_a - max_b$. The outline of our function is shown in Algorithm~\ref{algo:Fused Kernel}.
+maximum value, $max_b$ is the now higher maximum value, i.e., $max_b > max_a$, and $\Delta = max_a - max_b$. The outline of our function is shown in Algorithm~\ref{algo:Fused Kernel}.
 
@@
 
-In practise, however, efficiently using half-precision value requires a comprehensive redvelopment of the GPU code.
+In practise, however, efficiently using half-precision value requires a comprehensive redevelopment of the GPU code.
 We therefore make do with using the GPU's Tensor Core\footnote{https://devblogs.nvidia.com/programming-tensor-cores-cuda-9/} fast matrix multiplication routines which transparently converts 32-bit float point input matrices to 16-bit values and output a 32-bit float point product of the inputs.
@@
 
-Both of our submitted systems use the sequence-to-sequence model similar to that described in ~\citet{DBLP:journals/corr/BahdanauCB14}, containing a bidirectional RNN in the encoder and a two-layer RNN in the decoder.
+Both of our submitted systems use a sequence-to-sequence model similar to that described in ~\citet{DBLP:journals/corr/BahdanauCB14}, containing a bidirectional RNN in the encoder and a two-layer RNN in the decoder.
 We use byte pair encoding~\citep{sennrich-haddow-birch:2016:P16-12} to adjust the vocabulary size.
@@
 
-Our second system uses multiplicative-LSTM~\citep{krause2017} in the encoder and the first layer of a decder, and a GRU in the second layer, trained with an extension of the Nematus~\citep{sennrich-EtAl:2017:EACLDemo} toolkit which supports such models.
+Our second system uses multiplicative-LSTM~\citep{krause2017} in the encoder and the first layer of a decder, and a GRU in the second layer, trained with an extension of the Nematus~\citep{sennrich-EtAl:2017:EACLDemo} toolkit which supports such models; multiplicative-LSTM's suitability for use in NMT models has been previously demonstrated by~\citet{pinnis2017}.
 As with our first submission, Amun is used as inference engine.

### 2018-05-18 wnmt.2018/wnmt.tex

@@
 
-The Amun engine implements an efficient batching algorithm that takes into account the actual number of hypotheses that need to be decoded at each decoding step, Figure~\ref{algo:Mini-batching}.
+Amun implements an efficient batching algorithm that takes into account the actual number of hypotheses that need to be decoded at each decoding step, Figure~\ref{algo:Mini-batching}.
 
@@
 \end{eqnarray*}
-where $max_a$ is the previous
-%!dtr! original as in the one at the start of this iteration, or original as the one at the start of the process? 
-maximum value, $max_b$ is the now higher maximum value, i.e., $max_b > max_a$, and $\Delta = max_a - max_b$. The outline of our function is shown in Algorithm~\ref{algo:Fused Kernel}.
+where $max_a$ is the previous maximum value, $max_b$ is the now higher maximum value, i.e., $max_b > max_a$, and $\Delta = max_a - max_b$.
+The outline of our function is shown in Algorithm~\ref{algo:Fused Kernel}.
 

### 2018-05-18 wnmt.2018/wnmt.tex
bah

@@
 
+We will compare the effect of the two implementations in the Section~\ref{sec:Result}.
+
 \subsection{Softmax and Beam Search Fusion}
@@
 \end{eqnarray*}
+
 where $max_a$ is the previous maximum value, $max_b$ is the now higher maximum value, i.e., $max_b > max_a$, and $\Delta = max_a - max_b$.

### 2018-05-18 wnmt.2018/wnmt.tex
bah

@@
 \begin{tabular}{cc}
-{\includegraphics[scale=0.3]{pie-time.png}} 
+{\includegraphics[scale=0.35]{pie-time.png}} 
 \end{tabular}
@@
 \begin{tabular}{cc}
-{\includegraphics[scale=0.3]{decode-step.png}} 
+{\includegraphics[scale=0.35]{decode-step.png}} 
 \end{tabular}
@@
 \begin{tabular}{cc}
-{\includegraphics[scale=0.3]{batch-size.png}} 
+{\includegraphics[scale=0.35]{batch-size.png}} 
 \end{tabular}
@@
 \begin{tabular}{cc}
-{\includegraphics[scale=0.3]{fused.png}} 
+{\includegraphics[scale=0.35]{fused.png}} 
 \end{tabular}

### 2018-05-24 wnmt.2018/wnmt.tex
bah

@@
 As neural machine translation (NMT) models have become the new state-of-the-art, the challenge is to make their deployment efficient and economical.
-This is the challenge that this shared task is shining a spotlight on.
+This is the challenge that this shared task~\citep{birch2018wnmt} is shining a spotlight on.
 
@@
 
-We describe the main enhancements to Amun since the original 2016 publication.
+We describe the main enhancements to Amun since the original 2016 publication that has improved translation speed.
 
@@
 
-\subsection{Softmax and Beam Search Fusion}
+\subsection{Softmax and K-Best Fusion}
 
@@
    \item \vspace{-2 mm} applying the activation function, most commonly softmax $ p_i = \exp(p_i) / \sum \exp(p_i) $
-   \item \vspace{-2 mm} a beam search for the best (or n-best) output classes $\argmax_i p_i$
+   \item \vspace{-2 mm} a search for the best (or k-best) output classes $\argmax_i p_i$
 \end{enumerate}
@@
 We focus on the last three steps; their outline is shown in Algorithm~\ref{algo:original-softmax-beamsearch}.
-For brevity, we show the algorithm for 1-best, a beam search of the n-bests is a simple extension of this.
+For brevity, we show the algorithm for 1-best, a k-best search is a simple extension of this.
 
@@
 \end{algorithmic}
-\caption{Original softmax and beam Search Algorithm}
+\caption{Original softmax and k-best algorithm}
 \label{algo:original-softmax-beamsearch}
@@
 
-\caption{Fused softmax and beam search}
+\caption{Fused softmax and k-best}
 \label{algo:Fused Kernel}

### 2018-05-24 wnmt.2018/wnmt.tex
bah

@@
 \DeclareMathOperator*{\argmax}{argmax}
+
+% declaration of the new block
+\algblock{ParFor}{EndParFor}
+% customising the new block
+\algnewcommand\algorithmicparfor{\textbf{parfor}}
+\algnewcommand\algorithmicpardo{\textbf{do}}
+\algnewcommand\algorithmicendparfor{\textbf{end\ parfor}}
+\algrenewtext{ParFor}[1]{\algorithmicparfor\ #1\ \algorithmicpardo}
+\algrenewtext{EndParFor}{\algorithmicendparfor}
 
@@
 
+Since we are working on GPU optimization, it is essential to make full use of the many GPU cores available.
+This is accomplished by well-known parallelization methods which multi-thread the algorithms.
+For example, Algorithm~\ref{algo:Argmax only} is parallelized by sharding the vector $p$ and calculating $best$ and $max$ on each shard in parallel.
+The ultimate $best$ is found in the following reduction step, Algorithm~\ref{algo:Parallel Argmax only}.
+
+\begin{algorithm}
+\begin{algorithmic}
+\Procedure{Fused-kernel-1-best}{vector $p$, bias vector $b$}
+
+\Comment{parallelize}
+\State Create shards ${p^1...p^n}$ from $p$
+\ParFor{$p^j \in {p^1...p^n}$}
+  \State $max^j \gets - \infty$ 
+  \ForAll{$p_i^j$ in $p^j$}
+    \If{$p_i^j + b_i > max$}
+      \State $max^j \gets p_i^j + b_i$
+      \State $best^j \gets i$
+    \EndIf
+  \EndFor
+
+\EndParFor
+
+\Comment{reduce}
+\State $max \gets - \infty$ 
+\ForAll{$max^j \in max^1...max^n$}
+  \If{$max^j > max$}
+    \State $max \gets max^j$
+    \State $best \gets best^j$
+  \EndIf
+\EndFor
+
+\Return $best$
+
+\EndProcedure
+
+\end{algorithmic}
+\caption{Parallel find 1-best only}
+\label{algo:Parallel Argmax only}
+\end{algorithm}
+
 %These changes do not reduce the asymptotic runtime of our algorithm, which remains at $\mathcal{O}(|p|)$, but it does bring the number of iterations over $p$ down from 5 to 1.

### 2018-05-28 wnmt.2018/wnmt.tex
turn on timings
small numerical differences

@@
 We focus on efficient implementation of the recurrent deep-learning model as implemented in Amun, the fast inference engine for neural machine translation.
-We improve the performance with an efficient mini-batching algorithm and by fusing the softmax operation with the beam search algorithm.
+We improve the performance with an efficient mini-batching algorithm, and by fusing the softmax operation with the k-best extraction algorithm.
 
@@
 Softmax 	& 1.69		& 2.07 (-86.6\%)\\
-Beam search	& 12.53		&  \\ \hline
+K-best extr.	& 12.53		&  \\ \hline
 \multicolumn{3}{|c|}{\emph{Beam size 3}}	\\ \hline	
@@
 Softmax 	& 4.75		& 3.43 (-87.1\%)\\
-Beam search	& 18.23		&  \\ \hline
+K-best extr.	& 18.23		&  \\ \hline
 \multicolumn{3}{|c|}{\emph{Beam size 9}}	\\ \hline	
@@
 Softmax 	& 14.4		& 17.5 (-72.1\%)\\
-Beam search	& 36.7		&  \\ \hline
+K-best extr.	& 36.7		&  \\ \hline
 \end{tabular}

### 2018-05-28 wnmt.2018/wnmt.tex
small numerical differences

@@
 
-This paper describes the submissions to the efficiency track for GPUs by members of the University of Edinburgh, Adam Mickiewicz University, Tilde and University of Alicante.
+This paper describes the submissions to the efficiency track for GPUs at the Workshop for Neural Machine Translation and Generation by members of the University of Edinburgh, Adam Mickiewicz University, Tilde and University of Alicante.
 We focus on efficient implementation of the recurrent deep-learning model as implemented in Amun, the fast inference engine for neural machine translation.
@@
 
-\begin{table}
+\begin{table}[h]
 \begin{center}
